### 2020-02-21 manuscript/ERL_global_C_review.Rmd
started manuscript, rearranged folders

---
title: ""
output: 
  pdf_document:
    citation_package: natbib
    fig_caption: yes
bibliography: [library.bib,  packages.bib]
biblio-style: apalike
csl: apa.csl

header-includes:  \usepackage{float}
                  \usepackage{caption}
                  \captionsetup[table]{font=footnotesize}
                  \captionsetup[figure]{font=footnotesize}
                  \captionsetup[figure]{labelformat=empty}
                  \captionsetup[table]{labelformat=empty}
---
\raggedright

**Title:** Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global forest carbon database (ForC)

**Authors:** 
Kristina J. Anderson-Teixeira^1,2^*
Valentine Herrmann^1^
Becky Banbury Morgan
Ben Bond-Lamberty
Susan Cook-Patton
Abigail Ferson
Norbert Kunert
Jennifer McGarvey
Helene C. Muller-Landau^1^
Maria Wang

**Author Affiliations:**

1. Conservation Ecology Center; Smithsonian Conservation Biology Institute; National Zoological Park, Front Royal, VA 22630, USA
2. Center for Tropical Forest Science-Forest Global Earth Observatory; Smithsonian Tropical Research Institute; Panama, Republic of Panama

*corresponding author: teixeirak@si.edu; +1 540 635 6546

\newpage

### Summary
*(The Summary should be no longer than 300 words and divided into the following sections: Background, Methods/Design, Review results/ Synthesis and Discussion. Up to 7 Keywords should also be provided. The total length of the article is flexible.)*

*Background.* The fate of Earth’s climate closely linked to forests, which strongly influence atmospheric carbon dioxide (CO2) and climate through their influential role in the global carbon (C) cycle. Synthetic understanding of global forest C cycles is needed to constrain model estimates of forest feedbacks to climate change and to more accurately quantify the influence of land use decisions on climate.

*Methods/Design.* Here, we draw from the Global Forest C Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation. Specifically, we draw upon ## records from ## geographic locations representing ## C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests. For regrowth forests, we quantify age trends for all variables.

*Review Results/ Synthesis.*  The rate of C cycling generally increased from boreal to tropical regions, whereas C stocks showed less directional variation. The majority of flux variables, together with most live biomass pools, increased significantly with stand age, and the rate of increase again tended to increase from boreal to tropical regions.

*Discussion.* *[Discussion section will interpret results, highlighting new and significant findings, and discuss implications. Tentative headings are “Stand level C cycling in forests globally”, “Age trends in C cycling”, and “Implications for climate change mitigation”.]*

*Key words*: forest ecosystems; carbon cycle; stand age; productivity; respiration; biomass; global

### Background
*(The Background section needs to present the rationale for why a systematic review of this topic is needed along with a history of what has been done to date and an expectation of what new will emerge from the review, especially if quantitative meta-analyses of studies are being considered.)*

Forest ecosystems globally influence climate through their critical role in the global carbon (C) cycle (Fig. 1). Their annual gross CO2 sequestration (gross primary productivity, GPP) is estimated at 59 Gt C yr-1 (Beer et al 2010), or 6.3 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017). A small portion of global terrestrial GPP is retained in ecosystems (mainly forests), resulting in a C sink that averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017). The remaining ~98% of global GPP is counterbalanced by ecosystem respiration (Reco) or wildfire. Perturbation to the global GPP- Reco balance can substantially influence the growth rate of atmospheric CO2; for example, the 2015-2016 El Niño, which brought historically high temperatures and low precipitation to the tropics, released an extra 2.5 ± 0.3 Gt C to the atmosphere, resulting in the largest recorded atmospheric CO2 growth rate (Le Quéré et al 2016, Liu et al 2017). In addition, forests contain substantial C stocks that, when disturbed, release significant amounts of CO2 to the atmosphere. Although they cover only ~30% of the land surface, forests contain an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C (biomass and soils; Bonan 2008). Globally, gross tropical deforestation averaged 2.8 Gt C yr-1 from 2000-2007, but ~40% of this was offset by forest regrowth, resulting in a net source of ~1.1 Gt C yr-1 from tropical land use change (Pan et al 2011). This, coupled with minimal net deforestation in the extratropics and net uptake by intact forests (Pan et al 2011, Schimel et al 2015), resulted in a total gain in forest C of ~1.2-1.7 Gt C yr-1 (Le Quéré et al 2017, Schimel et al 2015), thereby substantially slowing the rate of increase of atmospheric CO2.

![**Figure 1 | The role of forests in the global carbon cycle. Values apply to the period 2000-2018.** Sources are as follows: biomass- (Pan et al 2013, Baccini et al 2012) IPCC 2000; intact forest sink, net forest sink, tropical deforestation and regrowth-(Pan et al 2011); terrestrial sink- (Le Quéré et al 2017); GPP- (Beer et al 2010); respiration+fire: calculated here; fossil fuel emissions- (Le Quéré et al 2017); atmospheric CO2 (2018 value from https://scripps.ucsd.edu/programs/keelingcurve/). ](tables_figures/Global_C_cycle/Global C cycle diagram.png)

### 2020-02-21 manuscript/ERL_global_C_review.Rmd
worked on transferring material to draft, added table

@@
 Values apply to the period 2000-2018.** Sources are as follows: biomass- (Pan et al 2013, Baccini et al 2012) IPCC 2000; intact forest sink, net forest sink, tropical deforestation and regrowth-(Pan et al 2011); terrestrial sink- (Le Quéré et al 2017); GPP- (Beer et al 2010); respiration+fire: calculated here; fossil fuel emissions- (Le Quéré et al 2017); atmospheric CO2 (2018 value from https://scripps.ucsd.edu/programs/keelingcurve/).
-](tables_figures/Global_C_cycle/Global C cycle diagram.png)
+**see issue 19: https://github.com/forc-db/ERL-review/issues/19** ](tables_figures/Global_C_cycle/Global C cycle diagram.png)
+
+Given their vital role in regulating atmospheric CO2, the future of Earth’s forests will strongly influence the course of climate change.
+If forests globally respond to the suite of contemporary and future global change drivers—including elevated CO2, climate change, and atmospheric deposition—with increased productivity and net biomass increases, they could continue to act as a significant buffer against anthropogenic emissions.
+In contrast, if factors such as higher temperatures and drought stress lead to net C losses, the C sink of forests could be lost and even reversed.
+Global coupled climate models vary substantially in predictions regarding the future of the global forest C sink, with uncertainty in end-of-century terrestrial NEP exceeding current annual anthropogenic CO2 emissions (Friedlingstein et al 2006, Cavaleri et al 2015).
+Anthropogenic land use decisions will also strongly influence the future course of atmospheric CO2.
+In the absence of measures to reduce deforestation, this will continue to be a significant component of anthropogenic CO2 emissions.
+On the other hand, forest conservation, reforestation, and forest management hold strong potential to help stabilize and eventually reduce total anthropogenic CO2 emissions (estimated total of 4.4 Gt C yr-1; Griscom et al 2017; *Cook-Patton et al.*; **other refs?**) and thereby contribute towards the goals of the Paris Climate Agreement (UNFCCC 2015, Houghton et al 2015; **REFS**).
+Indeed, forest-based climate mitigation was a key component of Paris Climate Agreement commitments, totaling approximately one-fourth of nations’ planned emissions reductions (Grassi et al 2017; IPCC1.5).
+Thus, forest will strongly influence the future course of climate change both through their responses to global change and through forest-related anthropogenic land use decisions.
+
+While forests will play a critical role in the future course of climate change, there remain significant uncertainties as to current and future C cycling the world’s forests.
+Even as the climate changes and forest age distributions shift, there are important gaps in our fundamental understanding of how C cycling varies with climate and stand age.
+A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling.
+Recent studies attempting to quantify the role of forests in the global C cycle (e.g., Pan et al 2011) have been hampered by insufficient data on regrowth rates of tropical forests worldwide, C in non-living pools (i.e., dead wood, litter, soil), and source/sink status of forests in some regions.
+Moreover, as discussed above, model representation of forest responses to global change remain highly uncertain (Friedlingstein et al 2006, Cavaleri et al 2015), with efforts to reduce model uncertainty in part limited by ready availability of appropriate benchmark data (Clark et al 2017, Luo et al 2012).
+To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018).
+ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (e.g., stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications.
+ForC currently contains **#** records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.
+
+**Here, we synthesize ForC data to provide a big-picture overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age. We address two broad hypotheses: 
+H1. The rate of C cycling generally decreases from tropical to boreal regions (i.e., tropical > temperate > boreal), whereas C stocks show less directional variation.
+H2. Both C fluxes and stocks generally increase with stand age, with particularly strong tends in autotrophic C fluxes. 
+While components of these hypotheses have been previously addressed (e.g., REFS(Luyssaert et al 2007, Anderson et al 2006, Anderson-Teixeira et al 2013), our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.**
+
+### Methods/ Design
+*(The Methods/Design section needs to describe how articles in the reviews were identified and what criteria were used for justifying inclusion in the review.
+While traditional reviews relying on past experience and expert knowledge are acceptable, editors should encourage reviews that are set up as so-called ‘systematic reviews’ (see the Cochrane Review procedure which initiated the process in the medical and health science: http://community.cochrane.org/ about-us/evidence-based-health-care).)*
+
+This review synthesizes data from the ForC database (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018).
+ForC amalgamates numerous intermediary data sets (e.g., REFS) and original studies.
+Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records.
+The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age.
+As such, there has been a focus on incorporating data from regrowth forests (e.g., Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible (83% of records in v.2.0; Anderson-Teixeira et al 2018).
+Particular attention was given to developing the database for tropical forests (Anderson-Teixeira et al 2016), yet these represented only approximately one-third of records in ForC v.2.0 (Anderson-Teixeira et al 2018).
+Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database (Bond-Lamberty and Thomson 2010; *hopefully Cook-Patton*; see [GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv)) ….
+We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (e.g.,: REFS).
+The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: XX).
+
+For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
+Climate zones were defined according to Köppen-Geiger zones, which were extracted based on site geographic coordinates from the ESRI Köppen-Geiger map (downloaded June 2014 from http://maps3.arcgisonline.com/ArcGIS/rest/services/A-16/Köppen- Geiger_Observed_and_Predicted_Climate_Shifts/MapServer; Rubel and Kottek 2010) and are recorded in the Koeppen field of the ForC SITES table.
+Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd).
+Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations (SYNMAP; Jung et al 2006) and recorded in the  dominant.veg field of the ForC MEAUREMENTS table.
+Stand age was obtained from the stand.age field of MEASUREMENTS table, and forests classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication).
+Records for which stand age was unknown were excluded from the analysis.
+These groupings were defined in March 2018, at which point they covered 91% of the primary variable records for forests of known age, or 86% of total records, in ForC (v.2.0; Anderson-Teixeira et al 2018).
+The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with <400 records total for stands of any age.
+
+We drew upon records for *#* annual flux and *#* C stock variables (Table 1).
+For this analysis, we combined some of ForC’s specific variables (e.g., multiple variables for net primary productivity including various components) into more broadly defined variables (Table 1, [this table](https://github.com/forc-db/ForC/blob/master/figures/C_cycle_diagrams/ForC_variables_mapping_for_C_cycle_diagrams.csv))).
+Although ForC contains information that may be used to standardize or control for methodological differences (e.g., area sampled, min stem diameter sampled, allometric equations applied; Anderson-Teixeira et al 2018), for this analysis we included all relevant data in ForC.
+Throughout ForC, for all measurements drawing from tree census data (e.g., biomass, productivity), the minimum stem diameter sampled was $\le$ 10cm.
+All records were measured directly or derived from field measurements (as opposed to modeled).
+
+\newpage
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+S1 <- read.csv("tables_figures/C_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
+kable(S1, format = "latex", booktabs = TRUE, caption = "Table 1.
+ForC variables included in this analysis.
+Shown are variable names and descriptions, the associated variable name(s) in the database, and number of plots with records.
+Complete list of variables with full definitions is available at [GitHub url].", escape = F)%>%
+  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 7) %>%
+  column_spec(1, width = "2cm") 
+```

### 2020-02-21 manuscript/ERL_global_C_review.Rmd
finished initial draft of table 1

@@
 Complete list of variables with full definitions is available at [GitHub url].", escape = F)%>%
-  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 7) %>%
-  column_spec(1, width = "2cm") 
+  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12)  %>%
+  pack_rows("Annual fluxes", 1, 21, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
+  pack_rows("Stocks", 22, 32, latex_gap_space = "1em", colnum = 1, hline_before = FALSE)
 ```

### 2020-02-21 manuscript/ERL_global_C_review.Rmd
transferred more content

@@
 
-\newpage
 ```{r eval = TRUE, echo=FALSE, warning=FALSE}
@@
 ```
+
+Analyses drew from ForC-simplified (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which is a rearrangement of ForC intended to facilitate analyses.
+In generating ForC-simplified, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006).
+Duplicate or otherwise conflicting records (indicated in conflicts field of MEASUREMENTS table) were reconciled as follows.
+Replicate measurements (i.e., replicates from within a single study) were averaged.
+Records that subsumed others— *i.e.*, the time period included that of $\ge$ 2 other records or dates were unknown and therefore conflicted with $\ge$ 2 other records—were removed.
+For each group of duplicate records—i.e., measurements of the same variable in the same plot at the same time—one record was assigned precedence (recorded in D.precedence field).
+When measurement periods overlapped or were not specified, precedence was given first to records representing longer measurement periods (*i.e.*, end.date - start.date) and then to more recently published values.
+We manually reviewed duplicates that differed only in methodology, assigning precedence to the record employing a more comprehensive approach (e.g., inclusion of understory, lianas, or bamboo as opposed to just trees) or using a favored methodology.
+
+Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand establishment (*i.e.*, that reflected by stand.age).
+Specifically, we removed all plots flagged as managed in ForC-simplified (managed field).
+This included plots with any record of managements manipulating CO2, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.
+Plots flagged as disturbed in ForC-simplified included stands that had undergone anthropogenic thinning or partial harvest (“Cut” or “Harvest” codes) unless this was very minor (percent.mortality= “minor”).
+We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods.
+We also removed all plots for which no stand history information had been retrieved.
+
+Data were analyzed to produce basic summaries of C cycle patterns across biomes and stand ages following an approach similar to that of Anderson-Teixeira et al (2016).
+For mature forests, to obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot, weighting flux measurements according to the length of measurement periods (i.e., end.date - start.date).
+Values were then averaged across plots clustered within 25 km of one another (geographic.area field of SITES table, sensu Anderson-Teixeira et al 2018), weighting by area.sampled (MEASUREMENTS table) or plot.area (PLOTS table) if available for all records.
+This step was taken to avoid pseudo-replication and to combine any records from sites with more than one name in ForC.
+Finally, we computed statistics with geographic.area as the unit of replication.
+To compare across biomes, [**Valentine, please describe**].
+There were enough data to run this analysis for all focal variables but XX.
+
+For young (<100yrs) forest types, we employed a mixed effects model (XX in R) with biome and log10[stand.age] as fixed effects and plot nested within geographic.area as a random effect.
+When the effect of stand.age was significant at p $\le$ 0.05 and when each biome had records for stands of at least 10 different ages, a biome - stand.age interaction was included in the model.
+In the C cycle schematics for young forests, we report equations based on these models.
+In cases where there was no significant effect of stand.age, records were averaged as for mature stands.
+
+All database manipulation, analyses, and figure production were fully automated in R (version, citation).
+Materials required to fully reproduce these analyses, including data, R scripts, and image files, are archived in Zenodo (DOI: TBD].
+Data, scripts, and results presented here are also available through the open-access ForC GitHub repository (https://github.com/forc-db/ForC), where many will be updated as the database develops.
+
+### Review Results/ Synthesis
+*(The Review Results/Synthesis section presents the findings of the analyses.)*
+
+*C stocks and fluxes of mature forests*
+	Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5, and statistics for each biome type are also summarized at: [GitHub URL].
+Of the # flux and # stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, # fluxes and # stocks in temperate broadleaf forests, #fluxes and # stocks in temperate conifer forests, and fluxes and # stocks in boreal forests.
+For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets were generally consistent.
+That is, component fluxes and stocks summed to within 1 std of more inclusive fluxes in all but one instance (in temperate conifer forests, $aboveground woody biomass + foliage biomass > aboveground biomass + 1std$; Fig.
+4).
+**This is based on my quick scan of the figures.
+Need to add a formal automated test: https://github.com/forc-db/ERL-review/issues/16 **

### 2020-02-21 manuscript/ERL_global_C_review.Rmd
changed hypotheses to questions

@@
 
-**Here, we synthesize ForC data to provide a big-picture overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age. We address two broad hypotheses: 
-H1. The rate of C cycling generally decreases from tropical to boreal regions (i.e., tropical > temperate > boreal), whereas C stocks show less directional variation.
-H2. Both C fluxes and stocks generally increase with stand age, with particularly strong tends in autotrophic C fluxes. 
-While components of these hypotheses have been previously addressed (e.g., REFS(Luyssaert et al 2007, Anderson et al 2006, Anderson-Teixeira et al 2013), our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.**
+Here, we synthesize ForC data to provide a big-picture overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age. We address two broad questions:
+
+1. How do C cycling vary across the worlds major forest biomes (i.e., tropical, temperate, boreal)?
+
+2. How does C cycling vary with stand age?
+
+While components of these questions have been previously addressed (e.g., REFS(Luyssaert et al 2007, Anderson et al 2006, Anderson-Teixeira et al 2013), our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.
 
@@
 
-*C stocks and fluxes of mature forests*
+#### C stocks and fluxes of mature forests
+
 Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5, and statistics for each biome type are also summarized at: [GitHub URL].

### 2020-02-21 manuscript/ERL_global_C_review.Rmd
Table 1 & Fig.2
added C cycle variables with no data to Table 1
started adding C cycle diagrams

@@
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12)  %>%
-  pack_rows("Annual fluxes", 1, 21, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
-  pack_rows("Stocks", 22, 32, latex_gap_space = "1em", colnum = 1, hline_before = FALSE)
+  pack_rows("Annual fluxes", 1, 23, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
+  pack_rows("Stocks", 24, 34, latex_gap_space = "1em", colnum = 1, hline_before = FALSE)
 ```
@@
 Need to add a formal automated test: https://github.com/forc-db/ERL-review/issues/16 **
+
+![**Fig.
+2 | C cycle diagram for mature tropical broadleaf forests.** Presented are mean±std, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha-1 yr-1 (fluxes) or Mg C ha-1 (stocks).
+Arrows indicate fluxes, boxes indicate stocks.
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.](tables_figures/Tropical broadleaf MATURE.png)

### 2020-02-21 manuscript/ERL_global_C_review.Rmd
added landscape orientation for C cycle diagrams

@@
                   \captionsetup[table]{labelformat=empty}
+                  \usepackage{pdflscape}
+                  \newcommand{\blandscape}{\begin{landscape}}
+                  \newcommand{\elandscape}{\end{landscape}}
+
 ---
@@
 
-**Title:** Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global forest carbon database (ForC)
+**Title:** Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database
 
@@
 kable(S1, format = "latex", booktabs = TRUE, caption = "Table 1.
-ForC variables included in this analysis.
-Shown are variable names and descriptions, the associated variable name(s) in the database, and number of plots with records.
+Carbon cycle variables incuded in this analysis.
 Complete list of variables with full definitions is available at [GitHub url].", escape = F)%>%
@@
 
-#### C stocks and fluxes of mature forests
+*C cycling in mature forests*
 
@@
 
+\newpage
+\blandscape
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 2 | C cycle diagram for mature tropical broadleaf forests.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
+Arrows indicate fluxes, boxes indicate stocks.
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/Tropical broadleaf MATURE.png")
+```
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 3 | C cycle diagram for mature temperate broadleaf forests.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
+Arrows indicate fluxes, boxes indicate stocks.
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/Temperate broadleaf MATURE.png")
+```
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 3 | C cycle diagram for mature temperate conifer forests.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
+Arrows indicate fluxes, boxes indicate stocks.
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/Temperate conifer MATURE.png")
+```
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 4 | C cycle diagram for mature boreal conifer forests.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
+Arrows indicate fluxes, boxes indicate stocks.
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/Boreal conifer MATURE.png")
+```
+\elandscape
+
+(**check paragraph with latest data**) The largest C fluxes—including $GPP$, $NPP$, $ANPP$, $BNPP$, $R_{soil_het}$, $R_{soil}$, $R_{eco}$, — were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests  ([ForC_variable_averages_per_Biome](https://github.com/forc-db/ForC/blob/master/numbers_and_facts/ForC_variable_averages_per_Biome.csv)) (Fig.
+6).
+The same held true for some of the subsidiary fluxes: ANPP_foliage, ANPP_woody, [**OTHERS?**].
+Other subsidiary fluxes—including NPP_woody, ANPP_repro, ANPP_stem, ANPP_branch, woody.mortality, BNPP_coarse, BNPP_fine,  [**OTHERS?**]—deviated from this pattern and/or lacked data for some biomes.
+Net ecosystem exchange (NEE) did not follow this pattern, with no significant differences across biomes but the largest (negative) average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests.
+Thus, C cycling rates generally decreased from tropical to temperate to boreal forests, but with less apparent trends for some of the subsidiary fluxes and an important exception in the overall C balance (NEE).
+
+There were less distinct trends in C stocks across biomes (Fig.
+7).
+Biome means for live aboveground and total biomass followed the same general trend as the major fluxes, with biomass of tropical= temperate broadleaf $\ge$ temperate conifer> boreal forests (Fig.
+7a).
+However, the relative differences in these means were much smaller than for the major C fluxes, and there was significant variation within biomes.
+Maximum aboveground biomass values followed a very different trend than the means: temperate broadleaf> temperate conifer>boreal>tropical.
+There were some statistically significant biome differences in less frequently sampled C stocks (e.g., woody biomass, foliage biomass, deadwood), but given high within-biome variability in C stocks and relatively low sample sizes, these were likely attributable to sampling biases and methodological differences than to true differences across biomes.
+
+*C cycling in young forests*
+
+(**check paragraph with latest data**) Average C cycles for forests <100 years old are presented in Figures 8-11.
+In general, ForC contained roughly comparable information on C fluxes in regrowth vs mature forests, with the notable exception of tropical forests, for which there were no fluxes that had been measured in $\ge$ 7 different geographic areas.
+C stocks were better documented for regrowth stands, with a total of # records from # distinct geographic areas.
+Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Fig.
+6-11; detailed below).
+Differences across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests (Figs.
+6-7).
+
+(**check paragraph with latest data**) ForC contained 14 flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig.
+6-7 and **S#- SI figures including plots for all variables**).
+Of these, 9 increased significantly with log10[stand.age]: GPP, NPP, ANPP, ANPP_foliage, ANPP_woody, ANPP_woody_stem, BNPP, BNPP_root_fine, Reco, and net C sequestration (-NEE).
+The remaining five--ANPP_woody_branch, BNPP_root_coarse, soil heterotrophic respiration, and total soil respiration–displayed no significant relationship to stand age, although all displayed a positive trend.
+In terms of C stocks, 10 variables had sufficient data to test for age trends.
+Six of these—total biomass, aboveground biomass, aboveground woody biomass, foliage biomass, root biomass, and coarse root biomass—increased significantly with log10[stand.age].
+The remaining four displayed non-significant positive trends:  fine root biomass, total dead wood, standing dead wood, and organic layer.
+
 ![**Fig.
-2 | C cycle diagram for mature tropical broadleaf forests.** Presented are mean±std, where geographically distinct areas are treated as the unit of replication.
-All units are Mg C ha-1 yr-1 (fluxes) or Mg C ha-1 (stocks).
+8 | C cycle diagram for mature tropical broadleaf forests.** Presented are mean±std, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha^-1^ yr^-1^ (fluxes) or Mg C ha^-1^ (stocks).
 Arrows indicate fluxes, boxes indicate stocks.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.](tables_figures/Tropical broadleaf MATURE.png)
+When age trends are signficant, they are presented with numbered equations; otherwise, means are presented.
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.](tables_figures/Tropical broadleaf YOUNG.png)
+
+![**Fig.
+9 | C cycle diagram for mature temperate broadleaf forests.** Presented are mean±std, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha^-1^ yr^-1^ (fluxes) or Mg C ha^-1^ (stocks).
+Arrows indicate fluxes, boxes indicate stocks.
+When age trends are signficant, they are presented with numbered equations; otherwise, means are presented.
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.](tables_figures/Temperate broadleaf YOUNG.png)
+
+![**Fig.
+10 | C cycle diagram for mature temperate deciduous forests.** Presented are mean±std, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha^-1^ yr^-1^ (fluxes) or Mg C ha^-1^ (stocks).
+Arrows indicate fluxes, boxes indicate stocks.
+When age trends are signficant, they are presented with numbered equations; otherwise, means are presented.
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.](tables_figures/Temperate conifer YOUNG.png)
+
+![**Fig.
+11 | C cycle diagram for mature boreal forests.** Presented are mean±std, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha^-1^ yr^-1^ (fluxes) or Mg C ha^-1^ (stocks).
+Arrows indicate fluxes, boxes indicate stocks.
+When age trends are signficant, they are presented with numbered equations; otherwise, means are presented.
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.](tables_figures/Boreal conifer YOUNG.png)
+
+### Discussion 
+(*The Discussion section highlights the general conclusions, emphasizes novel findings, puts them in context and identifies issues and inferences for future work or policy considerations.*)
+
+Carbon cycling rates generally increased from boreal to tropical regions and with stand age.
+Specifically, the major C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests—a pattern that generally held for regrowth as well as mature forests (Figs.
+6-7).
+In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs.
+2-5, 7).
+The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Figs.
+6-7).
+Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks of mature forests are influenced by a different set of drivers.
+
+*C cycling across biomes*
+
+*Age trends in C cycling*
+
+### Acknowledgements
+All researchers whose data is included in ForC and this analysis.
+A Smithsonian Scholarly Studies grant to KAT and HML.
+
+### References
+(*All references should be presented in alphabetical order and should also be bundled in ways that will help the reader to find those that are germane to each of the themes, topics or issues that the authors have used to organize the review.
+Such a matrix will be included
+as supplementary material. ERL has no standardized format for subdividing references by cross-cutting themes, but guidance from editors will be helpful to authors.*)

### 2020-02-21 manuscript/ERL_global_C_review.Rmd
landscape orientation for C cycle diagrams

@@
 
-![**Fig.
-8 | C cycle diagram for mature tropical broadleaf forests.** Presented are mean±std, where geographically distinct areas are treated as the unit of replication.
-All units are Mg C ha^-1^ yr^-1^ (fluxes) or Mg C ha^-1^ (stocks).
+\newpage
+\blandscape
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | C cycle diagram for young tropical broadleaf forests.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
 Arrows indicate fluxes, boxes indicate stocks.
 When age trends are signficant, they are presented with numbered equations; otherwise, means are presented.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.](tables_figures/Tropical broadleaf YOUNG.png)
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
 
-![**Fig.
-9 | C cycle diagram for mature temperate broadleaf forests.** Presented are mean±std, where geographically distinct areas are treated as the unit of replication.
-All units are Mg C ha^-1^ yr^-1^ (fluxes) or Mg C ha^-1^ (stocks).
+  knitr::include_graphics("tables_figures/Tropical broadleaf YOUNG.png")
+```
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 9 | C cycle diagram for young temperate broadleaf forests.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
 Arrows indicate fluxes, boxes indicate stocks.
 When age trends are signficant, they are presented with numbered equations; otherwise, means are presented.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.](tables_figures/Temperate broadleaf YOUNG.png)
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
 
-![**Fig.
-10 | C cycle diagram for mature temperate deciduous forests.** Presented are mean±std, where geographically distinct areas are treated as the unit of replication.
-All units are Mg C ha^-1^ yr^-1^ (fluxes) or Mg C ha^-1^ (stocks).
+  knitr::include_graphics("tables_figures/Temperate broadleaf YOUNG.png")
+```
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 10 | C cycle diagram for young temperate conifer forests.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
 Arrows indicate fluxes, boxes indicate stocks.
 When age trends are signficant, they are presented with numbered equations; otherwise, means are presented.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.](tables_figures/Temperate conifer YOUNG.png)
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
 
-![**Fig.
-11 | C cycle diagram for mature boreal forests.** Presented are mean±std, where geographically distinct areas are treated as the unit of replication.
-All units are Mg C ha^-1^ yr^-1^ (fluxes) or Mg C ha^-1^ (stocks).
+  knitr::include_graphics("tables_figures/Temperate conifer YOUNG.png")
+```
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 11 | C cycle diagram for young boreal conifer forests.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
 Arrows indicate fluxes, boxes indicate stocks.
 When age trends are signficant, they are presented with numbered equations; otherwise, means are presented.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.](tables_figures/Boreal conifer YOUNG.png)
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/Boreal conifer YOUNG.png")
+```
+\elandscape
 

### 2020-02-21 manuscript/ERL_global_C_review.Rmd
added Figs 6 & 7

@@
 
+(**check paragraph with latest data**) The largest C fluxes—including $GPP$, $NPP$, $ANPP$, $BNPP$, $R_{soil_het}$, $R_{soil}$, $R_{eco}$, — were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests  ([ForC_variable_averages_per_Biome](https://github.com/forc-db/ForC/blob/master/numbers_and_facts/ForC_variable_averages_per_Biome.csv)) (Fig.
+6).
+The same held true for some of the subsidiary fluxes: ANPP_foliage, ANPP_woody, [**OTHERS?**].
+Other subsidiary fluxes—including NPP_woody, ANPP_repro, ANPP_stem, ANPP_branch, woody.mortality, BNPP_coarse, BNPP_fine,  [**OTHERS?**]—deviated from this pattern and/or lacked data for some biomes.
+Net ecosystem exchange (NEE) did not follow this pattern, with no significant differences across biomes but the largest (negative) average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests.
+Thus, C cycling rates generally decreased from tropical to temperate to boreal forests, but with less apparent trends for some of the subsidiary fluxes and an important exception in the overall C balance (NEE).
+
 \newpage
@@
 
-(**check paragraph with latest data**) The largest C fluxes—including $GPP$, $NPP$, $ANPP$, $BNPP$, $R_{soil_het}$, $R_{soil}$, $R_{eco}$, — were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests  ([ForC_variable_averages_per_Biome](https://github.com/forc-db/ForC/blob/master/numbers_and_facts/ForC_variable_averages_per_Biome.csv)) (Fig.
-6).
-The same held true for some of the subsidiary fluxes: ANPP_foliage, ANPP_woody, [**OTHERS?**].
-Other subsidiary fluxes—including NPP_woody, ANPP_repro, ANPP_stem, ANPP_branch, woody.mortality, BNPP_coarse, BNPP_fine,  [**OTHERS?**]—deviated from this pattern and/or lacked data for some biomes.
-Net ecosystem exchange (NEE) did not follow this pattern, with no significant differences across biomes but the largest (negative) average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests.
-Thus, C cycling rates generally decreased from tropical to temperate to boreal forests, but with less apparent trends for some of the subsidiary fluxes and an important exception in the overall C balance (NEE).
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEE$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Lines...
+.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/Figure6.png")
+```
 
@@
 There were some statistically significant biome differences in less frequently sampled C stocks (e.g., woody biomass, foliage biomass, deadwood), but given high within-biome variability in C stocks and relatively low sample sizes, these were likely attributable to sampling biases and methodological differences than to true differences across biomes.
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/Figure7.png")
+```
 

### 2020-02-21 manuscript/ERL_global_C_review.Rmd
format cleanup

@@
                   \newcommand{\elandscape}{\end{landscape}}
+
+                   \usepackage{setspace}
+                  \onehalfspacing
 
@@
 
-*Background.* The fate of Earth’s climate closely linked to forests, which strongly influence atmospheric carbon dioxide (CO2) and climate through their influential role in the global carbon (C) cycle.
+*Background.* The fate of Earth’s climate closely linked to forests, which strongly influence atmospheric carbon dioxide (CO~2~) and climate through their influential role in the global carbon (C) cycle.
 Synthetic understanding of global forest C cycles is needed to constrain model estimates of forest feedbacks to climate change and to more accurately quantify the influence of land use decisions on climate.
@@
 1).
-Their annual gross CO2 sequestration (gross primary productivity, GPP) is estimated at 59 Gt C yr-1 (Beer et al 2010), or 6.3 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017).
-A small portion of global terrestrial GPP is retained in ecosystems (mainly forests), resulting in a C sink that averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017).
-The remaining ~98% of global GPP is counterbalanced by ecosystem respiration (Reco) or wildfire.
-Perturbation to the global GPP- Reco balance can substantially influence the growth rate of atmospheric CO2; for example, the 2015-2016 El Niño, which brought historically high temperatures and low precipitation to the tropics, released an extra 2.5 ± 0.3 Gt C to the atmosphere, resulting in the largest recorded atmospheric CO2 growth rate (Le Quéré et al 2016, Liu et al 2017).
-In addition, forests contain substantial C stocks that, when disturbed, release significant amounts of CO2 to the atmosphere.
+Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at 59 Gt C yr-1 (Beer et al 2010), or 6.3 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017).
+A small portion of global terrestrial $GPP$ is retained in ecosystems (mainly forests), resulting in a C sink that averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017).
+The remaining ~98% of global $GPP$ is counterbalanced by ecosystem respiration (Reco) or wildfire.
+Perturbation to the global $GPP$- Reco balance can substantially influence the growth rate of atmospheric CO~2~; for example, the 2015-2016 El Niño, which brought historically high temperatures and low precipitation to the tropics, released an extra 2.5 ± 0.3 Gt C to the atmosphere, resulting in the largest recorded atmospheric CO~2~ growth rate (Le Quéré et al 2016, Liu et al 2017).
+In addition, forests contain substantial C stocks that, when disturbed, release significant amounts of CO~2~ to the atmosphere.
 Although they cover only ~30% of the land surface, forests contain an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C (biomass and soils; Bonan 2008).
 Globally, gross tropical deforestation averaged 2.8 Gt C yr-1 from 2000-2007, but ~40% of this was offset by forest regrowth, resulting in a net source of ~1.1 Gt C yr-1 from tropical land use change (Pan et al 2011).
-This, coupled with minimal net deforestation in the extratropics and net uptake by intact forests (Pan et al 2011, Schimel et al 2015), resulted in a total gain in forest C of ~1.2-1.7 Gt C yr-1 (Le Quéré et al 2017, Schimel et al 2015), thereby substantially slowing the rate of increase of atmospheric CO2.
+This, coupled with minimal net deforestation in the extratropics and net uptake by intact forests (Pan et al 2011, Schimel et al 2015), resulted in a total gain in forest C of ~1.2-1.7 Gt C yr-1 (Le Quéré et al 2017, Schimel et al 2015), thereby substantially slowing the rate of increase of atmospheric CO~2~.
 
 ![**Figure 1 | The role of forests in the global carbon cycle.
-Values apply to the period 2000-2018.** Sources are as follows: biomass- (Pan et al 2013, Baccini et al 2012) IPCC 2000; intact forest sink, net forest sink, tropical deforestation and regrowth-(Pan et al 2011); terrestrial sink- (Le Quéré et al 2017); GPP- (Beer et al 2010); respiration+fire: calculated here; fossil fuel emissions- (Le Quéré et al 2017); atmospheric CO2 (2018 value from https://scripps.ucsd.edu/programs/keelingcurve/).
+Values apply to the period 2000-2018.** Sources are as follows: biomass- (Pan et al 2013, Baccini et al 2012) IPCC 2000; intact forest sink, net forest sink, tropical deforestation and regrowth-(Pan et al 2011); terrestrial sink- (Le Quéré et al 2017); $GPP$- (Beer et al 2010); respiration+fire: calculated here; fossil fuel emissions- (Le Quéré et al 2017); atmospheric CO~2~ (2018 value from https://scripps.ucsd.edu/programs/keelingcurve/).
 **see issue 19: https://github.com/forc-db/ERL-review/issues/19** ](tables_figures/Global_C_cycle/Global C cycle diagram.png)
 
-Given their vital role in regulating atmospheric CO2, the future of Earth’s forests will strongly influence the course of climate change.
-If forests globally respond to the suite of contemporary and future global change drivers—including elevated CO2, climate change, and atmospheric deposition—with increased productivity and net biomass increases, they could continue to act as a significant buffer against anthropogenic emissions.
+Given their vital role in regulating atmospheric CO~2~, the future of Earth’s forests will strongly influence the course of climate change.
+If forests globally respond to the suite of contemporary and future global change drivers—including elevated CO~2~, climate change, and atmospheric deposition—with increased productivity and net biomass increases, they could continue to act as a significant buffer against anthropogenic emissions.
 In contrast, if factors such as higher temperatures and drought stress lead to net C losses, the C sink of forests could be lost and even reversed.
-Global coupled climate models vary substantially in predictions regarding the future of the global forest C sink, with uncertainty in end-of-century terrestrial NEP exceeding current annual anthropogenic CO2 emissions (Friedlingstein et al 2006, Cavaleri et al 2015).
-Anthropogenic land use decisions will also strongly influence the future course of atmospheric CO2.
-In the absence of measures to reduce deforestation, this will continue to be a significant component of anthropogenic CO2 emissions.
-On the other hand, forest conservation, reforestation, and forest management hold strong potential to help stabilize and eventually reduce total anthropogenic CO2 emissions (estimated total of 4.4 Gt C yr-1; Griscom et al 2017; *Cook-Patton et al.*; **other refs?**) and thereby contribute towards the goals of the Paris Climate Agreement (UNFCCC 2015, Houghton et al 2015; **REFS**).
+Global coupled climate models vary substantially in predictions regarding the future of the global forest C sink, with uncertainty in end-of-century terrestrial NEP exceeding current annual anthropogenic CO~2~ emissions (Friedlingstein et al 2006, Cavaleri et al 2015).
+Anthropogenic land use decisions will also strongly influence the future course of atmospheric CO~2~.
+In the absence of measures to reduce deforestation, this will continue to be a significant component of anthropogenic CO~2~ emissions.
+On the other hand, forest conservation, reforestation, and forest management hold strong potential to help stabilize and eventually reduce total anthropogenic CO~2~ emissions (estimated total of 4.4 Gt C yr-1; Griscom et al 2017; *Cook-Patton et al.*; **other refs?**) and thereby contribute towards the goals of the Paris Climate Agreement (UNFCCC 2015, Houghton et al 2015; **REFS**).
 Indeed, forest-based climate mitigation was a key component of Paris Climate Agreement commitments, totaling approximately one-fourth of nations’ planned emissions reductions (Grassi et al 2017; IPCC1.5).
@@
 A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling.
-Recent studies attempting to quantify the role of forests in the global C cycle (e.g., Pan et al 2011) have been hampered by insufficient data on regrowth rates of tropical forests worldwide, C in non-living pools (i.e., dead wood, litter, soil), and source/sink status of forests in some regions.
+Recent studies attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011) have been hampered by insufficient data on regrowth rates of tropical forests worldwide, C in non-living pools (*i.e.*, dead wood, litter, soil), and source/sink status of forests in some regions.
 Moreover, as discussed above, model representation of forest responses to global change remain highly uncertain (Friedlingstein et al 2006, Cavaleri et al 2015), with efforts to reduce model uncertainty in part limited by ready availability of appropriate benchmark data (Clark et al 2017, Luo et al 2012).
 To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018).
-ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (e.g., stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications.
+ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (*e.g.*, stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications.
 ForC currently contains **#** records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.
@@
 
-1. How do C cycling vary across the worlds major forest biomes (i.e., tropical, temperate, boreal)?
+1. How do C cycling vary across the worlds major forest biomes (*i.e.*, tropical, temperate, boreal)?
 
@@
 
-While components of these questions have been previously addressed (e.g., REFS(Luyssaert et al 2007, Anderson et al 2006, Anderson-Teixeira et al 2013), our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.
+While components of these questions have been previously addressed (*e.g.*, REFS(Luyssaert et al 2007, Anderson et al 2006, Anderson-Teixeira et al 2013), our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.
 
@@
 This review synthesizes data from the ForC database (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018).
-ForC amalgamates numerous intermediary data sets (e.g., REFS) and original studies.
+ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies.
 Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records.
 The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age.
-As such, there has been a focus on incorporating data from regrowth forests (e.g., Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible (83% of records in v.2.0; Anderson-Teixeira et al 2018).
+As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible (83% of records in v.2.0; Anderson-Teixeira et al 2018).
 Particular attention was given to developing the database for tropical forests (Anderson-Teixeira et al 2016), yet these represented only approximately one-third of records in ForC v.2.0 (Anderson-Teixeira et al 2018).
 Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database (Bond-Lamberty and Thomson 2010; *hopefully Cook-Patton*; see [GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv)) ….
-We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (e.g.,: REFS).
+We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS).
 The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: XX).
@@
 We drew upon records for *#* annual flux and *#* C stock variables (Table 1).
-For this analysis, we combined some of ForC’s specific variables (e.g., multiple variables for net primary productivity including various components) into more broadly defined variables (Table 1, [this table](https://github.com/forc-db/ForC/blob/master/figures/C_cycle_diagrams/ForC_variables_mapping_for_C_cycle_diagrams.csv))).
-Although ForC contains information that may be used to standardize or control for methodological differences (e.g., area sampled, min stem diameter sampled, allometric equations applied; Anderson-Teixeira et al 2018), for this analysis we included all relevant data in ForC.
-Throughout ForC, for all measurements drawing from tree census data (e.g., biomass, productivity), the minimum stem diameter sampled was $\le$ 10cm.
+For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables (Table 1, [this table](https://github.com/forc-db/ForC/blob/master/figures/C_cycle_diagrams/ForC_variables_mapping_for_C_cycle_diagrams.csv))).
+Although ForC contains information that may be used to standardize or control for methodological differences (*e.g.*, area sampled, min stem diameter sampled, allometric equations applied; Anderson-Teixeira et al 2018), for this analysis we included all relevant data in ForC.
+Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum stem diameter sampled was $\le$ 10cm.
 All records were measured directly or derived from field measurements (as opposed to modeled).
@@
 Duplicate or otherwise conflicting records (indicated in conflicts field of MEASUREMENTS table) were reconciled as follows.
-Replicate measurements (i.e., replicates from within a single study) were averaged.
-Records that subsumed others— *i.e.*, the time period included that of $\ge$ 2 other records or dates were unknown and therefore conflicted with $\ge$ 2 other records—were removed.
-For each group of duplicate records—i.e., measurements of the same variable in the same plot at the same time—one record was assigned precedence (recorded in D.precedence field).
+Replicate measurements (*i.e.*, replicates from within a single study) were averaged.
+Records that subsumed others— **i.e.**, the time period included that of $\ge$ 2 other records or dates were unknown and therefore conflicted with $\ge$ 2 other records—were removed.
+For each group of duplicate records—*i.e.*, measurements of the same variable in the same plot at the same time—one record was assigned precedence (recorded in D.precedence field).
 When measurement periods overlapped or were not specified, precedence was given first to records representing longer measurement periods (*i.e.*, end.date - start.date) and then to more recently published values.
-We manually reviewed duplicates that differed only in methodology, assigning precedence to the record employing a more comprehensive approach (e.g., inclusion of understory, lianas, or bamboo as opposed to just trees) or using a favored methodology.
+We manually reviewed duplicates that differed only in methodology, assigning precedence to the record employing a more comprehensive approach (*e.g.*, inclusion of understory, lianas, or bamboo as opposed to just trees) or using a favored methodology.
 
@@
 Specifically, we removed all plots flagged as managed in ForC-simplified (managed field).
-This included plots with any record of managements manipulating CO2, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.
+This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.
 Plots flagged as disturbed in ForC-simplified included stands that had undergone anthropogenic thinning or partial harvest (“Cut” or “Harvest” codes) unless this was very minor (percent.mortality= “minor”).
@@
 Data were analyzed to produce basic summaries of C cycle patterns across biomes and stand ages following an approach similar to that of Anderson-Teixeira et al (2016).
-For mature forests, to obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot, weighting flux measurements according to the length of measurement periods (i.e., end.date - start.date).
+For mature forests, to obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot, weighting flux measurements according to the length of measurement periods (*i.e.*, end.date - start.date).
 Values were then averaged across plots clustered within 25 km of one another (geographic.area field of SITES table, sensu Anderson-Teixeira et al 2018), weighting by area.sampled (MEASUREMENTS table) or plot.area (PLOTS table) if available for all records.
@@
 4).
-**This is based on my quick scan of the figures.
-Need to add a formal automated test: https://github.com/forc-db/ERL-review/issues/16 **
-
-(**check paragraph with latest data**) The largest C fluxes—including $GPP$, $NPP$, $ANPP$, $BNPP$, $R_{soil_het}$, $R_{soil}$, $R_{eco}$, — were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests  ([ForC_variable_averages_per_Biome](https://github.com/forc-db/ForC/blob/master/numbers_and_facts/ForC_variable_averages_per_Biome.csv)) (Fig.
-6).
-The same held true for some of the subsidiary fluxes: ANPP_foliage, ANPP_woody, [**OTHERS?**].
-Other subsidiary fluxes—including NPP_woody, ANPP_repro, ANPP_stem, ANPP_branch, woody.mortality, BNPP_coarse, BNPP_fine,  [**OTHERS?**]—deviated from this pattern and/or lacked data for some biomes.
-Net ecosystem exchange (NEE) did not follow this pattern, with no significant differences across biomes but the largest (negative) average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests.
-Thus, C cycling rates generally decreased from tropical to temperate to boreal forests, but with less apparent trends for some of the subsidiary fluxes and an important exception in the overall C balance (NEE).
+**update this: https://github.com/forc-db/ERL-review/issues/16 **
 
@@
 
+(**check paragraph with latest data**) The largest C fluxes—including $GPP$, $NPP$, $ANPP$, $BNPP$, $R_{soil_het}$, $R_{soil}$, $R_{eco}$, — were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests  ([ForC_variable_averages_per_Biome](https://github.com/forc-db/ForC/blob/master/numbers_and_facts/ForC_variable_averages_per_Biome.csv)) (Fig.
+6).
+The same held true for some of the subsidiary fluxes: ANPP_foliage, ANPP_woody, [**OTHERS?**].
+Other subsidiary fluxes—including NPP_woody, ANPP_repro, ANPP_stem, ANPP_branch, woody.mortality, BNPP_coarse, BNPP_fine,  [**OTHERS?**]—deviated from this pattern and/or lacked data for some biomes.
+Net ecosystem exchange (NEE) did not follow this pattern, with no significant differences across biomes but the largest (negative) average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests.
+Thus, C cycling rates generally decreased from tropical to temperate to boreal forests, but with less apparent trends for some of the subsidiary fluxes and an important exception in the overall C balance (NEE).
+
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEE$.
@@
 Maximum aboveground biomass values followed a very different trend than the means: temperate broadleaf> temperate conifer>boreal>tropical.
-There were some statistically significant biome differences in less frequently sampled C stocks (e.g., woody biomass, foliage biomass, deadwood), but given high within-biome variability in C stocks and relatively low sample sizes, these were likely attributable to sampling biases and methodological differences than to true differences across biomes.
+There were some statistically significant biome differences in less frequently sampled C stocks (*e.g.*, woody biomass, foliage biomass, deadwood), but given high within-biome variability in C stocks and relatively low sample sizes, these were likely attributable to sampling biases and methodological differences than to true differences across biomes.
 
@@
 6-7 and **S#- SI figures including plots for all variables**).
-Of these, 9 increased significantly with log10[stand.age]: GPP, NPP, ANPP, ANPP_foliage, ANPP_woody, ANPP_woody_stem, BNPP, BNPP_root_fine, Reco, and net C sequestration (-NEE).
-The remaining five--ANPP_woody_branch, BNPP_root_coarse, soil heterotrophic respiration, and total soil respiration–displayed no significant relationship to stand age, although all displayed a positive trend.
+Of these, 9 increased significantly with log10[stand.age]: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($-NEE$).
+The remaining five--$ANPP_{woody-branch}$, $BNPP_{root-coarse}$, $R_{soil-het}$, and $R_{soil-het}$-–displayed no significant relationship to stand age, although all displayed a positive trend.
 In terms of C stocks, 10 variables had sufficient data to test for age trends.
-Six of these—total biomass, aboveground biomass, aboveground woody biomass, foliage biomass, root biomass, and coarse root biomass—increased significantly with log10[stand.age].
+Six of these—total biomass, aboveground biomass, aboveground woody biomass, foliage biomass, root biomass, and coarse root biomass—increased significantly with $log10[stand.age]$.
 The remaining four displayed non-significant positive trends:  fine root biomass, total dead wood, standing dead wood, and organic layer.

### 2020-02-22 manuscript/ERL_global_C_review.Rmd
Update ERL_global_C_review.Rmd
added Bagdley, #20

@@
 1).
-Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at 59 Gt C yr-1 (Beer et al 2010), or 6.3 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017).
+Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at 59 Gt C yr-1 (Beer et al 2010){**Bagdley et al.
+is better estimate: >69, also reviews other more recent estimates**}, or 6.3 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017).
 A small portion of global terrestrial $GPP$ is retained in ecosystems (mainly forests), resulting in a C sink that averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017).

### 2020-02-23 manuscript/ERL_global_C_review.Rmd
fixed up Table 1

@@
 Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd).
+Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis.
 Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations (SYNMAP; Jung et al 2006) and recorded in the  dominant.veg field of the ForC MEAUREMENTS table.
@@
 S1 <- read.csv("tables_figures/C_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S1, format = "latex", booktabs = TRUE, caption = "Table 1.
-Carbon cycle variables incuded in this analysis.
-Complete list of variables with full definitions is available at [GitHub url].", escape = F)%>%
+kable(S1, format = "latex", booktabs = TRUE, caption = "Table 1. Carbon cycle variables included in this analysis, their sample sizes, and summary of biome differences and age trends.", escape = F)%>%
+  # footnote(general = "Complete list of variables with full definitions and association to ForC variables is given in [XXXXX](https://github.com/forc-db/ForC/blob/master/figures/C_cycle_diagrams/ForC_variables_mapping_for_C_cycle_diagrams.csv).
+") %>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12)  %>%
+  add_header_above(c(" ", " ", "N records" = 3, " " , " " )) %>%
+  #column_spec(1, width = "4cm") %>%
+  column_spec(2, width = "7cm") %>%
+  column_spec(3, width = "1.1cm") %>%
+  column_spec(4, width = "1cm") %>%
+  column_spec(5, width = "1.8cm") %>%
   pack_rows("Annual fluxes", 1, 23, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%

### 2020-02-23 manuscript/ERL_global_C_review.Rmd
removed Fig.1, started reworking intro
#19

global C cycle figure files moved to: https://github.com/teixeirak/Global_C_cycle

@@
 
-Forest ecosystems globally influence climate through their critical role in the global carbon (C) cycle (Fig.
-1).
-Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at 59 Gt C yr-1 (Beer et al 2010){**Bagdley et al.
-is better estimate: >69, also reviews other more recent estimates**}, or 6.3 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017).
-A small portion of global terrestrial $GPP$ is retained in ecosystems (mainly forests), resulting in a C sink that averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017).
-The remaining ~98% of global $GPP$ is counterbalanced by ecosystem respiration (Reco) or wildfire.
-Perturbation to the global $GPP$- Reco balance can substantially influence the growth rate of atmospheric CO~2~; for example, the 2015-2016 El Niño, which brought historically high temperatures and low precipitation to the tropics, released an extra 2.5 ± 0.3 Gt C to the atmosphere, resulting in the largest recorded atmospheric CO~2~ growth rate (Le Quéré et al 2016, Liu et al 2017).
-In addition, forests contain substantial C stocks that, when disturbed, release significant amounts of CO~2~ to the atmosphere.
-Although they cover only ~30% of the land surface, forests contain an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C (biomass and soils; Bonan 2008).
-Globally, gross tropical deforestation averaged 2.8 Gt C yr-1 from 2000-2007, but ~40% of this was offset by forest regrowth, resulting in a net source of ~1.1 Gt C yr-1 from tropical land use change (Pan et al 2011).
-This, coupled with minimal net deforestation in the extratropics and net uptake by intact forests (Pan et al 2011, Schimel et al 2015), resulted in a total gain in forest C of ~1.2-1.7 Gt C yr-1 (Le Quéré et al 2017, Schimel et al 2015), thereby substantially slowing the rate of increase of atmospheric CO~2~.
+Carbon (C) cycling in forests is central to sustaining life on Earth.
+Forest ecosystems globally influence climate through their vital role in regulating atmospheric carbon dioxide (CO~2~).
+Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ (Bagdley et al.
+2019), or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**).
+While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion was retained in ecosystems over recent decades.
+The resulting in C sink averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**).
+Moreover, forests contain substantial C stocks: an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C (biomass and soils; Bonan 2008).
+Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), resulting in net C sink of ~1.2-1.7 Gt C yr^-1^ across Earth's forests (Le Quéré et al 2017, Schimel et al 2015) (**update, give range**).
+The future of this sink is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change.
+Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change (Friedlingstein et al 2006, Cavaleri et al 2015, Griscom et al 2017; Grassi et al 2017; IPCC1.5).
 
-![**Figure 1 | The role of forests in the global carbon cycle.
-Values apply to the period 2000-2018.** Sources are as follows: biomass- (Pan et al 2013, Baccini et al 2012) IPCC 2000; intact forest sink, net forest sink, tropical deforestation and regrowth-(Pan et al 2011); terrestrial sink- (Le Quéré et al 2017); $GPP$- (Beer et al 2010); respiration+fire: calculated here; fossil fuel emissions- (Le Quéré et al 2017); atmospheric CO~2~ (2018 value from https://scripps.ucsd.edu/programs/keelingcurve/).
-**see issue 19: https://github.com/forc-db/ERL-review/issues/19** ](tables_figures/Global_C_cycle/Global C cycle diagram.png)
+On an ecosystem level, forest C cycling is central to the flow of energy and materials, and as the basis of food webs...
 
-Given their vital role in regulating atmospheric CO~2~, the future of Earth’s forests will strongly influence the course of climate change.
-If forests globally respond to the suite of contemporary and future global change drivers—including elevated CO~2~, climate change, and atmospheric deposition—with increased productivity and net biomass increases, they could continue to act as a significant buffer against anthropogenic emissions.
-In contrast, if factors such as higher temperatures and drought stress lead to net C losses, the C sink of forests could be lost and even reversed.
-Global coupled climate models vary substantially in predictions regarding the future of the global forest C sink, with uncertainty in end-of-century terrestrial NEP exceeding current annual anthropogenic CO~2~ emissions (Friedlingstein et al 2006, Cavaleri et al 2015).
-Anthropogenic land use decisions will also strongly influence the future course of atmospheric CO~2~.
-In the absence of measures to reduce deforestation, this will continue to be a significant component of anthropogenic CO~2~ emissions.
-On the other hand, forest conservation, reforestation, and forest management hold strong potential to help stabilize and eventually reduce total anthropogenic CO~2~ emissions (estimated total of 4.4 Gt C yr-1; Griscom et al 2017; *Cook-Patton et al.*; **other refs?**) and thereby contribute towards the goals of the Paris Climate Agreement (UNFCCC 2015, Houghton et al 2015; **REFS**).
-Indeed, forest-based climate mitigation was a key component of Paris Climate Agreement commitments, totaling approximately one-fourth of nations’ planned emissions reductions (Grassi et al 2017; IPCC1.5).
-Thus, forest will strongly influence the future course of climate change both through their responses to global change and through forest-related anthropogenic land use decisions.
+Despite the centrality of forest C cycling in regulating both atmospheric CO~2~ and ecosystem function and diversity, we lack a comprehensive understanding of how C cycling varies with across biomes and with stand age.
+While satellite measurements are increasingly useful for estimating forest $GPP$ (Bagdley et al.
+2019), aboveground biomass ($B_{ag}$) (REFS), and possibly woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements.
+...
 
-While forests will play a critical role in the future course of climate change, there remain significant uncertainties as to current and future C cycling the world’s forests.
-Even as the climate changes and forest age distributions shift, there are important gaps in our fundamental understanding of how C cycling varies with climate and stand age.
-A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling.
+*A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling.
 Recent studies attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011) have been hampered by insufficient data on regrowth rates of tropical forests worldwide, C in non-living pools (*i.e.*, dead wood, litter, soil), and source/sink status of forests in some regions.
-Moreover, as discussed above, model representation of forest responses to global change remain highly uncertain (Friedlingstein et al 2006, Cavaleri et al 2015), with efforts to reduce model uncertainty in part limited by ready availability of appropriate benchmark data (Clark et al 2017, Luo et al 2012).
+Moreover, as discussed above, model representation of forest responses to global change remain highly uncertain (Friedlingstein et al 2006, Cavaleri et al 2015), with efforts to reduce model uncertainty in part limited by ready availability of appropriate benchmark data (Clark et al 2017, Luo et al 2012).*
+
 To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018).
@@
 
-Here, we synthesize ForC data to provide a big-picture overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age. We address two broad questions:
+Here, we synthesize ForC data to provide a macroscopic overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age. We address two broad questions:
 

### 2020-02-23 manuscript/ERL_global_C_review.Rmd
worked on intro

@@
 2019), aboveground biomass ($B_{ag}$) (REFS), and possibly woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements.
-...
+Moreover, data on stand characteristics--including species composition, age, and disturbance history--are critical to interpreting and projecting into the future, but [usually rely on ground-based measurements].
+Thus, ground-based measurements are--and will continue to be--central to characterizing forest C cycling.
+Tens of thousands of forest C measurements have been published, but their distribution across literally thousands of scientific articles-- along with variation in data formats, units, measurement methods, *etc.*--have made them effectively inaccesible for many global-scale analyses, including those attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011), and model evaluation(Clark et al 2017, Luo et al 2012).
+Important progress has been made in synthesizing global data to address how C cycling varies with respect to climate (Luyssaert et al.
+2007; REFS) and stand age (REFS),
 
-*A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling.
-Recent studies attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011) have been hampered by insufficient data on regrowth rates of tropical forests worldwide, C in non-living pools (*i.e.*, dead wood, litter, soil), and source/sink status of forests in some regions.
-Moreover, as discussed above, model representation of forest responses to global change remain highly uncertain (Friedlingstein et al 2006, Cavaleri et al 2015), with efforts to reduce model uncertainty in part limited by ready availability of appropriate benchmark data (Clark et al 2017, Luo et al 2012).*
+*A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling.*
 

### 2020-02-23 manuscript/ERL_global_C_review.Rmd
worked on intro

@@
 Thus, ground-based measurements are--and will continue to be--central to characterizing forest C cycling.
-Tens of thousands of forest C measurements have been published, but their distribution across literally thousands of scientific articles-- along with variation in data formats, units, measurement methods, *etc.*--have made them effectively inaccesible for many global-scale analyses, including those attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011), and model evaluation(Clark et al 2017, Luo et al 2012).
-Important progress has been made in synthesizing global data to address how C cycling varies with respect to climate (Luyssaert et al.
-2007; REFS) and stand age (REFS),
-
-*A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling.*
-
+Tens of thousands of forest C measurements have been published, but their distribution across literally thousands of scientific articles-- along with variation in data formats, units, measurement methods, *etc.*--have made them effectively inaccesible for many global-scale analyses, including those attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011), and model evaluation (Clark et al 2017, Luo et al 2012).
+Important progress has been made in synthesizing data to address how C cycling varies across forest biomes (Luyssaert et al.
+2007; REFS) and stand ages (REFS), yet generally consider only a limited set of C variables (but see Anderson-Teixeira et al 2016 for the tropics) and do not consider age-by-biome interactions.
 To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018).
@@
 
+A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling.
+
 ### Acknowledgements

### 2020-02-23 manuscript/ERL_global_C_review.Rmd
minor

@@
 
-*Background.* The fate of Earth’s climate closely linked to forests, which strongly influence atmospheric carbon dioxide (CO~2~) and climate through their influential role in the global carbon (C) cycle.
-Synthetic understanding of global forest C cycles is needed to constrain model estimates of forest feedbacks to climate change and to more accurately quantify the influence of land use decisions on climate.
+*Background.* *The fate of Earth’s climate closely linked to forests, which strongly influence atmospheric carbon dioxide (CO~2~) and climate through their influential role in the global carbon (C) cycle.
+Synthetic understanding of global forest C cycles is needed to constrain model estimates of forest feedbacks to climate change and to more accurately quantify the influence of land use decisions on climate.*
 

### 2020-02-25 manuscript/ERL_global_C_review.Rmd
update doc

@@
 Kristina J. Anderson-Teixeira^1,2^*
+
 Valentine Herrmann^1^
+
 Becky Banbury Morgan
+
 Ben Bond-Lamberty
+
 Susan Cook-Patton
+
 Abigail Ferson
+
 Norbert Kunert
+
 Jennifer McGarvey
+
 Helene C. Muller-Landau^1^
+
 Maria Wang
@@
 Despite the centrality of forest C cycling in regulating both atmospheric CO~2~ and ecosystem function and diversity, we lack a comprehensive understanding of how C cycling varies with across biomes and with stand age.
-While satellite measurements are increasingly useful for estimating forest $GPP$ (Bagdley et al.
-2019), aboveground biomass ($B_{ag}$) (REFS), and possibly woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements.
+While satellite measurements are increasingly useful for global estimates of forest $GPP$ (Bagdley et al.
+2019, [Li & Xiao 2019](https://www.mdpi.com/2072-4292/11/21/2563)), aboveground biomass ($B_{ag}$) (REFS), and possibly woody mortality (**?** *i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements.
 Moreover, data on stand characteristics--including species composition, age, and disturbance history--are critical to interpreting and projecting into the future, but [usually rely on ground-based measurements].
@@
 Arrows indicate fluxes, boxes indicate stocks.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
+Arrows are scaled as the **square root of flux divided by 5**.", out.width='100%', fig.pos='H'}
 

### 2020-02-28 manuscript/ERL_global_C_review.Rmd
tiny formatting change

@@
   column_spec(5, width = "1.8cm") %>%
-  pack_rows("Annual fluxes", 1, 23, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
-  pack_rows("Stocks", 24, 34, latex_gap_space = "1em", colnum = 1, hline_before = FALSE)
+  pack_rows("Annual fluxes", 1, 23, colnum = 2, hline_before = FALSE) %>%
+  pack_rows("Stocks", 24, 34, colnum = 2, hline_before = FALSE)
 ```

### 2020-02-29 manuscript/ERL_global_C_review.Rmd
added Table 1 footnotes

@@
 S1 <- read.csv("tables_figures/C_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
+names(S1) [6]  <- paste0(names(S1) [6], footnote_marker_symbol(1))
+names(S1) [7]  <- paste0(names(S1) [7], footnote_marker_symbol(2))
 kable(S1, format = "latex", booktabs = TRUE, caption = "Table 1. Carbon cycle variables included in this analysis, their sample sizes, and summary of biome differences and age trends.", escape = F)%>%
@@
   pack_rows("Annual fluxes", 1, 23, colnum = 2, hline_before = FALSE) %>%
-  pack_rows("Stocks", 24, 34, colnum = 2, hline_before = FALSE)
+  pack_rows("Stocks", 24, 34, colnum = 2, hline_before = FALSE) %>%
+kableExtra::footnote(symbol = c("Tr: Tropical, TeB: Temperate Broadleaf, TeN: Temperate Needleleaf, B: Boreal, n.s.: no significant differences", "+ or -: significant positive or negative trend, xB: significant age x biome interaction, n.s.: no significant age trend"), footnote_as_chunk = FALSE)
 ```

### 2020-03-10 manuscript/ERL_global_C_review.Rmd
minor edits

@@
 
-                  \usepackage{setspace}
-                  \onehalfspacing
+                  \usepackage{setspace}\onehalfspacing
+                  \usepackage{lineno}\linenumbers
 
@@
 Despite the centrality of forest C cycling in regulating both atmospheric CO~2~ and ecosystem function and diversity, we lack a comprehensive understanding of how C cycling varies with across biomes and with stand age.
-While satellite measurements are increasingly useful for global estimates of forest $GPP$ (Bagdley et al.
-2019, [Li & Xiao 2019](https://www.mdpi.com/2072-4292/11/21/2563)), aboveground biomass ($B_{ag}$) (REFS), and possibly woody mortality (**?** *i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements.
+While remote sensing measurements are increasingly useful for global- or regional-scale estimates of forest $GPP$ (Bagdley et al.
+2019, [Li & Xiao 2019](https://www.mdpi.com/2072-4292/11/21/2563)), aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) ([Clark et al.
+2004](https://www.onlinelibrary.wiley.com/doi/abs/10.1046/j.1461-0248.2003.00547.x); [Leithold et al 2018](https://nph-onlinelibrary-wiley-com.smithsonian.idm.oclc.org/doi/10.1111/nph.15110)), and to some extent net ecosystem exchange ($NEE$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements.
 Moreover, data on stand characteristics--including species composition, age, and disturbance history--are critical to interpreting and projecting into the future, but [usually rely on ground-based measurements].

### 2020-09-05 manuscript/ERL_global_C_SI.Rmd
manuscript updates

--- 2020-02-28 manuscript/ERL_global_C_review.Rmd
@@
   pdf_document:
-    citation_package: natbib
     fig_caption: yes
-bibliography: [library.bib,  packages.bib]
-biblio-style: apalike
-csl: apa.csl
+bibliography: references.bib
+csl: environmental-research-letters.csl
 
@@
 
-**Title:** Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database
+**Title:** Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database
 
-**Authors:** 
-Kristina J. Anderson-Teixeira^1,2^*
-
-Valentine Herrmann^1^
-
-Becky Banbury Morgan
-
-Ben Bond-Lamberty
-
-Susan Cook-Patton
-
-Abigail Ferson
-
-Norbert Kunert
-
-Jennifer McGarvey
-
-Helene C. Muller-Landau^1^
-
-Maria Wang
-
-**Author Affiliations:**
-
-1. Conservation Ecology Center; Smithsonian Conservation Biology Institute; National Zoological Park, Front Royal, VA 22630, USA
-2. Center for Tropical Forest Science-Forest Global Earth Observatory; Smithsonian Tropical Research Institute; Panama, Republic of Panama
-
-*corresponding author: teixeirak@si.edu; +1 540 635 6546
+Kristina J. Anderson-Teixeira,
+Valentine Herrmann,
+Becky Banbury Morgan,
+Ben Bond-Lamberty,
+Susan Cook-Patton,
+Abigail Ferson,
+Jennifer McGarvey,
+Helene C. Muller-Landau,
+Maria Wang,
 
 \newpage
+## Appendix A. Duplicates and Conflicting Records within ForC
 
-### Summary
-*(The Summary should be no longer than 300 words and divided into the following sections: Background, Methods/Design, Review results/ Synthesis and Discussion.
-Up to 7 Keywords should also be provided.
-The total length of the article is flexible.)*
+### Status of duplicates and conflicting records within ForC
 
-*Background.* *The fate of Earth’s climate closely linked to forests, which strongly influence atmospheric carbon dioxide (CO~2~) and climate through their influential role in the global carbon (C) cycle.
-Synthetic understanding of global forest C cycles is needed to constrain model estimates of forest feedbacks to climate change and to more accurately quantify the influence of land use decisions on climate.*
-
-*Methods/Design.* Here, we draw from the Global Forest C Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation.
-Specifically, we draw upon ## records from ## geographic locations representing ## C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests.
-For regrowth forests, we quantify age trends for all variables.
-
-*Review Results/ Synthesis.*  The rate of C cycling generally increased from boreal to tropical regions, whereas C stocks showed less directional variation.
-The majority of flux variables, together with most live biomass pools, increased significantly with stand age, and the rate of increase again tended to increase from boreal to tropical regions.
-
-*Discussion.* *[Discussion section will interpret results, highlighting new and significant findings, and discuss implications.
-Tentative headings are “Stand level C cycling in forests globally”, “Age trends in C cycling”, and “Implications for climate change mitigation”.]*
-
-*Key words*: forest ecosystems; carbon cycle; stand age; productivity; respiration; biomass; global
-
-### Background
-*(The Background section needs to present the rationale for why a systematic review of this topic is needed along with a history of what has been done to date and an expectation of what new will emerge from the review, especially if quantitative meta-analyses of studies are being considered.)*
-
-Carbon (C) cycling in forests is central to sustaining life on Earth.
-Forest ecosystems globally influence climate through their vital role in regulating atmospheric carbon dioxide (CO~2~).
-Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ (Bagdley et al.
-2019), or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**).
-While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion was retained in ecosystems over recent decades.
-The resulting in C sink averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**).
-Moreover, forests contain substantial C stocks: an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C (biomass and soils; Bonan 2008).
-Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), resulting in net C sink of ~1.2-1.7 Gt C yr^-1^ across Earth's forests (Le Quéré et al 2017, Schimel et al 2015) (**update, give range**).
-The future of this sink is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change.
-Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change (Friedlingstein et al 2006, Cavaleri et al 2015, Griscom et al 2017; Grassi et al 2017; IPCC1.5).
-
-On an ecosystem level, forest C cycling is central to the flow of energy and materials, and as the basis of food webs...
-
-Despite the centrality of forest C cycling in regulating both atmospheric CO~2~ and ecosystem function and diversity, we lack a comprehensive understanding of how C cycling varies with across biomes and with stand age.
-While satellite measurements are increasingly useful for global estimates of forest $GPP$ (Bagdley et al.
-2019, [Li & Xiao 2019](https://www.mdpi.com/2072-4292/11/21/2563)), aboveground biomass ($B_{ag}$) (REFS), and possibly woody mortality (**?** *i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements.
-Moreover, data on stand characteristics--including species composition, age, and disturbance history--are critical to interpreting and projecting into the future, but [usually rely on ground-based measurements].
-Thus, ground-based measurements are--and will continue to be--central to characterizing forest C cycling.
-Tens of thousands of forest C measurements have been published, but their distribution across literally thousands of scientific articles-- along with variation in data formats, units, measurement methods, *etc.*--have made them effectively inaccesible for many global-scale analyses, including those attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011), and model evaluation (Clark et al 2017, Luo et al 2012).
-Important progress has been made in synthesizing data to address how C cycling varies across forest biomes (Luyssaert et al.
-2007; REFS) and stand ages (REFS), yet generally consider only a limited set of C variables (but see Anderson-Teixeira et al 2016 for the tropics) and do not consider age-by-biome interactions.
-To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018).
-ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (*e.g.*, stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications.
-ForC currently contains **#** records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.
-
-Here, we synthesize ForC data to provide a macroscopic overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age. We address two broad questions:
-
-1. How do C cycling vary across the worlds major forest biomes (*i.e.*, tropical, temperate, boreal)?
-
-2. How does C cycling vary with stand age?
-
-While components of these questions have been previously addressed (*e.g.*, REFS(Luyssaert et al 2007, Anderson et al 2006, Anderson-Teixeira et al 2013), our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.
-
-### Methods/ Design
-*(The Methods/Design section needs to describe how articles in the reviews were identified and what criteria were used for justifying inclusion in the review.
-While traditional reviews relying on past experience and expert knowledge are acceptable, editors should encourage reviews that are set up as so-called ‘systematic reviews’ (see the Cochrane Review procedure which initiated the process in the medical and health science: http://community.cochrane.org/ about-us/evidence-based-health-care).)*
-
-This review synthesizes data from the ForC database (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018).
-ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies.
-Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records.
-The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age.
-As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible (83% of records in v.2.0; Anderson-Teixeira et al 2018).
-Particular attention was given to developing the database for tropical forests (Anderson-Teixeira et al 2016), yet these represented only approximately one-third of records in ForC v.2.0 (Anderson-Teixeira et al 2018).
-Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database (Bond-Lamberty and Thomson 2010; *hopefully Cook-Patton*; see [GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv)) ….
-We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS).
-The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: XX).
-
-For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
-Climate zones were defined according to Köppen-Geiger zones, which were extracted based on site geographic coordinates from the ESRI Köppen-Geiger map (downloaded June 2014 from http://maps3.arcgisonline.com/ArcGIS/rest/services/A-16/Köppen- Geiger_Observed_and_Predicted_Climate_Shifts/MapServer; Rubel and Kottek 2010) and are recorded in the Koeppen field of the ForC SITES table.
-Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd).
-Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis.
-Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations (SYNMAP; Jung et al 2006) and recorded in the  dominant.veg field of the ForC MEAUREMENTS table.
-Stand age was obtained from the stand.age field of MEASUREMENTS table, and forests classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication).
-Records for which stand age was unknown were excluded from the analysis.
-These groupings were defined in March 2018, at which point they covered 91% of the primary variable records for forests of known age, or 86% of total records, in ForC (v.2.0; Anderson-Teixeira et al 2018).
-The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with <400 records total for stands of any age.
-
-We drew upon records for *#* annual flux and *#* C stock variables (Table 1).
-For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables (Table 1, [this table](https://github.com/forc-db/ForC/blob/master/figures/C_cycle_diagrams/ForC_variables_mapping_for_C_cycle_diagrams.csv))).
-Although ForC contains information that may be used to standardize or control for methodological differences (*e.g.*, area sampled, min stem diameter sampled, allometric equations applied; Anderson-Teixeira et al 2018), for this analysis we included all relevant data in ForC.
-Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum stem diameter sampled was $\le$ 10cm.
-All records were measured directly or derived from field measurements (as opposed to modeled).
-
-```{r eval = TRUE, echo=FALSE, warning=FALSE}
-library(knitr)
-library(kableExtra)
-S1 <- read.csv("tables_figures/C_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S1, format = "latex", booktabs = TRUE, caption = "Table 1. Carbon cycle variables included in this analysis, their sample sizes, and summary of biome differences and age trends.", escape = F)%>%
-  # footnote(general = "Complete list of variables with full definitions and association to ForC variables is given in [XXXXX](https://github.com/forc-db/ForC/blob/master/figures/C_cycle_diagrams/ForC_variables_mapping_for_C_cycle_diagrams.csv).
-") %>%
-  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12)  %>%
-  add_header_above(c(" ", " ", "N records" = 3, " " , " " )) %>%
-  #column_spec(1, width = "4cm") %>%
-  column_spec(2, width = "7cm") %>%
-  column_spec(3, width = "1.1cm") %>%
-  column_spec(4, width = "1cm") %>%
-  column_spec(5, width = "1.8cm") %>%
-  pack_rows("Annual fluxes", 1, 23, colnum = 2, hline_before = FALSE) %>%
-  pack_rows("Stocks", 24, 34, colnum = 2, hline_before = FALSE)
-```
-
-Analyses drew from ForC-simplified (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which is a rearrangement of ForC intended to facilitate analyses.
-In generating ForC-simplified, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006).
-Duplicate or otherwise conflicting records (indicated in conflicts field of MEASUREMENTS table) were reconciled as follows.
+### Generating ForC_simplified
 Replicate measurements (*i.e.*, replicates from within a single study) were averaged.
@@
 We manually reviewed duplicates that differed only in methodology, assigning precedence to the record employing a more comprehensive approach (*e.g.*, inclusion of understory, lianas, or bamboo as opposed to just trees) or using a favored methodology.
-
-Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand establishment (*i.e.*, that reflected by stand.age).
-Specifically, we removed all plots flagged as managed in ForC-simplified (managed field).
-This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.
-Plots flagged as disturbed in ForC-simplified included stands that had undergone anthropogenic thinning or partial harvest (“Cut” or “Harvest” codes) unless this was very minor (percent.mortality= “minor”).
-We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods.
-We also removed all plots for which no stand history information had been retrieved.
-
-Data were analyzed to produce basic summaries of C cycle patterns across biomes and stand ages following an approach similar to that of Anderson-Teixeira et al (2016).
-For mature forests, to obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot, weighting flux measurements according to the length of measurement periods (*i.e.*, end.date - start.date).
-Values were then averaged across plots clustered within 25 km of one another (geographic.area field of SITES table, sensu Anderson-Teixeira et al 2018), weighting by area.sampled (MEASUREMENTS table) or plot.area (PLOTS table) if available for all records.
-This step was taken to avoid pseudo-replication and to combine any records from sites with more than one name in ForC.
-Finally, we computed statistics with geographic.area as the unit of replication.
-To compare across biomes, [**Valentine, please describe**].
-There were enough data to run this analysis for all focal variables but XX.
-
-For young (<100yrs) forest types, we employed a mixed effects model (XX in R) with biome and log10[stand.age] as fixed effects and plot nested within geographic.area as a random effect.
-When the effect of stand.age was significant at p $\le$ 0.05 and when each biome had records for stands of at least 10 different ages, a biome - stand.age interaction was included in the model.
-In the C cycle schematics for young forests, we report equations based on these models.
-In cases where there was no significant effect of stand.age, records were averaged as for mature stands.
-
-All database manipulation, analyses, and figure production were fully automated in R (version, citation).
-Materials required to fully reproduce these analyses, including data, R scripts, and image files, are archived in Zenodo (DOI: TBD].
-Data, scripts, and results presented here are also available through the open-access ForC GitHub repository (https://github.com/forc-db/ForC), where many will be updated as the database develops.
-
-### Review Results/ Synthesis
-*(The Review Results/Synthesis section presents the findings of the analyses.)*
-
-*C cycling in mature forests*
-
-Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5, and statistics for each biome type are also summarized at: [GitHub URL].
-Of the # flux and # stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, # fluxes and # stocks in temperate broadleaf forests, #fluxes and # stocks in temperate conifer forests, and fluxes and # stocks in boreal forests.
-For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets were generally consistent.
-That is, component fluxes and stocks summed to within 1 std of more inclusive fluxes in all but one instance (in temperate conifer forests, $aboveground woody biomass + foliage biomass > aboveground biomass + 1std$; Fig.
-4).
-**update this: https://github.com/forc-db/ERL-review/issues/16 **
-
-\newpage
-\blandscape
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 2 | C cycle diagram for mature tropical broadleaf forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrows are scaled as the **square root of flux divided by 5**.", out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics("tables_figures/Tropical broadleaf MATURE.png")
-```
-
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 3 | C cycle diagram for mature temperate broadleaf forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics("tables_figures/Temperate broadleaf MATURE.png")
-```
-
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 3 | C cycle diagram for mature temperate conifer forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics("tables_figures/Temperate conifer MATURE.png")
-```
-
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 4 | C cycle diagram for mature boreal conifer forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics("tables_figures/Boreal conifer MATURE.png")
-```
-\elandscape
-
-(**check paragraph with latest data**) The largest C fluxes—including $GPP$, $NPP$, $ANPP$, $BNPP$, $R_{soil_het}$, $R_{soil}$, $R_{eco}$, — were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests  ([ForC_variable_averages_per_Biome](https://github.com/forc-db/ForC/blob/master/numbers_and_facts/ForC_variable_averages_per_Biome.csv)) (Fig.
-6).
-The same held true for some of the subsidiary fluxes: ANPP_foliage, ANPP_woody, [**OTHERS?**].
-Other subsidiary fluxes—including NPP_woody, ANPP_repro, ANPP_stem, ANPP_branch, woody.mortality, BNPP_coarse, BNPP_fine,  [**OTHERS?**]—deviated from this pattern and/or lacked data for some biomes.
-Net ecosystem exchange (NEE) did not follow this pattern, with no significant differences across biomes but the largest (negative) average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests.
-Thus, C cycling rates generally decreased from tropical to temperate to boreal forests, but with less apparent trends for some of the subsidiary fluxes and an important exception in the overall C balance (NEE).
-
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEE$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
-Lines...
-.", out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics("tables_figures/Figure6.png")
-```
-
-There were less distinct trends in C stocks across biomes (Fig.
-7).
-Biome means for live aboveground and total biomass followed the same general trend as the major fluxes, with biomass of tropical= temperate broadleaf $\ge$ temperate conifer> boreal forests (Fig.
-7a).
-However, the relative differences in these means were much smaller than for the major C fluxes, and there was significant variation within biomes.
-Maximum aboveground biomass values followed a very different trend than the means: temperate broadleaf> temperate conifer>boreal>tropical.
-There were some statistically significant biome differences in less frequently sampled C stocks (*e.g.*, woody biomass, foliage biomass, deadwood), but given high within-biome variability in C stocks and relatively low sample sizes, these were likely attributable to sampling biases and methodological differences than to true differences across biomes.
-
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.", out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics("tables_figures/Figure7.png")
-```
-
-*C cycling in young forests*
-
-(**check paragraph with latest data**) Average C cycles for forests <100 years old are presented in Figures 8-11.
-In general, ForC contained roughly comparable information on C fluxes in regrowth vs mature forests, with the notable exception of tropical forests, for which there were no fluxes that had been measured in $\ge$ 7 different geographic areas.
-C stocks were better documented for regrowth stands, with a total of # records from # distinct geographic areas.
-Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Fig.
-6-11; detailed below).
-Differences across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests (Figs.
-6-7).
-
-(**check paragraph with latest data**) ForC contained 14 flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig.
-6-7 and **S#- SI figures including plots for all variables**).
-Of these, 9 increased significantly with log10[stand.age]: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($-NEE$).
-The remaining five--$ANPP_{woody-branch}$, $BNPP_{root-coarse}$, $R_{soil-het}$, and $R_{soil-het}$-–displayed no significant relationship to stand age, although all displayed a positive trend.
-In terms of C stocks, 10 variables had sufficient data to test for age trends.
-Six of these—total biomass, aboveground biomass, aboveground woody biomass, foliage biomass, root biomass, and coarse root biomass—increased significantly with $log10[stand.age]$.
-The remaining four displayed non-significant positive trends:  fine root biomass, total dead wood, standing dead wood, and organic layer.
-
-\newpage
-\blandscape
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | C cycle diagram for young tropical broadleaf forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
-When age trends are signficant, they are presented with numbered equations; otherwise, means are presented.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics("tables_figures/Tropical broadleaf YOUNG.png")
-```
-
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 9 | C cycle diagram for young temperate broadleaf forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
-When age trends are signficant, they are presented with numbered equations; otherwise, means are presented.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics("tables_figures/Temperate broadleaf YOUNG.png")
-```
-
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 10 | C cycle diagram for young temperate conifer forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
-When age trends are signficant, they are presented with numbered equations; otherwise, means are presented.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics("tables_figures/Temperate conifer YOUNG.png")
-```
-
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 11 | C cycle diagram for young boreal conifer forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
-When age trends are signficant, they are presented with numbered equations; otherwise, means are presented.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics("tables_figures/Boreal conifer YOUNG.png")
-```
-\elandscape
-
-### Discussion 
-(*The Discussion section highlights the general conclusions, emphasizes novel findings, puts them in context and identifies issues and inferences for future work or policy considerations.*)
-
-Carbon cycling rates generally increased from boreal to tropical regions and with stand age.
-Specifically, the major C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests—a pattern that generally held for regrowth as well as mature forests (Figs.
-6-7).
-In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs.
-2-5, 7).
-The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Figs.
-6-7).
-Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks of mature forests are influenced by a different set of drivers.
-
-*C cycling across biomes*
-
-*Age trends in C cycling*
-
-A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling.
-
-### Acknowledgements
-All researchers whose data is included in ForC and this analysis.
-A Smithsonian Scholarly Studies grant to KAT and HML.
-
-### References
-(*All references should be presented in alphabetical order and should also be bundled in ways that will help the reader to find those that are germane to each of the themes, topics or issues that the authors have used to organize the review.
-Such a matrix will be included
-as supplementary material. ERL has no standardized format for subdividing references by cross-cutting themes, but guidance from editors will be helpful to authors.*)

### 2020-09-05 manuscript/ERL_global_C_review.Rmd
manuscript updates

@@
 
-Norbert Kunert
-
 Jennifer McGarvey
@@
 ### Summary
-*(The Summary should be no longer than 300 words and divided into the following sections: Background, Methods/Design, Review results/ Synthesis and Discussion.
+```{r eval=FALSE, echo=FALSE}
+# The Summary should be no longer than 300 words and divided into the following sections: Background, Methods/Design, Review results/ Synthesis and Discussion.
 Up to 7 Keywords should also be provided.
-The total length of the article is flexible.)*
+The total length of the article is flexible.
+```
 
@@
 ### Background
-*(The Background section needs to present the rationale for why a systematic review of this topic is needed along with a history of what has been done to date and an expectation of what new will emerge from the review, especially if quantitative meta-analyses of studies are being considered.)*
+```{r eval=FALSE, echo=FALSE}
+# The Background section needs to present the rationale for why a systematic review of this topic is needed along with a history of what has been done to date and an expectation of what new will emerge from the review, especially if quantitative meta-analyses of studies are being considered.
+```
 
@@
 Moreover, forests contain substantial C stocks: an estimated 92% of terrestrial biomass  (Pan et al 2013) and 45% of terrestrial C (biomass and soils; Bonan 2008).
-Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), resulting in net C sink of ~1.2-1.7 Gt C yr^-1^ across Earth's forests (Le Quéré et al 2017, Schimel et al 2015) (**update, give range**).
+Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), resulting in net C sink of ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**).
 The future of this sink is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change.
-Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change (Friedlingstein et al 2006, Cavaleri et al 2015, Griscom et al 2017; Grassi et al 2017; IPCC1.5).
+Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change [@grassi_key_2017] (Friedlingstein et al 2006, Cavaleri et al 2015, Griscom et al 2017; IPCC1.5).
 
@@
 While remote sensing measurements are increasingly useful for global- or regional-scale estimates of forest $GPP$ (Bagdley et al.
-2019, [Li & Xiao 2019](https://www.mdpi.com/2072-4292/11/21/2563)), aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) ([Clark et al.
-2004](https://www.onlinelibrary.wiley.com/doi/abs/10.1046/j.1461-0248.2003.00547.x); [Leithold et al 2018](https://nph-onlinelibrary-wiley-com.smithsonian.idm.oclc.org/doi/10.1111/nph.15110)), and to some extent net ecosystem exchange ($NEE$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements.
+2019, [@li_mapping_2019]), aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) [@clark_quantifying_2004; @leitold_nino_2018], and to some extent net ecosystem exchange ($NEE$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements.
 Moreover, data on stand characteristics--including species composition, age, and disturbance history--are critical to interpreting and projecting into the future, but [usually rely on ground-based measurements].
-Thus, ground-based measurements are--and will continue to be--central to characterizing forest C cycling.
-Tens of thousands of forest C measurements have been published, but their distribution across literally thousands of scientific articles-- along with variation in data formats, units, measurement methods, *etc.*--have made them effectively inaccesible for many global-scale analyses, including those attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011), and model evaluation (Clark et al 2017, Luo et al 2012).
+Thus, ground-based measurements are, and will continue to be, central to characterizing forest C cycling.
+
+Tens of thousands of ground-based forest C measurements have been published, but their distribution across literally thousands of scientific articles-- along with variation in data formats, units, measurement methods, *etc.*--have made them effectively inaccessible for many global-scale analyses, including those attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011), and model evaluation (Clark et al 2017, Luo et al 2012).
 Important progress has been made in synthesizing data to address how C cycling varies across forest biomes (Luyssaert et al.
-2007; REFS) and stand ages (REFS), yet generally consider only a limited set of C variables (but see Anderson-Teixeira et al 2016 for the tropics) and do not consider age-by-biome interactions.
-To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018).
+2007; REFS) and stand ages (REFS), yet generally consider only a limited set of C variables [but see @anderson-teixeira_carbon_2016 for the tropics] and do not consider age-by-biome interactions.
+To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (@anderson-teixeira_carbon_2016, @anderson-teixeira_forc:_2018).
 ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (*e.g.*, stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications.
@@
 
-1. How do C cycling vary across the worlds major forest biomes (*i.e.*, tropical, temperate, boreal)?
+1. How do C cycling vary across the worlds major forest biomes (*i.e.*, tropical, temperate broadleaf and deciduous, boreal)?
 
@@
 
-While components of these questions have been previously addressed (*e.g.*, REFS(Luyssaert et al 2007, Anderson et al 2006, Anderson-Teixeira et al 2013), our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.
+While components of these questions have been previously addressed [@luyssaert_co2_2007; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc:_2018, @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.
 
 ### Methods/ Design
-*(The Methods/Design section needs to describe how articles in the reviews were identified and what criteria were used for justifying inclusion in the review.
-While traditional reviews relying on past experience and expert knowledge are acceptable, editors should encourage reviews that are set up as so-called ‘systematic reviews’ (see the Cochrane Review procedure which initiated the process in the medical and health science: http://community.cochrane.org/ about-us/evidence-based-health-care).)*
+```{r eval=FALSE, echo=FALSE}
+# The Methods/Design section needs to describe how articles in the reviews were identified and what criteria were used for justifying inclusion in the review.
+While traditional reviews relying on past experience and expert knowledge are acceptable, editors should encourage reviews that are set up as so-called ‘systematic reviews’ (see the Cochrane Review procedure which initiated the process in the medical and health science: http://community.cochrane.org/ about-us/evidence-based-health-care).
+```
 
-This review synthesizes data from the ForC database (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018).
+This review synthesizes data from the ForC database [https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc:_2018].
 ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies.
@@
 As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible (83% of records in v.2.0; Anderson-Teixeira et al 2018).
-Particular attention was given to developing the database for tropical forests (Anderson-Teixeira et al 2016), yet these represented only approximately one-third of records in ForC v.2.0 (Anderson-Teixeira et al 2018).
-Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database (Bond-Lamberty and Thomson 2010; *hopefully Cook-Patton*; see [GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv)) ….
+Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], yet these represented only approximately one-third of records in ForC v.2.0 [@anderson-teixeira_forc:_2018].
+Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment (GROA) database (Cook-Patton et al.
+in press), and [GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv).
 We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS). 
@@
 For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
-Climate zones were defined according to Köppen-Geiger zones, which were extracted based on site geographic coordinates from the ESRI Köppen-Geiger map (downloaded June 2014 from http://maps3.arcgisonline.com/ArcGIS/rest/services/A-16/Köppen- Geiger_Observed_and_Predicted_Climate_Shifts/MapServer; Rubel and Kottek 2010) and are recorded in the Koeppen field of the ForC SITES table.
+Climate zones were defined according to Köppen-Geiger zones, which were extracted based on site geographic coordinates from the ESRI Köppen-Geiger map (downloaded June 2014 from http://maps3.arcgisonline.com/ArcGIS/rest/services/A-16/Köppen- Geiger_Observed_and_Predicted_Climate_Shifts/MapServer; Rubel and Kottek 2010).
 Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd).
@@
 Records for which stand age was unknown were excluded from the analysis. 
-These groupings were defined in March 2018, at which point they covered 91% of the primary variable records for forests of known age, or 86% of total records, in ForC (v.2.0; Anderson-Teixeira et al 2018).
+These groupings were defined based on ForC v.2.0[@anderson-teixeira_forc:_2018], for which they covered 91% of the primary variable records for forests of known age, or 86% of total records. 
 The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with <400 records total for stands of any age.
@@
 In generating ForC-simplified, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006). 
-Duplicate or otherwise conflicting records (indicated in conflicts field of MEASUREMENTS table) were reconciled as follows.
-Replicate measurements (*i.e.*, replicates from within a single study) were averaged.
-Records that subsumed others— **i.e.**, the time period included that of $\ge$ 2 other records or dates were unknown and therefore conflicted with $\ge$ 2 other records—were removed.
-For each group of duplicate records—*i.e.*, measurements of the same variable in the same plot at the same time—one record was assigned precedence (recorded in D.precedence field).
-When measurement periods overlapped or were not specified, precedence was given first to records representing longer measurement periods (*i.e.*, end.date - start.date) and then to more recently published values.
-We manually reviewed duplicates that differed only in methodology, assigning precedence to the record employing a more comprehensive approach (*e.g.*, inclusion of understory, lianas, or bamboo as opposed to just trees) or using a favored methodology.
-
+Duplicate or otherwise conflicting records were reconciled as described in *APPENDIX A*. 
 Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand establishment (*i.e.*, that reflected by stand.age).
@@
 ### Review Results/ Synthesis
-*(The Review Results/Synthesis section presents the findings of the analyses.)*
+```{r eval=FALSE, echo=FALSE}
+# The Review Results/Synthesis section presents the findings of the analyses.
+```
 
@@
 ### Discussion 
-(*The Discussion section highlights the general conclusions, emphasizes novel findings, puts them in context and identifies issues and inferences for future work or policy considerations.*)
+```{r eval=FALSE, echo=FALSE}
+# The Discussion section highlights the general conclusions, emphasizes novel findings, puts them in context and identifies issues and inferences for future work or policy considerations.
+```
 
@@
 All researchers whose data is included in ForC and this analysis.
+Thanks to Norbert Kunert for helpful discussion.
 A Smithsonian Scholarly Studies grant to KAT and HML.
 
-### References
-(*All references should be presented in alphabetical order and should also be bundled in ways that will help the reader to find those that are germane to each of the themes, topics or issues that the authors have used to organize the review.
-Such a matrix will be included
-as supplementary material. ERL has no standardized format for subdividing references by cross-cutting themes, but guidance from editors will be helpful to authors.*)
+```{r eval=FALSE, echo=FALSE}
+# All references should be presented in alphabetical order and should also be bundled in ways that will help the reader to find those that are germane to each of the themes, topics or issues that the authors have used to organize the review.
+Such a matrix will be included as supplementary material.
+ERL has no standardized format for subdividing references by cross-cutting themes, but guidance from editors will be helpful to authors.
+```

### 2020-09-16 manuscript/ERL_global_C_review.Rmd
update with a bit a my part

@@
 This step was taken to avoid pseudo-replication and to combine any records from sites with more than one name in ForC.
-Finally, we computed statistics with geographic.area as the unit of replication.
-To compare across biomes, [**Valentine, please describe**].
+Finally, for figures 6 and 7, the orginal values were analyzed via a linear mixed effects model ('lmer' function in 'lme4' R package) with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
+When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another.
 There were enough data to run this analysis for all focal variables but XX.
+[** I think we can do all of them??*XXX**]
 
-For young (<100yrs) forest types, we employed a mixed effects model (XX in R) with biome and log10[stand.age] as fixed effects and plot nested within geographic.area as a random effect.
+For young (<100yrs) forest types, we employed a mixed effects model with biome and log10[stand.age] as fixed effects and plot nested within geographic.area as a random effect on the intercept.
 When the effect of stand.age was significant at p $\le$ 0.05 and when each biome had records for stands of at least 10 different ages, a biome - stand.age interaction was included in the model.

### 2020-09-17 manuscript/ERL_global_C_review.Rmd
Update ERL_global_C_review.Rmd

@@
 **Authors:**
+
 Kristina J. Anderson-Teixeira^1,2^*
@@
 
-*Review Results/ Synthesis.*  The rate of C cycling generally increased from boreal to tropical regions, whereas C stocks showed less directional variation.
+*Review Results/ Synthesis.*  ForC v.XX yielded a fairly comprehensive picture of C cycling in the world's major forest biomes, with broad closure of C budgets.
+The rate of C cycling generally increased from boreal to tropical regions, whereas C stocks showed less directional variation.
 The majority of flux variables, together with most live biomass pools, increased significantly with stand age, and the rate of increase again tended to increase from boreal to tropical regions.
 
-*Discussion.* *[Discussion section will interpret results, highlighting new and significant findings, and discuss implications.
+*Discussion.* This analysis yields a comprehensive, largely consistent picture of C cycling across the world's forests.
+*[Discussion section will interpret results, highlighting new and significant findings, and discuss implications.
 Tentative headings are “Stand level C cycling in forests globally”, “Age trends in C cycling”, and “Implications for climate change mitigation”.]*
@@
 
-Here, we synthesize ForC data to provide a macroscopic overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age. We address two broad questions:
+Here, we synthesize ForC data to provide a macroscopic overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age. We address three broad questions:
 
-1. How do C cycling vary across the worlds major forest biomes (*i.e.*, tropical, temperate broadleaf and deciduous, boreal)?
+1.
+To what extent can we fully represent, and "close", C budgets for each of the world's major forest biomes (*i.e.*, tropical, temperate broadleaf and deciduous, boreal) based on the current ForC data?
 
-2. How does C cycling vary with stand age?
+2. How do C cycling vary across the worlds major forest biomes?
+
+3. How does C cycling vary with stand age?
 
@@
 
+ForC v.XX yielded a fairly comprehensive picture of C cycling in the world's major forest biomes. 
 Carbon cycling rates generally increased from boreal to tropical regions and with stand age. 
-Specifically, the major C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests—a pattern that generally held for regrowth as well as mature forests (Figs.
+Specifically, the major C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for regrowth as well as mature forests (Figs.
 6-7).
-In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs.
-2-5, 7).
-The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Figs.
-6-7).
+In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs. 2-5, 7). 
+The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Figs. 6-7). 
 Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks of mature forests are influenced by a different set of drivers.
 
+*C variable coverage and budget closure*
+
+Coverage:
+
+* We have fairly good coverage of most major variables
+* (note how this improves upon previous data compilations/ for which variables does ForC make the greatest difference (e.g., not AGB or NEE/GPP/Reco, but by far the latest data compilation for dead wood, [variables])
+* Noteable holes include: fluxes: R_auto_ag, woody mortality, folivary/ herbivory and respiration of herbivores (and therefore total R_het), ANPP_repro; also fluxes in tropical regrowth forests
+* for the C stocks considered here, the most poorly covered is dead wood (none in E hemisphere!), despite a focused effort on this variable that has resulted in ForC being by far the largest collection of these data.
+* all variables poorly covered in Africa
+
+Thus, overall, we're lacking coverage of fluxes to herbivores and higher consumers, along with the woody mortality and dead wood.
+
+Closure:
+
+* C cycle closure is fairly good. 
+* However, SD's are often large, reflective of significant within-biome variation. This makes the standard for closure relatively loose.
+* Lack of closure, in the few instances where it occurs, is probably more reflective of differences in the representation of forest types (e.g., disproportionate representation of US Pacific NW for aboveground woody biomass relative to AGB; Fig.
+3) than of methodological accuracy.
+
+Thus, overall, a high degree of closure implies that ForC gives a consistent picture of C cycling within biomes. 
+While these means are unlikely to be accurate representations of C cycling within any particular forest, they provide a useful baseline for comparison.
+
 *C cycling across biomes*
+
+@banbury_morgan_global_nodate
 
@@
 
+*Relevance for climate change modeling and mitigation*
+
+Importance of forest C in climate change modeling and mitigation...
+
+For C can help models by serving as benchmark data.
+* (cite Dietze paper that we reviewed)
+
+For some variables, such as AGB of mature forests, we sources other than ForC can provide the best available data.
+However, for others, ForC provides by far the most comprehensive source:
+* agb age trends (GROA; Cook-Patton et al. 2020)
+* dead wood (cite need for this)
+* fluxes
+
 ### Acknowledgements
 All researchers whose data is included in ForC and this analysis.
-Thanks to Norbert Kunert for helpful discussion.
+Thanks to Norbert Kunert and [Helene's intern] for helpful input at an earlier phase.
 A Smithsonian Scholarly Studies grant to KAT and HML.
+
+### References
 

### 2020-09-17 manuscript/ERL_global_C_review.Rmd
small updates

@@
 
-*Background.* *The fate of Earth’s climate closely linked to forests, which strongly influence atmospheric carbon dioxide (CO~2~) and climate through their influential role in the global carbon (C) cycle.
+*Background.* *The fate of Earth’s climate is closely linked to forests, which strongly influence atmospheric carbon dioxide (CO~2~) and climate through their influential role in the global carbon (C) cycle.
 Synthetic understanding of global forest C cycles is needed to constrain model estimates of forest feedbacks to climate change and to more accurately quantify the influence of land use decisions on climate.*
@@
 On an ecosystem level, forest C cycling is central to the flow of energy and materials, and as the basis of food webs...
+(support of biodiversity)
+(human relience as source of food, fuel, fiber)
 
@@
 While remote sensing measurements are increasingly useful for global- or regional-scale estimates of forest $GPP$ (Bagdley et al.
-2019, [@li_mapping_2019]), aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) [@clark_quantifying_2004; @leitold_nino_2018], and to some extent net ecosystem exchange ($NEE$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements.
+2019, [@li_mapping_2019]), aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) [@clark_quantifying_2004; @leitold_nino_2018], and to some extent net ecosystem exchange ($NEP$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements.
 Moreover, data on stand characteristics--including species composition, age, and disturbance history--are critical to interpreting and projecting into the future, but [usually rely on ground-based measurements].
@@
 2007; REFS) and stand ages (REFS), yet generally consider only a limited set of C variables [but see @anderson-teixeira_carbon_2016 for the tropics] and do not consider age-by-biome interactions.
-To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (@anderson-teixeira_carbon_2016, @anderson-teixeira_forc:_2018).
+To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (@anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018).
 ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (*e.g.*, stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications.
-ForC currently contains **#** records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.
+ForC currently contains **#** (~49,000!) records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.
 
@@
 
-While components of these questions have been previously addressed [@luyssaert_co2_2007; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc:_2018, @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.
+While components of these questions have been previously addressed [@luyssaert_co2_2007; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018, @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.
 
@@
 
-This review synthesizes data from the ForC database [https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc:_2018].
+This review synthesizes data from the ForC database [https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018].
 ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies.
@@
 As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible (83% of records in v.2.0; Anderson-Teixeira et al 2018).
-Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], yet these represented only approximately one-third of records in ForC v.2.0 [@anderson-teixeira_forc:_2018].
-Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment (GROA) database (Cook-Patton et al.
-in press), and [GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv).
+Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], yet these represented only approximately one-third of records in ForC v.2.0 [@anderson-teixeira_forc_2018].
+Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database (GROA v.##; Cook-Patton et al.
+in press), and data from several publications (Taylor et al.
+2017, Lutz et al.
+2018, Johnson_2018_csss; see GitHub repo for complete list) ([GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv)).
 We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS). 
@@
 Records for which stand age was unknown were excluded from the analysis. 
-These groupings were defined based on ForC v.2.0[@anderson-teixeira_forc:_2018], for which they covered 91% of the primary variable records for forests of known age, or 86% of total records. 
+These groupings were defined based on ForC v.2.0 [@anderson-teixeira_forc_2018], for which they covered 91% of the primary variable records for forests of known age, or 86% of total records. 
 The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with <400 records total for stands of any age.
@@
 This step was taken to avoid pseudo-replication and to combine any records from sites with more than one name in ForC. 
-Finally, for figures 6 and 7, the orginal values were analyzed via a linear mixed effects model ('lmer' function in 'lme4' R package) with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
+Finally, for figures 6 and 7, the original values were analyzed via a linear mixed effects model ('lmer' function in 'lme4' R package) with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
 When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another.
-There were enough data to run this analysis for all focal variables but XX.
-[** I think we can do all of them??*XXX**]
 
@@
 (**check paragraph with latest data**) The largest C fluxes—including $GPP$, $NPP$, $ANPP$, $BNPP$, $R_{soil_het}$, $R_{soil}$, $R_{eco}$, — were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests  ([ForC_variable_averages_per_Biome](https://github.com/forc-db/ForC/blob/master/numbers_and_facts/ForC_variable_averages_per_Biome.csv)) (Fig.
-6).
+5).
 The same held true for some of the subsidiary fluxes: ANPP_foliage, ANPP_woody, [**OTHERS?**].
 Other subsidiary fluxes—including NPP_woody, ANPP_repro, ANPP_stem, ANPP_branch, woody.mortality, BNPP_coarse, BNPP_fine,  [**OTHERS?**]—deviated from this pattern and/or lacked data for some biomes.
-Net ecosystem exchange (NEE) did not follow this pattern, with no significant differences across biomes but the largest (negative) average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests.
-Thus, C cycling rates generally decreased from tropical to temperate to boreal forests, but with less apparent trends for some of the subsidiary fluxes and an important exception in the overall C balance (NEE).
+Net ecosystem production ($NEP$) did not follow this pattern, with no significant differences across biomes but the largest (negative) average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests.
+Thus, C cycling rates generally decreased from tropical to temperate to boreal forests, but with less apparent trends for some of the subsidiary fluxes and an important exception in the overall C balance ($NEP$).
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEE$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 5 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEP$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 There were less distinct trends in C stocks across biomes (Fig.
-7).
+6).
 Biome means for live aboveground and total biomass followed the same general trend as the major fluxes, with biomass of tropical= temperate broadleaf $\ge$ temperate conifer> boreal forests (Fig.
-7a).
+6a).
 However, the relative differences in these means were much smaller than for the major C fluxes, and there was significant variation within biomes.
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.", out.width='100%', fig.pos='H'}
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.", out.width='100%', fig.pos='H'}
 
@@
 
-(**check paragraph with latest data**) Average C cycles for forests <100 years old are presented in Figures 8-11.
+(**check paragraph with latest data**) Average C cycles for forests <100 years old are presented in Figures 7-10.
 In general, ForC contained roughly comparable information on C fluxes in regrowth vs mature forests, with the notable exception of tropical forests, for which there were no fluxes that had been measured in $\ge$ 7 different geographic areas.
@@
 Differences across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests (Figs.
-6-7).
+5-6).
 
 (**check paragraph with latest data**) ForC contained 14 flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig.
-6-7 and **S#- SI figures including plots for all variables**).
-Of these, 9 increased significantly with log10[stand.age]: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($-NEE$).
+5-6 and **S#- SI figures including plots for all variables**).
+Of these, 9 increased significantly with log10[stand.age]: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($NEP$).
 The remaining five--$ANPP_{woody-branch}$, $BNPP_{root-coarse}$, $R_{soil-het}$, and $R_{soil-het}$-–displayed no significant relationship to stand age, although all displayed a positive trend.
@@
 \blandscape
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | C cycle diagram for young tropical broadleaf forests.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | C cycle diagram for young tropical broadleaf forests.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 9 | C cycle diagram for young temperate broadleaf forests.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | C cycle diagram for young temperate broadleaf forests.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 10 | C cycle diagram for young temperate conifer forests.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 9 | C cycle diagram for young temperate conifer forests.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 11 | C cycle diagram for young boreal conifer forests.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 10 | C cycle diagram for young boreal conifer forests.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
@@
 Specifically, the major C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for regrowth as well as mature forests (Figs.
-6-7).
-In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs. 2-5, 7). 
-The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Figs. 6-7). 
+5-6).
+In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs. 1-4, 6). 
+The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Figs. 5-10). 
 Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks of mature forests are influenced by a different set of drivers.
@@
 * We have fairly good coverage of most major variables
-* (note how this improves upon previous data compilations/ for which variables does ForC make the greatest difference (e.g., not AGB or NEE/GPP/Reco, but by far the latest data compilation for dead wood, [variables])
+* (note how this improves upon previous data compilations/ for which variables does ForC make the greatest difference (e.g., not AGB or NEP/GPP/Reco, but by far the latest data compilation for dead wood, [variables])
 * Noteable holes include: fluxes: R_auto_ag, woody mortality, folivary/ herbivory and respiration of herbivores (and therefore total R_het), ANPP_repro; also fluxes in tropical regrowth forests
@@
 For C can help models by serving as benchmark data.
+
 * (cite Dietze paper that we reviewed)
@@
 However, for others, ForC provides by far the most comprehensive source:
+
 * agb age trends (GROA; Cook-Patton et al. 2020)
+
 * dead wood (cite need for this)
+
 * fluxes

### 2020-09-17 manuscript/ERL_global_C_review.Rmd
minor updates

@@
 (support of biodiversity)
-(human relience as source of food, fuel, fiber)
+(human reliance as source of food, fuel, fiber)
 
@@
 Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], yet these represented only approximately one-third of records in ForC v.2.0 [@anderson-teixeira_forc_2018].
-Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database (GROA v.##; Cook-Patton et al.
-in press), and data from several publications (Taylor et al.
+Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database (GROA v1.0; Cook-Patton et al.
+in press; ZenodoDOI), and data from several publications (Taylor et al.
 2017, Lutz et al.
@@
 We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS). 
-The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: XX).
+The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: TBD).
 
@@
 
-Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5, and statistics for each biome type are also summarized at: [GitHub URL].
-Of the # flux and # stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, # fluxes and # stocks in temperate broadleaf forests, #fluxes and # stocks in temperate conifer forests, and fluxes and # stocks in boreal forests.
+Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 1-4, and statistics for each biome type are also summarized at: [GitHub URL].
+Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, # fluxes and # stocks in temperate broadleaf forests, #fluxes and # stocks in temperate conifer forests, and fluxes and # stocks in boreal forests.
 For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets were generally consistent. 
@@
 Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Fig.
-6-11; detailed below).
+5-10; detailed below).
 Differences across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests (Figs.

### 2020-09-18 manuscript/ERL_global_C_review.Rmd
small updates

@@
 
-ForC v.XX yielded a fairly comprehensive picture of C cycling in the world's major forest biomes. 
+ForC v.XX yielded a fairly comprehensive and internally consistent picture of C cycling in the world's major forest biomes. 
 Carbon cycling rates generally increased from boreal to tropical regions and with stand age. 
@@
 
+NEP increases with log(age) to 100 --> strongest C sinks are established secondary forests
+
 ### Acknowledgements

### 2020-09-18 manuscript/ERL_global_C_review.Rmd
remove context on importance of forests beyond global C cycle/climate

@@
 
-*Background.* *The fate of Earth’s climate is closely linked to forests, which strongly influence atmospheric carbon dioxide (CO~2~) and climate through their influential role in the global carbon (C) cycle.
-Synthetic understanding of global forest C cycles is needed to constrain model estimates of forest feedbacks to climate change and to more accurately quantify the influence of land use decisions on climate.*
+*Background.* The fate of Earth’s climate is closely linked to forests, which strongly influence atmospheric carbon dioxide (CO~2~) and climate through their influential role in the global carbon (C) cycle.
+Synthetic understanding of global forest C cycles is needed to constrain model estimates of forest feedbacks to climate change and to more accurately quantify the influence of land use decisions on climate.
 
@@
 Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change [@grassi_key_2017] (Friedlingstein et al 2006, Cavaleri et al 2015, Griscom et al 2017; IPCC1.5).
-
-On an ecosystem level, forest C cycling is central to the flow of energy and materials, and as the basis of food webs...
-(support of biodiversity)
-(human reliance as source of food, fuel, fiber)
 

### 2020-09-18 manuscript/ERL_global_C_review.Rmd
update, we actually did not have enough data for some variables
Update ERL_global_C_review.Rmd

@@
 
-Here, we synthesize ForC data to provide a macroscopic overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age. We address three broad questions:
+Here, we synthesize ForC data (Fig.
+1) to provide a macroscopic overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age.
+We address three broad questions:
 
@@
 
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 1 | Map of sites included in this analysis.
+Symbols are colored according to the number of records at each site.
+Underlying map shows coverage of evergreen, deciduous, and mixed forests (from SYNMAP; Jung et al.
+2006) and biomes.
+Distribution of sites, plots, and records among biomes is shown in the inset.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/World_Map_records_in_Biomes.png")
+```
+
 ### Methods/ Design
@@
 
-This review synthesizes data from the ForC database [https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018].
+This review synthesizes data from the ForC database [Fig.
+1; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018].
 ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies.
@@
 For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables (Table 1, [this table](https://github.com/forc-db/ForC/blob/master/figures/C_cycle_diagrams/ForC_variables_mapping_for_C_cycle_diagrams.csv))).
-Although ForC contains information that may be used to standardize or control for methodological differences (*e.g.*, area sampled, min stem diameter sampled, allometric equations applied; Anderson-Teixeira et al 2018), for this analysis we included all relevant data in ForC.
+Although ForC contains information that may be used to standardize or control for methodological differences [*e.g.*, area sampled, min stem diameter sampled, allometric equations applied; @anderson-teixeira_forc_2018], for this analysis we included all relevant data in ForC.
 Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum stem diameter sampled was $\le$ 10cm.
@@
 When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another.
+There were enough data to run this analysis for all focal variables but **biomass_ag_understory, BNPP_root.turnover_fine, deadwood_down, total.ecosystem_2.**
 
@@
 
-Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 1-4, and statistics for each biome type are also summarized at: [GitHub URL].
-Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, # fluxes and # stocks in temperate broadleaf forests, #fluxes and # stocks in temperate conifer forests, and fluxes and # stocks in boreal forests.
+Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5, and statistics for each biome type are also summarized at: [GitHub URL].
+Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, *#* fluxes and *#* stocks in temperate broadleaf forests, *#* fluxes and *#* stocks in temperate conifer forests, and fluxes and *#* stocks in boreal forests.
 For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets were generally consistent. 
 That is, component fluxes and stocks summed to within 1 std of more inclusive fluxes in all but one instance (in temperate conifer forests, $aboveground woody biomass + foliage biomass > aboveground biomass + 1std$; Fig.
-4).
+5).
 **update this: https://github.com/forc-db/ERL-review/issues/16 **
 
-\newpage
 \blandscape
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 3 | C cycle diagram for mature temperate conifer forests.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 4 | C cycle diagram for mature temperate conifer forests.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 4 | C cycle diagram for mature boreal conifer forests.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 5 | C cycle diagram for mature boreal conifer forests.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
@@
 (**check paragraph with latest data**) The largest C fluxes—including $GPP$, $NPP$, $ANPP$, $BNPP$, $R_{soil_het}$, $R_{soil}$, $R_{eco}$, — were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests  ([ForC_variable_averages_per_Biome](https://github.com/forc-db/ForC/blob/master/numbers_and_facts/ForC_variable_averages_per_Biome.csv)) (Fig.
-5).
+6).
 The same held true for some of the subsidiary fluxes: ANPP_foliage, ANPP_woody, [**OTHERS?**].
 Other subsidiary fluxes—including NPP_woody, ANPP_repro, ANPP_stem, ANPP_branch, woody.mortality, BNPP_coarse, BNPP_fine,  [**OTHERS?**]—deviated from this pattern and/or lacked data for some biomes.
-Net ecosystem production ($NEP$) did not follow this pattern, with no significant differences across biomes but the largest (negative) average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests.
+Net ecosystem production ($NEP$) did not follow this pattern, with no significant differences across biomes but the largest average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests.
 Thus, C cycling rates generally decreased from tropical to temperate to boreal forests, but with less apparent trends for some of the subsidiary fluxes and an important exception in the overall C balance ($NEP$).
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 5 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEP$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEP$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 
-  knitr::include_graphics("tables_figures/Figure6.png")
+  knitr::include_graphics("tables_figures/Flux_age_trends.png")
 ```
 
-There were less distinct trends in C stocks across biomes (Fig.
-6).
+There were less distinct trends in C stocks across biomes (Fig. 7). 
 Biome means for live aboveground and total biomass followed the same general trend as the major fluxes, with biomass of tropical= temperate broadleaf $\ge$ temperate conifer> boreal forests (Fig.
-6a).
+7a).
 However, the relative differences in these means were much smaller than for the major C fluxes, and there was significant variation within biomes. 
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.", out.width='100%', fig.pos='H'}
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.", out.width='100%', fig.pos='H'}
 
-  knitr::include_graphics("tables_figures/Figure7.png")
+  knitr::include_graphics("tables_figures/Stock_age_trends.png")
 ```
@@
 
-(**check paragraph with latest data**) Average C cycles for forests <100 years old are presented in Figures 7-10.
+(**check paragraph with latest data**) Average C cycles for forests <100 years old are presented in Figures 8-11.
 In general, ForC contained roughly comparable information on C fluxes in regrowth vs mature forests, with the notable exception of tropical forests, for which there were no fluxes that had been measured in $\ge$ 7 different geographic areas.
@@
 Differences across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests (Figs.
-5-6).
+6-7).
 
 (**check paragraph with latest data**) ForC contained 14 flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig.
-5-6 and **S#- SI figures including plots for all variables**).
+6-7 and **S#- SI figures including plots for all variables**).
 Of these, 9 increased significantly with log10[stand.age]: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($NEP$).
@@
 \blandscape
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | C cycle diagram for young tropical broadleaf forests.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | C cycle diagram for young tropical broadleaf forests.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | C cycle diagram for young temperate broadleaf forests.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 9 | C cycle diagram for young temperate broadleaf forests.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 9 | C cycle diagram for young temperate conifer forests.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 10 | C cycle diagram for young temperate conifer forests.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 10 | C cycle diagram for young boreal conifer forests.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 11 | C cycle diagram for young boreal conifer forests.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
@@
 Specifically, the major C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for regrowth as well as mature forests (Figs.
-5-6).
-In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs. 1-4, 6). 
-The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Figs. 5-10). 
+6-7).
+In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs. 2-5, 7). 
+The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Figs. 6-11). 
 Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks of mature forests are influenced by a different set of drivers.
@@
 * Lack of closure, in the few instances where it occurs, is probably more reflective of differences in the representation of forest types (e.g., disproportionate representation of US Pacific NW for aboveground woody biomass relative to AGB; Fig.
-3) than of methodological accuracy.
+4) than of methodological accuracy.
 
@@
 While these means are unlikely to be accurate representations of C cycling within any particular forest, they provide a useful baseline for comparison.
+
+* note that our equations are unlikely to be the best possible for any given flux
 

### 2020-09-18 manuscript/ERL_global_C_review.Rmd
worked on discussion

@@
 
-Data were analyzed to produce basic summaries of C cycle patterns across biomes and stand ages following an approach similar to that of Anderson-Teixeira et al (2016). 
+Data were analyzed to produce basic summaries of C cycle patterns across biomes and stand ages following an approach similar to that of @anderson-teixeira_carbon_2016. 
 For mature forests, to obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot, weighting flux measurements according to the length of measurement periods (*i.e.*, end.date - start.date).
@@
 ```
+*Data Coverage*
+
+Of the *#* records in ForC v.#, *#* met the criteria for inclusion in this study (Fig. 1). 
+*(Give some stats on coverage by age and biome)*
 
@@
 
-Coverage:
+ForC provides fairly good coverage of most major variables.
+*(discuss  how this improves upon previous data compilations/ for which variables does ForC make the greatest difference (e.g., not AGB or NEP/GPP/Reco, but by far the latest data compilation for dead wood, [variables])*
+*(Noteable holes include: fluxes: R_auto_ag, woody mortality, folivary/ herbivory and respiration of herbivores (and therefore total R_het), ANPP_repro; also fluxes in tropical regrowth forests)* 
+*For the C stocks considered here, the most poorly covered is dead wood (none in E hemisphere!), despite a focused effort on this variable that has resulted in ForC being by far the largest collection of these data.*
+Thus, overall, we're lacking coverage of fluxes to herbivores and higher consumers, along with the woody mortality and dead wood. 
+Geographically, all variables poorly covered in Africa and Siberia.
 
-* We have fairly good coverage of most major variables
-* (note how this improves upon previous data compilations/ for which variables does ForC make the greatest difference (e.g., not AGB or NEP/GPP/Reco, but by far the latest data compilation for dead wood, [variables])
-* Noteable holes include: fluxes: R_auto_ag, woody mortality, folivary/ herbivory and respiration of herbivores (and therefore total R_het), ANPP_repro; also fluxes in tropical regrowth forests
-* for the C stocks considered here, the most poorly covered is dead wood (none in E hemisphere!), despite a focused effort on this variable that has resulted in ForC being by far the largest collection of these data.
-* all variables poorly covered in Africa
-
-Thus, overall, we're lacking coverage of fluxes to herbivores and higher consumers, along with the woody mortality and dead wood.
-
-Closure:
-
-* C cycle closure is fairly good. 
-* However, SD's are often large, reflective of significant within-biome variation. This makes the standard for closure relatively loose.
-* Lack of closure, in the few instances where it occurs, is probably more reflective of differences in the representation of forest types (e.g., disproportionate representation of US Pacific NW for aboveground woody biomass relative to AGB; Fig.
+Closure of the C cycle budgets for mature forests (Figs. 2-5) is fairly good. 
+However, SD's are often large, reflective of significant within-biome variation. This makes the standard for closure relatively loose.
+Lack of closure, in the few instances where it occurs, is probably more reflective of differences in the representation of forest types (e.g., disproportionate representation of US Pacific NW for aboveground woody biomass relative to AGB; Fig.
 4) than of methodological accuracy.
-
 Thus, overall, a high degree of closure implies that ForC gives a consistent picture of C cycling within biomes. 
@@
 
-* note that our equations are unlikely to be the best possible for any given flux
-
 *C cycling across biomes*
 
-@banbury_morgan_global_nodate
+THere is a general acceleration of carbon cycling from the tropics to the high latitudes. 
+For mature forests, this is consistent with lots of previous work, most recently @banbury_morgan_global_nodate.
+For regrowth forests, we see faster C accumulation in the tropics (*I assume!- analysis not yet complete*), consistent with the observation that biomass accumulation decreases with latitude (Anderson 2006, Cook-Patton et al.
+2020).
+For most C flux variables, this analysis is the first to examine flux trends in regrowth forests across biomes  (i.e., age x biome interaction).
+Data remain sparse, but for better-represented variables, we often see faster acceleration of C cycling in the warmer climates.
+Further work will be required to explore age x climate interactions, but our broad-brush overview indicates that C cycling of regrowth forests is not only higher in the tropics, parallel to fluxes in mature forests [ @banbury_morgan_global_nodate], but also that it accelerates more rapidly with stand age in the tropics, consistent with more rapid biomass accumulation.
 
-*Age trends in C cycling*
+*Age trends in C cycling*  (Just some rough notes at this point)
 
@@
 
-*Relevance for climate change modeling and mitigation*
+Moreover, as disturbances increase, understanding the carbon dynamics of regrowth forests will be increasingly important.
 
-Importance of forest C in climate change modeling and mitigation...
+NEP increases with log(age) to 100 --> strongest C sinks are established secondary forests.
 
-For C can help models by serving as benchmark data.
+*Relevance for climate change prediction and mitigation*
 
-* (cite Dietze paper that we reviewed)
+The future of forest C cycling will shape trends in atmospheric CO~2~ and the course of climate change.
+For a human society seeking to understand and mitigate climate change, the data contained in ForC and summarized here can help to meet two major challenges.
 
-For some variables, such as AGB of mature forests, we sources other than ForC can provide the best available data.
-However, for others, ForC provides by far the most comprehensive source:
+First, improved representation of forest C cycling in models is essential to improving predictions of the future course of climate change. 
+To ensure that models are giving the right answers for the right reasons, it is important benchmark against multiple components of the C cycle. 
+By making tens of thousands of records readily available in standardized format, ForC makes it feasible for the modeling community to draw upon these data to benchmark models.
+Integration of ForC with models is a goal (Fer et al., in revision).
+On a more cursory level, the values summarized here can serve as a sanity check for modelers to determine whether model predictions for multiple C cycle variables are reasonable--i.e., within the range of previous observations.
 
-* agb age trends (GROA; Cook-Patton et al. 2020)
+Second, ForC can serve as a pipeline through which forest science can inform forest-based climate change mitigation efforts.
+Such efforts will be most effective when informed by the best available data, yet it is not feasible for the individuals and organizations designing such efforts to sort through literature, often behind paywalls, with data reported in varying units, terminology, etc.
+One goal for ForC is to serve as a pipeline through which information can flow efficiently from forest researchers to decision-makers working to implement forest conservation strategies at global, national, or landscape scales.
+This is already happening! 
+ForC has already contributed to updating the IPCC guidelines for carbon accounting in forests (IPCC 2019, Requena-Suarez et al 2019, Rozendaal et al in prep), mapping C accumulation potential from natural forest regrowth globally (Cook-Patton et al.
+2020), and informing ecosystem conservation priorities (Goldstein et al.
+2020).
 
-* dead wood (cite need for this)
+There remain numerous data needs for improved accounting of forest carbon stocks and fluxes in GHG accounting.
+AGB is the largest stock, and most of the emphasis is on this variable. 
+Remote sensing, with calibration based on high-quality ground-based data (Schepaschenko et al. 2019, Chave et al. 2019), is the best approach for mapping forest carbon (REFS).
+However, it is limited in that it is not associated with stand age and disturbance history, except in recent decades (e.g., Hansen et al.; REFS).
+ForC  is therefore valuable in defining age-based trajectories in biomass, as in Cook-Patton et al. 2020.
 
-* fluxes
+Other variables cannot be remotely sensed.
+In terms of C stocks, there is a paucity of data on dead wood and organic layer (Pan et al. ?). 
+These can be significant.
+*(give some stats/ cite figures)*.
+ForC does not include soil carbon, which is covered by other efforts (REFS).
+For fluxes, Fluxnet is the keeper of the best data on NEE, GPP, Reco (REFS), and SRDB remains the authority on soil respiration (REFS).
+ForC includes recent data from both, but is not continuously integrated.
+For C is the best source for most of the subsidiary fluxes: NPP, woody mortality...
 
-NEP increases with log(age) to 100 --> strongest C sinks are established secondary forests
+We recommend that use of ForC data go to the original database, as opposed to useing "off-the-shelf" values from this publication. 
+This is because
+(1) ForC is constantly being updated,
+(2) analyses should be designed to match the application,
+(3) age equations presented here all fit a single functional form that is not necessarily the best possible for all the variables.
+
+As climate change accelerates, understanding and managing the carbon dynamics of forests will be critical to forecasting, mitigation, and adaptation.
+The C data in ForC, as summarized here, will be valuable to these efforts.
 
@@
 A Smithsonian Scholarly Studies grant to KAT and HML. 
+WLS grant to KAT.
 

### 2020-09-19 manuscript/ERL_global_C_review.Rmd
worked on paper

@@
 
-### Summary
+## Summary
 ```{r eval=FALSE, echo=FALSE}
@@
 *Methods/Design.* Here, we draw from the Global Forest C Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation.
-Specifically, we draw upon ## records from ## geographic locations representing ## C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests.
+Specifically, we draw upon *##* records from *##* geographic locations representing *##* C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests.
 For regrowth forests, we quantify age trends for all variables.
@@
 
-### Background
+## Background
 ```{r eval=FALSE, echo=FALSE}
@@
 
-Carbon (C) cycling in forests is central to sustaining life on Earth.
-Forest ecosystems globally influence climate through their vital role in regulating atmospheric carbon dioxide (CO~2~).
-Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ (Bagdley et al.
-2019), or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**).
+Forest ecosystems globally play a critical role in regulating atmospheric carbon dioxide (CO~2~), and thereby will shape the course of climate change (IPCC1.5). 
+Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**).
 While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion was retained in ecosystems over recent decades.
-The resulting in C sink averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**).
-Moreover, forests contain substantial C stocks: an estimated 92% of terrestrial biomass  (Pan et al 2013) and 45% of terrestrial C (biomass and soils; Bonan 2008).
+The resulting CO~2~ sink averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**). 
+Moreover, forests contain substantial carbon (C) stocks: an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C [biomass and soils; @bonan_forests_2008]. 
 Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), resulting in net C sink of ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**).
 The future of this sink is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change. 
-Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change [@grassi_key_2017] (Friedlingstein et al 2006, Cavaleri et al 2015, Griscom et al 2017; IPCC1.5).
+Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change [@grassi_key_2017; Cavaleri et al 2015, Griscom et al 2017].
 
-Despite the centrality of forest C cycling in regulating both atmospheric CO~2~ and ecosystem function and diversity, we lack a comprehensive understanding of how C cycling varies with across biomes and with stand age.
-While remote sensing measurements are increasingly useful for global- or regional-scale estimates of forest $GPP$ (Bagdley et al.
-2019, [@li_mapping_2019]), aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) [@clark_quantifying_2004; @leitold_nino_2018], and to some extent net ecosystem exchange ($NEP$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements.
-Moreover, data on stand characteristics--including species composition, age, and disturbance history--are critical to interpreting and projecting into the future, but [usually rely on ground-based measurements].
-Thus, ground-based measurements are, and will continue to be, central to characterizing forest C cycling.
+Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [Friedlingstein et al 2006; @krause_large_2018; REFS] and CO~2~ accounting frameworks [Pan, REFS] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history.
+These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published. 
+While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [@li_mapping_2019; **REFS for biomass, biomass change, net CO2 flux**], measurement of most forest C stocks and fluxes require intensive on-the-ground data collection.
+Original studies typically cover only a small numbers of sites at a time, with rare exceptions spanning regions or continents [e.g., @lutz_global_2018; FLUXNET_REF), typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015] or FLUXNET [@baldocchi_fluxnet_2001].
+The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*.
+In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle (*e.g.*, Pan et al 2011), or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere (REFS).
 
-Tens of thousands of ground-based forest C measurements have been published, but their distribution across literally thousands of scientific articles-- along with variation in data formats, units, measurement methods, *etc.*--have made them effectively inaccessible for many global-scale analyses, including those attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011), and model evaluation (Clark et al 2017, Luo et al 2012).
-Important progress has been made in synthesizing data to address how C cycling varies across forest biomes (Luyssaert et al.
-2007; REFS) and stand ages (REFS), yet generally consider only a limited set of C variables [but see @anderson-teixeira_carbon_2016 for the tropics] and do not consider age-by-biome interactions.
-To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (@anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018).
-ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (*e.g.*, stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications.
+To address the need for global-scale analyses of forest C cycling, we recently developed an open-access Global Forest Carbon database, ForC (@anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018).
+ForC contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables) based on ground-based measurements, along with associated data required for interpretation (*e.g.*, stand history, measurement methods).
+These data have been amalgamated from original peer-reviewed publications, either directly or via intermediary data compilations, with notable recent additions of the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020].
 ForC currently contains **#** (~49,000!) records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.
@@
 
-2. How do C cycling vary across the worlds major forest biomes?
+2. How do C cycling vary across the world's major forest biomes?
 
-3. How does C cycling vary with stand age?
+3. How does C cycling vary with stand age (in interaction with biome)?
 
-While components of these questions have been previously addressed [@luyssaert_co2_2007; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018, @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.
+While components of these questions have been previously addressed [@luyssaert_co2_2007; @anderson-teixeira_carbon_2016; @cook-patton_mapping_2020; @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.
 
@@
 
-### Methods/ Design
+## Methods/ Design
 ```{r eval=FALSE, echo=FALSE}
@@
 
-This review synthesizes data from the ForC database [Fig.
-1; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018].
+This review synthesizes data from the ForC database [Fig. 1; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018]. 
 ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies. 
@@
 The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age. 
-As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible (83% of records in v.2.0; Anderson-Teixeira et al 2018).
+As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible [83% of records in v.2.0; @anderson-teixeira_forc_2018].
 Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], yet these represented only approximately one-third of records in ForC v.2.0 [@anderson-teixeira_forc_2018].
@@
 
-### Review Results/ Synthesis
+## Review Results/ Synthesis
 ```{r eval=FALSE, echo=FALSE}
@@
 ```
-*Data Coverage*
+### Data Coverage
 
@@
 
-*C cycling in mature forests*
+### C cycling in mature forests
 
@@
 
-*C cycling in young forests*
+### C cycling in young forests
 
@@
 
-### Discussion 
+## Discussion 
 ```{r eval=FALSE, echo=FALSE}
@@
 
-*C variable coverage and budget closure*
+### C variable coverage and budget closure
 
@@
 
-*C cycling across biomes*
+### C cycling across biomes
 
-THere is a general acceleration of carbon cycling from the tropics to the high latitudes. 
-For mature forests, this is consistent with lots of previous work, most recently @banbury_morgan_global_nodate.
-For regrowth forests, we see faster C accumulation in the tropics (*I assume!- analysis not yet complete*), consistent with the observation that biomass accumulation decreases with latitude (Anderson 2006, Cook-Patton et al.
-2020).
+Our analysis reveals a general acceleration of carbon cycling from the tropics to the high latitudes. 
+For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude [e.g.,  @banbury_morgan_global_nodate].
+For regrowth forests, more rapid accumulation of biomass at lower latitudes has been well-established [@anderson_temperature-dependence_2006; @cook-patton_mapping_2020], whereas this is the first study to compare age trends in deadwood and organic layer across biomes [but see @cook-patton_mapping_2020].
 For most C flux variables, this analysis is the first to examine flux trends in regrowth forests across biomes  (i.e., age x biome interaction).
@@
 
-*Age trends in C cycling*  (Just some rough notes at this point)
+In contrast to C fluxes and accumulation rates in regrowth forests, stocks...
+
+### Age trends in C cycling
+
+*(Just some rough notes at this point)*
 
@@
 
-Moreover, as disturbances increase, understanding the carbon dynamics of regrowth forests will be increasingly important.
+Moreover, as disturbances increase [@andela_human-driven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests will be increasingly important.
+
+It's also important to understand secondary forest C sequestration to reduce uncertainty regarding the potential for carbon uptake by regrowth forests [@kruase_large_2018; @cook-patton_mapping_2020].
 
@@
 
-*Relevance for climate change prediction and mitigation*
+### Relevance for climate change prediction and mitigation
 
@@
 This is already happening! 
-ForC has already contributed to updating the IPCC guidelines for carbon accounting in forests (IPCC 2019, Requena-Suarez et al 2019, Rozendaal et al in prep), mapping C accumulation potential from natural forest regrowth globally (Cook-Patton et al.
-2020), and informing ecosystem conservation priorities (Goldstein et al.
-2020).
+ForC has already contributed to updating the IPCC guidelines for carbon accounting in forests [IPCC 2019; @requena_suarez_estimating_2019; Rozendaal et al in prep], mapping C accumulation potential from natural forest regrowth globally [@cook-patton_mapping_2020], and informing ecosystem conservation priorities [@goldstein_protecting_2020].
 
@@
 AGB is the largest stock, and most of the emphasis is on this variable. 
-Remote sensing, with calibration based on high-quality ground-based data (Schepaschenko et al. 2019, Chave et al. 2019), is the best approach for mapping forest carbon (REFS).
-However, it is limited in that it is not associated with stand age and disturbance history, except in recent decades (e.g., Hansen et al.; REFS).
-ForC  is therefore valuable in defining age-based trajectories in biomass, as in Cook-Patton et al. 2020.
+Remote sensing, with calibration based on high-quality ground-based data [@schepaschenko_forest_2019; @chave_ground_2019], is the best approach for mapping forest carbon (REFS).
+However, it is limited in that it is not associated with stand age and disturbance history, except in recent decades when satellite data can be used to detect forest loss, gain, and some of their dominant drivers [@hansen_high-resolution_2013; @song_global_2018; @curtis_classifying_2018].
+ForC  is therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020.
+
+ *remote sensing measurements are increasingly useful for global- or regional-scale estimates of forest $GPP$ (Bagdley et al.
+2019, [@li_mapping_2019]), aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) [@clark_quantifying_2004; @leitold_nino_2018], and to some extent net ecosystem exchange ($NEP$) (REFS),*
 
@@
 
-We recommend that use of ForC data go to the original database, as opposed to useing "off-the-shelf" values from this publication. 
+We recommend that use of ForC data go to the original database, as opposed to using "off-the-shelf" values from this publication. 
 This is because
@@
 
-### Acknowledgements
+## Acknowledgements
 All researchers whose data is included in ForC and this analysis.
@@
 
-### References
+## References
 

### 2020-09-19 manuscript/ERL_global_C_review.Rmd
worked on writing, refs

@@
 The rate of C cycling generally increased from boreal to tropical regions, whereas C stocks showed less directional variation.
-The majority of flux variables, together with most live biomass pools, increased significantly with stand age, and the rate of increase again tended to increase from boreal to tropical regions.
+The majority of flux variables, together with most live biomass pools, increased significantly with stand age, *and the rate of increase again tended to increase from boreal to tropical regions*.
 
-*Discussion.* This analysis yields a comprehensive, largely consistent picture of C cycling across the world's forests.
+*Discussion.* This analysis yields a comprehensive and broadly consistent picture of C cycling across the world's forests.
 *[Discussion section will interpret results, highlighting new and significant findings, and discuss implications.
@@
 Moreover, forests contain substantial carbon (C) stocks: an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C [biomass and soils; @bonan_forests_2008]. 
-Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), resulting in net C sink of ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**).
+Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), reducing the net forest sink to ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**).
 The future of this sink is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change. 
-Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change [@grassi_key_2017; Cavaleri et al 2015, Griscom et al 2017].
+Regrowth forests in particular will play an important role [@pugh_role_2019], as these represent a large (~#%) and growing proportion of Earth's forests.
+Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; Cavaleri et al 2015].
 
-Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [Friedlingstein et al 2006; @krause_large_2018; REFS] and CO~2~ accounting frameworks [Pan, REFS] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history.
+Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climate-carbon_2006; @krause_large_2018; REFS] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history.
 These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published. 
@@
 The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*.
-In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle (*e.g.*, Pan et al 2011), or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere (REFS).
+In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere (REFS).
 
@@
 ForC contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables) based on ground-based measurements, along with associated data required for interpretation (*e.g.*, stand history, measurement methods).
-These data have been amalgamated from original peer-reviewed publications, either directly or via intermediary data compilations, with notable recent additions of the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020].
-ForC currently contains **#** (~49,000!) records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.
+These data have been amalgamated from original peer-reviewed publications, either directly or via intermediary data compilations.
+Since the its most recent publication [@anderson-teixeira_forc_2018], ForC has been integrated with two large databases: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020], both of which have also synthesized published forest C data.
+Following these additions, ForC currently contains **#** (~49,000!) records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.
 
@@
 1) to provide a macroscopic overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age.
+Our primary goal is to provide a data-based summary of our current state of knowledge on broad trends in forest C cycling.
 We address three broad questions:
@@
 Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], yet these represented only approximately one-third of records in ForC v.2.0 [@anderson-teixeira_forc_2018].
-Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database (GROA v1.0; Cook-Patton et al.
-in press; ZenodoDOI), and data from several publications (Taylor et al.
-2017, Lutz et al.
-2018, Johnson_2018_csss; see GitHub repo for complete list) ([GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv)).
+Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database [GROA v1.0; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020], and data from several publications [e.g., @taylor_temperature_2017, @lutz_global_2018, @johnson_climate_2018; detailed list at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv].
 We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS). 
@@
 For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
-Climate zones were defined according to Köppen-Geiger zones, which were extracted based on site geographic coordinates from the ESRI Köppen-Geiger map (downloaded June 2014 from http://maps3.arcgisonline.com/ArcGIS/rest/services/A-16/Köppen- Geiger_Observed_and_Predicted_Climate_Shifts/MapServer; Rubel and Kottek 2010).
+Climate zones (Fig. 1) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
 Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd).
 Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis. 
-Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations (SYNMAP; Jung et al 2006) and recorded in the  dominant.veg field of the ForC MEAUREMENTS table.
-Stand age was obtained from the stand.age field of MEASUREMENTS table, and forests classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication).
-Records for which stand age was unknown were excluded from the analysis. 
-These groupings were defined based on ForC v.2.0 [@anderson-teixeira_forc_2018], for which they covered 91% of the primary variable records for forests of known age, or 86% of total records. 
-The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with <400 records total for stands of any age.
+Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
+Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication). 
+Records for which stand age was unknown (#% of total database) were excluded from the analysis. 
+*These groupings were defined based on ForC v.2.0 [@anderson-teixeira_forc_2018], for which they covered 91% of the primary variable records for forests of known age, or 86% of total records.
+The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with <400 records total for stands of any age.*
 

### 2020-09-20 manuscript/ERL_global_C_SI.Rmd
added age trends figures to SI
#33

@@
 ---
-title: ""
+title: "Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database"
 output: 
@@
     fig_caption: yes    
+    toc: true # table of content true
+    toc_depth: 2  
 bibliography: references.bib
@@
 
-**Title:** Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database
-
-Kristina J. Anderson-Teixeira,
-Valentine Herrmann,
-Becky Banbury Morgan,
-Ben Bond-Lamberty,
-Susan Cook-Patton,
-Abigail Ferson,
-Jennifer McGarvey,
-Helene C. Muller-Landau,
-Maria Wang,
-
 \newpage
-## Appendix A. Duplicates and Conflicting Records within ForC
+## Appendix S1. Duplicates and Conflicting Records within ForC
 
@@
 We manually reviewed duplicates that differed only in methodology, assigning precedence to the record employing a more comprehensive approach (*e.g.*, inclusion of understory, lianas, or bamboo as opposed to just trees) or using a favored methodology.
+
+\newpage
+
+## Table S#(variable mapping))
+
+[this table](https://github.com/forc-db/ForC/blob/master/figures/C_cycle_diagrams/ForC_variables_mapping_for_C_cycle_diagrams.csv))
+
+\newpage
+## Figure S1. Age trends and biome differences for $NEP$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S1 | Age trends and biome differences for $NEP$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/NEP.png")
+```
+
+\newpage
+## Figure S2. Age trends and biome differences for $GPP$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S2 | Age trends and biome differences for $GPP$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/GPP.png")
+```
+
+\newpage
+## Figure S3. Age trends and biome differences for $NPP$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S3 | Age trends and biome differences for $NPP$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/NPP.png")
+```
+
+\newpage
+## Figure S4. Age trends and biome differences for $ANPP$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S4 | Age trends and biome differences for $ANPP$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/ANPP.png")
+```
+
+\newpage
+## Figure S5. Age trends and biome differences for $ANPP_{woody}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S5 | Age trends and biome differences for $ANPP_{woody}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/ANPP_woody.png")
+```
+
+\newpage
+## Figure S6. Age trends and biome differences for $ANPP_{stem}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S6 | Age trends and biome differences for $ANPP_{stem}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/ANPP_woody_stem.png")
+```
+
+\newpage
+## Figure S7. Age trends and biome differences for $ANPP_{foliage}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S7 | Age trends and biome differences for $ANPP_{foliage}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/ANPP_foliage.png")
+```
+
+\newpage
+## Figure S8. Age trends and biome differences for $ANPP_{litterfall}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S8 | Age trends and biome differences for $ANPP_{litterfall}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/ANPP_litterfall.png")
+```
+
+\newpage
+## Figure S9. Age trends and biome differences for $BNPP$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S9 | Age trends and biome differences for $BNPP$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/BNPP_root.png")
+```
+
+\newpage
+## Figure S10. Age trends and biome differences for $BNPP_{coarse}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S10 | Age trends and biome differences for $BNPP_{coarse}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/BNPP_root_coarse.png")
+```
+
+\newpage
+## Figure S11. Age trends and biome differences for $BNPP_{fine}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S11 | Age trends and biome differences for $BNPP_{fine}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/BNPP_root_fine.png")
+```
+
+\newpage
+## Figure S12. Age trends and biome differences for $R_{eco}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S12 | Age trends and biome differences for $R_{eco}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/R_eco.png")
+```
+
+\newpage
+## Figure S13. Age trends and biome differences for $R_{root}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S13 | Age trends and biome differences for $R_{root}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/R_auto_root.png")
+```
+
+\newpage
+## Figure S14. Age trends and biome differences for $R_{soil}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S14 | Age trends and biome differences for $R_{soil}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/R_soil.png")
+```
+
+\newpage
+## Figure S15. Age trends and biome differences for $R_{het-soil}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S15 | Age trends and biome differences for $R_{het-soil}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/R_het_soil.png")
+```
+
+\newpage
+## Figure S16. Age trends and biome differences for $B_{tot}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S16 | Age trends and biome differences for $B_{tot}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass.png")
+```
+
+\newpage
+## Figure S17. Age trends and biome differences for $B_{ag}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S17 | Age trends and biome differences for $B_{ag}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass_ag.png")
+```
+
+\newpage
+## Figure S18. Age trends and biome differences for $B_{ag-wood}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S18 | Age trends and biome differences for $B_{ag-wood}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass_ag_woody.png")
+```
+
+\newpage
+## Figure S19. Age trends and biome differences for $B_{foliage}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S19 | Age trends and biome differences for $B_{foliage}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass_ag_foliage.png")
+```
+
+\newpage
+## Figure S20. Age trends and biome differences for $B_{root}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S20 | Age trends and biome differences for $B_{root}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass_root.png")
+```
+
+\newpage
+## Figure S21. Age trends and biome differences for $B_{root-coarse}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S21 | Age trends and biome differences for $B_{root-coarse}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass_root_coarse.png")
+```
+
+\newpage
+## Figure S22. Age trends and biome differences for $B_{root-fine}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S22 | Age trends and biome differences for $B_{root-fine}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass_root_fine.png")
+```
+
+\newpage
+## Figure S23. Age trends and biome differences for $DW_{tot}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S23 | Age trends and biome differences for $DW_{tot}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/deadwood.png")
+```
+
+\newpage
+## Figure S24. Age trends and biome differences for $DW_{standing}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S24 | Age trends and biome differences for $DW_{standing}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/deadwood_standing.png")
+```
+
+\newpage
+## Figure S25. Age trends and biome differences for $DW_{down}$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S25 | Age trends and biome differences for $DW_{down}$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/deadwood_down.png")
+```
+
+\newpage
+## Figure S26. Age trends and biome differences for $OL$
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S26 | Age trends and biome differences for $OL$.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/organic.layer.png")
+```

### 2020-09-20 manuscript/ERL_global_C_review.Rmd
worked on manuscript (mostly results)

@@
 
-Forest ecosystems globally play a critical role in regulating atmospheric carbon dioxide (CO~2~), and thereby will shape the course of climate change (IPCC1.5). 
+Forest ecosystems will play a critical role in shaping the course of climate change (IPCC1.5) through their influence on atmospheric carbon dioxide (CO~2~). 
 Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**).
 While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion was retained in ecosystems over recent decades.
-The resulting CO~2~ sink averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**). 
+The resulting CO~2~ sink averaged 3.0 ± 0.8 Gt C yr^{-1} from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**). 
 Moreover, forests contain substantial carbon (C) stocks: an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C [biomass and soils; @bonan_forests_2008]. 
@@
 Records for which stand age was unknown (#% of total database) were excluded from the analysis. 
-*These groupings were defined based on ForC v.2.0 [@anderson-teixeira_forc_2018], for which they covered 91% of the primary variable records for forests of known age, or 86% of total records.
-The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with <400 records total for stands of any age.*
+These groupings covered *#%* of the primary variable records for forests of known age, or *#%* of total records. 
+The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with *#* records total.
 
-We drew upon records for *#* annual flux and *#* C stock variables (Table 1).
-For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables (Table 1, [this table](https://github.com/forc-db/ForC/blob/master/figures/C_cycle_diagrams/ForC_variables_mapping_for_C_cycle_diagrams.csv))).
-Although ForC contains information that may be used to standardize or control for methodological differences [*e.g.*, area sampled, min stem diameter sampled, allometric equations applied; @anderson-teixeira_forc_2018], for this analysis we included all relevant data in ForC.
+We selected 23 annual flux and 11 C stock variables for inclusion in the analysis, although two flux variables ($R_{het-ag}$ and $R_{het}$) were included for conceptual completeness but had no records in ForC  (Table 1).
+For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables (Tables 1, S#(variable mapping)).
 Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum stem diameter sampled was $\le$ 10cm.
@@
 In generating ForC-simplified, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006). 
-Duplicate or otherwise conflicting records were reconciled as described in *APPENDIX A*. 
+Duplicate or otherwise conflicting records were reconciled as described in *APPENDIX S1*. 
 Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand establishment (*i.e.*, that reflected by stand.age). 
@@
 
-Data were analyzed to produce basic summaries of C cycle patterns across biomes and stand ages following an approach similar to that of @anderson-teixeira_carbon_2016. 
-For mature forests, to obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot, weighting flux measurements according to the length of measurement periods (*i.e.*, end.date - start.date).
-Values were then averaged across plots clustered within 25 km of one another (geographic.area field of SITES table, sensu Anderson-Teixeira et al 2018), weighting by area.sampled (MEASUREMENTS table) or plot.area (PLOTS table) if available for all records.
+Data were analyzed to produce basic summaries of C cycle patterns across biomes and stand ages. **(this par & following are confusing/ incomplete--needs work)**
+For mature forests, to obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot. 
+Values were then averaged across plots clustered within 25 km of one another [*sensu* @anderson-teixeira_forc_2018], weighting by area sampled if available for all records. 
 This step was taken to avoid pseudo-replication and to combine any records from sites with more than one name in ForC. 
@@
 When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another.
-There were enough data to run this analysis for all focal variables but **biomass_ag_understory, BNPP_root.turnover_fine, deadwood_down, total.ecosystem_2.**
+There were enough data to run this analysis for all focal variables but $DW_{down}$.
 
@@
 ```
+
+*NOTE TO COAUTHORS: THIS SECTION HAS A TON OF FIGURES, BUT VERY LITTLE TEXT. IS THERE ANYTHING THAT SHOULD BE ADDED?*
+
 ### Data Coverage
@@
 Of the *#* records in ForC v.#, *#* met the criteria for inclusion in this study (Fig. 1). 
-*(Give some stats on coverage by age and biome)*
+These records were distributed across *#* plots in *#* distinct geographic areas. 
+Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, *#* fluxes and *#* stocks in temperate broadleaf forests, *#* fluxes and *#* stocks in temperate conifer forests, and fluxes and *#* stocks in boreal forests.
+For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, *#* fluxes and *#* stocks in temperate broadleaf forests, *#* fluxes and *#* stocks in temperate conifer forests, and fluxes and *#* stocks in boreal forests.
 
@@
 
-Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5, and statistics for each biome type are also summarized at: [GitHub URL].
-Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, *#* fluxes and *#* stocks in temperate broadleaf forests, *#* fluxes and *#* stocks in temperate conifer forests, and fluxes and *#* stocks in boreal forests.
+Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5 (and available in tabular format in the ForC release accompanying this publication).
+
 For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets were generally consistent. 
-That is, component fluxes and stocks summed to within 1 std of more inclusive fluxes in all but one instance (in temperate conifer forests, $aboveground woody biomass + foliage biomass > aboveground biomass + 1std$; Fig.
+That is, component fluxes and stocks summed to within one standard deviation of more inclusive fluxes in all but two instances, both for living C stocks in temperate conifer forests (Fig.
 5).
-**update this: https://github.com/forc-db/ERL-review/issues/16 **
+**(check this with final results: https://github.com/forc-db/ForC/blob/master/numbers_and_facts/C_cycle_closure.csv) **
+Within this biome, $B_{ag}<B_{ag-wood}+B_{foliage}$ and $B_{root}<B_{root-coarse}+B_{root-fine}$ because $B_{ag-wood}$ and $B_{root-coarse}$ were very high, with strongly disproportionate numbers of records from the high-biomass forests of the US Pacific Northwest (Figs.
+S18, S21).
 
@@
 
-(**check paragraph with latest data**) The largest C fluxes—including $GPP$, $NPP$, $ANPP$, $BNPP$, $R_{soil_het}$, $R_{soil}$, $R_{eco}$, — were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests  ([ForC_variable_averages_per_Biome](https://github.com/forc-db/ForC/blob/master/numbers_and_facts/ForC_variable_averages_per_Biome.csv)) (Fig.
-6).
-The same held true for some of the subsidiary fluxes: ANPP_foliage, ANPP_woody, [**OTHERS?**].
-Other subsidiary fluxes—including NPP_woody, ANPP_repro, ANPP_stem, ANPP_branch, woody.mortality, BNPP_coarse, BNPP_fine,  [**OTHERS?**]—deviated from this pattern and/or lacked data for some biomes.
-Net ecosystem production ($NEP$) did not follow this pattern, with no significant differences across biomes but the largest average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests.
-Thus, C cycling rates generally decreased from tropical to temperate to boreal forests, but with less apparent trends for some of the subsidiary fluxes and an important exception in the overall C balance ($NEP$).
+**(The following covers only the variables with age trends figures; there maybe a few more without much age data)**
+With only # exceptions, C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs. 6, S1-S15).
+In this majority of cases, differences between tropical and boreal forests were always significant, with temperate forests intermediate and significantly different from one or both. 
+Fluxes tended to be greater in temperate broadleaf than conifer forests, but the difference was never significant.
+This pattern held for the following variables: $GPP$, $NPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{foliage}$, $ANPP_{litterfall}$, $BNPP$, $BNPP_fine$, $R_{eco}$, $R_{root}$, $R_{soil}$, and $R_{het-soil}$.
+The most notable exception to this pattern was $NEP$, with no significant differences across biomes but with the largest average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests (Figs.
+5,S1).
+Another exception was for $BNPP_{root-coarse}$, where all records came from high-biomass forests in the US Pacific Northwest, resulting in high values for the temperate conifer biome and no significant differences across biomes (Fig.
+S10).
+Thus, C cycling rates generally decreased from tropical to temperate to boreal forests, with the important exception in the overall C balance ($NEP$).
 
@@
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
-Lines...
-.", out.width='100%', fig.pos='H'}
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.
+Individual figures for each flux with sufficient data given in the Supplement (Figs.
+S1-S15).", out.width='100%', fig.pos='H'}
 
@@
 
-There were less distinct trends in C stocks across biomes (Fig. 7). 
-Biome means for live aboveground and total biomass followed the same general trend as the major fluxes, with biomass of tropical= temperate broadleaf $\ge$ temperate conifer> boreal forests (Fig.
-7a).
-However, the relative differences in these means were much smaller than for the major C fluxes, and there was significant variation within biomes. 
-Maximum aboveground biomass values followed a very different trend than the means: temperate broadleaf> temperate conifer>boreal>tropical. 
-There were some statistically significant biome differences in less frequently sampled C stocks (*e.g.*, woody biomass, foliage biomass, deadwood), but given high within-biome variability in C stocks and relatively low sample sizes, these were likely attributable to sampling biases and methodological differences than to true differences across biomes.
+There were fewer distinct trends in C stocks across biomes (Figs. 7, S16-S26). 
+Specifically, there were significant differences for only *#* of *#* variables: *LIST*. 
+In all of these cases, temperate conifer forests were in the highest significance grouping, and boreal forests in the lowest. 
+Tropical and temperate broadleaf forests fell in between--most commonly being significantly different from temperate conifer but not boreal forests.
+... 
+*(These differences are driven more by geographic distribution of sampling than by true differences.)*
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.", out.width='100%', fig.pos='H'}
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.
+Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.
+Individual figures for each stock with sufficient data given in the Supplement (Figs.
+S16-S26).", out.width='100%', fig.pos='H'}
 
@@
 (**check paragraph with latest data**) Average C cycles for forests <100 years old are presented in Figures 8-11.
-In general, ForC contained roughly comparable information on C fluxes in regrowth vs mature forests, with the notable exception of tropical forests, for which there were no fluxes that had been measured in $\ge$ 7 different geographic areas.
-C stocks were better documented for regrowth stands, with a total of # records from # distinct geographic areas.
-Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Fig.
-5-10; detailed below).
+Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Table 1, Figs.
+5-10, S1-S26; detailed below).
 Differences across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests (Figs.

### 2020-09-20 manuscript/ERL_global_C_review.Rmd
small update

@@
 This pattern held for the following variables: $GPP$, $NPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{foliage}$, $ANPP_{litterfall}$, $BNPP$, $BNPP_fine$, $R_{eco}$, $R_{root}$, $R_{soil}$, and $R_{het-soil}$.
+
 The most notable exception to this pattern was $NEP$, with no significant differences across biomes but with the largest average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests (Figs.
@@
 Tropical and temperate broadleaf forests fell in between--most commonly being significantly different from temperate conifer but not boreal forests.
-... 
-*(These differences are driven more by geographic distribution of sampling than by true differences.)*
+Once again, the high values for the temperate conifer biome were driven by the very high-biomass forests of the US Pacific Northwest, which are disproportionately represented in the current version of ForC.
+Thus, biome differences should be interpreted more as driven more by geographic distribution of sampling than by true differences.)*
 
@@
 Thus, overall, a high degree of closure implies that ForC gives a consistent picture of C cycling within biomes. 
-While these means are unlikely to be accurate representations of C cycling within any particular forest, they provide a useful baseline for comparison.
+While these means are unlikely to be accurate representations of C cycling within any particular forest, they provide a useful baseline for comparison, always keeping in mind that sample means do not necessarily represent the true mean of the entire biome.
 
@@
 In contrast to C fluxes and accumulation rates in regrowth forests, stocks...
+
+Thus, biome differences should be interpreted more as driven more by geographic distribution of sampling than by true differences.
 

### 2020-09-21 manuscript/ERL_global_C_review.Rmd
Update ERL_global_C_review.Rmd

@@
 *Review Results/ Synthesis.*  ForC v.XX yielded a fairly comprehensive picture of C cycling in the world's major forest biomes, with broad closure of C budgets.
-The rate of C cycling generally increased from boreal to tropical regions, whereas C stocks showed less directional variation.
+The rate of C cycling generally increased from boreal to tropical regions in both mature and regrowth forests, whereas C stocks showed less directional variation.
 The majority of flux variables, together with most live biomass pools, increased significantly with stand age, *and the rate of increase again tended to increase from boreal to tropical regions*.
@@
 
-(**check paragraph with latest data**) Average C cycles for forests <100 years old are presented in Figures 8-11.
-Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Table 1, Figs.
-5-10, S1-S26; detailed below).
-Differences across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests (Figs.
-6-7).
+Average C cycles for forests <100 years old are presented in Figures 8-11.  
+Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Table 1, Figs. 6-11, S1-S26; detailed below).
 
-(**check paragraph with latest data**) ForC contained 14 flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig.
-6-7 and **S#- SI figures including plots for all variables**).
-Of these, 9 increased significantly with log10[stand.age]: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($NEP$).
+ForC contained *14* flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig. 6-7 and **S#- SI figures including plots for all variables**). 
+Of these, *9* increased significantly with log10[stand.age]: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($NEP$).
 The remaining five--$ANPP_{woody-branch}$, $BNPP_{root-coarse}$, $R_{soil-het}$, and $R_{soil-het}$-–displayed no significant relationship to stand age, although all displayed a positive trend.
-In terms of C stocks, 10 variables had sufficient data to test for age trends.
-Six of these—total biomass, aboveground biomass, aboveground woody biomass, foliage biomass, root biomass, and coarse root biomass—increased significantly with $log10[stand.age]$.
-The remaining four displayed non-significant positive trends:  fine root biomass, total dead wood, standing dead wood, and organic layer.
+
+Differences in C fluxes across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests.  
+The single exception was $ANPP_{stem}$, for which temperate broadleaf forests and temperate conifer forests of age >*~30* had slightly higher flux rates than tropical forests (*represented by a single chronosequence*).
+Notably, the trend of tropical > temperate > boreal held for $NEP$ in regrowth forests, in contrast to the lack of biome differences in $NEP$ for mature forests (Fig. 6).
+
+There were only *#* flux variables with sufficient data to test for biome x age interactions: $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{litterfall}$, and $BNPP$.
+**(more could be added if age trends become significant after outliers are resolved)**
+For three of these ($ANPP$, $ANPP_{litterfall}$, $BNPP$), the increase in C flux with age was steepest increase in tropical forests, followed by temperate and then boreal forests (Figs S#). 
+Similarly, $ANPP_{woody}$ displayed a steeper increase with age in temperate than boreal boreal forests (no tropical data for this variable).
+In contrast, for $ANPP_{stem}$,  tropical and temperate broadleaf forests had the highest flux rates at young ages, but were surpassed by temperate conifer forests between ages 20 and 50 (Fig. S6).
+
+(**this needs to be updated with latest data**) 
+In terms of C stocks, *10* variables had sufficient data to test for age trends. 
+*Six* of these—*total biomass, aboveground biomass, aboveground woody biomass, foliage biomass, root biomass, and coarse root biomass*— increased significantly with $log10[stand.age]$. 
+*The remaining four displayed non-significant positive trends:  fine root biomass, total dead wood, standing dead wood, and organic layer.*
+*(discuss rates of increase)*
 
@@
 All researchers whose data is included in ForC and this analysis.
+Ian McGregor for help with the database.
 Thanks to Norbert Kunert and [Helene's intern] for helpful input at an earlier phase.

### 2020-09-21 manuscript/ERL_global_C_review.Rmd
knit in filled version of table 1

@@
 The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Figs. 6-11). 
-Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks of mature forests are influenced by a different set of drivers.
+Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks and $NEP$ of mature forests are influenced by a different set of drivers.
 

### 2020-09-23 manuscript/ERL_global_C_review.Rmd
switched to pulling C cycle diagrams from ForC
#23 -
@ValentineHerr , you don't need to save the diagrams in the ERL repo now-- just ForC is fine.

@@
 
-Susan Cook-Patton
+Susan C. Cook-Patton^5^
 
@@
 2. Center for Tropical Forest Science-Forest Global Earth Observatory; Smithsonian Tropical Research Institute; Panama, Republic of Panama
+3.
+4.
+5. The Nature Conservancy, Arlington VA 22203
+6.
 
@@
 
-  knitr::include_graphics("tables_figures/Tropical broadleaf MATURE.png")
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Tropical broadleaf MATURE.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/Temperate broadleaf MATURE.png")
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate broadleaf MATURE.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/Temperate conifer MATURE.png")
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate conifer MATURE.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/Boreal conifer MATURE.png")
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Boreal conifer MATURE.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/Tropical broadleaf YOUNG.png")
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Tropical broadleaf YOUNG.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/Temperate broadleaf YOUNG.png")
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate broadleaf YOUNG.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/Temperate conifer YOUNG.png")
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate conifer YOUNG.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/Boreal conifer YOUNG.png")
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Boreal conifer YOUNG.png")
 ```

### 2020-09-24 manuscript/ERL_global_C_review.Rmd
fill in full author names & affiliations

@@
 
-Becky Banbury Morgan
+Becky Banbury Morgan^3^
 
-Ben Bond-Lamberty
+Ben Bond-Lamberty^4
 
@@
 
-Abigail Ferson
+Abigail E. Ferson^1,6^
 
-Jennifer McGarvey
+Jennifer C. McGarvey^1^
 
@@
 
-Maria Wang
+Maria M. H. Wang^1,7^
 
@@
 2. Center for Tropical Forest Science-Forest Global Earth Observatory; Smithsonian Tropical Research Institute; Panama, Republic of Panama
-3.
-4.
-5. The Nature Conservancy, Arlington VA 22203
-6.
+3. School of Geography, University of Leeds, Leeds, UK
+4. Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park Maryland 20740, USA 
+5. The Nature Conservancy; Arlington VA 22203, USA
+6. College of Natural Resources, University of Idaho; Moscow, Idaho 83843, USA
+7. Grantham Centre for Sustainable Futures and Department of Animal and Plant Sciences, University of Sheffield, Western Bank, Sheffield, South Yorkshire S10 2TN, UK
 
@@
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
+Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
 Arrows indicate fluxes, boxes indicate stocks.
-When age trends are signficant, they are presented with numbered equations; otherwise, means are presented.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}

### 2020-09-24 manuscript/ERL_global_C_review.Rmd
auto-fill some stats
#29

@@
 
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+#READING IN TABLES TO GET STATS FOR TEXT
+#Table 1 for this manuscript
+T1 <- read.csv("tables_figures/C_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
+n_records <- sum(T1[,"records"])
+
+#ForC measurements
+ForC_measurements <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_measurements.csv", stringsAsFactors = FALSE, check.names = FALSE)
+ForC_plots <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_plots.csv", stringsAsFactors = FALSE, check.names = FALSE)
+ForC_sites <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_sites.csv", stringsAsFactors = FALSE, check.names = FALSE)
+n_areas <- max(ForC_sites[,"geographic.area"], na.rm = TRUE )
+```
+
 \newpage
@@
 *Methods/Design.* Here, we draw from the Global Forest C Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation.
-Specifically, we draw upon *##* records from *##* geographic locations representing *##* C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests.
+Specifically, we draw upon `r n_records` records from *##* geographic locations representing `r nrow(T1)` C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests.
 For regrowth forests, we quantify age trends for all variables.
@@
 Since the its most recent publication [@anderson-teixeira_forc_2018], ForC has been integrated with two large databases: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020], both of which have also synthesized published forest C data.
-Following these additions, ForC currently contains **#** (~49,000!) records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.
+Following these additions, ForC currently contains `r nrow(ForC_measurements)` records from `r nrow(ForC_plots)` plots and `r n_areas` distinct geographic areas representing all forested biogeographic and climate zones.
 
@@
 
-Of the *#* records in ForC v.#, *#* met the criteria for inclusion in this study (Fig. 1). 
+Of the `r nrow(ForC_measurements)` in ForC v.#, `r n_records` met the criteria for inclusion in this study (Fig. 1). 
 These records were distributed across *#* plots in *#* distinct geographic areas. 

### 2020-09-24 manuscript/ERL_global_C_review.Rmd
created variables for numbers needed in text, filled in more stats
#29

--- 2020-09-24 manuscript/ERL_global_C_review.Rmd
@@
 
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+#READING IN TABLES TO GET STATS FOR TEXT
+
+#Table 1 for this manuscript
+T1 <- read.csv("tables_figures/C_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
+n_records <- sum(T1[,"records"])
+
+#ForC raw
+ForC_measurements <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_measurements.csv", stringsAsFactors = FALSE, check.names = FALSE)
+ForC_plots <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_plots.csv", stringsAsFactors = FALSE, check.names = FALSE)
+ForC_sites <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_sites.csv", stringsAsFactors = FALSE, check.names = FALSE)
+n_areas <- max(ForC_sites[,"geographic.area"], na.rm = TRUE )
+
+#ForC simplified
+ForC_simplified <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/ForC_simplified/ForC_simplified.csv", stringsAsFactors = FALSE, check.names = FALSE)
+n_no_history <- sum(ForC_simplified[,"history.no.info"]) #(relative to ForC_simplified)
+n_managed_in_ForC_simplified <- sum(ForC_simplified[,"managed"])
+n_disturbed_in_ForC_simplified <- sum(ForC_simplified[,"disturbed"])
+records_eligible_for_inclusion <- NA  #number in ForC_simplified that are not disturbed, managed, or no.history
+
+#stats I haven't yet acquired:
+n_records_age_unk <- NA  #total n records no age info
+n_records_coord_unk <- NA  #total n records no geographic coordinates
+
+#total numbers of plots and areas (all variables combined)
+n_plots_tot <- NA
+n_areas_tot <- NA
+
+# percent eligible for inclusion in analysis
+
+n_focal_variable_records_eligible_for_inclusion <- NA #This is prior to assigning biome groupings
+
+# records by biome grouping
+n_records_BoB <- NA
+n_records_Bomixed <- NA
+n_records_Te_mixed <- NA
+
+#numbers of variables with records from ≥7 geographic areas, by biomes/ age class:
+n_fluxes_matureTrB_with_ge7_areas <- NA
+n_stocks_matureTrB_with_ge7_areas <- NA
+n_fluxes_matureTeB_with_ge7_areas <- NA
+n_stocks_matureTeB_with_ge7_areas <- NA
+n_fluxes_matureTeN_with_ge7_areas <- NA
+n_stocks_matureTeN_with_ge7_areas <- NA
+n_fluxes_matureBoN_with_ge7_areas <- NA
+n_stocks_matureBoN_with_ge7_areas <- NA
+
+n_fluxes_youngTrB_with_ge7_areas <- NA
+n_stocks_youngTrB_with_ge7_areas <- NA
+n_fluxes_youngTeB_with_ge7_areas <- NA
+n_stocks_youngTeB_with_ge7_areas <- NA
+n_fluxes_youngTeN_with_ge7_areas <- NA
+n_stocks_youngTeN_with_ge7_areas <- NA
+n_fluxes_youngBoN_with_ge7_areas <- NA
+n_stocks_youngBoN_with_ge7_areas <- NA
+
+```
+
 \newpage
@@
 *Methods/Design.* Here, we draw from the Global Forest C Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation.
-Specifically, we draw upon *##* records from *##* geographic locations representing *##* C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests.
+Specifically, we draw upon `r n_records` records from `r n_areas_tot` geographic locations representing `r nrow(T1)` C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests.
 For regrowth forests, we quantify age trends for all variables.
@@
 
-For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
-Climate zones (Fig. 1) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
-Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd).
-Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis. 
-Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
-Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication). 
-Records for which stand age was unknown (#% of total database) were excluded from the analysis. 
-These groupings covered *#%* of the primary variable records for forests of known age, or *#%* of total records. 
-The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with *#* records total.
+Analyses drew from ForC-simplified (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which is a rearrangement of ForC intended to facilitate analyses. 
+In generating ForC-simplified, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006). 
+Duplicate or otherwise conflicting records were reconciled as described in *APPENDIX S1*, resulting in a total of `r nrow(ForC_simplified)` records (`r round(nrow(ForC_simplified)/nrow(ForC_measurements)*100,digits=1)`% size of total database).
+Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand initiation event. 
+Specifically, we removed all plots flagged as managed in ForC-simplified (`r round(n_managed_in_ForC_simplified/nrow(ForC_simplified)*100,digits=1)`%). 
+This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.
+Plots flagged as disturbed in ForC-simplified (`r round(n_disturbed_in_ForC_simplified/nrow(ForC_simplified)*100,digits=1)`%) included stands that had undergone anthropogenic thinning or partial harvest unless this was very minor.
+We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods.
+We removed all plots for which no stand history information had been retrieved (`r round(n_no_history/nrow(ForC_simplified)*100,digits=1)`%). 
+In total, this resulted in `r records_eligible_for_inclusion` records (`r round(records_eligible_for_inclusion/nrow(ForC_measurements)*100,digits=1)`% of the records in the database) being eligible for inclusion in the analysis.
 
 We selected 23 annual flux and 11 C stock variables for inclusion in the analysis, although two flux variables ($R_{het-ag}$ and $R_{het}$) were included for conceptual completeness but had no records in ForC (Table 1).
+Records for these variables represented `r round(n_focal_variable_records_eligible_for_inclusion/records_eligible_for_inclusion*100,digits=1)`% of the total records eligible for inclusion.
 For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables (Tables 1, S#(variable mapping)).
@@
 
-Analyses drew from ForC-simplified (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which is a rearrangement of ForC intended to facilitate analyses. 
-In generating ForC-simplified, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006). 
-Duplicate or otherwise conflicting records were reconciled as described in *APPENDIX S1*. 
-Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand establishment (*i.e.*, that reflected by stand.age). 
-Specifically, we removed all plots flagged as managed in ForC-simplified (managed field). 
-This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.
-Plots flagged as disturbed in ForC-simplified included stands that had undergone anthropogenic thinning or partial harvest (“Cut” or “Harvest” codes) unless this was very minor (percent.mortality= “minor”).
-We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods. 
-We also removed all plots for which no stand history information had been retrieved.
+For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
+Climate zones (Fig. 1) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
+Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd).
+Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis. 
+Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
+Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication). 
+Assigning stands to these groupings required the exclusion of records for which ForC lacked records of stand age (n=`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of total database), or geographic coordinates (n=`r round(n_records_coord_unk/nrow(ForC_measurements)*100, digits=1)`% of total database).
+In total, our analysis retained `r round(n_records/n_focal_variable_records_eligible_for_inclusion*100, digits=1)` of the focal variable records for forests of known age. 
+The most well-represented forest types excluded were boreal broadleaf (`r n_records_BoB` records) and boreal and temperate mixed broadleaf-needleleaf (`r n_records_Bomixed` and `r n_records_Te_mixed` records, respectively).
 
@@
 
-Of the *#* records in ForC v.#, *#* met the criteria for inclusion in this study (Fig. 1). 
-These records were distributed across *#* plots in *#* distinct geographic areas. 
-Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, *#* fluxes and *#* stocks in temperate broadleaf forests, *#* fluxes and *#* stocks in temperate conifer forests, and fluxes and *#* stocks in boreal forests.
-For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, *#* fluxes and *#* stocks in temperate broadleaf forests, *#* fluxes and *#* stocks in temperate conifer forests, and fluxes and *#* stocks in boreal forests.
+Of the `r nrow(ForC_measurements)` in ForC v.#, `r n_records` met the criteria for inclusion in this study (Fig. 1). 
+These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas. 
+Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained mature forest estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_matureTrB_with_ge7_areas` fluxes and `r n_stocks_matureTrB_with_ge7_areas` stocks in tropical broadleaf forests, `r n_fluxes_matureTeB_with_ge7_areas` fluxes and `r n_stocks_matureTeB_with_ge7_areas` stocks in temperate broadleaf forests, `r n_fluxes_matureTeN_with_ge7_areas` fluxes and `r n_stocks_matureTeN_with_ge7_areas` stocks in temperate conifer forests, and `r n_fluxes_matureBoN_with_ge7_areas` fluxes and `r n_stocks_matureBoN_with_ge7_areas` stocks in boreal forests.
+For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_youngTrB_with_ge7_areas` fluxes and `r n_stocks_youngTrB_with_ge7_areas` stocks in tropical broadleaf forests, `r n_fluxes_youngTeB_with_ge7_areas` fluxes and `r n_stocks_youngTeB_with_ge7_areas` stocks in temperate broadleaf forests, `r n_fluxes_youngTeN_with_ge7_areas` fluxes and `r n_stocks_youngTeN_with_ge7_areas` stocks in temperate conifer forests, and `r n_fluxes_youngBoN_with_ge7_areas` fluxes and `r n_stocks_youngBoN_with_ge7_areas` stocks in boreal forests.
 
@@
 **(The following covers only the variables with age trends figures; there maybe a few more without much age data)**
-With only # exceptions, C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs. 6, S1-S15).
+With only ## exceptions, C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs. 6, S1-S15).
 In this majority of cases, differences between tropical and boreal forests were always significant, with temperate forests intermediate and significantly different from one or both. 
@@
 There were fewer distinct trends in C stocks across biomes (Figs. 7, S16-S26). 
-Specifically, there were significant differences for only *#* of *#* variables: *LIST*. 
+Specifically, there were significant differences for only *##* of *##* variables: *LIST*. 
 In all of these cases, temperate conifer forests were in the highest significance grouping, and boreal forests in the lowest. 
@@
 
-There were only *#* flux variables with sufficient data to test for biome x age interactions: $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{litterfall}$, and $BNPP$.
+There were only *##* flux variables with sufficient data to test for biome x age interactions: $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{litterfall}$, and $BNPP$.
 **(more could be added if age trends become significant after outliers are resolved)**

### 2020-09-24 manuscript/ERL_global_C_review.Rmd
small update

@@
 output: 
-  pdf_document:
+  word_document:
     fig_caption: yes
@@
 
-Of the `r nrow(ForC_measurements)` in ForC v.#, `r n_records` met the criteria for inclusion in this study (Fig. 1). 
+Of the `r nrow(ForC_measurements)` records in ForC v.#, `r n_records` met the criteria for inclusion in this study (Fig. 1). 
 These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas. 

### 2020-09-24 manuscript/ERL_global_C_review.Rmd
calculate stats
#29

@@
 ForC_measurements <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_measurements.csv", stringsAsFactors = FALSE, check.names = FALSE)
+  n_records_age_unk <- sum(is.na(as.numeric(ForC_measurements[, "stand.age"])))  #total n records no age info
 ForC_plots <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_plots.csv", stringsAsFactors = FALSE, check.names = FALSE)
 ForC_sites <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_sites.csv", stringsAsFactors = FALSE, check.names = FALSE)
+  n_sites_coord_unk <- sum(is.na(as.numeric(ForC_sites[, "lat"])))  #total n records no geographic coordinates
   n_areas <- max(ForC_sites[,"geographic.area"], na.rm = TRUE )
@@
 records_eligible_for_inclusion <- NA  #number in ForC_simplified that are not disturbed, managed, or no.history
-
-#stats I haven't yet acquired:
-n_records_age_unk <- NA  #total n records no age info
-n_records_coord_unk <- NA  #total n records no geographic coordinates
 
@@
 ```
+
+*(Abby has offered to update stats in this paragraph:)*
 
@@
 Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication). 
-Assigning stands to these groupings required the exclusion of records for which ForC lacked records of stand age (n=`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of total database), or geographic coordinates (n=`r round(n_records_coord_unk/nrow(ForC_measurements)*100, digits=1)`% of total database).
+Assigning stands to these groupings required the exclusion of records for which ForC lacked records of stand age (n=`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database), or geographic coordinates (n=`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database).
 In total, our analysis retained `r round(n_records/n_focal_variable_records_eligible_for_inclusion*100, digits=1)` of the focal variable records for forests of known age. 

### 2020-09-25 manuscript/ERL_global_C_review.Rmd
Update ERL_global_C_review.Rmd

@@
 
-#Table 1 for this manuscript
-T1 <- read.csv("tables_figures/C_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
-n_records <- sum(T1[,"records"])
-
 #ForC raw
@@
   n_records_age_unk <- sum(is.na(as.numeric(ForC_measurements[, "stand.age"])))  #total n records no age info
+  n_SRDB=NA
+  n_GROA=NA
 ForC_plots <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_plots.csv", stringsAsFactors = FALSE, check.names = FALSE)
@@
 
-#total numbers of plots and areas (all variables combined)
-n_plots_tot <- NA
-n_areas_tot <- NA
+# limiting to variables and biomes of interest:
+n_focal_variable_records_eligible_for_inclusion <- NA # records_eligible_for_inclusion, limited to focal variables
 
-# percent eligible for inclusion in analysis
-
-n_focal_variable_records_eligible_for_inclusion <- NA #This is prior to assigning biome groupings
-
-# records by biome grouping
+# records for excluded biome groupings ? 
 n_records_BoB <- NA
@@
 n_records_Te_mixed <- NA
+
+#Records included in the analysis
+T1 <- read.csv("tables_figures/C_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
+  n_records <- sum(T1[,"records"])
+
+n_plots_tot <- NA #total numbers of plots used in the analysis (all variables combined)
+n_areas_tot <- NA #total numbers of areas used in the analysis (all variables combined)
 
@@
 Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], yet these represented only approximately one-third of records in ForC v.2.0 [@anderson-teixeira_forc_2018].
-Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database [GROA v1.0; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020], and data from several publications [e.g., @taylor_temperature_2017, @lutz_global_2018, @johnson_climate_2018; detailed list at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv].
+Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##, `r n_SRDB` records; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database [GROA v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020], and data from several publications [e.g., @taylor_temperature_2017, @lutz_global_2018, @johnson_climate_2018; detailed list at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv].
 We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS). 
@@
 
-Data were analyzed to produce basic summaries of C cycle patterns across biomes and stand ages. **(this par & following are confusing/ incomplete--needs work)**
+Data were analyzed to produce ...
+*Schematics:* 
+Data were summarized to produce schematics of C cycling across the eight biome - age group combinations identified above.
 For mature forests, to obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot. 
@@
 This step was taken to avoid pseudo-replication and to combine any records from sites with more than one name in ForC.
+
+*FIgs 6-7*
 Finally, for figures 6 and 7, the original values were analyzed via a linear mixed effects model ('lmer' function in 'lme4' R package) with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
@@
 When the effect of stand.age was significant at p $\le$ 0.05 and when each biome had records for stands of at least 10 different ages, a biome - stand.age interaction was included in the model.
-In the C cycle schematics for young forests, we report equations based on these models. 
-In cases where there was no significant effect of stand.age, records were averaged as for mature stands.
 

### 2020-09-25 manuscript/ERL_global_C_review.Rmd
drop some stats
#29

@@
 n_focal_variable_records_eligible_for_inclusion <- NA # records_eligible_for_inclusion, limited to focal variables
-
-# records for excluded biome groupings ? 
-n_records_BoB <- NA
-n_records_Bomixed <- NA
-n_records_Te_mixed <- NA
 
@@
 In total, our analysis retained `r round(n_records/n_focal_variable_records_eligible_for_inclusion*100, digits=1)` of the focal variable records for forests of known age.
-The most well-represented forest types excluded were boreal broadleaf (`r n_records_BoB` records) and boreal and temperate mixed broadleaf-needleleaf (`r n_records_Bomixed` and `r n_records_Te_mixed` records, respectively).
 
-Data were analyzed to produce ...
+Data were analyzed to produce 
 *Schematics:* 

### 2020-09-25 manuscript/ERL_global_C_review.Rmd
update and add a couple summary tables #29

@@
 #READING IN TABLES TO GET STATS FOR TEXT
+path_to_ForC <- paste0(dirname(dirname(getwd())), "/ForC/") # this is to allow anybody to knit (assuming current working direcoty is in forc-db/ERL-review/manuscript and its "nesting" withing the ForC repo does not change)
 
 #ForC raw
-ForC_measurements <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_measurements.csv", stringsAsFactors = FALSE, check.names = FALSE)
+ForC_measurements <- read.csv(paste0(path_to_ForC, "/data/ForC_measurements.csv"), stringsAsFactors = FALSE, check.names = FALSE)
   n_records_age_unk <- sum(is.na(as.numeric(ForC_measurements[, "stand.age"])))  #total n records no age info
-  n_SRDB=NA
-  n_GROA=NA
-ForC_plots <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_plots.csv", stringsAsFactors = FALSE, check.names = FALSE)
-ForC_sites <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_sites.csv", stringsAsFactors = FALSE, check.names = FALSE)
+  n_records_age_0 <- sum(as.numeric(ForC_measurements[, "stand.age"]) %in% 0)
+
+  n_SRDB <- sum(ForC_measurements$ForC.investigator %in% "Ben Bond-Lamberty")
+  n_GROA <- sum(ForC_measurements$ForC.investigator %in% "Dr. Susan Cook-Patton")
+
+  ForC_plots <- read.csv(paste0(path_to_ForC, "/data/ForC_plots.csv"), stringsAsFactors = FALSE, check.names = FALSE)
+ForC_sites <- read.csv(paste0(path_to_ForC, "/data/ForC_sites.csv"), stringsAsFactors = FALSE, check.names = FALSE)
+
 n_sites_coord_unk <- sum(is.na(as.numeric(ForC_sites[, "lat"])))  #total n records no geographic coordinates
-  n_areas <- max(ForC_sites[,"geographic.area"], na.rm = TRUE )
+
+n_areas <- length(unique(na.omit(ForC_sites[,"geographic.area"])))
 
 #ForC simplified
-ForC_simplified <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/ForC_simplified/ForC_simplified.csv", stringsAsFactors = FALSE, check.names = FALSE)
+ForC_simplified <- read.csv(paste0(path_to_ForC, "ForC_simplified/ForC_simplified.csv"), stringsAsFactors = FALSE, check.names = FALSE)
+
 n_no_history <- sum(ForC_simplified[,"history.no.info"]) #(relative to ForC_simplified)
@@
 n_disturbed_in_ForC_simplified <- sum(ForC_simplified[,"disturbed"])
-records_eligible_for_inclusion <- NA  #number in ForC_simplified that are not disturbed, managed, or no.history
+records_eligible_for_inclusion <- sum(ForC_simplified$managed ==0 & ForC_simplified$disturbed ==0 & ForC_simplified$history.no.info == 0)  #number in ForC_simplified that are not disturbed, managed, or no.history
 
 # limiting to variables and biomes of interest:
-n_focal_variable_records_eligible_for_inclusion <- NA # records_eligible_for_inclusion, limited to focal variables
+ Biomes_n_records <- read.csv("tables_figures/Biomes_n_records_variables_of_interest.csv", check.names = F)
+n_focal_variable_records_eligible_for_inclusion <- sum(Biomes_n_records$n_records) # n records_eligible_for_inclusion, limited to focal variables
 
@@
 
-n_plots_tot <- NA #total numbers of plots used in the analysis (all variables combined)
-n_areas_tot <- NA #total numbers of areas used in the analysis (all variables combined)
+global_counts <- read.csv("tables_figures/Count_n_record_n_plot_n_area.csv", check.names = FALSE)
+n_records - global_counts$n_records # one count difference... not sure why. probably the filter thing... will ignore dor now.
+n_plots_tot <- global_counts$n_plots #total numbers of plots used in the analysis (all variables combined)
+n_areas_tot <- global_counts$n_areas #total numbers of areas used in the analysis (all variables combined)
 
 #numbers of variables with records from ≥7 geographic areas, by biomes/ age class:
-n_fluxes_matureTrB_with_ge7_areas <- NA
-n_stocks_matureTrB_with_ge7_areas <- NA
-n_fluxes_matureTeB_with_ge7_areas <- NA
-n_stocks_matureTeB_with_ge7_areas <- NA
-n_fluxes_matureTeN_with_ge7_areas <- NA
-n_stocks_matureTeN_with_ge7_areas <- NA
-n_fluxes_matureBoN_with_ge7_areas <- NA
-n_stocks_matureBoN_with_ge7_areas <- NA
+ForC_biome_averages <- read.csv(paste0(path_to_ForC, "numbers_and_facts/ForC_variable_averages_per_Biome.csv"), stringsAsFactors = F)
 
-n_fluxes_youngTrB_with_ge7_areas <- NA
-n_stocks_youngTrB_with_ge7_areas <- NA
-n_fluxes_youngTeB_with_ge7_areas <- NA
-n_stocks_youngTeB_with_ge7_areas <- NA
-n_fluxes_youngTeN_with_ge7_areas <- NA
-n_stocks_youngTeN_with_ge7_areas <- NA
-n_fluxes_youngBoN_with_ge7_areas <- NA
-n_stocks_youngBoN_with_ge7_areas <- NA
+idx_fluxes <- ForC_biome_averages$n.areas >= 7 & ForC_biome_averages$variable.type %in% "flux"
+idx_stocks <- ForC_biome_averages$n.areas >= 7 & ForC_biome_averages$variable.type %in% "stock"
+
+n_fluxes_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_fluxes])
+n_stocks_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_stocks])
 
@@
 Assigning stands to these groupings required the exclusion of records for which ForC lacked records of stand age (n=`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database), or geographic coordinates (n=`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database).
+We also excluded records of stand age = 0 year (n=`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
 In total, our analysis retained `r round(n_records/n_focal_variable_records_eligible_for_inclusion*100, digits=1)` of the focal variable records for forests of known age.
@@
 These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas. 
-Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained mature forest estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_matureTrB_with_ge7_areas` fluxes and `r n_stocks_matureTrB_with_ge7_areas` stocks in tropical broadleaf forests, `r n_fluxes_matureTeB_with_ge7_areas` fluxes and `r n_stocks_matureTeB_with_ge7_areas` stocks in temperate broadleaf forests, `r n_fluxes_matureTeN_with_ge7_areas` fluxes and `r n_stocks_matureTeN_with_ge7_areas` stocks in temperate conifer forests, and `r n_fluxes_matureBoN_with_ge7_areas` fluxes and `r n_stocks_matureBoN_with_ge7_areas` stocks in boreal forests.
-For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_youngTrB_with_ge7_areas` fluxes and `r n_stocks_youngTrB_with_ge7_areas` stocks in tropical broadleaf forests, `r n_fluxes_youngTeB_with_ge7_areas` fluxes and `r n_stocks_youngTeB_with_ge7_areas` stocks in temperate broadleaf forests, `r n_fluxes_youngTeN_with_ge7_areas` fluxes and `r n_stocks_youngTeN_with_ge7_areas` stocks in temperate conifer forests, and `r n_fluxes_youngBoN_with_ge7_areas` fluxes and `r n_stocks_youngBoN_with_ge7_areas` stocks in boreal forests.
+Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained mature forest estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests.
+For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.
 
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Tropical broadleaf MATURE.png")
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Tropical broadleaf MATURE.png"))
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate broadleaf MATURE.png")
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate broadleaf MATURE.png"))
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate conifer MATURE.png")
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate conifer MATURE.png"))
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Boreal conifer MATURE.png")
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Boreal conifer MATURE.png"))
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Tropical broadleaf YOUNG.png")
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Tropical broadleaf YOUNG.png"))
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate broadleaf YOUNG.png")
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate broadleaf YOUNG.png"))
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate conifer YOUNG.png")
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate conifer YOUNG.png"))
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Boreal conifer YOUNG.png")
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Boreal conifer YOUNG.png"))
 ```

### 2020-09-25 manuscript/ERL_global_C_review.Rmd
Update ERL_global_C_review.Rmd

@@
 n_focal_variable_records_eligible_for_inclusion <- sum(Biomes_n_records$n_records) # n records_eligible_for_inclusion, limited to focal variables
+# # records for excluded biome groupings ? 
+# n_records_BoB <- NA #Biomes_n_records$n_records[Biomes_n_records$Biome %in% "Boreal broadleaf"] # for all n, subset by Age if needed
+# n_records_Bomixed <- NA
+# n_records_Te_mixed <- NA
 

### 2020-09-26 manuscript/ERL_global_C_review.Rmd
knit in updated figures/stats
integrate BBL comments

@@
 output:
-  word_document:
+  pdf_document:
     fig_caption: yes
@@
 
-Ben Bond-Lamberty^4
+Ben Bond-Lamberty^4^
 
@@
 #READING IN TABLES TO GET STATS FOR TEXT
-path_to_ForC <- paste0(dirname(dirname(getwd())), "/ForC/") # this is to allow anybody to knit (assuming current working direcoty is in forc-db/ERL-review/manuscript and its "nesting" withing the ForC repo does not change)
+path_to_ForC <- paste0(dirname(dirname(getwd())), "/ForC/") # this is to allow anybody to knit (assuming current working directory is in forc-db/ERL-review/manuscript and its "nesting" withing the ForC repo does not change)
 
@@
 
+*NOTES TO COAUTHORS: *
+
+- We're still finalizing the data (and adding some data).
+Outliers in plots will all be checked/ resolved, and we'll be able to pull in more data (GROA needs classification by dominant vegetation before it can be pulled in, and we're working on that)
+- "**???**" indicates a reference that we have entered in the .Rmd file, but not yet in the bibliography file. Don't worry about those. (However, places with "REF" need references)
+- bold/ italic text indicates places where I know we need to go back and revisit/ fill in info
+
 \newpage
@@
 *Methods/Design.* Here, we draw from the Global Forest C Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation.
-Specifically, we draw upon `r n_records` records from `r n_areas_tot` geographic locations representing `r nrow(T1)` C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests.
+Specifically, we use `r n_records` ForC records from `r n_areas_tot` geographic locations representing `r nrow(T1)` C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests.
 For regrowth forests, we quantify age trends for all variables.
 
-*Review Results/ Synthesis.*  ForC v.XX yielded a fairly comprehensive picture of C cycling in the world's major forest biomes, with broad closure of C budgets.
+*Review Results/ Synthesis.*  ForC v.**XX** yielded a fairly comprehensive picture of C cycling in the world's major forest biomes, with broad closure of C budgets.
 The rate of C cycling generally increased from boreal to tropical regions in both mature and regrowth forests, whereas C stocks showed less directional variation.
@@
 
-*Discussion.* This analysis yields a comprehensive and broadly consistent picture of C cycling across the world's forests.
-*[Discussion section will interpret results, highlighting new and significant findings, and discuss implications.
-Tentative headings are “Stand level C cycling in forests globally”, “Age trends in C cycling”, and “Implications for climate change mitigation”.]*
+*Discussion.*  *NEED TO WRITE THIS!!!*
 
@@
 Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**).
-While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion was retained in ecosystems over recent decades.
+While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion has been retained in ecosystems over recent decades.
 The resulting CO~2~ sink averaged 3.0 ± 0.8 Gt C yr^{-1} from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**).
 Moreover, forests contain substantial carbon (C) stocks: an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C [biomass and soils; @bonan_forests_2008].
+Forests are also globally dominant as a source of soil respiration [@warner_spatial_2019]. 
 Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), reducing the net forest sink to ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**).
-The future of this sink is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change.
-Regrowth forests in particular will play an important role [@pugh_role_2019], as these represent a large (~#%) and growing proportion of Earth's forests.
-Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; Cavaleri et al 2015].
+The future of the current forest C is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change.
+Regrowing forests in particular will play an important role [@pugh_role_2019], as these represent a large (**~#%**) and growing proportion of Earth's forests [@mcdowell_pervasive_2020].
+Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thus central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; **Cavaleri et al 2015**].
 
-Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climate-carbon_2006; @krause_large_2018; REFS] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history.
+Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climate-carbon_2006; @krause_large_2018; @bonan_model_2019; @devittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history.
 These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published.
 While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [@li_mapping_2019; **REFS for biomass, biomass change, net CO2 flux**], measurement of most forest C stocks and fluxes require intensive on-the-ground data collection.
-Original studies typically cover only a small numbers of sites at a time, with rare exceptions spanning regions or continents [e.g., @lutz_global_2018; FLUXNET_REF), typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015] or FLUXNET [@baldocchi_fluxnet_2001].
+Original studies typically cover only a small numbers of sites at a time, with rare exceptions spanning regions or continents [e.g., @lutz_global_2018; **FLUXNET_REF**), typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015] or FLUXNET [@baldocchi_fluxnet_2001].
 The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*.
-In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere (REFS).
+In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere (**REFS**).
 
@@
 
-While components of these questions have been previously addressed [@luyssaert_co2_2007; @anderson-teixeira_carbon_2016; @cook-patton_mapping_2020; @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and thereby stands to serve as a foundation for improved understanding of global forest C cycling.
+While components of these questions have been previously addressed [@luyssaert_co2_2007; @anderson-teixeira_carbon_2016; @cook-patton_mapping_2020; @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and will serve as a foundation for improved understanding of global forest C cycling.
 
@@
 This review synthesizes data from the ForC database [Fig. 1; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018].
-ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies.
+ForC amalgamates numerous intermediary data sets (*e.g.*, **REFS**) and original studies.
 Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records.
@@
 As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible [83% of records in v.2.0; @anderson-teixeira_forc_2018].
-Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], yet these represented only approximately one-third of records in ForC v.2.0 [@anderson-teixeira_forc_2018].
-Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##, `r n_SRDB` records; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database [GROA v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020], and data from several publications [e.g., @taylor_temperature_2017, @lutz_global_2018, @johnson_climate_2018; detailed list at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv].
-We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS).
-The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: TBD).
+Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016].
+Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##, `r n_SRDB` records; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database [GROA v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020].
+We've also added data from individual publications (detailed list at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv), with a particular focus on productivity [@banbury_morgan_global_nodate; e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @lutz_global_2018, @johnson_climate_2018].
+We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: **REFS**).
+The database version used for this analysis has been tagged as a new release on Github (**XX**) and assigned a DOI through Zenodo (DOI: TBD).
 
@@
 For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables (Tables 1, S#(variable mapping)).
-Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum stem diameter sampled was $\le$ 10cm.
+Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum threshold for tree census was $\le$ 10cm. 
 All records were measured directly or derived from field measurements (as opposed to modeled).
@@
 Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication).
-Assigning stands to these groupings required the exclusion of records for which ForC lacked records of stand age (n=`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database), or geographic coordinates (n=`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database).
-We also excluded records of stand age = 0 year (n=`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
+Assigning stands to these groupings required the exclusion of records for which ForC lacked records of stand age (`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database), or geographic coordinates (`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database).
+We also excluded records of stand age = 0 year (`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
 In total, our analysis retained `r round(n_records/n_focal_variable_records_eligible_for_inclusion*100, digits=1)` of the focal variable records for forests of known age.
 
-Data were analyzed to produce
-*Schematics:*
 Data were summarized to produce schematics of C cycling across the eight biome - age group combinations identified above.
-For mature forests, to obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot.
+To obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot.
 Values were then averaged across plots clustered within 25 km of one another [*sensu* @anderson-teixeira_forc_2018], weighting by area sampled if available for all records.
@@
 
-*FIgs 6-7*
-Finally, for figures 6 and 7, the original values were analyzed via a linear mixed effects model ('lmer' function in 'lme4' R package) with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
+To test for differences across mature forest biomes, the original values were analyzed via a linear mixed effects model ['lmer' function in 'lme4' R package version **x.xx**; **REF**] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
 When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
-There were enough data to run this analysis for all focal variables but $DW_{down}$.
+There were enough data to run this analysis for all focal variables but **$DW_{down}$**.
 
-For young (<100yrs) forest types, we employed a mixed effects model with biome and log10[stand.age] as fixed effects and plot nested within geographic.area as a random effect on the intercept.
-When the effect of stand.age was significant at p $\le$ 0.05 and when each biome had records for stands of at least 10 different ages, a biome - stand.age interaction was included in the model.
+To test for age trends in young (<100yrs) forests, we employed a mixed effects model with biome and log10[stand.age] as fixed effects and plot nested within geographic.area as a random effect on the intercept.
+When the effect of stand.age was significant at p $\le$ 0.05 and when each biome had records for stands of at least 10 different ages, a biome * stand.age interaction was included in the model.
 
-All database manipulation, analyses, and figure production were fully automated in R (version, citation).
+All database manipulation, analyses, and figure production were fully automated in R (**version, citation**).
 Materials required to fully reproduce these analyses, including data, R scripts, and image files, are archived in Zenodo (DOI: TBD].
@@
 ```
-
-*NOTE TO COAUTHORS: THIS SECTION HAS A TON OF FIGURES, BUT VERY LITTLE TEXT. IS THERE ANYTHING THAT SHOULD BE ADDED?*
 
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 2 | C cycle diagram for mature tropical broadleaf forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.
 Arrows indicate fluxes, boxes indicate stocks.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrows are scaled as the **square root of flux divided by 5**.", out.width='100%', fig.pos='H'}
+Arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
 
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 3 | C cycle diagram for mature temperate broadleaf forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.
 Arrows indicate fluxes, boxes indicate stocks.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
+Arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
 
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 4 | C cycle diagram for mature temperate conifer forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.
 Arrows indicate fluxes, boxes indicate stocks.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
+Arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
 
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 5 | C cycle diagram for mature boreal conifer forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.
 Arrows indicate fluxes, boxes indicate stocks.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
+Arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
 
@@
 In this majority of cases, differences between tropical and boreal forests were always significant, with temperate forests intermediate and significantly different from one or both.
-Fluxes tended to be greater in temperate broadleaf than conifer forests, but the difference was never significant.
-This pattern held for the following variables: $GPP$, $NPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{foliage}$, $ANPP_{litterfall}$, $BNPP$, $BNPP_fine$, $R_{eco}$, $R_{root}$, $R_{soil}$, and $R_{het-soil}$.
+Fluxes tended to be numerically greater in temperate broadleaf than conifer forests, but the difference was never statistically significant.
+This pattern held for the following variables: ** $GPP$, $NPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{foliage}$, $ANPP_{litterfall}$, $BNPP$, $BNPP_fine$, $R_{eco}$, $R_{root}$, $R_{soil}$, and $R_{het-soil}$**.
 
@@
 There were fewer distinct trends in C stocks across biomes (Figs. 7, S16-S26).
-Specifically, there were significant differences for only *##* of *##* variables: *LIST*.
+Specifically, there were significant differences for only **##** of **##** variables: **LIST**.
 In all of these cases, temperate conifer forests were in the highest significance grouping, and boreal forests in the lowest.
@@
 Once again, the high values for the temperate conifer biome were driven by the very high-biomass forests of the US Pacific Northwest, which are disproportionately represented in the current version of ForC.
-Thus, biome differences should be interpreted more as driven more by geographic distribution of sampling than by true differences.)*
+Thus, biome differences should be interpreted more as driven more by geographic distribution of sampling than by true differences.
 
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | C cycle diagram for young tropical broadleaf forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
 Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
 Arrows indicate fluxes, boxes indicate stocks.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
+Arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
 
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 9 | C cycle diagram for young temperate broadleaf forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
 Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
 Arrows indicate fluxes, boxes indicate stocks.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
+Arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
 
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 10 | C cycle diagram for young temperate conifer forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
 Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
 Arrows indicate fluxes, boxes indicate stocks.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
+Arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
 
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 11 | C cycle diagram for young boreal conifer forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$.
 Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
 Arrows indicate fluxes, boxes indicate stocks.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.", out.width='100%', fig.pos='H'}
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
+Arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
 
@@
 The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Figs. 6-11).
-Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks and $NEP$ of mature forests are influenced by a different set of drivers.
+Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks and $NEP$ of mature forests are correlated with a different set of factors.
 
@@
 
-ForC provides fairly good coverage of most major variables.
+ForC provides provides unprecedented coverage of most major variables.
 *(discuss  how this improves upon previous data compilations/ for which variables does ForC make the greatest difference (e.g., not AGB or NEP/GPP/Reco, but by far the latest data compilation for dead wood, [variables])*
@@
 
-Closure of the C cycle budgets for mature forests (Figs. 2-5) is fairly good.
-However, SD's are often large, reflective of significant within-biome variation. This makes the standard for closure relatively loose.
+**notes from Ben on the above par:**
+Pregitzer and Euskirchen 2004 http://dx.doi.org/10.1111/j.1365-2486.2004.00866.x "Aggregated biome-level estimates of NPP and NEP were higher in intermediate-aged forests (e.g., 30–120 years), while older forests (e.g., 4120 years) were generally less productive.
+The mean NEP in the youngest forests (0–10 years) was negative (source to the atmosphere) in both boreal and temperate biomes…Forest age is a highly significant source of variability in NEP at the biome scale”
+
+Amiro et al. 2010 http://dx.doi.org/10.1029/2010JG001390
+Houghton et al. 2020 https://doi.org/10.1111/gcb.15050
+
+Turnover:
+Pugh et al. 2020 http://dx.doi.org/10.5194/bg-2019-491
+Yu et al. 2019 http://dx.doi.org/10.1073/pnas.1821387116
+
+Human footprint in global forests
+http://dx.doi.org/10.1038/nature05847
+http://dx.doi.org/10.1038/nature02619
+
+Mention consistency (or lack of) with e.g. GOLUM-CNP?
+http://dx.doi.org/10.5194/gmd-11-3903-2018
+
+The C cycle budgets for mature forests (Figs. 2-5) generally "close"--that is, the sums of component variables do not differ significantly from the larger fluxes.
+However, standard deviations are often large, reflective of significant within-biome variation. This makes the standard for closure relatively loose.
 Lack of closure, in the few instances where it occurs, is probably more reflective of differences in the representation of forest types (e.g., disproportionate representation of US Pacific NW for aboveground woody biomass relative to AGB; Fig.
@@
 
+Higher NEP in temperate forests – implications?
+Invariant NEP in older forests?
+This could be built out a bit; thinking of Luyssaert 2008 http://dx.doi.org/10.1038/nature07276 and following papers arguing about this.
+
 ### Age trends in C cycling
@@
 
-NEP increases with log(age) to 100 --> strongest C sinks are established secondary forests.
+NEP increases with log(age) to 100 --> strongest C sinks are established secondary forests. (But presumably this exact number is an artifact; don’t over-emphasize.)
 
@@
 
-The future of forest C cycling will shape trends in atmospheric CO~2~ and the course of climate change.
+The future of forest C cycling [@song_meta-analysis_2019] will shape trends in atmospheric CO~2~ and the course of climate change.
 For a human society seeking to understand and mitigate climate change, the data contained in ForC and summarized here can help to meet two major challenges.
@@
 First, improved representation of forest C cycling in models is essential to improving predictions of the future course of climate change.
-To ensure that models are giving the right answers for the right reasons, it is important benchmark against multiple components of the C cycle.
+To ensure that models are giving the right answers for the right reasons, it is important to benchmark against multiple components of the C cycle that are internally consistent with each other.
 By making tens of thousands of records readily available in standardized format, ForC makes it feasible for the modeling community to draw upon these data to benchmark models.
 Integration of ForC with models is a goal (Fer et al., in revision).
-On a more cursory level, the values summarized here can serve as a sanity check for modelers to determine whether model predictions for multiple C cycle variables are reasonable--i.e., within the range of previous observations.
 
@@
 
-Other variables cannot be remotely sensed.
+Other variables, in particular respiration fluxes, cannot be remotely sensed (@bond-Lamberty_estimating_2016), and efforts such as the Global Carbon Project (**le quere REF**) and NASA CMS (**citation: https://carbon.nasa.gov/pdfs/CMS_Phase-1_Report_Final_optimized.pdf but maybe better to cite open literature, one of the papers listed at https://cmsflux.jpl.nasa.gov/get-data/publication-data-sets**) typically compute them as residuals only.
+
 In terms of C stocks, there is a paucity of data on dead wood and organic layer (Pan et al. ?).
 These can be significant.
+(**Note that we've given a lot of emphasis to dead wood (work by Abby, and also Jenny), and as a result this work really advances knowledge of dead wood.
+We'll want to highlight that here.**)
 *(give some stats/ cite figures)*.
@@
 
+**Move to data availability statement, or methods?**:
 We recommend that use of ForC data go to the original database, as opposed to using "off-the-shelf" values from this publication.

### 2020-09-26 manuscript/ERL_global_C_review.Rmd
work on discussion

@@
 output:
-  pdf_document:
+  word_document:
     fig_caption: yes
@@
 All database manipulation, analyses, and figure production were fully automated in R (**version, citation**).
-Materials required to fully reproduce these analyses, including data, R scripts, and image files, are archived in Zenodo (DOI: TBD].
-Data, scripts, and results presented here are also available through the open-access ForC GitHub repository (https://github.com/forc-db/ForC), where many will be updated as the database develops.
 
@@
 
-ForC v.XX yielded a fairly comprehensive and internally consistent picture of C cycling in the world's major forest biomes.
+ForC v.XX brought together an unprecedented amount of data to yield an internally consistent picture of C cycling in the world's major forest biomes.
 Carbon cycling rates generally increased from boreal to tropical regions and with stand age.
@@
 http://dx.doi.org/10.5194/gmd-11-3903-2018
+
+In terms of C stocks, there is a paucity of data on dead wood and organic layer (Pan et al. ?).
+These can be significant.
+(**Note that we've given a lot of emphasis to dead wood (work by Abby, and also Jenny), and as a result this work really advances knowledge of dead wood.
+We'll want to highlight that here.**)
+*(give some stats/ cite figures)*.
+ForC does not include soil carbon, which is covered by other efforts (REFS).
+For fluxes, Fluxnet is the keeper of the best data on NEE, GPP, Reco (REFS), and SRDB remains the authority on soil respiration (REFS).
+ForC includes recent data from both, but is not continuously integrated.
+For C is the best source for most of the subsidiary fluxes: NPP, woody mortality...
 
@@
 
-There remain numerous data needs for improved accounting of forest carbon stocks and fluxes in GHG accounting.
+**ForC can complement remote sensing to provide a comprehensive picture of global forest C cycling.** (*here, discuss what is well-covered by remote sesnsing, where ForC is useful for calibrating remote sensing, and what ForC can get that remote sensing can't.*)
 AGB is the largest stock, and most of the emphasis is on this variable.
@@
 
- *remote sensing measurements are increasingly useful for global- or regional-scale estimates of forest $GPP$ (Bagdley et al.
-2019, [@li_mapping_2019]), aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) [@clark_quantifying_2004; @leitold_nino_2018], and to some extent net ecosystem exchange ($NEP$) (REFS),*
-
+*Remote sensing measurements are increasingly useful for global- or regional-scale estimates of forest $GPP$ (Bagdley et al.
+2019, [@li_mapping_2019]), aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) [@clark_quantifying_2004; @leitold_nino_2018], and to some extent net ecosystem exchange ($NEP$) (REFS).*
 Other variables, in particular respiration fluxes, cannot be remotely sensed (@bond-Lamberty_estimating_2016), and efforts such as the Global Carbon Project (**le quere REF**) and NASA CMS (**citation: https://carbon.nasa.gov/pdfs/CMS_Phase-1_Report_Final_optimized.pdf but maybe better to cite open literature, one of the papers listed at https://cmsflux.jpl.nasa.gov/get-data/publication-data-sets**) typically compute them as residuals only.
-
-In terms of C stocks, there is a paucity of data on dead wood and organic layer (Pan et al. ?).
-These can be significant.
-(**Note that we've given a lot of emphasis to dead wood (work by Abby, and also Jenny), and as a result this work really advances knowledge of dead wood.
-We'll want to highlight that here.**)
-*(give some stats/ cite figures)*.
-ForC does not include soil carbon, which is covered by other efforts (REFS).
-For fluxes, Fluxnet is the keeper of the best data on NEE, GPP, Reco (REFS), and SRDB remains the authority on soil respiration (REFS).
-ForC includes recent data from both, but is not continuously integrated.
-For C is the best source for most of the subsidiary fluxes: NPP, woody mortality...
+(**Ben, it woudl be particularly helpful if you could flesh this out some more.**)
 
@@
 
+## Data availability statement
+Materials required to fully reproduce these analyses, including data, R scripts, and image files, are archived in Zenodo (DOI: TBD].
+Data, scripts, and results presented here are also available through the open-access ForC GitHub repository (https://github.com/forc-db/ForC), where many will be updated as the database develops.
+
+## ORCID iD
+
 ## References

### 2020-09-26 manuscript/ERL_global_C_review.Rmd
knit in new table
closes #38

@@
 When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
-There were enough data to run this analysis for all focal variables but **$DW_{down}$**.
+There were enough data to run this analysis for all focal variables but **$R_{auto-ag}$ and $DW_{down}$**.
 

### 2020-09-26 manuscript/ERL_global_C_SI.Rmd
insert SI_age_trend_model_summaries.csv table
still need to finish formatting

@@
 [this table](https://github.com/forc-db/ForC/blob/master/figures/C_cycle_diagrams/ForC_variables_mapping_for_C_cycle_diagrams.csv))
+
+\newpage
+## Table S2. Model parameter estimates for age trends and biome differences in young forests
+
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+table_pt <- read.csv("tables_figures/SI_age_trend_model_summaries.csv", 
+                     #colClasses=c("NULL", NA, NA, NA, NA), 
+                     #col.names=c("NULL", "Parameter", "Estimate", "SE", "$t_{value}$"), 
+                     stringsAsFactors = FALSE, check.names = FALSE)
+kable(table_pt, format = "latex", booktabs = TRUE, longtable = TRUE,  escape = F, linesep = c('\\addlinespace'))%>%
+  pack_rows("NEP", 1, 5, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
+  pack_rows("GPP", 6, 10, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
+  pack_rows("NPP", 11, 15, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
+  pack_rows("ANPP", 16, 23, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
+  pack_rows("ANPP_woody", 24, 28, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
+  pack_rows("ANPP_stem", 29, 36, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
+  pack_rows("ANPP_branch", 37, 40, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
+  kable_styling(latex_options =c("repeat_header"), protect_latex = T, font_size = 7)
+```
 

### 2020-09-27 manuscript/ERL_global_C_SI.Rmd
add/ cite Table S1, misc cleanup

@@
 
-## Table S#(variable mapping))
-
-[this table](https://github.com/forc-db/ForC/blob/master/figures/C_cycle_diagrams/ForC_variables_mapping_for_C_cycle_diagrams.csv))
+## Table S1. Numbers of records by biome and age class
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+table <- read.csv("tables_figures/Biomes_n_records_variables_of_interest.csv",stringsAsFactors = FALSE, check.names = FALSE)
+table_noAgeNA <- subset(table, Age != 'NA') # remove records with age unknown
+table_wide <- reshape(table_noAgeNA, idvar = "Biome", timevar = "Age", direction = "wide")
+colnames(table_wide)[2] <- "Mature"
+colnames(table_wide)[3] <- "Young"
+kable(table_wide, format = "latex", booktabs = TRUE, escape = F)%>%
+  add_header_above(c(" ", "n records" = 2)) %>%
+  pack_rows("Boreal climate zones", 1, 3, latex_gap_space = "1em", colnum = 3, hline_before = FALSE) %>%
+  pack_rows("Excluded climate zones", 4, 6, latex_gap_space = "1em", colnum = 3, hline_before = FALSE) %>%
+  pack_rows("Temperate climate zones", 7, 9, latex_gap_space = "1em", colnum = 3, hline_before = FALSE) %>%
+  pack_rows("Tropical climate zones", 10, 12, latex_gap_space = "1em", colnum = 3, hline_before = FALSE) %>%
+  row_spec(2, bold = T) %>%
+  row_spec(7:8, bold = T) %>%
+  row_spec(10, bold = T) %>%
+  kable_styling( protect_latex = T, font_size = 10)
+```
+Sample sizes refer to data set after merging of duplicates, removal of stands with no age or history data, and removal of managed and disturbed stands.
+For vegetation type, "Other" refers to stands that are mixed broadleaf/ conifer or that have not been classified.
+Focal biomes are indicated in bold.
 

### 2020-09-27 manuscript/ERL_global_C_review.Rmd
add/ cite Table S1, misc cleanup

--- 2020-09-26 manuscript/ERL_global_C_review.Rmd
@@
 # n_records_BoB <- NA #Biomes_n_records$n_records[Biomes_n_records$Biome %in% "Boreal broadleaf"] # for all n, subset by Age if needed
-# n_records_Bomixed <- NA
-# n_records_Te_mixed <- NA
 
@@
 n_stocks_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_stocks])
-
 ```
@@
 
-*Review Results/ Synthesis.*  ForC v.**XX** yielded a fairly comprehensive picture of C cycling in the world's major forest biomes, with broad closure of C budgets.
+*Review Results/ Synthesis.*  ForC v3.0 yielded a fairly comprehensive picture of C cycling in the world's major forest biomes, with broad closure of C budgets.
 The rate of C cycling generally increased from boreal to tropical regions in both mature and regrowth forests, whereas C stocks showed less directional variation.
@@
 
-*Discussion.*  *NEED TO WRITE THIS!!!*
+*Discussion.*  **NEED TO WRITE THIS!!!**
 
 *Key words*: forest ecosystems; carbon cycle; stand age; productivity; respiration; biomass; global
+
+\newpage
 
@@
 Records for these variables represented `r round(n_focal_variable_records_eligible_for_inclusion/records_eligible_for_inclusion*100,digits=1)`% of the total records eligible for inclusion.
-For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables (Tables 1, S#(variable mapping)).
+For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables.
+Specifically, net ecosystem exchange (measured by eddy-covariance; **FLUXNET REF**) and biometric estimates of $NEP$ were combined into the single variable $NEP$ (Table 1).
+Furthermore, for $NPP$, $ANPP$, and $ANPP_{litterfall}$, ForC variables specifying inclusion of different components were combined.
 Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum threshold for tree census was $\le$ 10cm. 
@@
 In total, our analysis retained `r round(n_records/n_focal_variable_records_eligible_for_inclusion*100, digits=1)` of the focal variable records for forests of known age.
+Numbers of records by biome and age class are given in Table S1.
 
@@
 
-Of the `r nrow(ForC_measurements)` records in ForC v.#, `r n_records` met the criteria for inclusion in this study (Fig. 1).
+Of the `r nrow(ForC_measurements)` records in ForC v3.0, `r n_records` met the criteria for inclusion in this study (Fig. 1).
 These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas.
@@
 Average C cycles for forests <100 years old are presented in Figures 8-11.  
-Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Table 1, Figs. 6-11, S1-S26; detailed below).
+Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Tables 1, S2, Figs. 6-11, S1-S26; detailed below).
 
@@
 
-There were only *##* flux variables with sufficient data to test for biome x age interactions: $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{litterfall}$, and $BNPP$.
+There were only *##* flux variables with sufficient data to test for biome x age interactions: $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{litterfall}$, and $BNPP$ (Table S2).
 **(more could be added if age trends become significant after outliers are resolved)**
@@
 
-ForC v.XX brought together an unprecedented amount of data to yield an internally consistent picture of C cycling in the world's major forest biomes.
+ForC v3.0 brought together an unprecedented amount of data to yield an internally consistent picture of C cycling in the world's major forest biomes.
 Carbon cycling rates generally increased from boreal to tropical regions and with stand age.
@@
 ## ORCID iD
+Kristina J. Anderson-Teixeira: https://orcid.org/0000-0001-8461-9713
 

### 2020-09-27 manuscript/ERL_global_C_review.Rmd
work on results

@@
 When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
-There were enough data to run this analysis for all focal variables but **$R_{auto-ag}$ and $DW_{down}$**.
+*This analysis was run for variables with records for $\ge$7 distinct geographic areas in each biome, excluding any biomes that failed this criteria and dropping variable where less than two biomes met the criteria (Table 1).* **([issue 40](https://github.com/forc-db/ERL-review/issues/40)])**
 
@@
 For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.
+**([issue 29](https://github.com/forc-db/ERL-review/issues/29#issuecomment-699633801))**
 
@@
 
-Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5 (and available in tabular format in the ForC release accompanying this publication).
+Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5 (and available in tabular format in the ForC release accompanying this publication:  `ForC/numbers_and_facts/ForC_variable_averages_per_Biome.csv`).
 
 For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets were generally consistent.
-That is, component fluxes and stocks summed to within one standard deviation of more inclusive fluxes in all but two instances, both for living C stocks in temperate conifer forests (Fig.
-5).
+*That is, component fluxes and stocks summed to within one standard deviation of more inclusive fluxes in all but* **two** *instances, both in temperate conifer forests (Fig. 5).* 
 **(check this with final results: https://github.com/forc-db/ForC/blob/master/numbers_and_facts/C_cycle_closure.csv) ** 
-Within this biome, $B_{ag}<B_{ag-wood}+B_{foliage}$ and $B_{root}<B_{root-coarse}+B_{root-fine}$ because $B_{ag-wood}$ and $B_{root-coarse}$ were very high, with strongly disproportionate numbers of records from the high-biomass forests of the US Pacific Northwest (Figs.
-S18, S21).
+*Within this biome, $B_{ag}<B_{ag-wood}+B_{foliage}$ and $B_{root}<B_{root-coarse}+B_{root-fine}$ because $B_{ag-wood}$ and $B_{root-coarse}$ were very high, with strongly disproportionate numbers of records from the high-biomass forests of the US Pacific Northwest (Figs.
+S18, S21).*
 
@@
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
+Arrow size is proportional to the square root of corresponding flux.
+Asterisk after variable name indicates lack of C cycle closure.", out.width='100%', fig.pos='H'}
 
@@
 
-The C cycle budgets for mature forests (Figs. 2-5) generally "close"--that is, the sums of component variables do not differ significantly from the larger fluxes.
+The C cycle budgets for mature forests (Figs. 2-5) generally "close"--that is, the sums of component variables do not differ from the larger fluxes by more than one standard deviation.
 However, standard deviations are often large, reflective of significant within-biome variation. This makes the standard for closure relatively loose.
-Lack of closure, in the few instances where it occurs, is probably more reflective of differences in the representation of forest types (e.g., disproportionate representation of US Pacific NW for aboveground woody biomass relative to AGB; Fig.
-4) than of methodological accuracy.
+*Lack of closure, in the few instances where it occurs, is probably more reflective of differences in the representation of forest types (e.g., disproportionate representation of US Pacific NW for aboveground woody biomass relative to AGB; Fig.
+4) than of methodological accuracy.*
 Thus, overall, a high degree of closure implies that ForC gives a consistent picture of C cycling within biomes.

### 2020-09-27 manuscript/ERL_global_C_review.Rmd
worked on results

@@
 When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
-*This analysis was run for variables with records for $\ge$7 distinct geographic areas in each biome, excluding any biomes that failed this criteria and dropping variable where less than two biomes met the criteria (Table 1).* **([issue 40](https://github.com/forc-db/ERL-review/issues/40)])**
+*This analysis was run for variables with records for $\ge$ 7 distinct geographic areas in each biome, excluding any biomes that failed this criteria and dropping variable where less than two biomes met the criteria (Table 1).*
+**([issue 40](https://github.com/forc-db/ERL-review/issues/40)])**
 
@@
 
-**(The following covers only the variables with age trends figures; there maybe a few more without much age data)**
-With only ## exceptions, C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs. 6, S1-S15).
-In this majority of cases, differences between tropical and boreal forests were always significant, with temperate forests intermediate and significantly different from one or both.
+**check this with final data (based on Table 1):**
+There were sufficient data to assess mature forest biome differences for **20** flux variables, and significant differences among biomes were detected for **14** variables (Table 1).
+With only **one** exception ($R_{auto}$, which had low sample size; **[issue 40](https://github.com/forc-db/ERL-review/issues/40)**), C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs.
+6, S1-S15).
+Differences between tropical and boreal forests were always significant, with temperate forests intermediate and significantly different from one or both.
 Fluxes tended to be numerically greater in temperate broadleaf than conifer forests, but the difference was never statistically significant.
-This pattern held for the following variables: ** $GPP$, $NPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{foliage}$, $ANPP_{litterfall}$, $BNPP$, $BNPP_fine$, $R_{eco}$, $R_{root}$, $R_{soil}$, and $R_{het-soil}$**.
+This pattern held for the following variables: ** $GPP$, $NPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$,$ANPP_{branch}$, $ANPP_{foliage}$, $ANPP_{litterfall}$, $M_{woody}$, $BNPP$, $R_{eco}$, $R_{soil}$, and $R_{het-soil}$**.
+For variables without significant differences among biomes, the same general trends applied.
 
@@
 
-There were fewer distinct trends in C stocks across biomes (Figs. 7, S16-S26).
-Specifically, there were significant differences for only **##** of **##** variables: **LIST**.
-In all of these cases, temperate conifer forests were in the highest significance grouping, and boreal forests in the lowest.
-Tropical and temperate broadleaf forests fell in between--most commonly being significantly different from temperate conifer but not boreal forests.
+**check this with final data (based on Table 1):**
+There were sufficient data to assess mature forest biome differences for **9** stock variables, and significant differences among biomes were detected for **6** variables (**$B_{tot}$, $B_{ag}$, $B_{ag-wood}$, $B_{foliage}$, $B_{root-coarse}$, $DW_{tot}$**; Table 1).
+C stocks had less consistent patterns across biomes (Figs. 7, S16-S26).
+In **four of the six** cases (**$B_{tot}$, $B_{ag-wood}$,  $B_{root-coarse}$, $DW_{tot}$**), temperate conifer forests were in the highest significance grouping, and boreal forests in the lowest, with tropical and temperate broadleaf forests  in between--most commonly being significantly different from temperate conifer but not boreal forests.
+For $B_{ag}$, which had by far the highest sample size, tropical forests exceeded temperate conifer forests (but not significantly). 
+For $B_{foliage}$,temperate broadleaf forests were lowest (again, not significantly).
 Once again, the high values for the temperate conifer biome were driven by the very high-biomass forests of the US Pacific Northwest, which are disproportionately represented in the current version of ForC.

### 2020-09-27 manuscript/ERL_global_C_review.Rmd
Update ERL_global_C_review.Rmd

@@
 For $B_{foliage}$,temperate broadleaf forests were lowest (again, not significantly).
-Once again, the high values for the temperate conifer biome were driven by the very high-biomass forests of the US Pacific Northwest, which are disproportionately represented in the current version of ForC.
+The high values for the temperate conifer biome were driven by the very high-biomass forests of the US Pacific Northwest, which are disproportionately represented in the current version of ForC.
 Thus, biome differences should be interpreted more as driven more by geographic distribution of sampling than by true differences.

### 2020-09-27 manuscript/ERL_global_C_SI.Rmd
incorporated new changes
#32
#34
#40

@@
 kable(table_pt, format = "latex", booktabs = TRUE, longtable = TRUE,  escape = F, linesep = c('\\addlinespace'))%>%
-  pack_rows("NEP", 1, 5, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
-  pack_rows("GPP", 6, 10, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
-  pack_rows("NPP", 11, 15, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
-  pack_rows("ANPP", 16, 23, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
-  pack_rows("ANPP_woody", 24, 28, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
-  pack_rows("ANPP_stem", 29, 36, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
-  pack_rows("ANPP_branch", 37, 40, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
+  #pack_rows("NEP", 1, 5, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
+  #pack_rows("GPP", 6, 10, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
+  #pack_rows("NPP", 11, 15, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
+  #pack_rows("ANPP", 16, 23, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
+  #pack_rows("ANPP_woody", 24, 28, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
+  #pack_rows("ANPP_stem", 29, 36, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
+  #pack_rows("ANPP_branch", 37, 40, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
   kable_styling(latex_options =c("repeat_header"), protect_latex = T, font_size = 7)

### 2020-09-27 manuscript/ERL_global_C_review.Rmd
incorporated new changes
#32
#34
#40

@@
 Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
+For young tropical forests imported from GROA but not yet classified by leaf type, we assumed that all were broadleaf, consistent with the rarity of naturally regenerating needleleaf forests in the tropics (**REF**).
 Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication).

### 2020-09-28 manuscript/ERL_global_C_review.Rmd
fix fluxes vs stocks for Boreal conifer YOUNG

@@
 Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained mature forest estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests.
-For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.
+For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.
 **([issue 29](https://github.com/forc-db/ERL-review/issues/29#issuecomment-699633801))**

### 2020-09-28 manuscript/ERL_global_C_SI.Rmd
#34

@@
 
+ForC v3.0 contains potential duplicates that have arisen from importing intermediary data sets with overlap in data. 
+In particular, the recent imports of SRDB and GROA resulted in the introduction of numerous potential duplicates. 
+SRDB and ForC were developed independently (but with different foci), resulting in numerous duplicates.
+The GROA database incorporated ForC data. 
+For records retaining the same site names, GROA duplicates were not imported into ForC (although they were checked against corresponding ForC records, and differences reconciled).
+There were, however, some records entered into GROA independently of ForC, resulting in mismatched names.
+
+Prior to the merging of the three databases (ForC v2.0, SRDB, and GROA), there were potential duplicates within the ForC and SRDB portions, the latter an artifact of how data were recorded in SRDB and imported into ForC.
+GROA has no known duplicates, with several steps having been taken to ensure independence of the records within the database.
+
+### Detecting and reconciling duplicate sites
+
+To the extent possible, we used automated scripts in R to detect duplicate records.
+Potential duplicates were defined as geographically proximate records for stands of **similar age** with the same variable measured in the same year (if known).
+In cases where site name, plot name, and stand age matched, duplicate status could be automatically detected and reconciled (see below).
+In cases where site and plot names or reported age differed, our script detected potential duplicate sites that were **geographically proximate**.
+When all such potential duplicate sites were within GROA, they were assumed independent.
+
+Suspected duplicates require manual review and merging of records, and this process has not been completed for all records. 
+To handle potential duplicates at sites that have yet to be reviewed, we implemented the following procedure. 
+In cases where a single location -- generally an established research site where multiple investigators have worked -- contained multiple plots in nested or unknown relation to one another, we grouped multiple sites into a "supersite" (e.g., Harvard Forest, Barro Colorado Island, Pasoh Forest Reserve), and duplicates within a supersite were handled in the same way as records with matching site and plot names.
+(**VALENTINE, IS THIS ACCURATE?**)
+For suspected duplicate groups that were not flagged as supersites and had not yet been reviewed, we retained only one potential duplicate record, assigning precedence as follows: (1) original GROA record(s), (2) record(s) in ForC prior to SRDB and GROA import, (3) SRDB record(s).
+
+Finally, ....
+https://github.com/forc-db/ERL-review/issues/34
+
 ### Generating ForC_simplified
+Detected duplicates were reconciled as follows.
 Replicate measurements (*i.e.*, replicates from within a single study) were averaged. 
@@
 For each group of duplicate records—*i.e.*, measurements of the same variable in the same plot at the same time—one record was assigned precedence (recorded in D.precedence field). 
+When duplicates were exact matches, **one was dropped**. 
 When measurement periods overlapped or were not specified, precedence was given first to records representing longer measurement periods (*i.e.*, end.date - start.date) and then to more recently published values.

### 2020-09-28 manuscript/ERL_global_C_review.Rmd
knit in new figs (now drawn from ForC), changed captions
#40
also, deleted unnecessary figs in ERL repo

@@
 When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
-*This analysis was run for variables with records for $\ge$ 7 distinct geographic areas in each biome, excluding any biomes that failed this criteria and dropping variable where less than two biomes met the criteria (Table 1).*
-**([issue 40](https://github.com/forc-db/ERL-review/issues/40)])**
+This analysis was run for variables with records for $\ge$7 distinct geographic areas in more than one biome, excluding any biomes that failed this criteria (Table 1).
 
 To test for age trends in young (<100yrs) forests, we employed a mixed effects model with biome and log10[stand.age] as fixed effects and plot nested within geographic.area as a random effect on the intercept.
-When the effect of stand.age was significant at p $\le$ 0.05 and when each biome had records for stands of at least 10 different ages, a biome * stand.age interaction was included in the model.
-
+This analysis was run for variables with records for $\ge$3 distinct geographic areas in more than one biome, excluding any biomes that failed this criteria (Table 1). 
+When the effect of stand age was significant at p $\le$ 0.05 and when each biome had records for stands of at least 10 different ages, a biome * stand.age interaction was included in the model.
 All database manipulation, analyses, and figure production were fully automated in R (**version, citation**).
@@
 Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.
+Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).
 Individual figures for each flux with sufficient data given in the Supplement (Figs.
@@
 
-  knitr::include_graphics("tables_figures/Flux_age_trends.png")
+  knitr::include_graphics(paste0(path_to_ForC, "figures/age_trends/for_ERL_review/Flux_age_trends.png"))
 ```
@@
 Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.
+Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).
 Individual figures for each stock with sufficient data given in the Supplement (Figs.
@@
 
-  knitr::include_graphics("tables_figures/Stock_age_trends.png")
+  knitr::include_graphics(paste0(path_to_ForC, "figures/age_trends/for_ERL_review/Stock_age_trends.png"))
 ```

### 2020-09-28 manuscript/ERL_global_C_review.tex
knit in new figs (now drawn from ForC), changed captions
#40
also, deleted unnecessary figs in ERL repo

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{float} \usepackage{caption} \captionsetup[table]{font=footnotesize} \captionsetup[figure]{font=footnotesize} \captionsetup[figure]{labelformat=empty} \captionsetup[table]{labelformat=empty} \usepackage{pdflscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}}
\usepackage{setspace}\onehalfspacing \usepackage{lineno}\linenumbers
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\raggedright

\textbf{Title:} Carbon cycling in mature and regrowth forests globally:
a macroecological synthesis based on the global Forest Carbon (ForC)
database

\textbf{Authors:}

Kristina J. Anderson-Teixeira\textsuperscript{1,2}*

Valentine Herrmann\textsuperscript{1}

Becky Banbury Morgan\textsuperscript{3}

Ben Bond-Lamberty\textsuperscript{4}

Susan C. Cook-Patton\textsuperscript{5}

Abigail E. Ferson\textsuperscript{1,6}

Jennifer C. McGarvey\textsuperscript{1}

Helene C. Muller-Landau\textsuperscript{1}

Maria M. H. Wang\textsuperscript{1,7}

\textbf{Author Affiliations:}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Conservation Ecology Center; Smithsonian Conservation Biology
  Institute; National Zoological Park, Front Royal, VA 22630, USA
\item
  Center for Tropical Forest Science-Forest Global Earth Observatory;
  Smithsonian Tropical Research Institute; Panama, Republic of Panama
\item
  School of Geography, University of Leeds, Leeds, UK
\item
  Joint Global Change Research Institute, Pacific Northwest National
  Laboratory, College Park Maryland 20740, USA
\item
  The Nature Conservancy; Arlington VA 22203, USA
\item
  College of Natural Resources, University of Idaho; Moscow, Idaho
  83843, USA
\item
  Grantham Centre for Sustainable Futures and Department of Animal and
  Plant Sciences, University of Sheffield, Western Bank, Sheffield,
  South Yorkshire S10 2TN, UK
\end{enumerate}

*corresponding author:
\href{mailto:teixeirak@si.edu}{\nolinkurl{teixeirak@si.edu}}; +1 540 635
6546

\begin{verbatim}
## [1] 0
\end{verbatim}

\emph{NOTES TO COAUTHORS: }

\begin{itemize}
\tightlist
\item
  We're still finalizing the data (and adding some data). Outliers in
  plots will all be checked/ resolved, and we'll be able to pull in more
  data (GROA needs classification by dominant vegetation before it can
  be pulled in, and we're working on that)
\item
  ``\textbf{???}'' indicates a reference that we have entered in the
  .Rmd file, but not yet in the bibliography file. Don't worry about
  those. (However, places with ``REF'' need references)
\item
  bold/ italic text indicates places where I know we need to go back and
  revisit/ fill in info
\end{itemize}

\newpage

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

\emph{Background.} The fate of Earth's climate is closely linked to
forests, which strongly influence atmospheric carbon dioxide
(CO\textsubscript{2}) and climate through their influential role in the
global carbon (C) cycle. Synthetic understanding of global forest C
cycles is needed to constrain model estimates of forest feedbacks to
climate change and to more accurately quantify the influence of land use
decisions on climate.

\emph{Methods/Design.} Here, we draw from the Global Forest C Database,
ForC, to provide a macroscopic overview of C cycling in the world's
forests, giving special attention to stand age-related variation.
Specifically, we use 14847 ForC records from 874 geographic locations
representing 34 C cycle variables to characterize ensemble C budgets for
four broad forest types (tropical broadleaf evergreen, temperate
broadleaf, temperate conifer, and taiga), including estimates for both
mature and regrowth (age \textless100 years) forests. For regrowth
forests, we quantify age trends for all variables.

\emph{Review Results/ Synthesis.} ForC v3.0 yielded a fairly
comprehensive picture of C cycling in the world's major forest biomes,
with broad closure of C budgets. The rate of C cycling generally
increased from boreal to tropical regions in both mature and regrowth
forests, whereas C stocks showed less directional variation. The
majority of flux variables, together with most live biomass pools,
increased significantly with stand age, \emph{and the rate of increase
again tended to increase from boreal to tropical regions}.

\emph{Discussion.} \textbf{NEED TO WRITE THIS!!!}

\emph{Key words}: forest ecosystems; carbon cycle; stand age;
productivity; respiration; biomass; global

\newpage

\hypertarget{background}{%
\subsection{Background}\label{background}}

\emph{(Abby has offered to update stats in this paragraph:)}

Forest ecosystems will play a critical role in shaping the course of
climate change (IPCC1.5) through their influence on atmospheric carbon
dioxide (CO\textsubscript{2}). Their annual gross CO\textsubscript{2}
sequestration (gross primary productivity, \(GPP\)) is estimated at
\textgreater69 Gt C yr\textsuperscript{-1} ({\textbf{???}}), or
\textgreater7 times average annual fossil fuel emissions from 2007-2016
(9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (\textbf{update}). While most
of this enormous C flux is couterbalanced by CO\textsubscript{2}
releases to the atmosphere through ecosystem respiration (\(R_{eco}\))
or fire, a small portion has been retained in ecosystems over recent
decades. The resulting CO\textsubscript{2} sink averaged 3.0 ± 0.8 Gt C
yr\^{}\{-1\} from 2007-2016, offsetting 32\% of anthropogenic fossil
fuel emissions (Le Quéré et al 2017) (\textbf{update, give range}).
Moreover, forests contain substantial carbon (C) stocks: an estimated
92\% of terrestrial biomass (Pan et al 2013) and 45\% of terrestrial C
(biomass and soils; Bonan 2008). Forests are also globally dominant as a
source of soil respiration ({\textbf{???}}). Globally, net deforestation
(\emph{i.e.}, gross deforestation - regrowth) has been a source of
CO\textsubscript{2} emissions, estimated at \textasciitilde1.1 Gt C yr-1
from YEAR-YEAR (Pan et al 2011), reducing the net forest sink to
\textasciitilde1.2-1.7 Gt C yr\textsuperscript{-1} across Earth's
forests (Le Quéré et al 2017, Schimel \emph{et al} 2015)
(\textbf{update, give range}). The future of the current forest C is
dependent both upon forest responses to a broad suite of global change
drivers and to future land use decisions, and will strongly influence
the course of climate change. Regrowing forests in particular will play
an important role (Pugh \emph{et al} 2019), as these represent a large
(\textbf{\textasciitilde\#\%}) and growing proportion of Earth's forests
(McDowell \emph{et al} 2020). Understanding, modeling, and managing
forest-atmosphere CO\textsubscript{2} exchange is thus central to
efforts to mitigate climate change {[}Grassi \emph{et al} (2017);
Griscom \emph{et al} (2017); \textbf{Cavaleri et al 2015}{]}.

Despite the centrality of forest C cycling in regulating atmospheric
CO\textsubscript{2}, important uncertainties in climate models
({\textbf{???}}, {\textbf{???}}, {\textbf{???}}, Krause \emph{et al}
2018) and CO\textsubscript{2} accounting frameworks (Pan \emph{et al}
2011) can be traced to lack of accessible, comprehensive data on how C
cycling varies across forest types and in relation to stand history.
These require large-scale databases with global coverage, which runs
contrary to the nature in which forest C stocks and fluxes are measured
and published. While remote sensing measurements are increasingly useful
for global- or regional-scale estimates of a few critical variables
{[}Li and Xiao (2019); \textbf{REFS for biomass, biomass change, net CO2
flux}{]}, measurement of most forest C stocks and fluxes require
intensive on-the-ground data collection. Original studies typically
cover only a small numbers of sites at a time, with rare exceptions
spanning regions or continents {[}e.g., Lutz \emph{et al} (2018);
\textbf{FLUXNET\_REF}), typically coordinated through research networks
such as ForestGEO (Anderson-Teixeira \emph{et al} 2015) or FLUXNET
(Baldocchi \emph{et al} 2001). The result of decades of research on
forest C cycling is that tens of thousands of records have been
distributed across literally thousands of scientific articles --often
behind paywalls-- along with variation in data formats, units,
measurement methods, \emph{etc.}. In this format, the data are
effectively inaccessible for many global-scale analyses, including those
attempting to benchmark model performance with global data (Clark et al
2017, Luo et al 2012), quantify the the role of forests in the global C
cycle (\emph{e.g.}, Pan \emph{et al} 2011), or use book-keeping methods
to quantify actual or scenario-based exchanges of CO\textsubscript{2}
between forests and the atmosphere (\textbf{REFS}).

To address the need for global-scale analyses of forest C cycling, we
recently developed an open-access Global Forest Carbon database, ForC
(Anderson-Teixeira \emph{et al} (2016), Anderson-Teixeira \emph{et al}
(2018)). ForC contains published estimates of forest ecosystem C stocks
and annual fluxes (\textgreater50 variables) based on ground-based
measurements, along with associated data required for interpretation
(\emph{e.g.}, stand history, measurement methods). These data have been
amalgamated from original peer-reviewed publications, either directly or
via intermediary data compilations. Since the its most recent
publication (Anderson-Teixeira \emph{et al} 2018), ForC has been
integrated with two large databases: the Global Soil Respiration
Database (SRDB; Bond-Lamberty and Thomson 2010) and the Global
Reforestation Opportunity Assessment database (GROA; Cook-Patton
\emph{et al} 2020), both of which have also synthesized published forest
C data. Following these additions, ForC currently contains 47846 records
from 10609 plots and 1532 distinct geographic areas representing all
forested biogeographic and climate zones.

Here, we synthesize ForC data (Fig. 1) to provide a macroscopic overview
of stand-level carbon cycling of the world's major forest biomes and how
it varies with stand age. Our primary goal is to provide a data-based
summary of our current state of knowledge on broad trends in forest C
cycling. We address three broad questions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  To what extent can we fully represent, and ``close'', C budgets for
  each of the world's major forest biomes (\emph{i.e.}, tropical,
  temperate broadleaf and deciduous, boreal) based on the current ForC
  data?
\item
  How do C cycling vary across the world's major forest biomes?
\item
  How does C cycling vary with stand age (in interaction with biome)?
\end{enumerate}

While components of these questions have been previously addressed
(Luyssaert \emph{et al} 2007, Anderson-Teixeira \emph{et al} 2016,
Cook-Patton \emph{et al} 2020, Banbury Morgan \emph{et al} n.d.), our
analysis represents by far the most comprehensive analysis of C cycling
in global forests, and will serve as a foundation for improved
understanding of global forest C cycling.

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/World_Map_records_in_Biomes}

}

\caption{Figure 1 | Map of sites included in this analysis. Symbols are colored according to the number of records at each site. Underlying map shows coverage of evergreen, deciduous, and mixed forests (from SYNMAP; Jung et al. 2006) and biomes. Distribution of sites, plots, and records among biomes is shown in the inset.}\label{fig:unnamed-chunk-4}
\end{figure}

\hypertarget{methods-design}{%
\subsection{Methods/ Design}\label{methods-design}}

This review synthesizes data from the ForC database (Fig. 1;
\url{https://github.com/forc-db/ForC}; Anderson-Teixeira \emph{et al}
2016, pp @anderson--teixeira\_forc\_2018). ForC amalgamates numerous
intermediary data sets (\emph{e.g.}, \textbf{REFS}) and original
studies. Original publications were referenced to check values and
obtain information not contained in intermediary data sets, although
this process has not been completed for all records. The database was
developed with goals of understanding how C cycling in forests varies
across broad geographic scales and as a function of stand age. As such,
there has been a focus on incorporating data from regrowth forests
(\emph{e.g.}, Anderson et al 2006, Martin et al 2013, Bonner et al 2013)
and obtaining stand age data when possible (83\% of records in v.2.0;
Anderson-Teixeira \emph{et al} 2018). Particular attention was given to
developing the database for tropical forests (Anderson-Teixeira \emph{et
al} 2016). Since publication of ForC v.2.0, we added the following data
to ForC: the Global Database of Soil Respiration Database (SRDB v.\#\#,
9497 records; Bond-Lamberty and Thomson 2010), the Global Reforestation
Opportunity Assessment database (GROA v1.0, 18143 records; Cook-Patton
\emph{et al} 2020, Anderson-Teixeira \emph{et al} 2020). We've also
added data from individual publications (detailed list at
\url{https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv}),
with a particular focus on productivity (Banbury Morgan \emph{et al}
n.d., e.g., Taylor \emph{et al} 2017), dead wood, and ForestGEO sites
(e.g., Lutz \emph{et al} 2018, p @johnson\_climate\_2018). We note that
there remains a significant amount of relevant data that is not yet
included in ForC, particularly biomass data from national forest
inventories (\emph{e.g.},: \textbf{REFS}). The database version used for
this analysis has been tagged as a new release on Github (\textbf{XX})
and assigned a DOI through Zenodo (DOI: TBD).

Analyses drew from ForC-simplified
(\url{https://github.com/forc-db/ForC/blob/master/ForC_simplified}),
which is a rearrangement of ForC intended to facilitate analyses. In
generating ForC-simplified, all measurements originally expressed in
units of dry organic matter (\(OM\)) were converted to units of C using
the IPCC default of \(C = 0.47 * OM\) (IPCC 2006). Duplicate or
otherwise conflicting records were reconciled as described in
\emph{APPENDIX S1}, resulting in a total of 30927 records (64.6\% size
of total database). Records were filtered to remove plots that had
undergone significant anthropogenic management or major disturbance
since the most recent stand initiation event. Specifically, we removed
all plots flagged as managed in ForC-simplified (19\%). This included
plots with any record of managements manipulating CO\textsubscript{2},
temperature, hydrology, nutrients, or biota, as well as any plots whose
site or plot name contained the terms ``plantation'', ``planted'',
``managed'', ``irrigated'', or ``fertilized''. Plots flagged as
disturbed in ForC-simplified (5.9\%) included stands that had undergone
anthropogenic thinning or partial harvest unless this was very minor. We
retained sites that were grazed or had undergone low severity natural
disturbances (\textless10\% mortality) including droughts, major storms,
fires, and floods. We removed all plots for which no stand history
information had been retrieved (7.7\%). In total, this resulted in 21872
records (45.7\% of the records in the database) being eligible for
inclusion in the analysis.

We selected 23 annual flux and 11 C stock variables for inclusion in the
analysis, although two flux variables (\(R_{het-ag}\) and \(R_{het}\))
were included for conceptual completeness but had no records in ForC
(Table 1). Records for these variables represented 89.1\% of the total
records eligible for inclusion. For this analysis, we combined some of
ForC's specific variables (\emph{e.g.}, multiple variables for net
primary productivity including various components) into more broadly
defined variables. Specifically, net ecosystem exchange (measured by
eddy-covariance; \textbf{FLUXNET REF}) and biometric estimates of
\(NEP\) were combined into the single variable \(NEP\) (Table 1).
Furthermore, for \(NPP\), \(ANPP\), and \(ANPP_{litterfall}\), ForC
variables specifying inclusion of different components were combined.
Throughout ForC, for all measurements drawing from tree census data
(\emph{e.g.}, biomass, productivity), the minimum threshold for tree
census was \(\le\) 10cm. All records were measured directly or derived
from field measurements (as opposed to modeled).

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-6}Table 1. Carbon cycle variables included in this analysis, their sample sizes, and summary of biome differences and age trends.}
\centering
\resizebox{\linewidth}{!}{
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{7cm}>{\raggedleft\arraybackslash}p{1.1cm}>{\raggedleft\arraybackslash}p{1cm}>{\raggedleft\arraybackslash}p{1.8cm}ll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{3}{c}{N records} & \multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){3-5}
Variable & Description & records & plots & geographic areas & biome differences\textsuperscript{*} & age trend\textsuperscript{\dag}\\
\midrule
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Annual fluxes}}\\
\hspace{1em}$NEP$ & net ecosystem production or net ecosystem exchange (+ indicates C sink) & 377 & 163 & 89 & n.s. & +; xB\\
\hspace{1em}$GPP$ & gross primary production ($NPP + R_{auto}$ or $R_{eco}-NEE$) & 309 & 117 & 72 & TrB > TeB $\ge$ TeN $\ge$ BoN & +; xB\\
\hspace{1em}$NPP$ & net primary production ($ANPP$ + $BNPP$) & 292 & 165 & 74 & TrB > TeB $\ge$ TeN > BoN & +\\
\hspace{1em}$ANPP$ & aboveground $NPP$ & 566 & 384 & 135 & TrB > TeB $\ge$ TeN $\ge$ BoN & +; xB\\
\hspace{1em}$ANPP_{woody}$ & woody production ($ANPP_{stem}$ + $ANPP_{branch}$) & 126 & 114 & 37 & n.s. & +; xB\\
\hspace{1em}$ANPP_{stem}$ & woody stem production & 458 & 411 & 118 & TrB > TeN $\ge$ TeB > BoN & -; xB\\
\hspace{1em}$ANPP_{branch}$ & branch turnover & 125 & 112 & 42 & TrB > TeB $\ge$ TeN & n.s.\\
\hspace{1em}$ANPP_{foliage}$ & foliage production, typically estimated as annual leaf litterfall & 275 & 231 & 89 & TrB > TeB $\ge$ TeN > BoN & +\\
\hspace{1em}$ANPP_{litterfall}$ & litterfall, including leaves, reproductive structures, twigs, and sometimes branches & 147 & 117 & 55 & TrB > TeB & +\\
\hspace{1em}$ANPP_{repro}$ & production of reproductive structures (flowers, fruits, seeds) & 72 & 65 & 35 & n.s. & -\\
\hspace{1em}$ANPP_{folivory}$ & foliar biomass consumed by folivores & 24 & 16 & 11 & n.s. & -\\
\hspace{1em}$M_{woody}$ & woody mortality--i.e., $B_{ag}$ of trees that die & 35 & 35 & 19 & n.s. & -\\
\hspace{1em}$BNPP$ & belowground NPP ($BNPP_{coarse}$+ $BNPP_{fine}$) & 210 & 178 & 76 & TrB > TeB $\ge$ TeN $\ge$ BoN & +\\
\hspace{1em}$BNPP_{coarse}$ & coarse root production & 144 & 111 & 36 & n.s. & n.s.\\
\hspace{1em}$BNPP_{fine}$ & fine root production & 167 & 136 & 62 & n.s. & +\\
\hspace{1em}$R_{eco}$ & ecosystem respiration ($R_{auto}$+ $R_{het}$) & 197 & 81 & 59 & TrB > TeN & +; xB\\
\hspace{1em}$R_{auto}$ & autotrophic respiration ($R_{auto-ag}+ R_{root}$) & 36 & 35 & 15 & TrB > TeN & -\\
\hspace{1em}$R_{auto-ag}$ & aboveground autotrophic respiration (i.e., leaves and stems) & 3 & 3 & 2 & TrB > TeN & -\\
\hspace{1em}$R_{root}$ & root respiration & 227 & 183 & 98 & n.s. & +\\
\hspace{1em}$R_{soil}$ & soil respiration ($R_{het-soil} + R_{root}$) & 644 & 468 & 219 & TrB > TeB $\ge$ TeN > BoN & n.s.\\
\hspace{1em}$R_{het-soil}$ & soil heterotrophic respiration & 213 & 178 & 93 & TrB > TeB $\ge$ TeN & -; xB\\
\hspace{1em}$R_{het-ag}$ & aboveground heterotrophic respiration & 0 & 0 & 0 & - & -\\
\hspace{1em}$R_{het}$ & heterotrophic respiration ($R_{het-ag} +R_{het-soil}$) & 0 & 0 & 0 & - & -\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Stocks}}\\
\hspace{1em}$B_{tot}$ & total live biomass ($B_{ag}$+$B_{root}$) & 299 & 252 & 85 & TeN > TeB $\ge$ TrB $\ge$ BoN & +; xB\\
\hspace{1em}$B_{ag}$ & aboveground live biomass  ($B_{ag-wood}$+$B_{foliage}$) & 5371 & 4761 & 621 & TrB $\ge$ TeN $\ge$ TeB > BoN & +; xB\\
\hspace{1em}$B_{ag-wood}$ & woody component of aboveground biomass & 224 & 210 & 66 & TeN > TrB $\ge$ TeB $\ge$ BoN & +; xB\\
\hspace{1em}$B_{foliage}$ & foliage biomass & 260 & 216 & 74 & TeN > TrB $\ge$ BoN $\ge$ TeB & +; xB\\
\hspace{1em}$B_{root}$ & total root biomass ($B_{root-coarse}$+$B_{root-fine}$) & 2380 & 2337 & 359 & n.s. & +; xB\\
\hspace{1em}$B_{root-coarse}$ & coarse root biomass & 204 & 190 & 68 & TeN > TeB $\ge$ BoN & +; xB\\
\hspace{1em}$B_{root-fine}$ & fine root biomass & 335 & 266 & 107 & n.s. & n.s.\\
\hspace{1em}$DW_{tot}$ & deadwood ($DW_{standing}$+$DW_{down}$) & 223 & 212 & 42 & TeB > TrB & +; xB\\
\hspace{1em}$DW_{standing}$ & standing dead wood & 65 & 61 & 22 & TeB > TrB & -\\
\hspace{1em}$DW_{down}$ & fallen dead wood, including coarse and sometimes fine woody debris & 284 & 271 & 37 & TeB > TrB & +; xB\\
\hspace{1em}$OL$ & organic layer $/$ litter$/$ forest floor & 555 & 482 & 113 & n.s. & +; xB\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{*} Tr: Tropical, TeB: Temperate Broadleaf, TeN: Temperate Needleleaf, B: Boreal, n.s.: no significant differences}\\
\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{\dag} + or -: significant positive or negative trend, xB: significant age x biome interaction, n.s.: no significant age trend}\\
\end{tabular}}
\end{table}

For this analysis, we grouped forests into four broad biome types
(tropical broadleaf, temperate broadleaf, temperate needleleaf, and
boreal needleleaf) and two age classifications (young and mature).
Climate zones (Fig. 1) were defined based on site geographic coordinates
according to Köppen-Geiger zones (Rubel and Kottek 2010). Tropical
climates were defined to include all equatorial (A) zones, temperate
climates were defined to include all warm temperate (C) zones and warmer
snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates
were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd,
Dfc, and Dfd). Any forests in dry (B) and polar (E) Köppen-Geiger zones
were excluded from the analysis. Leaf type (broadleaf / needleleaf) was
defined based on descriptions in original publications (prioritized) or
values extracted from a global map based on satellite observations
(SYNMAP; {\textbf{???}}). For young tropical forests imported from GROA
but not yet classified by leaf type, we assumed that all were broadleaf,
consistent with the rarity of naturally regenerating needleleaf forests
in the tropics (\textbf{REF}). Forests were classified as ``young''
(\textless{} 100 years) or ``mature'' (\(\ge\) 100 years or classified
as ``mature'', ``old growth'', ``intact'', or ``undisturbed'' in
original publication). Assigning stands to these groupings required the
exclusion of records for which ForC lacked records of stand age (4.8\%
of records in full database), or geographic coordinates (0.4\% of sites
in full database). We also excluded records of stand age = 0 year (0.8\%
of records in full database). In total, our analysis retained 76.2 of
the focal variable records for forests of known age. Numbers of records
by biome and age class are given in Table S1.

Data were summarized to produce schematics of C cycling across the eight
biome - age group combinations identified above. To obtain the values
reported in the C cycle schematics, we first averaged any repeated
measurements within a plot. Values were then averaged across plots
clustered within 25 km of one another (\emph{sensu} Anderson-Teixeira
\emph{et al} 2018), weighting by area sampled if available for all
records. This step was taken to avoid pseudo-replication and to combine
any records from sites with more than one name in ForC.

To test for differences across mature forest biomes, the original values
were analyzed via a linear mixed effects model {[}`lmer' function in
`lme4' R package version \textbf{x.xx}; \textbf{REF}{]} with biome as
fixed effect and plot nested within geographic.area as random effects on
the intercept. When Biome had a significant effect, we looked at a
Tukey's pairwise comparison to see which biomes were significantly
different from one another. This analysis was run for variables with
records for \$\ge\$7 distinct geographic areas in more than one biome,
excluding any biomes that failed this criteria (Table 1).

To test for age trends in young (\textless100yrs) forests, we employed a
mixed effects model with biome and log10{[}stand.age{]} as fixed effects
and plot nested within geographic.area as a random effect on the
intercept. This analysis was run for variables with records for \$\ge\$3
distinct geographic areas in more than one biome, excluding any biomes
that failed this criteria (Table 1). When the effect of stand age was
significant at p \(\le\) 0.05 and when each biome had records for stands
of at least 10 different ages, a biome * stand.age interaction was
included in the model. All database manipulation, analyses, and figure
production were fully automated in R (\textbf{version, citation}).

\hypertarget{review-results-synthesis}{%
\subsection{Review Results/ Synthesis}\label{review-results-synthesis}}

\hypertarget{data-coverage}{%
\subsubsection{Data Coverage}\label{data-coverage}}

Of the 47846 records in ForC v3.0, 14847 met the criteria for inclusion
in this study (Fig. 1). These records were distributed across 5889 plots
in 874 distinct geographic areas. Of the 23 flux and 11 stock variables
mapped in these diagrams, ForC contained mature forest estimates from
\(\ge\) 7 distinct geographic areas for 20 fluxes and 9 stocks in
tropical broadleaf forests, 14 fluxes and 9 stocks in temperate
broadleaf forests, 15 fluxes and 8 stocks in temperate conifer forests,
and 8 fluxes and 7 stocks in boreal forests. For regrowth forests
(\textless100 yrs), ForC contained estimates from \(\ge\) 7 distinct
geographic areas for 3 fluxes and 9 stocks in tropical broadleaf
forests, 12 fluxes and 8 stocks in temperate broadleaf forests, 15
fluxes and 8 stocks in temperate conifer forests, and 12 fluxes and 7
stocks in boreal forests.
\textbf{(\href{https://github.com/forc-db/ERL-review/issues/29\#issuecomment-699633801}{issue
29})}

\hypertarget{c-cycling-in-mature-forests}{%
\subsubsection{C cycling in mature
forests}\label{c-cycling-in-mature-forests}}

Average C cycles for tropical broadleaf, temperate broadleaf, temperate
conifer, and boreal forests \(\ge\) 100 years old and with no known
major disturbance or significant anthropogenic management are presented
in Figures 2-5 (and available in tabular format in the ForC release
accompanying this publication:
\texttt{ForC/numbers\_and\_facts/ForC\_variable\_averages\_per\_Biome.csv}).

For variables with records from \(\ge\) 7 distinct geographic areas,
these ensemble C budgets were generally consistent. \emph{That is,
component fluxes and stocks summed to within one standard deviation of
more inclusive fluxes in all but} \textbf{two} \emph{instances, both in
temperate conifer forests (Fig. 5).} \textbf{(check this with final
results:
\url{https://github.com/forc-db/ForC/blob/master/numbers_and_facts/C_cycle_closure.csv})
} \emph{Within this biome, \(B_{ag}<B_{ag-wood}+B_{foliage}\) and
\(B_{root}<B_{root-coarse}+B_{root-fine}\) because \(B_{ag-wood}\) and
\(B_{root-coarse}\) were very high, with strongly disproportionate
numbers of records from the high-biomass forests of the US Pacific
Northwest (Figs. S18, S21).}

\begin{landscape}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Tropical broadleaf MATURE}

}

\caption{Figure 2 | C cycle diagram for mature tropical broadleaf forests. All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.  Arrows indicate fluxes, boxes indicate stocks. Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data. Arrow size is proportional to the square root of corresponding flux. Asterisk after variable name indicates lack of C cycle closure.}\label{fig:unnamed-chunk-8}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate broadleaf MATURE}

}

\caption{Figure 3 | C cycle diagram for mature temperate broadleaf forests. All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.  Arrows indicate fluxes, boxes indicate stocks. Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data. Arrow size is proportional to the square root of corresponding flux. Asterisk after variable name indicates lack of C cycle closure.}\label{fig:unnamed-chunk-9}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate conifer MATURE}

}

\caption{Figure 4 | C cycle diagram for mature temperate conifer forests. All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.  Arrows indicate fluxes, boxes indicate stocks. Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data. Arrow size is proportional to the square root of corresponding flux. Asterisk after variable name indicates lack of C cycle closure.}\label{fig:unnamed-chunk-10}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Boreal conifer MATURE}

}

\caption{Figure 5 | C cycle diagram for mature boreal conifer forests. All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.  Arrows indicate fluxes, boxes indicate stocks. Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data. Arrow size is proportional to the square root of corresponding flux. Asterisk after variable name indicates lack of C cycle closure.}\label{fig:unnamed-chunk-11}
\end{figure}
\end{landscape}

\textbf{check this with final data (based on Table 1):} There were
sufficient data to assess mature forest biome differences for
\textbf{20} flux variables, and significant differences among biomes
were detected for \textbf{14} variables (Table 1). With only
\textbf{one} exception (\(R_{auto}\), which had low sample size;
\textbf{\href{https://github.com/forc-db/ERL-review/issues/40}{issue
40}}), C fluxes were highest in tropical forests, intermediate in
temperate (broadleaf or conifer) forests, and lowest in boreal forests
(Table 1, Figs. 6, S1-S15). Differences between tropical and boreal
forests were always significant, with temperate forests intermediate and
significantly different from one or both. Fluxes tended to be
numerically greater in temperate broadleaf than conifer forests, but the
difference was never statistically significant. This pattern held for
the following variables: ** \(GPP\), \(NPP\), \(ANPP\),
\(ANPP_{woody}\), \(ANPP_{stem}\),\(ANPP_{branch}\), \(ANPP_{foliage}\),
\(ANPP_{litterfall}\), \(M_{woody}\), \(BNPP\), \(R_{eco}\),
\(R_{soil}\), and \(R_{het-soil}\)**. For variables without significant
differences among biomes, the same general trends applied.

The most notable exception to this pattern was \(NEP\), with no
significant differences across biomes but with the largest average in
temperate broadleaf forests, followed by temperate conifer, boreal, and
tropical forests (Figs. 5,S1). Another exception was for
\(BNPP_{root-coarse}\), where all records came from high-biomass forests
in the US Pacific Northwest, resulting in high values for the temperate
conifer biome and no significant differences across biomes (Fig. S10).\\
Thus, C cycling rates generally decreased from tropical to temperate to
boreal forests, with the important exception in the overall C balance
(\(NEP\)).

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/for_ERL_review/Flux_age_trends}

}

\caption{Figure 6 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEP$. Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively). Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome. Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests). Individual figures for each flux with sufficient data given in the Supplement (Figs. S1-S15).}\label{fig:unnamed-chunk-12}
\end{figure}

\textbf{check this with final data (based on Table 1):} There were
sufficient data to assess mature forest biome differences for \textbf{9}
stock variables, and significant differences among biomes were detected
for \textbf{6} variables (\textbf{\(B_{tot}\), \(B_{ag}\),
\(B_{ag-wood}\), \(B_{foliage}\), \(B_{root-coarse}\), \(DW_{tot}\)};
Table 1). C stocks had less consistent patterns across biomes (Figs. 7,
S16-S26). In \textbf{four of the six} cases (\textbf{\(B_{tot}\),
\(B_{ag-wood}\), \(B_{root-coarse}\), \(DW_{tot}\)}), temperate conifer
forests were in the highest significance grouping, and boreal forests in
the lowest, with tropical and temperate broadleaf forests in
between--most commonly being significantly different from temperate
conifer but not boreal forests. For \(B_{ag}\), which had by far the
highest sample size, tropical forests exceeded temperate conifer forests
(but not significantly). For \(B_{foliage}\),temperate broadleaf forests
were lowest (again, not significantly). The high values for the
temperate conifer biome were driven by the very high-biomass forests of
the US Pacific Northwest, which are disproportionately represented in
the current version of ForC. Thus, biome differences should be
interpreted more as driven more by geographic distribution of sampling
than by true differences.

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/for_ERL_review/Stock_age_trends}

}

\caption{Figure 7 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood. Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively). Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome. Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests). Individual figures for each stock with sufficient data given in the Supplement (Figs. S16-S26).}\label{fig:unnamed-chunk-13}
\end{figure}

\hypertarget{c-cycling-in-young-forests}{%
\subsubsection{C cycling in young
forests}\label{c-cycling-in-young-forests}}

Average C cycles for forests \textless100 years old are presented in
Figures 8-11.\\
Both C stocks and fluxes commonly displayed significant trends with
stand age for within-biome analyses (Tables 1, S2, Figs. 6-11, S1-S26;
detailed below).

ForC contained \emph{14} flux variables with sufficient data for
cross-biome analyses of age trends in regrowth forests (see Methods)
(Fig. 6-7 and \textbf{S\#- SI figures including plots for all
variables}). Of these, \emph{9} increased significantly with
log10{[}stand.age{]}: \(GPP\), \(NPP\), \(ANPP\), \(ANPP_{foliage}\),
\(ANPP_{woody}\), \(ANPP_{woody-stem}\), \(BNPP\), \(BNPP_{root-fine}\),
\(R_{eco}\), and net C sequestration (\(NEP\)). The remaining
five--\(ANPP_{woody-branch}\), \(BNPP_{root-coarse}\), \(R_{soil-het}\),
and \(R_{soil-het}\)-\/--displayed no significant relationship to stand
age, although all displayed a positive trend.

Differences in C fluxes across biomes typically paralleled those
observed for mature forests, with C cycling generally most rapid in the
tropics and slowest in boreal forests.\\
The single exception was \(ANPP_{stem}\), for which temperate broadleaf
forests and temperate conifer forests of age
\textgreater{}\emph{\textasciitilde30} had slightly higher flux rates
than tropical forests (\emph{represented by a single chronosequence}).
Notably, the trend of tropical \textgreater{} temperate \textgreater{}
boreal held for \(NEP\) in regrowth forests, in contrast to the lack of
biome differences in \(NEP\) for mature forests (Fig. 6).

There were only \emph{\#\#} flux variables with sufficient data to test
for biome x age interactions: \(ANPP\), \(ANPP_{woody}\),
\(ANPP_{stem}\), \(ANPP_{litterfall}\), and \(BNPP\) (Table S2).
\textbf{(more could be added if age trends become significant after
outliers are resolved)} For three of these (\(ANPP\),
\(ANPP_{litterfall}\), \(BNPP\)), the increase in C flux with age was
steepest increase in tropical forests, followed by temperate and then
boreal forests (Figs S\#). Similarly, \(ANPP_{woody}\) displayed a
steeper increase with age in temperate than boreal boreal forests (no
tropical data for this variable). In contrast, for \(ANPP_{stem}\),
tropical and temperate broadleaf forests had the highest flux rates at
young ages, but were surpassed by temperate conifer forests between ages
20 and 50 (Fig. S6).

(\textbf{this needs to be updated with latest data}) In terms of C
stocks, \emph{10} variables had sufficient data to test for age trends.
\emph{Six} of these---\emph{total biomass, aboveground biomass,
aboveground woody biomass, foliage biomass, root biomass, and coarse
root biomass}--- increased significantly with \(log10[stand.age]\).
\emph{The remaining four displayed non-significant positive trends: fine
root biomass, total dead wood, standing dead wood, and organic layer.}
\emph{(discuss rates of increase)}

\newpage
\begin{landscape}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Tropical broadleaf YOUNG}

}

\caption{Figure 8 | C cycle diagram for young tropical broadleaf forests. Presented are observed ranges, where geographically distinct areas are treated as the unit of replication. All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks). Arrows indicate fluxes, boxes indicate stocks. Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data. Arrow size is proportional to the square root of corresponding flux.}\label{fig:unnamed-chunk-14}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate broadleaf YOUNG}

}

\caption{Figure 9 | C cycle diagram for young temperate broadleaf forests. Presented are observed ranges, where geographically distinct areas are treated as the unit of replication. All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks). Arrows indicate fluxes, boxes indicate stocks. Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data. Arrow size is proportional to the square root of corresponding flux.}\label{fig:unnamed-chunk-15}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate conifer YOUNG}

}

\caption{Figure 10 | C cycle diagram for young temperate conifer forests. Presented are observed ranges, where geographically distinct areas are treated as the unit of replication. All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks). Arrows indicate fluxes, boxes indicate stocks. Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data. Arrow size is proportional to the square root of corresponding flux.}\label{fig:unnamed-chunk-16}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Boreal conifer YOUNG}

}

\caption{Figure 11 | C cycle diagram for young boreal conifer forests. Presented are observed ranges, where geographically distinct areas are treated as the unit of replication. All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks). Arrows indicate fluxes, boxes indicate stocks. Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data. Arrow size is proportional to the square root of corresponding flux.}\label{fig:unnamed-chunk-17}
\end{figure}
\end{landscape}

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

ForC v3.0 brought together an unprecedented amount of data to yield an
internally consistent picture of C cycling in the world's major forest
biomes. Carbon cycling rates generally increased from boreal to tropical
regions and with stand age. Specifically, the major C fluxes were
highest in tropical forests, intermediate in temperate (broadleaf or
conifer) forests, and lowest in boreal forests -- a pattern that
generally held for regrowth as well as mature forests (Figs. 6-7). In
contrast to C fluxes, there was little directional variation in mature
forest C stocks across biomes (Figs. 2-5, 7). The majority of flux
variables, together with most live biomass pools, increased
significantly with stand age (Figs. 6-11). Together, these results
indicate that, moving from cold to tropical climates and from young to
old stands, there is a general acceleration of C cycling, whereas C
stocks and \(NEP\) of mature forests are correlated with a different set
of factors.

\hypertarget{c-variable-coverage-and-budget-closure}{%
\subsubsection{C variable coverage and budget
closure}\label{c-variable-coverage-and-budget-closure}}

ForC provides provides unprecedented coverage of most major variables.
\emph{(discuss how this improves upon previous data compilations/ for
which variables does ForC make the greatest difference (e.g., not AGB or
NEP/GPP/Reco, but by far the latest data compilation for dead wood,
{[}variables{]})} \emph{(Noteable holes include: fluxes: R\_auto\_ag,
woody mortality, folivary/ herbivory and respiration of herbivores (and
therefore total R\_het), ANPP\_repro; also fluxes in tropical regrowth
forests)} \emph{For the C stocks considered here, the most poorly
covered is dead wood (none in E hemisphere!), despite a focused effort
on this variable that has resulted in ForC being by far the largest
collection of these data.} Thus, overall, we're lacking coverage of
fluxes to herbivores and higher consumers, along with the woody
mortality and dead wood. Geographically, all variables poorly covered in
Africa and Siberia.

\textbf{notes from Ben on the above par:} Pregitzer and Euskirchen 2004
\url{http://dx.doi.org/10.1111/j.1365-2486.2004.00866.x} ``Aggregated
biome-level estimates of NPP and NEP were higher in intermediate-aged
forests (e.g., 30--120 years), while older forests (e.g., 4120 years)
were generally less productive. The mean NEP in the youngest forests
(0--10 years) was negative (source to the atmosphere) in both boreal and
temperate biomes\ldots Forest age is a highly significant source of
variability in NEP at the biome scale''

Amiro et al.~2010 \url{http://dx.doi.org/10.1029/2010JG001390} Houghton
et al.~2020 \url{https://doi.org/10.1111/gcb.15050}

Turnover: Pugh et al.~2020 \url{http://dx.doi.org/10.5194/bg-2019-491}
Yu et al.~2019 \url{http://dx.doi.org/10.1073/pnas.1821387116}

Human footprint in global forests
\url{http://dx.doi.org/10.1038/nature05847}
\url{http://dx.doi.org/10.1038/nature02619}

Mention consistency (or lack of) with e.g.~GOLUM-CNP?
\url{http://dx.doi.org/10.5194/gmd-11-3903-2018}

In terms of C stocks, there is a paucity of data on dead wood and
organic layer (Pan et al.~?). These can be significant. (\textbf{Note
that we've given a lot of emphasis to dead wood (work by Abby, and also
Jenny), and as a result this work really advances knowledge of dead
wood. We'll want to highlight that here.}) \emph{(give some stats/ cite
figures)}. ForC does not include soil carbon, which is covered by other
efforts (REFS). For fluxes, Fluxnet is the keeper of the best data on
NEE, GPP, Reco (REFS), and SRDB remains the authority on soil
respiration (REFS). ForC includes recent data from both, but is not
continuously integrated. For C is the best source for most of the
subsidiary fluxes: NPP, woody mortality\ldots{}

The C cycle budgets for mature forests (Figs. 2-5) generally
``close''--that is, the sums of component variables do not differ from
the larger fluxes by more than one standard deviation. However, standard
deviations are often large, reflective of significant within-biome
variation. This makes the standard for closure relatively loose.
\emph{Lack of closure, in the few instances where it occurs, is probably
more reflective of differences in the representation of forest types
(e.g., disproportionate representation of US Pacific NW for aboveground
woody biomass relative to AGB; Fig. 4) than of methodological accuracy.}
Thus, overall, a high degree of closure implies that ForC gives a
consistent picture of C cycling within biomes. While these means are
unlikely to be accurate representations of C cycling within any
particular forest, they provide a useful baseline for comparison, always
keeping in mind that sample means do not necessarily represent the true
mean of the entire biome.

\hypertarget{c-cycling-across-biomes}{%
\subsubsection{C cycling across biomes}\label{c-cycling-across-biomes}}

Our analysis reveals a general acceleration of carbon cycling from the
tropics to the high latitudes. For mature forests, this is consistent
with a large body of previous work demonstrating that C fluxes generally
decline with latitude (e.g., Banbury Morgan \emph{et al} n.d.). For
regrowth forests, more rapid accumulation of biomass at lower latitudes
has been well-established (Anderson \emph{et al} 2006, Cook-Patton
\emph{et al} 2020), whereas this is the first study to compare age
trends in deadwood and organic layer across biomes (but see Cook-Patton
\emph{et al} 2020). For most C flux variables, this analysis is the
first to examine flux trends in regrowth forests across biomes (i.e.,
age x biome interaction). Data remain sparse, but for better-represented
variables, we often see faster acceleration of C cycling in the warmer
climates. Further work will be required to explore age x climate
interactions, but our broad-brush overview indicates that C cycling of
regrowth forests is not only higher in the tropics, parallel to fluxes
in mature forests (Banbury Morgan \emph{et al} n.d.), but also that it
accelerates more rapidly with stand age in the tropics, consistent with
more rapid biomass accumulation.

In contrast to C fluxes and accumulation rates in regrowth forests,
stocks\ldots{}

Thus, biome differences should be interpreted more as driven more by
geographic distribution of sampling than by true differences.

Higher NEP in temperate forests -- implications? Invariant NEP in older
forests? This could be built out a bit; thinking of Luyssaert 2008
\url{http://dx.doi.org/10.1038/nature07276} and following papers arguing
about this.

\hypertarget{age-trends-in-c-cycling}{%
\subsubsection{Age trends in C cycling}\label{age-trends-in-c-cycling}}

\emph{(Just some rough notes at this point)}

A relative dearth of data on C cycling in secondary forests,
particularly in the tropics (Anderson-Teixeira et al 2016), is
problematic in that almost 2/3 of the world's forests were secondary as
of 2010 (FAO 2010), implying an under-filled need to characterize
age-related trends in forest C cycling.

Moreover, as disturbances increase ({\textbf{???}}, McDowell \emph{et
al} 2020), understanding the carbon dynamics of regrowth forests will be
increasingly important.

It's also important to understand secondary forest C sequestration to
reduce uncertainty regarding the potential for carbon uptake by regrowth
forests ({\textbf{???}}, Cook-Patton \emph{et al} 2020).

NEP increases with log(age) to 100 --\textgreater{} strongest C sinks
are established secondary forests. (But presumably this exact number is
an artifact; don't over-emphasize.)

\hypertarget{relevance-for-climate-change-prediction-and-mitigation}{%
\subsubsection{Relevance for climate change prediction and
mitigation}\label{relevance-for-climate-change-prediction-and-mitigation}}

The future of forest C cycling ({\textbf{???}}) will shape trends in
atmospheric CO\textsubscript{2} and the course of climate change. For a
human society seeking to understand and mitigate climate change, the
data contained in ForC and summarized here can help to meet two major
challenges.

First, improved representation of forest C cycling in models is
essential to improving predictions of the future course of climate
change. To ensure that models are giving the right answers for the right
reasons, it is important to benchmark against multiple components of the
C cycle that are internally consistent with each other. By making tens
of thousands of records readily available in standardized format, ForC
makes it feasible for the modeling community to draw upon these data to
benchmark models. Integration of ForC with models is a goal (Fer et al.,
in revision).

Second, ForC can serve as a pipeline through which forest science can
inform forest-based climate change mitigation efforts. Such efforts will
be most effective when informed by the best available data, yet it is
not feasible for the individuals and organizations designing such
efforts to sort through literature, often behind paywalls, with data
reported in varying units, terminology, etc. One goal for ForC is to
serve as a pipeline through which information can flow efficiently from
forest researchers to decision-makers working to implement forest
conservation strategies at global, national, or landscape scales. This
is already happening! ForC has already contributed to updating the IPCC
guidelines for carbon accounting in forests {[}IPCC 2019; Requena Suarez
\emph{et al} (2019); Rozendaal et al in prep{]}, mapping C accumulation
potential from natural forest regrowth globally (Cook-Patton \emph{et
al} 2020), and informing ecosystem conservation priorities (Goldstein
\emph{et al} 2020).

\textbf{ForC can complement remote sensing to provide a comprehensive
picture of global forest C cycling.} (\emph{here, discuss what is
well-covered by remote sesnsing, where ForC is useful for calibrating
remote sensing, and what ForC can get that remote sensing can't.}) AGB
is the largest stock, and most of the emphasis is on this variable.
Remote sensing, with calibration based on high-quality ground-based data
(Schepaschenko \emph{et al} 2019, Chave \emph{et al} 2019), is the best
approach for mapping forest carbon (REFS). However, it is limited in
that it is not associated with stand age and disturbance history, except
in recent decades when satellite data can be used to detect forest loss,
gain, and some of their dominant drivers (Hansen \emph{et al} 2013, Song
\emph{et al} 2018, Curtis \emph{et al} 2018). ForC is therefore valuable
in defining age-based trajectories in biomass, as in Cook-Patton
\emph{et al} (2020).

\emph{Remote sensing measurements are increasingly useful for global- or
regional-scale estimates of forest \(GPP\) (Bagdley et al.~2019, (Li and
Xiao 2019)), aboveground biomass (\(B_{ag}\)) (REFS), woody mortality
(}i.e.\emph{, \(B_{ag}\) losses to mortality \(M_{woody}\)) (Clark
\emph{et al} 2004, Leitold \emph{et al} 2018), and to some extent net
ecosystem exchange (\(NEP\)) (REFS).} Other variables, in particular
respiration fluxes, cannot be remotely sensed (({\textbf{???}})), and
efforts such as the Global Carbon Project (\textbf{le quere REF}) and
NASA CMS (\textbf{citation:
\url{https://carbon.nasa.gov/pdfs/CMS_Phase-1_Report_Final_optimized.pdf}
but maybe better to cite open literature, one of the papers listed at
\url{https://cmsflux.jpl.nasa.gov/get-data/publication-data-sets}})
typically compute them as residuals only. (\textbf{Ben, it woudl be
particularly helpful if you could flesh this out some more.})

\textbf{Move to data availability statement, or methods?}: We recommend
that use of ForC data go to the original database, as opposed to using
``off-the-shelf'' values from this publication. This is because (1) ForC
is constantly being updated, (2) analyses should be designed to match
the application, (3) age equations presented here all fit a single
functional form that is not necessarily the best possible for all the
variables.

As climate change accelerates, understanding and managing the carbon
dynamics of forests will be critical to forecasting, mitigation, and
adaptation. The C data in ForC, as summarized here, will be valuable to
these efforts.

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

All researchers whose data is included in ForC and this analysis. Ian
McGregor for help with the database. Thanks to Norbert Kunert and
{[}Helene's intern{]} for helpful input at an earlier phase. A
Smithsonian Scholarly Studies grant to KAT and HML. WLS grant to KAT.

\hypertarget{data-availability-statement}{%
\subsection{Data availability
statement}\label{data-availability-statement}}

Materials required to fully reproduce these analyses, including data, R
scripts, and image files, are archived in Zenodo (DOI: TBD{]}. Data,
scripts, and results presented here are also available through the
open-access ForC GitHub repository
(\url{https://github.com/forc-db/ForC}), where many will be updated as
the database develops.

\hypertarget{orcid-id}{%
\subsection{ORCID iD}\label{orcid-id}}

Kristina J. Anderson-Teixeira:
\url{https://orcid.org/0000-0001-8461-9713}

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-anderson_temperature-dependence_2006}{}%
Anderson K J, Allen A P, Gillooly J F and Brown J H 2006
Temperature-dependence of biomass accumulation rates during secondary
succession \emph{Ecology Letters} \textbf{9} 673--82 Online:
\url{http://www.blackwell-synergy.com/doi/abs/10.1111/j.1461-0248.2006.00914.x}

\leavevmode\hypertarget{ref-anderson-teixeira_forc-dbgroa_2020}{}%
Anderson-Teixeira K, Herrmann V, CookPatton, Ferson A and Lister K 2020
Forc-db/GROA: Release with Cook-Patton et al. 2020, Nature. Online:
\url{https://zenodo.org/record/3983644}

\leavevmode\hypertarget{ref-anderson-teixeira_ctfs-forestgeo_2015}{}%
Anderson-Teixeira K J, Davies S J, Bennett A C, Gonzalez-Akre E B,
Muller-Landau H C, Joseph Wright S, Abu Salim K, Almeyda Zambrano A M,
Alonso A, Baltzer J L, Basset Y, Bourg N A, Broadbent E N, Brockelman W
Y, Bunyavejchewin S, Burslem D F R P, Butt N, Cao M, Cardenas D, Chuyong
G B, Clay K, Cordell S, Dattaraja H S, Deng X, Detto M, Du X, Duque A,
Erikson D L, Ewango C E N, Fischer G A, Fletcher C, Foster R B, Giardina
C P, Gilbert G S, Gunatilleke N, Gunatilleke S, Hao Z, Hargrove W W,
Hart T B, Hau B C H, He F, Hoffman F M, Howe R W, Hubbell S P,
Inman-Narahari F M, Jansen P A, Jiang M, Johnson D J, Kanzaki M, Kassim
A R, Kenfack D, Kibet S, Kinnaird M F, Korte L, Kral K, Kumar J, Larson
A J, Li Y, Li X, Liu S, Lum S K Y, Lutz J A, Ma K, Maddalena D M, Makana
J-R, Malhi Y, Marthews T, Mat Serudin R, McMahon S M, McShea W J,
Memiaghe H R, Mi X, Mizuno T, Morecroft M, Myers J A, Novotny V,
Oliveira A A de, Ong P S, Orwig D A, Ostertag R, Ouden J den, Parker G
G, Phillips R P, Sack L, Sainge M N, Sang W, Sri-ngernyuang K, Sukumar
R, Sun I-F, Sungpalee W, Suresh H S, Tan S, Thomas S C, Thomas D W,
Thompson J, Turner B L, Uriarte M, Valencia R, et al 2015
CTFS-ForestGEO: A worldwide network monitoring forests in an era of
global change \emph{Global Change Biology} \textbf{21} 528--49 Online:
\url{http://onlinelibrary.wiley.com/doi/10.1111/gcb.12712/abstract}

\leavevmode\hypertarget{ref-anderson-teixeira_forc_2018}{}%
Anderson-Teixeira K J, Wang M M H, McGarvey J C, Herrmann V, Tepley A J,
Bond-Lamberty B and LeBauer D S 2018 ForC: A global database of forest
carbon stocks and fluxes \emph{Ecology} \textbf{99} 1507--7 Online:
\url{http://doi.wiley.com/10.1002/ecy.2229}

\leavevmode\hypertarget{ref-anderson-teixeira_carbon_2016}{}%
Anderson-Teixeira K J, Wang M M H, McGarvey J C and LeBauer D S 2016
Carbon dynamics of mature and regrowth tropical forests derived from a
pantropical database (TropForC-db) \emph{Global Change Biology}
\textbf{22} 1690--709 Online:
\url{http://onlinelibrary.wiley.com/doi/10.1111/gcb.13226/abstract}

\leavevmode\hypertarget{ref-baldocchi_fluxnet_2001}{}%
Baldocchi D, Falge E, Gu L, Olson R, Hollinger D, Running S, Anthoni P,
Bernhofer C, Davis K, Evans R, Fuentes J, Goldstein A, Katul G, Law B,
Lee X, Malhi Y, Meyers T, Munger W, Oechel W, Paw K T, Pilegaard K,
Schmid H P, Valentini R, Verma S, Vesala T, Wilson K and Wofsy S 2001
FLUXNET: A New Tool to Study the Temporal and Spatial Variability of
Ecosystem--Scale Carbon Dioxide, Water Vapor, and Energy Flux Densities
\emph{Bulletin of the American Meteorological Society} \textbf{82}
2415--34 Online:
\url{http://journals.ametsoc.org/doi/abs/10.1175/1520-0477(2001)082\%3C2415:FANTTS\%3E2.3.CO;2}

\leavevmode\hypertarget{ref-banbury_morgan_global_nodate}{}%
Banbury Morgan B, Herrmann V, Kunert N, Bond-Lamberty B, Muller-Landau H
C and Anderson-Teixeira K J Global patterns of forest autotrophic carbon
fluxes \emph{Global Change Biology}

\leavevmode\hypertarget{ref-bonan_forests_2008}{}%
Bonan G B 2008 Forests and Climate Change: Forcings, Feedbacks, and the
Climate Benefits of Forests \emph{Science} \textbf{320} 1444--9 Online:
\url{http://www.sciencemag.org/cgi/content/abstract/320/5882/1444}

\leavevmode\hypertarget{ref-bond-lamberty_global_2010}{}%
Bond-Lamberty B and Thomson A 2010 A global database of soil respiration
data \emph{Biogeosciences} \textbf{7} 1915--26 Online:
\url{http://www.biogeosciences.net/7/1915/2010/}

\leavevmode\hypertarget{ref-chave_ground_2019}{}%
Chave J, Davies S J, Phillips O L, Lewis S L, Sist P, Schepaschenko D,
Armston J, Baker T R, Coomes D, Disney M, Duncanson L, Hérault B,
Labrière N, Meyer V, Réjou-Méchain M, Scipal K and Saatchi S 2019 Ground
Data are Essential for Biomass Remote Sensing Missions \emph{Surveys in
Geophysics} \textbf{40} 863--80 Online:
\url{https://doi.org/10.1007/s10712-019-09528-w}

\leavevmode\hypertarget{ref-clark_quantifying_2004}{}%
Clark D B, Castro C S, Alvarado L D A and Read J M 2004 Quantifying
mortality of tropical rain forest trees using high-spatial-resolution
satellite data \emph{Ecology Letters} \textbf{7} 52--9 Online:
\url{https://onlinelibrary.wiley.com/doi/abs/10.1046/j.1461-0248.2003.00547.x}

\leavevmode\hypertarget{ref-cook-patton_mapping_2020}{}%
Cook-Patton S C, Leavitt S M, Gibbs D, Harris N L, Lister K,
Anderson-Teixeira K J, Briggs R D, Chazdon R L, Crowther T W, Ellis P W,
Griscom H P, Herrmann V, Holl K D, Houghton R A, Larrosa C, Lomax G,
Lucas R, Madsen P, Malhi Y, Paquette A, Parker J D, Paul K, Routh D,
Roxburgh S, Saatchi S, Hoogen J van den, Walker W S, Wheeler C E, Wood S
A, Xu L and Griscom B W 2020 Mapping carbon accumulation potential from
global natural forest regrowth \emph{Nature} \textbf{585} 545--50
Online: \url{http://www.nature.com/articles/s41586-020-2686-x}

\leavevmode\hypertarget{ref-curtis_classifying_2018}{}%
Curtis P G, Slay C M, Harris N L, Tyukavina A and Hansen M C 2018
Classifying drivers of global forest loss \emph{Science} \textbf{361}
1108--11 Online:
\url{http://science.sciencemag.org/content/361/6407/1108}

\leavevmode\hypertarget{ref-goldstein_protecting_2020}{}%
Goldstein A, Turner W R, Spawn S A, Anderson-Teixeira K J, Cook-Patton
S, Fargione J, Gibbs H K, Griscom B, Hewson J H, Howard J F, Ledezma J
C, Page S, Koh L P, Rockström J, Sanderman J and Hole D G 2020
Protecting irrecoverable carbon in Earth's ecosystems \emph{Nature
Climate Change} 1--9 Online:
\url{http://www.nature.com/articles/s41558-020-0738-8}

\leavevmode\hypertarget{ref-grassi_key_2017}{}%
Grassi G, House J, Dentener F, Federici S, Elzen M den and Penman J 2017
The key role of forests in meeting climate targets requires science for
credible mitigation \emph{Nature Climate Change} \textbf{7} 220--6
Online: \url{https://www.nature.com/articles/nclimate3227}

\leavevmode\hypertarget{ref-griscom_natural_2017}{}%
Griscom B W, Adams J, Ellis P W, Houghton R A, Lomax G, Miteva D A,
Schlesinger W H, Shoch D, Siikamäki J V, Smith P, Woodbury P, Zganjar C,
Blackman A, Campari J, Conant R T, Delgado C, Elias P, Gopalakrishna T,
Hamsik M R, Herrero M, Kiesecker J, Landis E, Laestadius L, Leavitt S M,
Minnemeyer S, Polasky S, Potapov P, Putz F E, Sanderman J, Silvius M,
Wollenberg E and Fargione J 2017 Natural climate solutions
\emph{Proceedings of the National Academy of Sciences} \textbf{114}
11645--50 Online:
\url{http://www.pnas.org/lookup/doi/10.1073/pnas.1710465114}

\leavevmode\hypertarget{ref-hansen_high-resolution_2013}{}%
Hansen M C, Potapov P V, Moore R, Hancher M, Turubanova S A, Tyukavina
A, Thau D, Stehman S V, Goetz S J, Loveland T R, Kommareddy A, Egorov A,
Chini L, Justice C O and Townshend J R G 2013 High-Resolution Global
Maps of 21st-Century Forest Cover Change \emph{Science} \textbf{342}
850--3 Online:
\url{http://www.sciencemag.org/cgi/doi/10.1126/science.1244693}

\leavevmode\hypertarget{ref-johnson_climate_2018}{}%
Johnson D J, Needham J, Xu C, Massoud E C, Davies S J, Anderson-Teixeira
K J, Bunyavejchewin S, Chambers J Q, Chang-Yang C-H, Chiang J-M, Chuyong
G B, Condit R, Cordell S, Fletcher C, Giardina C P, Giambelluca T W,
Gunatilleke N, Gunatilleke S, Hsieh C-F, Hubbell S, Inman-Narahari F,
Kassim A R, Katabuchi M, Kenfack D, Litton C M, Lum S, Mohamad M,
Nasardin M, Ong P S, Ostertag R, Sack L, Swenson N G, Sun I F, Tan S,
Thomas D W, Thompson J, Umaña M N, Uriarte M, Valencia R, Yap S,
Zimmerman J, McDowell N G and McMahon S M 2018 Climate sensitive
size-dependent survival in tropical trees \emph{Nature Ecology \&
Evolution} \textbf{2} 1436--42 Online:
\url{http://www.nature.com/articles/s41559-018-0626-z}

\leavevmode\hypertarget{ref-krause_large_2018}{}%
Krause A, Pugh T A M, Bayer A D, Li W, Leung F, Bondeau A, Doelman J C,
Humpenöder F, Anthoni P, Bodirsky B L, Ciais P, Müller C,
Murray‐Tortarolo G, Olin S, Popp A, Sitch S, Stehfest E and Arneth A
2018 Large uncertainty in carbon uptake potential of land-based
climate-change mitigation efforts \emph{Global Change Biology}
\textbf{24} 3025--38 Online:
\url{https://onlinelibrary.wiley.com/doi/abs/10.1111/gcb.14144}

\leavevmode\hypertarget{ref-leitold_nino_2018}{}%
Leitold V, Morton D C, Longo M, dos-Santos M N, Keller M and Scaranello
M 2018 El Niño drought increased canopy turnover in Amazon forests
\emph{New Phytologist} \textbf{219} 959--71 Online:
\url{http://doi.wiley.com/10.1111/nph.15110}

\leavevmode\hypertarget{ref-li_mapping_2019}{}%
Li X and Xiao J 2019 Mapping Photosynthesis Solely from Solar-Induced
Chlorophyll Fluorescence: A Global, Fine-Resolution Dataset of Gross
Primary Production Derived from OCO-2 \emph{Remote Sensing} \textbf{11}
2563 Online: \url{https://www.mdpi.com/2072-4292/11/21/2563}

\leavevmode\hypertarget{ref-lutz_global_2018}{}%
Lutz J A, Furniss T J, Johnson D J, Davies S J, Allen D, Alonso A,
Anderson‐Teixeira K J, Andrade A, Baltzer J, Becker K M L, Blomdahl E M,
Bourg N A, Bunyavejchewin S, Burslem D F R P, Cansler C A, Cao K, Cao M,
Cárdenas D, Chang L-W, Chao K-J, Chao W-C, Chiang J-M, Chu C, Chuyong G
B, Clay K, Condit R, Cordell S, Dattaraja H S, Duque A, Ewango C E N,
Fischer G A, Fletcher C, Freund J A, Giardina C, Germain S J, Gilbert G
S, Hao Z, Hart T, Hau B C H, He F, Hector A, Howe R W, Hsieh C-F, Hu
Y-H, Hubbell S P, Inman‐Narahari F M, Itoh A, Janík D, Kassim A R,
Kenfack D, Korte L, Král K, Larson A J, Li Y, Lin Y, Liu S, Lum S, Ma K,
Makana J-R, Malhi Y, McMahon S M, McShea W J, Memiaghe H R, Mi X,
Morecroft M, Musili P M, Myers J A, Novotny V, Oliveira A de, Ong P,
Orwig D A, Ostertag R, Parker G G, Patankar R, Phillips R P, Reynolds G,
Sack L, Song G-Z M, Su S-H, Sukumar R, Sun I-F, Suresh H S, Swanson M E,
Tan S, Thomas D W, Thompson J, Uriarte M, Valencia R, Vicentini A, Vrška
T, Wang X, Weiblen G D, Wolf A, Wu S-H, Xu H, Yamakura T, Yap S and
Zimmerman J K 2018 Global importance of large-diameter trees
\emph{Global Ecology and Biogeography} \textbf{27} 849--64 Online:
\url{https://onlinelibrary.wiley.com/doi/abs/10.1111/geb.12747}

\leavevmode\hypertarget{ref-luyssaert_co2_2007}{}%
Luyssaert S, Inglima I, Jung M, Richardson A D, Reichstein M, Papale D,
Piao S L, Schulze E-D, Wingate L, Matteucci G, Aragao L, Aubinet M, Beer
C, Bernhofer C, Black K G, Bonal D, Bonnefond J-M, Chambers J, Ciais P,
Cook B, Davis K J, Dolman A J, Gielen B, Goulden M, Grace J, Granier A,
Grelle A, Griffis T, Grünwald T, Guidolotti G, Hanson P J, Harding R,
Hollinger D Y, Hutyra L R, Kolari P, Kruijt B, Kutsch W, Lagergren F,
Laurila T, Law B E, Maire G L, Lindroth A, Loustau D, Malhi Y, Mateus J,
Migliavacca M, Misson L, Montagnani L, Moncrieff J, Moors E, Munger J W,
Nikinmaa E, Ollinger S V, Pita G, Rebmann C, Roupsard O, Saigusa N, Sanz
M J, Seufert G, Sierra C, Smith M-L, Tang J, Valentini R, Vesala T and
Janssens I A 2007 CO2 balance of boreal, temperate, and tropical forests
derived from a global database \emph{Global Change Biology} \textbf{13}
2509--37 Online:
\url{http://onlinelibrary.wiley.com/doi/abs/10.1111/j.1365-2486.2007.01439.x}

\leavevmode\hypertarget{ref-mcdowell_pervasive_2020}{}%
McDowell N G, Allen C D, Anderson-Teixeira K, Aukema B H, Bond-Lamberty
B, Chini L, Clark J S, Dietze M, Grossiord C, Hanbury-Brown A, Hurtt G
C, Jackson R B, Johnson D J, Kueppers L, Lichstein J W, Ogle K, Poulter
B, Pugh T A M, Seidl R, Turner M G, Uriarte M, Walker A P and Xu C 2020
Pervasive shifts in forest dynamics in a changing world \emph{Science}
\textbf{368} Online:
\url{https://science.sciencemag.org/content/368/6494/eaaz9463}

\leavevmode\hypertarget{ref-pan_large_2011}{}%
Pan Y, Birdsey R A, Fang J, Houghton R, Kauppi P E, Kurz W A, Phillips O
L, Shvidenko A, Lewis S L, Canadell J G, Ciais P, Jackson R B, Pacala S,
McGuire A D, Piao S, Rautiainen A, Sitch S and Hayes D 2011 A Large and
Persistent Carbon Sink in the World's Forests \emph{Science}
\textbf{333} 988--93 Online:
\url{http://www.sciencemag.org/content/early/2011/07/27/science.1201609.abstract}

\leavevmode\hypertarget{ref-pugh_role_2019}{}%
Pugh T A M, Lindeskog M, Smith B, Poulter B, Arneth A, Haverd V and
Calle L 2019 Role of forest regrowth in global carbon sink dynamics
\emph{Proceedings of the National Academy of Sciences} \textbf{116}
4382--7 Online:
\url{http://www.pnas.org/lookup/doi/10.1073/pnas.1810512116}

\leavevmode\hypertarget{ref-requena_suarez_estimating_2019}{}%
Requena Suarez D, Rozendaal D M A, Sy V D, Phillips O L, Alvarez‐Dávila
E, Anderson‐Teixeira K, Araujo‐Murakami A, Arroyo L, Baker T R, Bongers
F, Brienen R J W, Carter S, Cook‐Patton S C, Feldpausch T R, Griscom B
W, Harris N, Hérault B, Coronado E N H, Leavitt S M, Lewis S L, Marimon
B S, Mendoza A M, N'dja J K, N'Guessan A E, Poorter L, Qie L,
Rutishauser E, Sist P, Sonké B, Sullivan M J P, Vilanova E, Wang M M H,
Martius C and Herold M 2019 Estimating aboveground net biomass change
for tropical and subtropical forests: Refinement of IPCC default rates
using forest plot data \emph{Global Change Biology} \textbf{25} 3609--24
Online: \url{http://onlinelibrary.wiley.com/doi/abs/10.1111/gcb.14767}

\leavevmode\hypertarget{ref-schepaschenko_forest_2019}{}%
Schepaschenko D, Chave J, Phillips O L, Lewis S L, Davies S J,
Réjou-Méchain M, Sist P, Scipal K, Perger C, Herault B, Labrière N,
Hofhansl F, Affum-Baffoe K, Aleinikov A, Alonso A, Amani C,
Araujo-Murakami A, Armston J, Arroyo L, Ascarrunz N, Azevedo C, Baker T,
Bałazy R, Bedeau C, Berry N, Bilous A M, Bilous S Y, Bissiengou P, Blanc
L, Bobkova K S, Braslavskaya T, Brienen R, Burslem D F R P, Condit R,
Cuni-Sanchez A, Danilina D, Torres D del C, Derroire G, Descroix L,
Sotta E D, d'Oliveira M V N, Dresel C, Erwin T, Evdokimenko M D, Falck
J, Feldpausch T R, Foli E G, Foster R, Fritz S, Garcia-Abril A D, Gornov
A, Gornova M, Gothard-Bassébé E, Gourlet-Fleury S, Guedes M, Hamer K C,
Susanty F H, Higuchi N, Coronado E N H, Hubau W, Hubbell S, Ilstedt U,
Ivanov V V, Kanashiro M, Karlsson A, Karminov V N, Killeen T, Koffi J-C
K, Konovalova M, Kraxner F, Krejza J, Krisnawati H, Krivobokov L V,
Kuznetsov M A, Lakyda I, Lakyda P I, Licona J C, Lucas R M, Lukina N,
Lussetti D, Malhi Y, Manzanera J A, Marimon B, Junior B H M, Martinez R
V, Martynenko O V, Matsala M, Matyashuk R K, Mazzei L, Memiaghe H,
Mendoza C, Mendoza A M, Moroziuk O V, Mukhortova L, Musa S, Nazimova D
I, Okuda T, Oliveira L C, et al 2019 The Forest Observation System,
building a global reference dataset for remote sensing of forest biomass
\emph{Scientific Data} \textbf{6} 1--11 Online:
\url{http://www.nature.com/articles/s41597-019-0196-1}

\leavevmode\hypertarget{ref-schimel_observing_2015}{}%
Schimel D, Pavlick R, Fisher J B, Asner G P, Saatchi S, Townsend P,
Miller C, Frankenberg C, Hibbard K and Cox P 2015 Observing terrestrial
ecosystems and the carbon cycle from space \emph{Global Change Biology}
\textbf{21} 1762--76 Online:
\url{http://doi.wiley.com/10.1111/gcb.12822}

\leavevmode\hypertarget{ref-song_global_2018}{}%
Song X-P, Hansen M C, Stehman S V, Potapov P V, Tyukavina A, Vermote E F
and Townshend J R 2018 Global land change from 1982 to 2016
\emph{Nature} \textbf{560} 639--43 Online:
\url{http://www.nature.com/articles/s41586-018-0411-9/}

\leavevmode\hypertarget{ref-taylor_temperature_2017}{}%
Taylor P G, Cleveland C C, Wieder W R, Sullivan B W, Doughty C E,
Dobrowski S Z and Townsend A R 2017 Temperature and rainfall interact to
control carbon cycling in tropical forests ed L Liu \emph{Ecology
Letters} \textbf{20} 779--88 Online:
\url{http://doi.wiley.com/10.1111/ele.12765}

\end{document}

### 2020-09-28 manuscript/ERL_global_C_review.Rmd
(oops, files didn't actually knit on the last commit)

@@
 When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
-This analysis was run for variables with records for $\ge$7 distinct geographic areas in more than one biome, excluding any biomes that failed this criteria (Table 1).
+This analysis was run for variables with records for at least seven distinct geographic areas in more than one biome, excluding any biomes that failed this criteria (Table 1).
 
 To test for age trends in young (<100yrs) forests, we employed a mixed effects model with biome and log10[stand.age] as fixed effects and plot nested within geographic.area as a random effect on the intercept.
-This analysis was run for variables with records for $\ge$3 distinct geographic areas in more than one biome, excluding any biomes that failed this criteria (Table 1). 
+This analysis was run for variables with records for at least three distinct geographic areas in more than one biome, excluding any biomes that failed this criteria (Table 1). 
 When the effect of stand age was significant at p $\le$ 0.05 and when each biome had records for stands of at least 10 different ages, a biome * stand.age interaction was included in the model.

### 2020-09-28 manuscript/ERL_global_C_review.Rmd
incorporate SCP comments in intro

@@
 
-*Background.* The fate of Earth’s climate is closely linked to forests, which strongly influence atmospheric carbon dioxide (CO~2~) and climate through their influential role in the global carbon (C) cycle.
-Synthetic understanding of global forest C cycles is needed to constrain model estimates of forest feedbacks to climate change and to more accurately quantify the influence of land use decisions on climate.
+*Background.* Earth’s climate is closely linked to forests, which strongly influence atmospheric carbon dioxide (CO2) and climate by their impact  on the global carbon (C) cycle.
+However, efforts to incorporate forests into climate models and CO2 accounting frameworks have been constrained by a lack of accessible, global-scale data on how C cycling varies across forest types and stand ages.
 
-*Methods/Design.* Here, we draw from the Global Forest C Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation.
+*Methods/Design.* Here, we draw from the Global Forest Carbon Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation.
 Specifically, we use `r n_records` ForC records from `r n_areas_tot` geographic locations representing `r nrow(T1)` C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests.
@@
 
+Forest ecosystems are shaping the course of climate change (IPCC1.5) through their influence on atmospheric carbon dioxide (CO~2~).
+Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climate-carbon_2006; @krause_large_2018; @bonan_model_2019; @devittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history.
+These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published.
+While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [@li_mapping_2019; **REFS for biomass, biomass change, net CO2 flux**], measurement of most forest C stocks and fluxes require intensive on-the-ground data collection.
+Here, we provide a robust and comprehensive analysis of carbon cycling from a stand to global level, and by biome and stand age, using the largest global compilation of forest carbon data, which is available in our open source Global Carbon Forest database (ForC; Fig.
+1).
+
 *(Abby has offered to update stats in this paragraph:)*
 
-Forest ecosystems will play a critical role in shaping the course of climate change (IPCC1.5) through their influence on atmospheric carbon dioxide (CO~2~).
-Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**).
-While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion has been retained in ecosystems over recent decades.
-The resulting CO~2~ sink averaged 3.0 ± 0.8 Gt C yr^{-1} from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**).
-Moreover, forests contain substantial carbon (C) stocks: an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C [biomass and soils; @bonan_forests_2008].
-Forests are also globally dominant as a source of soil respiration [@warner_spatial_2019]. 
-Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), reducing the net forest sink to ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**).
-The future of the current forest C is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change.
+A robust understanding of forest impacts on global C cycling is essential. 
+Annual gross CO~2~ sequestration in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**).
+Most of this enormous C sequestration is counterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) *or fire*, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] *and fire emissions (REF)*.
+(**check if deforestation statistic below includes natural fire, exclude here if it does.**)
+In recent years, the remaining CO~2~ sink averaged 3.0 ± 0.8 Gt C yr^{-1} from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**).
+Yet, this sink is reduced by *deforestation/ forest losses to anthropogenic and natural disturbances*.
+Recent net deforestation (*i.e.*, gross deforestation minus regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), reducing the net forest sink to ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**).
+
+The future of the current forest C sink is dependent both upon forest responses to climate change itself and human land use decisions, which will feedback and strongly influence the course of climate change.
 Regrowing forests in particular will play an important role [@pugh_role_2019], as these represent a large (**~#%**) and growing proportion of Earth's forests [@mcdowell_pervasive_2020].
@@
 
-Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climate-carbon_2006; @krause_large_2018; @bonan_model_2019; @devittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history.
-These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published.
-While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [@li_mapping_2019; **REFS for biomass, biomass change, net CO2 flux**], measurement of most forest C stocks and fluxes require intensive on-the-ground data collection.
-Original studies typically cover only a small numbers of sites at a time, with rare exceptions spanning regions or continents [e.g., @lutz_global_2018; **FLUXNET_REF**), typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015] or FLUXNET [@baldocchi_fluxnet_2001].
+Despite the importance of forests, comprehensive global studies have historically been limited by the scattered and more local nature of research studies. 
+Primary research articles typically cover only a small numbers of sites at a time.
+The rare exceptions that span regions or continents with rare exceptions spanning regions or continents are typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; e.g., @lutz_global_2018] or FLUXNET [@baldocchi_fluxnet_2001; e.g., **FLUXNET_REF**].
 The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*.
-In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere (**REFS**).
+In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; **REFS**].
+Scattered data are not conducive for advancing science nor for making decisions about how to best manage our forests as a tool for constraining the climate crisis.
 
-To address the need for global-scale analyses of forest C cycling, we recently developed an open-access Global Forest Carbon database, ForC (@anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018).
-ForC contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables) based on ground-based measurements, along with associated data required for interpretation (*e.g.*, stand history, measurement methods).
+To address the need for global-scale analyses of forest C cycling, we recently developed ForC [@anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018].
+ForC contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables), with the different variables capturing the unique ecosystem pools (e.g., woody, foliage, and root biomass; dead wood) and flux types (e.g, gross and net primary productivity; soil, root, and ecosystem respiration).
+These data are ground-based measurements, and ForC contains associated data required for interpretation (*e.g.*, stand history, measurement methods).
 These data have been amalgamated from original peer-reviewed publications, either directly or via intermediary data compilations.
-Since the its most recent publication [@anderson-teixeira_forc_2018], ForC has been integrated with two large databases: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020], both of which have also synthesized published forest C data.
+Since the its most recent publication [@anderson-teixeira_forc_2018], ForC has grown to include two additional large databases: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020] that also synthesized published forest C data.
 Following these additions, ForC currently contains `r nrow(ForC_measurements)` records from `r nrow(ForC_plots)` plots and `r n_areas` distinct geographic areas representing all forested biogeographic and climate zones.
+This represents an `r round((nrow(ForC_measurements) - 17367)/17367*100, digits=0)`% increase in records from the prior publication [@anderson-teixeira_forc_2018].
 
-Here, we synthesize ForC data (Fig.
-1) to provide a macroscopic overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age.
-Our primary goal is to provide a data-based summary of our current state of knowledge on broad trends in forest C cycling.
-We address three broad questions:
+Here, we analyze the more extensive ForC data (Fig. 1) to provide a robust overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age.
+Our primary goal is to provide a data-driven summary of our current state of knowledge on broad trends in forest C cycling. 
+Specifically, we address three broad questions:
 
-1.
-To what extent can we fully represent, and "close", C budgets for each of the world's major forest biomes (*i.e.*, tropical, temperate broadleaf and deciduous, boreal) based on the current ForC data?
+1. How thoroughly can we represent C budgets for each of the world's major forest biomes (*i.e.*, tropical, temperate broadleaf and deciduous, boreal) based on the current ForC data?
 
@@
 
-While components of these questions have been previously addressed [@luyssaert_co2_2007; @anderson-teixeira_carbon_2016; @cook-patton_mapping_2020; @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and will serve as a foundation for improved understanding of global forest C cycling.
+While components of these questions have been previously addressed [@luyssaert_co2_2007; @anderson-teixeira_carbon_2016; @cook-patton_mapping_2020; @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and will serve as a foundation for improved understanding of global forest C cycling and highlight where key sources of uncertainty still reside.
 

### 2020-09-28 manuscript/ERL_global_C_review.Rmd
incorporating SC comments

@@
 - We're still finalizing the data (and adding some data).
-Outliers in plots will all be checked/ resolved, and we'll be able to pull in more data (GROA needs classification by dominant vegetation before it can be pulled in, and we're working on that)
+Outliers in plots will all be checked/ resolved, and we'll be able to pull in more data (*GROA* needs classification by dominant vegetation before it can be pulled in, and we're working on that)
 - "**???**" indicates a reference that we have entered in the .Rmd file, but not yet in the bibliography file. Don't worry about those. (However, places with "REF" need references)
@@
 
-*Background.* Earth’s climate is closely linked to forests, which strongly influence atmospheric carbon dioxide (CO2) and climate by their impact  on the global carbon (C) cycle.
+*Background.* Earth’s climate is closely linked to forests, which strongly influence atmospheric carbon dioxide (CO~2~) and climate by their impact  on the global carbon (C) cycle.
 However, efforts to incorporate forests into climate models and CO2 accounting frameworks have been constrained by a lack of accessible, global-scale data on how C cycling varies across forest types and stand ages.
@@
 *Methods/Design.* Here, we draw from the Global Forest Carbon Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation.
-Specifically, we use `r n_records` ForC records from `r n_areas_tot` geographic locations representing `r nrow(T1)` C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests.
+Specifically, we use `r n_records` *ForC* records from `r n_areas_tot` geographic locations representing `r nrow(T1)` C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests.
 For regrowth forests, we quantify age trends for all variables.
 
-*Review Results/ Synthesis.*  ForC v3.0 yielded a fairly comprehensive picture of C cycling in the world's major forest biomes, with broad closure of C budgets.
+*Review Results/ Synthesis.*  *ForC v3.0* yielded a fairly comprehensive picture of C cycling in the world's major forest biomes, with broad closure of C budgets.
 The rate of C cycling generally increased from boreal to tropical regions in both mature and regrowth forests, whereas C stocks showed less directional variation.
@@
 
-To address the need for global-scale analyses of forest C cycling, we recently developed ForC [@anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018].
-ForC contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables), with the different variables capturing the unique ecosystem pools (e.g., woody, foliage, and root biomass; dead wood) and flux types (e.g, gross and net primary productivity; soil, root, and ecosystem respiration).
-These data are ground-based measurements, and ForC contains associated data required for interpretation (*e.g.*, stand history, measurement methods).
+To address the need for global-scale analyses of forest C cycling, we recently developed *ForC* [@anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018].
+*ForC* contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables), with the different variables capturing the unique ecosystem pools (e.g., woody, foliage, and root biomass; dead wood) and flux types (e.g, gross and net primary productivity; soil, root, and ecosystem respiration).
+These data are ground-based measurements, and *ForC* contains associated data required for interpretation (*e.g.*, stand history, measurement methods).
 These data have been amalgamated from original peer-reviewed publications, either directly or via intermediary data compilations.
-Since the its most recent publication [@anderson-teixeira_forc_2018], ForC has grown to include two additional large databases: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020] that also synthesized published forest C data.
-Following these additions, ForC currently contains `r nrow(ForC_measurements)` records from `r nrow(ForC_plots)` plots and `r n_areas` distinct geographic areas representing all forested biogeographic and climate zones.
+Since the its most recent publication [@anderson-teixeira_forc_2018], *ForC* has grown to include two additional large databases: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020] that also synthesized published forest C data.
+Following these additions, *ForC* currently contains `r nrow(ForC_measurements)` records from `r nrow(ForC_plots)` plots and `r n_areas` distinct geographic areas representing all forested biogeographic and climate zones.
 This represents an `r round((nrow(ForC_measurements) - 17367)/17367*100, digits=0)`% increase in records from the prior publication [@anderson-teixeira_forc_2018].
 
-Here, we analyze the more extensive ForC data (Fig. 1) to provide a robust overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age.
+Here, we analyze the more extensive *ForC* data (Fig. 1) to provide a robust overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age.
 Our primary goal is to provide a data-driven summary of our current state of knowledge on broad trends in forest C cycling. 
@@
 
-1. How thoroughly can we represent C budgets for each of the world's major forest biomes (*i.e.*, tropical, temperate broadleaf and deciduous, boreal) based on the current ForC data?
+1. How thoroughly can we represent C budgets for each of the world's major forest biomes (*i.e.*, tropical, temperate broadleaf and deciduous, boreal) based on the current *ForC* data?
 
@@
 
-This review synthesizes data from the ForC database [Fig. 1; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018].
-ForC amalgamates numerous intermediary data sets (*e.g.*, **REFS**) and original studies.
+This review synthesizes data from the *ForC* database [Fig. 1; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018].
+*ForC* amalgamates numerous intermediary data sets [*e.g.*, @luyssaert_co2_2007; @bond-lamberty_global_2010; @cook-patton_mapping_2020] and original studies.
 Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records.
@@
 As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible [83% of records in v.2.0; @anderson-teixeira_forc_2018].
-Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016].
-Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##, `r n_SRDB` records; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database [GROA v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020].
+Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], which represented roughly one-third of records in *ForC* v.2.0* [@anderson-teixeira_forc_2018].
+Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [*SRDB* v.##, `r n_SRDB` records; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database [*GROA* v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020].
 We've also added data from individual publications (detailed list at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv), with a particular focus on productivity [@banbury_morgan_global_nodate; e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @lutz_global_2018, @johnson_climate_2018].
@@
 
-Analyses drew from ForC-simplified (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which is a rearrangement of ForC intended to facilitate analyses.
-In generating ForC-simplified, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006).
+To facilitate analyses, we created a simplified version of ForC, *ForC-simplified* (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which we analyzed here.
+In generating *ForC-simplified*, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006).
 Duplicate or otherwise conflicting records were reconciled as described in *APPENDIX S1*, resulting in a total of `r nrow(ForC_simplified)` records (`r round(nrow(ForC_simplified)/nrow(ForC_measurements)*100,digits=1)`% size of total database).
@@
 This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.
-Plots flagged as disturbed in ForC-simplified (`r round(n_disturbed_in_ForC_simplified/nrow(ForC_simplified)*100,digits=1)`%) included stands that had undergone anthropogenic thinning or partial harvest unless this was very minor.
+Plots flagged as disturbed in ForC-simplified (`r round(n_disturbed_in_ForC_simplified/nrow(ForC_simplified)*100,digits=1)`%) included stands that had undergone any notable anthropogenic thinning or partial harvest.
 We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods.
@@
 
-We selected 23 annual flux and 11 C stock variables for inclusion in the analysis, although two flux variables ($R_{het-ag}$ and $R_{het}$) were included for conceptual completeness but had no records in ForC (Table 1).
+We selected 23 annual flux and 11 C stock variables for inclusion in the analysis, although two flux variables ($R_{het-ag}$ and $R_{het}$) were included for conceptual completeness but had no records in *ForC* (Table 1).
 Records for these variables represented `r round(n_focal_variable_records_eligible_for_inclusion/records_eligible_for_inclusion*100,digits=1)`% of the total records eligible for inclusion.
@@
 Specifically, net ecosystem exchange (measured by eddy-covariance; **FLUXNET REF**) and biometric estimates of $NEP$ were combined into the single variable $NEP$ (Table 1).
-Furthermore, for $NPP$, $ANPP$, and $ANPP_{litterfall}$, ForC variables specifying inclusion of different components were combined.
+Furthermore, for $NPP$, $ANPP$, and $ANPP_{litterfall}$, *ForC* variables specifying inclusion of different components were combined.
 Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum threshold for tree census was $\le$ 10cm. 
@@
 Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
-For young tropical forests imported from GROA but not yet classified by leaf type, we assumed that all were broadleaf, consistent with the rarity of naturally regenerating needleleaf forests in the tropics (**REF**).
+For young tropical forests imported from *GROA* but not yet classified by leaf type, we assumed that all were broadleaf, consistent with the rarity of naturally regenerating needleleaf forests in the tropics (**REF**).
 Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication).
-Assigning stands to these groupings required the exclusion of records for which ForC lacked records of stand age (`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database), or geographic coordinates (`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database).
+Assigning stands to these groupings required the exclusion of records for which *ForC* lacked records of stand age (`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database), or geographic coordinates (`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database).
 We also excluded records of stand age = 0 year (`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
@@
 
-Of the `r nrow(ForC_measurements)` records in ForC v3.0, `r n_records` met the criteria for inclusion in this study (Fig. 1).
+Of the `r nrow(ForC_measurements)` records in *ForC* v3.0, `r n_records` met the criteria for inclusion in this study (Fig. 1).
 These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas.
-Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained mature forest estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests.
-For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.
+Of the 23 flux and 11 stock variables mapped in these diagrams, *ForC* contained mature forest estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests.
+For regrowth forests (<100 yrs), *ForC* contained estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.
 **([issue 29](https://github.com/forc-db/ERL-review/issues/29#issuecomment-699633801))**
@@
 
-Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5 (and available in tabular format in the ForC release accompanying this publication:  `ForC/numbers_and_facts/ForC_variable_averages_per_Biome.csv`).
+Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5 (and available in tabular format in the *ForC* release accompanying this publication:  `ForC/numbers_and_facts/ForC_variable_averages_per_Biome.csv`).
 
@@
 
-ForC contained *14* flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig. 6-7 and **S#- SI figures including plots for all variables**).
+*ForC* contained *14* flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig. 6-7 and **S#- SI figures including plots for all variables**).
 Of these, *9* increased significantly with log10[stand.age]: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($NEP$).
@@
 
-ForC v3.0 brought together an unprecedented amount of data to yield an internally consistent picture of C cycling in the world's major forest biomes.
+*ForC* v3.0 brought together an unprecedented amount of data to yield an internally consistent picture of C cycling in the world's major forest biomes.
 Carbon cycling rates generally increased from boreal to tropical regions and with stand age.
@@
 
-ForC provides provides unprecedented coverage of most major variables.
+*ForC* provides provides unprecedented coverage of most major variables.
 *(discuss  how this improves upon previous data compilations/ for which variables does ForC make the greatest difference (e.g., not AGB or NEP/GPP/Reco, but by far the latest data compilation for dead wood, [variables])*
@@
 *(give some stats/ cite figures)*.
-ForC does not include soil carbon, which is covered by other efforts (REFS).
-For fluxes, Fluxnet is the keeper of the best data on NEE, GPP, Reco (REFS), and SRDB remains the authority on soil respiration (REFS).
-ForC includes recent data from both, but is not continuously integrated.
+*ForC* does not include soil carbon, which is covered by other efforts (REFS).
+For fluxes, Fluxnet is the keeper of the best data on NEE, GPP, Reco (REFS), and *SRDB* remains the authority on soil respiration (REFS).
+*ForC* includes recent data from both, but is not continuously integrated.
 For C is the best source for most of the subsidiary fluxes: NPP, woody mortality...
@@
 4) than of methodological accuracy.*
-Thus, overall, a high degree of closure implies that ForC gives a consistent picture of C cycling within biomes.
+Thus, overall, a high degree of closure implies that *ForC* gives a consistent picture of C cycling within biomes.
 While these means are unlikely to be accurate representations of C cycling within any particular forest, they provide a useful baseline for comparison, always keeping in mind that sample means do not necessarily represent the true mean of the entire biome.
@@
 The future of forest C cycling [@song_meta-analysis_2019] will shape trends in atmospheric CO~2~ and the course of climate change.
-For a human society seeking to understand and mitigate climate change, the data contained in ForC and summarized here can help to meet two major challenges.
+For a human society seeking to understand and mitigate climate change, the data contained in *ForC* and summarized here can help to meet two major challenges.
 
@@
 To ensure that models are giving the right answers for the right reasons, it is important to benchmark against multiple components of the C cycle that are internally consistent with each other.
-By making tens of thousands of records readily available in standardized format, ForC makes it feasible for the modeling community to draw upon these data to benchmark models.
-Integration of ForC with models is a goal (Fer et al., in revision).
+By making tens of thousands of records readily available in standardized format, *ForC* makes it feasible for the modeling community to draw upon these data to benchmark models.
+Integration of *ForC* with models is a goal (Fer et al., in revision).
 
-Second, ForC can serve as a pipeline through which forest science can inform forest-based climate change mitigation efforts.
+Second, *ForC* can serve as a pipeline through which forest science can inform forest-based climate change mitigation efforts.
 Such efforts will be most effective when informed by the best available data, yet it is not feasible for the individuals and organizations designing such efforts to sort through literature, often behind paywalls, with data reported in varying units, terminology, etc.
-One goal for ForC is to serve as a pipeline through which information can flow efficiently from forest researchers to decision-makers working to implement forest conservation strategies at global, national, or landscape scales.
+One goal for *ForC* is to serve as a pipeline through which information can flow efficiently from forest researchers to decision-makers working to implement forest conservation strategies at global, national, or landscape scales.
 This is already happening!
-ForC has already contributed to updating the IPCC guidelines for carbon accounting in forests [IPCC 2019; @requena_suarez_estimating_2019; Rozendaal et al in prep], mapping C accumulation potential from natural forest regrowth globally [@cook-patton_mapping_2020], and informing ecosystem conservation priorities [@goldstein_protecting_2020].
+*ForC* has already contributed to updating the IPCC guidelines for carbon accounting in forests [IPCC 2019; @requena_suarez_estimating_2019; Rozendaal et al in prep], mapping C accumulation potential from natural forest regrowth globally [@cook-patton_mapping_2020], and informing ecosystem conservation priorities [@goldstein_protecting_2020].
 
-**ForC can complement remote sensing to provide a comprehensive picture of global forest C cycling.** (*here, discuss what is well-covered by remote sesnsing, where ForC is useful for calibrating remote sensing, and what ForC can get that remote sensing can't.*)
+***ForC* can complement remote sensing to provide a comprehensive picture of global forest C cycling.** (*here, discuss what is well-covered by remote sesnsing, where ForC is useful for calibrating remote sensing, and what ForC can get that remote sensing can't.*)
 AGB is the largest stock, and most of the emphasis is on this variable.
@@
 However, it is limited in that it is not associated with stand age and disturbance history, except in recent decades when satellite data can be used to detect forest loss, gain, and some of their dominant drivers [@hansen_high-resolution_2013; @song_global_2018; @curtis_classifying_2018].
-ForC  is therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020.
+*ForC*  is therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020.
 
@@
 **Move to data availability statement, or methods?**:
-We recommend that use of ForC data go to the original database, as opposed to using "off-the-shelf" values from this publication.
+We recommend that use of *ForC* data go to the original database, as opposed to using "off-the-shelf" values from this publication.
 This is because
-(1) ForC is constantly being updated,
+(1) *ForC* is constantly being updated,
 (2) analyses should be designed to match the application,
@@
 ## Acknowledgements
-All researchers whose data is included in ForC and this analysis.
+All researchers whose data is included in *ForC* and this analysis.
 Ian McGregor for help with the database.
@@
 Materials required to fully reproduce these analyses, including data, R scripts, and image files, are archived in Zenodo (DOI: TBD].
-Data, scripts, and results presented here are also available through the open-access ForC GitHub repository (https://github.com/forc-db/ForC), where many will be updated as the database develops.
+Data, scripts, and results presented here are also available through the open-access *ForC* GitHub repository (https://github.com/forc-db/ForC), where many will be updated as the database develops.
 

### 2020-09-28 manuscript/ERL_global_C_review.Rmd
integrate SCP comments- methods section

@@
 
-We selected 23 annual flux and 11 C stock variables for inclusion in the analysis, although two flux variables ($R_{het-ag}$ and $R_{het}$) were included for conceptual completeness but had no records in *ForC* (Table 1).
+We selected 23 annual flux and 11 C stock variables for inclusion in the analysis.
+These different flux and stock variables (Table 1) represent different pools (e.g., aboveground biomass, dead wood, ecosystem total) and levels of combination (e.g., total aboveground net primary productivity ($ANPP$) versus the ANPP of individual elements such as foliage, roots, and branches.
+Note that two flux variables, aboveground heterotropic ($R_{het-ag}$) and total ($R_{het}$) respiration, were included for conceptual completeness but had no records in *ForC* (Table 1).
 Records for these variables represented `r round(n_focal_variable_records_eligible_for_inclusion/records_eligible_for_inclusion*100,digits=1)`% of the total records eligible for inclusion.
-For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables.
+For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity, such as measurements including or excluding fruit and flower production and herbivory) into more broadly defined variables.
 Specifically, net ecosystem exchange (measured by eddy-covariance; **FLUXNET REF**) and biometric estimates of $NEP$ were combined into the single variable $NEP$ (Table 1).
 Furthermore, for $NPP$, $ANPP$, and $ANPP_{litterfall}$, *ForC* variables specifying inclusion of different components were combined.
-Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum threshold for tree census was $\le$ 10cm. 
+Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum diameter breast height (DBH) threshold for tree census was $\le$ 10cm. 
 All records were measured directly or derived from field measurements (as opposed to modeled).
@@
 
-For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
+For this analysis, we grouped forests into four broad biome types based on climate zones and dominant vegetation type (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
 Climate zones (Fig. 1) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
-Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd).
+We defined the tropical biome as including all equatorial (A) zones, temperate biomes as including all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and the boreal biome as including the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd).
 Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis.
-Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
+We defined leaf type (broadleaf / needleleaf) was based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
 For young tropical forests imported from *GROA* but not yet classified by leaf type, we assumed that all were broadleaf, consistent with the rarity of naturally regenerating needleleaf forests in the tropics (**REF**).
-Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication).
-Assigning stands to these groupings required the exclusion of records for which *ForC* lacked records of stand age (`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database), or geographic coordinates (`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database).
+We also classified forests as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication).
+Assigning stands to these groupings required the exclusion of records for which *ForC* lacked geographic coordinates (`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database) or records of stand age (`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
 We also excluded records of stand age = 0 year (`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
@@
 
-Data were summarized to produce schematics of C cycling across the eight biome - age group combinations identified above.
+Data were summarized to produce schematics of C cycling across the eight biome by age group combinations identified above.
 To obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot.
@@
 
-To test for differences across mature forest biomes, the original values were analyzed via a linear mixed effects model ['lmer' function in 'lme4' R package version **x.xx**; **REF**] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
+To test for differences across mature forest biomes, we also examined how stand age impacted fluxes and stocks, employing a mixed effects model ['lmer' function in 'lme4' R package version **x.xx**; **REF**] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
 When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
@@
 This analysis was run for variables with records for at least three distinct geographic areas in more than one biome, excluding any biomes that failed this criteria (Table 1). 
-When the effect of stand age was significant at p $\le$ 0.05 and when each biome had records for stands of at least 10 different ages, a biome * stand.age interaction was included in the model.
-All database manipulation, analyses, and figure production were fully automated in R (**version, citation**).
+When the effect of stand age was significant at p $\le$ 0.05 and when each biome had records for stands of at least 10 different ages, a biome $\times$ stand.age interaction was included in the model.
+
+To facilitate the accessibility of our results and data, and to allow for rapid updates as additional data become available, we have automated all database manipulation, analyses, and figure production in R (**version, citation**).
 

### 2020-09-28 manuscript/ERL_global_C_review.Rmd
fixed results in Data coverage section

@@
 
-idx_fluxes <- ForC_biome_averages$n.areas >= 7 & ForC_biome_averages$variable.type %in% "flux"
-idx_stocks <- ForC_biome_averages$n.areas >= 7 & ForC_biome_averages$variable.type %in% "stock"
+idx_fluxes7 <- ForC_biome_averages$n.areas >= 7 & ForC_biome_averages$variable.type %in% "flux"
+idx_stocks7 <- ForC_biome_averages$n.areas >= 7 & ForC_biome_averages$variable.type %in% "stock"
+idx_fluxes3 <- ForC_biome_averages$n.areas >= 3 & ForC_biome_averages$variable.type %in% "flux"
+idx_stocks3 <- ForC_biome_averages$n.areas >= 3 & ForC_biome_averages$variable.type %in% "stock"
 
-n_fluxes_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_fluxes])
-n_stocks_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_stocks])
+n_fluxes_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_fluxes7])  
+n_stocks_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_stocks7])  
+n_fluxes_with_ge3_areas <- table(ForC_biome_averages$Biome[idx_fluxes3])  
+n_stocks_with_ge3_areas <- table(ForC_biome_averages$Biome[idx_stocks3])  
 ```
@@
 
-Of the `r nrow(ForC_measurements)` records in *ForC* v3.0, `r n_records` met the criteria for inclusion in this study (Fig. 1).
+Of the `r nrow(ForC_measurements)` records in *ForC* v3.0, `r n_records` met our strict criteria for inclusion in this study (Fig. 1).
 These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas.
-Of the 23 flux and 11 stock variables mapped in these diagrams, *ForC* contained mature forest estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests.
-For regrowth forests (<100 yrs), *ForC* contained estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.
-**([issue 29](https://github.com/forc-db/ERL-review/issues/29#issuecomment-699633801))**
+Of the 23 flux and 11 stock variables mapped in these diagrams, *ForC* contained sufficient mature forest data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ 7 distinct geographic areas) for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests.
+For regrowth forests (<100 yrs), *ForC* contained sufficient data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ **3** distinct geographic areas) for `r n_fluxes_with_ge3_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge3_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.
 

### 2020-09-28 manuscript/ERL_global_C_SI.Rmd

@@
 For each group of duplicate records—*i.e.*, measurements of the same variable in the same plot at the same time—one record was assigned precedence (recorded in D.precedence field). 
-When duplicates were exact matches, **one was dropped**. 
+If duplicates were exact matches, **one was dropped**. 
 When measurement periods overlapped or were not specified, precedence was given first to records representing longer measurement periods (*i.e.*, end.date - start.date) and then to more recently published values.
@@
 Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
-Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.", out.width='100%', fig.pos='H'}
+Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
+Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).", out.width='100%', fig.pos='H'}
 

### 2020-09-28 manuscript/ERL_global_C_review.Rmd

@@
 To obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot.
-Values were then averaged across plots clustered within 25 km of one another [*sensu* @anderson-teixeira_forc_2018], weighting by area sampled if available for all records.
+Values were then averaged across geographically distinct areas, defined as plots clustered within 25 km of one another [*sensu* @anderson-teixeira_forc_2018], weighting by area sampled if available for all records.
 This step was taken to avoid pseudo-replication and to combine any records from sites with more than one name in ForC.
+
+We tested whether the C budgets described above "closed"--*i.e.*, whether they were internally consistent.
+Specifically, we first defined relationships among variables: for example, $NEP =	GPP - R_{eco}$, $BNPP = BNPP_{coarse} + BNPP_{fine}$, $DW_{tot} = DW_{standing} + DW_{down}$).
+**([issue #44](https://github.com/forc-db/ERL-review/issues/44)--but just delete this chunk if not resolved before submission)**
+Henceforth, we refer to the variables on the left side of the equation as "aggregate" fluxes or stocks, and those that are summed as "component" fluxes or stocks, noting that the same variable can take both aggregate and component positions in different relationships.
+We considered the C budget for a given relationship "closed" when component variables summed to within one standard deviation of the aggregate variable.
 
@@
 
-Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5 (and available in tabular format in the *ForC* release accompanying this publication:  `ForC/numbers_and_facts/ForC_variable_averages_per_Biome.csv`).
+Average C cycles for mature tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major natural or anthropogenic disturbance are presented in Figures 2-5 (and available in tabular format in the *ForC* release accompanying this publication:  `ForC/numbers_and_facts/ForC_variable_averages_per_Biome.csv`).
 
 For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets were generally consistent.
-*That is, component fluxes and stocks summed to within one standard deviation of more inclusive fluxes in all but* **two** *instances, both in temperate conifer forests (Fig. 5).* 
-**(check this with final results: https://github.com/forc-db/ForC/blob/master/numbers_and_facts/C_cycle_closure.csv) ** 
+*That is, component variables summed to within one standard deviation of their respective aggregate variables in all but* **two** *instances, both in temperate conifer forests (Fig. 5).* 
+**(check all this with final results: https://github.com/forc-db/ForC/blob/master/numbers_and_facts/C_cycle_closure.csv, and be sure to write out variable names at first occurance, and provide a bit more detail) **
+*For the temperate conifer biome, the average composite measure of aboveground biomass (Bag) was less than the combined average value of woody biomass (Bag-wood) and foliage biomass (Bfoliage), partly due to the very high estimates of Bag-wood.*
 *Within this biome, $B_{ag}<B_{ag-wood}+B_{foliage}$ and $B_{root}<B_{root-coarse}+B_{root-fine}$ because $B_{ag-wood}$ and $B_{root-coarse}$ were very high, with strongly disproportionate numbers of records from the high-biomass forests of the US Pacific Northwest (Figs.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 2 | C cycle diagram for mature tropical broadleaf forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.
 Asterisk after variable name indicates lack of C cycle closure.", out.width='100%', fig.pos='H'}
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 3 | C cycle diagram for mature temperate broadleaf forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.
 Asterisk after variable name indicates lack of C cycle closure.", out.width='100%', fig.pos='H'}
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 4 | C cycle diagram for mature temperate conifer forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.
 Asterisk after variable name indicates lack of C cycle closure.", out.width='100%', fig.pos='H'}
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 5 | C cycle diagram for mature boreal conifer forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.
 Asterisk after variable name indicates lack of C cycle closure.", out.width='100%', fig.pos='H'}
@@
 There were sufficient data to assess mature forest biome differences for **20** flux variables, and significant differences among biomes were detected for **14** variables (Table 1).
-With only **one** exception ($R_{auto}$, which had low sample size; **[issue 40](https://github.com/forc-db/ERL-review/issues/40)**), C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs.
+In all cases--including C fluxes into, within, and out of the ecosystem--C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs.
 6, S1-S15).
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | C cycle diagram for young tropical broadleaf forests.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
 Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
-Arrows indicate fluxes, boxes indicate stocks.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
 
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 9 | C cycle diagram for young temperate broadleaf forests.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
 Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
-Arrows indicate fluxes, boxes indicate stocks.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
 
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 10 | C cycle diagram for young temperate conifer forests.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
 Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
-Arrows indicate fluxes, boxes indicate stocks.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
 
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 11 | C cycle diagram for young boreal conifer forests.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
 Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
-Arrows indicate fluxes, boxes indicate stocks.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
 

### 2020-09-29 manuscript/ERL_global_C_SI.Rmd
SCP changes- results

@@
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
-Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
-Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
+In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+The fitted line indicates the effect of age on flux (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant age x biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.

### 2020-09-29 manuscript/ERL_global_C_review.Rmd
SCP changes- results

@@
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
-Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
-Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
-Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.
+In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+The fitted line indicates the effect of age on flux (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant age x biome interaction.
+The boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
 Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).
-Individual figures for each flux with sufficient data given in the Supplement (Figs.
+Individual figures for each flux with sufficient data are given in the Supplement (Figs.
 S1-S15).", out.width='100%', fig.pos='H'}
@@
 C stocks had less consistent patterns across biomes (Figs. 7, S16-S26).
-In **four of the six** cases (**$B_{tot}$, $B_{ag-wood}$,  $B_{root-coarse}$, $DW_{tot}$**), temperate conifer forests were in the highest significance grouping, and boreal forests in the lowest, with tropical and temperate broadleaf forests  in between--most commonly being significantly different from temperate conifer but not boreal forests.
+In **four of the six** cases (**$B_{tot}$, $B_{ag-wood}$,  $B_{root-coarse}$, $DW_{tot}$**), temperate conifer forests had significantly higher stocks than the other biomes, and boreal forests in the lowest, with tropical and temperate broadleaf forests in between.
 For $B_{ag}$, which had by far the highest sample size, tropical forests exceeded temperate conifer forests (but not significantly). 
@@
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
-Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
-Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
-Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.
+In each panel, the left scatterplot  shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+The fitted line indicates the effect of age on flux (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant age x biome interaction.
+The boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.
 Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).
-Individual figures for each stock with sufficient data given in the Supplement (Figs.
+Individual figures for each stock with sufficient data are given in the Supplement (Figs.
 S16-S26).", out.width='100%', fig.pos='H'}
@@
 Average C cycles for forests <100 years old are presented in Figures 8-11.  
-Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Tables 1, S2, Figs. 6-11, S1-S26; detailed below).
+Both C stocks and fluxes commonly increased signficantly with stand age (Tables 1, S2, Figs. 6-11, S1-S26; detailed below).
 
 *ForC* contained *14* flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig. 6-7 and **S#- SI figures including plots for all variables**).
-Of these, *9* increased significantly with log10[stand.age]: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($NEP$).
+Of these, *9* increased significantly with age: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($NEP$).
 The remaining five--$ANPP_{woody-branch}$, $BNPP_{root-coarse}$, $R_{soil-het}$, and $R_{soil-het}$-–displayed no significant relationship to stand age, although all displayed a positive trend.
@@
 
+(discuss NEP well- intuition for some would be that NEP is higher in tropics.
+Note biome differences for young mut not mature forests, link to latitudinal differences in C sequestration potential (Anderson 2006, @cook-patton_mapping_2020)
+
 ### Age trends in C cycling
@@
 
+(discuss NEP well, including)
 NEP increases with log(age) to 100 --> strongest C sinks are established secondary forests. (But presumably this exact number is an artifact; don’t over-emphasize.)

### 2020-09-29 manuscript/ERL_global_C_review.Rmd
finish SCP changes
incorporate Ben's discussion

@@
 
-*ForC* v3.0 brought together an unprecedented amount of data to yield an internally consistent picture of C cycling in the world's major forest biomes.
+*ForC* v3.0 provided unprecedented coverage of most major variables, yielding an internally consistent picture of C cycling in the world's major forest biomes.
 Carbon cycling rates generally increased from boreal to tropical regions and with stand age.
@@
 
-*ForC* provides provides unprecedented coverage of most major variables.
-*(discuss  how this improves upon previous data compilations/ for which variables does ForC make the greatest difference (e.g., not AGB or NEP/GPP/Reco, but by far the latest data compilation for dead wood, [variables])*
-*(Noteable holes include: fluxes: R_auto_ag, woody mortality, folivary/ herbivory and respiration of herbivores (and therefore total R_het), ANPP_repro; also fluxes in tropical regrowth forests)*
-*For the C stocks considered here, the most poorly covered is dead wood (none in E hemisphere!), despite a focused effort on this variable that has resulted in ForC being by far the largest collection of these data.*
-Thus, overall, we're lacking coverage of fluxes to herbivores and higher consumers, along with the woody mortality and dead wood.
-Geographically, all variables poorly covered in Africa and Siberia.
+The large number of C cycle variables covered by ForC, and the general consistency between them, provide confidence that our overall reported means provide accurate and useful baselines for analysis (with the caveats that they are unlikely to be accurate representations of C cycling for any particular forest, and that these sample means almost certainly do not represent true biome means).
 
-**notes from Ben on the above par:**
-Pregitzer and Euskirchen 2004 http://dx.doi.org/10.1111/j.1365-2486.2004.00866.x "Aggregated biome-level estimates of NPP and NEP were higher in intermediate-aged forests (e.g., 30–120 years), while older forests (e.g., 4120 years) were generally less productive.
-The mean NEP in the youngest forests (0–10 years) was negative (source to the atmosphere) in both boreal and temperate biomes…Forest age is a highly significant source of variability in NEP at the biome scale”
+There are of course notable holes in the ForC variable coverage, as discussed by Anderson-Teixeira et al.
+(xxxx), that limit the scope of our inferences here.
+Notably, ForC lacks coverage of fluxes to herbivores and higher consumers, along with the woody mortality and dead wood stocks.
+Geographically, all variables are poorly covered in Africa and Siberia, a common problem in the carbon-cycle community (Xu and Shang 2016 10.1016/j.jplph.2016.08.007, Schimel et al.
+2015 10.1073/pnas.1407302112).
+ForC does not include soil carbon, which is covered by other efforts (e.g. Köchy et al. 2015 10.5194/soil-1-351-2015). 
+ForC is not intended to replace databases that are specialized for particular parts of the C cycle analyses, e.g., aboveground biomass (REFS), land-atmosphere fluxes (Baldocchi et al.
+2001 10.1175/1520-0477(2001)082<2415:FANTTS>2.3.CO;2), soil respiration (Jian et al.
+2020 10.5194/essd-2020-136), or the human footprint in global forests (Magnani et al.
+2007 10.1038/nature05847).
 
-Amiro et al. 2010 http://dx.doi.org/10.1029/2010JG001390
-Houghton et al. 2020 https://doi.org/10.1111/gcb.15050
-
-Turnover:
-Pugh et al. 2020 http://dx.doi.org/10.5194/bg-2019-491
-Yu et al. 2019 http://dx.doi.org/10.1073/pnas.1821387116
-
-Human footprint in global forests
-http://dx.doi.org/10.1038/nature05847
-http://dx.doi.org/10.1038/nature02619
-
-Mention consistency (or lack of) with e.g. GOLUM-CNP?
-http://dx.doi.org/10.5194/gmd-11-3903-2018
-
-In terms of C stocks, there is a paucity of data on dead wood and organic layer (Pan et al. ?).
-These can be significant.
-(**Note that we've given a lot of emphasis to dead wood (work by Abby, and also Jenny), and as a result this work really advances knowledge of dead wood.
-We'll want to highlight that here.**)
-*(give some stats/ cite figures)*.
-*ForC* does not include soil carbon, which is covered by other efforts (REFS).
-For fluxes, Fluxnet is the keeper of the best data on NEE, GPP, Reco (REFS), and *SRDB* remains the authority on soil respiration (REFS).
-*ForC* includes recent data from both, but is not continuously integrated.
-For C is the best source for most of the subsidiary fluxes: NPP, woody mortality...
-
-The C cycle budgets for mature forests (Figs. 2-5) generally "close"--that is, the sums of component variables do not differ from the larger fluxes by more than one standard deviation.
-However, standard deviations are often large, reflective of significant within-biome variation. This makes the standard for closure relatively loose.
-*Lack of closure, in the few instances where it occurs, is probably more reflective of differences in the representation of forest types (e.g., disproportionate representation of US Pacific NW for aboveground woody biomass relative to AGB; Fig.
-4) than of methodological accuracy.*
-Thus, overall, a high degree of closure implies that *ForC* gives a consistent picture of C cycling within biomes.
-While these means are unlikely to be accurate representations of C cycling within any particular forest, they provide a useful baseline for comparison, always keeping in mind that sample means do not necessarily represent the true mean of the entire biome.
+In this analysis, the C cycle budgets for mature forests (Figs.
+2-5) generally “close”–that is, the sums of component variables do not differ from the larger fluxes by more than one standard deviation.
+On the one hand, this reflects the general fact that ecosystem-scale measurements tend to close the C budget more easily and consistently than, for example, for energy balance (Stoy et al.
+2013 10.1016/j.agrformet.2012.11.004).
+On the other, however, as noted above ForC derives data from multiple heterogeneous sources, often with large errors (standard deviations); as a result, the standard for C closure is relatively loose (cf Houghton 2020 10.1111/gcb.15050).
+Nonetheless, the lack of closure, in the few instances where it occurs, is probably more reflective of differences in the representation of forest types (e.g., disproportionate representation of US Pacific NW for aboveground woody biomass relative to AGB; Fig.
+4) than of methodological accuracy.
+The overall high degree of closure implies that ForC gives a consistent picture of C cycling within biomes. 
+This is an important and useful test, because it allows for consistency checks within the C cycle, for example leveraging separate and independently-measured fluxes to constrain errors in another (Phillips et al.
+2017 10.1007/s11104-016-3084-x, Williams et al.
+2014 10.1016/j.rse.2013.10.034, Harmon et al.
+2011 10.1029/2010JG001495), or producing internally consistent global data products (Wang et al.
+2018 10.5194/gmd-11-3903-2018).
 
@@
 In contrast to C fluxes and accumulation rates in regrowth forests, stocks...
+Patterns are consistent with others studies.
+Highest biomass forests are found in temperate oceanic climates (REF- Keith, something in GEB, some global forest C map).
 
@@
 NEP increases with log(age) to 100 --> strongest C sinks are established secondary forests. (But presumably this exact number is an artifact; don’t over-emphasize.)
+
+Our findings are largely consistent with, but built from a far larger dataset than, those of Pregitzer and Euskirchen (2004 http://dx.doi.org/10.1111/j.1365-2486.2004.00866.x), who found that NPP and NEP to be higher in intermediate-aged forests than older forests, and emphasize the importance of forest age at the biome scale.
+Quickly-changing and age-dependent fluxes were also found in a number of previous syntheses (Amiro et al.
+2010 10.1029/2010JG001390, Magnani et al.
+2007 10.1038/nature05847).
 
@@
 (**Ben, it woudl be particularly helpful if you could flesh this out some more.**)
+
+In terms of C stocks, there is a paucity of data on dead wood and organic layer (Pan et al. ?).
+These can be significant.
+(**Note that we've given a lot of emphasis to dead wood (work by Abby, and also Jenny), and as a result this work really advances knowledge of dead wood.
+We'll want to highlight that here.**)
+*(give some stats/ cite figures)*.
 

### 2020-09-29 manuscript/ERL_global_C_review.Rmd
incorporate Abby's stats in intro

@@
 
-*(Abby has offered to update stats in this paragraph:)*
-
 A robust understanding of forest impacts on global C cycling is essential. 
-Annual gross CO~2~ sequestration in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**).
+Annual gross CO~2~ sequestration in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
 Most of this enormous C sequestration is counterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) *or fire*, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] *and fire emissions (REF)*.
 (**check if deforestation statistic below includes natural fire, exclude here if it does.**)
-In recent years, the remaining CO~2~ sink averaged 3.0 ± 0.8 Gt C yr^{-1} from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**).
+In recent years, the remaining CO~2~ sink averaged 3.2 $\pm$ 0.6 GtC yr^−1^ from 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
 Yet, this sink is reduced by *deforestation/ forest losses to anthropogenic and natural disturbances*.
-Recent net deforestation (*i.e.*, gross deforestation minus regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), reducing the net forest sink to ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**).
+Recent net deforestation (*i.e.*, gross deforestation minus regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr^-1^ from YEAR-YEAR (Pan et al 2011; **UPDATE**), reducing the net forest sink to ~1.1-2.2 Gt C yr^-1^ across Earth's forests [@friedlingstein_global_2019].
 

### 2020-09-29 manuscript/ERL_global_C_review.Rmd
working on discussion

@@
 These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published.
-While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [@li_mapping_2019; **REFS for biomass, biomass change, net CO2 flux**], measurement of most forest C stocks and fluxes require intensive on-the-ground data collection.
+While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass: @hu_mapping_2016; **REF**; gross primary productivity, $GPP$: @li_mapping_2019; **REFS for biomass change, net CO2 flux**], measurement of most forest C stocks and fluxes require intensive on-the-ground data collection.
 Here, we provide a robust and comprehensive analysis of carbon cycling from a stand to global level, and by biome and stand age, using the largest global compilation of forest carbon data, which is available in our open source Global Carbon Forest database (ForC; Fig.
@@
 Our analysis reveals a general acceleration of carbon cycling from the tropics to the high latitudes.
-For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude [e.g.,  @banbury_morgan_global_nodate].
-For regrowth forests, more rapid accumulation of biomass at lower latitudes has been well-established [@anderson_temperature-dependence_2006; @cook-patton_mapping_2020], whereas this is the first study to compare age trends in deadwood and organic layer across biomes [but see @cook-patton_mapping_2020].
-For most C flux variables, this analysis is the first to examine flux trends in regrowth forests across biomes  (i.e., age x biome interaction).
+For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude (or increase with temperature) on a global scale [e.g., @luyssaert_co_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
+The consistency with which this occurs across numerous fluxes is not surprising, but has never been simultaneously assessed across such a large number of variables [but see @banbury_morgan_global_nodate for nine autotrophic fluxes].
+Thus, our analysis reveals that carbon cycling is most rapid in the tropics and slowest in boreal regions, including C fluxes into ($GPP$), within (e.g., $NPP$ and its components), and out of (e.g., $R_{soil}$, $R_{eco}$) the ecosystem.
+
+The notable exception to the pattern of fluxes decreasing from tropical to boreal regions is $NEP$ (Fig. 6f), which showed no significant differences across biomes.
+Unlike the other C flux variables, $NEP$ does not represent the rapidity with which C cycles through the ecosystem, but is the balance between C sequestration ($GPP$) and respiratory losses ($R_{eco}$) and represents net CO~2~ sequestration (or release) by the ecosystem.
+$NEP$ tends to be relatively small in mature forest stands [@Luyssaert_old-growth_2008, **MORE REFS?**; discussed further below], which accumulate carbon slowly relative to younger stands [@Luyssaert_old-growth_2008; **REFS**), if at all (**REFS**).
+It is therefore unsurprising that there are no pronounced differences across biomes, suggesting that variation in $NEP$ of mature forests is controlled less by climate and more by other factors including moderate disturbances (**REFS**) or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition; **REFS**].
+
+In contrast to the patterns observed for $NEP$ in mature stands, $NEP$ of young forests ...
+(discuss NEP well- intuition for some would be that NEP is higher in tropics.
+Note biome differences for young mut not mature forests, link to latitudinal differences in C sequestration potential [@anderson_temperature-dependence_2006; @cook-patton_mapping_2020]
+Lack of synthesis comparing deadwood and organic layer across biomes, but see @cook-patton_mapping_2020 for age trends.
+
+More rapid accumulation of biomass at lower latitudes has been well-established [@anderson_temperature-dependence_2006; @cook-patton_mapping_2020].
+
+For regrowth forests, far less is known about cross-biome differences in carbon cycling. 
+However, for most C flux variables, this analysis is the first to examine flux trends in regrowth forests across biomes  (i.e., age x biome interaction).
 Data remain sparse, but for better-represented variables, we often see faster acceleration of C cycling in the warmer climates.
@@
 
-In contrast to C fluxes and accumulation rates in regrowth forests, stocks...
-Patterns are consistent with others studies.
-Highest biomass forests are found in temperate oceanic climates (REF- Keith, something in GEB, some global forest C map).
+In contrast to C fluxes and biomass accumulation rates in regrowth forests, stocks showed less systematic variation across biomes.
+For aboveground biomass, this Patterns are consistent with others studies showing that variation in forest biomass across broad climatic gradients is modest and constrained more by moisture than temperature [@stegen_variation_2011].
+@hu_mapping_2016, using spaceborne lidar, showed a decline in aboveground biomass (all forests, including secondary) with latitude in the N hemisphere, but values exceeding tropical forests in coastal climates of both the southern and northern hemisphere.
+Highest biomass forests are found in temperate oceanic climates (REF- , something in GEB, some global forest C map) [@keith_re-evaluation_2009].
+Lack of synthesis comparing deadwood and organic layer across biomes, but see @cook-patton_mapping_2020 for age trends.
 
 Thus, biome differences should be interpreted more as driven more by geographic distribution of sampling than by true differences.
-
-Higher NEP in temperate forests – implications?
-Invariant NEP in older forests?
-This could be built out a bit; thinking of Luyssaert 2008 http://dx.doi.org/10.1038/nature07276 and following papers arguing about this.
-
-(discuss NEP well- intuition for some would be that NEP is higher in tropics.
-Note biome differences for young mut not mature forests, link to latitudinal differences in C sequestration potential (Anderson 2006, @cook-patton_mapping_2020)
 
@@
 Our findings are largely consistent with, but built from a far larger dataset than, those of Pregitzer and Euskirchen (2004 http://dx.doi.org/10.1111/j.1365-2486.2004.00866.x), who found that NPP and NEP to be higher in intermediate-aged forests than older forests, and emphasize the importance of forest age at the biome scale.
-Quickly-changing and age-dependent fluxes were also found in a number of previous syntheses (Amiro et al.
-2010 10.1029/2010JG001390, Magnani et al.
-2007 10.1038/nature05847).
+Quickly-changing and age-dependent fluxes were also found in a number of previous syntheses (Amiro et al. 2010 10.1029/2010JG001390, Magnani et al. 2007 10.1038/nature05847).
+
+In contrast to most fluxes, $NEP$ is highest at intermediate ages
 
@@
 
-*Remote sensing measurements are increasingly useful for global- or regional-scale estimates of forest $GPP$ (Bagdley et al.
-2019, [@li_mapping_2019]), aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) [@clark_quantifying_2004; @leitold_nino_2018], and to some extent net ecosystem exchange ($NEP$) (REFS).*
+*Remote sensing measurements are increasingly useful for global- or regional-scale estimates of forest $GPP$ [@bagdley_terrestrial_2019; @li_mapping_2019], aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) [@clark_quantifying_2004; @leitold_nino_2018], and to some extent net ecosystem exchange ($NEP$) (REFS).*
 Other variables, in particular respiration fluxes, cannot be remotely sensed (@bond-Lamberty_estimating_2016), and efforts such as the Global Carbon Project (**le quere REF**) and NASA CMS (**citation: https://carbon.nasa.gov/pdfs/CMS_Phase-1_Report_Final_optimized.pdf but maybe better to cite open literature, one of the papers listed at https://cmsflux.jpl.nasa.gov/get-data/publication-data-sets**) typically compute them as residuals only.

### 2020-09-29 manuscript/ERL_global_C_review.Rmd
worked on discussion, tried to knit in new results
@ValentineHerr , the Rmd script is now breaking on line 115. Can you help?

@@
 n_fluxes_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_fluxes7])  
-n_stocks_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_stocks7])  
+n_stocks_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_stocks7]) #BREAKS HERE
 n_fluxes_with_ge3_areas <- table(ForC_biome_averages$Biome[idx_fluxes3])  
@@
 
-Our analysis reveals a general acceleration of carbon cycling from the tropics to the high latitudes.
+Our analysis reveals a general deceleration of carbon cycling from the tropics to boreal regions.
 For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude (or increase with temperature) on a global scale [e.g., @luyssaert_co_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
@@
 
-In contrast to the patterns observed for $NEP$ in mature stands, $NEP$ of young forests ...
-(discuss NEP well- intuition for some would be that NEP is higher in tropics.
-Note biome differences for young mut not mature forests, link to latitudinal differences in C sequestration potential [@anderson_temperature-dependence_2006; @cook-patton_mapping_2020]
-Lack of synthesis comparing deadwood and organic layer across biomes, but see @cook-patton_mapping_2020 for age trends.
+In contrast to the patterns observed for $NEP$ in mature stands, $NEP$ of stands between 20 and 100 years of age varied across biomes, being lowest in boreal forests, intermediate in temperate broadleaf forests, and highest in temperate conifer forests (with insufficient data to assess tropical forests; Figs.
+6f, S1).
+This is consistent with findings that live biomass accumulation rates (*e.g.*, $\Delta B_{ag}$ or $\Delta B_{tot}$) during early secondary succession decrease with latitude [@anderson_temperature-dependence_2006; @cook-patton_mapping_2020; Figs.
+7a, S16-S22].
+Note, though, that $NEP$ includes not only $\Delta B_{tot}$, but also changes in $DW_{tot}$, $OL$, and soil carbon, and biome differences in the accumulation rates of these variables have not been detected, in part because these variables do not consistently increase with stand age [@cook-patton_mapping_2020; Figs.
+7, S23-S26; see discussion below].
 
-More rapid accumulation of biomass at lower latitudes has been well-established [@anderson_temperature-dependence_2006; @cook-patton_mapping_2020].
-
-For regrowth forests, far less is known about cross-biome differences in carbon cycling. 
-However, for most C flux variables, this analysis is the first to examine flux trends in regrowth forests across biomes  (i.e., age x biome interaction).
-Data remain sparse, but for better-represented variables, we often see faster acceleration of C cycling in the warmer climates.
-Further work will be required to explore age x climate interactions, but our broad-brush overview indicates that C cycling of regrowth forests is not only higher in the tropics, parallel to fluxes in mature forests [ @banbury_morgan_global_nodate], but also that it accelerates more rapidly with stand age in the tropics, consistent with more rapid biomass accumulation.
+For regrowth forests, little is known about cross-biome differences in carbon fluxes, and we are not aware of any previous large-scale comparisons of C fluxes that have been limited to regrowth forests [although they're commonly mixed in with mature forests; e.g., **REFS**].
+Thus, this analysis was the first to examine flux trends in regrowth forests across biomes.
+The observed tendency for young forest fluxes to decrease from tropical to boreal regions paralleled patterns in mature forests (Figs.
+6, S1-S15), suggesting that regrowth forests follow latitudinal trends in carbon cycling similar to those of mature forests [e.g., @banbury_morgan_global_nodate].
+*Yet, data remain sparse, and further work will be required to explore age x climate interactions.Nevertheless, our broad-brush overview indicates that C cycling of regrowth forests is not only higher in the tropics, parallel to fluxes in mature forests [@banbury_morgan_global_nodate], but also that it accelerates more rapidly with stand age in the tropics, consistent with more rapid biomass accumulation.*
 

### 2020-09-29 manuscript/ERL_global_C_review.Rmd
update method to specify when models were not run

@@
 n_fluxes_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_fluxes7])  
-n_stocks_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_stocks7]) #BREAKS HERE
+n_stocks_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_stocks7])
 n_fluxes_with_ge3_areas <- table(ForC_biome_averages$Biome[idx_fluxes3])  
@@
 These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas.
-Of the 23 flux and 11 stock variables mapped in these diagrams, *ForC* contained sufficient mature forest data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ 7 distinct geographic areas) for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests.
-For regrowth forests (<100 yrs), *ForC* contained sufficient data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ **3** distinct geographic areas) for `r n_fluxes_with_ge3_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge3_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.
+Of the 23 flux and 11 stock variables mapped in these diagrams, *ForC* contained sufficient mature forest data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ 7 distinct geographic areas, and number of distinct plots < number of observations) for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests.
+For regrowth forests (<100 yrs), *ForC* contained sufficient data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ **3** distinct geographic areas, and number of distinct plots < number of observations) for `r n_fluxes_with_ge3_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge3_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.
 

### 2020-09-29 manuscript/ERL_global_C_review.Rmd
knit in new results

@@
 (**check if deforestation statistic below includes natural fire, exclude here if it does.**)
-In recent years, the remaining CO~2~ sink averaged 3.2 $\pm$ 0.6 GtC yr^−1^ from 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
+In recent years, the remaining CO~2~ sink averaged 3.2 $\pm$ 0.6 GtC yr^-1^ from 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
 Yet, this sink is reduced by *deforestation/ forest losses to anthropogenic and natural disturbances*.

### 2020-09-29 manuscript/ERL_global_C_SI.Rmd
knit in new results, working on text

@@
                      stringsAsFactors = FALSE, check.names = FALSE)
-kable(table_pt, format = "latex", booktabs = TRUE, longtable = TRUE,  escape = F, linesep = c('\\addlinespace'))%>%
+kable(table_pt, format = "latex", booktabs = TRUE, longtable = TRUE,  escape = F)%>%
   #pack_rows("NEP", 1, 5, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%

### 2020-09-29 manuscript/ERL_global_C_review.Rmd
knit in new results, working on text

--- 2020-09-29 manuscript/ERL_global_C_review.Rmd
@@
 
-- We're still finalizing the data (and adding some data).
-Outliers in plots will all be checked/ resolved, and we'll be able to pull in more data (*GROA* needs classification by dominant vegetation before it can be pulled in, and we're working on that)
 - "**???**" indicates a reference that we have entered in the .Rmd file, but not yet in the bibliography file. Don't worry about those. (However, places with "REF" need references)
@@
 For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets were generally consistent.
-*That is, component variables summed to within one standard deviation of their respective aggregate variables in all but* **two** *instances, both in temperate conifer forests (Fig. 5).* 
-**(check all this with final results: https://github.com/forc-db/ForC/blob/master/numbers_and_facts/C_cycle_closure.csv, and be sure to write out variable names at first occurance, and provide a bit more detail) **
-*For the temperate conifer biome, the average composite measure of aboveground biomass (Bag) was less than the combined average value of woody biomass (Bag-wood) and foliage biomass (Bfoliage), partly due to the very high estimates of Bag-wood.*
-*Within this biome, $B_{ag}<B_{ag-wood}+B_{foliage}$ and $B_{root}<B_{root-coarse}+B_{root-fine}$ because $B_{ag-wood}$ and $B_{root-coarse}$ were very high, with strongly disproportionate numbers of records from the high-biomass forests of the US Pacific Northwest (Figs.
-S18, S21).*
+That is, component variables summed to within one standard deviation of their respective aggregate variables in all but one instance.
+In the temperate conifer biome, the average composite measure of root biomass ($B_{root}$) was less than the combined average value of coarse and fine root biomass ($B_{root-coarse}$ and $B_{root-fine}$, respectively).
+This lack of closure was driven by very high estimates of $B_{root-coarse}$ from high-biomass forests of the US Pacific Northwest.
 

### 2020-09-29 manuscript/ERL_global_C_SI.Rmd
knit in new results

@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/ANPP_woody_stem.png")
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/ANPP_stem.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/BNPP_root.png")
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/BNPP.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/BNPP_root_coarse.png")
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/BNPP_coarse.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/BNPP_root_fine.png")
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/BNPP_fine.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/R_auto_root.png")
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/R_root.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass_ag_foliage.png")
+  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass_foliage.png")
 ```

### 2020-09-29 manuscript/ERL_global_C_review.Rmd
hide output from knitted product

@@
 
-```{r eval = TRUE, echo=FALSE, warning=FALSE}
+```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
 #READING IN TABLES TO GET STATS FOR TEXT

### 2020-09-29 manuscript/ERL_global_C_review.Rmd
knit in fixed table
closes #50

(also, Jenny opted out of coauthorship)

--- 2020-09-29 manuscript/ERL_global_C_review.Rmd
@@
 Abigail E. Ferson^1,6^
-
-Jennifer C. McGarvey^1^
 
@@
   pack_rows("Stocks", 24, 34, colnum = 2, hline_before = FALSE) %>%
-kableExtra::footnote(symbol = c("Tr: Tropical, TeB: Temperate Broadleaf, TeN: Temperate Needleleaf, B: Boreal, n.s.: no significant differences", "+ or -: significant positive or negative trend, xB: significant age x biome interaction, n.s.: no significant age trend"), footnote_as_chunk = FALSE)
+kableExtra::footnote(symbol = c("Tr: Tropical, TeB: Temperate Broadleaf, TeN: Temperate Needleleaf, B: Boreal, n.s.: no significant differences, n.t.: not tested ", "+ or -: significant positive or negative trend, xB: significant age x biome interaction, n.s.: no significant age trend, n.t.: not tested"), footnote_as_chunk = FALSE)
 ```
@@
 
-**check this with final data (based on Table 1):**
-There were sufficient data to assess mature forest biome differences for **20** flux variables, and significant differences among biomes were detected for **14** variables (Table 1).
-In all cases--including C fluxes into, within, and out of the ecosystem--C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs.
+There were sufficient data to assess mature forest biome differences for 15 flux variables, and significant differences among biomes were detected for 12 variables (Table 1).
+In all of these cases--including C fluxes into, within, and out of the ecosystem--C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs.
 6, S1-S15).
@@
 Fluxes tended to be numerically greater in temperate broadleaf than conifer forests, but the difference was never statistically significant.
-This pattern held for the following variables: ** $GPP$, $NPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$,$ANPP_{branch}$, $ANPP_{foliage}$, $ANPP_{litterfall}$, $M_{woody}$, $BNPP$, $R_{eco}$, $R_{soil}$, and $R_{het-soil}$**.
-For variables without significant differences among biomes, the same general trends applied.
+This pattern held for the following variables: GPP$, $NPP$, $ANPP$, $ANPP_{stem}$,$ANPP_{branch}$, $ANPP_{foliage}$, $BNPP$, $R_{eco}$, $R_{root}$, $R_{soil}$, and $R_{het-soil}$. 
+For two of the variables without significant differences among biomes ($ANPP_{litterfall}$ and $BNPP_{fine}$; Figs.
+S8 and S11, respectively), the same general trends applied but were not statistically significant.
+Another exception was for $BNPP_{root-coarse}$, where all records came from high-biomass forests in the US Pacific Northwest, resulting in marginally higher values for the temperate conifer biome (Table 1, Fig.
+S10; differences significant in mixed effects model but not in post-hoc pairwise comparison).
 
-The most notable exception to this pattern was $NEP$, with no significant differences across biomes but with the largest average in temperate broadleaf forests, followed by temperate conifer, boreal, and tropical forests (Figs.
+The most notable exception to the pattern of decreasing flux from tropical to boreal biomes was $NEP$, with no significant differences across biomes but with the largest average in temperate broadleaf forests, followed by tropical, boreal, and temperate conifer forests (Figs.
 5,S1).
-Another exception was for $BNPP_{root-coarse}$, where all records came from high-biomass forests in the US Pacific Northwest, resulting in high values for the temperate conifer biome and no significant differences across biomes (Fig.
-S10).
-Thus, C cycling rates generally decreased from tropical to temperate to boreal forests, with the important exception in the overall C balance ($NEP$).
 
@@
 These can be significant.
-(**Note that we've given a lot of emphasis to dead wood (work by Abby, and also Jenny), and as a result this work really advances knowledge of dead wood.
+(**Note that we've given a lot of emphasis to dead wood (work by Abby), and as a result this work really advances knowledge of dead wood.
 We'll want to highlight that here.**)
@@
 ## Acknowledgements
-All researchers whose data is included in *ForC* and this analysis.
-Ian McGregor for help with the database.
-Thanks to Norbert Kunert and [Helene's intern] for helpful input at an earlier phase.
-A Smithsonian Scholarly Studies grant to KAT and HML.
-WLS grant to KAT.
+Thanks to all researchers whose data are included in *ForC* and this analysis, to Jennifer McGarvey and Ian McGregor for help with the database, and to Norbert Kunert for helpful discussion.
+Funding sources included a Smithsonian Scholarly Studies grant to KAT and HML and a Smithsonian Working Land and Seascapes grant to KAT.
 

### 2020-09-29 manuscript/ERL_global_C_review.Rmd
revising results with final results

@@
 *Background.* Earth’s climate is closely linked to forests, which strongly influence atmospheric carbon dioxide (CO~2~) and climate by their impact  on the global carbon (C) cycle.
-However, efforts to incorporate forests into climate models and CO2 accounting frameworks have been constrained by a lack of accessible, global-scale data on how C cycling varies across forest types and stand ages.
+However, efforts to incorporate forests into climate models and CO~2~ accounting frameworks have been constrained by a lack of accessible, global-scale data on how C cycling varies across forest types and stand ages.
 
@@
 
-**check this with final data (based on Table 1):**
-There were sufficient data to assess mature forest biome differences for **9** stock variables, and significant differences among biomes were detected for **6** variables (**$B_{tot}$, $B_{ag}$, $B_{ag-wood}$, $B_{foliage}$, $B_{root-coarse}$, $DW_{tot}$**; Table 1).
+There were sufficient data to assess mature forest biome differences for nine stock variables, and significant differences among biomes were detected for five variables ($B_{tot}$, $B_{ag}$, $B_{ag-wood}$, $B_{foliage}$, $B_{root-coarse}$; Table 1).
 C stocks had less consistent patterns across biomes (Figs. 7, S16-S26).
-In **four of the six** cases (**$B_{tot}$, $B_{ag-wood}$,  $B_{root-coarse}$, $DW_{tot}$**), temperate conifer forests had significantly higher stocks than the other biomes, and boreal forests in the lowest, with tropical and temperate broadleaf forests in between.
-For $B_{ag}$, which had by far the highest sample size, tropical forests exceeded temperate conifer forests (but not significantly). 
-For $B_{foliage}$,temperate broadleaf forests were lowest (again, not significantly).
-The high values for the temperate conifer biome were driven by the very high-biomass forests of the US Pacific Northwest, which are disproportionately represented in the current version of ForC.
-Thus, biome differences should be interpreted more as driven more by geographic distribution of sampling than by true differences.
+For $B_{tot}$ and $B_{ag}$, tropical broadleaf forests had the highest biomass and boreal forests the lowest, with temperate broadleaf and needleleaf ($B_{ag}$ only) intermediate.
+For three variables that had been disproportionately sampled in the high-biomass forests of the US Pacific Northwest ($B_{ag-wood}$, $B_{foliage}$, and $B_{root-coarse}$), temperate conifer forests had significantly higher stocks than the other biomes, which were not significantly different from one another.
 
@@
 Average C cycles for forests <100 years old are presented in Figures 8-11.  
-Both C stocks and fluxes commonly increased signficantly with stand age (Tables 1, S2, Figs. 6-11, S1-S26; detailed below).
+Both C stocks and fluxes commonly increased significantly with stand age (Tables 1, S2, Figs. 6-11, S1-S26; detailed below).
 

### 2020-09-29 manuscript/ERL_global_C_review.Rmd
updated results, knit

@@
 
-*ForC* contained *14* flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig. 6-7 and **S#- SI figures including plots for all variables**).
-Of these, *9* increased significantly with age: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($NEP$).
-The remaining five--$ANPP_{woody-branch}$, $BNPP_{root-coarse}$, $R_{soil-het}$, and $R_{soil-het}$-–displayed no significant relationship to stand age, although all displayed a positive trend.
+*ForC* contained 16 C flux variables with sufficient data for analyses of age trends in young forests (see Methods) (Figs. 6, S1-S15).
+Of these, ten increased significantly with age: $NEP$, $GPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{foliage}$, $ANPP_{litterfal}$, $BNPP$, $BNPP_{fine}$, $R_{eco}$, and $R_{root}$.
+The remaining six--$NPP$, $ANPP_{stem}$, $ANPP_{branch}$, $BNPP_{coarse}$, $R_{soil-het}$, and $R_{soil-het}$-–displayed no significant relationship to stand age.
 
 Differences in C fluxes across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests.  
-The single exception was $ANPP_{stem}$, for which temperate broadleaf forests and temperate conifer forests of age >*~30* had slightly higher flux rates than tropical forests (*represented by a single chronosequence*).
-Notably, the trend of tropical > temperate > boreal held for $NEP$ in regrowth forests, in contrast to the lack of biome differences in $NEP$ for mature forests (Fig. 6).
+The single exception was $ANPP_{stem}$, for which temperate broadleaf and conifer forests had similar flux rates than tropical forests.
+Notably, and in contrast to the lack of biome differences in $NEP$ for mature forests (Fig.
+6), the tendency for temperate forests to have greater fluxes than boreal forests held for $NEP$ in regrowth forests (tropical forests excluded because of insufficient data).
 
-There were only *##* flux variables with sufficient data to test for biome x age interactions: $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{litterfall}$, and $BNPP$ (Table S2).
-**(more could be added if age trends become significant after outliers are resolved)**
-For three of these ($ANPP$, $ANPP_{litterfall}$, $BNPP$), the increase in C flux with age was steepest increase in tropical forests, followed by temperate and then boreal forests (Figs S#).
-Similarly, $ANPP_{woody}$ displayed a steeper increase with age in temperate than boreal boreal forests (no tropical data for this variable).
-In contrast, for $ANPP_{stem}$,  tropical and temperate broadleaf forests had the highest flux rates at young ages, but were surpassed by temperate conifer forests between ages 20 and 50 (Fig. S6).
-
-(**this needs to be updated with latest data**)
-In terms of C stocks, *10* variables had sufficient data to test for age trends.
-*Six* of these—*total biomass, aboveground biomass, aboveground woody biomass, foliage biomass, root biomass, and coarse root biomass*— increased significantly with $log10[stand.age]$.
-*The remaining four displayed non-significant positive trends:  fine root biomass, total dead wood, standing dead wood, and organic layer.*
-*(discuss rates of increase)*
+In terms of C stocks, ten variables (all but standing deadwood, $DW_{standing}$) had sufficient data to test for age trends (Table 1, Figs. 7, S16-26).
+All of these increased significantly with $log10[stand.age]$.
+There were sufficient data to model age x biome interactions were also significant for all ten of these C stock variables (Table S2). 
+(**return here after Valentine checks whether differences in Table 1 were signficant! **)
+In the case of two non-living C stocks ($DW_{down}$ and $OL$)
 

### 2020-09-29 manuscript/ERL_global_C_review.Rmd
worked on discussion, references

@@
 Forest ecosystems are shaping the course of climate change (IPCC1.5) through their influence on atmospheric carbon dioxide (CO~2~).
-Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climate-carbon_2006; @krause_large_2018; @bonan_model_2019; @devittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history.
+Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climatecarbon_2006; @krause_large_2018; @bonan_model_2019; @di_vittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history.
 These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published.
@@
 
-The large number of C cycle variables covered by ForC, and the general consistency between them, provide confidence that our overall reported means provide accurate and useful baselines for analysis (with the caveats that they are unlikely to be accurate representations of C cycling for any particular forest, and that these sample means almost certainly do not represent true biome means).
+The large number of C cycle variables covered by ForC, and the general consistency between them, provide confidence that our overall reported means provide accurate and useful baselines for analysis -- with the caveats that they are unlikely to be accurate representations of C cycling for any particular forest, and that these sample means almost certainly do not represent true biome means (particularly for temperate conifer forests where high-biomass stands are over-represented in *ForC*).
 
-There are of course notable holes in the ForC variable coverage, as discussed by Anderson-Teixeira et al.
-(xxxx), that limit the scope of our inferences here.
-Notably, ForC lacks coverage of fluxes to herbivores and higher consumers, along with the woody mortality and dead wood stocks.
-Geographically, all variables are poorly covered in Africa and Siberia, a common problem in the carbon-cycle community (Xu and Shang 2016 10.1016/j.jplph.2016.08.007, Schimel et al.
-2015 10.1073/pnas.1407302112).
-ForC does not include soil carbon, which is covered by other efforts (e.g. Köchy et al. 2015 10.5194/soil-1-351-2015). 
-ForC is not intended to replace databases that are specialized for particular parts of the C cycle analyses, e.g., aboveground biomass (REFS), land-atmosphere fluxes (Baldocchi et al.
-2001 10.1175/1520-0477(2001)082<2415:FANTTS>2.3.CO;2), soil respiration (Jian et al.
-2020 10.5194/essd-2020-136), or the human footprint in global forests (Magnani et al.
-2007 10.1038/nature05847).
+There are of course notable holes in the ForC variable coverage (Fig. 1), that limit the scope of our inferences here. 
+Notably, *ForC* currently has sparse--if any--coverage of fluxes to herbivores and higher consumers, along with the woody mortality ($M_{woody}$) and dead wood stocks (Tables 1, Figs. S23-S25). 
+Geographically, all variables are poorly covered in Africa and Siberia, a common problem in the carbon-cycle community [@xu_contribution_2016; @schimel_effect_2015]. 
+*ForC* does not include soil carbon, which is covered by other efforts [e.g., @kochy_global_2015]. 
+*ForC* is not intended to replace databases that are specialized for particular parts of the C cycle analyses, e.g., aboveground biomass (**REFS**), land-atmosphere fluxes [@baldocchi_fluxnet_2001], soil respiration [@jian_reconstructed_2020], or the human footprint in global forests [@magnani_human_2007].
 
 In this analysis, the C cycle budgets for mature forests (Figs.
-2-5) generally “close”–that is, the sums of component variables do not differ from the larger fluxes by more than one standard deviation.
-On the one hand, this reflects the general fact that ecosystem-scale measurements tend to close the C budget more easily and consistently than, for example, for energy balance (Stoy et al.
-2013 10.1016/j.agrformet.2012.11.004).
-On the other, however, as noted above ForC derives data from multiple heterogeneous sources, often with large errors (standard deviations); as a result, the standard for C closure is relatively loose (cf Houghton 2020 10.1111/gcb.15050).
-Nonetheless, the lack of closure, in the few instances where it occurs, is probably more reflective of differences in the representation of forest types (e.g., disproportionate representation of US Pacific NW for aboveground woody biomass relative to AGB; Fig.
+2-5) generally “close”–-that is, the sums of component variables do not differ from the larger fluxes by more than one standard deviation.
+On the one hand, this reflects the general fact that ecosystem-scale measurements tend to close the C budget more easily and consistently than, for example, for energy balance [@stoy_data-driven_2013].
+On the other, however, *ForC* derives data from multiple heterogeneous sources, and standard deviations within each biome are high; as a result, the standard for C closure is relatively loose [*c.f.* @houghton_terrestrial_2020].
+Nonetheless, the lack of closure, in the one instance where it occurs, is probably more reflective of differences in the representation of forest types (*i.e.*, disproportionate representation of US Pacific NW for  $B_{root-coarse}$ relative to  $B_{root}$; Fig.
 4) than of methodological accuracy.
-The overall high degree of closure implies that ForC gives a consistent picture of C cycling within biomes. 
-This is an important and useful test, because it allows for consistency checks within the C cycle, for example leveraging separate and independently-measured fluxes to constrain errors in another (Phillips et al.
-2017 10.1007/s11104-016-3084-x, Williams et al.
-2014 10.1016/j.rse.2013.10.034, Harmon et al.
-2011 10.1029/2010JG001495), or producing internally consistent global data products (Wang et al.
-2018 10.5194/gmd-11-3903-2018).
+The overall high degree of closure implies that *ForC* gives a consistent picture of C cycling within biomes. 
+This is an important and useful test, because it allows for consistency checks within the C cycle, for example leveraging separate and independently-measured fluxes to constrain errors in another [@phillips_value_2017; @williams_impacts_2014; @harmon_heterotrophic_2011], or producing internally consistent global data products [@wang_golum-cnp_2018].
 
@@
 Unlike the other C flux variables, $NEP$ does not represent the rapidity with which C cycles through the ecosystem, but is the balance between C sequestration ($GPP$) and respiratory losses ($R_{eco}$) and represents net CO~2~ sequestration (or release) by the ecosystem.
-$NEP$ tends to be relatively small in mature forest stands [@Luyssaert_old-growth_2008, **MORE REFS?**; discussed further below], which accumulate carbon slowly relative to younger stands [@Luyssaert_old-growth_2008; **REFS**), if at all (**REFS**).
+$NEP$ tends to be relatively small in mature forest stands [@luyssaert_old-growth_2008, **MORE REFS?**; discussed further below], which accumulate carbon slowly relative to younger stands [@luyssaert_old-growth_2008; **REFS**), if at all (**REFS**).
 It is therefore unsurprising that there are no pronounced differences across biomes, suggesting that variation in $NEP$ of mature forests is controlled less by climate and more by other factors including moderate disturbances (**REFS**) or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition; **REFS**].
@@
 6, S1-S15), suggesting that regrowth forests follow latitudinal trends in carbon cycling similar to those of mature forests [e.g., @banbury_morgan_global_nodate].
-*Yet, data remain sparse, and further work will be required to explore age x climate interactions.Nevertheless, our broad-brush overview indicates that C cycling of regrowth forests is not only higher in the tropics, parallel to fluxes in mature forests [@banbury_morgan_global_nodate], but also that it accelerates more rapidly with stand age in the tropics, consistent with more rapid biomass accumulation.*
 
 In contrast to C fluxes and biomass accumulation rates in regrowth forests, stocks showed less systematic variation across biomes.
-For aboveground biomass, this Patterns are consistent with others studies showing that variation in forest biomass across broad climatic gradients is modest and constrained more by moisture than temperature [@stegen_variation_2011].
+This is consistent with others studies showing that variation in forest biomass across broad climatic gradients is modest and constrained more by moisture than temperature [@stegen_variation_2011].
+Yet, differences across biomes did exist.
 @hu_mapping_2016, using spaceborne lidar, showed a decline in aboveground biomass (all forests, including secondary) with latitude in the N hemisphere, but values exceeding tropical forests in coastal climates of both the southern and northern hemisphere.
@@
 *Remote sensing measurements are increasingly useful for global- or regional-scale estimates of forest $GPP$ [@bagdley_terrestrial_2019; @li_mapping_2019], aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) [@clark_quantifying_2004; @leitold_nino_2018], and to some extent net ecosystem exchange ($NEP$) (REFS).*
-Other variables, in particular respiration fluxes, cannot be remotely sensed (@bond-Lamberty_estimating_2016), and efforts such as the Global Carbon Project (**le quere REF**) and NASA CMS (**citation: https://carbon.nasa.gov/pdfs/CMS_Phase-1_Report_Final_optimized.pdf but maybe better to cite open literature, one of the papers listed at https://cmsflux.jpl.nasa.gov/get-data/publication-data-sets**) typically compute them as residuals only.
+Other variables, in particular respiration fluxes, cannot be remotely sensed (@bond-lamberty_estimating_2016), and efforts such as the Global Carbon Project (**le quere REF**) and NASA CMS (**citation: https://carbon.nasa.gov/pdfs/CMS_Phase-1_Report_Final_optimized.pdf but maybe better to cite open literature, one of the papers listed at https://cmsflux.jpl.nasa.gov/get-data/publication-data-sets**) typically compute them as residuals only.
 (**Ben, it woudl be particularly helpful if you could flesh this out some more.**)

### 2020-09-30 manuscript/ERL_global_C_review.Rmd
work on discussion

@@
 
-For regrowth forests, little is known about cross-biome differences in carbon fluxes, and we are not aware of any previous large-scale comparisons of C fluxes that have been limited to regrowth forests [although they're commonly mixed in with mature forests; e.g., **REFS**].
+For regrowth forests, little is known about cross-biome differences in carbon fluxes, and we are not aware of any previous large-scale comparisons of C fluxes that have been limited to regrowth forests [although they're commonly mixed in with mature forests; e.g., @luyssaert_co_2007 **REFS**].
 Thus, this analysis was the first to examine flux trends in regrowth forests across biomes.
@@
 Yet, differences across biomes did exist.
-@hu_mapping_2016, using spaceborne lidar, showed a decline in aboveground biomass (all forests, including secondary) with latitude in the N hemisphere, but values exceeding tropical forests in coastal climates of both the southern and northern hemisphere.
-Highest biomass forests are found in temperate oceanic climates (REF- , something in GEB, some global forest C map) [@keith_re-evaluation_2009].
-Lack of synthesis comparing deadwood and organic layer across biomes, but see @cook-patton_mapping_2020 for age trends.
+For aboveground biomass, which is the variable in *ForC* with broadest geographical representation, the trend of declining biomass from tropical to boreal regions mirrors observations from spaceborne lidar that reveal a decline in aboveground biomass (for all forests, including secondary) with latitude across the N hemisphere [@hu_mapping_2016].
+Yet, the highest- biomass forests on Earth are found in coastal temperate climates of both the southern and northern hemisphere [@keith_re-evaluation_2009; @smithwick_potential_2002; **REF??**; @hu_mapping_2016].
+Disproportionate representation of forests in one such region--the US Pacific Northwest--inflated estimates of temperate conifer fluxes and stocks for some variables and was responsible for all of the anomalous results described here (e.g., lack of complete C budget closure, anomalous trend across biomes for $BNPP_{coarse}$).
+Thus, biome differences should always be interpreted relative to the geographic distribution of sampling, which only rarely covers the majority of forested area within a biome.
 
-Thus, biome differences should be interpreted more as driven more by geographic distribution of sampling than by true differences.
+Whereas biomass can be remotely sensed and receives a lot of research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
+Although these stocks can be important--exceeding 100 t C ha^-1^ in some stands (Figs.
+7d-e, S23-S25), this study is the first to synthesize deadwood data on a global scale [but see @cook-patton_mapping_2020 for young forests].
+Unfortunately, data remain too sparse for statistical comparison across biomes (Figs.
+7, S23-S25; but see below for age trends), pointing to a need for more widespread quantification of both standing and downed deadwood.
+*ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that slower decomposition in colder climates results in more buildup of organic matter [@allen_2015].
+Further research on non-living C stocks in the world's forests will be essential to completing the picture.
 
@@
 In contrast to most fluxes, $NEP$ is highest at intermediate ages
+
+In terms of stocks, our study reveals consistent increases in live biomass stocks with stand age--a pattern that is well-known and expected [e.g., **REFS**]--and more variable age trends in deadwood and $OL$.
+The latter are particularly sensitive to the type of disturbance, where disturbances that remove deadwood (e.g., logging, agriculture) result in negligible deadwood in young stands, followed by a buildup over time [e.g., tropical stands in Fig.
+7e **REFS**].
+In contrast, natural disturbances (e.g., fire, drought) can produce large amounts of deadwood (mostly $DW_{standing}$) that slowly decomposes as the stand recovers, resulting in declines across young stand ages [e.g., temperate and boreal stands in Fig.
+7e; **REF**].
+Again, further study and synthesis of non-living C stocks across biomes and stand ages will be valuable to giving a more comprehensive picture.
 

### 2020-09-30 manuscript/ERL_global_C_review.Rmd
add Ben's text

@@
 The future of the current forest C sink is dependent both upon forest responses to climate change itself and human land use decisions, which will feedback and strongly influence the course of climate change.
-Regrowing forests in particular will play an important role [@pugh_role_2019], as these represent a large (**~#%**) and growing proportion of Earth's forests [@mcdowell_pervasive_2020].
+Regrowing forests in particular will play an important role [@pugh_role_2019], as almost two-thirds of the world’s forests were secondary as of 2010 (**FAO 2010**)
+As anthropogenic and climate-driven disturbances impact an growing proportion of Earth's forests [@andela_human-driven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests is increasingly important.
+Although age trends in aboveground biomass have been relatively well-studied and recently synthesized globally [@cook-patton_mapping_2020], a relative dearth of data and synthesis on other C stocks and fluxes in secondary forests points to an under-filled need to characterize age-related trends in forest C cycling.
+Such understanding is particularly critical for reducing uncertainty regarding the potential for carbon uptake and climate change mitigation by regrowth forests [@kruase_large_2018; @cook-patton_mapping_2020].
 Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thus central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; **Cavaleri et al 2015**].
@@
 
-*(Just some rough notes at this point)*
+**Our study reveals that most C fluxes quickly increase to a plateau as stands age...**
+@anderson-teixeira_altered_2013
 
-A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling.
+**(paragraph on NEP): **
 
-Moreover, as disturbances increase [@andela_human-driven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests will be increasingly important.
-
-It's also important to understand secondary forest C sequestration to reduce uncertainty regarding the potential for carbon uptake by regrowth forests [@kruase_large_2018; @cook-patton_mapping_2020].
-
-(discuss NEP well, including)
 NEP increases with log(age) to 100 --> strongest C sinks are established secondary forests. (But presumably this exact number is an artifact; don’t over-emphasize.)
@@
 
-The future of forest C cycling [@song_meta-analysis_2019] will shape trends in atmospheric CO~2~ and the course of climate change.
-For a human society seeking to understand and mitigate climate change, the data contained in *ForC* and summarized here can help to meet two major challenges.
+The future of forest C cycling [@song_meta-analysis_2019] will shape trends in atmospheric CO~2~ and the course of climate change (**Schimel et al. 2015 10.1073/pnas.1407302112**).
+Our findings, and more generally the data contained in *ForC* and summarized here, can help to meet two major challenges.
 
-First, improved representation of forest C cycling in models is essential to improving predictions of the future course of climate change.
-To ensure that models are giving the right answers for the right reasons, it is important to benchmark against multiple components of the C cycle that are internally consistent with each other.
-By making tens of thousands of records readily available in standardized format, *ForC* makes it feasible for the modeling community to draw upon these data to benchmark models.
-Integration of *ForC* with models is a goal (Fer et al., in revision).
+First, improved representation of forest C cycling in models is essential to improving predictions of the future course of climate change, for the simple reason that by definition future projections extend our existing observations and understanding to conditions that do not currently exist on Earth (**https://doi.org/10.1016/j.tree.2017.10.002, https://doi.org/10.1126/science.aam8328, https://link.springer.com/article/10.1186/s13717-018-0142-8**).
+To ensure that models are giving the right answers for the right reasons (Sulman et al.
+2018 10.1007/s10533-018-0509-z), it is important to benchmark against multiple components of the C cycle that are internally consistent with each other (Collier et al.
+2018 10.1029/2018MS001354, Wang et al.
+2018 10.5194/gmd-11-3903-2018).
+*ForC*’s tens of thousands of records are readily available in a standardized format, and our analyses here indicate their internal consistency is reasonably high.
+Integration of ForC with models is a goal (Fer et al., in revision).
 
+**(Ben would remove this paragraph:)**
 Second, *ForC* can serve as a pipeline through which forest science can inform forest-based climate change mitigation efforts. 
@@
 One goal for *ForC* is to serve as a pipeline through which information can flow efficiently from forest researchers to decision-makers working to implement forest conservation strategies at global, national, or landscape scales.
-This is already happening!
-*ForC* has already contributed to updating the IPCC guidelines for carbon accounting in forests [IPCC 2019; @requena_suarez_estimating_2019; Rozendaal et al in prep], mapping C accumulation potential from natural forest regrowth globally [@cook-patton_mapping_2020], and informing ecosystem conservation priorities [@goldstein_protecting_2020].
+This is already happening: *ForC* has contributed to updating the IPCC guidelines for carbon accounting in forests [**IPCC 2019**; @requena_suarez_estimating_2019], mapping C accumulation potential from natural forest regrowth globally [@cook-patton_mapping_2020], and informing ecosystem conservation priorities [@goldstein_protecting_2020].
 
-***ForC* can complement remote sensing to provide a comprehensive picture of global forest C cycling.** (*here, discuss what is well-covered by remote sesnsing, where ForC is useful for calibrating remote sensing, and what ForC can get that remote sensing can't.*)
-AGB is the largest stock, and most of the emphasis is on this variable.
-Remote sensing, with calibration based on high-quality ground-based data [@schepaschenko_forest_2019; @chave_ground_2019], is the best approach for mapping forest carbon (REFS).
-However, it is limited in that it is not associated with stand age and disturbance history, except in recent decades when satellite data can be used to detect forest loss, gain, and some of their dominant drivers [@hansen_high-resolution_2013; @song_global_2018; @curtis_classifying_2018].
-*ForC*  is therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020.
+It is also interesting to consider the complementary utility of global-scale but spatially discontinuous databases such as ForC and remote wall-to-wall remote sensing products.
+The latter provide unparalleled insight into aboveground carbon stocks, but less constraint on belowground stocks or carbon fluxes in general (Bond-Lamberty et al.
+2016, http://dx.doi.org/10.1002/ecs2.1380, Anav et al.
+2015 http://dx.doi.org/10.1002/2015RG000483).
+Combining observational data and remote observations may provide a much more comprehensive and accurate picture of global forest C cycling, particularly when used in formal data assimilation systems (Konings et all 2019 10.5194/bg-16-2269-2019, Liu et al.
+2018 10.1088/1748-9326/aad5ef).
+biomass is the largest stock, and most of the emphasis has traditionally been on this variable. 
+Remote-sensing driven biomass estimates (e.g.
+Saatchi et al.
+2011 http://dx.doi.org/10.1073/pnas.1019576108), calibrated based on high-quality ground-based data [@schepaschenko_forest_2019; @chave_ground_2019], are well suited for this task.
+Note however that factors such as stand age and disturbance history (except in very recent decades are important but difficult to detect this way [@hansen_high-resolution_2013; @song_global_2018; @curtis_classifying_2018].
+Ground-based data such as *ForC* are therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020, and thus constraining variables such as carbon sink potential (Luyssaert et al.
+2007 http://dx.doi.org/10.1038/nature07276).
 
-*Remote sensing measurements are increasingly useful for global- or regional-scale estimates of forest $GPP$ [@bagdley_terrestrial_2019; @li_mapping_2019], aboveground biomass ($B_{ag}$) (REFS), woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) [@clark_quantifying_2004; @leitold_nino_2018], and to some extent net ecosystem exchange ($NEP$) (REFS).*
-Other variables, in particular respiration fluxes, cannot be remotely sensed (@bond-lamberty_estimating_2016), and efforts such as the Global Carbon Project (**le quere REF**) and NASA CMS (**citation: https://carbon.nasa.gov/pdfs/CMS_Phase-1_Report_Final_optimized.pdf but maybe better to cite open literature, one of the papers listed at https://cmsflux.jpl.nasa.gov/get-data/publication-data-sets**) typically compute them as residuals only.
-(**Ben, it woudl be particularly helpful if you could flesh this out some more.**)
-
-In terms of C stocks, there is a paucity of data on dead wood and organic layer (Pan et al. ?).
-These can be significant.
-(**Note that we've given a lot of emphasis to dead wood (work by Abby), and as a result this work really advances knowledge of dead wood.
-We'll want to highlight that here.**)
-*(give some stats/ cite figures)*.
-
-**Move to data availability statement, or methods?**:
-We recommend that use of *ForC* data go to the original database, as opposed to using "off-the-shelf" values from this publication.
-This is because
-(1) *ForC* is constantly being updated,
-(2) analyses should be designed to match the application,
-(3) age equations presented here all fit a single functional form that is not necessarily the best possible for all the variables.
+In contrast, respiration fluxes cannot be remotely sensed. 
+Efforts such as the Global Carbon Project (le quere http://dx.doi.org/10.5194/essd-11-1783-2019) and NASA’s Carbon Monitoring System (**Liu et al.
+2018 10.1088/1748-9326/aad5ef**) typically compute respiration as residuals of all other terms [@bond-lamberty_estimating_2016;  Harmon et al.
+2011 http://dx.doi.org/10.1029/2010JG001495].
+This means that the errors on respiration outputs are likely to be large and certainly poorly constrained, offering a unique opportunity for databases such as ForC and SRDB [@jian_restructured_2020] to provide observational benchmarks.
+For example, **Konings et al.
+(2019 http://dx.doi.org/10.5194/bg-16-2269-2019)** produced a unique top-down estimate of global heterotrophic respiration that can both be compared with extant bottom-up estimates **(Bond-Lamberty et al.
+2018 http://dx.doi.org/10.1029/2018EF000866)** and used as an internal consistency check on other parts of the carbon cycle **(Phillips et al.
+2016)**.
 
 As climate change accelerates, understanding and managing the carbon dynamics of forests will be critical to forecasting, mitigation, and adaptation.
-The C data in ForC, as summarized here, will be valuable to these efforts.
+The C data in *ForC*, as summarized here, will be valuable to these efforts.
 

### 2020-09-30 manuscript/ERL_global_C_review.Rmd
complete draft!

@@
 
-**Our study reveals that most C fluxes quickly increase to a plateau as stands age...**
-@anderson-teixeira_altered_2013
+Our study reveals that most C fluxes quickly increase to a plateau as stands age (Fig.
+6), consistent with current understanding of age trends in forest C cycling [e.g., @anderson-teixeira_altered_2013].
+*Quickly-changing and age-dependent fluxes were also found in a number of previous syntheses (Amiro et al. 2010 10.1029/2010JG001390, Magnani et al. 2007 10.1038/nature05847).*
+While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_auto$) would be minimal immediately following a stand-clearing disturbance.
+These would be expected to increase rapidly with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Fig. 7).
+In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$) are expected to be non-zero following stand-clearing disturbance, although these may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots and litter [@ribeiro-kumara_how_2020; @mauer_carbon_2020; @bond-lamberty_contribution_2004].
+In this study, we detect no significant age trends in either variable.
 
-**(paragraph on NEP): **
-
-NEP increases with log(age) to 100 --> strongest C sinks are established secondary forests. (But presumably this exact number is an artifact; don’t over-emphasize.)
-
-Our findings are largely consistent with, but built from a far larger dataset than, those of Pregitzer and Euskirchen (2004 http://dx.doi.org/10.1111/j.1365-2486.2004.00866.x), who found that NPP and NEP to be higher in intermediate-aged forests than older forests, and emphasize the importance of forest age at the biome scale.
-Quickly-changing and age-dependent fluxes were also found in a number of previous syntheses (Amiro et al. 2010 10.1029/2010JG001390, Magnani et al. 2007 10.1038/nature05847).
-
-In contrast to most fluxes, $NEP$ is highest at intermediate ages
+Notably, net carbon sequestration ($NEP$) increases with age up to the 100-yr threshold examined here, with more pronounced patterns in temperate than boreal forests (Fig. 6f). 
+This finding is largely consistent with, but built from a far larger dataset than, previous studies showing an increase in $NEP$ across relatively young stand ages [**Pregitzer and Euskirchen (2004 http://dx.doi.org/10.1111/j.1365-2486.2004.00866.x); Baldocchi; Luyssaert**].
+However, $NEP$ has been observed to decline from intermediate to old stands **[Luyssaert 2008]**, and the $NEP$ estimated by our model for 100-year-old temperate conifer stands (~5 Mg C ha^-1^ yr^-1^) exceeds the mean of mature forests in the same biome (0.7 Mg C ha^-1^ yr^-1^; Fig.4).
+A decrease in $NEP$ is consistent with the observed decelleration of biomass accumulation as stands age, although both biomass and non-living C stocks will often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands (**REFS**).
 

### 2020-09-30 manuscript/ERL_global_C_review.Rmd
integrate bbl round 2

@@
 
-*Review Results/ Synthesis.*  *ForC v3.0* yielded a fairly comprehensive picture of C cycling in the world's major forest biomes, with broad closure of C budgets.
+*Review Results/ Synthesis.*  *ForC v3.0* yielded a comprehensive picture of C cycling in the world's major forest biomes.
 The rate of C cycling generally increased from boreal to tropical regions in both mature and regrowth forests, whereas C stocks showed less directional variation.
-The majority of flux variables, together with most live biomass pools, increased significantly with stand age, *and the rate of increase again tended to increase from boreal to tropical regions*.
+The majority of flux variables, together with most live biomass pools, increased significantly with stand age.
+Importantly, there was generally good closure of C budgets, *i.e.,* internal consistency in the *ForC* data.
 
@@
 These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published.
-While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass: @hu_mapping_2016; **REF**; gross primary productivity, $GPP$: @li_mapping_2019; **REFS for biomass change, net CO2 flux**], measurement of most forest C stocks and fluxes require intensive on-the-ground data collection.
-Here, we provide a robust and comprehensive analysis of carbon cycling from a stand to global level, and by biome and stand age, using the largest global compilation of forest carbon data, which is available in our open source Global Carbon Forest database (ForC; Fig.
-1).
+While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass: @hu_mapping_2016; **REF**; gross primary productivity, $GPP$: @li_mapping_2019; **Saatchi 2011; REFS for biomass change, net CO2 flux**], measurement and validation of most forest C stocks and fluxes require intensive on-the-ground data collection.
+
+The future of the current forest C sink is dependent both upon forest responses to climate change itself and human land use decisions, which will feedback and strongly influence the course of climate change.
+Regrowing forests in particular will play an important role [@pugh_role_2019], as almost two-thirds of the world’s forests were secondary as of 2010 (**FAO 2010**)
+As anthropogenic and climate-driven disturbances impact an growing proportion of Earth's forests [@andela_human-driven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests is increasingly important.
+Although age trends in aboveground biomass have been relatively well-studied and synthesized globally [@cook-patton_mapping_2020], a relative dearth of data and synthesis on other C stocks and fluxes in secondary forests points to an under-filled need to characterize age-related trends in forest C cycling.
+Such understanding is particularly critical for reducing uncertainty regarding the potential for carbon uptake and climate change mitigation by regrowth forests [@kruase_large_2018; @cook-patton_mapping_2020].
+Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thus central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; **Cavaleri et al 2015**].
 
@@
 Annual gross CO~2~ sequestration in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
-Most of this enormous C sequestration is counterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) *or fire*, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] *and fire emissions (REF)*.
+Most of this enormous C sequestration is counterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) *or fire*, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] *and fire emissions [@van_der_werf_global_2017]*.
 (**check if deforestation statistic below includes natural fire, exclude here if it does.**)
@@
 
-The future of the current forest C sink is dependent both upon forest responses to climate change itself and human land use decisions, which will feedback and strongly influence the course of climate change.
-Regrowing forests in particular will play an important role [@pugh_role_2019], as almost two-thirds of the world’s forests were secondary as of 2010 (**FAO 2010**)
-As anthropogenic and climate-driven disturbances impact an growing proportion of Earth's forests [@andela_human-driven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests is increasingly important.
-Although age trends in aboveground biomass have been relatively well-studied and recently synthesized globally [@cook-patton_mapping_2020], a relative dearth of data and synthesis on other C stocks and fluxes in secondary forests points to an under-filled need to characterize age-related trends in forest C cycling.
-Such understanding is particularly critical for reducing uncertainty regarding the potential for carbon uptake and climate change mitigation by regrowth forests [@kruase_large_2018; @cook-patton_mapping_2020].
-Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thus central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; **Cavaleri et al 2015**].
-
 Despite the importance of forests, comprehensive global studies have historically been limited by the scattered and more local nature of research studies. 
 Primary research articles typically cover only a small numbers of sites at a time.
-The rare exceptions that span regions or continents with rare exceptions spanning regions or continents are typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; e.g., @lutz_global_2018] or FLUXNET [@baldocchi_fluxnet_2001; e.g., **FLUXNET_REF**].
+The rare exceptions that span regions or continents with rare exceptions spanning regions or continents are typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; e.g., @lutz_global_2018], NEON [@schimel_NEON_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
 The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*.
-In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; **REFS**].
-Scattered data are not conducive for advancing science nor for making decisions about how to best manage our forests as a tool for constraining the climate crisis.
+In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
 
@@
 
-Here, we analyze the more extensive *ForC* data (Fig. 1) to provide a robust overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age.
+Here, we provide a robust and comprehensive analysis of carbon cycling from a stand to global level, and by biome and stand age, using the largest global compilation of forest carbon data, which is available in our open source Global Carbon Forest database (*ForC*; Fig.
+1).
 Our primary goal is to provide a data-driven summary of our current state of knowledge on broad trends in forest C cycling. 
@@
 We've also added data from individual publications (detailed list at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv), with a particular focus on productivity [@banbury_morgan_global_nodate; e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @lutz_global_2018, @johnson_climate_2018].
-We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: **REFS**).
-The database version used for this analysis has been tagged as a new release on Github (**XX**) and assigned a DOI through Zenodo (DOI: TBD).
+The database version used for this analysis has been tagged as a new release on Github (v3.0) and assigned a DOI through Zenodo (DOI: TBD).
 
@@
 
-For this analysis, we grouped forests into four broad biome types based on climate zones and dominant vegetation type (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
+We grouped forests into four broad biome types based on climate zones and dominant vegetation type (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
 Climate zones (Fig. 1) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
@@
 
-To test for differences across mature forest biomes, we also examined how stand age impacted fluxes and stocks, employing a mixed effects model ['lmer' function in 'lme4' R package version **x.xx**; **REF**] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
+To test for differences across mature forest biomes, we also examined how stand age impacted fluxes and stocks, employing a mixed effects model ['lmer' function in 'lme4' R package version **x.xx**; @bates_fitting_2015] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
 When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
@@
 There were sufficient data to model age x biome interactions were also significant for all ten of these C stock variables (Table S2). 
-(**return here after Valentine checks whether differences in Table 1 were signficant! **)
-In the case of two non-living C stocks ($DW_{down}$ and $OL$)
+(**FINISH THIS!!!**)
+In the case of two non-living C stocks ($DW_{down}$ and $OL$), ..**FINISH THIS!!!**
 
@@
 
-There are of course notable holes in the ForC variable coverage (Fig. 1), that limit the scope of our inferences here. 
+There are of course notable holes in the ForC variable coverage (Fig. 1) that limit the scope of our inferences here. 
 Notably, *ForC* currently has sparse--if any--coverage of fluxes to herbivores and higher consumers, along with the woody mortality ($M_{woody}$) and dead wood stocks (Tables 1, Figs. S23-S25). 
@@
 *ForC* does not include soil carbon, which is covered by other efforts [e.g., @kochy_global_2015]. 
-*ForC* is not intended to replace databases that are specialized for particular parts of the C cycle analyses, e.g., aboveground biomass (**REFS**), land-atmosphere fluxes [@baldocchi_fluxnet_2001], soil respiration [@jian_reconstructed_2020], or the human footprint in global forests [@magnani_human_2007].
+*ForC* is not intended to replace databases that are specialized for particular parts of the C cycle analyses, e.g., aboveground biomass [@spawn_harmonized_2020], land-atmosphere fluxes [@baldocchi_fluxnet_2001], soil respiration [@jian_reconstructed_2020], or the human footprint in global forests [@magnani_human_2007].
 
@@
 Unlike the other C flux variables, $NEP$ does not represent the rapidity with which C cycles through the ecosystem, but is the balance between C sequestration ($GPP$) and respiratory losses ($R_{eco}$) and represents net CO~2~ sequestration (or release) by the ecosystem.
-$NEP$ tends to be relatively small in mature forest stands [@luyssaert_old-growth_2008, **MORE REFS?**; discussed further below], which accumulate carbon slowly relative to younger stands [@luyssaert_old-growth_2008; **REFS**), if at all (**REFS**).
-It is therefore unsurprising that there are no pronounced differences across biomes, suggesting that variation in $NEP$ of mature forests is controlled less by climate and more by other factors including moderate disturbances (**REFS**) or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition; **REFS**].
+$NEP$ tends to be relatively small in mature forest stands (discussed further below), which accumulate carbon slowly relative to younger stands, if at all [@luyssaert_old-growth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
+It is therefore unsurprising that there are no pronounced differences across biomes, suggesting that variation in $NEP$ of mature forests is controlled less by climate and more by other factors including moderate disturbances [@curtis_forest_2018] or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition; @wilson_stability_2018].
 
@@
 
-As climate change accelerates, understanding and managing the carbon dynamics of forests will be critical to forecasting, mitigation, and adaptation.
+### Conclusions
+As climate change accelerates, understanding and managing the carbon dynamics of forests--notably including dynamics and fluxes that cannot be observed by satellites--will be critical to forecasting, mitigation, and adaptation.
 The C data in *ForC*, as summarized here, will be valuable to these efforts.
+Notably, the fact that tropical forests tend to have both the highest rates of C sequestration in young stands [Fig.
+7; @cook-patton_mapping_2020], fueled by their generally high C flux rates (Table 1; Fig.
+6), and the highest mean biomass [Fig.
+7; Table 1; @hu_mapping_2016; @jian_reconstructed_2020] reinforces the concept that conservation and restoration of these forests is a priority for climate change mitigation **(REFS)**, along with high-biomass old-growth temperate stands [@goldstein_protecting_2020].
+It is also important to note the trade-off in climate mitigation potential of restoration of young forests, with high rates of CO~2~ sequestration [$NEP$; @cook-patton_mapping_2020], versus conservation and management of mature forests, with low $NEP$ but high C stocks that could not be recovered on a time scale relevant to climate change mitigation [@goldstein_protecting_2020].
+Generally speaking, the conservation of mature forests will yield greater climate benefits [@anderson-teixeira_greenhouse_2011], but both approaches are critical to avoiding catastrophic climate change (**REFS**).
 

### 2020-09-30 manuscript/ERL_global_C_SI.Rmd
decent draft of SI
- wrote appendix S1
- removed repeating variable names in table S2 (could still use some formatting)

@@
 
-ForC v3.0 contains potential duplicates that have arisen from importing intermediary data sets with overlap in data. 
-In particular, the recent imports of SRDB and GROA resulted in the introduction of numerous potential duplicates. 
-SRDB and ForC were developed independently (but with different foci), resulting in numerous duplicates.
-The GROA database incorporated ForC data. 
-For records retaining the same site names, GROA duplicates were not imported into ForC (although they were checked against corresponding ForC records, and differences reconciled).
-There were, however, some records entered into GROA independently of ForC, resulting in mismatched names.
+*ForC* v3.0 contains potential duplicates that have arisen from importing intermediary data sets with overlap in data. 
+In particular, the recent imports of *SRDB* and *GROA* resulted in the introduction of numerous potential duplicates. 
+*SRDB* and *ForC* were developed independently (but with different foci), resulting in numerous duplicates.
+The *GROA* database was developed semi-independently, incorporating *ForC* data partway through its development. 
+For records retaining the same site names, *GROA* duplicates were not imported into *ForC* (although they were checked against corresponding *ForC* records, and differences reconciled).
+There were, however, some records entered into *GROA* independently of *ForC*, resulting in mismatched names.
 
-Prior to the merging of the three databases (ForC v2.0, SRDB, and GROA), there were potential duplicates within the ForC and SRDB portions, the latter an artifact of how data were recorded in SRDB and imported into ForC.
-GROA has no known duplicates, with several steps having been taken to ensure independence of the records within the database.
+Prior to the merging of the three databases (*ForC* v2.0, *SRDB*, and *GROA*), there were potential duplicates within the *ForC* and *SRDB* portions, the latter an artifact of how data were recorded in *SRDB* and imported into *ForC.*
+*GROA* has no known duplicates, with several steps having been taken to ensure independence of the records within the database.
 
-### Detecting and reconciling duplicate sites
+To the extent possible, we used automated scripts in R to detect and remove duplicate records in the creation of *ForC-simplified*.
 
-To the extent possible, we used automated scripts in R to detect duplicate records.
-Potential duplicates were defined as geographically proximate records for stands of **similar age** with the same variable measured in the same year (if known).
-In cases where site name, plot name, and stand age matched, duplicate status could be automatically detected and reconciled (see below).
-In cases where site and plot names or reported age differed, our script detected potential duplicate sites that were **geographically proximate**.
-When all such potential duplicate sites were within GROA, they were assumed independent.
+### Handling within-plot duplicates
 
-Suspected duplicates require manual review and merging of records, and this process has not been completed for all records. 
-To handle potential duplicates at sites that have yet to be reviewed, we implemented the following procedure. 
-In cases where a single location -- generally an established research site where multiple investigators have worked -- contained multiple plots in nested or unknown relation to one another, we grouped multiple sites into a "supersite" (e.g., Harvard Forest, Barro Colorado Island, Pasoh Forest Reserve), and duplicates within a supersite were handled in the same way as records with matching site and plot names.
-(**VALENTINE, IS THIS ACCURATE?**)
-For suspected duplicate groups that were not flagged as supersites and had not yet been reviewed, we retained only one potential duplicate record, assigning precedence as follows: (1) original GROA record(s), (2) record(s) in ForC prior to SRDB and GROA import, (3) SRDB record(s).
+Duplicate records within a plot and taken in the same year were reconciled as follows.
+Replicate measurements (*i.e.*, replicates from within a single study) were averaged. 
+Records that subsumed others—- *i.e.*, the time period included that of $\ge$ 2 other records or dates were unknown and therefore conflicted with $\ge$ 2 other records—were removed. 
+For each group of duplicate records—*i.e.*, measurements of the same variable in the same plot at the same time—one record was assigned precedence as follows. 
+If duplicates were exact matches, one was dropped. 
+When measurement periods overlapped or were not specified, precedence was given first to records representing longer measurement periods (*i.e.*, multi-year mean over a single year) and then to more recently published values.
+We manually reviewed duplicates that differed only in methodology, assigning precedence to the record employing a more comprehensive approach (*e.g.*, inclusion of understory, lianas, or bamboo as opposed to just trees) or using a favored methodology.
 
-Finally, ....
-https://github.com/forc-db/ERL-review/issues/34
+### Handling duplicate sites
 
-### Generating ForC_simplified
-Detected duplicates were reconciled as follows.
-Replicate measurements (*i.e.*, replicates from within a single study) were averaged. 
-Records that subsumed others— **i.e.**, the time period included that of $\ge$ 2 other records or dates were unknown and therefore conflicted with $\ge$ 2 other records—were removed. 
-For each group of duplicate records—*i.e.*, measurements of the same variable in the same plot at the same time—one record was assigned precedence (recorded in D.precedence field). 
-If duplicates were exact matches, **one was dropped**. 
-When measurement periods overlapped or were not specified, precedence was given first to records representing longer measurement periods (*i.e.*, end.date - start.date) and then to more recently published values.
-We manually reviewed duplicates that differed only in methodology, assigning precedence to the record employing a more comprehensive approach (*e.g.*, inclusion of understory, lianas, or bamboo as opposed to just trees) or using a favored methodology.
+To handle cases where duplicate records had different site and/or plot names (and often slightly different geographic coordinates), we detected potential duplicates and, for those that had not yet been resolved, included only one in *ForC-simplified* and the analyses that were based on it.
+Potential duplicates were defined as records for the same variable made in the same year (when known) on stands of similar age and in the same geographic area (see Methods).
+Because records of stand age often lack precision and  were sometimes recorded differently in different intermediary databases, we considered records to be potential duplicates when ages fell within one of several broad age groups: <10, 10-50, 50-100, and >100 years (or mature and of undefined age).
+We then generated groups of potential duplicates grouped by variable, geographic area, stand age grouping, and measurement year.
+
+Within each group, we flagged records as potential duplicates--and excluded them from *ForC-simplified* until they can be reconciled--as follows.
+Because *GROA* is believed to contain no duplicates, all records originating from *GROA* were assumed independent, and any other records within the group flagged as potential duplicates.
+If all records originated from *ForC v2.0* or *SRDB*, we gave precedence to *ForC* records, flagging *SRDB* records as potential duplicates.
+If duplicates remained (within *ForC v2.0* or *SRDB*), we selected a record for retention based on citation year, keeping the oldest (presumably original) citations for most variables. 
+For eddy-covariance variables ($NEP$, $GPP$, $R_{eco}$), we retained the record associated with the most recent publication, as these data are commonly re-assessed using new analysis methods.
+If publication years were the same, we randomly selected one record.
+
+This approach was conservative in that it removed some legitimately independent records. 
+Future refinement of this process and manual review of potential duplicates will be needed to include legitimately-independent records in ForC-simplified and in analyses.
 
@@
 library(kableExtra)
-table_pt <- read.csv("tables_figures/SI_age_trend_model_summaries.csv", 
+table_pt <- read.csv("tables_figures/SI_age_trend_model_summaries_manually_edited.csv", 
                      #colClasses=c("NULL", NA, NA, NA, NA), 

### 2020-09-30 manuscript/ERL_global_C_review.Rmd
minor cleanup, citations

@@
 
-*Discussion.*  **NEED TO WRITE THIS!!!**
+*Discussion.* As climate change accelerates, understanding and managing the carbon dynamics of forests is critical to forecasting, mitigation, and adaptation.
+The synthetic and internally consistent overview of C stocks and fluxes across biomes and stand ages will help to advance these efforts.
 
@@
 
-Forest ecosystems are shaping the course of climate change (IPCC1.5) through their influence on atmospheric carbon dioxide (CO~2~).
+Forest ecosystems are shaping the course of climate change [@ipcc_global_2018] through their influence on atmospheric carbon dioxide (CO~2~).
 Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climatecarbon_2006; @krause_large_2018; @bonan_model_2019; @di_vittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history.
 These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published.
-While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass: @hu_mapping_2016; **REF**; gross primary productivity, $GPP$: @li_mapping_2019; **Saatchi 2011; REFS for biomass change, net CO2 flux**], measurement and validation of most forest C stocks and fluxes require intensive on-the-ground data collection.
+While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass: @hu_mapping_2016; @spawn_harmonized_2020; gross primary productivity, $GPP$: @li_mapping_2019; @saatchi_benchmark_2011], measurement and validation of most forest C stocks and fluxes require intensive on-the-ground data collection.
 
 The future of the current forest C sink is dependent both upon forest responses to climate change itself and human land use decisions, which will feedback and strongly influence the course of climate change.
-Regrowing forests in particular will play an important role [@pugh_role_2019], as almost two-thirds of the world’s forests were secondary as of 2010 (**FAO 2010**)
+Regrowing forests in particular will play an important role [@pugh_role_2019], as almost two-thirds of the world’s forests were secondary as of 2010 [@fao_global_2010].
 As anthropogenic and climate-driven disturbances impact an growing proportion of Earth's forests [@andela_human-driven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests is increasingly important.
@@
 Such understanding is particularly critical for reducing uncertainty regarding the potential for carbon uptake and climate change mitigation by regrowth forests [@kruase_large_2018; @cook-patton_mapping_2020].
-Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thus central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; **Cavaleri et al 2015**].
+Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thus central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; @cavaleri_urgent_2015].
 
@@
 Annual gross CO~2~ sequestration in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
-Most of this enormous C sequestration is counterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) *or fire*, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] *and fire emissions [@van_der_werf_global_2017]*.
-(**check if deforestation statistic below includes natural fire, exclude here if it does.**)
+Most of this enormous C sequestration is counterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] and fire emissions [@van_der_werf_global_2017].
 In recent years, the remaining CO~2~ sink averaged 3.2 $\pm$ 0.6 GtC yr^-1^ from 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
-Yet, this sink is reduced by *deforestation/ forest losses to anthropogenic and natural disturbances*.
-Recent net deforestation (*i.e.*, gross deforestation minus regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr^-1^ from YEAR-YEAR (Pan et al 2011; **UPDATE**), reducing the net forest sink to ~1.1-2.2 Gt C yr^-1^ across Earth's forests [@friedlingstein_global_2019].
+Yet, this sink is reduced by deforestation, estimated at 3.7 Gt CO~2~ yr-1 ranging from 1990-2020
+Recent net deforestation (*i.e.*, gross deforestation minus regrowth) has been a source of CO~2~ emissions, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reducing the net forest sink to ~1.1-2.2 Gt C yr^-1^ across Earth's forests [@friedlingstein_global_2019].
 
@@
 To facilitate analyses, we created a simplified version of ForC, *ForC-simplified* (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which we analyzed here.
-In generating *ForC-simplified*, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006).
+In generating *ForC-simplified*, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ [@ipcc_global_2018].
 Duplicate or otherwise conflicting records were reconciled as described in *APPENDIX S1*, resulting in a total of `r nrow(ForC_simplified)` records (`r round(nrow(ForC_simplified)/nrow(ForC_measurements)*100,digits=1)`% size of total database).
@@
 For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity, such as measurements including or excluding fruit and flower production and herbivory) into more broadly defined variables.
-Specifically, net ecosystem exchange (measured by eddy-covariance; **FLUXNET REF**) and biometric estimates of $NEP$ were combined into the single variable $NEP$ (Table 1).
+Specifically, net ecosystem exchange [measured by eddy-covariance; @baldocchi_fluxnet_2001] and biometric estimates of $NEP$ were combined into the single variable $NEP$ (Table 1).
 Furthermore, for $NPP$, $ANPP$, and $ANPP_{litterfall}$, *ForC* variables specifying inclusion of different components were combined.
@@
 We defined leaf type (broadleaf / needleleaf) was based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
-For young tropical forests imported from *GROA* but not yet classified by leaf type, we assumed that all were broadleaf, consistent with the rarity of naturally regenerating needleleaf forests in the tropics (**REF**).
+For young tropical forests imported from *GROA* but not yet classified by leaf type, we assumed that all were broadleaf, consistent with the rarity of naturally regenerating needleleaf forests in the tropics.
 We also classified forests as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication).
@@
 In terms of C stocks, ten variables (all but standing deadwood, $DW_{standing}$) had sufficient data to test for age trends (Table 1, Figs. 7, S16-26).
-All of these increased significantly with $log10[stand.age]$.
-There were sufficient data to model age x biome interactions were also significant for all ten of these C stock variables (Table S2). 
-(**FINISH THIS!!!**)
-In the case of two non-living C stocks ($DW_{down}$ and $OL$), ..**FINISH THIS!!!**
+All of these displayed a significant overall increase with with $log10[stand.age]$.
+There were sufficient data to model age x biome interactions were also significant for all ten of these C stock variables (Table S2), with living C stocks tending to accumulate more rapidly during the early stages of forest regrowth in tropical forests (Figs.
+7, S16-S22).
+In the case of two non-living C stocks ($DW_{down}$ and $OL$), age x biome interactions were such that 
+Specifically, $DW_{down}$ declined with age in temperate and boreal forests, compared to an increase with age in tropical forests (Figs. 7, S25).
+Similarly, $OL$ declined slightly with age in temperate broadleaf forests, contrasting an increase in the other three biomes (Figs. 7, S26).
 
@@
 
-For regrowth forests, little is known about cross-biome differences in carbon fluxes, and we are not aware of any previous large-scale comparisons of C fluxes that have been limited to regrowth forests [although they're commonly mixed in with mature forests; e.g., @luyssaert_co_2007 **REFS**].
+For regrowth forests, little is known about cross-biome differences in carbon fluxes, and we are not aware of any previous large-scale comparisons of C fluxes that have been limited to regrowth forests.
 Thus, this analysis was the first to examine flux trends in regrowth forests across biomes.
@@
 For aboveground biomass, which is the variable in *ForC* with broadest geographical representation, the trend of declining biomass from tropical to boreal regions mirrors observations from spaceborne lidar that reveal a decline in aboveground biomass (for all forests, including secondary) with latitude across the N hemisphere [@hu_mapping_2016].
-Yet, the highest- biomass forests on Earth are found in coastal temperate climates of both the southern and northern hemisphere [@keith_re-evaluation_2009; @smithwick_potential_2002; **REF??**; @hu_mapping_2016].
+The highest- biomass forests on Earth are however found in coastal temperate climates of both the southern and northern hemisphere [@keith_re-evaluation_2009; @smithwick_potential_2002; **REF??**; @hu_mapping_2016].
 Disproportionate representation of forests in one such region--the US Pacific Northwest--inflated estimates of temperate conifer fluxes and stocks for some variables and was responsible for all of the anomalous results described here (e.g., lack of complete C budget closure, anomalous trend across biomes for $BNPP_{coarse}$).
@@
 
-Whereas biomass can be remotely sensed and receives a lot of research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
+Whereas biomass can be remotely sensed and receives significant research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
 Although these stocks can be important--exceeding 100 t C ha^-1^ in some stands (Figs.
@@
 Our study reveals that most C fluxes quickly increase to a plateau as stands age (Fig.
-6), consistent with current understanding of age trends in forest C cycling [e.g., @anderson-teixeira_altered_2013].
-*Quickly-changing and age-dependent fluxes were also found in a number of previous syntheses (Amiro et al. 2010 10.1029/2010JG001390, Magnani et al. 2007 10.1038/nature05847).*
+6), consistent with current understanding of age trends in forest C cycling [e.g., @anderson-teixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
 While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_auto$) would be minimal immediately following a stand-clearing disturbance.
@@
 Notably, net carbon sequestration ($NEP$) increases with age up to the 100-yr threshold examined here, with more pronounced patterns in temperate than boreal forests (Fig. 6f). 
-This finding is largely consistent with, but built from a far larger dataset than, previous studies showing an increase in $NEP$ across relatively young stand ages [**Pregitzer and Euskirchen (2004 http://dx.doi.org/10.1111/j.1365-2486.2004.00866.x); Baldocchi; Luyssaert**].
-However, $NEP$ has been observed to decline from intermediate to old stands **[Luyssaert 2008]**, and the $NEP$ estimated by our model for 100-year-old temperate conifer stands (~5 Mg C ha^-1^ yr^-1^) exceeds the mean of mature forests in the same biome (0.7 Mg C ha^-1^ yr^-1^; Fig.4).
-A decrease in $NEP$ is consistent with the observed decelleration of biomass accumulation as stands age, although both biomass and non-living C stocks will often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands (**REFS**).
+This finding is largely consistent with, but built from a far larger dataset than, previous studies showing an increase in $NEP$ across relatively young stand ages [@pregitzer_carbon_2004; @baldocchi_fluxnet_2001;@luyssaert_old-growth_2008].
+However, $NEP$ has been observed to decline from intermediate to old stands [@luyssaert_old-growth_2008], and the $NEP$ estimated by our model for 100-year-old temperate conifer stands (~5 Mg C ha^-1^ yr^-1^) exceeds the mean of mature forests in the same biome (0.7 Mg C ha^-1^ yr^-1^; Fig.4).
+A decrease in $NEP$ is consistent with the observed deceleration of biomass accumulation as stands age, although both biomass and non-living C stocks will often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands [@luyssaert_old-growth_2008; @mc_garvey_carbon_2014; @lichstein_biomass_2009].
 
-In terms of stocks, our study reveals consistent increases in live biomass stocks with stand age--a pattern that is well-known and expected [e.g., **REFS**]--and more variable age trends in deadwood and $OL$.
-The latter are particularly sensitive to the type of disturbance, where disturbances that remove deadwood (e.g., logging, agriculture) result in negligible deadwood in young stands, followed by a buildup over time [e.g., tropical stands in Fig.
-7e **REFS**].
+In terms of stocks, our study reveals consistent increases in live biomass stocks with stand age--a pattern that is well-known and expected [e.g., @lichstein_biomass_2009; @yang_global_2011]--and more variable age trends in deadwood and $OL$.
+The latter are particularly sensitive to the type of disturbance, where disturbances that remove most organic material (e.g., logging, agriculture) result in negligible deadwood in young stands, followed by a buildup over time [tropical stands in Fig.
+7e; e.g., @vargas_biomass_2008].
 In contrast, natural disturbances (e.g., fire, drought) can produce large amounts of deadwood (mostly $DW_{standing}$) that slowly decomposes as the stand recovers, resulting in declines across young stand ages [e.g., temperate and boreal stands in Fig.
-7e; **REF**].
+7e; e.g., @carmona_coarse_2002].
 Again, further study and synthesis of non-living C stocks across biomes and stand ages will be valuable to giving a more comprehensive picture.
@@
 
-The future of forest C cycling [@song_meta-analysis_2019] will shape trends in atmospheric CO~2~ and the course of climate change (**Schimel et al. 2015 10.1073/pnas.1407302112**).
+The future of forest C cycling [@song_meta-analysis_2019] will shape trends in atmospheric CO~2~ and the course of climate change [@schimel_effect_2015].
 Our findings, and more generally the data contained in *ForC* and summarized here, can help to meet two major challenges.
 
-First, improved representation of forest C cycling in models is essential to improving predictions of the future course of climate change, for the simple reason that by definition future projections extend our existing observations and understanding to conditions that do not currently exist on Earth (**https://doi.org/10.1016/j.tree.2017.10.002, https://doi.org/10.1126/science.aam8328, https://link.springer.com/article/10.1186/s13717-018-0142-8**).
-To ensure that models are giving the right answers for the right reasons (Sulman et al.
-2018 10.1007/s10533-018-0509-z), it is important to benchmark against multiple components of the C cycle that are internally consistent with each other (Collier et al.
-2018 10.1029/2018MS001354, Wang et al.
-2018 10.5194/gmd-11-3903-2018).
+First, improved representation of forest C cycling in models is essential to improving predictions of the future course of climate change, for the simple reason that by definition future projections extend our existing observations and understanding to conditions that do not currently exist on Earth [@mcdowell_predicting_2018; @bonan_climate_2018; @gustafson_extrapolating_2018].
+To ensure that models are giving the right answers for the right reasons [@sulman_multiple_2018], it is important to benchmark against multiple components of the C cycle that are internally consistent with each other [@collier_international_2018; @wang_golum-cnp_2018].
 *ForC*’s tens of thousands of records are readily available in a standardized format, and our analyses here indicate their internal consistency is reasonably high.
-Integration of ForC with models is a goal (Fer et al., in revision).
+Integration of *ForC* with models will be valuable to improving the accuracy and reliability of models.
 
-**(Ben would remove this paragraph:)**
 Second, *ForC* can serve as a pipeline through which forest science can inform forest-based climate change mitigation efforts. 
@@
 One goal for *ForC* is to serve as a pipeline through which information can flow efficiently from forest researchers to decision-makers working to implement forest conservation strategies at global, national, or landscape scales.
-This is already happening: *ForC* has contributed to updating the IPCC guidelines for carbon accounting in forests [**IPCC 2019**; @requena_suarez_estimating_2019], mapping C accumulation potential from natural forest regrowth globally [@cook-patton_mapping_2020], and informing ecosystem conservation priorities [@goldstein_protecting_2020].
+This is already happening: *ForC* has contributed to updating the IPCC guidelines for carbon accounting in forests [@ipcc_2019_2019; @requena_suarez_estimating_2019], mapping C accumulation potential from natural forest regrowth globally [@cook-patton_mapping_2020], and informing ecosystem conservation priorities [@goldstein_protecting_2020].
 
 It is also interesting to consider the complementary utility of global-scale but spatially discontinuous databases such as ForC and remote wall-to-wall remote sensing products.
-The latter provide unparalleled insight into aboveground carbon stocks, but less constraint on belowground stocks or carbon fluxes in general (Bond-Lamberty et al.
-2016, http://dx.doi.org/10.1002/ecs2.1380, Anav et al.
-2015 http://dx.doi.org/10.1002/2015RG000483).
-Combining observational data and remote observations may provide a much more comprehensive and accurate picture of global forest C cycling, particularly when used in formal data assimilation systems (Konings et all 2019 10.5194/bg-16-2269-2019, Liu et al.
-2018 10.1088/1748-9326/aad5ef).
-biomass is the largest stock, and most of the emphasis has traditionally been on this variable. 
-Remote-sensing driven biomass estimates (e.g.
-Saatchi et al.
-2011 http://dx.doi.org/10.1073/pnas.1019576108), calibrated based on high-quality ground-based data [@schepaschenko_forest_2019; @chave_ground_2019], are well suited for this task.
-Note however that factors such as stand age and disturbance history (except in very recent decades are important but difficult to detect this way [@hansen_high-resolution_2013; @song_global_2018; @curtis_classifying_2018].
-Ground-based data such as *ForC* are therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020, and thus constraining variables such as carbon sink potential (Luyssaert et al.
-2007 http://dx.doi.org/10.1038/nature07276).
+The latter provide unparalleled insight into aboveground carbon stocks, but less constraint on belowground stocks or carbon fluxes in general [@bond-lamberty_estimating_2016; @anav_spatiotemporal_2016].
+Combining observational data and remote observations may provide a much more comprehensive and accurate picture of global forest C cycling, particularly when used in formal data assimilation systems [@konings_global_2019; @liu_detection_2018].
+Biomass is the largest C stock in most forests, and most of the emphasis has traditionally been on this variable. 
+Remote-sensing driven biomass estimates [e.g., @saatchi_benchmark_2011], calibrated based on high-quality ground-based data [@schepaschenko_forest_2019; @chave_ground_2019], are well suited for this task.
+Note, however, that factors such as stand age and disturbance history are difficult, if possible, to detect remotely, and can only be characterized for very recent decades [@hansen_high-resolution_2013; @song_global_2018; @curtis_classifying_2018].
+Ground-based data such as *ForC* are therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020, and thus constraining variables such as carbon sink potential [@luyssaert_old-growth_2008].
 
-In contrast, respiration fluxes cannot be remotely sensed. 
-Efforts such as the Global Carbon Project (le quere http://dx.doi.org/10.5194/essd-11-1783-2019) and NASA’s Carbon Monitoring System (**Liu et al.
-2018 10.1088/1748-9326/aad5ef**) typically compute respiration as residuals of all other terms [@bond-lamberty_estimating_2016;  Harmon et al.
-2011 http://dx.doi.org/10.1029/2010JG001495].
+In contrast, carbon allocation within forest ecosystems and respiration fluxes cannot be remotely sensed. 
+Efforts such as the Global Carbon Project [@friedlingstein_global_2019] and NASA’s Carbon Monitoring System [@liu_detection_2018] typically compute respiration as residuals of all other terms [@bond-lamberty_estimating_2016; @harmon_heterotrophic_2011].
 This means that the errors on respiration outputs are likely to be large and certainly poorly constrained, offering a unique opportunity for databases such as ForC and SRDB [@jian_restructured_2020] to provide observational benchmarks.
-For example, **Konings et al.
-(2019 http://dx.doi.org/10.5194/bg-16-2269-2019)** produced a unique top-down estimate of global heterotrophic respiration that can both be compared with extant bottom-up estimates **(Bond-Lamberty et al.
-2018 http://dx.doi.org/10.1029/2018EF000866)** and used as an internal consistency check on other parts of the carbon cycle **(Phillips et al.
-2016)**.
+For example, @konings_global_2019 produced a unique top-down estimate of global heterotrophic respiration that can both be compared with extant bottom-up estimates [@bond-lamberty_new_2018] and used as an internal consistency check on other parts of the carbon cycle [@phillips_value_2016].
 
 ### Conclusions
-As climate change accelerates, understanding and managing the carbon dynamics of forests--notably including dynamics and fluxes that cannot be observed by satellites--will be critical to forecasting, mitigation, and adaptation.
+As climate change accelerates, understanding and managing the carbon dynamics of forests--notably including dynamics and fluxes that cannot be observed by satellites--is critical to forecasting, mitigation, and adaptation.
 The C data in *ForC*, as summarized here, will be valuable to these efforts.
@@
 6), and the highest mean biomass [Fig.
-7; Table 1; @hu_mapping_2016; @jian_reconstructed_2020] reinforces the concept that conservation and restoration of these forests is a priority for climate change mitigation **(REFS)**, along with high-biomass old-growth temperate stands [@goldstein_protecting_2020].
+7; Table 1; @hu_mapping_2016; @jian_reconstructed_2020] reinforces the concept that conservation and restoration of these forests is a priority for climate change mitigation [@grassi_key_2017], along with high-biomass old-growth temperate stands [@goldstein_protecting_2020].
 It is also important to note the trade-off in climate mitigation potential of restoration of young forests, with high rates of CO~2~ sequestration [$NEP$; @cook-patton_mapping_2020], versus conservation and management of mature forests, with low $NEP$ but high C stocks that could not be recovered on a time scale relevant to climate change mitigation [@goldstein_protecting_2020].
-Generally speaking, the conservation of mature forests will yield greater climate benefits [@anderson-teixeira_greenhouse_2011], but both approaches are critical to avoiding catastrophic climate change (**REFS**).
+Generally speaking, the conservation of mature forests will yield greater climate benefits [@anderson-teixeira_greenhouse_2011], but both approaches are critical to avoiding catastrophic climate change [@ipcc_global_2018].
 
@@
 
-## ORCID iD
-Kristina J. Anderson-Teixeira: https://orcid.org/0000-0001-8461-9713
-
 ## References

### 2020-09-30 manuscript/ERL_global_C_review.Rmd
worked on refs

@@
 Although age trends in aboveground biomass have been relatively well-studied and synthesized globally [@cook-patton_mapping_2020], a relative dearth of data and synthesis on other C stocks and fluxes in secondary forests points to an under-filled need to characterize age-related trends in forest C cycling.
-Such understanding is particularly critical for reducing uncertainty regarding the potential for carbon uptake and climate change mitigation by regrowth forests [@kruase_large_2018; @cook-patton_mapping_2020].
+Such understanding is particularly critical for reducing uncertainty regarding the potential for carbon uptake and climate change mitigation by regrowth forests [@krause_large_2018; @cook-patton_mapping_2020].
 Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thus central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; @cavaleri_urgent_2015].
@@
 Primary research articles typically cover only a small numbers of sites at a time.
-The rare exceptions that span regions or continents with rare exceptions spanning regions or continents are typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; e.g., @lutz_global_2018], NEON [@schimel_NEON_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
+The rare exceptions that span regions or continents with rare exceptions spanning regions or continents are typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015
+], NEON [@schimel_neon_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
 The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*.
@@
 *ForC* does not include soil carbon, which is covered by other efforts [e.g., @kochy_global_2015]. 
-*ForC* is not intended to replace databases that are specialized for particular parts of the C cycle analyses, e.g., aboveground biomass [@spawn_harmonized_2020], land-atmosphere fluxes [@baldocchi_fluxnet_2001], soil respiration [@jian_reconstructed_2020], or the human footprint in global forests [@magnani_human_2007].
+*ForC* is not intended to replace databases that are specialized for particular parts of the C cycle analyses, e.g., aboveground biomass [@spawn_harmonized_2020], land-atmosphere fluxes [@baldocchi_fluxnet_2001], soil respiration [@jian_restructured_2020], or the human footprint in global forests [@magnani_human_2007].
 
@@
 Our analysis reveals a general deceleration of carbon cycling from the tropics to boreal regions.
-For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude (or increase with temperature) on a global scale [e.g., @luyssaert_co_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
+For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude (or increase with temperature) on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
 The consistency with which this occurs across numerous fluxes is not surprising, but has never been simultaneously assessed across such a large number of variables [but see @banbury_morgan_global_nodate for nine autotrophic fluxes].
@@
 $NEP$ tends to be relatively small in mature forest stands (discussed further below), which accumulate carbon slowly relative to younger stands, if at all [@luyssaert_old-growth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
-It is therefore unsurprising that there are no pronounced differences across biomes, suggesting that variation in $NEP$ of mature forests is controlled less by climate and more by other factors including moderate disturbances [@curtis_forest_2018] or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition; @wilson_stability_2018].
+It is therefore unsurprising that there are no pronounced differences across biomes, suggesting that variation in $NEP$ of mature forests is controlled less by climate and more by other factors including moderate disturbances [@curtis_forest_2018] or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition; @wilson_stability_2016].
 
@@
 In contrast to C fluxes and biomass accumulation rates in regrowth forests, stocks showed less systematic variation across biomes.
-This is consistent with others studies showing that variation in forest biomass across broad climatic gradients is modest and constrained more by moisture than temperature [@stegen_variation_2011].
-Yet, differences across biomes did exist.
-For aboveground biomass, which is the variable in *ForC* with broadest geographical representation, the trend of declining biomass from tropical to boreal regions mirrors observations from spaceborne lidar that reveal a decline in aboveground biomass (for all forests, including secondary) with latitude across the N hemisphere [@hu_mapping_2016].
-The highest- biomass forests on Earth are however found in coastal temperate climates of both the southern and northern hemisphere [@keith_re-evaluation_2009; @smithwick_potential_2002; **REF??**; @hu_mapping_2016].
+For aboveground biomass, which is the variable in *ForC* with broadest geographical representation, the modest trend of declining biomass from tropical to boreal regions mirrors observations from spaceborne lidar that reveal a decline in aboveground biomass (for all forests, including secondary) with latitude across the N hemisphere [@hu_mapping_2016].
+The highest- biomass forests on Earth are however found in coastal temperate climates of both the southern and northern hemisphere [@keith_re-evaluation_2009; @smithwick_potential_2002; @hu_mapping_2016].
 Disproportionate representation of forests in one such region--the US Pacific Northwest--inflated estimates of temperate conifer fluxes and stocks for some variables and was responsible for all of the anomalous results described here (e.g., lack of complete C budget closure, anomalous trend across biomes for $BNPP_{coarse}$).
@@
 These would be expected to increase rapidly with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Fig. 7).
-In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$) are expected to be non-zero following stand-clearing disturbance, although these may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots and litter [@ribeiro-kumara_how_2020; @mauer_carbon_2020; @bond-lamberty_contribution_2004].
+In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$) are expected to be non-zero following stand-clearing disturbance, although these may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots and litter [@ribeiro-kumara_how_2020; @maurer_carbon_2020; @bond-lamberty_contribution_2004].
 In this study, we detect no significant age trends in either variable.
@@
 Notably, net carbon sequestration ($NEP$) increases with age up to the 100-yr threshold examined here, with more pronounced patterns in temperate than boreal forests (Fig. 6f). 
-This finding is largely consistent with, but built from a far larger dataset than, previous studies showing an increase in $NEP$ across relatively young stand ages [@pregitzer_carbon_2004; @baldocchi_fluxnet_2001;@luyssaert_old-growth_2008].
+This finding is largely consistent with, but built from a far larger dataset than, previous studies showing an increase in $NEP$ across relatively young stand ages [@pregitzer_carbon_2004; @baldocchi_fluxnet_2001; @luyssaert_old-growth_2008].
 However, $NEP$ has been observed to decline from intermediate to old stands [@luyssaert_old-growth_2008], and the $NEP$ estimated by our model for 100-year-old temperate conifer stands (~5 Mg C ha^-1^ yr^-1^) exceeds the mean of mature forests in the same biome (0.7 Mg C ha^-1^ yr^-1^; Fig.4).
-A decrease in $NEP$ is consistent with the observed deceleration of biomass accumulation as stands age, although both biomass and non-living C stocks will often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands [@luyssaert_old-growth_2008; @mc_garvey_carbon_2014; @lichstein_biomass_2009].
+A decrease in $NEP$ is consistent with the observed deceleration of biomass accumulation as stands age, although both biomass and non-living C stocks will often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands [@luyssaert_old-growth_2008; @mcgarvey_carbon_2014; @lichstein_biomass_2009].
 
@@
 The latter provide unparalleled insight into aboveground carbon stocks, but less constraint on belowground stocks or carbon fluxes in general [@bond-lamberty_estimating_2016; @anav_spatiotemporal_2016].
-Combining observational data and remote observations may provide a much more comprehensive and accurate picture of global forest C cycling, particularly when used in formal data assimilation systems [@konings_global_2019; @liu_detection_2018].
+Combining observational data and remote observations may provide a much more comprehensive and accurate picture of global forest C cycling, particularly when used in formal data assimilation systems [@konings_global_2019; @liu_detecting_2018].
 Biomass is the largest C stock in most forests, and most of the emphasis has traditionally been on this variable. 
@@
 In contrast, carbon allocation within forest ecosystems and respiration fluxes cannot be remotely sensed. 
-Efforts such as the Global Carbon Project [@friedlingstein_global_2019] and NASA’s Carbon Monitoring System [@liu_detection_2018] typically compute respiration as residuals of all other terms [@bond-lamberty_estimating_2016; @harmon_heterotrophic_2011].
+Efforts such as the Global Carbon Project [@friedlingstein_global_2019] and NASA’s Carbon Monitoring System [@liu_detecting_2018] typically compute respiration as residuals of all other terms [@bond-lamberty_estimating_2016; @harmon_heterotrophic_2011].
 This means that the errors on respiration outputs are likely to be large and certainly poorly constrained, offering a unique opportunity for databases such as ForC and SRDB [@jian_restructured_2020] to provide observational benchmarks.

### 2020-09-30 manuscript/ERL_global_C_review.Rmd
refs

@@
 7, S23-S25; but see below for age trends), pointing to a need for more widespread quantification of both standing and downed deadwood.
-*ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that slower decomposition in colder climates results in more buildup of organic matter [@allen_2015].
+*ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that slower decomposition in colder climates results in more buildup of organic matter [@allen_global_2005].
 Further research on non-living C stocks in the world's forests will be essential to completing the picture.
@@
 Our study reveals that most C fluxes quickly increase to a plateau as stands age (Fig.
-6), consistent with current understanding of age trends in forest C cycling [e.g., @anderson-teixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
+6), consistent with current understanding of age trends in forest C cycling [e.g., @andersonteixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
 While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_auto$) would be minimal immediately following a stand-clearing disturbance.
@@
 It is also interesting to consider the complementary utility of global-scale but spatially discontinuous databases such as ForC and remote wall-to-wall remote sensing products.
-The latter provide unparalleled insight into aboveground carbon stocks, but less constraint on belowground stocks or carbon fluxes in general [@bond-lamberty_estimating_2016; @anav_spatiotemporal_2016].
+The latter provide unparalleled insight into aboveground carbon stocks, but less constraint on belowground stocks or carbon fluxes in general [@bondlamberty_estimating_2016; @anav_spatiotemporal_2015].
 Combining observational data and remote observations may provide a much more comprehensive and accurate picture of global forest C cycling, particularly when used in formal data assimilation systems [@konings_global_2019; @liu_detecting_2018].
@@
 In contrast, carbon allocation within forest ecosystems and respiration fluxes cannot be remotely sensed. 
-Efforts such as the Global Carbon Project [@friedlingstein_global_2019] and NASA’s Carbon Monitoring System [@liu_detecting_2018] typically compute respiration as residuals of all other terms [@bond-lamberty_estimating_2016; @harmon_heterotrophic_2011].
+Efforts such as the Global Carbon Project [@friedlingstein_global_2019] and NASA’s Carbon Monitoring System [@liu_detecting_2018] typically compute respiration as residuals of all other terms [@bondlamberty_estimating_2016; @harmon_heterotrophic_2011].
 This means that the errors on respiration outputs are likely to be large and certainly poorly constrained, offering a unique opportunity for databases such as ForC and SRDB [@jian_restructured_2020] to provide observational benchmarks.
-For example, @konings_global_2019 produced a unique top-down estimate of global heterotrophic respiration that can both be compared with extant bottom-up estimates [@bond-lamberty_new_2018] and used as an internal consistency check on other parts of the carbon cycle [@phillips_value_2016].
+For example, @konings_global_2019 produced a unique top-down estimate of global heterotrophic respiration that can both be compared with extant bottom-up estimates [@bondlamberty_new_2018] and used as an internal consistency check on other parts of the carbon cycle [@phillips_value_2017].
 
@@
 6), and the highest mean biomass [Fig.
-7; Table 1; @hu_mapping_2016; @jian_reconstructed_2020] reinforces the concept that conservation and restoration of these forests is a priority for climate change mitigation [@grassi_key_2017], along with high-biomass old-growth temperate stands [@goldstein_protecting_2020].
+7; Table 1; @hu_mapping_2016; @jian_restructured_2020] reinforces the concept that conservation and restoration of these forests is a priority for climate change mitigation [@grassi_key_2017], along with high-biomass old-growth temperate stands [@goldstein_protecting_2020].
 It is also important to note the trade-off in climate mitigation potential of restoration of young forests, with high rates of CO~2~ sequestration [$NEP$; @cook-patton_mapping_2020], versus conservation and management of mature forests, with low $NEP$ but high C stocks that could not be recovered on a time scale relevant to climate change mitigation [@goldstein_protecting_2020].

### 2020-09-30 manuscript/ERL_global_C_review.Rmd
cleanup

@@
 
-**Title:** Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database
+**Title:** Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the Global Forest Carbon (ForC) database
 
@@
 
-Becky Banbury Morgan^3^
+Becky Banbury Morgan^1,3^
 
@@
 ForC_simplified <- read.csv(paste0(path_to_ForC, "ForC_simplified/ForC_simplified.csv"), stringsAsFactors = FALSE, check.names = FALSE)
+ForC_simplified_duplicates_removed <- sum(ForC_simplified$suspected.duplicate ==0)
 
-n_no_history <- sum(ForC_simplified[,"history.no.info"]) #(relative to ForC_simplified)
-n_managed_in_ForC_simplified <- sum(ForC_simplified[,"managed"])
-n_disturbed_in_ForC_simplified <- sum(ForC_simplified[,"disturbed"])
-records_eligible_for_inclusion <- sum(ForC_simplified$managed ==0 & ForC_simplified$disturbed ==0 & ForC_simplified$history.no.info == 0)  #number in ForC_simplified that are not disturbed, managed, or no.history
+n_no_history <- sum(ForC_simplified$suspected.duplicate ==0 & ForC_simplified$history.no.info ==1) #(relative to ForC_simplified)
+n_managed_in_ForC_simplified <- sum(ForC_simplified$suspected.duplicate ==0 & ForC_simplified$managed ==1)
+n_disturbed_in_ForC_simplified <- sum(ForC_simplified$suspected.duplicate ==0 & ForC_simplified$disturbed ==1)
+records_eligible_for_inclusion <- sum(ForC_simplified$suspected.duplicate ==0 & ForC_simplified$managed ==0 & ForC_simplified$disturbed ==0 & ForC_simplified$history.no.info == 0)  #number in ForC_simplified that are not disturbed, managed, or no.history
 
@@
 
-*NOTES TO COAUTHORS: *
-
-- "**???**" indicates a reference that we have entered in the .Rmd file, but not yet in the bibliography file. Don't worry about those. (However, places with "REF" need references)
-- bold/ italic text indicates places where I know we need to go back and revisit/ fill in info
-
 \newpage
@@
 Specifically, we use `r n_records` *ForC* records from `r n_areas_tot` geographic locations representing `r nrow(T1)` C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests.
-For regrowth forests, we quantify age trends for all variables.
+For regrowth forests, we quantify age trends for all variables with sufficient data.
 
@@
 *Discussion.* As climate change accelerates, understanding and managing the carbon dynamics of forests is critical to forecasting, mitigation, and adaptation.
-The synthetic and internally consistent overview of C stocks and fluxes across biomes and stand ages will help to advance these efforts.
+This synthetic and internally consistent global overview of C stocks and fluxes across biomes and stand ages will help to advance these efforts.
 
@@
 
-Forest ecosystems are shaping the course of climate change [@ipcc_global_2018] through their influence on atmospheric carbon dioxide (CO~2~).
+Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide [CO~2~; @bonan_forests_2008; @friedlingstein_global_2019; @ipcc_global_2018].
 Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climatecarbon_2006; @krause_large_2018; @bonan_model_2019; @di_vittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history.
@@
 While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass: @hu_mapping_2016; @spawn_harmonized_2020; gross primary productivity, $GPP$: @li_mapping_2019; @saatchi_benchmark_2011], measurement and validation of most forest C stocks and fluxes require intensive on-the-ground data collection.
-
-The future of the current forest C sink is dependent both upon forest responses to climate change itself and human land use decisions, which will feedback and strongly influence the course of climate change.
-Regrowing forests in particular will play an important role [@pugh_role_2019], as almost two-thirds of the world’s forests were secondary as of 2010 [@fao_global_2010].
-As anthropogenic and climate-driven disturbances impact an growing proportion of Earth's forests [@andela_human-driven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests is increasingly important.
-Although age trends in aboveground biomass have been relatively well-studied and synthesized globally [@cook-patton_mapping_2020], a relative dearth of data and synthesis on other C stocks and fluxes in secondary forests points to an under-filled need to characterize age-related trends in forest C cycling.
-Such understanding is particularly critical for reducing uncertainty regarding the potential for carbon uptake and climate change mitigation by regrowth forests [@krause_large_2018; @cook-patton_mapping_2020].
-Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thus central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; @cavaleri_urgent_2015].
 
@@
 In recent years, the remaining CO~2~ sink averaged 3.2 $\pm$ 0.6 GtC yr^-1^ from 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
-Yet, this sink is reduced by deforestation, estimated at 3.7 Gt CO~2~ yr-1 ranging from 1990-2020
-Recent net deforestation (*i.e.*, gross deforestation minus regrowth) has been a source of CO~2~ emissions, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reducing the net forest sink to ~1.1-2.2 Gt C yr^-1^ across Earth's forests [@friedlingstein_global_2019].
+Yet, this sink is reduced by deforestation, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reducing the net forest sink to ~1.1-2.2 Gt C yr^-1^ across Earth's forests [@friedlingstein_global_2019].
+
+The future of the current forest C sink is dependent both upon forest responses to climate change itself and human land use decisions, which will feedback and strongly influence the course of climate change.
+Regrowing forests in particular will play an important role [@pugh_role_2019], as almost two-thirds of the world’s forests were secondary as of 2010 [@fao_global_2010].
+As anthropogenic and climate-driven disturbances impact an growing proportion of Earth's forests [@andela_human-driven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests is increasingly important [@andersonteixeira_altered_2013].
+Although age trends in aboveground biomass have been relatively well-studied and synthesized globally [@cook-patton_mapping_2020], a relative dearth of data and synthesis on other C stocks and fluxes in secondary forests points to an under-filled need to characterize age-related trends in forest C cycling.
+Such understanding is particularly critical for reducing uncertainty regarding the potential for carbon uptake and climate change mitigation by regrowth forests [@krause_large_2018; @cook-patton_mapping_2020].
+Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thus central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; @cavaleri_urgent_2015].
 
@@
 Primary research articles typically cover only a small numbers of sites at a time.
-The rare exceptions that span regions or continents with rare exceptions spanning regions or continents are typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015
-], NEON [@schimel_neon_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
-The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*.
+The rare exceptions that span regions or continents with rare exceptions spanning regions or continents are typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015, e.g., @lutz_global_2018], NEON [@schimel_neon_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
+The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*
 In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
 
-To address the need for global-scale analyses of forest C cycling, we recently developed *ForC* [@anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018].
+To address the need for global-scale analyses of forest C cycling, we recently developed *ForC* [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
 *ForC* contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables), with the different variables capturing the unique ecosystem pools (e.g., woody, foliage, and root biomass; dead wood) and flux types (e.g, gross and net primary productivity; soil, root, and ecosystem respiration).
 These data are ground-based measurements, and *ForC* contains associated data required for interpretation (*e.g.*, stand history, measurement methods).
-These data have been amalgamated from original peer-reviewed publications, either directly or via intermediary data compilations.
+Data have been amalgamated from original peer-reviewed publications, either directly or via intermediary data compilations.
 Since the its most recent publication [@anderson-teixeira_forc_2018], *ForC* has grown to include two additional large databases: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020] that also synthesized published forest C data.
@@
 
-This review synthesizes data from the *ForC* database [Fig. 1; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018].
+This review synthesizes data from the *ForC* database [Fig. 1; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
 *ForC* amalgamates numerous intermediary data sets [*e.g.*, @luyssaert_co2_2007; @bond-lamberty_global_2010; @cook-patton_mapping_2020] and original studies.
@@
 As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible [83% of records in v.2.0; @anderson-teixeira_forc_2018].
-Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], which represented roughly one-third of records in *ForC* v.2.0* [@anderson-teixeira_forc_2018].
-Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [*SRDB* v.##, `r n_SRDB` records; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database [*GROA* v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020].
-We've also added data from individual publications (detailed list at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv), with a particular focus on productivity [@banbury_morgan_global_nodate; e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @lutz_global_2018, @johnson_climate_2018].
+Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], which represented roughly one-third of records in *ForC* v2.0 [@anderson-teixeira_forc_2018].
+Since publication of ForC v2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [*SRDB* v4, `r n_SRDB` records; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database [*GROA* v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020].
+We have also added data from individual publications, with a particular focus on productivity [@banbury_morgan_global_nodate; e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @lutz_global_2018, @johnson_climate_2018].
 The database version used for this analysis has been tagged as a new release on Github (v3.0) and assigned a DOI through Zenodo (DOI: TBD).
@@
 In generating *ForC-simplified*, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ [@ipcc_global_2018].
-Duplicate or otherwise conflicting records were reconciled as described in *APPENDIX S1*, resulting in a total of `r nrow(ForC_simplified)` records (`r round(nrow(ForC_simplified)/nrow(ForC_measurements)*100,digits=1)`% size of total database).
+Duplicate or otherwise conflicting records were reconciled as described in Appendix S1, resulting in a total of `r ForC_simplified_duplicates_removed` records (`r round(ForC_simplified_duplicates_removed/nrow(ForC_measurements)*100,digits=1)`% size of total database).
 Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand initiation event.
-Specifically, we removed all plots flagged as managed in ForC-simplified (`r round(n_managed_in_ForC_simplified/nrow(ForC_simplified)*100,digits=1)`%).
+Specifically, we removed all plots flagged as managed in ForC-simplified (`r round(n_managed_in_ForC_simplified/ForC_simplified_duplicates_removed*100,digits=1)`%).
 This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.
-Plots flagged as disturbed in ForC-simplified (`r round(n_disturbed_in_ForC_simplified/nrow(ForC_simplified)*100,digits=1)`%) included stands that had undergone any notable anthropogenic thinning or partial harvest.
+Plots flagged as disturbed in ForC-simplified (`r round(n_disturbed_in_ForC_simplified/ForC_simplified_duplicates_removed*100,digits=1)`%) included stands that had undergone any notable anthropogenic thinning or partial harvest.
 We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods.
-We removed all plots for which no stand history information had been retrieved (`r round(n_no_history/nrow(ForC_simplified)*100,digits=1)`%).
+We removed all plots for which no stand history information had been retrieved (`r round(n_no_history/ForC_simplified_duplicates_removed*100,digits=1)`%).
 In total, this resulted in `r records_eligible_for_inclusion` records (`r round(records_eligible_for_inclusion/nrow(ForC_measurements)*100,digits=1)`% of the records in the database) being eligible for inclusion in the analysis.
@@
 
-To test for differences across mature forest biomes, we also examined how stand age impacted fluxes and stocks, employing a mixed effects model ['lmer' function in 'lme4' R package version **x.xx**; @bates_fitting_2015] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
+To test for differences across mature forest biomes, we also examined how stand age impacted fluxes and stocks, employing a mixed effects model ['lmer' function in 'lme4' R; @bates_fitting_2015] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
 When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
@@
 
-To facilitate the accessibility of our results and data, and to allow for rapid updates as additional data become available, we have automated all database manipulation, analyses, and figure production in R (**version, citation**).
+To facilitate the accessibility of our results and data, and to allow for rapid updates as additional data become available, we have automated all database manipulation, analyses, and figure production in R [@r_core_team_r_2020].
 
@@
 Of the 23 flux and 11 stock variables mapped in these diagrams, *ForC* contained sufficient mature forest data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ 7 distinct geographic areas) for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests.
-For regrowth forests (<100 yrs), *ForC* contained sufficient data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ **3** distinct geographic areas) for `r n_fluxes_with_ge3_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge3_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.
+For regrowth forests (<100 yrs), *ForC* contained sufficient data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ 3 distinct geographic areas) for `r n_fluxes_with_ge3_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge3_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.
 
@@
 7, S23-S25; but see below for age trends), pointing to a need for more widespread quantification of both standing and downed deadwood.
-*ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that slower decomposition in colder climates results in more buildup of organic matter [@allen_global_2005].
+*ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that slower decomposition in colder climates results in more buildup of organic matter [@allen_global_2002].
 Further research on non-living C stocks in the world's forests will be essential to completing the picture.
@@
 These would be expected to increase rapidly with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Fig. 7).
-In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$) are expected to be non-zero following stand-clearing disturbance, although these may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots and litter [@ribeiro-kumara_how_2020; @maurer_carbon_2020; @bond-lamberty_contribution_2004].
+In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$) are expected to be non-zero following stand-clearing disturbance, although these may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots and litter [@ribeiro-kumara_how_2020; @maurer_carbon_2016; @bond-lamberty_contribution_2004].
 In this study, we detect no significant age trends in either variable.
@@
 
-In terms of stocks, our study reveals consistent increases in live biomass stocks with stand age--a pattern that is well-known and expected [e.g., @lichstein_biomass_2009; @yang_global_2011]--and more variable age trends in deadwood and $OL$.
+In terms of stocks, our study reveals consistent increases in live biomass stocks with stand age--a pattern that is well-known and expected [e.g., @lichstein_biomass_2009; @yang_carbon_2011]--and more variable age trends in deadwood and $OL$.
 The latter are particularly sensitive to the type of disturbance, where disturbances that remove most organic material (e.g., logging, agriculture) result in negligible deadwood in young stands, followed by a buildup over time [tropical stands in Fig.

### 2020-09-30 manuscript/ERL_global_C_review.Rmd
cleanup

@@
 Primary research articles typically cover only a small numbers of sites at a time.
-The rare exceptions that span regions or continents with rare exceptions spanning regions or continents are typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015, e.g., @lutz_global_2018], NEON [@schimel_neon_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
+The rare exceptions that span regions or continents with rare exceptions spanning regions or continents are typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; e.g., @lutz_global_2018], NEON [@schimel_neon_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
 The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*
@@
 Since publication of ForC v2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [*SRDB* v4, `r n_SRDB` records; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database [*GROA* v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020].
-We have also added data from individual publications, with a particular focus on productivity [@banbury_morgan_global_nodate; e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @lutz_global_2018, @johnson_climate_2018].
+We have also added data from individual publications, with a particular focus on productivity [e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @lutz_global_2018, @johnson_climate_2018].
 The database version used for this analysis has been tagged as a new release on Github (v3.0) and assigned a DOI through Zenodo (DOI: TBD).
@@
 
-We selected 23 annual flux and 11 C stock variables for inclusion in the analysis.
-These different flux and stock variables (Table 1) represent different pools (e.g., aboveground biomass, dead wood, ecosystem total) and levels of combination (e.g., total aboveground net primary productivity ($ANPP$) versus the ANPP of individual elements such as foliage, roots, and branches.
+We selected 23 annual flux and 11 C stock variables for inclusion in the analysis (Table 1).
+These different flux and stock variables represent different pools (e.g., aboveground biomass, root biomass, dead wood) and levels of combination (e.g., total net primary productivity, $NPP$, versus the individual elements of $NPP$ such as foliage, roots, and branches).
 Note that two flux variables, aboveground heterotropic ($R_{het-ag}$) and total ($R_{het}$) respiration, were included for conceptual completeness but had no records in *ForC* (Table 1).
-Records for these variables represented `r round(n_focal_variable_records_eligible_for_inclusion/records_eligible_for_inclusion*100,digits=1)`% of the total records eligible for inclusion.
-For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity, such as measurements including or excluding fruit and flower production and herbivory) into more broadly defined variables.
+Records for our focal variables represented `r round(n_focal_variable_records_eligible_for_inclusion/records_eligible_for_inclusion*100,digits=1)`% of the total records eligible for inclusion.
+For this analysis, we combined some of ForC’s specific variables into more broadly defined variables. 
 Specifically, net ecosystem exchange [measured by eddy-covariance; @baldocchi_fluxnet_2001] and biometric estimates of $NEP$ were combined into the single variable $NEP$ (Table 1).
-Furthermore, for $NPP$, $ANPP$, and $ANPP_{litterfall}$, *ForC* variables specifying inclusion of different components were combined.
+Furthermore, for $NPP$, aboveground $NPP$ ($ANPP$), and the litterfall component of $ANPP$ ($ANPP_{litterfall}$), *ForC* variables specifying inclusion of different components were combined (*e.g.*, measurements including or excluding fruit and flower production and herbivory).
 Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum diameter breast height (DBH) threshold for tree census was $\le$ 10cm. 
@@
 Values were then averaged across geographically distinct areas, defined as plots clustered within 25 km of one another [*sensu* @anderson-teixeira_forc_2018], weighting by area sampled if available for all records.
-This step was taken to avoid pseudo-replication and to combine any records from sites with more than one name in ForC.
+This step was taken to avoid pseudo-replication.
 
@@
 Specifically, we first defined relationships among variables: for example, $NEP =	GPP - R_{eco}$, $BNPP = BNPP_{coarse} + BNPP_{fine}$, $DW_{tot} = DW_{standing} + DW_{down}$).
-**([issue #44](https://github.com/forc-db/ERL-review/issues/44)--but just delete this chunk if not resolved before submission)**
 Henceforth, we refer to the variables on the left side of the equation as "aggregate" fluxes or stocks, and those that are summed as "component" fluxes or stocks, noting that the same variable can take both aggregate and component positions in different relationships.
@@
 
-To test for differences across mature forest biomes, we also examined how stand age impacted fluxes and stocks, employing a mixed effects model ['lmer' function in 'lme4' R; @bates_fitting_2015] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
+To test for differences across mature forest biomes, we also examined how stand age impacted fluxes and stocks, employing a mixed effects model ['lmer' function in 'lme4' R package; @bates_fitting_2015] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
 When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
@@
 Fluxes tended to be numerically greater in temperate broadleaf than conifer forests, but the difference was never statistically significant.
-This pattern held for the following variables: GPP$, $NPP$, $ANPP$, $ANPP_{stem}$,$ANPP_{branch}$, $ANPP_{foliage}$, $BNPP$, $R_{eco}$, $R_{root}$, $R_{soil}$, and $R_{het-soil}$. 
+This pattern held for the following variables: $GPP$, $NPP$, $ANPP$, $ANPP_{stem}$,$ANPP_{branch}$, $ANPP_{foliage}$, $BNPP$, $R_{eco}$, $R_{root}$, $R_{soil}$, and $R_{het-soil}$. 
 For two of the variables without significant differences among biomes ($ANPP_{litterfall}$ and $BNPP_{fine}$; Figs.
@@
 All of these displayed a significant overall increase with with $log10[stand.age]$.
-There were sufficient data to model age x biome interactions were also significant for all ten of these C stock variables (Table S2), with living C stocks tending to accumulate more rapidly during the early stages of forest regrowth in tropical forests (Figs.
+There were sufficient data to model age $\times$ biome interactions were also significant for all ten of these C stock variables (Table S2), with living C stocks tending to accumulate more rapidly during the early stages of forest regrowth in tropical forests (Figs.
 7, S16-S22).
-In the case of two non-living C stocks ($DW_{down}$ and $OL$), age x biome interactions were such that 
+In the case of two non-living C stocks ($DW_{down}$ and $OL$), age $\times$ biome interactions were such that 
 Specifically, $DW_{down}$ declined with age in temperate and boreal forests, compared to an increase with age in tropical forests (Figs. 7, S25).

### 2020-09-30 manuscript/ERL_global_C_SI.Rmd
fixed up table S2

@@
 kable(table_pt, format = "latex", booktabs = TRUE, longtable = TRUE,  escape = F)%>%
-  #pack_rows("NEP", 1, 5, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
-  #pack_rows("GPP", 6, 10, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
-  #pack_rows("NPP", 11, 15, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
-  #pack_rows("ANPP", 16, 23, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
-  #pack_rows("ANPP_woody", 24, 28, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
-  #pack_rows("ANPP_stem", 29, 36, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
-  #pack_rows("ANPP_branch", 37, 40, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
+  pack_rows("", 1, 6, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #NEP
+  pack_rows("", 7, 12, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #GPP
+  pack_rows("", 13, 17, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #NPP
+  pack_rows("", 18, 25, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #ANPP
+  pack_rows("", 26, 29, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #ANPP_woody
+  pack_rows("", 30, 34, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #ANPP_stem
+  pack_rows("", 35, 37, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #ANPP_branch
+  pack_rows("", 38, 42, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #ANPP_foliage
+  pack_rows("", 43, 45, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #ANPP_litterfall
+  pack_rows("", 46, 48, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #skipped variables
+  pack_rows("", 49, 53, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #
+  pack_rows("", 54, 57, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #
+  pack_rows("", 58, 62, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #
+  pack_rows("", 63, 67, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #
+  pack_rows("", 68, 69, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% # 2 skipped
+  pack_rows("", 70, 74, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #
+  pack_rows("", 75, 79, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #
+  pack_rows("", 80, 84, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #
+  pack_rows("", 85, 92, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #B_tot
+  pack_rows("", 93, 100, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% # B_ag
+  pack_rows("", 101, 108, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% # 
+  pack_rows("", 109, 116, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% # 
+  pack_rows("", 117, 124, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% # 
+  pack_rows("", 125, 132, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% # 
+  pack_rows("", 133, 140, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% # 
+  pack_rows("", 141, 144, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% # 
+  pack_rows("", 145, 145, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% # 
+  pack_rows("", 146, 153, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% # 
+  pack_rows("", 154, 161, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>% #
   kable_styling(latex_options =c("repeat_header"), protect_latex = T, font_size = 7)

### 2020-09-30 manuscript/ERL_global_C_review.Rmd
fixup

@@
 
-Helene C. Muller-Landau^1^
+Helene C. Muller-Landau^2^
 
@@
 The notable exception to the pattern of fluxes decreasing from tropical to boreal regions is $NEP$ (Fig. 6f), which showed no significant differences across biomes.
-Unlike the other C flux variables, $NEP$ does not represent the rapidity with which C cycles through the ecosystem, but is the balance between C sequestration ($GPP$) and respiratory losses ($R_{eco}$) and represents net CO~2~ sequestration (or release) by the ecosystem.
+Unlike the other C flux variables, $NEP$ does not characterize the rate at which C cycles through the ecosystem, but is the balance between C sequestration ($GPP$) and respiratory losses ($R_{eco}$) and represents net CO~2~ sequestration (or release) by the ecosystem.
 $NEP$ tends to be relatively small in mature forest stands (discussed further below), which accumulate carbon slowly relative to younger stands, if at all [@luyssaert_old-growth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
@@
 6f, S1).
-This is consistent with findings that live biomass accumulation rates (*e.g.*, $\Delta B_{ag}$ or $\Delta B_{tot}$) during early secondary succession decrease with latitude [@anderson_temperature-dependence_2006; @cook-patton_mapping_2020; Figs.
-7a, S16-S22].
-Note, though, that $NEP$ includes not only $\Delta B_{tot}$, but also changes in $DW_{tot}$, $OL$, and soil carbon, and biome differences in the accumulation rates of these variables have not been detected, in part because these variables do not consistently increase with stand age [@cook-patton_mapping_2020; Figs.
-7, S23-S26; see discussion below].
+This is consistent with findings that live biomass accumulation rates (*e.g.*, $\Delta B_{ag}$ or $\Delta B_{tot}$) during early secondary succession decrease with latitude [Figs.
+7a, S16-S22; @anderson_temperature-dependence_2006; @cook-patton_mapping_2020].
+Note, though, that $NEP$ includes not only $\Delta B_{tot}$, but also changes in $DW_{tot}$, $OL$, and soil carbon, and biome differences in the accumulation rates of these variables have not been detected, in part because these variables do not consistently increase with stand age [Figs.
+7, S23-S26, and see discussion below; @cook-patton_mapping_2020].
 
@@
 Whereas biomass can be remotely sensed and receives significant research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
-Although these stocks can be important--exceeding 100 t C ha^-1^ in some stands (Figs.
+Although these stocks can be important--exceeding 100 Mg C ha^-1^ in some stands (Figs.
 7d-e, S23-S25), this study is the first to synthesize deadwood data on a global scale [but see @cook-patton_mapping_2020 for young forests].
@@
 
-Second, *ForC* can serve as a pipeline through which forest science can inform forest-based climate change mitigation efforts. 
-Such efforts will be most effective when informed by the best available data, yet it is not feasible for the individuals and organizations designing such efforts to sort through literature, often behind paywalls, with data reported in varying units, terminology, etc.
-One goal for *ForC* is to serve as a pipeline through which information can flow efficiently from forest researchers to decision-makers working to implement forest conservation strategies at global, national, or landscape scales.
+Second, *ForC* can serve as a pipeline through which information can flow efficiently from forest researchers to decision-makers working to implement forest conservation strategies at global, national, or landscape scales.
 This is already happening: *ForC* has contributed to updating the IPCC guidelines for carbon accounting in forests [@ipcc_2019_2019; @requena_suarez_estimating_2019], mapping C accumulation potential from natural forest regrowth globally [@cook-patton_mapping_2020], and informing ecosystem conservation priorities [@goldstein_protecting_2020].
@@
 6), and the highest mean biomass [Fig.
-7; Table 1; @hu_mapping_2016; @jian_restructured_2020] reinforces the concept that conservation and restoration of these forests is a priority for climate change mitigation [@grassi_key_2017], along with high-biomass old-growth temperate stands [@goldstein_protecting_2020].
+7; Table 1; @hu_mapping_2016; @jian_restructured_2020] reinforces the concept that conservation and restoration of these forests is a priority for climate change mitigation, along with high-biomass old-growth temperate stands [@grassi_key_2017; @goldstein_protecting_2020].
 It is also important to note the trade-off in climate mitigation potential of restoration of young forests, with high rates of CO~2~ sequestration [$NEP$; @cook-patton_mapping_2020], versus conservation and management of mature forests, with low $NEP$ but high C stocks that could not be recovered on a time scale relevant to climate change mitigation [@goldstein_protecting_2020].

### 2020-09-30 manuscript/ERL_global_C_review.Rmd
cleanup, final plots

@@
 
-Our analysis reveals a general deceleration of carbon cycling from the tropics to boreal regions.
+Our analysis reveals that carbon cycling is most rapid in the tropics and slowest in boreal regions, including C fluxes into ($GPP$), within (e.g., $NPP$ and its components), and out of (e.g., $R_{soil}$, $R_{eco}$) the ecosystem.
 For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude (or increase with temperature) on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
 The consistency with which this occurs across numerous fluxes is not surprising, but has never been simultaneously assessed across such a large number of variables [but see @banbury_morgan_global_nodate for nine autotrophic fluxes].
-Thus, our analysis reveals that carbon cycling is most rapid in the tropics and slowest in boreal regions, including C fluxes into ($GPP$), within (e.g., $NPP$ and its components), and out of (e.g., $R_{soil}$, $R_{eco}$) the ecosystem.
 
@@
 For aboveground biomass, which is the variable in *ForC* with broadest geographical representation, the modest trend of declining biomass from tropical to boreal regions mirrors observations from spaceborne lidar that reveal a decline in aboveground biomass (for all forests, including secondary) with latitude across the N hemisphere [@hu_mapping_2016].
-The highest- biomass forests on Earth are however found in coastal temperate climates of both the southern and northern hemisphere [@keith_re-evaluation_2009; @smithwick_potential_2002; @hu_mapping_2016].
+The highest- biomass forests on Earth are, however, found in coastal temperate climates of both the southern and northern hemisphere [@keith_re-evaluation_2009; @smithwick_potential_2002; @hu_mapping_2016].
 Disproportionate representation of forests in one such region--the US Pacific Northwest--inflated estimates of temperate conifer fluxes and stocks for some variables and was responsible for all of the anomalous results described here (e.g., lack of complete C budget closure, anomalous trend across biomes for $BNPP_{coarse}$).
@@
 To ensure that models are giving the right answers for the right reasons [@sulman_multiple_2018], it is important to benchmark against multiple components of the C cycle that are internally consistent with each other [@collier_international_2018; @wang_golum-cnp_2018].
-*ForC*’s tens of thousands of records are readily available in a standardized format, and our analyses here indicate their internal consistency is reasonably high.
+*ForC*’s tens of thousands of records are readily available in a standardized format, and our analyses here indicate that their internal consistency is reasonably high.
 Integration of *ForC* with models will be valuable to improving the accuracy and reliability of models.

### 2020-10-01 manuscript/ERL_global_C_review.Rmd
reduced file size for submission
https://github.com/forc-db/ERL-review/issues/55

@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Tropical broadleaf MATURE.png"))
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Tropical broadleaf MATURE.jpeg"))
 ```
@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate broadleaf MATURE.png"))
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate broadleaf MATURE.jpeg"))
 ```
@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate conifer MATURE.png"))
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate conifer MATURE.jpeg"))
 ```
@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Boreal conifer MATURE.png"))
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Boreal conifer MATURE.jpeg"))
 ```
@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/age_trends/for_ERL_review/Flux_age_trends.png"))
+  knitr::include_graphics(paste0(path_to_ForC, "figures/age_trends/for_ERL_review/Flux_age_trends.jpeg"))
 ```
@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/age_trends/for_ERL_review/Stock_age_trends.png"))
+  knitr::include_graphics(paste0(path_to_ForC, "figures/age_trends/for_ERL_review/Stock_age_trends.jpeg"))
 ```
@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Tropical broadleaf YOUNG.png"))
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Tropical broadleaf YOUNG.jpeg"))
 ```
@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate broadleaf YOUNG.png"))
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate broadleaf YOUNG.jpeg"))
 ```
@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate conifer YOUNG.png"))
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate conifer YOUNG.jpeg"))
 ```
@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Boreal conifer YOUNG.png"))
+  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Boreal conifer YOUNG.jpeg"))
 ```

### 2020-10-06 manuscript/ERL_global_C_review.Rmd
knit Word doc without figs/ tables, pdf with

@@
 *corresponding author: teixeirak@si.edu; +1 540 635 6546
+```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
+generate_figures=FALSE
+```
 
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 1 | Map of sites included in this analysis.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 1 | Map of sites included in this analysis.
 Symbols are colored according to the number of records at each site.
@@
 
-```{r eval = TRUE, echo=FALSE, warning=FALSE}
+```{r eval = generate_figures, echo=FALSE, warning=FALSE}
 library(knitr)
@@
 \blandscape
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 2 | C cycle diagram for mature tropical broadleaf forests.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 2 | C cycle diagram for mature tropical broadleaf forests.
 Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 3 | C cycle diagram for mature temperate broadleaf forests.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 3 | C cycle diagram for mature temperate broadleaf forests.
 Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 4 | C cycle diagram for mature temperate conifer forests.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 4 | C cycle diagram for mature temperate conifer forests.
 Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 5 | C cycle diagram for mature boreal conifer forests.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 5 | C cycle diagram for mature boreal conifer forests.
 Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEP$.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEP$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \blandscape
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | C cycle diagram for young tropical broadleaf forests.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | C cycle diagram for young tropical broadleaf forests.
 Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 9 | C cycle diagram for young temperate broadleaf forests.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 9 | C cycle diagram for young temperate broadleaf forests.
 Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 10 | C cycle diagram for young temperate conifer forests.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 10 | C cycle diagram for young temperate conifer forests.
 Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure 11 | C cycle diagram for young boreal conifer forests.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 11 | C cycle diagram for young boreal conifer forests.
 Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.

### 2020-10-06 manuscript/ERL_global_C_review.Rmd
Update ERL_global_C_review.Rmd

@@
 ```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
-generate_figures=FALSE
+generate_figures=TRUE
 ```

### 2020-12-14 manuscript/response_to_reviews.Rmd
add decision letter, response to reviews document

---
title: ""
output:
  pdf_document:
    fig_caption: yes
bibliography: references.bib
csl: environmental-research-letters.csl

header-includes:  \usepackage{float}
                  \usepackage{caption}
                  \captionsetup[table]{font=footnotesize}
                  \captionsetup[figure]{font=footnotesize}
                  \captionsetup[figure]{labelformat=empty}
                  \captionsetup[table]{labelformat=empty}
                  \usepackage{pdflscape}
                  \newcommand{\blandscape}{\begin{landscape}}
                  \newcommand{\elandscape}{\end{landscape}}

                  \usepackage{setspace}\onehalfspacing
                  \usepackage{lineno}\linenumbers

---
\raggedright

**Title:** Response to reviews

Re: "Carbon cycling in mature and regrowth forests globally" by Anderson-Teixeira, Kristina; Herrmann, Valentine; Banbury Morgan , Rebecca ; Bond-Lamberty, Benjamin; Cook-Patton, Susan; Ferson , Abigail; Muller-Landau, Helene; Wang, Maria
Article reference: ERL-109898

## REFEREE REPORT(S):

# Referee: 2

COMMENTS TO THE AUTHOR(S)

The manuscript provides an important contribution to carbon cycle research. It provides an update to a previously published data base of forest carbon stocks and fluxes (ForC) and analyses the dataset across forests biomes and in relation to stand age. This is a valuable dataset for assessing the current state of the carbon cycle and comparing across biomes. The dataset presents carbon cycle modelers with an opportunity to rigorously compare their models with observations. Figures 2-5, 8-11 are excellent summaries of the dataset. Yet I have some concerns and clarifications that must be addressed.

1. It is not clear what the various sources of data are. Perhaps this is because the authors have already described data sources in previous documentation of the dataset. However, I think some summary is needed for readers new to the dataset. I need to know what I am looking at when I see the various numbers in Figures 2-5, 8-11.

a. For example, is Fluxnet the source of GPP and Reco? The authors state that "ForC amalgamates numerous intermediary data sets" (line 119) and cite Luyssaert et al (2007). Is that the source of GPP and Reco, or have the fluxes been updated to newer Fluxnet data products? Similarly for the soil respiration data. Is this the data from Bond-Lamberty and Thomson (2010) or has it been updated?

b. I need information on how Figures 2-5, 8-11 were prepared. Carbon stocks and fluxes such as NPP, litterfall, etc. are probably plot data. There is a detailed discussion of how the data were filtered to remove, for example, disturbance effects (lines 140-149). The GPP, Reco, and soil respiration databases seem to me to be different and may not be collocated with the plot data. Was disturbance filtering applied to the GPP, Reco, and soil respiration databases, too?

c. If Luyssaert et al (2007) is a source of data, is there circularity in the comparison with latitudinal trends found in other datasets, for which Luyssaert et al is cited (lines 317-318)?

2. Figures 2-5, 8-11 are excellent summaries of the dataset. Yet there is much unexplained in the figure and the reported data values. No guidance is given on how to use the data, and especially how to resolve discrepancies in the data.

a. For example, Figure 2 lists foliage and woody aboveground biomass in tropical broadleaf forests. These do not sum to aboveground biomass (3.53 + 125.42 = 128.95 Mg C/ha vs 146.69). Nor do coarse root biomass and fine root biomass sum to root biomass (23.15 + 9.29 = 32.44 vs. 21.86). Total biomass (147.96) is not the sum of aboveground (146.69) and root (21.86).

b. Figure 4 for temperate conifer forests has large discrepancies in carbon stocks: e.g., 315.61 Mg C/ha for woody aboveground biomass, but only 167.46 for aboveground biomass and 226.58 for total biomass. The authors have a detailed discussion about lack of closure for root biomass, but do not discuss aboveground biomass. Presumably, this is because the data fall within the closure criteria, but what are we to make of the inconsistent data?

I understand that the data comes from many different sources and so may not always be compatible and that they have large standard deviations, but how are we to interpret and use the various data entries? For example, what has higher confidence: the component variables or the aggregate variables? What is the best estimate for aboveground biomass or root biomass? The authors take great pride in that component variables sum to within one standard deviation of the aggregate variables in all but one instance (lines 225-226, 301-302, 309-310). This is somewhat remarkable! But users of the dataset need guidance on how to interpret and use the data.

3. There is a general sloppiness to the manuscript that makes me think the authors did not proofread the manuscript carefully.

a. Line 58: The authors say GPP is estimated to be >69 Gt C/year, and cite Badgley et al. (2019). This statement is factually correct, but is quite misleading. Badgley et al. "estimate global annual terrestrial photosynthesis to be 147 Pg C/year (95% credible interval 131-163 Pg C/year)". The value of 69 Gt C is less than one-half that reported by Badgley. What are the authors trying to say? Such a large discrepancy does not instill confidence in the other numbers reported in the manuscript.

b. Line 81: "rare exceptions that span regions or continents" is repeated twice.

c. Figure 1: Explain the gray scale for "forest cover". It seems like this refers to gradations of the various biomes, but the figure caption does not provide an explanation.

d. Line 162: Clarify what is meant by "the minimum diameter breast height (DBH) threshold for tree census was <10cm"?

e. Table 1: (i) The definition of GPP refers to NEE, but this in never defined. The preceding entry defines NEP not NEE. (ii) The description of the biome labels is inaccurate. The table uses "TrB" but the note refers to "Tr". The table uses "BoN", but the note refers to "B".

f. Line 210: I am confused by the statement that of the 39762 records in ForC v3.0, 11923 were included in this study. Previously (lines 135-149), the authors described creating "ForC-simplified" with 17349 records. Clarify the difference between ForC-simplified and ForC v3.0.

g. Line 212: The authors refer to ForC, where in the previous sentence they referred to ForC v3.0, and previously used ForC-simplified. Clarify what database is used in the analyses.

4. This is not a review article in the traditional sense that a review is a critical assessment of recent papers in the field of carbon cycle research or identifies future research priorities. Instead, the paper documents a database and shows the utility of the database for carbon cycle research. My comments on the review aspect of the manuscript are:

a. Yes, there is a need for this database; but

b. No, this is not a critical and authoritative review of the carbon cycle; and

c. The authors reference many other publications and datasets, but they do not critically evaluate their own dataset or other datasets.

# Referee: 3

COMMENTS TO THE AUTHOR(S)

This papers uses a database recently created and updated by the author team – For C – to understand how forest carbon stocks and pools varies across broad biome classifications (e.g. boreal, temperature, and tropical) and stand age. The find that the rate of C cycling is faster in warmer climates, and that many C fluxes and pools increase with stand age (at least up to 100 years of age).

There were several things I really liked about this paper. First, I applaud the author’s ambition in creating (and maintaining) this database, which has already been used in multiple high-profile papers. It is a novel idea to curate all the C fluxes and pools together in one virtual location, and the carbon cycle budgets illustrated in Figures 2-5, and 8-11, will likely function as useful “reality checks” against which both empirical and modeling results can be assessed. I also appreciated the focus on understanding how carbon cycles varies with stand age, as this is an important unknown that limits understanding of the usefulness of reforestation as a natural climate solution (among other unknowns). Overall I found the manuscript to be clearly written.

However, I also found that many aspects of the paper gave me considerable pause.

First, the paper covers an awful lot of ground. It strikes me that each of the three research questions (bottom of page 4) could easily motivate an entire paper on their own. By attempting to address all three in one (relatively short) manuscript, it is not possible to discuss in any detail the mechanisms and processes that determine the results. There are no process-oriented hypotheses or frameworks against which the results are evaluated.

For example, concerning expectations about how various C cycles and pools vary with stand age? There is a fairly extensive literature on this topic … Odum’s classic paper on the topic (“The Strategy of Ecosystem Development, 1969) has been cited thousands of times. But this literature is not referenced or cited in the current manuscript. Odum’s hypothesis suggests that NEP (arguably the most important flux, at least from a climate mitigation perspective) should increase initially with stand age and then eventually decrease as forests continue to mature. However, this framework/literature is not referenced, and the way the results are presented make it difficult to understand whether those expectations were borne out in the data. On that note, I didn’t understand why the authors chose to show flux trends explicitly as a function of age up to 100 years (Figure 6), but then group all the forests >100 years old into a single “bar). Why not plot the mature forests explicitly on the flux versus age axis? This would allow a clearer assessment of whether flux trends with age are really linear.  It would also help the reader understand better one of the most striking results from this manuscript: that NEP of mature forests is indistinguishable across biomes. The authors describe this as unsurprising, but I think it is a bit unexpected, especially given the results of some of the more synthetic work from FLUXNET (e.g. Luyssaert et al. 2007). The authors explain this result in a couple of sentences (Page 25, lines 22-26), speculating that the result is driver by “moderate disturbances” or “disequilibrium of Rsoil relative to C inputs). A deeper dive into the results, combined with some mechanistic grounding, might reveal to what extent this result represents a climate-age interaction that is predicted from the existing theory.

Finally, I wondered about the interacting effects of changing climate (rising CO2, warming temperatures) and stand age, especially in determining trends in the pools. Mature forests will have experienced a much wider range of climate conditions than younger stands. How does this complicate the comparison of live biomass across forest of different age?

Second, I had some questions about the representativeness of the dataset. While I appreciate that the authors choose to use distinct geographic areas as the unit of analysis (avoiding some issues of pseudo-replication from many observations from a single site), I still wondered about the extent to which the distinct geographic areas were representative of the climate space within each biome. For example, if mean annual climate versus presentation for all boreal, temperate and tropical locations are shown in a scatterplot (for example, using reanalysis data), and then mean annual temperature and precipation of the observations are shown on top, how much of the “climate space” is covered by the dataset?

Third, I found the presentation of the results made it difficult to see clearly the major differences in C fluxes and pools across biomes and age classes. The illustrated C budgets (the majority of the figures, 8 in total) are visually very appealing, but the reader has to do a lot of flipping back and forth to see how any particular flux or pool varies across biomes and age class. Figure 6 is more synthetic, but each panel is very small and the differences from one group to the next are hard to see. It is also difficult to compare results for young forests (as scatterplots) with the box plots for the mature forests.  My advice is to move some of the budgets to SI, and include in the main manuscript more figures that clearly illustrate the most interesting trends with biome, and to allow an expansion of the results in Figure 6 (for example, by first showing scatterplots of all forests, young and old, as a function of stand age), and then perhaps another that is a box plot comparison of mature versus old forests in each biome.

A few other comments:

Page 6, first paragraph: I wondered about the extent to which filtering the data for “managed” affected the results. In the Eastern US, for example, its difficult to find any forests on public land that aren’t managed to some extent (for example, through periodic selective harvests), and many of them regenerated from “planted” stands back in the 1930s. I would be curious to know if including “managed” forests substantially altered results.

The binary age classification (two ages, less than 100 years or greater than 100 years) was difficult to accept, as forests that are in the 80-100 year age range are often considered to be mature (at least in the temperate zone). I realize that even with a dataset as rich as For C, data availability will limit stratification into too many bins. Nonetheless, at least for some measurements, I wonder if it's possible to consider a greater number of age classes (for example, young, maturing, and mature).

The consistency check (e.g. do component fluxes/pools sum to within one standard deviation of the aggregate flux or pool) seems like it is destined to provide a favorable assessment of the degree of closure so long as the data within each grouping represent a wide range of natural climatic and soil variability. As long as the aggregate variable has a large standard deviation, the results are likely guaranteed to be “consistent.” It seems this metric would benefit from some simulations (perhaps with artificially generated data) to understand exactly how poor closure needs to be at the site-level to generate an inconsistent results when aggregated across sites.

# Referee: 4

COMMENTS TO THE AUTHOR(S)
General:
This paper presents the findings of a rather comprehensive modeling analysis of C fluxes and stocks in the world’s major forest biomes. An important strength of this work is its reliance on a fairly large empirical data set to calculate a comprehensive suite of fluxes and stocks to close (or come close) C budgets in these systems. Another unique component of this work is the contrasts of young vs mature forests in each biome. An important contribution of this work is its highlighting of gaps in data (e.g., deadwood) and non-random distribution of empirical data (though this, of course is well-known) and how these factors influence C accounting efforts. Overall, the paper is well written with beautiful figures and the analyses and data sets are sound. I think this work makes an important and timely contribution to fields of C cycle science, forest ecology, and ecosystem ecology and is likely to be of interest to many ERL readers. I have no fundamental concerns with this manuscript. However, in addition to the detailed comments/suggestions below, I think it would greatly benefit the paper if the authors could include some discussion of how the forest ecosystems they characterize here compare to the forest ecosystems that actually exist. The data here represent generally interior forest ecosystems, which of course are incredibly important. But, work over the last 5 years or so highlights the extent to which forest fragmentation influences a large proportion of the world’s forests (Haddad et al., 2015). Fragmentation and the creation of edges has been shown to have important implications for C stocks and fluxes with regional and global implications (e.g., (Chaplin-Kramer et al., 2015; Remy et al., 2016; Reinmann & Hutyra, 2017; Smith et al., 2019; Ordway & Asner, 2020; Reinmann et al., 2020: FULL REFS BELOW). Logging and other forms of management also influence a large proportion of the world’s forests. I am not suggesting this be included in modeling efforts here, but in placing this work in the broader context of C stocks and fluxes of the world’s forests I think it would do the scientific community a great service to more explicitly recognize what is being modeled and perhaps the proportion of the world’s forests these data might represent…even if discussed in a qualitative sense.

Introduction:

L81: “exceptions” in sentence twice.

L97: Typo “Since the its most”

Methods:

L144: If there is a non-trivial proportion of the world’s forests (especially in certain biomes) is plantation or planted forests, does removing such plots from the dataset bias the results of a global modeling product?

Results:

L243-245: I think this is per unit forest area, correct? If so, I think it would be helpful to specify that here. Also, the reader should be referred to Figs 2-5, not just 5, right?

L270-271: The wording of the sentence “There were sufficient data to model….” Seems a bit awkward. Should it read “…WHICH were also significant…”?

Discussion:
You might consider reiterating in the first paragraph of the Discussion section that your findings indicate that Temperate Broadleaf Forests are actually the most productive forest biome in terms of NEP. I think this is a surprising finding (we normally think of Tropical forests as being more productive) that warrants attention.

L297: The fact that ForC does not include soil C is important. It is not a flaw in the model or the approach here, but throughout you discuss C stocks, which many would interpret as being inclusive of soil C. I think the authors should consider clearly indicating in the Methods section, and perhaps reiterating in the Results and Discussion, that stocks here are defined as litter layer, biomass, and necromass, but excludes C in soil.

L322-323: As you point out in the Results section, while there are no statistically sig differences in NEP across biomes, there are large differences in the means. Is this an artifact of the data sources used (i.e. distribution and number of sites with empirical data)? Can you speak a little more to this point in this section?

L382: You might consider changing text to “increases with age AT LEAST up to the 100-yr threshold examined here” so that it does not come across as suggesting that NEP only increases for the first hundred years of stand development, which we know is not true.

L445-450: I come back to the data in Figs 2-5 and text in the results section (L243-245), which indicates no sig differences in NEP, but highest means in temperate forests. Of course, the high biomass in tropical forests makes them critical to protect from a C storage perspective, but if NEP (i.e. rates of C sequestration) are highest, at least as a mean, in temperate broadleaf forests how should those ecosystems factor into conservation priorities. Related to this point, is the high NEP of temperate forests driven by the relatively young nature of temperate broadleaf forests in the eastern U.S., where a lot of data exist?

References mentioned above:

Chaplin-Kramer R., Ramler I., Sharp R., Haddad N. M., Gerber J. S., West P. C., … King H. (2015). Degradation in carbon stocks near tropical forest edges. Nature Communications, 6, 1–6. https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1038%2Fncomms10158&amp;data=04%7C01%7Cteixeirak%40si.edu%7Cd27c53d6ec434c8ae42108d8a03017cf%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637435475707545147%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C1000&amp;sdata=maYY3mHl%2F%2FX0jgkAOK5Kc9K1TpaoTVkFM52HsOxeMjU%3D&amp;reserved=0.

Haddad N. M., Brudvig L. a., Clobert J., Davies K. F., Gonzalez A., Holt W. M., … Townshend J. R. (2015). Habitat fragmentation and its lasting impact on Earth’s ecosystems. Science Advances, (March), e1500052. https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.4028%2Fwww.scientific.net%2FAMM.315.108&amp;data=04%7C01%7Cteixeirak%40si.edu%7Cd27c53d6ec434c8ae42108d8a03017cf%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637435475707545147%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C1000&amp;sdata=1L35xf2RdBj6Y9PjsYhSY5tct4RhnJLE8Jd50QJdzBM%3D&amp;reserved=0.

Ordway E. M. & Asner G. P. (2020). Carbon declines along tropical forest edges correspond to heterogeneous effects on canopy structure and function. Proceedings of the National Academy of Sciences of the United States of America, 117(14), 7863–7870. https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1073%2Fpnas.1914420117&amp;data=04%7C01%7Cteixeirak%40si.edu%7Cd27c53d6ec434c8ae42108d8a03017cf%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637435475707545147%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C1000&amp;sdata=DnTMc4LCIoRcHMLUEAcpvNOn%2FeJLO3%2Bh8pXVYSUb%2BI4%3D&amp;reserved=0.

Reinmann A. B. & Hutyra L. R. (2017). Edge effects enhance carbon uptake and its vulnerability to climate change in temperate broadleaf forests. Proceedings of the National Academy of Sciences, 114(1), 107–112. https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1073%2Fpnas.1612369114&amp;data=04%7C01%7Cteixeirak%40si.edu%7Cd27c53d6ec434c8ae42108d8a03017cf%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637435475707545147%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C1000&amp;sdata=GncJJ%2FLa2f96c0tjkl5RZZv2JLJFoAoQnOTm0b%2BYyNo%3D&amp;reserved=0.

Reinmann A. B., Smith I. A., Thompson J. R. & Hutyra L. R. (2020). Urbanization and fragmentation mediate temperate forest carbon cycle response to climate. Environmental Research Letters, 15, 114036.

Remy E., Wuyts K., Boeckx P., Ginzburg S., Gundersen P., Demey A., … Verheyen K. (2016). Strong gradients in nitrogen and carbon stocks at temperate forest edges. Forest Ecology and Management, 376(2016), 45–58. https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1016%2Fj.foreco.2016.05.040&amp;data=04%7C01%7Cteixeirak%40si.edu%7Cd27c53d6ec434c8ae42108d8a03017cf%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637435475707545147%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C1000&amp;sdata=u2ylmC4K4VIxHWamrCnk%2BNfu%2FsZDUkSPfABgfN%2BCzT4%3D&amp;reserved=0.

Smith I. A., Hutyra L. R., Reinmann A. B. & Thompson J. R. (2019). Evidence for Edge Enhancements of Soil Respiration in Temperate Forests Geophysical Research Letters. Geophysical Research Letters, 46, 1–10. https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1029%2F2019GL082459&amp;data=04%7C01%7Cteixeirak%40si.edu%7Cd27c53d6ec434c8ae42108d8a03017cf%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637435475707545147%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C1000&amp;sdata=92jjRZSUzQrNX6SGlV%2BZRfHlUN%2B6AL96%2FJIl4PIK3a8%3D&amp;reserved=0.

## References

```{r eval=FALSE, echo=FALSE}
# All references should be presented in alphabetical order and should also be bundled in ways that will help the reader to find those that are germane to each of the themes, topics or issues that the authors have used to organize the review. Such a matrix will be included as supplementary material. ERL has no standardized format for subdividing references by cross-cutting themes, but guidance from editors will be helpful to authors.
```

### 2020-12-14 manuscript/ERL_global_C_review.tex
add decision letter, response to reviews document

@@
 \usepackage[margin=1in]{geometry}
-\usepackage{graphicx,grffile}
+\usepackage{graphicx}
 \makeatletter
@@
 \usepackage{makecell}
+\usepackage{xcolor}
+\ifluatex
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {\par}
 
@@
 \textbf{Title:} Carbon cycling in mature and regrowth forests globally:
-a macroecological synthesis based on the global Forest Carbon (ForC)
+a macroecological synthesis based on the Global Forest Carbon (ForC)
 database
@@
 
-Becky Banbury Morgan\textsuperscript{3}
+Becky Banbury Morgan\textsuperscript{1,3}
 
@@
 
-Jennifer C. McGarvey\textsuperscript{1}
-
-Helene C. Muller-Landau\textsuperscript{1}
+Helene C. Muller-Landau\textsuperscript{2}
 
@@
 
-\begin{verbatim}
-## [1] 0
-\end{verbatim}
-
-\emph{NOTES TO COAUTHORS: }
-
-\begin{itemize}
-\tightlist
-\item
-  We're still finalizing the data (and adding some data). Outliers in
-  plots will all be checked/ resolved, and we'll be able to pull in more
-  data (GROA needs classification by dominant vegetation before it can
-  be pulled in, and we're working on that)
-\item
-  ``\textbf{???}'' indicates a reference that we have entered in the
-  .Rmd file, but not yet in the bibliography file. Don't worry about
-  those. (However, places with ``REF'' need references)
-\item
-  bold/ italic text indicates places where I know we need to go back and
-  revisit/ fill in info
-\end{itemize}
-
 \newpage
@@
 
-\emph{Background.} The fate of Earth's climate is closely linked to
-forests, which strongly influence atmospheric carbon dioxide
-(CO\textsubscript{2}) and climate through their influential role in the
-global carbon (C) cycle. Synthetic understanding of global forest C
-cycles is needed to constrain model estimates of forest feedbacks to
-climate change and to more accurately quantify the influence of land use
-decisions on climate.
-
-\emph{Methods/Design.} Here, we draw from the Global Forest C Database,
-ForC, to provide a macroscopic overview of C cycling in the world's
-forests, giving special attention to stand age-related variation.
-Specifically, we use 14847 ForC records from 874 geographic locations
-representing 34 C cycle variables to characterize ensemble C budgets for
-four broad forest types (tropical broadleaf evergreen, temperate
-broadleaf, temperate conifer, and taiga), including estimates for both
-mature and regrowth (age \textless100 years) forests. For regrowth
-forests, we quantify age trends for all variables.
-
-\emph{Review Results/ Synthesis.} ForC v3.0 yielded a fairly
-comprehensive picture of C cycling in the world's major forest biomes,
-with broad closure of C budgets. The rate of C cycling generally
-increased from boreal to tropical regions in both mature and regrowth
-forests, whereas C stocks showed less directional variation. The
-majority of flux variables, together with most live biomass pools,
-increased significantly with stand age, \emph{and the rate of increase
-again tended to increase from boreal to tropical regions}.
-
-\emph{Discussion.} \textbf{NEED TO WRITE THIS!!!}
+\emph{Background.} Earth's climate is closely linked to forests, which
+strongly influence atmospheric carbon dioxide (CO\textsubscript{2}) and
+climate by their impact on the global carbon (C) cycle. However, efforts
+to incorporate forests into climate models and CO\textsubscript{2}
+accounting frameworks have been constrained by a lack of accessible,
+global-scale data on how C cycling varies across forest types and stand
+ages.
+
+\emph{Methods/Design.} Here, we draw from the Global Forest Carbon
+Database, ForC, to provide a macroscopic overview of C cycling in the
+world's forests, giving special attention to stand age-related
+variation. Specifically, we use 11923 \emph{ForC} records from 865
+geographic locations representing 34 C cycle variables to characterize
+ensemble C budgets for four broad forest types (tropical broadleaf
+evergreen, temperate broadleaf, temperate conifer, and taiga), including
+estimates for both mature and regrowth (age \textless100 years) forests.
+For regrowth forests, we quantify age trends for all variables with
+sufficient data.
+
+\emph{Review Results/ Synthesis.} \emph{ForC v3.0} yielded a
+comprehensive picture of C cycling in the world's major forest biomes.
+The rate of C cycling generally increased from boreal to tropical
+regions in both mature and regrowth forests, whereas C stocks showed
+less directional variation. The majority of flux variables, together
+with most live biomass pools, increased significantly with stand age.
+Importantly, there was generally good closure of C budgets, \emph{i.e.,}
+internal consistency in the \emph{ForC} data.
+
+\emph{Discussion.} As climate change accelerates, understanding and
+managing the carbon dynamics of forests is critical to forecasting,
+mitigation, and adaptation. This synthetic and internally consistent
+global overview of C stocks and fluxes across biomes and stand ages will
+help to advance these efforts.
 
@@
 
-\emph{(Abby has offered to update stats in this paragraph:)}
-
-Forest ecosystems will play a critical role in shaping the course of
-climate change (IPCC1.5) through their influence on atmospheric carbon
-dioxide (CO\textsubscript{2}). Their annual gross CO\textsubscript{2}
-sequestration (gross primary productivity, \(GPP\)) is estimated at
-\textgreater69 Gt C yr\textsuperscript{-1} ({\textbf{???}}), or
-\textgreater7 times average annual fossil fuel emissions from 2007-2016
-(9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (\textbf{update}). While most
-of this enormous C flux is couterbalanced by CO\textsubscript{2}
-releases to the atmosphere through ecosystem respiration (\(R_{eco}\))
-or fire, a small portion has been retained in ecosystems over recent
-decades. The resulting CO\textsubscript{2} sink averaged 3.0 ± 0.8 Gt C
-yr\^{}\{-1\} from 2007-2016, offsetting 32\% of anthropogenic fossil
-fuel emissions (Le Quéré et al 2017) (\textbf{update, give range}).
-Moreover, forests contain substantial carbon (C) stocks: an estimated
-92\% of terrestrial biomass (Pan et al 2013) and 45\% of terrestrial C
-(biomass and soils; Bonan 2008). Forests are also globally dominant as a
-source of soil respiration ({\textbf{???}}). Globally, net deforestation
-(\emph{i.e.}, gross deforestation - regrowth) has been a source of
-CO\textsubscript{2} emissions, estimated at \textasciitilde1.1 Gt C yr-1
-from YEAR-YEAR (Pan et al 2011), reducing the net forest sink to
-\textasciitilde1.2-1.7 Gt C yr\textsuperscript{-1} across Earth's
-forests (Le Quéré et al 2017, Schimel \emph{et al} 2015)
-(\textbf{update, give range}). The future of the current forest C is
-dependent both upon forest responses to a broad suite of global change
-drivers and to future land use decisions, and will strongly influence
-the course of climate change. Regrowing forests in particular will play
-an important role (Pugh \emph{et al} 2019), as these represent a large
-(\textbf{\textasciitilde\#\%}) and growing proportion of Earth's forests
-(McDowell \emph{et al} 2020). Understanding, modeling, and managing
+Forest ecosystems are shaping the course of climate change through their
+influence on atmospheric carbon dioxide (CO\textsubscript{2}; Bonan
+2008, Friedlingstein \emph{et al} 2019, IPCC 2018). Despite the
+centrality of forest C cycling in regulating atmospheric
+CO\textsubscript{2}, important uncertainties in climate models
+(Friedlingstein \emph{et al} 2006, Krause \emph{et al} 2018, Bonan
+\emph{et al} 2019, Di Vittorio \emph{et al} 2020) and
+CO\textsubscript{2} accounting frameworks (Pan \emph{et al} 2011) can be
+traced to lack of accessible, comprehensive data on how C cycling varies
+across forest types and in relation to stand history. These require
+large-scale databases with global coverage, which runs contrary to the
+nature in which forest C stocks and fluxes are measured and published.
+While remote sensing measurements are increasingly useful for global- or
+regional-scale estimates of a few critical variables (e.g., aboveground
+biomass: Hu \emph{et al} 2016, Spawn \emph{et al} 2020, gross primary
+productivity, \(GPP\): Li and Xiao 2019, Saatchi \emph{et al} 2011),
+measurement and validation of most forest C stocks and fluxes require
+intensive on-the-ground data collection.
+
+A robust understanding of forest impacts on global C cycling is
+essential. Annual gross CO\textsubscript{2} sequestration in forests
+(gross primary productivity, \(GPP\)) is estimated at \textgreater69 Gt
+C yr\textsuperscript{-1} (Badgley \emph{et al} 2019), or \textgreater7
+times average annual fossil fuel emissions from 2009-2018 (9.5 \(\pm\)
+0.5 Gt C yr\textsuperscript{-1}; Friedlingstein \emph{et al} 2019). Most
+of this enormous C sequestration is counterbalanced by
+CO\textsubscript{2} releases to the atmosphere through ecosystem
+respiration (\(R_{eco}\)) or fire, with forests globally dominant as
+sources of both soil respiration (Warner \emph{et al} 2019) and fire
+emissions (Werf \emph{et al} 2017). In recent years, the remaining
+CO\textsubscript{2} sink averaged 3.2 \(\pm\) 0.6 GtC
+yr\textsuperscript{-1} from 2009-2018, offsetting 29\% of anthropogenic
+fossil fuel emissions (Friedlingstein \emph{et al} 2019). Yet, this sink
+is reduced by deforestation, estimated at \textasciitilde1 Gt C
+yr\textsuperscript{-1} in recent decades (Pan \emph{et al} 2011,
+Tubiello \emph{et al} 2020), reducing the net forest sink to
+\textasciitilde1.1-2.2 Gt C yr\textsuperscript{-1} across Earth's
+forests (Friedlingstein \emph{et al} 2019).
+
+The future of the current forest C sink is dependent both upon forest
+responses to climate change itself and human land use decisions, which
+will feedback and strongly influence the course of climate change.
+Regrowing forests in particular will play an important role (Pugh
+\emph{et al} 2019), as almost two-thirds of the world's forests were
+secondary as of 2010 (FAO 2010). As anthropogenic and climate-driven
+disturbances impact an growing proportion of Earth's forests (Andela
+\emph{et al} 2017, McDowell \emph{et al} 2020), understanding the carbon
+dynamics of regrowth forests is increasingly important
+(Anderson‐Teixeira \emph{et al} 2013). Although age trends in
+aboveground biomass have been relatively well-studied and synthesized
+globally (Cook-Patton \emph{et al} 2020), a relative dearth of data and
+synthesis on other C stocks and fluxes in secondary forests points to an
+under-filled need to characterize age-related trends in forest C
+cycling. Such understanding is particularly critical for reducing
+uncertainty regarding the potential for carbon uptake and climate change
+mitigation by regrowth forests (Krause \emph{et al} 2018, Cook-Patton
+\emph{et al} 2020). Understanding, modeling, and managing
 forest-atmosphere CO\textsubscript{2} exchange is thus central to
-efforts to mitigate climate change {[}Grassi \emph{et al} (2017);
-Griscom \emph{et al} (2017); \textbf{Cavaleri et al 2015}{]}.
-
-Despite the centrality of forest C cycling in regulating atmospheric
-CO\textsubscript{2}, important uncertainties in climate models
-({\textbf{???}}, {\textbf{???}}, {\textbf{???}}, Krause \emph{et al}
-2018) and CO\textsubscript{2} accounting frameworks (Pan \emph{et al}
-2011) can be traced to lack of accessible, comprehensive data on how C
-cycling varies across forest types and in relation to stand history.
-These require large-scale databases with global coverage, which runs
-contrary to the nature in which forest C stocks and fluxes are measured
-and published. While remote sensing measurements are increasingly useful
-for global- or regional-scale estimates of a few critical variables
-{[}Li and Xiao (2019); \textbf{REFS for biomass, biomass change, net CO2
-flux}{]}, measurement of most forest C stocks and fluxes require
-intensive on-the-ground data collection. Original studies typically
-cover only a small numbers of sites at a time, with rare exceptions
-spanning regions or continents {[}e.g., Lutz \emph{et al} (2018);
-\textbf{FLUXNET\_REF}), typically coordinated through research networks
-such as ForestGEO (Anderson-Teixeira \emph{et al} 2015) or FLUXNET
-(Baldocchi \emph{et al} 2001). The result of decades of research on
-forest C cycling is that tens of thousands of records have been
+efforts to mitigate climate change (Grassi \emph{et al} 2017, Griscom
+\emph{et al} 2017, Cavaleri \emph{et al} 2015).
+
+Despite the importance of forests, comprehensive global studies have
+historically been limited by the scattered and more local nature of
+research studies. Primary research articles typically cover only a small
+numbers of sites at a time. The rare exceptions that span regions or
+continents with rare exceptions spanning regions or continents are
+typically coordinated through research networks such as ForestGEO
+(Anderson-Teixeira \emph{et al} 2015, e.g., Lutz \emph{et al} 2018),
+NEON (Schimel \emph{et al} 2007), or FLUXNET (Baldocchi \emph{et al}
+2001, e.g., Novick \emph{et al} 2018). The result of decades of research
+on forest C cycling is that tens of thousands of records have been
 distributed across literally thousands of scientific articles --often
 behind paywalls-- along with variation in data formats, units,
-measurement methods, \emph{etc.}. In this format, the data are
+measurement methods, \emph{etc.} In this format, the data are
 effectively inaccessible for many global-scale analyses, including those
@@
 to quantify actual or scenario-based exchanges of CO\textsubscript{2}
-between forests and the atmosphere (\textbf{REFS}).
+between forests and the atmosphere (Griscom \emph{et al} 2017, Houghton
+2020).
 
 To address the need for global-scale analyses of forest C cycling, we
-recently developed an open-access Global Forest Carbon database, ForC
-(Anderson-Teixeira \emph{et al} (2016), Anderson-Teixeira \emph{et al}
-(2018)). ForC contains published estimates of forest ecosystem C stocks
-and annual fluxes (\textgreater50 variables) based on ground-based
-measurements, along with associated data required for interpretation
-(\emph{e.g.}, stand history, measurement methods). These data have been
-amalgamated from original peer-reviewed publications, either directly or
-via intermediary data compilations. Since the its most recent
-publication (Anderson-Teixeira \emph{et al} 2018), ForC has been
-integrated with two large databases: the Global Soil Respiration
-Database (SRDB; Bond-Lamberty and Thomson 2010) and the Global
-Reforestation Opportunity Assessment database (GROA; Cook-Patton
-\emph{et al} 2020), both of which have also synthesized published forest
-C data. Following these additions, ForC currently contains 47846 records
-from 10609 plots and 1532 distinct geographic areas representing all
-forested biogeographic and climate zones.
-
-Here, we synthesize ForC data (Fig. 1) to provide a macroscopic overview
-of stand-level carbon cycling of the world's major forest biomes and how
-it varies with stand age. Our primary goal is to provide a data-based
-summary of our current state of knowledge on broad trends in forest C
-cycling. We address three broad questions:
+recently developed \emph{ForC} (Anderson-Teixeira \emph{et al} 2016,
+2018). \emph{ForC} contains published estimates of forest ecosystem C
+stocks and annual fluxes (\textgreater50 variables), with the different
+variables capturing the unique ecosystem pools (e.g., woody, foliage,
+and root biomass; dead wood) and flux types (e.g, gross and net primary
+productivity; soil, root, and ecosystem respiration). These data are
+ground-based measurements, and \emph{ForC} contains associated data
+required for interpretation (\emph{e.g.}, stand history, measurement
+methods). Data have been amalgamated from original peer-reviewed
+publications, either directly or via intermediary data compilations.
+Since the its most recent publication (Anderson-Teixeira \emph{et al}
+2018), \emph{ForC} has grown to include two additional large databases:
+the Global Soil Respiration Database (SRDB; Bond-Lamberty and Thomson
+2010) and the Global Reforestation Opportunity Assessment database
+(GROA; Cook-Patton \emph{et al} 2020) that also synthesized published
+forest C data. Following these additions, \emph{ForC} currently contains
+39762 records from 10608 plots and 1532 distinct geographic areas
+representing all forested biogeographic and climate zones. This
+represents an 129\% increase in records from the prior publication
+(Anderson-Teixeira \emph{et al} 2018).
+
+Here, we provide a robust and comprehensive analysis of carbon cycling
+from a stand to global level, and by biome and stand age, using the
+largest global compilation of forest carbon data, which is available in
+our open source Global Carbon Forest database (\emph{ForC}; Fig. 1). Our
+primary goal is to provide a data-driven summary of our current state of
+knowledge on broad trends in forest C cycling. Specifically, we address
+three broad questions:
 
@@
 \item
-  To what extent can we fully represent, and ``close'', C budgets for
-  each of the world's major forest biomes (\emph{i.e.}, tropical,
-  temperate broadleaf and deciduous, boreal) based on the current ForC
-  data?
+  How thoroughly can we represent C budgets for each of the world's
+  major forest biomes (\emph{i.e.}, tropical, temperate broadleaf and
+  deciduous, boreal) based on the current \emph{ForC} data?
 \item
@@
 in global forests, and will serve as a foundation for improved
-understanding of global forest C cycling.
+understanding of global forest C cycling and highlight where key sources
+of uncertainty still reside.
 
@@
 2006) and biomes.
-Distribution of sites, plots, and records among biomes is shown in the inset.}\label{fig:unnamed-chunk-4}
+Distribution of sites, plots, and records among biomes is shown in the inset.}\label{fig:unnamed-chunk-5}
 \end{figure}
@@
 
-This review synthesizes data from the ForC database (Fig. 1;
+This review synthesizes data from the \emph{ForC} database (Fig. 1;
 \url{https://github.com/forc-db/ForC}; Anderson-Teixeira \emph{et al}
-2016, pp @anderson--teixeira\_forc\_2018). ForC amalgamates numerous
-intermediary data sets (\emph{e.g.}, \textbf{REFS}) and original
-studies. Original publications were referenced to check values and
-obtain information not contained in intermediary data sets, although
-this process has not been completed for all records. The database was
-developed with goals of understanding how C cycling in forests varies
-across broad geographic scales and as a function of stand age. As such,
-there has been a focus on incorporating data from regrowth forests
-(\emph{e.g.}, Anderson et al 2006, Martin et al 2013, Bonner et al 2013)
-and obtaining stand age data when possible (83\% of records in v.2.0;
-Anderson-Teixeira \emph{et al} 2018). Particular attention was given to
-developing the database for tropical forests (Anderson-Teixeira \emph{et
-al} 2016). Since publication of ForC v.2.0, we added the following data
-to ForC: the Global Database of Soil Respiration Database (SRDB v.\#\#,
-9497 records; Bond-Lamberty and Thomson 2010), the Global Reforestation
-Opportunity Assessment database (GROA v1.0, 18143 records; Cook-Patton
-\emph{et al} 2020, Anderson-Teixeira \emph{et al} 2020). We've also
-added data from individual publications (detailed list at
-\url{https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv}),
-with a particular focus on productivity (Banbury Morgan \emph{et al}
-n.d., e.g., Taylor \emph{et al} 2017), dead wood, and ForestGEO sites
-(e.g., Lutz \emph{et al} 2018, p @johnson\_climate\_2018). We note that
-there remains a significant amount of relevant data that is not yet
-included in ForC, particularly biomass data from national forest
-inventories (\emph{e.g.},: \textbf{REFS}). The database version used for
-this analysis has been tagged as a new release on Github (\textbf{XX})
-and assigned a DOI through Zenodo (DOI: TBD).
-
-Analyses drew from ForC-simplified
+2016, 2018). \emph{ForC} amalgamates numerous intermediary data sets
+(\emph{e.g.}, Luyssaert \emph{et al} 2007, Bond-Lamberty and Thomson
+2010, Cook-Patton \emph{et al} 2020) and original studies. Original
+publications were referenced to check values and obtain information not
+contained in intermediary data sets, although this process has not been
+completed for all records. The database was developed with goals of
+understanding how C cycling in forests varies across broad geographic
+scales and as a function of stand age. As such, there has been a focus
+on incorporating data from regrowth forests (\emph{e.g.}, Anderson et al
+2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data
+when possible (83\% of records in v.2.0; Anderson-Teixeira \emph{et al}
+2018). Particular attention was given to developing the database for
+tropical forests (Anderson-Teixeira \emph{et al} 2016), which
+represented roughly one-third of records in \emph{ForC} v2.0
+(Anderson-Teixeira \emph{et al} 2018). Since publication of ForC v2.0,
+we added the following data to ForC: the Global Database of Soil
+Respiration Database (\emph{SRDB} v4, 9488 records; Bond-Lamberty and
+Thomson 2010), the Global Reforestation Opportunity Assessment database
+(\emph{GROA} v1.0, 10116 records; Cook-Patton \emph{et al} 2020,
+Anderson-Teixeira \emph{et al} 2020). We have also added data from
+individual publications, with a particular focus on productivity (e.g.,
+Taylor \emph{et al} 2017), dead wood, and ForestGEO sites (e.g., Lutz
+\emph{et al} 2018, p @johnson\_climate\_2018). The database version used
+for this analysis has been tagged as a new release on Github (v3.0) and
+assigned a DOI through Zenodo (DOI: TBD).
+
+To facilitate analyses, we created a simplified version of ForC,
+\emph{ForC-simplified}
 (\url{https://github.com/forc-db/ForC/blob/master/ForC_simplified}),
-which is a rearrangement of ForC intended to facilitate analyses. In
-generating ForC-simplified, all measurements originally expressed in
-units of dry organic matter (\(OM\)) were converted to units of C using
-the IPCC default of \(C = 0.47 * OM\) (IPCC 2006). Duplicate or
-otherwise conflicting records were reconciled as described in
-\emph{APPENDIX S1}, resulting in a total of 30927 records (64.6\% size
-of total database). Records were filtered to remove plots that had
-undergone significant anthropogenic management or major disturbance
-since the most recent stand initiation event. Specifically, we removed
-all plots flagged as managed in ForC-simplified (19\%). This included
-plots with any record of managements manipulating CO\textsubscript{2},
-temperature, hydrology, nutrients, or biota, as well as any plots whose
-site or plot name contained the terms ``plantation'', ``planted'',
-``managed'', ``irrigated'', or ``fertilized''. Plots flagged as
-disturbed in ForC-simplified (5.9\%) included stands that had undergone
-anthropogenic thinning or partial harvest unless this was very minor. We
-retained sites that were grazed or had undergone low severity natural
-disturbances (\textless10\% mortality) including droughts, major storms,
-fires, and floods. We removed all plots for which no stand history
-information had been retrieved (7.7\%). In total, this resulted in 21872
-records (45.7\% of the records in the database) being eligible for
-inclusion in the analysis.
+which we analyzed here. In generating \emph{ForC-simplified}, all
+measurements originally expressed in units of dry organic matter
+(\(OM\)) were converted to units of C using the IPCC default of
+\(C = 0.47 * OM\) (IPCC 2018). Duplicate or otherwise conflicting
+records were reconciled as described in Appendix S1, resulting in a
+total of 22265 records (56\% size of total database). Records were
+filtered to remove plots that had undergone significant anthropogenic
+management or major disturbance since the most recent stand initiation
+event. Specifically, we removed all plots flagged as managed in
+ForC-simplified (13.9\%). This included plots with any record of
+managements manipulating CO\textsubscript{2}, temperature, hydrology,
+nutrients, or biota, as well as any plots whose site or plot name
+contained the terms ``plantation'', ``planted'', ``managed'',
+``irrigated'', or ``fertilized''. Plots flagged as disturbed in
+ForC-simplified (5.6\%) included stands that had undergone any notable
+anthropogenic thinning or partial harvest. We retained sites that were
+grazed or had undergone low severity natural disturbances (\textless10\%
+mortality) including droughts, major storms, fires, and floods. We
+removed all plots for which no stand history information had been
+retrieved (5.7\%). In total, this resulted in 17349 records (43.6\% of
+the records in the database) being eligible for inclusion in the
+analysis.
 
 We selected 23 annual flux and 11 C stock variables for inclusion in the
-analysis, although two flux variables (\(R_{het-ag}\) and \(R_{het}\))
-were included for conceptual completeness but had no records in ForC
-(Table 1). Records for these variables represented 89.1\% of the total
-records eligible for inclusion. For this analysis, we combined some of
-ForC's specific variables (\emph{e.g.}, multiple variables for net
-primary productivity including various components) into more broadly
-defined variables. Specifically, net ecosystem exchange (measured by
-eddy-covariance; \textbf{FLUXNET REF}) and biometric estimates of
-\(NEP\) were combined into the single variable \(NEP\) (Table 1).
-Furthermore, for \(NPP\), \(ANPP\), and \(ANPP_{litterfall}\), ForC
-variables specifying inclusion of different components were combined.
-Throughout ForC, for all measurements drawing from tree census data
-(\emph{e.g.}, biomass, productivity), the minimum threshold for tree
-census was \(\le\) 10cm. All records were measured directly or derived
-from field measurements (as opposed to modeled).
+analysis (Table 1). These different flux and stock variables represent
+different pools (e.g., aboveground biomass, root biomass, dead wood) and
+levels of combination (e.g., total net primary productivity, \(NPP\),
+versus the individual elements of \(NPP\) such as foliage, roots, and
+branches). Note that two flux variables, aboveground heterotropic
+(\(R_{het-ag}\)) and total (\(R_{het}\)) respiration, were included for
+conceptual completeness but had no records in \emph{ForC} (Table 1).
+Records for our focal variables represented 90.3\% of the total records
+eligible for inclusion. For this analysis, we combined some of ForC's
+specific variables into more broadly defined variables. Specifically,
+net ecosystem exchange (measured by eddy-covariance; Baldocchi \emph{et
+al} 2001) and biometric estimates of \(NEP\) were combined into the
+single variable \(NEP\) (Table 1). Furthermore, for \(NPP\), aboveground
+\(NPP\) (\(ANPP\)), and the litterfall component of \(ANPP\)
+(\(ANPP_{litterfall}\)), \emph{ForC} variables specifying inclusion of
+different components were combined (\emph{e.g.}, measurements including
+or excluding fruit and flower production and herbivory). Throughout
+ForC, for all measurements drawing from tree census data (\emph{e.g.},
+biomass, productivity), the minimum diameter breast height (DBH)
+threshold for tree census was \(\le\) 10cm. All records were measured
+directly or derived from field measurements (as opposed to modeled).
 
@@
 
-\caption{\label{tab:unnamed-chunk-6}Table 1. Carbon cycle variables included in this analysis, their sample sizes, and summary of biome differences and age trends.}
+\caption{\label{tab:unnamed-chunk-7}Table 1. Carbon cycle variables included in this analysis, their sample sizes, and summary of biome differences and age trends.}
 \centering
@@
 \multicolumn{2}{l}{\textbf{Annual fluxes}}\\
-\hspace{1em}$NEP$ & net ecosystem production or net ecosystem exchange (+ indicates C sink) & 377 & 163 & 89 & n.s. & +; xB\\
-\hspace{1em}$GPP$ & gross primary production ($NPP + R_{auto}$ or $R_{eco}-NEE$) & 309 & 117 & 72 & TrB > TeB $\ge$ TeN $\ge$ BoN & +; xB\\
-\hspace{1em}$NPP$ & net primary production ($ANPP$ + $BNPP$) & 292 & 165 & 74 & TrB > TeB $\ge$ TeN > BoN & +\\
-\hspace{1em}$ANPP$ & aboveground $NPP$ & 566 & 384 & 135 & TrB > TeB $\ge$ TeN $\ge$ BoN & +; xB\\
-\hspace{1em}$ANPP_{woody}$ & woody production ($ANPP_{stem}$ + $ANPP_{branch}$) & 126 & 114 & 37 & n.s. & +; xB\\
-\hspace{1em}$ANPP_{stem}$ & woody stem production & 458 & 411 & 118 & TrB > TeN $\ge$ TeB > BoN & -; xB\\
-\hspace{1em}$ANPP_{branch}$ & branch turnover & 125 & 112 & 42 & TrB > TeB $\ge$ TeN & n.s.\\
-\hspace{1em}$ANPP_{foliage}$ & foliage production, typically estimated as annual leaf litterfall & 275 & 231 & 89 & TrB > TeB $\ge$ TeN > BoN & +\\
-\hspace{1em}$ANPP_{litterfall}$ & litterfall, including leaves, reproductive structures, twigs, and sometimes branches & 147 & 117 & 55 & TrB > TeB & +\\
-\hspace{1em}$ANPP_{repro}$ & production of reproductive structures (flowers, fruits, seeds) & 72 & 65 & 35 & n.s. & -\\
-\hspace{1em}$ANPP_{folivory}$ & foliar biomass consumed by folivores & 24 & 16 & 11 & n.s. & -\\
-\hspace{1em}$M_{woody}$ & woody mortality--i.e., $B_{ag}$ of trees that die & 35 & 35 & 19 & n.s. & -\\
-\hspace{1em}$BNPP$ & belowground NPP ($BNPP_{coarse}$+ $BNPP_{fine}$) & 210 & 178 & 76 & TrB > TeB $\ge$ TeN $\ge$ BoN & +\\
-\hspace{1em}$BNPP_{coarse}$ & coarse root production & 144 & 111 & 36 & n.s. & n.s.\\
-\hspace{1em}$BNPP_{fine}$ & fine root production & 167 & 136 & 62 & n.s. & +\\
-\hspace{1em}$R_{eco}$ & ecosystem respiration ($R_{auto}$+ $R_{het}$) & 197 & 81 & 59 & TrB > TeN & +; xB\\
-\hspace{1em}$R_{auto}$ & autotrophic respiration ($R_{auto-ag}+ R_{root}$) & 36 & 35 & 15 & TrB > TeN & -\\
-\hspace{1em}$R_{auto-ag}$ & aboveground autotrophic respiration (i.e., leaves and stems) & 3 & 3 & 2 & TrB > TeN & -\\
-\hspace{1em}$R_{root}$ & root respiration & 227 & 183 & 98 & n.s. & +\\
-\hspace{1em}$R_{soil}$ & soil respiration ($R_{het-soil} + R_{root}$) & 644 & 468 & 219 & TrB > TeB $\ge$ TeN > BoN & n.s.\\
-\hspace{1em}$R_{het-soil}$ & soil heterotrophic respiration & 213 & 178 & 93 & TrB > TeB $\ge$ TeN & -; xB\\
+\hspace{1em}$NEP$ & net ecosystem production or net ecosystem exchange (+ indicates C sink) & 329 & 146 & 88 & n.s. & +; xB\\
+\hspace{1em}$GPP$ & gross primary production ($NPP + R_{auto}$ or $R_{eco}-NEE$) & 303 & 115 & 84 & TrB > TeB $\ge$ TeN $\ge$ BoN & +; xB\\
+\hspace{1em}$NPP$ & net primary production ($ANPP$ + $BNPP$) & 214 & 112 & 74 & TrB > TeB $\ge$ TeN > BoN & n.s.\\
+\hspace{1em}$ANPP$ & aboveground $NPP$ & 343 & 236 & 131 & TrB > TeB $\ge$ TeN > BoN & +; xB\\
+\hspace{1em}$ANPP_{woody}$ & woody production ($ANPP_{stem}$ + $ANPP_{branch}$) & 64 & 53 & 37 & n.s. & +\\
+\hspace{1em}$ANPP_{stem}$ & woody stem production & 217 & 190 & 117 & TrB > TeN $\ge$ TeB $\ge$ BoN & n.s.\\
+\hspace{1em}$ANPP_{branch}$ & branch turnover & 69 & 59 & 42 & TrB > TeB $\ge$ TeN & n.s.\\
+\hspace{1em}$ANPP_{foliage}$ & foliage production, typically estimated as annual leaf litterfall & 162 & 132 & 88 & TrB > TeB $\ge$ TeN > BoN & +\\
+\hspace{1em}$ANPP_{litterfall}$ & litterfall, including leaves, reproductive structures, twigs, and sometimes branches & 82 & 70 & 55 & n.s. & +\\
+\hspace{1em}$ANPP_{repro}$ & production of reproductive structures (flowers, fruits, seeds) & 51 & 44 & 34 & n.t. & n.t.\\
+\hspace{1em}$ANPP_{folivory}$ & foliar biomass consumed by folivores & 20 & 12 & 11 & n.t. & n.t.\\
+\hspace{1em}$M_{woody}$ & woody mortality--i.e., $B_{ag}$ of trees that die & 18 & 18 & 18 & n.t. & n.t.\\
+\hspace{1em}$BNPP$ & belowground NPP ($BNPP_{coarse}$+ $BNPP_{fine}$) & 148 & 116 & 79 & TrB > TeN $\ge$ TeB $\ge$ BoN & +\\
+\hspace{1em}$BNPP_{coarse}$ & coarse root production & 77 & 56 & 36 & TeN $\ge$ TrB & n.s.\\
+\hspace{1em}$BNPP_{fine}$ & fine root production & 123 & 99 & 66 & n.s. & +\\
+\hspace{1em}$R_{eco}$ & ecosystem respiration ($R_{auto}$+ $R_{het}$) & 213 & 98 & 70 & TrB > TeB $\ge$ TeN & +\\
+\hspace{1em}$R_{auto}$ & autotrophic respiration ($R_{auto-ag}+ R_{root}$) & 24 & 23 & 15 & n.t. & n.t.\\
+\hspace{1em}$R_{auto-ag}$ & aboveground autotrophic respiration (i.e., leaves and stems) & 2 & 2 & 1 & n.t. & n.t.\\
+\hspace{1em}$R_{root}$ & root respiration & 181 & 139 & 95 & TrB $\ge$ TeB & +\\
+\hspace{1em}$R_{soil}$ & soil respiration ($R_{het-soil} + R_{root}$) & 627 & 411 & 229 & TrB > TeB > TeN $\ge$ BoN & n.s.\\
+\hspace{1em}$R_{het-soil}$ & soil heterotrophic respiration & 197 & 156 & 100 & TrB > TeB $\ge$ TeN & n.s.\\
 \hspace{1em}$R_{het-ag}$ & aboveground heterotrophic respiration & 0 & 0 & 0 & - & -\\
@@
 \multicolumn{2}{l}{\textbf{Stocks}}\\
-\hspace{1em}$B_{tot}$ & total live biomass ($B_{ag}$+$B_{root}$) & 299 & 252 & 85 & TeN > TeB $\ge$ TrB $\ge$ BoN & +; xB\\
-\hspace{1em}$B_{ag}$ & aboveground live biomass  ($B_{ag-wood}$+$B_{foliage}$) & 5371 & 4761 & 621 & TrB $\ge$ TeN $\ge$ TeB > BoN & +; xB\\
-\hspace{1em}$B_{ag-wood}$ & woody component of aboveground biomass & 224 & 210 & 66 & TeN > TrB $\ge$ TeB $\ge$ BoN & +; xB\\
-\hspace{1em}$B_{foliage}$ & foliage biomass & 260 & 216 & 74 & TeN > TrB $\ge$ BoN $\ge$ TeB & +; xB\\
-\hspace{1em}$B_{root}$ & total root biomass ($B_{root-coarse}$+$B_{root-fine}$) & 2380 & 2337 & 359 & n.s. & +; xB\\
-\hspace{1em}$B_{root-coarse}$ & coarse root biomass & 204 & 190 & 68 & TeN > TeB $\ge$ BoN & +; xB\\
-\hspace{1em}$B_{root-fine}$ & fine root biomass & 335 & 266 & 107 & n.s. & n.s.\\
-\hspace{1em}$DW_{tot}$ & deadwood ($DW_{standing}$+$DW_{down}$) & 223 & 212 & 42 & TeB > TrB & +; xB\\
-\hspace{1em}$DW_{standing}$ & standing dead wood & 65 & 61 & 22 & TeB > TrB & -\\
-\hspace{1em}$DW_{down}$ & fallen dead wood, including coarse and sometimes fine woody debris & 284 & 271 & 37 & TeB > TrB & +; xB\\
-\hspace{1em}$OL$ & organic layer $/$ litter$/$ forest floor & 555 & 482 & 113 & n.s. & +; xB\\
+\hspace{1em}$B_{tot}$ & total live biomass ($B_{ag}$+$B_{root}$) & 188 & 157 & 87 & TrB $\ge$ TeB > BoN & +; xB\\
+\hspace{1em}$B_{ag}$ & aboveground live biomass  ($B_{ag-wood}$+$B_{foliage}$) & 4466 & 4072 & 621 & TrB $\ge$ TeN $\ge$ TeB > BoN & +; xB\\
+\hspace{1em}$B_{ag-wood}$ & woody component of aboveground biomass & 115 & 102 & 64 & TeN > TrB $\ge$ BoN & +; xB\\
+\hspace{1em}$B_{foliage}$ & foliage biomass & 134 & 115 & 72 & TeN > TrB $\ge$ BoN $\ge$ TeB & +; xB\\
+\hspace{1em}$B_{root}$ & total root biomass ($B_{root-coarse}$+$B_{root-fine}$) & 2329 & 2298 & 360 & n.s. & +; xB\\
+\hspace{1em}$B_{root-coarse}$ & coarse root biomass & 134 & 120 & 73 & TeN > TeB $\ge$ BoN & +; xB\\
+\hspace{1em}$B_{root-fine}$ & fine root biomass & 226 & 180 & 109 & n.s. & +; xB\\
+\hspace{1em}$DW_{tot}$ & deadwood ($DW_{standing}$+$DW_{down}$) & 79 & 73 & 42 & n.t. & +; xB\\
+\hspace{1em}$DW_{standing}$ & standing dead wood & 36 & 35 & 22 & n.t. & n.t.\\
+\hspace{1em}$DW_{down}$ & fallen dead wood, including coarse and sometimes fine woody debris & 278 & 265 & 37 & n.t. & +; xB\\
+\hspace{1em}$OL$ & organic layer $/$ litter$/$ forest floor & 474 & 413 & 115 & n.s. & +; xB\\
 \bottomrule
-\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{*} Tr: Tropical, TeB: Temperate Broadleaf, TeN: Temperate Needleleaf, B: Boreal, n.s.: no significant differences}\\
-\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{\dag} + or -: significant positive or negative trend, xB: significant age x biome interaction, n.s.: no significant age trend}\\
+\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{*} Tr: Tropical, TeB: Temperate Broadleaf, TeN: Temperate Needleleaf, B: Boreal, n.s.: no significant differences, n.t.: not tested }\\
+\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{\dag} + or -: significant positive or negative trend, xB: significant age x biome interaction, n.s.: no significant age trend, n.t.: not tested}\\
 \end{tabular}}
@@
 
-For this analysis, we grouped forests into four broad biome types
-(tropical broadleaf, temperate broadleaf, temperate needleleaf, and
-boreal needleleaf) and two age classifications (young and mature).
-Climate zones (Fig. 1) were defined based on site geographic coordinates
-according to Köppen-Geiger zones (Rubel and Kottek 2010). Tropical
-climates were defined to include all equatorial (A) zones, temperate
-climates were defined to include all warm temperate (C) zones and warmer
-snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates
-were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd,
+We grouped forests into four broad biome types based on climate zones
+and dominant vegetation type (tropical broadleaf, temperate broadleaf,
+temperate needleleaf, and boreal needleleaf) and two age classifications
+(young and mature). Climate zones (Fig. 1) were defined based on site
+geographic coordinates according to Köppen-Geiger zones (Rubel and
+Kottek 2010). We defined the tropical biome as including all equatorial
+(A) zones, temperate biomes as including all warm temperate (C) zones
+and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and the
+boreal biome as including the colder snow climates (Dsc, Dsd, Dwc, Dwd,
 Dfc, and Dfd). Any forests in dry (B) and polar (E) Köppen-Geiger zones
-were excluded from the analysis. Leaf type (broadleaf / needleleaf) was
-defined based on descriptions in original publications (prioritized) or
-values extracted from a global map based on satellite observations
-(SYNMAP; {\textbf{???}}). For young tropical forests imported from GROA
-but not yet classified by leaf type, we assumed that all were broadleaf,
-consistent with the rarity of naturally regenerating needleleaf forests
-in the tropics (\textbf{REF}). Forests were classified as ``young''
-(\textless{} 100 years) or ``mature'' (\(\ge\) 100 years or classified
-as ``mature'', ``old growth'', ``intact'', or ``undisturbed'' in
-original publication). Assigning stands to these groupings required the
-exclusion of records for which ForC lacked records of stand age (4.8\%
-of records in full database), or geographic coordinates (0.4\% of sites
-in full database). We also excluded records of stand age = 0 year (0.8\%
-of records in full database). In total, our analysis retained 76.2 of
-the focal variable records for forests of known age. Numbers of records
-by biome and age class are given in Table S1.
+were excluded from the analysis. We defined leaf type (broadleaf /
+needleleaf) was based on descriptions in original publications
+(prioritized) or values extracted from a global map based on satellite
+observations (SYNMAP; Jung \emph{et al} 2006). For young tropical
+forests imported from \emph{GROA} but not yet classified by leaf type,
+we assumed that all were broadleaf, consistent with the rarity of
+naturally regenerating needleleaf forests in the tropics. We also
+classified forests as ``young'' (\textless{} 100 years) or ``mature''
+(\(\ge\) 100 years or classified as ``mature'', ``old growth'',
+``intact'', or ``undisturbed'' in original publication). Assigning
+stands to these groupings required the exclusion of records for which
+\emph{ForC} lacked geographic coordinates (0.4\% of sites in full
+database) or records of stand age (5.7\% of records in full database).
+We also excluded records of stand age = 0 year (0.8\% of records in full
+database). In total, our analysis retained 76.1 of the focal variable
+records for forests of known age. Numbers of records by biome and age
+class are given in Table S1.
 
 Data were summarized to produce schematics of C cycling across the eight
-biome - age group combinations identified above. To obtain the values
+biome by age group combinations identified above. To obtain the values
 reported in the C cycle schematics, we first averaged any repeated
-measurements within a plot. Values were then averaged across plots
-clustered within 25 km of one another (\emph{sensu} Anderson-Teixeira
-\emph{et al} 2018), weighting by area sampled if available for all
-records. This step was taken to avoid pseudo-replication and to combine
-any records from sites with more than one name in ForC.
-
-To test for differences across mature forest biomes, the original values
-were analyzed via a linear mixed effects model {[}`lmer' function in
-`lme4' R package version \textbf{x.xx}; \textbf{REF}{]} with biome as
-fixed effect and plot nested within geographic.area as random effects on
-the intercept. When Biome had a significant effect, we looked at a
-Tukey's pairwise comparison to see which biomes were significantly
-different from one another. This analysis was run for variables with
-records for \$\ge\$7 distinct geographic areas in more than one biome,
-excluding any biomes that failed this criteria (Table 1).
+measurements within a plot. Values were then averaged across
+geographically distinct areas, defined as plots clustered within 25 km
+of one another (\emph{sensu} Anderson-Teixeira \emph{et al} 2018),
+weighting by area sampled if available for all records. This step was
+taken to avoid pseudo-replication.
+
+We tested whether the C budgets described above ``closed''--\emph{i.e.},
+whether they were internally consistent. Specifically, we first defined
+relationships among variables: for example, \(NEP = GPP - R_{eco}\),
+\(BNPP = BNPP_{coarse} + BNPP_{fine}\),
+\(DW_{tot} = DW_{standing} + DW_{down}\)). Henceforth, we refer to the
+variables on the left side of the equation as ``aggregate'' fluxes or
+stocks, and those that are summed as ``component'' fluxes or stocks,
+noting that the same variable can take both aggregate and component
+positions in different relationships. We considered the C budget for a
+given relationship ``closed'' when component variables summed to within
+one standard deviation of the aggregate variable.
+
+To test for differences across mature forest biomes, we also examined
+how stand age impacted fluxes and stocks, employing a mixed effects
+model (``lmer'' function in ``lme4'' R package; Bates \emph{et al} 2015)
+with biome as fixed effect and plot nested within geographic.area as
+random effects on the intercept. When Biome had a significant effect, we
+looked at a Tukey's pairwise comparison to see which biomes were
+significantly different from one another. This analysis was run for
+variables with records for at least seven distinct geographic areas in
+more than one biome, excluding any biomes that failed this criteria
+(Table 1).
 
@@
 and plot nested within geographic.area as a random effect on the
-intercept. This analysis was run for variables with records for \$\ge\$3
-distinct geographic areas in more than one biome, excluding any biomes
-that failed this criteria (Table 1). When the effect of stand age was
-significant at p \(\le\) 0.05 and when each biome had records for stands
-of at least 10 different ages, a biome * stand.age interaction was
-included in the model. All database manipulation, analyses, and figure
-production were fully automated in R (\textbf{version, citation}).
+intercept. This analysis was run for variables with records for at least
+three distinct geographic areas in more than one biome, excluding any
+biomes that failed this criteria (Table 1). When the effect of stand age
+was significant at p \(\le\) 0.05 and when each biome had records for
+stands of at least 10 different ages, a biome \(\times\) stand.age
+interaction was included in the model.
+
+To facilitate the accessibility of our results and data, and to allow
+for rapid updates as additional data become available, we have automated
+all database manipulation, analyses, and figure production in R (R Core
+Team 2020).
 
@@
 
-Of the 47846 records in ForC v3.0, 14847 met the criteria for inclusion
-in this study (Fig. 1). These records were distributed across 5889 plots
-in 874 distinct geographic areas. Of the 23 flux and 11 stock variables
-mapped in these diagrams, ForC contained mature forest estimates from
-\(\ge\) 7 distinct geographic areas for 20 fluxes and 9 stocks in
-tropical broadleaf forests, 14 fluxes and 9 stocks in temperate
-broadleaf forests, 15 fluxes and 8 stocks in temperate conifer forests,
-and 8 fluxes and 7 stocks in boreal forests. For regrowth forests
-(\textless100 yrs), ForC contained estimates from \(\ge\) 7 distinct
-geographic areas for 3 fluxes and 9 stocks in tropical broadleaf
-forests, 12 fluxes and 8 stocks in temperate broadleaf forests, 15
-fluxes and 8 stocks in temperate conifer forests, and 12 fluxes and 7
-stocks in boreal forests.
-\textbf{(\href{https://github.com/forc-db/ERL-review/issues/29\#issuecomment-699633801}{issue
-29})}
+Of the 39762 records in \emph{ForC} v3.0, 11923 met our strict criteria
+for inclusion in this study (Fig. 1). These records were distributed
+across 5062 plots in 865 distinct geographic areas. Of the 23 flux and
+11 stock variables mapped in these diagrams, \emph{ForC} contained
+sufficient mature forest data for inclusion in our statistical analyses
+(\emph{i.e.}, records from \(\ge\) 7 distinct geographic areas) for 20
+fluxes and 9 stocks in tropical broadleaf forests, 15 fluxes and 8
+stocks in temperate broadleaf forests, 14 fluxes and 7 stocks in
+temperate conifer forests, and 8 fluxes and 7 stocks in boreal forests.
+For regrowth forests (\textless100 yrs), \emph{ForC} contained
+sufficient data for inclusion in our statistical analyses (\emph{i.e.},
+records from \(\ge\) 3 distinct geographic areas) for 11 fluxes and 10
+stocks in tropical broadleaf forests, 16 fluxes and 10 stocks in
+temperate broadleaf forests, 16 fluxes and 10 stocks in temperate
+conifer forests, and 14 fluxes and 9 stocks in boreal forests.
 
@@
 
-Average C cycles for tropical broadleaf, temperate broadleaf, temperate
-conifer, and boreal forests \(\ge\) 100 years old and with no known
-major disturbance or significant anthropogenic management are presented
-in Figures 2-5 (and available in tabular format in the ForC release
+Average C cycles for mature tropical broadleaf, temperate broadleaf,
+temperate conifer, and boreal forests \(\ge\) 100 years old and with no
+known major natural or anthropogenic disturbance are presented in
+Figures 2-5 (and available in tabular format in the \emph{ForC} release
 accompanying this publication:
@@
 For variables with records from \(\ge\) 7 distinct geographic areas,
-these ensemble C budgets were generally consistent. \emph{That is,
-component fluxes and stocks summed to within one standard deviation of
-more inclusive fluxes in all but} \textbf{two} \emph{instances, both in
-temperate conifer forests (Fig. 5).} \textbf{(check this with final
-results:
-\url{https://github.com/forc-db/ForC/blob/master/numbers_and_facts/C_cycle_closure.csv})
-} \emph{Within this biome, \(B_{ag}<B_{ag-wood}+B_{foliage}\) and
-\(B_{root}<B_{root-coarse}+B_{root-fine}\) because \(B_{ag-wood}\) and
-\(B_{root-coarse}\) were very high, with strongly disproportionate
-numbers of records from the high-biomass forests of the US Pacific
-Northwest (Figs. S18, S21).}
+these ensemble C budgets were generally consistent. That is, component
+variables summed to within one standard deviation of their respective
+aggregate variables in all but one instance. In the temperate conifer
+biome, the average composite measure of root biomass (\(B_{root}\)) was
+less than the combined average value of coarse and fine root biomass
+(\(B_{root-coarse}\) and \(B_{root-fine}\), respectively). This lack of
+closure was driven by very high estimates of \(B_{root-coarse}\) from
+high-biomass forests of the US Pacific Northwest.
 
@@
 \caption{Figure 2 | C cycle diagram for mature tropical broadleaf forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.
-Asterisk after variable name indicates lack of C cycle closure.}\label{fig:unnamed-chunk-8}
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate broadleaf MATURE}
-
-}
-
-\caption{Figure 3 | C cycle diagram for mature temperate broadleaf forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.
 Asterisk after variable name indicates lack of C cycle closure.}\label{fig:unnamed-chunk-9}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate conifer MATURE}
+{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate broadleaf MATURE}
 
@@
 
-\caption{Figure 4 | C cycle diagram for mature temperate conifer forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
+\caption{Figure 3 | C cycle diagram for mature temperate broadleaf forests.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.
 Asterisk after variable name indicates lack of C cycle closure.}\label{fig:unnamed-chunk-10}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Boreal conifer MATURE}
+{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate conifer MATURE}
 
@@
 
-\caption{Figure 5 | C cycle diagram for mature boreal conifer forests.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks), presented as mean ± std, where geographically distinct areas are treated as the unit of replication.
-Arrows indicate fluxes, boxes indicate stocks.
+\caption{Figure 4 | C cycle diagram for mature temperate conifer forests.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.
 Asterisk after variable name indicates lack of C cycle closure.}\label{fig:unnamed-chunk-11}
 \end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\linewidth]{/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Boreal conifer MATURE}
+
+}
+
+\caption{Figure 5 | C cycle diagram for mature boreal conifer forests.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.
+Asterisk after variable name indicates lack of C cycle closure.}\label{fig:unnamed-chunk-12}
+\end{figure}
 \end{landscape}
 
-\textbf{check this with final data (based on Table 1):} There were
-sufficient data to assess mature forest biome differences for
-\textbf{20} flux variables, and significant differences among biomes
-were detected for \textbf{14} variables (Table 1). With only
-\textbf{one} exception (\(R_{auto}\), which had low sample size;
-\textbf{\href{https://github.com/forc-db/ERL-review/issues/40}{issue
-40}}), C fluxes were highest in tropical forests, intermediate in
-temperate (broadleaf or conifer) forests, and lowest in boreal forests
-(Table 1, Figs. 6, S1-S15). Differences between tropical and boreal
-forests were always significant, with temperate forests intermediate and
-significantly different from one or both. Fluxes tended to be
-numerically greater in temperate broadleaf than conifer forests, but the
-difference was never statistically significant. This pattern held for
-the following variables: ** \(GPP\), \(NPP\), \(ANPP\),
-\(ANPP_{woody}\), \(ANPP_{stem}\),\(ANPP_{branch}\), \(ANPP_{foliage}\),
-\(ANPP_{litterfall}\), \(M_{woody}\), \(BNPP\), \(R_{eco}\),
-\(R_{soil}\), and \(R_{het-soil}\)**. For variables without significant
-differences among biomes, the same general trends applied.
-
-The most notable exception to this pattern was \(NEP\), with no
-significant differences across biomes but with the largest average in
-temperate broadleaf forests, followed by temperate conifer, boreal, and
-tropical forests (Figs. 5,S1). Another exception was for
+There were sufficient data to assess mature forest biome differences for
+15 flux variables, and significant differences among biomes were
+detected for 12 variables (Table 1). In all of these cases--including C
+fluxes into, within, and out of the ecosystem--C fluxes were highest in
+tropical forests, intermediate in temperate (broadleaf or conifer)
+forests, and lowest in boreal forests (Table 1, Figs. 6, S1-S15).
+Differences between tropical and boreal forests were always significant,
+with temperate forests intermediate and significantly different from one
+or both. Fluxes tended to be numerically greater in temperate broadleaf
+than conifer forests, but the difference was never statistically
+significant. This pattern held for the following variables: \(GPP\),
+\(NPP\), \(ANPP\), \(ANPP_{stem}\),\(ANPP_{branch}\),
+\(ANPP_{foliage}\), \(BNPP\), \(R_{eco}\), \(R_{root}\), \(R_{soil}\),
+and \(R_{het-soil}\). For two of the variables without significant
+differences among biomes (\(ANPP_{litterfall}\) and \(BNPP_{fine}\);
+Figs. S8 and S11, respectively), the same general trends applied but
+were not statistically significant. Another exception was for
 \(BNPP_{root-coarse}\), where all records came from high-biomass forests
-in the US Pacific Northwest, resulting in high values for the temperate
-conifer biome and no significant differences across biomes (Fig. S10).\\
-Thus, C cycling rates generally decreased from tropical to temperate to
-boreal forests, with the important exception in the overall C balance
-(\(NEP\)).
+in the US Pacific Northwest, resulting in marginally higher values for
+the temperate conifer biome (Table 1, Fig. S10; differences significant
+in mixed effects model but not in post-hoc pairwise comparison).
+
+The most notable exception to the pattern of decreasing flux from
+tropical to boreal biomes was \(NEP\), with no significant differences
+across biomes but with the largest average in temperate broadleaf
+forests, followed by tropical, boreal, and temperate conifer forests
+(Figs. 5,S1).
 
@@
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
-Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
-Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
-Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.
+In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+The fitted line indicates the effect of age on flux (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant age x biome interaction.
+The boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
 Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).
-Individual figures for each flux with sufficient data given in the Supplement (Figs.
-S1-S15).}\label{fig:unnamed-chunk-12}
+Individual figures for each flux with sufficient data are given in the Supplement (Figs.
+S1-S15).}\label{fig:unnamed-chunk-13}
 \end{figure}
 
-\textbf{check this with final data (based on Table 1):} There were
-sufficient data to assess mature forest biome differences for \textbf{9}
-stock variables, and significant differences among biomes were detected
-for \textbf{6} variables (\textbf{\(B_{tot}\), \(B_{ag}\),
-\(B_{ag-wood}\), \(B_{foliage}\), \(B_{root-coarse}\), \(DW_{tot}\)};
-Table 1). C stocks had less consistent patterns across biomes (Figs. 7,
-S16-S26). In \textbf{four of the six} cases (\textbf{\(B_{tot}\),
-\(B_{ag-wood}\), \(B_{root-coarse}\), \(DW_{tot}\)}), temperate conifer
-forests were in the highest significance grouping, and boreal forests in
-the lowest, with tropical and temperate broadleaf forests in
-between--most commonly being significantly different from temperate
-conifer but not boreal forests. For \(B_{ag}\), which had by far the
-highest sample size, tropical forests exceeded temperate conifer forests
-(but not significantly). For \(B_{foliage}\),temperate broadleaf forests
-were lowest (again, not significantly). The high values for the
-temperate conifer biome were driven by the very high-biomass forests of
-the US Pacific Northwest, which are disproportionately represented in
-the current version of ForC. Thus, biome differences should be
-interpreted more as driven more by geographic distribution of sampling
-than by true differences.
+There were sufficient data to assess mature forest biome differences for
+nine stock variables, and significant differences among biomes were
+detected for five variables (\(B_{tot}\), \(B_{ag}\), \(B_{ag-wood}\),
+\(B_{foliage}\), \(B_{root-coarse}\); Table 1). C stocks had less
+consistent patterns across biomes (Figs. 7, S16-S26). For \(B_{tot}\)
+and \(B_{ag}\), tropical broadleaf forests had the highest biomass and
+boreal forests the lowest, with temperate broadleaf and needleleaf
+(\(B_{ag}\) only) intermediate. For three variables that had been
+disproportionately sampled in the high-biomass forests of the US Pacific
+Northwest (\(B_{ag-wood}\), \(B_{foliage}\), and \(B_{root-coarse}\)),
+temperate conifer forests had significantly higher stocks than the other
+biomes, which were not significantly different from one another.
 
@@
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
-Left plot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
-Solid lines indicate signficant effect of age, non-pareallel lines indicate a significant age x biome interaction.
-Boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.
+In each panel, the left scatterplot  shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
+The fitted line indicates the effect of age on flux (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant age x biome interaction.
+The boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.
 Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).
-Individual figures for each stock with sufficient data given in the Supplement (Figs.
-S16-S26).}\label{fig:unnamed-chunk-13}
+Individual figures for each stock with sufficient data are given in the Supplement (Figs.
+S16-S26).}\label{fig:unnamed-chunk-14}
 \end{figure}
@@
 Average C cycles for forests \textless100 years old are presented in
-Figures 8-11.\\
-Both C stocks and fluxes commonly displayed significant trends with
-stand age for within-biome analyses (Tables 1, S2, Figs. 6-11, S1-S26;
-detailed below).
-
-ForC contained \emph{14} flux variables with sufficient data for
-cross-biome analyses of age trends in regrowth forests (see Methods)
-(Fig. 6-7 and \textbf{S\#- SI figures including plots for all
-variables}). Of these, \emph{9} increased significantly with
-log10{[}stand.age{]}: \(GPP\), \(NPP\), \(ANPP\), \(ANPP_{foliage}\),
-\(ANPP_{woody}\), \(ANPP_{woody-stem}\), \(BNPP\), \(BNPP_{root-fine}\),
-\(R_{eco}\), and net C sequestration (\(NEP\)). The remaining
-five--\(ANPP_{woody-branch}\), \(BNPP_{root-coarse}\), \(R_{soil-het}\),
-and \(R_{soil-het}\)-\/--displayed no significant relationship to stand
-age, although all displayed a positive trend.
+Figures 8-11. Both C stocks and fluxes commonly increased significantly
+with stand age (Tables 1, S2, Figs. 6-11, S1-S26; detailed below).
+
+\emph{ForC} contained 16 C flux variables with sufficient data for
+analyses of age trends in young forests (see Methods) (Figs. 6, S1-S15).
+Of these, ten increased significantly with age: \(NEP\), \(GPP\),
+\(ANPP\), \(ANPP_{woody}\), \(ANPP_{foliage}\), \(ANPP_{litterfal}\),
+\(BNPP\), \(BNPP_{fine}\), \(R_{eco}\), and \(R_{root}\). The remaining
+six--\(NPP\), \(ANPP_{stem}\), \(ANPP_{branch}\), \(BNPP_{coarse}\),
+\(R_{soil-het}\), and \(R_{soil-het}\)-\/--displayed no significant
+relationship to stand age.
 
@@
 observed for mature forests, with C cycling generally most rapid in the
-tropics and slowest in boreal forests.\\
-The single exception was \(ANPP_{stem}\), for which temperate broadleaf
-forests and temperate conifer forests of age
-\textgreater{}\emph{\textasciitilde30} had slightly higher flux rates
-than tropical forests (\emph{represented by a single chronosequence}).
-Notably, the trend of tropical \textgreater{} temperate \textgreater{}
-boreal held for \(NEP\) in regrowth forests, in contrast to the lack of
-biome differences in \(NEP\) for mature forests (Fig. 6).
-
-There were only \emph{\#\#} flux variables with sufficient data to test
-for biome x age interactions: \(ANPP\), \(ANPP_{woody}\),
-\(ANPP_{stem}\), \(ANPP_{litterfall}\), and \(BNPP\) (Table S2).
-\textbf{(more could be added if age trends become significant after
-outliers are resolved)} For three of these (\(ANPP\),
-\(ANPP_{litterfall}\), \(BNPP\)), the increase in C flux with age was
-steepest increase in tropical forests, followed by temperate and then
-boreal forests (Figs S\#). Similarly, \(ANPP_{woody}\) displayed a
-steeper increase with age in temperate than boreal boreal forests (no
-tropical data for this variable). In contrast, for \(ANPP_{stem}\),
-tropical and temperate broadleaf forests had the highest flux rates at
-young ages, but were surpassed by temperate conifer forests between ages
-20 and 50 (Fig. S6).
-
-(\textbf{this needs to be updated with latest data}) In terms of C
-stocks, \emph{10} variables had sufficient data to test for age trends.
-\emph{Six} of these---\emph{total biomass, aboveground biomass,
-aboveground woody biomass, foliage biomass, root biomass, and coarse
-root biomass}--- increased significantly with \(log10[stand.age]\).
-\emph{The remaining four displayed non-significant positive trends: fine
-root biomass, total dead wood, standing dead wood, and organic layer.}
-\emph{(discuss rates of increase)}
+tropics and slowest in boreal forests. The single exception was
+\(ANPP_{stem}\), for which temperate broadleaf and conifer forests had
+similar flux rates than tropical forests. Notably, and in contrast to
+the lack of biome differences in \(NEP\) for mature forests (Fig. 6),
+the tendency for temperate forests to have greater fluxes than boreal
+forests held for \(NEP\) in regrowth forests (tropical forests excluded
+because of insufficient data).
+
+In terms of C stocks, ten variables (all but standing deadwood,
+\(DW_{standing}\)) had sufficient data to test for age trends (Table 1,
+Figs. 7, S16-26). All of these displayed a significant overall increase
+with with \(log10[stand.age]\). There were sufficient data to model age
+\(\times\) biome interactions were also significant for all ten of these
+C stock variables (Table S2), with living C stocks tending to accumulate
+more rapidly during the early stages of forest regrowth in tropical
+forests (Figs. 7, S16-S22). In the case of two non-living C stocks
+(\(DW_{down}\) and \(OL\)), age \(\times\) biome interactions were such
+that Specifically, \(DW_{down}\) declined with age in temperate and
+boreal forests, compared to an increase with age in tropical forests
+(Figs. 7, S25). Similarly, \(OL\) declined slightly with age in
+temperate broadleaf forests, contrasting an increase in the other three
+biomes (Figs. 7, S26).
 
@@
 \caption{Figure 8 | C cycle diagram for young tropical broadleaf forests.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
 Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
-Arrows indicate fluxes, boxes indicate stocks.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.}\label{fig:unnamed-chunk-14}
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.}\label{fig:unnamed-chunk-15}
 \end{figure}
@@
 \caption{Figure 9 | C cycle diagram for young temperate broadleaf forests.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
 Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
-Arrows indicate fluxes, boxes indicate stocks.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.}\label{fig:unnamed-chunk-15}
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.}\label{fig:unnamed-chunk-16}
 \end{figure}
@@
 \caption{Figure 10 | C cycle diagram for young temperate conifer forests.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
 Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
-Arrows indicate fluxes, boxes indicate stocks.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.}\label{fig:unnamed-chunk-16}
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.}\label{fig:unnamed-chunk-17}
 \end{figure}
@@
 \caption{Figure 11 | C cycle diagram for young boreal conifer forests.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
 Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
-Arrows indicate fluxes, boxes indicate stocks.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-Arrow size is proportional to the square root of corresponding flux.}\label{fig:unnamed-chunk-17}
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.}\label{fig:unnamed-chunk-18}
 \end{figure}
@@
 
-ForC v3.0 brought together an unprecedented amount of data to yield an
-internally consistent picture of C cycling in the world's major forest
-biomes. Carbon cycling rates generally increased from boreal to tropical
-regions and with stand age. Specifically, the major C fluxes were
-highest in tropical forests, intermediate in temperate (broadleaf or
-conifer) forests, and lowest in boreal forests -- a pattern that
-generally held for regrowth as well as mature forests (Figs. 6-7). In
-contrast to C fluxes, there was little directional variation in mature
-forest C stocks across biomes (Figs. 2-5, 7). The majority of flux
-variables, together with most live biomass pools, increased
-significantly with stand age (Figs. 6-11). Together, these results
-indicate that, moving from cold to tropical climates and from young to
-old stands, there is a general acceleration of C cycling, whereas C
-stocks and \(NEP\) of mature forests are correlated with a different set
-of factors.
+\emph{ForC} v3.0 provided unprecedented coverage of most major
+variables, yielding an internally consistent picture of C cycling in the
+world's major forest biomes. Carbon cycling rates generally increased
+from boreal to tropical regions and with stand age. Specifically, the
+major C fluxes were highest in tropical forests, intermediate in
+temperate (broadleaf or conifer) forests, and lowest in boreal forests
+-- a pattern that generally held for regrowth as well as mature forests
+(Figs. 6-7). In contrast to C fluxes, there was little directional
+variation in mature forest C stocks across biomes (Figs. 2-5, 7). The
+majority of flux variables, together with most live biomass pools,
+increased significantly with stand age (Figs. 6-11). Together, these
+results indicate that, moving from cold to tropical climates and from
+young to old stands, there is a general acceleration of C cycling,
+whereas C stocks and \(NEP\) of mature forests are correlated with a
+different set of factors.
 
@@
 
-ForC provides provides unprecedented coverage of most major variables.
-\emph{(discuss how this improves upon previous data compilations/ for
-which variables does ForC make the greatest difference (e.g., not AGB or
-NEP/GPP/Reco, but by far the latest data compilation for dead wood,
-{[}variables{]})} \emph{(Noteable holes include: fluxes: R\_auto\_ag,
-woody mortality, folivary/ herbivory and respiration of herbivores (and
-therefore total R\_het), ANPP\_repro; also fluxes in tropical regrowth
-forests)} \emph{For the C stocks considered here, the most poorly
-covered is dead wood (none in E hemisphere!), despite a focused effort
-on this variable that has resulted in ForC being by far the largest
-collection of these data.} Thus, overall, we're lacking coverage of
-fluxes to herbivores and higher consumers, along with the woody
-mortality and dead wood. Geographically, all variables poorly covered in
-Africa and Siberia.
-
-\textbf{notes from Ben on the above par:} Pregitzer and Euskirchen 2004
-\url{http://dx.doi.org/10.1111/j.1365-2486.2004.00866.x} ``Aggregated
-biome-level estimates of NPP and NEP were higher in intermediate-aged
-forests (e.g., 30--120 years), while older forests (e.g., 4120 years)
-were generally less productive. The mean NEP in the youngest forests
-(0--10 years) was negative (source to the atmosphere) in both boreal and
-temperate biomes\ldots Forest age is a highly significant source of
-variability in NEP at the biome scale''
-
-Amiro et al.~2010 \url{http://dx.doi.org/10.1029/2010JG001390} Houghton
-et al.~2020 \url{https://doi.org/10.1111/gcb.15050}
-
-Turnover: Pugh et al.~2020 \url{http://dx.doi.org/10.5194/bg-2019-491}
-Yu et al.~2019 \url{http://dx.doi.org/10.1073/pnas.1821387116}
-
-Human footprint in global forests
-\url{http://dx.doi.org/10.1038/nature05847}
-\url{http://dx.doi.org/10.1038/nature02619}
-
-Mention consistency (or lack of) with e.g.~GOLUM-CNP?
-\url{http://dx.doi.org/10.5194/gmd-11-3903-2018}
-
-In terms of C stocks, there is a paucity of data on dead wood and
-organic layer (Pan et al.~?). These can be significant. (\textbf{Note
-that we've given a lot of emphasis to dead wood (work by Abby, and also
-Jenny), and as a result this work really advances knowledge of dead
-wood. We'll want to highlight that here.}) \emph{(give some stats/ cite
-figures)}. ForC does not include soil carbon, which is covered by other
-efforts (REFS). For fluxes, Fluxnet is the keeper of the best data on
-NEE, GPP, Reco (REFS), and SRDB remains the authority on soil
-respiration (REFS). ForC includes recent data from both, but is not
-continuously integrated. For C is the best source for most of the
-subsidiary fluxes: NPP, woody mortality\ldots{}
-
-The C cycle budgets for mature forests (Figs. 2-5) generally
-``close''--that is, the sums of component variables do not differ from
-the larger fluxes by more than one standard deviation. However, standard
-deviations are often large, reflective of significant within-biome
-variation. This makes the standard for closure relatively loose.
-\emph{Lack of closure, in the few instances where it occurs, is probably
-more reflective of differences in the representation of forest types
-(e.g., disproportionate representation of US Pacific NW for aboveground
-woody biomass relative to AGB; Fig. 4) than of methodological accuracy.}
-Thus, overall, a high degree of closure implies that ForC gives a
-consistent picture of C cycling within biomes. While these means are
-unlikely to be accurate representations of C cycling within any
-particular forest, they provide a useful baseline for comparison, always
-keeping in mind that sample means do not necessarily represent the true
-mean of the entire biome.
+The large number of C cycle variables covered by ForC, and the general
+consistency between them, provide confidence that our overall reported
+means provide accurate and useful baselines for analysis -- with the
+caveats that they are unlikely to be accurate representations of C
+cycling for any particular forest, and that these sample means almost
+certainly do not represent true biome means (particularly for temperate
+conifer forests where high-biomass stands are over-represented in
+\emph{ForC}).
+
+There are of course notable holes in the ForC variable coverage (Fig. 1)
+that limit the scope of our inferences here. Notably, \emph{ForC}
+currently has sparse--if any--coverage of fluxes to herbivores and
+higher consumers, along with the woody mortality (\(M_{woody}\)) and
+dead wood stocks (Tables 1, Figs. S23-S25). Geographically, all
+variables are poorly covered in Africa and Siberia, a common problem in
+the carbon-cycle community (Xu and Shang 2016, Schimel \emph{et al}
+2015). \emph{ForC} does not include soil carbon, which is covered by
+other efforts (e.g., Köchy \emph{et al} 2015). \emph{ForC} is not
+intended to replace databases that are specialized for particular parts
+of the C cycle analyses, e.g., aboveground biomass (Spawn \emph{et al}
+2020), land-atmosphere fluxes (Baldocchi \emph{et al} 2001), soil
+respiration (Jian \emph{et al} 2020), or the human footprint in global
+forests (Magnani \emph{et al} 2007).
+
+In this analysis, the C cycle budgets for mature forests (Figs. 2-5)
+generally ``close''---that is, the sums of component variables do not
+differ from the larger fluxes by more than one standard deviation. On
+the one hand, this reflects the general fact that ecosystem-scale
+measurements tend to close the C budget more easily and consistently
+than, for example, for energy balance (Stoy \emph{et al} 2013). On the
+other, however, \emph{ForC} derives data from multiple heterogeneous
+sources, and standard deviations within each biome are high; as a
+result, the standard for C closure is relatively loose (\emph{c.f.}
+Houghton 2020). Nonetheless, the lack of closure, in the one instance
+where it occurs, is probably more reflective of differences in the
+representation of forest types (\emph{i.e.}, disproportionate
+representation of US Pacific NW for \(B_{root-coarse}\) relative to
+\(B_{root}\); Fig. 4) than of methodological accuracy. The overall high
+degree of closure implies that \emph{ForC} gives a consistent picture of
+C cycling within biomes. This is an important and useful test, because
+it allows for consistency checks within the C cycle, for example
+leveraging separate and independently-measured fluxes to constrain
+errors in another (Phillips \emph{et al} 2017, Williams \emph{et al}
+2014, Harmon \emph{et al} 2011), or producing internally consistent
+global data products (Wang \emph{et al} 2018).
 
@@
 
-Our analysis reveals a general acceleration of carbon cycling from the
-tropics to the high latitudes. For mature forests, this is consistent
-with a large body of previous work demonstrating that C fluxes generally
-decline with latitude (e.g., Banbury Morgan \emph{et al} n.d.). For
-regrowth forests, more rapid accumulation of biomass at lower latitudes
-has been well-established (Anderson \emph{et al} 2006, Cook-Patton
-\emph{et al} 2020), whereas this is the first study to compare age
-trends in deadwood and organic layer across biomes (but see Cook-Patton
-\emph{et al} 2020). For most C flux variables, this analysis is the
-first to examine flux trends in regrowth forests across biomes (i.e.,
-age x biome interaction). Data remain sparse, but for better-represented
-variables, we often see faster acceleration of C cycling in the warmer
-climates. Further work will be required to explore age x climate
-interactions, but our broad-brush overview indicates that C cycling of
-regrowth forests is not only higher in the tropics, parallel to fluxes
-in mature forests (Banbury Morgan \emph{et al} n.d.), but also that it
-accelerates more rapidly with stand age in the tropics, consistent with
-more rapid biomass accumulation.
-
-In contrast to C fluxes and accumulation rates in regrowth forests,
-stocks\ldots{}
-
-Thus, biome differences should be interpreted more as driven more by
-geographic distribution of sampling than by true differences.
-
-Higher NEP in temperate forests -- implications? Invariant NEP in older
-forests? This could be built out a bit; thinking of Luyssaert 2008
-\url{http://dx.doi.org/10.1038/nature07276} and following papers arguing
-about this.
+Our analysis reveals that carbon cycling is most rapid in the tropics
+and slowest in boreal regions, including C fluxes into (\(GPP\)), within
+(e.g., \(NPP\) and its components), and out of (e.g., \(R_{soil}\),
+\(R_{eco}\)) the ecosystem. For mature forests, this is consistent with
+a large body of previous work demonstrating that C fluxes generally
+decline with latitude (or increase with temperature) on a global scale
+(e.g., Luyssaert \emph{et al} 2007, Gillman \emph{et al} 2015, Li and
+Xiao 2019, Banbury Morgan \emph{et al} n.d.). The consistency with which
+this occurs across numerous fluxes is not surprising, but has never been
+simultaneously assessed across such a large number of variables (but see
+Banbury Morgan \emph{et al} n.d. for nine autotrophic fluxes).
+
+The notable exception to the pattern of fluxes decreasing from tropical
+to boreal regions is \(NEP\) (Fig. 6f), which showed no significant
+differences across biomes. Unlike the other C flux variables, \(NEP\)
+does not characterize the rate at which C cycles through the ecosystem,
+but is the balance between C sequestration (\(GPP\)) and respiratory
+losses (\(R_{eco}\)) and represents net CO\textsubscript{2}
+sequestration (or release) by the ecosystem. \(NEP\) tends to be
+relatively small in mature forest stands (discussed further below),
+which accumulate carbon slowly relative to younger stands, if at all
+(Luyssaert \emph{et al} 2008, Amiro \emph{et al} 2010, Besnard \emph{et
+al} 2018). It is therefore unsurprising that there are no pronounced
+differences across biomes, suggesting that variation in \(NEP\) of
+mature forests is controlled less by climate and more by other factors
+including moderate disturbances (Curtis and Gough 2018) or
+disequilibrium of \(R_{soil}\) relative to C inputs (e.g., in peatlands
+where anoxic conditions inhibit decomposition; Wilson \emph{et al}
+2016).
+
+In contrast to the patterns observed for \(NEP\) in mature stands,
+\(NEP\) of stands between 20 and 100 years of age varied across biomes,
+being lowest in boreal forests, intermediate in temperate broadleaf
+forests, and highest in temperate conifer forests (with insufficient
+data to assess tropical forests; Figs. 6f, S1). This is consistent with
+findings that live biomass accumulation rates (\emph{e.g.},
+\(\Delta B_{ag}\) or \(\Delta B_{tot}\)) during early secondary
+succession decrease with latitude (Figs. 7a, S16-S22; Anderson \emph{et
+al} 2006, Cook-Patton \emph{et al} 2020). Note, though, that \(NEP\)
+includes not only \(\Delta B_{tot}\), but also changes in \(DW_{tot}\),
+\(OL\), and soil carbon, and biome differences in the accumulation rates
+of these variables have not been detected, in part because these
+variables do not consistently increase with stand age (Figs. 7, S23-S26,
+and see discussion below; Cook-Patton \emph{et al} 2020).
+
+For regrowth forests, little is known about cross-biome differences in
+carbon fluxes, and we are not aware of any previous large-scale
+comparisons of C fluxes that have been limited to regrowth forests.
+Thus, this analysis was the first to examine flux trends in regrowth
+forests across biomes. The observed tendency for young forest fluxes to
+decrease from tropical to boreal regions paralleled patterns in mature
+forests (Figs. 6, S1-S15), suggesting that regrowth forests follow
+latitudinal trends in carbon cycling similar to those of mature forests
+(e.g., Banbury Morgan \emph{et al} n.d.).
+
+In contrast to C fluxes and biomass accumulation rates in regrowth
+forests, stocks showed less systematic variation across biomes. For
+aboveground biomass, which is the variable in \emph{ForC} with broadest
+geographical representation, the modest trend of declining biomass from
+tropical to boreal regions mirrors observations from spaceborne lidar
+that reveal a decline in aboveground biomass (for all forests, including
+secondary) with latitude across the N hemisphere (Hu \emph{et al} 2016).
+The highest- biomass forests on Earth are, however, found in coastal
+temperate climates of both the southern and northern hemisphere (Keith
+\emph{et al} 2009, Smithwick \emph{et al} 2002, Hu \emph{et al} 2016).
+Disproportionate representation of forests in one such region--the US
+Pacific Northwest--inflated estimates of temperate conifer fluxes and
+stocks for some variables and was responsible for all of the anomalous
+results described here (e.g., lack of complete C budget closure,
+anomalous trend across biomes for \(BNPP_{coarse}\)). Thus, biome
+differences should always be interpreted relative to the geographic
+distribution of sampling, which only rarely covers the majority of
+forested area within a biome.
+
+Whereas biomass can be remotely sensed and receives significant research
+attention, far less is known about geographical variation in deadwood
+and organic layer (\(OL\)) across biomes, which has proved a limitation
+for C accounting efforts (Pan \emph{et al} 2011). Although these stocks
+can be important--exceeding 100 Mg C ha\textsuperscript{-1} in some
+stands (Figs. 7d-e, S23-S25), this study is the first to synthesize
+deadwood data on a global scale (but see Cook-Patton \emph{et al} 2020
+for young forests). Unfortunately, data remain too sparse for
+statistical comparison across biomes (Figs. 7, S23-S25; but see below
+for age trends), pointing to a need for more widespread quantification
+of both standing and downed deadwood. \emph{ForC} coverage of \(OL\)
+stocks is more comprehensive, revealing no significant differences
+across temperate and tropical biomes, but a tendency towards higher
+\(OL\) in boreal forests, consistent with the idea that slower
+decomposition in colder climates results in more buildup of organic
+matter (Allen \emph{et al} 2002). Further research on non-living C
+stocks in the world's forests will be essential to completing the
+picture.
 
@@
 
-\emph{(Just some rough notes at this point)}
-
-A relative dearth of data on C cycling in secondary forests,
-particularly in the tropics (Anderson-Teixeira et al 2016), is
-problematic in that almost 2/3 of the world's forests were secondary as
-of 2010 (FAO 2010), implying an under-filled need to characterize
-age-related trends in forest C cycling.
-
-Moreover, as disturbances increase ({\textbf{???}}, McDowell \emph{et
-al} 2020), understanding the carbon dynamics of regrowth forests will be
-increasingly important.
-
-It's also important to understand secondary forest C sequestration to
-reduce uncertainty regarding the potential for carbon uptake by regrowth
-forests ({\textbf{???}}, Cook-Patton \emph{et al} 2020).
-
-NEP increases with log(age) to 100 --\textgreater{} strongest C sinks
-are established secondary forests. (But presumably this exact number is
-an artifact; don't over-emphasize.)
+Our study reveals that most C fluxes quickly increase to a plateau as
+stands age (Fig. 6), consistent with current understanding of age trends
+in forest C cycling (e.g., Anderson‐Teixeira \emph{et al} 2013, Amiro
+\emph{et al} 2010, Magnani \emph{et al} 2007). While limited records in
+very young (\emph{i.e.}, \textless5 year old) stands resulted in poor
+resolution of the earliest phases of this increase for many variables
+(sometimes detecting no age trend; Table 1), any autotrophic C flux
+(e.g., \(GPP\), \(NPP\) and its components, \(R_auto\)) would be minimal
+immediately following a stand-clearing disturbance. These would be
+expected to increase rapidly with the most metabolically active
+components of biomass, foliage and fine roots, which also increase
+rapidly with stand age (Fig. 7). In contrast, soil heterotrophic
+respiration (\(R_{het-soil}\)) and total soil respiration (\(R_{soil}\))
+are expected to be non-zero following stand-clearing disturbance,
+although these may decrease with a reduction of root respiration
+(\(R_{soil}\) only) and C exudates or increase in response to an influx
+of dead roots and litter (Ribeiro-Kumara \emph{et al} 2020, Maurer
+\emph{et al} 2016, Bond-Lamberty \emph{et al} 2004). In this study, we
+detect no significant age trends in either variable.
+
+Notably, net carbon sequestration (\(NEP\)) increases with age up to the
+100-yr threshold examined here, with more pronounced patterns in
+temperate than boreal forests (Fig. 6f). This finding is largely
+consistent with, but built from a far larger dataset than, previous
+studies showing an increase in \(NEP\) across relatively young stand
+ages (Pregitzer and Euskirchen 2004, Baldocchi \emph{et al} 2001,
+Luyssaert \emph{et al} 2008). However, \(NEP\) has been observed to
+decline from intermediate to old stands (Luyssaert \emph{et al} 2008),
+and the \(NEP\) estimated by our model for 100-year-old temperate
+conifer stands (\textasciitilde5 Mg C ha\textsuperscript{-1}
+yr\textsuperscript{-1}) exceeds the mean of mature forests in the same
+biome (0.7 Mg C ha\textsuperscript{-1} yr\textsuperscript{-1}; Fig.4). A
+decrease in \(NEP\) is consistent with the observed deceleration of
+biomass accumulation as stands age, although both biomass and non-living
+C stocks will often continue to increase well beyond the 100-yr
+threshold used here to delimit young and mature stands (Luyssaert
+\emph{et al} 2008, McGarvey \emph{et al} 2014, Lichstein \emph{et al}
+2009).
+
+In terms of stocks, our study reveals consistent increases in live
+biomass stocks with stand age--a pattern that is well-known and expected
+(e.g., Lichstein \emph{et al} 2009, Yang \emph{et al} 2011)--and more
+variable age trends in deadwood and \(OL\). The latter are particularly
+sensitive to the type of disturbance, where disturbances that remove
+most organic material (e.g., logging, agriculture) result in negligible
+deadwood in young stands, followed by a buildup over time (tropical
+stands in Fig. 7e; e.g., Vargas \emph{et al} 2008). In contrast, natural
+disturbances (e.g., fire, drought) can produce large amounts of deadwood
+(mostly \(DW_{standing}\)) that slowly decomposes as the stand recovers,
+resulting in declines across young stand ages (e.g., temperate and
+boreal stands in Fig. 7e; e.g., Carmona \emph{et al} 2002). Again,
+further study and synthesis of non-living C stocks across biomes and
+stand ages will be valuable to giving a more comprehensive picture.
 
@@
 
-The future of forest C cycling ({\textbf{???}}) will shape trends in
-atmospheric CO\textsubscript{2} and the course of climate change. For a
-human society seeking to understand and mitigate climate change, the
-data contained in ForC and summarized here can help to meet two major
-challenges.
+The future of forest C cycling (Song \emph{et al} 2019) will shape
+trends in atmospheric CO\textsubscript{2} and the course of climate
+change (Schimel \emph{et al} 2015). Our findings, and more generally the
+data contained in \emph{ForC} and summarized here, can help to meet two
+major challenges.
 
@@
 essential to improving predictions of the future course of climate
-change. To ensure that models are giving the right answers for the right
-reasons, it is important to benchmark against multiple components of the
-C cycle that are internally consistent with each other. By making tens
-of thousands of records readily available in standardized format, ForC
-makes it feasible for the modeling community to draw upon these data to
-benchmark models. Integration of ForC with models is a goal (Fer et al.,
-in revision).
-
-Second, ForC can serve as a pipeline through which forest science can
-inform forest-based climate change mitigation efforts. Such efforts will
-be most effective when informed by the best available data, yet it is
-not feasible for the individuals and organizations designing such
-efforts to sort through literature, often behind paywalls, with data
-reported in varying units, terminology, etc. One goal for ForC is to
-serve as a pipeline through which information can flow efficiently from
-forest researchers to decision-makers working to implement forest
-conservation strategies at global, national, or landscape scales. This
-is already happening! ForC has already contributed to updating the IPCC
-guidelines for carbon accounting in forests {[}IPCC 2019; Requena Suarez
-\emph{et al} (2019); Rozendaal et al in prep{]}, mapping C accumulation
+change, for the simple reason that by definition future projections
+extend our existing observations and understanding to conditions that do
+not currently exist on Earth (McDowell \emph{et al} 2018, Bonan and
+Doney 2018, Gustafson \emph{et al} 2018). To ensure that models are
+giving the right answers for the right reasons (Sulman \emph{et al}
+2018), it is important to benchmark against multiple components of the C
+cycle that are internally consistent with each other (Collier \emph{et
+al} 2018, Wang \emph{et al} 2018). \emph{ForC}'s tens of thousands of
+records are readily available in a standardized format, and our analyses
+here indicate that their internal consistency is reasonably high.
+Integration of \emph{ForC} with models will be valuable to improving the
+accuracy and reliability of models.
+
+Second, \emph{ForC} can serve as a pipeline through which information
+can flow efficiently from forest researchers to decision-makers working
+to implement forest conservation strategies at global, national, or
+landscape scales. This is already happening: \emph{ForC} has contributed
+to updating the IPCC guidelines for carbon accounting in forests (IPCC
+2019, Requena Suarez \emph{et al} 2019), mapping C accumulation
 potential from natural forest regrowth globally (Cook-Patton \emph{et
@@
 
-\textbf{ForC can complement remote sensing to provide a comprehensive
-picture of global forest C cycling.} (\emph{here, discuss what is
-well-covered by remote sesnsing, where ForC is useful for calibrating
-remote sensing, and what ForC can get that remote sensing can't.}) AGB
-is the largest stock, and most of the emphasis is on this variable.
-Remote sensing, with calibration based on high-quality ground-based data
-(Schepaschenko \emph{et al} 2019, Chave \emph{et al} 2019), is the best
-approach for mapping forest carbon (REFS). However, it is limited in
-that it is not associated with stand age and disturbance history, except
-in recent decades when satellite data can be used to detect forest loss,
-gain, and some of their dominant drivers (Hansen \emph{et al} 2013, Song
-\emph{et al} 2018, Curtis \emph{et al} 2018). ForC is therefore valuable
-in defining age-based trajectories in biomass, as in Cook-Patton
-\emph{et al} (2020).
-
-\emph{Remote sensing measurements are increasingly useful for global- or
-regional-scale estimates of forest \(GPP\) (Bagdley et al.~2019, (Li and
-Xiao 2019)), aboveground biomass (\(B_{ag}\)) (REFS), woody mortality
-(}i.e.\emph{, \(B_{ag}\) losses to mortality \(M_{woody}\)) (Clark
-\emph{et al} 2004, Leitold \emph{et al} 2018), and to some extent net
-ecosystem exchange (\(NEP\)) (REFS).} Other variables, in particular
-respiration fluxes, cannot be remotely sensed (({\textbf{???}})), and
-efforts such as the Global Carbon Project (\textbf{le quere REF}) and
-NASA CMS (\textbf{citation:
-\url{https://carbon.nasa.gov/pdfs/CMS_Phase-1_Report_Final_optimized.pdf}
-but maybe better to cite open literature, one of the papers listed at
-\url{https://cmsflux.jpl.nasa.gov/get-data/publication-data-sets}})
-typically compute them as residuals only. (\textbf{Ben, it woudl be
-particularly helpful if you could flesh this out some more.})
-
-\textbf{Move to data availability statement, or methods?}: We recommend
-that use of ForC data go to the original database, as opposed to using
-``off-the-shelf'' values from this publication. This is because (1) ForC
-is constantly being updated, (2) analyses should be designed to match
-the application, (3) age equations presented here all fit a single
-functional form that is not necessarily the best possible for all the
-variables.
+It is also interesting to consider the complementary utility of
+global-scale but spatially discontinuous databases such as ForC and
+remote wall-to-wall remote sensing products. The latter provide
+unparalleled insight into aboveground carbon stocks, but less constraint
+on belowground stocks or carbon fluxes in general (Bond‐Lamberty
+\emph{et al} 2016, Anav \emph{et al} 2015). Combining observational data
+and remote observations may provide a much more comprehensive and
+accurate picture of global forest C cycling, particularly when used in
+formal data assimilation systems (Konings \emph{et al} 2019, Liu
+\emph{et al} 2018). Biomass is the largest C stock in most forests, and
+most of the emphasis has traditionally been on this variable.
+Remote-sensing driven biomass estimates (e.g., Saatchi \emph{et al}
+2011), calibrated based on high-quality ground-based data (Schepaschenko
+\emph{et al} 2019, Chave \emph{et al} 2019), are well suited for this
+task. Note, however, that factors such as stand age and disturbance
+history are difficult, if possible, to detect remotely, and can only be
+characterized for very recent decades (Hansen \emph{et al} 2013, Song
+\emph{et al} 2018, Curtis \emph{et al} 2018). Ground-based data such as
+\emph{ForC} are therefore valuable in defining age-based trajectories in
+biomass, as in Cook-Patton \emph{et al} (2020), and thus constraining
+variables such as carbon sink potential (Luyssaert \emph{et al} 2008).
+
+In contrast, carbon allocation within forest ecosystems and respiration
+fluxes cannot be remotely sensed. Efforts such as the Global Carbon
+Project (Friedlingstein \emph{et al} 2019) and NASA's Carbon Monitoring
+System (Liu \emph{et al} 2018) typically compute respiration as
+residuals of all other terms (Bond‐Lamberty \emph{et al} 2016, Harmon
+\emph{et al} 2011). This means that the errors on respiration outputs
+are likely to be large and certainly poorly constrained, offering a
+unique opportunity for databases such as ForC and SRDB (Jian \emph{et
+al} 2020) to provide observational benchmarks. For example, Konings
+\emph{et al} (2019) produced a unique top-down estimate of global
+heterotrophic respiration that can both be compared with extant
+bottom-up estimates (Bond‐Lamberty 2018) and used as an internal
+consistency check on other parts of the carbon cycle (Phillips \emph{et
+al} 2017).
+
+\hypertarget{conclusions}{%
+\subsubsection{Conclusions}\label{conclusions}}
 
 As climate change accelerates, understanding and managing the carbon
-dynamics of forests will be critical to forecasting, mitigation, and
-adaptation. The C data in ForC, as summarized here, will be valuable to
-these efforts.
+dynamics of forests--notably including dynamics and fluxes that cannot
+be observed by satellites--is critical to forecasting, mitigation, and
+adaptation. The C data in \emph{ForC}, as summarized here, will be
+valuable to these efforts. Notably, the fact that tropical forests tend
+to have both the highest rates of C sequestration in young stands (Fig.
+7; Cook-Patton \emph{et al} 2020), fueled by their generally high C flux
+rates (Table 1; Fig. 6), and the highest mean biomass (Fig. 7; Table 1;
+Hu \emph{et al} 2016, Jian \emph{et al} 2020) reinforces the concept
+that conservation and restoration of these forests is a priority for
+climate change mitigation, along with high-biomass old-growth temperate
+stands (Grassi \emph{et al} 2017, Goldstein \emph{et al} 2020). It is
+also important to note the trade-off in climate mitigation potential of
+restoration of young forests, with high rates of CO\textsubscript{2}
+sequestration (\(NEP\); Cook-Patton \emph{et al} 2020), versus
+conservation and management of mature forests, with low \(NEP\) but high
+C stocks that could not be recovered on a time scale relevant to climate
+change mitigation (Goldstein \emph{et al} 2020). Generally speaking, the
+conservation of mature forests will yield greater climate benefits
+(Anderson-Teixeira and DeLucia 2011), but both approaches are critical
+to avoiding catastrophic climate change (IPCC 2018).
 
@@
 
-All researchers whose data is included in ForC and this analysis. Ian
-McGregor for help with the database. Thanks to Norbert Kunert and
-{[}Helene's intern{]} for helpful input at an earlier phase. A
-Smithsonian Scholarly Studies grant to KAT and HML. WLS grant to KAT.
+Thanks to all researchers whose data are included in \emph{ForC} and
+this analysis, to Jennifer McGarvey and Ian McGregor for help with the
+database, and to Norbert Kunert for helpful discussion. Funding sources
+included a Smithsonian Scholarly Studies grant to KAT and HML and a
+Smithsonian Working Land and Seascapes grant to KAT.
 
@@
 scripts, and results presented here are also available through the
-open-access ForC GitHub repository
+open-access \emph{ForC} GitHub repository
 (\url{https://github.com/forc-db/ForC}), where many will be updated as
 the database develops.
-
-\hypertarget{orcid-id}{%
-\subsection{ORCID iD}\label{orcid-id}}
-
-Kristina J. Anderson-Teixeira:
-\url{https://orcid.org/0000-0001-8461-9713}
 
@@
 \hypertarget{refs}{}
+\begin{cslreferences}
+\leavevmode\hypertarget{ref-allen_global_2002}{}%
+Allen A, Brown J and Gillooly J 2002 Global biodiversity, biochemical
+kinetics, and the energetic-equivalence rule \emph{SCIENCE} \textbf{297}
+1545--8 Online:
+\url{http://linkseeker.lanl.gov/unm?genre=article\&issn=0036-8075\&date=2002\&volume=297\&issue=5586\&spage=1545\&atitle=Global+biodiversity\%2C+biochemical+kinetics\%2C+and+the+energetic\%2Dequivalence+rule\&aulast=Allen\&auinit=AP}
+
+\leavevmode\hypertarget{ref-amiro_ecosystem_2010}{}%
+Amiro B D, Barr A G, Barr J G, Black T A, Bracho R, Brown M, Chen J,
+Clark K L, Davis K J, Desai A R, Dore S, Engel V, Fuentes J D, Goldstein
+A H, Goulden M L, Kolb T E, Lavigne M B, Law B E, Margolis H A, Martin
+T, McCaughey J H, Misson L, Montes‐Helu M, Noormets A, Randerson J T,
+Starr G and Xiao J 2010 Ecosystem carbon dioxide fluxes after
+disturbance in forests of North America \emph{Journal of Geophysical
+Research: Biogeosciences} \textbf{115} Online:
+\url{http://agupubs.onlinelibrary.wiley.com/doi/abs/10.1029/2010JG001390}
+
+\leavevmode\hypertarget{ref-anav_spatiotemporal_2015}{}%
+Anav A, Friedlingstein P, Beer C, Ciais P, Harper A, Jones C,
+Murray‐Tortarolo G, Papale D, Parazoo N C, Peylin P, Piao S, Sitch S,
+Viovy N, Wiltshire A and Zhao M 2015 Spatiotemporal patterns of
+terrestrial gross primary production: A review \emph{Reviews of
+Geophysics} \textbf{53} 785--818 Online:
+\url{http://agupubs.onlinelibrary.wiley.com/doi/abs/10.1002/2015RG000483}
+
+\leavevmode\hypertarget{ref-andela_human-driven_2017}{}%
+Andela N, Morton D C, Giglio L, Chen Y, Werf G R van der, Kasibhatla P
+S, DeFries R S, Collatz G J, Hantson S, Kloster S, Bachelet D, Forrest
+M, Lasslop G, Li F, Mangeon S, Melton J R, Yue C and Randerson J T 2017
+A human-driven decline in global burned area \emph{Science} \textbf{356}
+1356--62 Online:
+\url{http://science.sciencemag.org/content/356/6345/1356}
+
 \leavevmode\hypertarget{ref-anderson_temperature-dependence_2006}{}%
@@
 
+\leavevmode\hypertarget{ref-anderson-teixeira_greenhouse_2011}{}%
+Anderson-Teixeira K J and DeLucia E H 2011 The greenhouse gas value of
+ecosystems \emph{Global Change Biology} \textbf{17} 425--38 Online:
+\url{http://doi.wiley.com/10.1111/j.1365-2486.2010.02220.x}
+
+\leavevmode\hypertarget{ref-andersonteixeira_altered_2013}{}%
+Anderson‐Teixeira K J, Miller A D, Mohan J E, Hudiburg T W, Duval B D
+and DeLucia E H 2013 Altered dynamics of forest recovery under a
+changing climate \emph{Global Change Biology} \textbf{19} 2001--21
+Online: \url{http://onlinelibrary.wiley.com/doi/abs/10.1111/gcb.12194}
+
 \leavevmode\hypertarget{ref-anderson-teixeira_forc_2018}{}%
@@
 \url{http://onlinelibrary.wiley.com/doi/10.1111/gcb.13226/abstract}
+
+\leavevmode\hypertarget{ref-badgley_terrestrial_2019}{}%
+Badgley G, Anderegg L D L, Berry J A and Field C B 2019 Terrestrial
+gross primary production: Using NIRV to scale from site to globe
+\emph{Global Change Biology} \textbf{25} 3731--40 Online:
+\url{http://onlinelibrary.wiley.com/doi/abs/10.1111/gcb.14729}
 
@@
 
+\leavevmode\hypertarget{ref-bates_fitting_2015}{}%
+Bates D, Mächler M, Bolker B and Walker S 2015 Fitting Linear
+Mixed-Effects Models Using \textbf{lme4} \emph{Journal of Statistical
+Software} \textbf{67} Online: \url{http://www.jstatsoft.org/v67/i01/}
+
+\leavevmode\hypertarget{ref-besnard_quantifying_2018}{}%
+Besnard S, Carvalhais N, Arain M A, Black A, Bruin S de, Buchmann N,
+Cescatti A, Chen J, Clevers J G P W, Desai A R, Gough C M, Havrankova K,
+Herold M, Hörtnagl L, Jung M, Knohl A, Kruijt B, Krupkova L, Law B E,
+Lindroth A, Noormets A, Roupsard O, Steinbrecher R, Varlagin A, Vincke C
+and Reichstein M 2018 Quantifying the effect of forest age in annual net
+forest carbon balance \emph{Environmental Research Letters} \textbf{13}
+124018 Online:
+\url{https://iopscience.iop.org/article/10.1088/1748-9326/aaeaeb}
+
 \leavevmode\hypertarget{ref-bonan_forests_2008}{}%
@@
 
+\leavevmode\hypertarget{ref-bonan_climate_2018}{}%
+Bonan G B and Doney S C 2018 Climate, ecosystems, and planetary futures:
+The challenge to predict life in Earth system models \emph{Science}
+\textbf{359} Online:
+\url{http://science.sciencemag.org/content/359/6375/eaam8328}
+
+\leavevmode\hypertarget{ref-bonan_model_2019}{}%
+Bonan G B, Lombardozzi D L, Wieder W R, Oleson K W, Lawrence D M,
+Hoffman F M and Collier N 2019 Model Structure and Climate Data
+Uncertainty in Historical Simulations of the Terrestrial Carbon Cycle
+(1850--2014) \emph{Global Biogeochemical Cycles} \textbf{33} 1310--26
+Online:
+\url{https://onlinelibrary.wiley.com/doi/abs/10.1029/2019GB006175}
+
+\leavevmode\hypertarget{ref-bondlamberty_new_2018}{}%
+Bond‐Lamberty B 2018 New Techniques and Data for Understanding the
+Global Soil Respiration Flux \emph{Earth's Future} \textbf{6} 1176--80
+Online:
+\url{http://agupubs.onlinelibrary.wiley.com/doi/abs/10.1029/2018EF000866}
+
+\leavevmode\hypertarget{ref-bondlamberty_estimating_2016}{}%
+Bond‐Lamberty B, Epron D, Harden J, Harmon M E, Hoffman F, Kumar J,
+David McGuire A and Vargas R 2016 Estimating heterotrophic respiration
+at large scales: Challenges, approaches, and next steps \emph{Ecosphere}
+\textbf{7} Online:
+\url{https://onlinelibrary.wiley.com/doi/abs/10.1002/ecs2.1380}
+
 \leavevmode\hypertarget{ref-bond-lamberty_global_2010}{}%
@@
 \url{http://www.biogeosciences.net/7/1915/2010/}
+
+\leavevmode\hypertarget{ref-bond-lamberty_contribution_2004}{}%
+Bond-Lamberty B, Wang C and Gower S T 2004 Contribution of root
+respiration to soil surface CO 2 flux in a boreal black spruce
+chronosequence \emph{Tree Physiology} \textbf{24} 1387--95 Online:
+\url{http://academic.oup.com/treephys/article/24/12/1387/1612153}
+
+\leavevmode\hypertarget{ref-carmona_coarse_2002}{}%
+Carmona M R, Armesto J J, Aravena J C and Pérez C A 2002 Coarse woody
+debris biomass in successional and primary temperate forests in Chiloé
+Island, Chile \emph{Forest Ecology and Management} \textbf{164} 265--75
+Online:
+\url{http://www.sciencedirect.com/science/article/pii/S0378112701006028}
+
+\leavevmode\hypertarget{ref-cavaleri_urgent_2015}{}%
+Cavaleri M A, Reed S C, Smith W K and Wood T E 2015 Urgent need for
+warming experiments in tropical forests \emph{Global Change Biology}
+\textbf{21} 2111--21 Online:
+\url{http://onlinelibrary.wiley.com/doi/10.1111/gcb.12860/abstract}
 
@@
 
-\leavevmode\hypertarget{ref-clark_quantifying_2004}{}%
-Clark D B, Castro C S, Alvarado L D A and Read J M 2004 Quantifying
-mortality of tropical rain forest trees using high-spatial-resolution
-satellite data \emph{Ecology Letters} \textbf{7} 52--9 Online:
-\url{https://onlinelibrary.wiley.com/doi/abs/10.1046/j.1461-0248.2003.00547.x}
+\leavevmode\hypertarget{ref-collier_international_2018}{}%
+Collier N, Hoffman F M, Lawrence D M, Keppel‐Aleks G, Koven C D, Riley W
+J, Mu M and Randerson J T 2018 The International Land Model Benchmarking
+(ILAMB) System: Design, Theory, and Implementation \emph{Journal of
+Advances in Modeling Earth Systems} \textbf{10} 2731--54 Online:
+\url{http://agupubs.onlinelibrary.wiley.com/doi/abs/10.1029/2018MS001354}
 
@@
 \url{http://science.sciencemag.org/content/361/6407/1108}
+
+\leavevmode\hypertarget{ref-curtis_forest_2018}{}%
+Curtis P S and Gough C M 2018 Forest aging, disturbance and the carbon
+cycle \emph{New Phytologist} \textbf{219} 1188--93 Online:
+\url{http://nph.onlinelibrary.wiley.com/doi/abs/10.1111/nph.15227}
+
+\leavevmode\hypertarget{ref-di_vittorio_initial_2020}{}%
+Di Vittorio A V, Shi X, Bond‐Lamberty B, Calvin K and Jones A 2020
+Initial Land Use/Cover Distribution Substantially Affects Global Carbon
+and Local Temperature Projections in the Integrated Earth System Model
+\emph{Global Biogeochemical Cycles} \textbf{34} Online:
+\url{https://onlinelibrary.wiley.com/doi/abs/10.1029/2019GB006383}
+
+\leavevmode\hypertarget{ref-fao_global_2010}{}%
+FAO 2010 \emph{Global Forest Resources Assessment 2010} (Rome, Italy:
+Food; Agriculture Organization of the United Nations)
+
+\leavevmode\hypertarget{ref-friedlingstein_climatecarbon_2006}{}%
+Friedlingstein P, Cox P, Betts R, Bopp L, Bloh W von, Brovkin V, Cadule
+P, Doney S, Eby M, Fung I, Bala G, John J, Jones C, Joos F, Kato T,
+Kawamiya M, Knorr W, Lindsay K, Matthews H D, Raddatz T, Rayner P, Reick
+C, Roeckner E, Schnitzler K-G, Schnur R, Strassmann K, Weaver A J,
+Yoshikawa C and Zeng N 2006 Climate--Carbon Cycle Feedback Analysis:
+Results from the C4MIP Model Intercomparison \emph{Journal of Climate}
+\textbf{19} 3337--53 Online:
+\url{https://journals.ametsoc.org/doi/abs/10.1175/JCLI3800.1}
+
+\leavevmode\hypertarget{ref-friedlingstein_global_2019}{}%
+Friedlingstein P, Jones M W, O'Sullivan M, Andrew R M, Hauck J, Peters G
+P, Peters W, Pongratz J, Sitch S, Le Quéré C, Bakker D C E, Canadell J
+G, Ciais P, Jackson R B, Anthoni P, Barbero L, Bastos A, Bastrikov V,
+Becker M, Bopp L, Buitenhuis E, Chandra N, Chevallier F, Chini L P,
+Currie K I, Feely R A, Gehlen M, Gilfillan D, Gkritzalis T, Goll D S,
+Gruber N, Gutekunst S, Harris I, Haverd V, Houghton R A, Hurtt G, Ilyina
+T, Jain A K, Joetzjer E, Kaplan J O, Kato E, Klein Goldewijk K,
+Korsbakken J I, Landschützer P, Lauvset S K, Lefèvre N, Lenton A,
+Lienert S, Lombardozzi D, Marland G, McGuire P C, Melton J R, Metzl N,
+Munro D R, Nabel J E M S, Nakaoka S-I, Neill C, Omar A M, Ono T, Peregon
+A, Pierrot D, Poulter B, Rehder G, Resplandy L, Robertson E, Rödenbeck
+C, Séférian R, Schwinger J, Smith N, Tans P P, Tian H, Tilbrook B,
+Tubiello F N, Werf G R van der, Wiltshire A J and Zaehle S 2019 Global
+Carbon Budget 2019 \emph{Earth System Science Data} \textbf{11}
+1783--838 Online:
+\url{https://essd.copernicus.org/articles/11/1783/2019/}
+
+\leavevmode\hypertarget{ref-gillman_latitude_2015}{}%
+Gillman L N, Wright S D, Cusens J, McBride P D, Malhi Y and Whittaker R
+J 2015 Latitude, productivity and species richness \emph{Global Ecology
+and Biogeography} \textbf{24} 107--17 Online:
+\url{https://onlinelibrary.wiley.com/doi/abs/10.1111/geb.12245}
 
@@
 
+\leavevmode\hypertarget{ref-gustafson_extrapolating_2018}{}%
+Gustafson E J, Kubiske M E, Miranda B R, Hoshika Y and Paoletti E 2018
+Extrapolating plot-scale CO2 and ozone enrichment experimental results
+to novel conditions and scales using mechanistic modeling
+\emph{Ecological Processes} \textbf{7} 31 Online:
+\url{https://doi.org/10.1186/s13717-018-0142-8}
+
 \leavevmode\hypertarget{ref-hansen_high-resolution_2013}{}%
@@
 \url{http://www.sciencemag.org/cgi/doi/10.1126/science.1244693}
+
+\leavevmode\hypertarget{ref-harmon_heterotrophic_2011}{}%
+Harmon M E, Bond-Lamberty B, Tang J and Vargas R 2011 Heterotrophic
+respiration in disturbed forests: A review with examples from North
+America \emph{Journal of Geophysical Research} \textbf{116} Online:
+\url{http://doi.wiley.com/10.1029/2010JG001495}
+
+\leavevmode\hypertarget{ref-houghton_terrestrial_2020}{}%
+Houghton R A 2020 Terrestrial fluxes of carbon in GCP carbon budgets
+\emph{Global Change Biology} \textbf{26} 3006--14 Online:
+\url{https://onlinelibrary.wiley.com/doi/abs/10.1111/gcb.15050}
+
+\leavevmode\hypertarget{ref-hu_mapping_2016}{}%
+Hu T, Su Y, Xue B, Liu J, Zhao X, Fang J and Guo Q 2016 Mapping Global
+Forest Aboveground Biomass with Spaceborne LiDAR, Optical Imagery, and
+Forest Inventory Data \emph{Remote Sensing} \textbf{8} 565 Online:
+\url{http://www.mdpi.com/2072-4292/8/7/565}
+
+\leavevmode\hypertarget{ref-ipcc_2019_2019}{}%
+IPCC 2019 \emph{2019 Refinement to the 2006 IPCC Guidelines for National
+Greenhouse Gas Inventories} Online:
+\url{https://www.ipcc.ch/report/2019-refinement-to-the-2006-ipcc-guidelines-for-national-greenhouse-gas-inventories/}
+
+\leavevmode\hypertarget{ref-ipcc_global_2018}{}%
+IPCC 2018 \emph{Global Warming of 1.5°C. An IPCC Special Report on the
+impacts of global warming of 1.5°C above pre-industrial levels and
+related global greenhouse gas emission pathways, in the context of
+strengthening the global response to the threat of climate change,
+sustainable development, and efforts to eradicate poverty
+{[}Masson-Delmotte, V., P. Zhai, H.-O. Pörtner, D. Roberts, J. Skea,
+P.R. Shukla, A. Pirani, W. Moufouma-Okia, C. Péan, R. Pidcock, S.
+Connors, J.B.R. Matthews, Y. Chen, X. Zhou, M.I. Gomis, E. Lonnoy, T.
+Maycock, M. Tignor, and T. Waterfield (eds.){]}.} Online:
+\url{https://www.ipcc.ch/sr15/}
+
+\leavevmode\hypertarget{ref-jian_restructured_2020}{}%
+Jian J, Vargas R, Anderson-Teixeira K, Stell E, Herrmann V, Horn M,
+Kholod N, Manzon J, Marchesi R, Paredes D and Bond-Lamberty B 2020
+\emph{A restructured and updated global soil respiration database
+(SRDB-V5)} (Data, Algorithms,; Models) Online:
+\url{https://essd.copernicus.org/preprints/essd-2020-136/}
 
@@
 
+\leavevmode\hypertarget{ref-jung_exploiting_2006}{}%
+Jung M, Henkel K, Herold M and Churkina G 2006 Exploiting synergies of
+global land cover products for carbon cycle modeling \emph{Remote
+Sensing of Environment} \textbf{101} 534--53 Online:
+\url{http://www.sciencedirect.com/science/article/pii/S0034425706000514}
+
+\leavevmode\hypertarget{ref-keith_re-evaluation_2009}{}%
+Keith H, Mackey B G and Lindenmayer D B 2009 Re-evaluation of forest
+biomass carbon stocks and lessons from the world's most carbon-dense
+forests \emph{Proceedings of the National Academy of Sciences}
+\textbf{106} 11635--40 Online:
+\url{http://www.pnas.org/content/106/28/11635.abstract}
+
+\leavevmode\hypertarget{ref-konings_global_2019}{}%
+Konings A G, Bloom A A, Liu J, Parazoo N C, Schimel D S and Bowman K W
+2019 Global satellite-driven estimates of heterotrophic respiration
+\emph{Biogeosciences} \textbf{16} 2269--84 Online:
+\url{https://bg.copernicus.org/articles/16/2269/2019/}
+
+\leavevmode\hypertarget{ref-kochy_global_2015}{}%
+Köchy M, Hiederer R and Freibauer A 2015 Global distribution of soil
+organic carbon -- Part 1: Masses and frequency distributions of SOC
+stocks for the tropics, permafrost regions, wetlands, and the world
+\emph{SOIL} \textbf{1} 351--65 Online:
+\url{https://soil.copernicus.org/articles/1/351/2015/}
+
 \leavevmode\hypertarget{ref-krause_large_2018}{}%
@@
 
-\leavevmode\hypertarget{ref-leitold_nino_2018}{}%
-Leitold V, Morton D C, Longo M, dos-Santos M N, Keller M and Scaranello
-M 2018 El Niño drought increased canopy turnover in Amazon forests
-\emph{New Phytologist} \textbf{219} 959--71 Online:
-\url{http://doi.wiley.com/10.1111/nph.15110}
-
 \leavevmode\hypertarget{ref-li_mapping_2019}{}%
@@
 2563 Online: \url{https://www.mdpi.com/2072-4292/11/21/2563}
+
+\leavevmode\hypertarget{ref-lichstein_biomass_2009}{}%
+Lichstein J W, Wirth C, Horn H S and Pacala S W 2009 Biomass
+Chronosequences of United States Forests: Implications for Carbon
+Storage and Forest Management \emph{Old-Growth Forests} Ecological
+Studies ed C Wirth, G Gleixner and M Heimann (Springer Berlin
+Heidelberg) pp 301--41 Online:
+\url{http://link.springer.com/chapter/10.1007/978-3-540-92706-8_14}
+
+\leavevmode\hypertarget{ref-liu_detecting_2018}{}%
+Liu J, Bowman K, Parazoo N C, Bloom A A, Wunch D, Jiang Z, Gurney K R
+and Schimel D 2018 Detecting drought impact on terrestrial biosphere
+carbon fluxes over contiguous US with satellite observations
+\emph{Environmental Research Letters} \textbf{13} 095003 Online:
+\url{https://doi.org/10.1088\%2F1748-9326\%2Faad5ef}
 
@@
 
+\leavevmode\hypertarget{ref-luyssaert_old-growth_2008}{}%
+Luyssaert S, Schulze E D, Borner A, Knohl A, Hessenmoller D, Law B E,
+Ciais P and Grace J 2008 Old-growth forests as global carbon sinks
+\emph{Nature} \textbf{455} 213 Online:
+\url{http://dx.doi.org/10.1038/nature07276\%20http://www.nature.com/nature/journal/v455/n7210/suppinfo/nature07276_S1.html}
+
+\leavevmode\hypertarget{ref-magnani_human_2007}{}%
+Magnani F, Mencuccini M, Borghetti M, Berbigier P, Berninger F, Delzon
+S, Grelle A, Hari P, Jarvis P G, Kolari P, Kowalski A S, Lankreijer H,
+Law B E, Lindroth A, Loustau D, Manca G, Moncrieff J B, Rayment M,
+Tedeschi V, Valentini R and Grace J 2007 The human footprint in the
+carbon cycle of temperate and boreal forests \emph{Nature} \textbf{447}
+849--51 Online: \url{http://www.nature.com/articles/nature05847}
+
+\leavevmode\hypertarget{ref-maurer_carbon_2016}{}%
+Maurer G E, Chan A M, Trahan N A, Moore D J P and Bowling D R 2016
+Carbon isotopic composition of forest soil respiration in the decade
+following bark beetle and stem girdling disturbances in the Rocky
+Mountains \emph{Plant, Cell \& Environment} \textbf{39} 1513--23 Online:
+\url{http://onlinelibrary.wiley.com/doi/abs/10.1111/pce.12716}
+
 \leavevmode\hypertarget{ref-mcdowell_pervasive_2020}{}%
@@
 
+\leavevmode\hypertarget{ref-mcdowell_predicting_2018}{}%
+McDowell N G, Michaletz S T, Bennett K E, Solander K C, Xu C, Maxwell R
+M and Middleton R S 2018 Predicting Chronic Climate-Driven Disturbances
+and Their Mitigation \emph{Trends in Ecology \& Evolution} \textbf{33}
+15--27 Online:
+\url{https://www.cell.com/trends/ecology-evolution/abstract/S0169-5347(17)30261-6}
+
+\leavevmode\hypertarget{ref-mcgarvey_carbon_2014}{}%
+McGarvey J C, Thompson J R, Epstein H E and Shugart H H 2014 Carbon
+storage in old-growth forests of the Mid-Atlantic: Toward better
+understanding the eastern forest carbon sink \emph{Ecology} \textbf{96}
+311--7 Online:
+\url{http://www.esajournals.org/doi/abs/10.1890/14-1154.1}
+
+\leavevmode\hypertarget{ref-novick_ameriflux_2018}{}%
+Novick K A, Biederman J A, Desai A R, Litvak M E, Moore D J P, Scott R L
+and Torn M S 2018 The AmeriFlux network: A coalition of the willing
+\emph{Agricultural and Forest Meteorology} \textbf{249} 444--56 Online:
+\url{http://www.sciencedirect.com/science/article/pii/S0168192317303295}
+
 \leavevmode\hypertarget{ref-pan_large_2011}{}%
@@
 
+\leavevmode\hypertarget{ref-phillips_value_2017}{}%
+Phillips C L, Bond-Lamberty B, Desai A R, Lavoie M, Risk D, Tang J,
+Todd-Brown K and Vargas R 2017 The value of soil respiration
+measurements for interpreting and modeling terrestrial carbon cycling
+\emph{Plant and Soil} \textbf{413} 1--25 Online:
+\url{http://link.springer.com/10.1007/s11104-016-3084-x}
+
+\leavevmode\hypertarget{ref-pregitzer_carbon_2004}{}%
+Pregitzer K S and Euskirchen E S 2004 Carbon cycling and storage in
+world forests: Biome patterns related to forest age \emph{Global Change
+Biology} \textbf{10} 2052--77 Online:
+\url{http://dx.doi.org/10.1111/j.1365-2486.2004.00866.x}
+
 \leavevmode\hypertarget{ref-pugh_role_2019}{}%
@@
 \url{http://www.pnas.org/lookup/doi/10.1073/pnas.1810512116}
+
+\leavevmode\hypertarget{ref-r_core_team_r_2020}{}%
+R Core Team 2020 R: A language and environment for statistical
+computing. R Foundation for Statistical Computing, Vienna, Austria. URL
+http://www.R-project.org/.
 
@@
 Online: \url{http://onlinelibrary.wiley.com/doi/abs/10.1111/gcb.14767}
+
+\leavevmode\hypertarget{ref-ribeiro-kumara_how_2020}{}%
+Ribeiro-Kumara C, Köster E, Aaltonen H and Köster K 2020 How do forest
+fires affect soil greenhouse gas emissions in upland boreal forests? A
+review \emph{Environmental Research} \textbf{184} 109328 Online:
+\url{http://www.sciencedirect.com/science/article/pii/S0013935120302218}
+
+\leavevmode\hypertarget{ref-saatchi_benchmark_2011}{}%
+Saatchi S S, Harris N L, Brown S, Lefsky M, Mitchard E T A, Salas W,
+Zutta B R, Buermann W, Lewis S L, Hagen S, Petrova S, White L, Silman M
+and Morel A 2011 Benchmark map of forest carbon stocks in tropical
+regions across three continents \emph{Proceedings of the National
+Academy of Sciences} \textbf{108} 9899--904 Online:
+\url{http://www.pnas.org/content/108/24/9899}
 
@@
 
-\leavevmode\hypertarget{ref-schimel_observing_2015}{}%
-Schimel D, Pavlick R, Fisher J B, Asner G P, Saatchi S, Townsend P,
-Miller C, Frankenberg C, Hibbard K and Cox P 2015 Observing terrestrial
-ecosystems and the carbon cycle from space \emph{Global Change Biology}
-\textbf{21} 1762--76 Online:
-\url{http://doi.wiley.com/10.1111/gcb.12822}
+\leavevmode\hypertarget{ref-schimel_neon_2007}{}%
+Schimel D, Hargrove W, Hoffman F and MacMahon J 2007 NEON: A
+hierarchically designed national ecological network \emph{Frontiers in
+Ecology and the Environment} \textbf{5} 59--9 Online:
+\url{http://esajournals.onlinelibrary.wiley.com/doi/abs/10.1890/1540-9295\%282007\%295\%5B59\%3ANAHDNE\%5D2.0.CO\%3B2}
+
+\leavevmode\hypertarget{ref-schimel_effect_2015}{}%
+Schimel D, Stephens B B and Fisher J B 2015 Effect of increasing CO
+\(_{\textrm{2}}\) on the terrestrial carbon cycle \emph{Proceedings of
+the National Academy of Sciences} \textbf{112} 436--41 Online:
+\url{http://www.pnas.org/lookup/doi/10.1073/pnas.1407302112}
+
+\leavevmode\hypertarget{ref-smithwick_potential_2002}{}%
+Smithwick E A H, Harmon M E, Remillard S M, Acker S A and Franklin J F
+2002 Potential upper bounds of carbon stores in forests of the Pacific
+Northwest \emph{Ecological Applications} \textbf{12} 1303--17 Online:
+\url{http://doi.wiley.com/10.1890/1051-0761(2002)012\%5B1303:PUBOCS\%5D2.0.CO;2}
+
+\leavevmode\hypertarget{ref-song_meta-analysis_2019}{}%
+Song J, Wan S, Piao S, Knapp A K, Classen A T, Vicca S, Ciais P,
+Hovenden M J, Leuzinger S, Beier C, Kardol P, Xia J, Liu Q, Ru J, Zhou
+Z, Luo Y, Guo D, Adam Langley J, Zscheischler J, Dukes J S, Tang J, Chen
+J, Hofmockel K S, Kueppers L M, Rustad L, Liu L, Smith M D, Templer P H,
+Quinn Thomas R, Norby R J, Phillips R P, Niu S, Fatichi S, Wang Y, Shao
+P, Han H, Wang D, Lei L, Wang J, Li X, Zhang Q, Li X, Su F, Liu B, Yang
+F, Ma G, Li G, Liu Y, Liu Y, Yang Z, Zhang K, Miao Y, Hu M, Yan C, Zhang
+A, Zhong M, Hui Y, Li Y and Zheng M 2019 A meta-analysis of 1,119
+manipulative experiments on terrestrial carbon-cycling responses to
+global change \emph{Nature Ecology \& Evolution} \textbf{3} 1309--20
+Online: \url{http://www.nature.com/articles/s41559-019-0958-3}
 
@@
 \url{http://www.nature.com/articles/s41586-018-0411-9/}
+
+\leavevmode\hypertarget{ref-spawn_harmonized_2020}{}%
+Spawn S A, Sullivan C C, Lark T J and Gibbs H K 2020 Harmonized global
+maps of above and belowground biomass carbon density in the year 2010
+\emph{Scientific Data} \textbf{7} 112 Online:
+\url{http://www.nature.com/articles/s41597-020-0444-4}
+
+\leavevmode\hypertarget{ref-stoy_data-driven_2013}{}%
+Stoy P C, Mauder M, Foken T, Marcolla B, Boegh E, Ibrom A, Arain M A,
+Arneth A, Aurela M, Bernhofer C, Cescatti A, Dellwik E, Duce P, Gianelle
+D, Gorsel E van, Kiely G, Knohl A, Margolis H, McCaughey H, Merbold L,
+Montagnani L, Papale D, Reichstein M, Saunders M, Serrano-Ortiz P,
+Sottocornola M, Spano D, Vaccari F and Varlagin A 2013 A data-driven
+analysis of energy balance closure across FLUXNET research sites: The
+role of landscape scale heterogeneity \emph{Agricultural and Forest
+Meteorology} \textbf{171-172} 137--52 Online:
+\url{https://linkinghub.elsevier.com/retrieve/pii/S0168192312003413}
+
+\leavevmode\hypertarget{ref-sulman_multiple_2018}{}%
+Sulman B N, Moore J A M, Abramoff R, Averill C, Kivlin S, Georgiou K,
+Sridhar B, Hartman M D, Wang G, Wieder W R, Bradford M A, Luo Y, Mayes M
+A, Morrison E, Riley W J, Salazar A, Schimel J P, Tang J and Classen A T
+2018 Multiple models and experiments underscore large uncertainty in
+soil carbon dynamics \emph{Biogeochemistry} \textbf{141} 109--23 Online:
+\url{https://doi.org/10.1007/s10533-018-0509-z}
 
@@
 
+\leavevmode\hypertarget{ref-tubiello_carbon_2020}{}%
+Tubiello F N, Pekkarinen A, Marklund L, Wanner N, Conchedda G, Federici
+S, Rossi S and Grassi G 2020 Carbon Emissions and Removals by Forests:
+New Estimates 1990\&ndash;2020 \emph{Earth System Science Data
+Discussions} 1--21 Online:
+\url{https://essd.copernicus.org/preprints/essd-2020-203/}
+
+\leavevmode\hypertarget{ref-vargas_biomass_2008}{}%
+Vargas R, Allen M F and Allen E B 2008 Biomass and carbon accumulation
+in a fire chronosequence of a seasonally dry tropical forest
+\emph{Global Change Biology} \textbf{14} 109--24 Online:
+\url{http://onlinelibrary.wiley.com/doi/abs/10.1111/j.1365-2486.2007.01462.x}
+
+\leavevmode\hypertarget{ref-wang_golum-cnp_2018}{}%
+Wang Y, Ciais P, Goll D, Huang Y, Luo Y, Wang Y-P, Bloom A A, Broquet G,
+Hartmann J, Peng S, Penuelas J, Piao S, Sardans J, Stocker B D, Wang R,
+Zaehle S and Zechmeister-Boltenstern S 2018 GOLUM-CNP v1.0: A
+data-driven modeling of carbon, nitrogen and phosphorus cycles in major
+terrestrial biomes \emph{Geoscientific Model Development} \textbf{11}
+3903--28 Online: \url{https://gmd.copernicus.org/articles/11/3903/2018/}
+
+\leavevmode\hypertarget{ref-warner_spatial_2019}{}%
+Warner D L, Bond‐Lamberty B, Jian J, Stell E and Vargas R 2019 Spatial
+Predictions and Associated Uncertainty of Annual Soil Respiration at the
+Global Scale \emph{Global Biogeochemical Cycles} \textbf{33} 1733--45
+Online:
+\url{https://onlinelibrary.wiley.com/doi/abs/10.1029/2019GB006264}
+
+\leavevmode\hypertarget{ref-van_der_werf_global_2017}{}%
+Werf G R van der, Randerson J T, Giglio L, Leeuwen T T van, Chen Y,
+Rogers B M, Mu M, Marle M J E van, Morton D C, Collatz G J, Yokelson R J
+and Kasibhatla P S 2017 Global fire emissions estimates during
+1997--2016 \emph{Earth System Science Data} \textbf{9} 697--720 Online:
+\url{https://essd.copernicus.org/articles/9/697/2017/}
+
+\leavevmode\hypertarget{ref-williams_impacts_2014}{}%
+Williams C A, Collatz G J, Masek J, Huang C and Goward S N 2014 Impacts
+of disturbance history on forest carbon stocks and fluxes: Merging
+satellite disturbance mapping with forest inventory data in a carbon
+cycle model framework \emph{Remote Sensing of Environment} \textbf{151}
+57--71 Online:
+\url{https://linkinghub.elsevier.com/retrieve/pii/S0034425713004185}
+
+\leavevmode\hypertarget{ref-wilson_stability_2016}{}%
+Wilson R M, Hopple A M, Tfaily M M, Sebestyen S D, Schadt C W,
+Pfeifer-Meister L, Medvedeff C, McFarlane K J, Kostka J E, Kolton M,
+Kolka R K, Kluber L A, Keller J K, Guilderson T P, Griffiths N A,
+Chanton J P, Bridgham S D and Hanson P J 2016 Stability of peatland
+carbon to rising temperatures \emph{Nature Communications} \textbf{7}
+13723 Online: \url{http://www.nature.com/articles/ncomms13723}
+
+\leavevmode\hypertarget{ref-xu_contribution_2016}{}%
+Xu M and Shang H 2016 Contribution of soil respiration to the global
+carbon equation \emph{Journal of Plant Physiology} \textbf{203} 16--28
+Online:
+\url{https://linkinghub.elsevier.com/retrieve/pii/S0176161716301742}
+
+\leavevmode\hypertarget{ref-yang_carbon_2011}{}%
+Yang Y, Luo Y and Finzi A C 2011 Carbon and nitrogen dynamics during
+forest stand development: A global synthesis \emph{New Phytologist}
+\textbf{190} 977 Online:
+\url{http://dx.doi.org/10.1111/j.1469-8137.2011.03645.x}
+\end{cslreferences}
+
 \end{document}

### 2021-01-13 manuscript/response_to_reviews.Rmd
start on response to reviews

@@
 
+*This requires careful consideration and response*
+
 # Referee: 3
@@
 
+**Thank you**
+
 However, I also found that many aspects of the paper gave me considerable pause.
@@
 There are no process-oriented hypotheses or frameworks against which the results are evaluated.
+
+(*add some broad-brush hypotheses/ frameworks*)
 
@@
 A deeper dive into the results, combined with some mechanistic grounding, might reveal to what extent this result represents a climate-age interaction that is predicted from the existing theory.
+
+**Regarding reference to Odum, ...**, *add citation to Odum*
+
+**Regarding the question of showing flux trends as a function of age beyond 100 years, this unfortunately doesn't make sense within the context of the database because tropical forests can rarely be aged beyond 100 years (if that).
+This is because tropical trees do not form annual rings.
+...**
 

### 2021-01-14 manuscript/response_to_reviews.Rmd
thorough review of reviewer comments, planning responses

@@
 
+*This just requires adding some explanation.*
+
 a.
@@
 Is this the data from Bond-Lamberty and Thomson (2010) or has it been updated?
+
+*This just requires adding some explanation.*
 
@@
 
+*This just requires adding some explanation.*
+
 c. If Luyssaert et al (2007) is a source of data, is there circularity in the comparison with latitudinal trends found in other datasets, for which Luyssaert et al is cited (lines 317-318)?
+
+*This just requires some re-wording. We are building on/ updating results of Luyssaert.*
 
@@
 
+*This just requires adding some explanation.*
+
 b.
@@
 Presumably, this is because the data fall within the closure criteria, but what are we to make of the inconsistent data?
+
+*This just requires adding some explanation.*
 
@@
 
+*This just requires adding some explanation. Note that R2 and R3 have very different interpretations of the "closure"*
+
 3. There is a general sloppiness to the manuscript that makes me think the authors did not proofread the manuscript carefully.
+
+*This is true. Need to fix all points below. None difficult.*
 
@@
 **Regarding the question of showing flux trends as a function of age beyond 100 years, this unfortunately doesn't make sense within the context of the database because tropical forests can rarely be aged beyond 100 years (if that).
-This is because tropical trees do not form annual rings.
-...**
+This is because tropical trees rarely form annual rings, which are used to age (older) extratropical forests.
+It would not make sense to treat tropical and extra-tropical forests differently.**
+
+**Regarding the NEP result for mature forests, ...** *(Note that R4 also found this surprising. I think an Odum-esque diagram can help clarify.)*
 
@@
 
+**This is a good point, but unlikely to have much influence at this relatively coarse scale of analysis. ...**
+
 Second, I had some questions about the representativeness of the dataset.
@@
 For example, if mean annual climate versus presentation for all boreal, temperate and tropical locations are shown in a scatterplot (for example, using reanalysis data), and then mean annual temperature and precipation of the observations are shown on top, how much of the “climate space” is covered by the dataset?
+
+*[Issue #57](https://github.com/forc-db/ERL-review/issues/57)*
 
@@
 My advice is to move some of the budgets to SI, and include in the main manuscript more figures that clearly illustrate the most interesting trends with biome, and to allow an expansion of the results in Figure 6 (for example, by first showing scatterplots of all forests, young and old, as a function of stand age), and then perhaps another that is a box plot comparison of mature versus old forests in each biome.
+
+- *move young forest C budgets to SI?*
+
+- *remove maps from age trend figures?*
+
+**Regarding the suggestion of showing age of old forests as a scatterplot, this is not possible give unknown ages of most mature tropical forests, as explained above.**
+
+**Regarding the proposed figure rearrangement, ...** *[issue #58](https://github.com/forc-db/ERL-review/issues/58)*
 
@@
 
+*No changes here; just respond....*
+
+**We prefer not to separate forests of known age into categories, but rather to represent age as a continuous variable. ...**
+
 The consistency check (e.g.
@@
 As long as the aggregate variable has a large standard deviation, the results are likely guaranteed to be “consistent.” It seems this metric would benefit from some simulations (perhaps with artificially generated data) to understand exactly how poor closure needs to be at the site-level to generate an inconsistent results when aggregated across sites.
+
+*(No! Too much work!)*
+
+*I'm inclined to place less--not more--emphasis on representativeness: [Issue #60](https://github.com/forc-db/ERL-review/issues/60) *
 
@@
 
+*add discussion of the more intact stands studied here, vs what actually exists. straightforward.*
+
 Introduction:
@@
 
+*text edit*
+
 L97: Typo “Since the its most”
+
+*text edit*
 
@@
 
+*straightforward--just a bit of text*
+
 Results:
@@
 
+*straightforward--just a bit of text edit.*
+
 L270-271: The wording of the sentence “There were sufficient data to model….” Seems a bit awkward. Should it read “…WHICH were also significant…”?
+
+*straightforward--just a bit of text edit.*
 
@@
 
+*It strikes me that both R3 and R4 find this surprising. I don't. An Odum-esque conceptual diagram + some discussion should help to explain this.*
+
 L297: The fact that ForC does not include soil C is important.
@@
 I think the authors should consider clearly indicating in the Methods section, and perhaps reiterating in the Results and Discussion, that stocks here are defined as litter layer, biomass, and necromass, but excludes C in soil.
+
+*straightforward--just a bit of text edit.*
 
@@
 
+*straightforward--just a bit of text edit.*
+
 L382: You might consider changing text to “increases with age AT LEAST up to the 100-yr threshold examined here” so that it does not come across as suggesting that NEP only increases for the first hundred years of stand development, which we know is not true.
+
+*straightforward--just a bit of text edit.*
 
@@
 Related to this point, is the high NEP of temperate forests driven by the relatively young nature of temperate broadleaf forests in the eastern U.S., where a lot of data exist?
+
+*straightforward--just a bit of text edit.*
 

### 2021-01-14 manuscript/response_to_reviews.Rmd
add points for letter to editor

@@
                   \usepackage{setspace}\onehalfspacing
-                  \usepackage{lineno}\linenumbers
 
@@
 
-## REFEREE REPORT(S):
+## LETTER TO EDITOR:
+
+We have made the following substantive changes:
+
+- Changed the figure set in accordance with comments from R3. Specifically, 
+(1) moved C cycle diagrams for young forests to SI; 
+(2) removed maps from age trend diagrams (available in SI) from age trend diagrams, which makes the figures less busy/ easier to read;
+(3) added figures showing changes in fluxes through time
+(4) added a Odum-esque schematic, for framing review
+
+- Improved conceptual framework and reference to previous literature (e.g., Odum)
+
+- Shifted emphasis more towards the ecological questions (trends across biomes and with age), while retaining (with less emphasis) the analysis of database representativeness
+
+\newpage
+
+## REFEREE REPORT(S) & RESPONSES:
 
@@
 
+\newpage
 ## References

### 2021-01-14 manuscript/ERL_global_C_review.Rmd
starting some rearrangement

--- 2020-10-06 manuscript/ERL_global_C_review.Rmd
@@
 ```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
-generate_figures=FALSE
-```
 
-```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
+# KNIT IN FIGURES? 
+generate_figures=TRUE
+#note: table fails when knitting in word
+
+# DISPLAY ITEMS
+#figures
+schematic =1
+figure_map=2
+diagram_TrB_mature= 3
+diagram_TeB_mature=4
+diagram_TeN_mature=5
+diagram_BoN_mature=6
+flux_age_trends=7
+stock_age_trends=8
+
+#tables
+variables_table=1
+
+#SI
+diagram_TrB_young= "S1"
+diagram_TeB_young= "S2"
+diagram_TeN_young= "S3"
+diagram_BoN_young= "S4"
+
 #READING IN TABLES TO GET STATS FOR TEXT
@@
 Here, we provide a robust and comprehensive analysis of carbon cycling from a stand to global level, and by biome and stand age, using the largest global compilation of forest carbon data, which is available in our open source Global Carbon Forest database (*ForC*; Fig.
-1).
+`r figure_map`).
 Our primary goal is to provide a data-driven summary of our current state of knowledge on broad trends in forest C cycling. 
@@
 
-1. How thoroughly can we represent C budgets for each of the world's major forest biomes (*i.e.*, tropical, temperate broadleaf and deciduous, boreal) based on the current *ForC* data?
+~~1. How thoroughly can we represent C budgets for each of the world's major forest biomes (*i.e.*, tropical, temperate broadleaf and deciduous, boreal) based on the current *ForC* data?~~
 
@@
 
-While components of these questions have been previously addressed [@luyssaert_co2_2007; @anderson-teixeira_carbon_2016; @cook-patton_mapping_2020; @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and will serve as a foundation for improved understanding of global forest C cycling and highlight where key sources of uncertainty still reside.
+While researchers have been addressing components of these questions for more than half a century [@odum_strategy_1969; @luyssaert_co2_2007; @anderson-teixeira_carbon_2016; @cook-patton_mapping_2020; @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and will serve as a foundation for improved understanding of global forest C cycling and highlight where key sources of uncertainty still reside.
 
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 1 | Map of sites included in this analysis.
+```{r schematic, eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 1 | DRAFT SCHEMATIC, to inlcude bar plots of mature biome differences as a second column.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/schematic_opt2.png")
+```
+
+```{r figure_map, eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 2 | Map of sites included in this analysis.
 Symbols are colored according to the number of records at each site.
@@
 
-This review synthesizes data from the *ForC* database [Fig. 1; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
+This review synthesizes data from the *ForC* database [Fig. `r figure_map`; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
 *ForC* amalgamates numerous intermediary data sets [*e.g.*, @luyssaert_co2_2007; @bond-lamberty_global_2010; @cook-patton_mapping_2020] and original studies.
@@
 We grouped forests into four broad biome types based on climate zones and dominant vegetation type (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
-Climate zones (Fig. 1) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
+Climate zones (Fig. `r figure_map`) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
 We defined the tropical biome as including all equatorial (A) zones, temperate biomes as including all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and the boreal biome as including the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd).
@@
 
-Of the `r nrow(ForC_measurements)` records in *ForC* v3.0, `r n_records` met our strict criteria for inclusion in this study (Fig. 1).
+Of the `r nrow(ForC_measurements)` records in *ForC* v3.0, `r n_records` met our strict criteria for inclusion in this study (Fig. `r figure_map`).
 These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas.
@@
 \blandscape
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 2 | C cycle diagram for mature tropical broadleaf forests.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 3 | C cycle diagram for mature tropical broadleaf forests.
 Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
@@
 
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 3 | C cycle diagram for mature temperate broadleaf forests.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 4 | C cycle diagram for mature temperate broadleaf forests.
 Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
@@
 
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 4 | C cycle diagram for mature temperate conifer forests.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 5 | C cycle diagram for mature temperate conifer forests.
 Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
@@
 
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 5 | C cycle diagram for mature boreal conifer forests.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | C cycle diagram for mature boreal conifer forests.
 Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.

### 2021-01-14 manuscript/tables_figures/schematic/plotBarStackGroups.m
start at schematic
#61

@@ -1,0 +1,52 @@
+function [] = plotBarStackGroups(stackData, groupLabels,bar_colors)
+%% Plot a set of stacked bars, but group them according to labels provided.
+%%
+%% Params: 
+%%      stackData is a 3D matrix (i.e., stackData(i, j, k) => (Group, Stack, StackElement)) 
+%%      groupLabels is a CELL type (i.e., { 'a', 1 , 20, 'because' };)
+%%
+%% Copyright 2011 Evan Bollig (bollig at scs DOT fsu ANOTHERDOT edu
+%%
+%% 
+NumGroupsPerAxis = size(stackData, 1);
+NumStacksPerGroup = size(stackData, 2);
+
+
+% Count off the number of bins
+groupBins = 1:NumGroupsPerAxis;
+MaxGroupWidth = 0.65; % Fraction of 1. If 1, then we have all bars in groups touching
+groupOffset = MaxGroupWidth/NumStacksPerGroup;
+%figure
+    hold on; 
+for i=1:NumStacksPerGroup
+
+    Y = squeeze(stackData(:,i,:));
+    
+    % Center the bars:
+    
+    internalPosCount = i - ((NumStacksPerGroup+1) / 2);
+    
+    % Offset the group draw positions:
+    groupDrawPos = (internalPosCount)* groupOffset + groupBins;
+    
+    h(i,:) = bar(Y, 'stacked');
+    for n=1:size(stackData,3)
+        if sum(Y(:,n))==0
+            set(get(get(h(i,n),'Annotation'),'LegendInformation'),'IconDisplayStyle','off');
+        end
+    end
+    set(h(i,:),'BarWidth',groupOffset);
+    set(h(i,:),'XData',groupDrawPos);
+end
+hold off;
+set(gca,'XTickMode','manual');
+set(gca,'XTick',1:NumGroupsPerAxis);
+set(gca,'XTickLabelMode','manual');
+set(gca,'XTickLabel',groupLabels);
+
+%set(h(1,1),'FaceColor',[11 102 35]/255); %KAT
+%for n=1:size(stackData,3)  %KAT
+%    set(h(2,n),'FaceColor',bar_colors(n,:)); %KAT
+%end %KAT
+
+end 

### 2021-01-14 manuscript/tables_figures/schematic/schematic.m

@@ -1,0 +1,114 @@
+% schematic figure for ERL review
+% Kristina Anderson Teixeira
+% January 2021
+
+clf; clear; close all;
+set(0,'DefaultLegendAutoUpdate','off');
+
+
+%plot settings
+
+    %subplot positions
+    plot_width=0.32;
+    plot_height=0.15;
+    plot_space_vertical=(1-plot_height*5)/5-.01;
+    plot_space_horizontal=(1-plot_width*2.5)/2-.05;
+
+    %color
+    facecolor_cat=[.1 .4 .9; 0.8 0 0;  .9 .527 .527 ; 1 .9 .9 ]; %blue/ red
+
+    
+    %font
+    fs= 10; %font size
+
+
+
+%% ~~~~~~~CREATE DATA FOR PLOTTING~~~~~~~~~~
+age = linspace (1,100,100);
+biome_names={'tropical', 'temperate', 'boreal'};
+
+%FLUXES
+% 1. by age
+%for stacked plot
+R_auto = 2* log10(age);
+NPP = R_auto;
+
+fluxes = [R_auto; NPP]; %matrix with all fluxes for stacked plot
+flux_names = {'R_{auto}', 'NPP'};
+
+%others
+GPP = sum(fluxes,1);
+
+% 2. mature fluxes
+Flux_ratio_temp_trop=0.5;
+Flux_ratio_boreaal_trop=0.25;
+
+%for stacked plot
+mature_fluxes = [fluxes(:,end)'; Flux_ratio_temp_trop*fluxes(:,end)' ; Flux_ratio_boreaal_trop*fluxes(:,end)'];
+
+%others
+
+%STOCKS
+% 1. by age
+%for stacked plot
+biomass=54* log10(age); 
+DW= 2.6*log10(age);
+
+stocks = [biomass; DW]; %matrix with all stocks for stacked plot
+stock_names = {'B_{tot}', 'DW'};
+
+%others
+
+
+
+% 2. mature forests
+Stock_ratio_temp_trop=0.9;
+Stock_ratio_boreaal_trop=0.7;
+
+mature_stocks = [stocks(:,end)'; Stock_ratio_temp_trop*stocks(:,end)' ; Stock_ratio_boreaal_trop*stocks(:,end)'];
+
+
+%% ~~~~~~~PLOTTING~~~~~~~~~~
+
+figure (1)
+subplot (2,3,1:2)
+%flux plot:
[…]

### 2021-01-14 manuscript/ERL_global_C_SI.Rmd
starting some rearrangement
fixed SI knitting problem

@@
                   \usepackage{setspace}\onehalfspacing
-                  \usepackage{lineno}\linenumbers
 
 ---
+
+```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
+path_to_dir <- dirname(dirname(getwd())) # get path to repo
+
+# MOVE FIGURE FILES (Need to do this because I got an error when using the absolute path to call figures. However, it's not working at the moment, so did manual copy paste. )
+
+#C cycle diagrams
+#do.call(file.remove, list(list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/tables_figures/C_cycle_diagrams/", full.names = TRUE))) # first remove old files
+
+#flist <- list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/", full.names = TRUE)
+#file.copy(flist, "/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/tables_figures/C_cycle_diagrams/")
+
+#Age diagrams
+#do.call(file.remove, list(list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/tables_figures/age_trends/", full.names = TRUE))) # first remove old files
+
+#flist <- list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForC-db/ForC/figures/age_trends/", full.names = TRUE)
+#file.copy(flist, "/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/tables_figures/age_trends/")
+
+# FIGURE NUMBERS
+
+#SI
+diagram_TrB_young= "S1"
+diagram_TeB_young= "S2"
+diagram_TeN_young= "S3"
+diagram_BoN_young= "S4"
+NEP_fig= "S5"
+GPP_fig= "S6"
+NPP_fig= "S7"
+ANPP_fig= "S8"
+ANPP_woody_fig= "S9"
+ANPP_stem_fig= "S10"
+ANPP_foliage_fig= "S11"
+ANPP_litterfall_fig= "S12"
+BNPP_fig= "S13"
+BNPP_coarse_fig= "S14"
+BNPP_fine_fig= "S15"
+R_eco_fig= "S16"
+R_root_fig= "S17"
+R_soil_fig= "S18"
+R_het_soil_fig= "S19"
+
+```
+
 \raggedright
@@
 \newpage
-## Figure S1. Age trends and biome differences for $NEP$
+## Figure `r NEP_fig`. Age trends and biome differences for $NEP$
 
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/NEP.png")
+  knitr::include_graphics("tables_figures/age_trends/NEP.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/GPP.png")
+  knitr::include_graphics("tables_figures/age_trends/GPP.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/NPP.png")
+  knitr::include_graphics("tables_figures/age_trends/NPP.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/ANPP.png")
+  knitr::include_graphics("tables_figures/age_trends/ANPP.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/ANPP_woody.png")
+  knitr::include_graphics("tables_figures/age_trends/ANPP_woody.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/ANPP_stem.png")
+  knitr::include_graphics("tables_figures/age_trends/ANPP_stem.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/ANPP_foliage.png")
+  knitr::include_graphics("tables_figures/age_trends/ANPP_foliage.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/ANPP_litterfall.png")
+  knitr::include_graphics("tables_figures/age_trends/ANPP_litterfall.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/BNPP.png")
+  knitr::include_graphics("tables_figures/age_trends/BNPP.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/BNPP_coarse.png")
+  knitr::include_graphics("tables_figures/age_trends/BNPP_coarse.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/BNPP_fine.png")
+  knitr::include_graphics("tables_figures/age_trends/BNPP_fine.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/R_eco.png")
+  knitr::include_graphics("tables_figures/age_trends/R_eco.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/R_root.png")
+  knitr::include_graphics("tables_figures/age_trends/R_root.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/R_soil.png")
+  knitr::include_graphics("tables_figures/age_trends/R_soil.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/R_het_soil.png")
+  knitr::include_graphics("tables_figures/age_trends/R_het_soil.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass.png")
+  knitr::include_graphics("tables_figures/age_trends/biomass.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass_ag.png")
+  knitr::include_graphics("tables_figures/age_trends/biomass_ag.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass_ag_woody.png")
+  knitr::include_graphics("tables_figures/age_trends/biomass_ag_woody.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass_foliage.png")
+  knitr::include_graphics("tables_figures/age_trends/biomass_foliage.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass_root.png")
+  knitr::include_graphics("tables_figures/age_trends/biomass_root.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass_root_coarse.png")
+  knitr::include_graphics("tables_figures/age_trends/biomass_root_coarse.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/biomass_root_fine.png")
+  knitr::include_graphics("tables_figures/age_trends/biomass_root_fine.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/deadwood.png")
+  knitr::include_graphics("tables_figures/age_trends/deadwood.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/deadwood_standing.png")
+  knitr::include_graphics("tables_figures/age_trends/deadwood_standing.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/deadwood_down.png")
+  knitr::include_graphics("tables_figures/age_trends/deadwood_down.png")
 ```
@@
 
-  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/age_trends/organic.layer.png")
+  knitr::include_graphics("tables_figures/age_trends/organic_layer.png")
 ```

### 2021-01-14 manuscript/ERL_global_C_SI.Rmd
update /automate numbering on age trend figures

@@
 R_het_soil_fig= "S19"
+B_tot_fig="S20"
+B_ag_fig="S21"
+B_ag_wood_fig="S22"
+B_foliage_fig="S23"
+B_root_fig="S24"
+B_root_coarse_fig="S25"
+B_root_fine_fig="S26"
+DW_tot_fig="S27"
+DW_standing_fig="S28"
+DW_down_fig="S29"
+OL_fig="S30"
 
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S1 | Age trends and biome differences for $NEP$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $NEP$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S2. Age trends and biome differences for $GPP$
+## Figure `r GPP_fig`. Age trends and biome differences for $GPP$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S2 | Age trends and biome differences for $GPP$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $GPP$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S3. Age trends and biome differences for $NPP$
+## Figure `r NPP_fig`. Age trends and biome differences for $NPP$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S3 | Age trends and biome differences for $NPP$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $NPP$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S4. Age trends and biome differences for $ANPP$
+## Figure `r ANPP_fig`. Age trends and biome differences for $ANPP$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S4 | Age trends and biome differences for $ANPP$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $ANPP$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S5. Age trends and biome differences for $ANPP_{woody}$
+## Figure `r ANPP_woody_fig`. Age trends and biome differences for $ANPP_{woody}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S5 | Age trends and biome differences for $ANPP_{woody}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $ANPP_{woody}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S6. Age trends and biome differences for $ANPP_{stem}$
+## Figure `r ANPP_stem_fig`. Age trends and biome differences for $ANPP_{stem}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S6 | Age trends and biome differences for $ANPP_{stem}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $ANPP_{stem}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S7. Age trends and biome differences for $ANPP_{foliage}$
+## Figure `r ANPP_foliage_fig`. Age trends and biome differences for $ANPP_{foliage}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S7 | Age trends and biome differences for $ANPP_{foliage}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $ANPP_{foliage}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S8. Age trends and biome differences for $ANPP_{litterfall}$
+## Figure `r ANPP_litterfall_fig`. Age trends and biome differences for $ANPP_{litterfall}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S8 | Age trends and biome differences for $ANPP_{litterfall}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $ANPP_{litterfall}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S9. Age trends and biome differences for $BNPP$
+## Figure `r BNPP_fig`. Age trends and biome differences for $BNPP$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S9 | Age trends and biome differences for $BNPP$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $BNPP$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S10. Age trends and biome differences for $BNPP_{coarse}$
+## Figure `r BNPP_coarse_fig`. Age trends and biome differences for $BNPP_{coarse}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S10 | Age trends and biome differences for $BNPP_{coarse}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $BNPP_{coarse}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S11. Age trends and biome differences for $BNPP_{fine}$
+## Figure `r BNPP_fine_fig`. Age trends and biome differences for $BNPP_{fine}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S11 | Age trends and biome differences for $BNPP_{fine}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $BNPP_{fine}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S12. Age trends and biome differences for $R_{eco}$
+## Figure `r R_eco_fig`. Age trends and biome differences for $R_{eco}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S12 | Age trends and biome differences for $R_{eco}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $R_{eco}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S13. Age trends and biome differences for $R_{root}$
+## Figure `r R_root_fig`. Age trends and biome differences for $R_{root}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S13 | Age trends and biome differences for $R_{root}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $R_{root}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S14. Age trends and biome differences for $R_{soil}$
+## Figure `r R_soil_fig`. Age trends and biome differences for $R_{soil}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S14 | Age trends and biome differences for $R_{soil}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $R_{soil}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S15. Age trends and biome differences for $R_{het-soil}$
+## Figure `r R_het_soil_fig`. Age trends and biome differences for $R_{het-soil}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S15 | Age trends and biome differences for $R_{het-soil}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $R_{het-soil}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S16. Age trends and biome differences for $B_{tot}$
+## Figure `r B_tot_fig`. Age trends and biome differences for $B_{tot}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S16 | Age trends and biome differences for $B_{tot}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $B_{tot}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S17. Age trends and biome differences for $B_{ag}$
+## Figure `r B_ag_fig`. Age trends and biome differences for $B_{ag}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S17 | Age trends and biome differences for $B_{ag}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $B_{ag}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S18. Age trends and biome differences for $B_{ag-wood}$
+## Figure `r B_ag_wood_fig`. Age trends and biome differences for $B_{ag-wood}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S18 | Age trends and biome differences for $B_{ag-wood}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $B_{ag-wood}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S19. Age trends and biome differences for $B_{foliage}$
+## Figure `r B_foliage_fig`. Age trends and biome differences for $B_{foliage}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S19 | Age trends and biome differences for $B_{foliage}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $B_{foliage}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S20. Age trends and biome differences for $B_{root}$
+## Figure `r B_root_fig`. Age trends and biome differences for $B_{root}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S20 | Age trends and biome differences for $B_{root}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $B_{root}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S21. Age trends and biome differences for $B_{root-coarse}$
+## Figure `r B_root_coarse_fig`. Age trends and biome differences for $B_{root-coarse}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S21 | Age trends and biome differences for $B_{root-coarse}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $B_{root-coarse}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S22. Age trends and biome differences for $B_{root-fine}$
+## Figure `r B_root_fine_fig`. Age trends and biome differences for $B_{root-fine}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S22 | Age trends and biome differences for $B_{root-fine}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $B_{root-fine}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S23. Age trends and biome differences for $DW_{tot}$
+## Figure `r DW_tot_fig`. Age trends and biome differences for $DW_{tot}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S23 | Age trends and biome differences for $DW_{tot}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $DW_{tot}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S24. Age trends and biome differences for $DW_{standing}$
+## Figure `r DW_standing_fig`. Age trends and biome differences for $DW_{standing}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S24 | Age trends and biome differences for $DW_{standing}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $DW_{standing}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S25. Age trends and biome differences for $DW_{down}$
+## Figure `r DW_down_fig`. Age trends and biome differences for $DW_{down}$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S25 | Age trends and biome differences for $DW_{down}$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $DW_{down}$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 \newpage
-## Figure S26. Age trends and biome differences for $OL$
+## Figure `r OL_fig`. Age trends and biome differences for $OL$
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S26 | Age trends and biome differences for $OL$.
+```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $OL$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).

### 2021-01-14 manuscript/ERL_global_C_SI.Rmd
added temperate C cycle diagrams to SI
had to copy figures to this repo + rename

#58

@@
 ```
+
+\blandscape
+## Figure `r diagram_TrB_young`. C cycle diagram for young tropical broadleaf forests.
+```{r eval = TRUE, echo=FALSE, fig.align = "center", fig.cap = "C cycle diagram for young tropical broadleaf forests.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
+Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.", out.width='90%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/C_cycle_diagrams/Tropical_broadleaf_YOUNG.jpeg")
+```
+
+## Figure `r diagram_TeB_young`. C cycle diagram for young temperate broadleaf forests.
+```{r eval = TRUE, echo=FALSE, fig.align = "center", fig.cap = "C cycle diagram for young temperate broadleaf forests.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
+Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.", out.width='90%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/C_cycle_diagrams/Temperate_broadleaf_YOUNG.jpeg")
+```
+
+## Figure `r diagram_TeN_young`. C cycle diagram for young temperate conifer forests.
+```{r eval = TRUE, echo=FALSE, fig.align = "center", fig.cap = " C cycle diagram for young temperate conifer forests.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
+Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.", out.width='90%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/C_cycle_diagrams/Temperate_conifer_YOUNG.jpeg")
+```
+
+## Figure `r diagram_BoN_young`. C cycle diagram for young boreal conifer forests.
+```{r eval = TRUE, echo=FALSE, fig.align = "center", fig.cap = "C cycle diagram for young boreal conifer forests.
+Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
+Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
+All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.", out.width='90%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/C_cycle_diagrams/Boreal_conifer_YOUNG.jpeg")
+```
+\elandscape
 

### 2021-01-14 manuscript/ERL_global_C_review.Rmd
removed young forest diagrams, redid (automated) fig numbering
#58

@@
 diagram_BoN_mature=6
+all_diagrams_mature="3-6"
 flux_age_trends=7
@@
 #SI
+
+all_diagrams_young="S1-S4"
+all_flux_age_trends="S5-S19"
+all_stock_age_trends= "S20-S30"
+all_age_trends="S5-S30"
+
+#copied from SI
 diagram_TrB_young= "S1"
@@
 diagram_BoN_young= "S4"
+NEP_fig= "S5"
+GPP_fig= "S6"
+NPP_fig= "S7"
+ANPP_fig= "S8"
+ANPP_woody_fig= "S9"
+ANPP_stem_fig= "S10"
+ANPP_foliage_fig= "S11"
+ANPP_litterfall_fig= "S12"
+BNPP_fig= "S13"
+BNPP_coarse_fig= "S14"
+BNPP_fine_fig= "S15"
+R_eco_fig= "S16"
+R_root_fig= "S17"
+R_soil_fig= "S18"
+R_het_soil_fig= "S19"
+B_tot_fig="S20"
+B_ag_fig="S21"
+B_ag_wood_fig="S22"
+B_foliage_fig="S23"
+B_root_fig="S24"
+B_root_coarse_fig="S25"
+B_root_fine_fig="S26"
+DW_tot_fig="S27"
+DW_standing_fig="S28"
+DW_down_fig="S29"
+OL_fig="S30"
 
@@
 
+```{r schematic, eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 1 | DRAFT SCHEMATIC.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/schematic/schematic.png")
+```
+
 Here, we provide a robust and comprehensive analysis of carbon cycling from a stand to global level, and by biome and stand age, using the largest global compilation of forest carbon data, which is available in our open source Global Carbon Forest database (*ForC*; Fig.
@@
 While researchers have been addressing components of these questions for more than half a century [@odum_strategy_1969; @luyssaert_co2_2007; @anderson-teixeira_carbon_2016; @cook-patton_mapping_2020; @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and will serve as a foundation for improved understanding of global forest C cycling and highlight where key sources of uncertainty still reside.
-
-```{r schematic, eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 1 | DRAFT SCHEMATIC, to inlcude bar plots of mature biome differences as a second column.", out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics("tables_figures/schematic_opt2.png")
-```
 
@@
 In all of these cases--including C fluxes into, within, and out of the ecosystem--C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs.
-6, S1-S15).
+`r flux_age_trends`, `r all_flux_age_trends`).
 Differences between tropical and boreal forests were always significant, with temperate forests intermediate and significantly different from one or both.
@@
 For two of the variables without significant differences among biomes ($ANPP_{litterfall}$ and $BNPP_{fine}$; Figs.
-S8 and S11, respectively), the same general trends applied but were not statistically significant.
+`r ANPP_litterfall_fig` and `r BNPP_fine_fig`, respectively), the same general trends applied but were not statistically significant.
 Another exception was for $BNPP_{root-coarse}$, where all records came from high-biomass forests in the US Pacific Northwest, resulting in marginally higher values for the temperate conifer biome (Table 1, Fig.
-S10; differences significant in mixed effects model but not in post-hoc pairwise comparison).
+`r BNPP_coarse_fig`; differences significant in mixed effects model but not in post-hoc pairwise comparison).
 
 The most notable exception to the pattern of decreasing flux from tropical to boreal biomes was $NEP$, with no significant differences across biomes but with the largest average in temperate broadleaf forests, followed by tropical, boreal, and temperate conifer forests (Figs.
-5,S1).
+`r flux_age_trends`, `rNEP_fig`).
 
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEP$.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEP$.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 Individual figures for each flux with sufficient data are given in the Supplement (Figs.
-S1-S15).", out.width='100%', fig.pos='H'}
+S4-S19).", out.width='100%', fig.pos='H'}
 
@@
 There were sufficient data to assess mature forest biome differences for nine stock variables, and significant differences among biomes were detected for five variables ($B_{tot}$, $B_{ag}$, $B_{ag-wood}$, $B_{foliage}$, $B_{root-coarse}$; Table 1).
-C stocks had less consistent patterns across biomes (Figs. 7, S16-S26).
+C stocks had less consistent patterns across biomes (Figs. `r stock_age_trends`, `r all_stock_age_trends`).
 For $B_{tot}$ and $B_{ag}$, tropical broadleaf forests had the highest biomass and boreal forests the lowest, with temperate broadleaf and needleleaf ($B_{ag}$ only) intermediate.
@@
 
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
@@
 Individual figures for each stock with sufficient data are given in the Supplement (Figs.
-S16-S26).", out.width='100%', fig.pos='H'}
+S20-S30).", out.width='100%', fig.pos='H'}
 
@@
 Average C cycles for forests <100 years old are presented in Figures 8-11.
-Both C stocks and fluxes commonly increased significantly with stand age (Tables 1, S2, Figs. 6-11, S1-S26; detailed below).
+Both C stocks and fluxes commonly increased significantly with stand age (Tables 1, S2, Figs. `r flux_age_trends`- `r stock_age_trends`, `r all_age_trends`; detailed below).
 
-*ForC* contained 16 C flux variables with sufficient data for analyses of age trends in young forests (see Methods) (Figs. 6, S1-S15).
+*ForC* contained 16 C flux variables with sufficient data for analyses of age trends in young forests (see Methods) (Figs. `r flux_age_trends`, `r all_flux_age_trends`).
 Of these, ten increased significantly with age: $NEP$, $GPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{foliage}$, $ANPP_{litterfal}$, $BNPP$, $BNPP_{fine}$, $R_{eco}$, and $R_{root}$.
@@
 Notably, and in contrast to the lack of biome differences in $NEP$ for mature forests (Fig.
-6), the tendency for temperate forests to have greater fluxes than boreal forests held for $NEP$ in regrowth forests (tropical forests excluded because of insufficient data).
+`r flux_age_trends`), the tendency for temperate forests to have greater fluxes than boreal forests held for $NEP$ in regrowth forests (tropical forests excluded because of insufficient data).
 
-In terms of C stocks, ten variables (all but standing deadwood, $DW_{standing}$) had sufficient data to test for age trends (Table 1, Figs. 7, S16-26).
+In terms of C stocks, ten variables (all but standing deadwood, $DW_{standing}$) had sufficient data to test for age trends (Table 1, Figs. `r stock_age_trends`, `r all_stock_age_trends`).
 All of these displayed a significant overall increase with with $log10[stand.age]$.
 There were sufficient data to model age $\times$ biome interactions were also significant for all ten of these C stock variables (Table S2), with living C stocks tending to accumulate more rapidly during the early stages of forest regrowth in tropical forests (Figs.
-7, S16-S22).
+`r stock_age_trends`, `r all_stock_age_trends`).
 In the case of two non-living C stocks ($DW_{down}$ and $OL$), age $\times$ biome interactions were such that 
-Specifically, $DW_{down}$ declined with age in temperate and boreal forests, compared to an increase with age in tropical forests (Figs. 7, S25).
-Similarly, $OL$ declined slightly with age in temperate broadleaf forests, contrasting an increase in the other three biomes (Figs. 7, S26).
+Specifically, $DW_{down}$ declined with age in temperate and boreal forests, compared to an increase with age in tropical forests (Figs. `r stock_age_trends`, `r DW_down_fig`).
+Similarly, $OL$ declined slightly with age in temperate broadleaf forests, contrasting an increase in the other three biomes (Figs. `r stock_age_trends`, `r OL_fig`).
 
 \newpage
-\blandscape
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | C cycle diagram for young tropical broadleaf forests.
-Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
-Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Tropical broadleaf YOUNG.jpeg"))
-```
-
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 9 | C cycle diagram for young temperate broadleaf forests.
-Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
-Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate broadleaf YOUNG.jpeg"))
-```
-
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 10 | C cycle diagram for young temperate conifer forests.
-Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
-Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate conifer YOUNG.jpeg"))
-```
-
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 11 | C cycle diagram for young boreal conifer forests.
-Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
-Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
-All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.", out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Boreal conifer YOUNG.jpeg"))
-```
-\elandscape
 
@@
 Specifically, the major C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for regrowth as well as mature forests (Figs.
-6-7).
-In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs. 2-5, 7).
-The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Figs. 6-11).
+`r flux_age_trends`- `r stock_age_trends`).
+In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs. 2-5, `r stock_age_trends`).
+The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Table 1; Figs. `r flux_age_trends`- `r stock_age_trends`, `r all_age_trends`).
 Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks and $NEP$ of mature forests are correlated with a different set of factors.
@@
 
-There are of course notable holes in the ForC variable coverage (Fig. 1) that limit the scope of our inferences here. 
-Notably, *ForC* currently has sparse--if any--coverage of fluxes to herbivores and higher consumers, along with the woody mortality ($M_{woody}$) and dead wood stocks (Tables 1, Figs. S23-S25). 
+There are of course notable holes in the ForC variable coverage (Fig. `r figure_map`) that limit the scope of our inferences here. 
+Notably, *ForC* currently has sparse--if any--coverage of fluxes to herbivores and higher consumers, along with the woody mortality ($M_{woody}$) and dead wood stocks (Table 1, Figs.
+`r DW_tot_fig`-`r DW_down_fig`).
 Geographically, all variables are poorly covered in Africa and Siberia, a common problem in the carbon-cycle community [@xu_contribution_2016; @schimel_effect_2015]. 
@@
 In this analysis, the C cycle budgets for mature forests (Figs.
-2-5) generally “close”–-that is, the sums of component variables do not differ from the larger fluxes by more than one standard deviation.
+`r all_diagrams_mature`) generally “close”–-that is, the sums of component variables do not differ from the larger fluxes by more than one standard deviation.
 On the one hand, this reflects the general fact that ecosystem-scale measurements tend to close the C budget more easily and consistently than, for example, for energy balance [@stoy_data-driven_2013].
@@
 Nonetheless, the lack of closure, in the one instance where it occurs, is probably more reflective of differences in the representation of forest types (*i.e.*, disproportionate representation of US Pacific NW for  $B_{root-coarse}$ relative to  $B_{root}$; Fig.
-4) than of methodological accuracy.
+`r diagram_TeN_mature`) than of methodological accuracy.
 The overall high degree of closure implies that *ForC* gives a consistent picture of C cycling within biomes. 
@@
 
-The notable exception to the pattern of fluxes decreasing from tropical to boreal regions is $NEP$ (Fig. 6f), which showed no significant differences across biomes.
+The notable exception to the pattern of fluxes decreasing from tropical to boreal regions is $NEP$ (Fig. `r flux_age_trends`), which showed no significant differences across biomes.
 Unlike the other C flux variables, $NEP$ does not characterize the rate at which C cycles through the ecosystem, but is the balance between C sequestration ($GPP$) and respiratory losses ($R_{eco}$) and represents net CO~2~ sequestration (or release) by the ecosystem.
@@
 In contrast to the patterns observed for $NEP$ in mature stands, $NEP$ of stands between 20 and 100 years of age varied across biomes, being lowest in boreal forests, intermediate in temperate broadleaf forests, and highest in temperate conifer forests (with insufficient data to assess tropical forests; Figs.
-6f, S1).
+`r flux_age_trends`, `r NEP_fig`).
 This is consistent with findings that live biomass accumulation rates (*e.g.*, $\Delta B_{ag}$ or $\Delta B_{tot}$) during early secondary succession decrease with latitude [Figs.
@@
 Note, though, that $NEP$ includes not only $\Delta B_{tot}$, but also changes in $DW_{tot}$, $OL$, and soil carbon, and biome differences in the accumulation rates of these variables have not been detected, in part because these variables do not consistently increase with stand age [Figs.
-7, S23-S26, and see discussion below; @cook-patton_mapping_2020].
+`r stock_age_trends`, `r DW_tot_fig`-`r_OL_fig`, and see discussion below; @cook-patton_mapping_2020].
 
@@
 The observed tendency for young forest fluxes to decrease from tropical to boreal regions paralleled patterns in mature forests (Figs.
-6, S1-S15), suggesting that regrowth forests follow latitudinal trends in carbon cycling similar to those of mature forests [e.g., @banbury_morgan_global_nodate].
+`r flux_age_trends`, `r all_flux_age_trends`), suggesting that regrowth forests follow latitudinal trends in carbon cycling similar to those of mature forests [e.g., @banbury_morgan_global_nodate].
 
@@
 Although these stocks can be important--exceeding 100 Mg C ha^-1^ in some stands (Figs.
-7d-e, S23-S25), this study is the first to synthesize deadwood data on a global scale [but see @cook-patton_mapping_2020 for young forests].
+`r stock_age_trends`, `r DW_tot_fig`-`r DW_down_fig`), this study is the first to synthesize deadwood data on a global scale [but see @cook-patton_mapping_2020 for young forests].
 Unfortunately, data remain too sparse for statistical comparison across biomes (Figs.
-7, S23-S25; but see below for age trends), pointing to a need for more widespread quantification of both standing and downed deadwood.
+`r stock_age_trends`, `r DW_tot_fig`-`r DW_down_fig`; but see below for age trends), pointing to a need for more widespread quantification of both standing and downed deadwood.
 *ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that slower decomposition in colder climates results in more buildup of organic matter [@allen_global_2002].
@@
 Our study reveals that most C fluxes quickly increase to a plateau as stands age (Fig.
-6), consistent with current understanding of age trends in forest C cycling [e.g., @andersonteixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
+`r flux_age_trends`), consistent with current understanding of age trends in forest C cycling [e.g., @andersonteixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
 While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_auto$) would be minimal immediately following a stand-clearing disturbance.
-These would be expected to increase rapidly with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Fig. 7).
+These would be expected to increase rapidly with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Fig. `r stock_age_trends`).
 In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$) are expected to be non-zero following stand-clearing disturbance, although these may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots and litter [@ribeiro-kumara_how_2020; @maurer_carbon_2016; @bond-lamberty_contribution_2004].
@@
 
-Notably, net carbon sequestration ($NEP$) increases with age up to the 100-yr threshold examined here, with more pronounced patterns in temperate than boreal forests (Fig. 6f). 
+Notably, net carbon sequestration ($NEP$) increases with age up to the 100-yr threshold examined here, with more pronounced patterns in temperate than boreal forests (Fig. `r flux_age_trends`). 
 This finding is largely consistent with, but built from a far larger dataset than, previous studies showing an increase in $NEP$ across relatively young stand ages [@pregitzer_carbon_2004; @baldocchi_fluxnet_2001; @luyssaert_old-growth_2008].
-However, $NEP$ has been observed to decline from intermediate to old stands [@luyssaert_old-growth_2008], and the $NEP$ estimated by our model for 100-year-old temperate conifer stands (~5 Mg C ha^-1^ yr^-1^) exceeds the mean of mature forests in the same biome (0.7 Mg C ha^-1^ yr^-1^; Fig.4).
+However, $NEP$ has been observed to decline from intermediate to old stands [@luyssaert_old-growth_2008], and the $NEP$ estimated by our model for 100-year-old temperate conifer stands (~5 Mg C ha^-1^ yr^-1^) exceeds the mean of mature forests in the same biome (0.7 Mg C ha^-1^ yr^-1^; Fig.
+`r diagram_TeN_mature`).
 A decrease in $NEP$ is consistent with the observed deceleration of biomass accumulation as stands age, although both biomass and non-living C stocks will often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands [@luyssaert_old-growth_2008; @mcgarvey_carbon_2014; @lichstein_biomass_2009].
@@
 The latter are particularly sensitive to the type of disturbance, where disturbances that remove most organic material (e.g., logging, agriculture) result in negligible deadwood in young stands, followed by a buildup over time [tropical stands in Fig.
-7e; e.g., @vargas_biomass_2008].
+`r stock_age_trends`; e.g., @vargas_biomass_2008].
 In contrast, natural disturbances (e.g., fire, drought) can produce large amounts of deadwood (mostly $DW_{standing}$) that slowly decomposes as the stand recovers, resulting in declines across young stand ages [e.g., temperate and boreal stands in Fig.
-7e; e.g., @carmona_coarse_2002].
+`r stock_age_trends`; e.g., @carmona_coarse_2002].
 Again, further study and synthesis of non-living C stocks across biomes and stand ages will be valuable to giving a more comprehensive picture.
@@
 Notably, the fact that tropical forests tend to have both the highest rates of C sequestration in young stands [Fig.
-7; @cook-patton_mapping_2020], fueled by their generally high C flux rates (Table 1; Fig.
-6), and the highest mean biomass [Fig.
-7; Table 1; @hu_mapping_2016; @jian_restructured_2020] reinforces the concept that conservation and restoration of these forests is a priority for climate change mitigation, along with high-biomass old-growth temperate stands [@grassi_key_2017; @goldstein_protecting_2020].
+`r stock_age_trends`; @cook-patton_mapping_2020], fueled by their generally high C flux rates (Table 1; Fig.
+`r flux_age_trends`), and the highest mean biomass [Fig.
+`r stock_age_trends`; Table `r variables_table`; @hu_mapping_2016; @jian_restructured_2020] reinforces the concept that conservation and restoration of these forests is a priority for climate change mitigation, along with high-biomass old-growth temperate stands [@grassi_key_2017; @goldstein_protecting_2020].
 It is also important to note the trade-off in climate mitigation potential of restoration of young forests, with high rates of CO~2~ sequestration [$NEP$; @cook-patton_mapping_2020], versus conservation and management of mature forests, with low $NEP$ but high C stocks that could not be recovered on a time scale relevant to climate change mitigation [@goldstein_protecting_2020].

### 2021-01-15 manuscript/ERL_global_C_review.Rmd
a bit of cleanup

@@
 Primary research articles typically cover only a small numbers of sites at a time.
-The rare exceptions that span regions or continents with rare exceptions spanning regions or continents are typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; e.g., @lutz_global_2018], NEON [@schimel_neon_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
+The rare exceptions spanning regions or continents are typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; e.g., @lutz_global_2018], NEON [@schimel_neon_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
 The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*
@@
 Data have been amalgamated from original peer-reviewed publications, either directly or via intermediary data compilations.
-Since the its most recent publication [@anderson-teixeira_forc_2018], *ForC* has grown to include two additional large databases: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020] that also synthesized published forest C data.
+Since its most recent publication [@anderson-teixeira_forc_2018], *ForC* has grown to include two additional large databases: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020] that also synthesized published forest C data.
 Following these additions, *ForC* currently contains `r nrow(ForC_measurements)` records from `r nrow(ForC_plots)` plots and `r n_areas` distinct geographic areas representing all forested biogeographic and climate zones.
@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Tropical broadleaf MATURE.jpeg"))
+  knitr::include_graphics("tables_figures/C_cycle_diagrams/Tropical_broadleaf_MATURE.jpeg")
 ```
@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate broadleaf MATURE.jpeg"))
+  knitr::include_graphics("tables_figures/C_cycle_diagrams/Temperate_broadleaf_MATURE.jpeg")
 ```
@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate conifer MATURE.jpeg"))
+  knitr::include_graphics("tables_figures/C_cycle_diagrams/Temperate_conifer_MATURE.jpeg")
 ```
@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Boreal conifer MATURE.jpeg"))
+  knitr::include_graphics("tables_figures/C_cycle_diagrams/Boreal_conifer_MATURE.jpeg")
 ```
@@
 
-The most notable exception to the pattern of decreasing flux from tropical to boreal biomes was $NEP$, with no significant differences across biomes but with the largest average in temperate broadleaf forests, followed by tropical, boreal, and temperate conifer forests (Figs.
+The most notable exception to the pattern of decreasing flux per unit area from tropical to boreal biomes was $NEP$, with no significant differences across biomes but with the largest average in temperate broadleaf forests, followed by tropical, boreal, and temperate conifer forests (Figs.
 `r flux_age_trends`, `rNEP_fig`).
@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/age_trends/for_ERL_review/Flux_age_trends.jpeg"))
+  knitr::include_graphics("tables_figures/age_trends/for_ERL_review/Flux_age_trends.jpeg")
 ```
@@
 
-  knitr::include_graphics(paste0(path_to_ForC, "figures/age_trends/for_ERL_review/Stock_age_trends.jpeg"))
+  knitr::include_graphics("tables_figures//age_trends/for_ERL_review/Stock_age_trends.jpeg")
 ```

### 2021-01-15 manuscript/ERL_global_C_review.Rmd
rearrange intro

@@
 
+*including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].*
+
+### Forests in the global C cycle: current and future
+
 A robust understanding of forest impacts on global C cycling is essential. 
@@
 
-Despite the importance of forests, comprehensive global studies have historically been limited by the scattered and more local nature of research studies. 
-Primary research articles typically cover only a small numbers of sites at a time.
-The rare exceptions spanning regions or continents are typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; e.g., @lutz_global_2018], NEON [@schimel_neon_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
+### Evolution of forest C cycle research
+
+For more than half a century, researchers have sought to understand how forest carbon cycling varies across stands, including those of different biomes [e.g., Leith, @luyssaert_co2_2007] and stand ages [e.g., @odum_strategy_1969 Odum, Luyssaert 2008].
+- increasingly refined understanding of the elements of ecosystem C cycles,  
+- standardization of definitions, (which were vague at the time of Odum)
+- Expanding set of variables, notably including the development of the eddy-covariance technique to measure ecosystem-atmosphere CO~2~ exchange (REFS), NEP, soil respiration, Also various techniques for measuring belowground C cycling.
+- critical review of methods and their sources of errors
+- Increasing standardization of  methods: allometries, which are highly influential to any measurement including woody biomass or changes therein, but until recently have not been standardized
+- Further standardization made possible through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; e.g., @lutz_global_2018], GEM (Malhi), NEON [@schimel_neon_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
+
+Alongside these conceptual and methodological developments, there has been a proliferation of measurements across the world's forests. 
 The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*
-In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
+In this format, the data are effectively inaccessible for many global-scale analyses, 
+Data synthesis began,...
 
@@
 
+### Biome differences
+Forest C cycling varies enormously across biomes...
+Biome is an artificial categorical distinction convenient for categorizing the world's forests according to major differences in climate and vegetation.
+
+The most important gradient for defining biomes is climate. 
+Since (Humboldt?), it has been recognized that climate plays a dominant role in shaping differences among forests.
+
+@lieth_primary_1973 synthesized differences in NPP across gradients of mean annual temperture and precipitation.
+
+@banbury_morgan_global_nodate
+
+Recently, machine learning draws on global maps in combination with ground-based data to create global maps of C cycling [e.g., REFS; @cook-patton_mapping_2020].
+
+While current data and technology allow increasingly fine-scale resolution for understanding differences across forests, cateogorization into biomes remains useful for applications where data do not yet support robust analysis across finer-scale axes of variation.
+
+### Age trends and their variation across biomes
+In 1969, Eugene Odum published ... [@odum_strategy_1969].
+Although the conceptualization of the C cycle in this paper is simplistic by current standards, the paper was foundational... [@corman_foundations_2019].
+...
+*(Need for a synthetic update.)*
+
+In the past *couple of/several* decades, researchers have started asking how age trends--mostly in biomass accumulation-- vary across biomes.
+*(look up early work on this--citations in @anderson_temperature-dependence_2006)*
+@anderson_temperature-dependence_2006
+@chazdon_carbon_2016
+@cook-patton_mapping_2020
+There has been little synthesis of differences in variables other than C stocks or their accumulation rates.
+
 ```{r schematic, eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 1 | DRAFT SCHEMATIC.", out.width='100%', fig.pos='H'}
@@
 
-Here, we provide a robust and comprehensive analysis of carbon cycling from a stand to global level, and by biome and stand age, using the largest global compilation of forest carbon data, which is available in our open source Global Carbon Forest database (*ForC*; Fig.
+Here, conduct a data-based review of carbon cycling from a stand to global level, and by biome and stand age, using the largest global compilation of forest carbon data, which is available in our open source Global Carbon Forest database (*ForC*; Fig.
 `r figure_map`).
-Our primary goal is to provide a data-driven summary of our current state of knowledge on broad trends in forest C cycling. 
-Specifically, we address three broad questions:
-
-~~1. How thoroughly can we represent C budgets for each of the world's major forest biomes (*i.e.*, tropical, temperate broadleaf and deciduous, boreal) based on the current *ForC* data?~~
-
-2. How do C cycling vary across the world's major forest biomes?
-
-3. How does C cycling vary with stand age (in interaction with biome)?
-
-While researchers have been addressing components of these questions for more than half a century [@odum_strategy_1969; @luyssaert_co2_2007; @anderson-teixeira_carbon_2016; @cook-patton_mapping_2020; @banbury_morgan_global_nodate], our analysis represents by far the most comprehensive analysis of C cycling in global forests, and will serve as a foundation for improved understanding of global forest C cycling and highlight where key sources of uncertainty still reside.
+Our primary goal is to provide a comprehensive synthesis on broad trends in forest C cycling that can serve as a foundation for improved understanding of global forest C cycling and highlight where key sources of uncertainty still reside.
 
@@
 To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.
-Asterisk after variable name indicates lack of C cycle closure.", out.width='100%', fig.pos='H'}
+Asterisk after variable name indicates lack of C cycle closure.", out.width='90%', fig.pos='H'}
 

### 2021-01-15 manuscript/tables_figures/schematic/schematic.m
work on schematic
#61

@@ -30,21 +30,31 @@ biome_names={'tropical', 'temperate', 'boreal'};
 %FLUXES
 % 1. by age
 %for stacked plot
-R_auto = 2* log10(age);
-NPP = R_auto;
+NEP=2*log10(age)-.037*(age-1); %placeholder
+GPP= 4* log10(age)+NEP;
+R_auto = 0.5*GPP;
+NPP = GPP-R_auto;
+R_het=GPP-R_auto;
 
-fluxes = [R_auto; NPP]; %matrix with all fluxes for stacked plot
-flux_names = {'R_{auto}', 'NPP'};
 
-%others
-GPP = sum(fluxes,1);
+%categorize by "in" (GPP components, including NEP) and "out" (Reco components)
+in_fluxes = [R_auto; NPP; NEP]; %matrix with all fluxes for stacked plot
+in_flux_names = {'R_{auto}', 'NPP', 'NEP'};
+out_fluxes = [R_auto; R_het];
+out_flux_names = {'R_{auto}', 'R_{het}'};
+
+%sum in and out fluxes
+in_sum = sum(in_fluxes,1);
+out_sum=sum(out_fluxes,1);
 
 % 2. mature fluxes
 Flux_ratio_temp_trop=0.5;
-Flux_ratio_boreaal_trop=0.25;
+Flux_ratio_boreal_trop=0.25;
 
-%for stacked plot
-mature_fluxes = [fluxes(:,end)'; Flux_ratio_temp_trop*fluxes(:,end)' ; Flux_ratio_boreaal_trop*fluxes(:,end)'];
+%for grouped stacked plot
+mature_fluxes=0*ones(3,2,size(in_fluxes,1)); %empty 3D matrix (Group=biome, Stack=bar(GPP vs Reco), StackElement=component)
+mature_fluxes(:,1,1:size(in_fluxes,1))=[in_fluxes(:,end)'; Flux_ratio_temp_trop*in_fluxes(:,end)' ; Flux_ratio_boreal_trop*in_fluxes(:,end)']; %components of GPP
+mature_fluxes(:,2,1:size(out_fluxes,1))=[out_fluxes(:,end)'; Flux_ratio_temp_trop*out_fluxes(:,end)' ; Flux_ratio_boreal_trop*out_fluxes(:,end)'];
 
 %others
 
[…]

### 2021-01-15 manuscript/ERL_global_C_review.Rmd
small updates

@@
 These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published.
-While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass: @hu_mapping_2016; @spawn_harmonized_2020; gross primary productivity, $GPP$: @li_mapping_2019; @saatchi_benchmark_2011], measurement and validation of most forest C stocks and fluxes require intensive on-the-ground data collection.
-
-*including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].*
+Large-scale synthesis is critical to benchmarking model performance with global data (Clark et al 2017, Luo et al 2012), quantifying the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or using book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
 
@@
 - Further standardization made possible through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; e.g., @lutz_global_2018], GEM (Malhi), NEON [@schimel_neon_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
+- While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass: @hu_mapping_2016; @spawn_harmonized_2020; gross primary productivity, $GPP$: @li_mapping_2019; @saatchi_benchmark_2011], measurement and validation of most forest C stocks and fluxes require intensive on-the-ground data collection.
 
@@
 
-@lieth_primary_1973 synthesized differences in NPP across gradients of mean annual temperture and precipitation.
+@lieth_primary_1973 synthesized differences in NPP across gradients of mean annual temperature and precipitation.
 
@@
 There has been little synthesis of differences in variables other than C stocks or their accumulation rates.
+Given the important role of secondary forests in the current and future global C cycle, concrete understanding of age trends in C fluxes and stocks and how these vary across biomes is critical to better understandign of the global C cycle.
 
-```{r schematic, eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 1 | DRAFT SCHEMATIC.", out.width='100%', fig.pos='H'}
+```{r schematic, eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 1 | DRAFT Schematic diagram summarizing current understanding of biome differences and age trends in forest C cycling.
+Plotted values are realistic with respect to the results of this synthesis, but schematic is not intended to be quantitatively precise.
+", out.width='100%', fig.pos='H'}
 
@@
 The most notable exception to the pattern of decreasing flux per unit area from tropical to boreal biomes was $NEP$, with no significant differences across biomes but with the largest average in temperate broadleaf forests, followed by tropical, boreal, and temperate conifer forests (Figs.
-`r flux_age_trends`, `rNEP_fig`).
+`r flux_age_trends`, `r NEP_fig`).
 
@@
 Note, though, that $NEP$ includes not only $\Delta B_{tot}$, but also changes in $DW_{tot}$, $OL$, and soil carbon, and biome differences in the accumulation rates of these variables have not been detected, in part because these variables do not consistently increase with stand age [Figs.
-`r stock_age_trends`, `r DW_tot_fig`-`r_OL_fig`, and see discussion below; @cook-patton_mapping_2020].
+`r stock_age_trends`, `r DW_tot_fig`-`r OL_fig`, and see discussion below; @cook-patton_mapping_2020].
 
@@
 `r stock_age_trends`, `r DW_tot_fig`-`r DW_down_fig`; but see below for age trends), pointing to a need for more widespread quantification of both standing and downed deadwood.
-*ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that slower decomposition in colder climates results in more buildup of organic matter [@allen_global_2002].
+*ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that proportionally slower decomposition in colder climates results in more buildup of organic matter [@allen_global_2002; @andersonteixeira_differential_2011].
 Further research on non-living C stocks in the world's forests will be essential to completing the picture.
@@
 `r flux_age_trends`), consistent with current understanding of age trends in forest C cycling [e.g., @andersonteixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
-While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_auto$) would be minimal immediately following a stand-clearing disturbance.
+While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_{auto}$) would be minimal immediately following a stand-clearing disturbance.
 These would be expected to increase rapidly with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Fig. `r stock_age_trends`).
@@
 *ForC*’s tens of thousands of records are readily available in a standardized format, and our analyses here indicate that their internal consistency is reasonably high.
-Integration of *ForC* with models will be valuable to improving the accuracy and reliability of models.
+Integration of *ForC* with models will be valuable to improving the accuracy and reliability of models [@fer_beyond_2021].
 

### 2021-01-15 manuscript/tables_figures/schematic/schematic.m
worked on schematic

@@ -30,16 +30,20 @@ biome_names={'tropical', 'temperate', 'boreal'};
 %FLUXES
 % 1. by age
 %for stacked plot
-NEP=2*log10(age)-.037*(age-1); %placeholder
-GPP= 4* log10(age)+NEP;
-R_auto = 0.5*GPP;
-NPP = GPP-R_auto;
-R_het=GPP-R_auto;
+NEP=2*log10(age)-.037*(age-1); %placeholder (peaks around 2 at 23 yrs age)
+GPP= 11.15 + (1.99+1.44)* log10(age)+NEP;
+ANPP_foliage = 1.66* log10(age) +.18;
+ANPP_woody = 2.31* log10(age) -.76;
+ANPP_woody_turnover=ANPP_woody-NEP;
+BNPP=0.97*log10(age)+.67;
+NPP=ANPP_foliage+ANPP_woody+BNPP;
+R_auto = NPP.*(1+.001*age); %insufficient data
+R_het=ANPP_foliage+ANPP_woody_turnover+BNPP;
 
 
 %categorize by "in" (GPP components, including NEP) and "out" (Reco components)
-in_fluxes = [R_auto; NPP; NEP]; %matrix with all fluxes for stacked plot
-in_flux_names = {'R_{auto}', 'NPP', 'NEP'};
+in_fluxes = [R_auto; BNPP; ANPP_foliage;  ANPP_woody_turnover; NEP]; %matrix with all fluxes for stacked plot
+in_flux_names = {'R_{auto}', 'BNPP', 'ANPP_{foliage}', 'ANPP_{woody.turnover}',  'NEP'};
 out_fluxes = [R_auto; R_het];
 out_flux_names = {'R_{auto}', 'R_{het}'};
 

### 2021-01-15 manuscript/tables_figures/schematic/schematic.m
minor changes

@@ -31,10 +31,10 @@ biome_names={'tropical', 'temperate', 'boreal'};
 % 1. by age
 %for stacked plot
 NEP=2*log10(age)-.037*(age-1); %placeholder (peaks around 2 at 23 yrs age)
-GPP= 11.15 + (1.99+1.44)* log10(age)+NEP;
+GPP= NEP + (1.99+1.44)* log10(age);
 ANPP_foliage = 1.66* log10(age) +.18;
 ANPP_woody = 2.31* log10(age) -.76;
-ANPP_woody_turnover=ANPP_woody-NEP;
+ANPP_woody_turnover=max(0,ANPP_woody-NEP);
 BNPP=0.97*log10(age)+.67;
 NPP=ANPP_foliage+ANPP_woody+BNPP;
 R_auto = NPP.*(1+.001*age); %insufficient data
@@ -71,15 +71,14 @@ DW= 2.6*log10(age);
 stocks = [biomass; DW]; %matrix with all stocks for stacked plot
 stock_names = {'B_{tot}', 'DW'};
 
-%others
 
 
 
 % 2. mature forests
 Stock_ratio_temp_trop=0.9;
-Stock_ratio_boreaal_trop=0.7;
+Stock_ratio_boreal_trop=0.7;
 
-mature_stocks = [stocks(:,end)'; Stock_ratio_temp_trop*stocks(:,end)' ; Stock_ratio_boreaal_trop*stocks(:,end)'];
+mature_stocks = [stocks(:,end)'; Stock_ratio_temp_trop*stocks(:,end)' ; Stock_ratio_boreal_trop*stocks(:,end)'];
 
 
 %% ~~~~~~~PLOTTING~~~~~~~~~~

### 2021-01-15 manuscript/tables_figures/schematic/stacked_area_results.m
Create stacked_area_results.m
start at stacked area plots for results visualization. Not sure if it will work well...

#58

@@ -1,0 +1,101 @@
+% age trend summaries for ERL review
+% Kristina Anderson Teixeira
+% January 2021
+
+clf; clear; close all;
+set(0,'DefaultLegendAutoUpdate','off');
+
+%% ~~~~~~~READ IN DATA~~~~~~~~~~
+
+age = linspace (1,100,100);
+biome_names={'tropical', 'temperate', 'boreal'};
+
+%FLUXES
+% 1. by age
+%for stacked plot
+NEP=4*log10(age)-.08*(age-1); %placeholder (peaks around 2 at 23 yrs age)
+GPP= 11.15 + (1.99+1.44)* log10(age);
+ANPP_foliage = 1.66* log10(age) +.18;
+ANPP_woody = 2.31* log10(age) -.76;
+ANPP_woody_turnover=max(0,ANPP_woody-NEP);
+BNPP=0.97*log10(age)+.67;
+NPP=ANPP_foliage+ANPP_woody+BNPP;
+R_auto = NPP.*(1+.001*age); %insufficient data
+R_het=ANPP_foliage+ANPP_woody_turnover+BNPP;
+
+
+%categorize by "in" (GPP components, including NEP) and "out" (Reco components)
+in_fluxes = [R_auto; BNPP; ANPP_foliage;  ANPP_woody_turnover; NEP]; %matrix with all fluxes for stacked plot
+in_flux_names = {'R_{auto}', 'BNPP', 'ANPP_{foliage}', 'ANPP_{woody.turnover}',  'NEP'};
+out_fluxes = [R_auto; R_het];
+out_flux_names = {'R_{auto}', 'R_{het}'};
+
+%sum in and out fluxes
+in_sum = sum(in_fluxes,1);
+out_sum=sum(out_fluxes,1);
+
+% 2. mature fluxes
+Flux_ratio_temp_trop=0.5;
+Flux_ratio_boreal_trop=0.25;
+
+%for grouped stacked plot
+mature_fluxes=0*ones(3,2,size(in_fluxes,1)); %empty 3D matrix (Group=biome, Stack=bar(GPP vs Reco), StackElement=component)
+mature_fluxes(:,1,1:size(in_fluxes,1))=[in_fluxes(:,end)'; Flux_ratio_temp_trop*in_fluxes(:,end)' ; Flux_ratio_boreal_trop*in_fluxes(:,end)']; %components of GPP
[…]

### 2021-01-15 manuscript/tables_figures/schematic/stacked_area_results.m
Update stacked_area_results.m
#58

@@ -67,21 +67,21 @@ for b =3:3
 
 figure (b)
  
-
 subplot (2,4,1:3)
 %flux plot:
 h=area (age, in_fluxes'); %, 'LineStyle','-'); 
 hold on;
+h=area (age, -1*out_fluxes'); %, 'LineStyle','-'); 
+hold on;
 plot(age, in_sum, '-b', 'LineWidth', 3); hold on;
 plot(age, GPP, '--b', 'LineWidth', 3); hold on;
-%plot(age, out_sum, '-r', 'LineWidth', 3);
+plot(age, NEP, '-w', 'LineWidth', 3);
 t = title(biomes(b));
+ylabel ('C  stocks (Mg C ha^{-1})')
 
 subplot (2,4,4) %mature C fluxes
-bar(mature_stocks, 'stacked');
+bar(mature_stocks(b,:), 'stacked');
 legend (stock_names);
-xlabel ('biome');
-ylabel ('C  stocks (Mg C ha^{-1})')
 
 
 subplot (2,4,5:7)
@@ -90,12 +90,12 @@ h=area (age, stocks'); %, 'LineStyle','-');
 hold on;
 plot(age, biomass, '-k', 'LineWidth', 3);
 xlabel ('stand age');
-
+ylabel ('C  stocks (Mg C ha^{-1})')
 
 subplot (2,4,8) %mature C stocks
 bar(mature_stocks, 'stacked');
 legend (stock_names);
-xlabel ('biome');
-ylabel ('C  stocks (Mg C ha^{-1})')
+xlabel ('mature stands');
+
 
 end

### 2021-01-16 manuscript/ERL_global_C_SI.Rmd
cropped figures & adjusted size for fit on one page
Update ERL_global_C_SI.Rmd

@@
 #C cycle diagrams
-#do.call(file.remove, list(list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/tables_figures/C_cycle_diagrams/", full.names = TRUE))) # first remove old files
+#do.call(file.remove, list(list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/manuscript/tables_figures/C_cycle_diagrams/", full.names = TRUE))) # first remove old files
 
 #flist <- list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/", full.names = TRUE)
-#file.copy(flist, "/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/tables_figures/C_cycle_diagrams/")
+#file.copy(flist, "/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/manuscript/tables_figures/C_cycle_diagrams/")
 
 #Age diagrams
-#do.call(file.remove, list(list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/tables_figures/age_trends/", full.names = TRUE))) # first remove old files
+#do.call(file.remove, list(list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/manuscript/tables_figures/age_trends/", full.names = TRUE))) # first remove old files
 
 #flist <- list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForC-db/ForC/figures/age_trends/", full.names = TRUE)
-#file.copy(flist, "/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/tables_figures/age_trends/")
+#file.copy(flist, "/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/manuscript/tables_figures/age_trends/")
 
@@
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.", out.width='90%', fig.pos='H'}
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.", out.width='87%', fig.pos='H'}
 

### 2021-01-17 manuscript/tables_figures/schematic/stacked_area_results.m
fill in parameters matrix

@@ -13,7 +13,12 @@ working_dir='/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/ma
 %read in data:
 cd(figures_dir)
 model_summaries=readtable('SI_age_trend_model_summaries.csv');
-%probably need to remove specieal formatting from model_summaries.Variable
+%remove specieal formatting from model_summaries.Variable:
+for n=1:size(model_summaries,1)
+    s=cell2mat(model_summaries.Variable(n));
+    s(regexp(s,'[$,{,}]'))=[];
+    model_summaries.Variable(n)=cellstr(s);
+end
 
 %prepare some variables
 variables = unique(model_summaries.Variable);
@@ -23,7 +28,61 @@ params_matrix=cell2table(num2cell(NaN*ones(length(variables), 10)),'VariableName
 params_matrix.variable=variables;
 
 for v=1:length(variables) %cycle through variables and fill in params_matrix
-    
+    var=variables(v);
+    %beta
+    if max(strcmp(model_summaries.Variable, var) + strcmp(model_summaries.Parameter, 'log10(stand.age)'))==2;
+        params_matrix.beta(v)=table2array(model_summaries(strcmp(model_summaries.Variable, var) & strcmp(model_summaries.Parameter, 'log10(stand.age)'),3));
+    else
+        params_matrix.beta(v)=NaN;
+    end
+    %betaTrB
+    if max(strcmp(model_summaries.Variable, var) + strcmp(model_summaries.Parameter, 'log10(stand.age):BiomeTropical broadleaf'))==2;
+        params_matrix.betaTrB(v)=table2array(model_summaries(strcmp(model_summaries.Variable, var) & strcmp(model_summaries.Parameter, 'log10(stand.age):BiomeTropical broadleaf'),3));
+    else
+        params_matrix.betaTrB(v)=NaN;
+    end
+    %betaTeB
+    if max(strcmp(model_summaries.Variable, var) + strcmp(model_summaries.Parameter, 'log10(stand.age):BiomeTemperate broadleaf'))==2;
+        params_matrix.betaTeB(v)=table2array(model_summaries(strcmp(model_summaries.Variable, var) & strcmp(model_summaries.Parameter, 'log10(stand.age):BiomeTemperate broadleaf'),3));
+    else
+        params_matrix.betaTeB(v)=NaN;
+    end
+    %betaTeN
[…]

### 2021-01-17 manuscript/tables_figures/schematic/stacked_area_results.m
plot real results

@@ -39,73 +39,92 @@ for v=1:length(variables) %cycle through variables and fill in params_matrix
     if max(strcmp(model_summaries.Variable, var) + strcmp(model_summaries.Parameter, 'log10(stand.age):BiomeTropical broadleaf'))==2;
         params_matrix.betaTrB(v)=table2array(model_summaries(strcmp(model_summaries.Variable, var) & strcmp(model_summaries.Parameter, 'log10(stand.age):BiomeTropical broadleaf'),3));
     else
-        params_matrix.betaTrB(v)=NaN;
+        params_matrix.betaTrB(v)=0;
     end
     %betaTeB
     if max(strcmp(model_summaries.Variable, var) + strcmp(model_summaries.Parameter, 'log10(stand.age):BiomeTemperate broadleaf'))==2;
         params_matrix.betaTeB(v)=table2array(model_summaries(strcmp(model_summaries.Variable, var) & strcmp(model_summaries.Parameter, 'log10(stand.age):BiomeTemperate broadleaf'),3));
     else
-        params_matrix.betaTeB(v)=NaN;
+        params_matrix.betaTeB(v)=0;
     end
     %betaTeN
     if max(strcmp(model_summaries.Variable, var) + strcmp(model_summaries.Parameter, 'log10(stand.age):BiomeTemperate conifer'))==2;
         params_matrix.betaTeN(v)=table2array(model_summaries(strcmp(model_summaries.Variable, var) & strcmp(model_summaries.Parameter, 'log10(stand.age):BiomeTemperate conifer'),3));
     else
-        params_matrix.betaTeN(v)=NaN;
+        params_matrix.betaTeN(v)=0;
     end
     %betaBoN
     if max(strcmp(model_summaries.Variable, var) + strcmp(model_summaries.Parameter, 'log10(stand.age):BiomeBoreal conifer'))==2;
         params_matrix.betaBoN(v)=table2array(model_summaries(strcmp(model_summaries.Variable, var) & strcmp(model_summaries.Parameter, 'log10(stand.age):BiomeBoreal conifer'),3));
     else
-        params_matrix.betaBoN(v)=NaN;
[…]

### 2021-01-17 manuscript/tables_figures/schematic/stacked_area_results.m
Update stacked_area_results.m

@@ -165,6 +165,12 @@ B_tot=max(0,(params_matrix.beta(B_tot_index)+beta_matrix(B_tot_index,b))* log10(
 B_ag_index=find(strcmp(params_matrix.variable,'B_ag'));
 B_ag=max(0,(params_matrix.beta(B_ag_index)+beta_matrix(B_ag_index,b))* log10(age)+ int_matrix(B_ag_index,b));
 
+B_root_index=find(strcmp(params_matrix.variable,'B_root'));
+B_root=max(0,(params_matrix.beta(B_root_index)+beta_matrix(B_root_index,b))* log10(age)+ int_matrix(B_root_index,b));
+
+B_foliage_index=find(strcmp(params_matrix.variable,'B_foliage'));
+B_foliage=max(0,(params_matrix.beta(B_foliage_index)+beta_matrix(B_foliage_index,b))* log10(age)+ int_matrix(B_foliage_index,b));
+
 DW_tot_index=find(strcmp(params_matrix.variable,'DW_tot'));
 DW_tot=max(0,(params_matrix.beta(DW_tot_index)+beta_matrix(DW_tot_index,b))* log10(age)+ int_matrix(DW_tot_index,b));
 
@@ -179,8 +185,8 @@ OL=max(0,(params_matrix.beta(OL_index)+beta_matrix(OL_index,b))* log10(age)+ int
 DW_standing=max(0,DW_tot-DW_down);
 
 % 2.2 group for plotting
-stock_names = {'B_{tot}', 'DW_{standing}', 'DW_{down}', 'OL'};
-stocks = [B_tot; DW_standing; DW_down; OL]; %matrix with all stocks for stacked plot
+stock_names = {'B_{ag}', 'B_{root}', 'DW_{standing}', 'DW_{down}', 'OL'};
+stocks = [B_ag; B_root ; DW_standing; DW_down; OL]; %matrix with all stocks for stacked plot
 
 
 % 2. mature forests

### 2021-01-17 manuscript/ERL_global_C_review.Rmd
minor changes
change dir name

@@
 
-  knitr::include_graphics("tables_figures/schematic/schematic.png")
+  knitr::include_graphics("tables_figures/stacked_age_trends/schematic.png")
 ```

### 2021-01-17 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m

@@ -7,7 +7,7 @@ set(0,'DefaultLegendAutoUpdate','off');
 
 %% SETTINGS
 figures_dir='/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/manuscript/tables_figures';
-working_dir='/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/manuscript/tables_figures/schematic';
+working_dir='/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/manuscript/tables_figures/stacked_age_trends';
 
 %% READ IN & PREPARE DATA
 %read in data:

### 2021-01-18 manuscript/ERL_global_C_review.Rmd
update
incorporated @bpbond 's feedback on intro, fixed missing references

@@
 Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide [CO~2~; @bonan_forests_2008; @friedlingstein_global_2019; @ipcc_global_2018].
-Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climatecarbon_2006; @krause_large_2018; @bonan_model_2019; @di_vittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history.
-These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published.
-Large-scale synthesis is critical to benchmarking model performance with global data (Clark et al 2017, Luo et al 2012), quantifying the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or using book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
+Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climatecarbon_2006; @krause_large_2018; @bonan_model_2019; @di_vittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of understanding on how C cycling varies across forest types and in relation to stand history.
+This requires accessible, comprehensive, and large-scale databases with global coverage, which runs contrary to the traditional way forest C stocks and fluxes have been measured and published.
+Large-scale synthesis is critical to benchmarking model performance with global data (Clark et al 2017, Luo et al 2012), quantifying the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], and using book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
 
@@
 In recent years, the remaining CO~2~ sink averaged 3.2 $\pm$ 0.6 GtC yr^-1^ from 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
-Yet, this sink is reduced by deforestation, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reducing the net forest sink to ~1.1-2.2 Gt C yr^-1^ across Earth's forests [@friedlingstein_global_2019].
+However, deforestation, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reduces the net forest sink to ~1.1-2.2 Gt C yr^-1^ across Earth's forests [@friedlingstein_global_2019].
 
@@
 
-For more than half a century, researchers have sought to understand how forest carbon cycling varies across stands, including those of different biomes [e.g., Leith, @luyssaert_co2_2007] and stand ages [e.g., @odum_strategy_1969 Odum, Luyssaert 2008].
-- increasingly refined understanding of the elements of ecosystem C cycles,  
-- standardization of definitions, (which were vague at the time of Odum)
-- Expanding set of variables, notably including the development of the eddy-covariance technique to measure ecosystem-atmosphere CO~2~ exchange (REFS), NEP, soil respiration, Also various techniques for measuring belowground C cycling.
-- critical review of methods and their sources of errors
-- Increasing standardization of  methods: allometries, which are highly influential to any measurement including woody biomass or changes therein, but until recently have not been standardized
-- Further standardization made possible through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; e.g., @lutz_global_2018], GEM (Malhi), NEON [@schimel_neon_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
-- While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass: @hu_mapping_2016; @spawn_harmonized_2020; gross primary productivity, $GPP$: @li_mapping_2019; @saatchi_benchmark_2011], measurement and validation of most forest C stocks and fluxes require intensive on-the-ground data collection.
+For more than half a century, researchers have sought to understand how forest carbon cycling varies across stands, including those of different biomes [e.g., @lieth_primary_1973; @luyssaert_co2_2007] and stand ages [e.g., @odum_strategy_1969; @luyssaert_old-growth_2008].
+Over this time, an increasingly refined conceptual understanding of the elements of ecosystem C cycles has developed, as a growing number of variables have been defined [e.g., @chapin_reconciling_2006], along with appropriate measurement methods [e.g., @clark_measuring_2001].
+New technology has also enabled researches to directly measure an expanding set of variables, notably including the development of continuous measurements of soil CO~2~ efflux (**REF-BEN**) and ecosystem-atmosphere CO~2~ exchange (**REFS**).
+Standardization of estimation techniques has also increased; for example, there has been increasing standardization of the biomass allometries strongly influence estimates of any C cycle variable including woody biomass or changes therein [e.g., @chojnacky_updated_2014; @chave_improved_2014].
+Further standardization has been made possible through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; @davies_forestgeo_2021], NEON [@schimel_neon_2007], and FLUXNET [@baldocchi_fluxnet_2001; @novick_ameriflux_2018].
+Remote sensing technology has become increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass: @hu_mapping_2016; @spawn_harmonized_2020; gross primary productivity, $GPP$: @li_mapping_2019; @saatchi_benchmark_2011], yet measurement and validation of most forest C stocks and fluxes necessarily requires intensive on-the-ground data collection.
 
 Alongside these conceptual and methodological developments, there has been a proliferation of measurements across the world's forests. 
-The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*
-In this format, the data are effectively inaccessible for many global-scale analyses, 
-Data synthesis began,...
-
-To address the need for global-scale analyses of forest C cycling, we recently developed *ForC* [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
-*ForC* contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables), with the different variables capturing the unique ecosystem pools (e.g., woody, foliage, and root biomass; dead wood) and flux types (e.g, gross and net primary productivity; soil, root, and ecosystem respiration).
+The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles, along with variation in data formats, units, measurement methods, *etc.*
+To use these data to address global-scale questions, researchers began synthesizing data into increasingly large databases [e.g., @lieth_primary_1973; @luyssaert_co2_2007;  @bond-lamberty_global_2010; @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018; @cook-patton_mapping_2020].
+The current largest, most comprehensive database on forest C cycling is *ForC* [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], which contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables), with the different variables capturing the unique ecosystem pools (e.g., woody, foliage, and root biomass; dead wood) and flux types (e.g, gross and net primary productivity; soil, root, and ecosystem respiration).
 These data are ground-based measurements, and *ForC* contains associated data required for interpretation (*e.g.*, stand history, measurement methods).
-Data have been amalgamated from original peer-reviewed publications, either directly or via intermediary data compilations.
-Since its most recent publication [@anderson-teixeira_forc_2018], *ForC* has grown to include two additional large databases: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020] that also synthesized published forest C data.
-Following these additions, *ForC* currently contains `r nrow(ForC_measurements)` records from `r nrow(ForC_plots)` plots and `r n_areas` distinct geographic areas representing all forested biogeographic and climate zones.
-This represents an `r round((nrow(ForC_measurements) - 17367)/17367*100, digits=0)`% increase in records from the prior publication [@anderson-teixeira_forc_2018].
+Since its most recent publication [@anderson-teixeira_forc_2018], *ForC* has grown `r round((nrow(ForC_measurements) - 17367)/17367*100, digits=0)`% through the incorporation of two additional large databases that also synthesized published forest C data: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010; @jian_restructured_2020] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020].
+Following these additions, *ForC* currently contains `r nrow(ForC_measurements)` records from `r nrow(ForC_plots)` plots and `r n_areas` distinct geographic areas representing all forested biogeographic and climate zones, making it ideal for assessing how forest C cycling varies across biomes and with respect to stand age.
 

### 2021-01-18 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m
refining stacked age trend figures

@@ -99,7 +99,7 @@ age = linspace (1,100,100);
 NEP_index=find(strcmp(params_matrix.variable,'NEP'));
 NEP=(params_matrix.beta(NEP_index)+beta_matrix(NEP_index,b))* log10(age)+ int_matrix(NEP_index,b);
 
-GPP_index=find(strcmp(params_matrix.variable,'GPP'));
+GPP_index=find(strcmp(params_matrix.variable,'GPP')); %not available for tropical forests
 GPP=max(0,(params_matrix.beta(GPP_index)+beta_matrix(GPP_index,b))* log10(age)+ int_matrix(GPP_index,b));
 
 ANPP_foliage_index=find(strcmp(params_matrix.variable,'ANPP_foliage'));
@@ -110,6 +110,10 @@ ANPP_woody=max(0,(params_matrix.beta(ANPP_woody_index)+beta_matrix(ANPP_woody_in
 
 BNPP_index=find(strcmp(params_matrix.variable,'BNPP'));
 BNPP=max(0,(params_matrix.beta(BNPP_index)+beta_matrix(BNPP_index,b))* log10(age)+ int_matrix(BNPP_index,b));
+
+BNPP_fine_index=find(strcmp(params_matrix.variable,'BNPP_fine'));
+BNPP_fine=max(0,(params_matrix.beta(BNPP_fine_index)+beta_matrix(BNPP_fine_index,b))* log10(age)+ int_matrix(BNPP_fine_index,b));
+
 
 NPP_index=find(strcmp(params_matrix.variable,'NPP'));
 NPP=max(0,(params_matrix.beta(NPP_index)+beta_matrix(NPP_index,b))* log10(age)+ int_matrix(NPP_index,b));
@@ -126,17 +130,18 @@ NEP_calc=GPP-R_eco;
 NPP_calc=ANPP_foliage+ANPP_woody+BNPP;
 R_auto_calc = 0.5*GPP; % this is the one that's used
 R_auto_calc1 = GPP.*(0.5+.001*age); %insufficient data
-R_auto_calc2 = NPP_calc.*(1+.001*age); %insufficient data
+R_auto_calc = NPP_calc.*(1+.001*age); %insufficient data
 R_auto_calc3 = R_eco-R_het_soil; %insufficient data
+BNPP_coarse_calc=max(0,BNPP-BNPP_fine);
 
 % 1.2 group for plotting
 % "in" (GPP components) 
[…]

### 2021-01-18 manuscript/ERL_global_C_review.Rmd
added stacked age trends figures to manuscript
#58

@@
 # DISPLAY ITEMS
+#tables
+variables_table=1
+
 #figures
@@
 stock_age_trends=8
+TrB_age_trends=9
 
-#tables
-variables_table=1
+age_trends_caption_generic= "Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
+Asterisks indicate variables whose age trends were calculated based on other variables, as follows: $R_{auto}=NPP $ ,$B_{ag-wood}=B_{ag}-B_{foliage}$, $B_{root-coarse}=B_{root}-B_{root-fine}$."
+TrB_age_trends_caption = paste("Figure 9 | Age trends in C cycling in tropical broadleaf forests.", age_trends_caption_generic) 
+TeB_age_trends_caption = paste("Figure 10 | Age trends in C cycling in temperate broadleaf forests.", age_trends_caption_generic) 
+TeN_age_trends_caption = paste("Figure 11 | Age trends in C cycling in temperate conifer forests.", age_trends_caption_generic) 
+BoN_age_trends_caption = paste("Figure 12 | Age trends in C cycling in boreal conifer forests.", age_trends_caption_generic)
 
@@
 ### Biome differences
-Forest C cycling varies enormously across biomes...
-Biome is an artificial categorical distinction convenient for categorizing the world's forests according to major differences in climate and vegetation.
 
-The most important gradient for defining biomes is climate. 
-Since (Humboldt?), it has been recognized that climate plays a dominant role in shaping differences among forests.
-
-@lieth_primary_1973 synthesized differences in NPP across gradients of mean annual temperature and precipitation.
+Forest C cycling varies enormously across biomes, which are artificial categorical distinctions convenient for categorizing the world's forests according to major differences in climate, vegetation, *etc.*
+Since *(Humboldt?)*, it has been recognized that climate plays a dominant role in shaping differences among forests.
+Early quantitative syntheses revealed that C fluxes such as net primary productivity ($NPP$) decrease dramatically with latitude, or increase across gradients of mean annual temperature and, to a lesser extent, precipitation [@lieth_primary_1973].
 
@@
 
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap =TrB_age_trends_caption, out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/stacked_age_trends/Tropical broadleaf_age_trends.png")
+```
+
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap =TeB_age_trends_caption, out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/stacked_age_trends/Temperate broadleaf_age_trends.png")
+```
+
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap =TeN_age_trends_caption, out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/stacked_age_trends/Temperate conifer_age_trends.png")
+```
+
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap =BoN_age_trends_caption, out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/stacked_age_trends/Boreal conifer_age_trends.png")
+```
+
 \newpage

### 2021-01-18 manuscript/ERL_global_C_review.Rmd
Update ERL_global_C_review.Rmd

@@
 TrB_age_trends=9
+TeB_age_trends=10
+TeN_age_trends=11
+BoN_age_trends=12
 

### 2021-01-18 manuscript/ERL_global_C_review.Rmd
working on fluxes
#64

@@
 age_trends_caption_generic= "Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
-Asterisks indicate variables whose age trends were calculated based on other variables, as follows: $R_{auto}=NPP $ ,$B_{ag-wood}=B_{ag}-B_{foliage}$, $B_{root-coarse}=B_{root}-B_{root-fine}$."
+Asterisks indicate variables whose age trends were calculated based on other variables, as follows: $R_{auto}=NPP(1/CUE-1)$, where $CUE$ is the age trend from DeLucia et al.
+(2007); $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
 TrB_age_trends_caption = paste("Figure 9 | Age trends in C cycling in tropical broadleaf forests.", age_trends_caption_generic) 

### 2021-01-18 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m

@@ -121,17 +121,28 @@ NPP=max(0,(params_matrix.beta(NPP_index)+beta_matrix(NPP_index,b))* log10(age)+ 
 R_eco_index=find(strcmp(params_matrix.variable,'R_eco'));
 R_eco=max(0,(params_matrix.beta(R_eco_index)+beta_matrix(R_eco_index,b))* log10(age)+ int_matrix(R_eco_index,b));
 
+
+R_soil_index=find(strcmp(params_matrix.variable,'R_soil'));
+R_soil=max(0,(params_matrix.beta(R_soil_index)+beta_matrix(R_soil_index,b))* log10(age)+ int_matrix(R_soil_index,b));
+
 R_het_soil_index=find(strcmp(params_matrix.variable,'R_het-soil'));
 R_het_soil=max(0,(params_matrix.beta(R_het_soil_index)+beta_matrix(R_het_soil_index,b))* log10(age)+ int_matrix(R_het_soil_index,b));
+
+R_root_index=find(strcmp(params_matrix.variable,'R_root'));
+R_root=max(0,(params_matrix.beta(R_root_index)+beta_matrix(R_root_index,b))* log10(age)+ int_matrix(R_root_index,b));
 
 
 % 1.1.2 calculated fluxes
 NEP_calc=GPP-R_eco;
 NPP_calc=ANPP_foliage+ANPP_woody+BNPP;
-R_auto_calc = 0.5*GPP; % this is the one that's used
-R_auto_calc1 = GPP.*(0.5+.001*age); %insufficient data
-R_auto_calc = NPP_calc.*(1+.001*age); %insufficient data
-R_auto_calc3 = R_eco-R_het_soil; %insufficient data
+%options for R_auto_calc (insufficient data for regression or to calculate as sum/difference). 
+%Last one in list is the one that will be used.
+CUE=0.679-0.153*log10(age); %from DeLucia et al. 2007 (note that equation given mistakenly leaves of the log10(age))
+R_auto_calc = CUE.*GPP; 
+R_auto_calc= NPP_calc;
+R_auto_calc = NPP_calc.*(1./CUE-1); %Rauto=NPP*(1/CUE-1), and CUE equation is above
+
+
 BNPP_coarse_calc=max(0,BNPP-BNPP_fine);
 
 % 1.2 group for plotting
[…]

### 2021-01-18 manuscript/ERL_global_C_review.Rmd
worked on equation selection
I think I've optimized which sets of variables/ equations to plot. #64

--- 2021-01-18 manuscript/ERL_global_C_review.Rmd
@@
 
+TrB_age_trends_caption = "Figure 9 | Age trends in C cycling in tropical broadleaf forests.
+Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
+Asterisks indicate variables whose age trends were calculated based on other variables, as follows: $R_{auto}=NPP(1/CUE-1)$, where $CUE$ is TEMPORARILY 0.5; $ANPP_{woody}=max(0,ANPP-ANPP_{foliage})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
 age_trends_caption_generic= "Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
-Asterisks indicate variables whose age trends were calculated based on other variables, as follows: $R_{auto}=NPP $ ,$B_{ag-wood}=B_{ag}-B_{foliage}$, $B_{root-coarse}=B_{root}-B_{root-fine}$."
-TrB_age_trends_caption = paste("Figure 9 | Age trends in C cycling in tropical broadleaf forests.", age_trends_caption_generic) 
+Asterisks indicate variables whose age trends were calculated based on other variables, as follows: $R_{auto-ag}=R_{eco}-R{soil}$; ANPP_{woody}=min(ANPP_{stem},ANPP_{woody}); $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
 TeB_age_trends_caption = paste("Figure 10 | Age trends in C cycling in temperate broadleaf forests.", age_trends_caption_generic) 

### 2021-01-18 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m

@@ -102,8 +102,14 @@ NEP=(params_matrix.beta(NEP_index)+beta_matrix(NEP_index,b))* log10(age)+ int_ma
 GPP_index=find(strcmp(params_matrix.variable,'GPP')); %not available for tropical forests
 GPP=max(0,(params_matrix.beta(GPP_index)+beta_matrix(GPP_index,b))* log10(age)+ int_matrix(GPP_index,b));
 
+ANPP_index=find(strcmp(params_matrix.variable,'ANPP'));
+ANPP=max(0,(params_matrix.beta(ANPP_index)+beta_matrix(ANPP_index,b))* log10(age)+ int_matrix(ANPP_index,b));
+
 ANPP_foliage_index=find(strcmp(params_matrix.variable,'ANPP_foliage'));
 ANPP_foliage=max(0,(params_matrix.beta(ANPP_foliage_index)+beta_matrix(ANPP_foliage_index,b))* log10(age)+ int_matrix(ANPP_foliage_index,b));
+
+ANPP_stem_index=find(strcmp(params_matrix.variable,'ANPP_stem'));
+ANPP_stem=max(0,(params_matrix.beta(ANPP_stem_index)+beta_matrix(ANPP_stem_index,b))* log10(age)+ int_matrix(ANPP_stem_index,b));
 
 ANPP_woody_index=find(strcmp(params_matrix.variable,'ANPP_woody'));
 ANPP_woody=max(0,(params_matrix.beta(ANPP_woody_index)+beta_matrix(ANPP_woody_index,b))* log10(age)+ int_matrix(ANPP_woody_index,b));
@@ -134,26 +140,41 @@ R_root=max(0,(params_matrix.beta(R_root_index)+beta_matrix(R_root_index,b))* log
 
 % 1.1.2 calculated fluxes
 NEP_calc=GPP-R_eco;
-NPP_calc=ANPP_foliage+ANPP_woody+BNPP;
-%options for R_auto_calc (insufficient data for regression or to calculate as sum/difference). 
-%Last one in list is the one that will be used.
+
+
+%options for CUE. lat one in list is the one that will be used.
 CUE=0.679-0.153*log10(age); %from DeLucia et al. 2007 (note that equation given mistakenly leaves of the log10(age))
-R_auto_calc = CUE.*GPP; 
[…]

### 2021-01-18 manuscript/ERL_global_C_review.Rmd
added results on age trends/ budget closure
#64

@@
 Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
-Asterisks indicate variables whose age trends were calculated based on other variables, as follows: $R_{auto}=NPP(1/CUE-1)$, where $CUE$ is TEMPORARILY 0.5; $ANPP_{woody}=max(0,ANPP-ANPP_{foliage})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
+Asterisks indicate variables whose age trends were calculated based on other variables, as follows: $R_{auto-ag}=R_{auto}-R{root}$, where $R_{auto}=NPP(1/CUE-1)$, where $CUE$ is TEMPORARILY 0.5; $ANPP_{woody}=max(0,ANPP-ANPP_{foliage})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
 age_trends_caption_generic= "Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
@@
 
-Average C cycles for forests <100 years old are presented in Figures 8-11.
-Both C stocks and fluxes commonly increased significantly with stand age (Tables 1, S2, Figs. `r flux_age_trends`- `r stock_age_trends`, `r all_age_trends`; detailed below).
+Both C stocks and fluxes commonly increased significantly with stand age (Tables 1, S2, Figs.
+`r flux_age_trends`- `r stock_age_trends`, `r TrB_age_trends`-`r BoN_age_trends`, `r all_age_trends`; detailed below).
 
 *ForC* contained 16 C flux variables with sufficient data for analyses of age trends in young forests (see Methods) (Figs. `r flux_age_trends`, `r all_flux_age_trends`).
-Of these, ten increased significantly with age: $NEP$, $GPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{foliage}$, $ANPP_{litterfal}$, $BNPP$, $BNPP_{fine}$, $R_{eco}$, and $R_{root}$.
+Of these, ten increased significantly with $log10[age]$: $NEP$, $GPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{foliage}$, $ANPP_{litterfal}$, $BNPP$, $BNPP_{fine}$, $R_{eco}$, and $R_{root}$.
 The remaining six--$NPP$, $ANPP_{stem}$, $ANPP_{branch}$, $BNPP_{coarse}$, $R_{soil-het}$, and $R_{soil-het}$-–displayed no significant relationship to stand age.
@@
 `r flux_age_trends`), the tendency for temperate forests to have greater fluxes than boreal forests held for $NEP$ in regrowth forests (tropical forests excluded because of insufficient data).
+
+"Closure" and internal consistency of the C flux budget was less successful for young than mature forests (Figs. `r TrB_age_trends`-`r BoN_age_trends`).
+Summed regression equations for $R_{soil-het}$ and $R_{root}$ were generally very close to $R_{soil}$.
+We calculated $R_{auto-ag}$ as the difference between $R_{eco}$ and $R_{soil}$ (except for tropical forests, which had insufficient $R_{eco}$ data), effectively guaranteeing near-closure of the CO~2~ efflux (respiration) portion of the budget (negative values in Figs.
+`r TrB_age_trends`-`r BoN_age_trends`).
+In contrast, the  CO~2~ influx portion of the budget generally did not "close": the sum of $R_{auto}$ ($R_{root}+R_{auto-ag}$, as described above) and components of $NPP$ consistently fell short of $GPP$ by **##-##** Mg C ha^-1^ yr^-1^ , particularly in in young stands.
+Moreover, there was not consistent budget closure among the components of $NPP$, and substantially different age trends resulting from the sum of components versus total $NPP$ (Figs.
+`r TrB_age_trends`-`r BoN_age_trends`).
 

### 2021-01-18 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m

@@ -156,7 +156,7 @@ elseif b==1  %tropical forests
     NPP_calc=ANPP_foliage+ANPP_woody_calc+BNPP;
     %options for R_auto_calc (insufficient data for regression or to calculate as sum/difference). 
     %Last one in list is the one that will be used.
-    R_auto_calc = NPP_calc.*(1./CUE-1); %Rauto=NPP*(1/CUE-1), and CUE equation is above
+    R_auto_calc = max(R_root,NPP_calc.*(1./CUE-1)); %Rauto=NPP*(1/CUE-1), and CUE equation is above
     R_auto_ag_calc=max(0,R_auto_calc-R_root);
 end
 
@@ -170,6 +170,8 @@ BNPP_coarse_calc=max(0,BNPP-BNPP_fine);
 % "in" (GPP components) 
 in_flux_names = {'R_{root}', 'R_{auto-ag}*', 'BNPP', 'ANPP_{foliage}', 'ANPP_{woody}*'};
 in_fluxes = [R_root; R_auto_ag_calc; BNPP; ANPP_foliage; ANPP_woody_calc ]; %matrix with all fluxes for stacked plot
+%in_fluxes = [R_root; R_auto_ag_calc; NPP ]; %matrix with all fluxes for stacked plot
+
 in_sum = sum(in_fluxes,1);
 
 % "out" (Reco components)
@@ -246,7 +248,6 @@ h=area (age, in_fluxes'); %, 'LineStyle','-');
 hold on;
 h=area (age, -1*out_fluxes'); %, 'LineStyle','-'); 
 hold on;
-%plot(age, in_sum, '--b', 'LineWidth', 3); hold on;
 if b~= 1 
     plot(age, GPP, '-b', 'LineWidth', 3); hold on; % eddy flux: insufficient data for tropics
     plot(age, -R_eco, '-r', 'LineWidth', 3); % eddy flux: insufficient data for tropics
@@ -254,6 +255,8 @@ if b~= 1
     plot(age, NEP, '-w', 'LineWidth', 3);hold on; % eddy flux: insufficient data for tropics
     plot(age, NEP_calc, '--w', 'LineWidth', 3);hold on;
 end
+plot(age, in_sum, '--k', 'LineWidth', 3); hold on;
+plot(age, R_auto_calc+NPP, '-k', 'LineWidth', 3);hold on;
 plot(age, -R_soil, '-k', 'LineWidth', 3);hold on;
[…]

### 2021-01-18 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m
replaced mature forest flux values with real values
#64

@@ -8,9 +8,10 @@ set(0,'DefaultLegendAutoUpdate','off');
 %% SETTINGS
 figures_dir='/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/manuscript/tables_figures';
 working_dir='/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/manuscript/tables_figures/stacked_age_trends';
+ForC_dir='/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC';
 
 %% READ IN & PREPARE DATA
-%read in data:
+%read in age trends modesl summaries:
 cd(figures_dir)
 model_summaries=readtable('SI_age_trend_model_summaries.csv');
 %remove specieal formatting from model_summaries.Variable:
@@ -87,6 +88,10 @@ end
 
 int_matrix=[params_matrix.intTrB, params_matrix.intTeB, params_matrix.intTeN, params_matrix.intBoN];
 beta_matrix=[params_matrix.betaTrB, params_matrix.betaTeB, params_matrix.betaTeN, params_matrix.betaBoN];
+
+%read in ForC variable averages
+cd(strcat(ForC_dir,'/numbers_and_facts'))
+ForC_variable_averages=readtable('ForC_variable_averages_per_Biome.csv');
 
 for b =1:4
 %% GENERATE EQUATIONS FOR PLOTTING
@@ -170,7 +175,6 @@ BNPP_coarse_calc=max(0,BNPP-BNPP_fine);
 % "in" (GPP components) 
 in_flux_names = {'R_{root}', 'R_{auto-ag}*', 'BNPP', 'ANPP_{foliage}', 'ANPP_{woody}*'};
 in_fluxes = [R_root; R_auto_ag_calc; BNPP; ANPP_foliage; ANPP_woody_calc ]; %matrix with all fluxes for stacked plot
-%in_fluxes = [R_root; R_auto_ag_calc; NPP ]; %matrix with all fluxes for stacked plot
 
 in_sum = sum(in_fluxes,1);
 
@@ -185,12 +189,41 @@ fluxes = [in_fluxes; -1*out_fluxes];
 
 
 % 1.2. define mature fluxes 
[…]

### 2021-01-19 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m
real results for stocks
#64

@@ -216,9 +216,8 @@ GPP_mature=ForC_variable_averages.mean(GPP_index);
 R_het_soil_index=find(strcmp(ForC_variable_averages.variable_diagram,'R_het_soil') + strcmp(ForC_variable_averages.Biome,strcat(biomes(b),' MATURE'))==2);
 R_het_soil_mature=ForC_variable_averages.mean(R_het_soil_index);
 
-%THIS IS CURRENTLY JUST THE 100-YEAR VALUE FOR FLUXES. NEEDS TO BE REPLACED WITH MATURE BIOME MEANS!
-
-
+R_eco_index=find(strcmp(ForC_variable_averages.variable_diagram,'R_eco') + strcmp(ForC_variable_averages.Biome,strcat(biomes(b),' MATURE'))==2);
+R_eco_mature=ForC_variable_averages.mean(R_eco_index);
 
 %for grouped stacked plot
 mature_fluxes= [R_root_mature R_auto_ag_mature BNPP_mature ANPP_foliage_mature ANPP_woody_mature... %in fluxes
@@ -268,8 +267,37 @@ stocks = [ B_root_coarse ; B_root_fine ; B_ag_wood_calc;  B_foliage; DW_standing
 
 
 % 2. mature forests
+% 2.2.1 ForC means
+B_root_coarse_index=find(strcmp(ForC_variable_averages.variable_diagram,'biomass_root_coarse') + strcmp(ForC_variable_averages.Biome,strcat(biomes(b),' MATURE'))==2);
+B_root_coarse_mature=ForC_variable_averages.mean(B_root_coarse_index);
 
-mature_stocks = [stocks(:,end)'; stocks(:,end)' ; stocks(:,end)'; stocks(:,end)'];
+B_root_fine_index=find(strcmp(ForC_variable_averages.variable_diagram,'biomass_root_fine') + strcmp(ForC_variable_averages.Biome,strcat(biomes(b),' MATURE'))==2);
+B_root_fine_mature=ForC_variable_averages.mean(B_root_fine_index);
+
+B_ag_wood_index=find(strcmp(ForC_variable_averages.variable_diagram,'biomass_ag_woody') + strcmp(ForC_variable_averages.Biome,strcat(biomes(b),' MATURE'))==2);
+B_ag_wood_mature=ForC_variable_averages.mean(B_ag_wood_index);
+
[…]

### 2021-01-19 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m
combined young and mature into one plot
#64
(messes up flux legend)

@@ -165,13 +165,9 @@ elseif b==1  %tropical forests
     R_auto_ag_calc=max(0,R_auto_calc-R_root);
 end
 
-
-
-
-
 BNPP_coarse_calc=max(0,BNPP-BNPP_fine);
 
-% 1.2 group for plotting
+% 1.1.3 group for plotting
 % "in" (GPP components) 
 in_flux_names = {'R_{root}', 'R_{auto-ag}*', 'BNPP', 'ANPP_{foliage}', 'ANPP_{woody}*'};
 in_fluxes = [R_root; R_auto_ag_calc; BNPP; ANPP_foliage; ANPP_woody_calc ]; %matrix with all fluxes for stacked plot
@@ -188,7 +184,7 @@ flux_names= [in_flux_names, out_flux_names];
 fluxes = [in_fluxes; -1*out_fluxes];
 
 
-% 1.2. define mature fluxes 
+% 1.2.1 define mature fluxes 
 NEP_index=find(strcmp(ForC_variable_averages.variable_diagram,'NEP') + strcmp(ForC_variable_averages.Biome,strcat(biomes(b),' MATURE'))==2);
 NEP_mature=ForC_variable_averages.mean(NEP_index);
 
@@ -213,16 +209,23 @@ ANPP_stem_mature=ForC_variable_averages.mean(ANPP_stem_index);
 GPP_index=find(strcmp(ForC_variable_averages.variable_diagram,'GPP') + strcmp(ForC_variable_averages.Biome,strcat(biomes(b),' MATURE'))==2);
 GPP_mature=ForC_variable_averages.mean(GPP_index);
 
+R_soil_index=find(strcmp(ForC_variable_averages.variable_diagram,'R_soil') + strcmp(ForC_variable_averages.Biome,strcat(biomes(b),' MATURE'))==2);
+R_soil_mature=ForC_variable_averages.mean(R_soil_index);
+
 R_het_soil_index=find(strcmp(ForC_variable_averages.variable_diagram,'R_het_soil') + strcmp(ForC_variable_averages.Biome,strcat(biomes(b),' MATURE'))==2);
 R_het_soil_mature=ForC_variable_averages.mean(R_het_soil_index);
 
[…]

### 2021-01-19 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m
fix color scheme
#64

@@ -169,14 +169,14 @@ BNPP_coarse_calc=max(0,BNPP-BNPP_fine);
 
 % 1.1.3 group for plotting
 % "in" (GPP components) 
-in_flux_names = {'R_{root}', 'R_{auto-ag}*', 'BNPP', 'ANPP_{foliage}', 'ANPP_{woody}*'};
-in_fluxes = [R_root; R_auto_ag_calc; BNPP; ANPP_foliage; ANPP_woody_calc ]; %matrix with all fluxes for stacked plot
+in_flux_names = {'R_{auto-ag}*', 'R_{root}',  'BNPP',  'ANPP_{woody}*','ANPP_{foliage}'};
+in_fluxes = [ R_auto_ag_calc; R_root; BNPP;  ANPP_woody_calc ;ANPP_foliage]; %matrix with all fluxes for stacked plot
 
 in_sum = sum(in_fluxes,1);
 
 % "out" (Reco components)
-out_flux_names = { 'R_{het-soil}', 'R_{root}', 'R_{auto-ag}*'};
-out_fluxes = [R_het_soil; R_root; R_auto_ag_calc ]; %matrix with all fluxes for stacked plot
+out_flux_names = { 'R_{auto-ag}*', 'R_{root}', 'R_{het-soil}'};
+out_fluxes = [ R_auto_ag_calc ; R_root; R_het_soil]; %matrix with all fluxes for stacked plot
 out_sum = sum(out_fluxes,1);
 
 % all fluxes
@@ -222,8 +222,8 @@ R_eco_mature=ForC_variable_averages.mean(R_eco_index);
 R_auto_ag_mature_calc=R_eco_mature-R_soil_mature;
 
 %for grouped stacked plot (%THIS NEEDS TO MATCH YOUNG STANDS, ABOVE)
-mature_in_fluxes= [R_root_mature R_auto_ag_mature_calc BNPP_mature ANPP_foliage_mature ANPP_woody_mature]; %in fluxes
-mature_out_fluxes= [-R_het_soil_mature -R_root_mature -R_auto_ag_mature_calc]; % out fluxes
+mature_in_fluxes= [ R_auto_ag_mature_calc R_root_mature  BNPP_mature ANPP_woody_mature ANPP_foliage_mature ]; %in fluxes
+mature_out_fluxes= [-R_auto_ag_mature_calc -R_root_mature  -R_het_soil_mature ]; % out fluxes
[…]

### 2021-01-19 manuscript/ERL_global_C_review.Rmd
check mature forest values, explain calculations in legend
#64

@@
 Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
-Asterisks indicate variables whose age trends were calculated based on other variables, as follows: $R_{auto-ag}=R_{auto}-R{root}$, where $R_{auto}=NPP(1/CUE-1)$, where $CUE$ is TEMPORARILY 0.5; $ANPP_{woody}=max(0,ANPP-ANPP_{foliage})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
+Asterisks indicate variables whose age trends were calculated based on other variables (*young and mature forests; **young forests only; ***mature forests only), as follows: $R_{auto-ag}=R_{auto}-R{root}$, where $R_{auto}=NPP(1/CUE-1)$, where $CUE$ is TEMPORARILY 0.5; $ANPP_{woody}=max(0,ANPP-ANPP_{foliage})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
 age_trends_caption_generic= "Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
-Asterisks indicate variables whose age trends were calculated based on other variables, as follows: $R_{auto-ag}=R_{eco}-R{soil}$; ANPP_{woody}=min(ANPP_{stem},ANPP_{woody}); $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
+Asterisks indicate variables whose age trends were calculated based on other variables (*young and mature forests; **young forests only; ***mature forests only), as follows: $R_{auto-ag}=R_{eco}-R{soil}$; ANPP_{woody}=min(ANPP_{stem},ANPP_{woody}); $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
 TeB_age_trends_caption = paste("Figure 10 | Age trends in C cycling in temperate broadleaf forests.", age_trends_caption_generic) 

### 2021-01-19 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m

@@ -169,7 +169,7 @@ BNPP_coarse_calc=max(0,BNPP-BNPP_fine);
 
 % 1.1.3 group for plotting
 % "in" (GPP components) 
-in_flux_names = {'R_{auto-ag}*', 'R_{root}',  'BNPP',  'ANPP_{woody}*','ANPP_{foliage}'};
+in_flux_names = {'R_{auto-ag}*', 'R_{root}',  'BNPP',  'ANPP_{woody}**','ANPP_{foliage}'};
 in_fluxes = [ R_auto_ag_calc; R_root; BNPP;  ANPP_woody_calc ;ANPP_foliage]; %matrix with all fluxes for stacked plot
 
 in_sum = sum(in_fluxes,1);
@@ -260,17 +260,20 @@ OL=max(0,(params_matrix.beta(OL_index)+beta_matrix(OL_index,b))* log10(age)+ int
 
 % 2.1.2 caculated stocks
 DW_standing=max(0,DW_tot-DW_down);
-B_root_coarse=max(0,B_root-B_root_fine);
-B_tot_calc=B_ag+B_root_coarse+B_root_fine;
+B_root_coarse_calc=max(0,B_root-B_root_fine);
+B_tot_calc=B_ag+B_root_coarse_calc+B_root_fine;
 B_ag_wood_calc=max(0,B_ag-B_foliage);
 
 % 2.2 group for plotting
-stock_names = { 'B_{root-coarse}*', 'B_{root-fine}','B_{ag-wood}*', 'B_{foliage}','DW_{standing}*', 'DW_{down}', 'OL'};
-stocks = [ B_root_coarse ; B_root_fine ; B_ag_wood_calc;  B_foliage; DW_standing; DW_down; OL]; %matrix with all stocks for stacked plot
+stock_names = { 'B_{root-coarse}*', 'B_{root-fine}','B_{ag-wood}*', 'B_{foliage}','DW_{standing}**', 'DW_{down}***', 'OL'};
+stocks = [ B_root_coarse_calc ; B_root_fine ; B_ag_wood_calc;  B_foliage; DW_standing; DW_down; OL]; %matrix with all stocks for stacked plot
 
 
 % 2. mature forests
 % 2.2.1 ForC means
+B_root_index=find(strcmp(ForC_variable_averages.variable_diagram,'biomass_root') + strcmp(ForC_variable_averages.Biome,strcat(biomes(b),' MATURE'))==2);
+B_root_mature=ForC_variable_averages.mean(B_root_index);
+
[…]

### 2021-01-19 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m
added GPP, NEP, Reco to mature forests + legend
#64

@@ -319,16 +319,22 @@ subplot (2,1,1) %FLUXES
 h1a=area (age, in_fluxes'); hold on;
 h2a=area(112:116, mature_in_fluxes.*ones(5,size(mature_in_fluxes,1)),'HandleVisibility','off'); hold on;
 
-h1b=area (age, -1*out_fluxes'); %, 'LineStyle','-'); 
+h2bl=area(age, -1*out_fluxes(3,:)'); hold on; %not shown, just for legend
+h1b=area (age, -1*out_fluxes','HandleVisibility','off'); 
 h2b=area(112:116, mature_out_fluxes.*ones(5,size(mature_out_fluxes,1)),'HandleVisibility','off'); hold on;
 
 if b~= 1 
-    plot(age, GPP, '-b', 'LineWidth', 3); hold on; % eddy flux: insufficient data for tropics
-    plot(age, -R_eco, '-r', 'LineWidth', 3); % eddy flux: insufficient data for tropics
+    plot(age, GPP, '-b', 'LineWidth', 2,'HandleVisibility','off'); hold on; % eddy flux: insufficient data for tropics
+    plot(age, -R_eco, '-r', 'LineWidth', 2,'HandleVisibility','off'); % eddy flux: insufficient data for tropics
     %plot(age, -R_root-R_het_soil-R_auto_ag_calc, '--r', 'LineWidth', 3); % eddy flux: insufficient data for tropics
-    plot(age, NEP, '-w', 'LineWidth', 3);hold on; % eddy flux: insufficient data for tropics
+    plot(age, NEP, '-w', 'LineWidth', 2,'HandleVisibility','off');hold on; % eddy flux: insufficient data for tropics
     %plot(age, NEP_calc, '--w', 'LineWidth', 3);hold on;
 end
+plot (114, GPP_mature, 'ok', 'MarkerFaceColor', 'b'); hold on;
+plot (114, NEP_mature, 'ok', 'MarkerFaceColor', 'w'); hold on;
+plot (114, -R_eco_mature, 'ok', 'MarkerFaceColor', 'r'); hold on;
+
+
 %plot(age, in_sum, '--k', 'LineWidth', 3); hold on;
 %plot(age, R_auto_calc+NPP, '-k', 'LineWidth', 3);hold on;
 %plot(age, -R_auto_ag_calc-R_soil, '-y', 'LineWidth', 3);hold on;
[…]

### 2021-01-19 manuscript/ERL_global_C_review.Rmd
presentation of age trend figures in manuscript

@@
 
-Both C stocks and fluxes commonly increased significantly with stand age (Tables 1, S2, Figs.
-`r flux_age_trends`- `r stock_age_trends`, `r TrB_age_trends`-`r BoN_age_trends`, `r all_age_trends`; detailed below).
-
-*ForC* contained 16 C flux variables with sufficient data for analyses of age trends in young forests (see Methods) (Figs. `r flux_age_trends`, `r all_flux_age_trends`).
+C fluxes commonly increased significantly with stand age (Tables 1, S2, Figs. `r flux_age_trends`, `r TrB_age_trends`-`r BoN_age_trends`, `r all_age_trends`).
+*ForC* contained 16 C flux variables with sufficient data for analyses of age trends in young forests (see Methods).
 Of these, ten increased significantly with $log10[age]$: $NEP$, $GPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{foliage}$, $ANPP_{litterfal}$, $BNPP$, $BNPP_{fine}$, $R_{eco}$, and $R_{root}$.
-The remaining six--$NPP$, $ANPP_{stem}$, $ANPP_{branch}$, $BNPP_{coarse}$, $R_{soil-het}$, and $R_{soil-het}$-–displayed no significant relationship to stand age.
-
-Differences in C fluxes across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests.
-The single exception was $ANPP_{stem}$, for which temperate broadleaf and conifer forests had similar flux rates than tropical forests.
-Notably, and in contrast to the lack of biome differences in $NEP$ for mature forests (Fig.
-`r flux_age_trends`), the tendency for temperate forests to have greater fluxes than boreal forests held for $NEP$ in regrowth forests (tropical forests excluded because of insufficient data).
-
-"Closure" and internal consistency of the C flux budget was less successful for young than mature forests (Figs. `r TrB_age_trends`-`r BoN_age_trends`).
-Summed regression equations for $R_{soil-het}$ and $R_{root}$ were generally very close to $R_{soil}$.
-We calculated $R_{auto-ag}$ as the difference between $R_{eco}$ and $R_{soil}$ (except for tropical forests, which had insufficient $R_{eco}$ data), effectively guaranteeing near-closure of the CO~2~ efflux (respiration) portion of the budget (negative values in Figs.
-`r TrB_age_trends`-`r BoN_age_trends`).
-In contrast, the  CO~2~ influx portion of the budget generally did not "close": the sum of $R_{auto}$ ($R_{root}+R_{auto-ag}$, as described above) and components of $NPP$ consistently fell short of $GPP$ by **##-##** Mg C ha^-1^ yr^-1^ , particularly in in young stands.
-Moreover, there was not consistent budget closure among the components of $NPP$, and substantially different age trends resulting from the sum of components versus total $NPP$ (Figs.
-`r TrB_age_trends`-`r BoN_age_trends`).
-
-In terms of C stocks, ten variables (all but standing deadwood, $DW_{standing}$) had sufficient data to test for age trends (Table 1, Figs. `r stock_age_trends`, `r all_stock_age_trends`).
-All of these displayed a significant overall increase with with $log10[stand.age]$.
-There were sufficient data to model age $\times$ biome interactions were also significant for all ten of these C stock variables (Table S2), with living C stocks tending to accumulate more rapidly during the early stages of forest regrowth in tropical forests (Figs.
-`r stock_age_trends`, `r all_stock_age_trends`).
-In the case of two non-living C stocks ($DW_{down}$ and $OL$), age $\times$ biome interactions were such that 
-Specifically, $DW_{down}$ declined with age in temperate and boreal forests, compared to an increase with age in tropical forests (Figs. `r stock_age_trends`, `r DW_down_fig`).
-Similarly, $OL$ declined slightly with age in temperate broadleaf forests, contrasting an increase in the other three biomes (Figs. `r stock_age_trends`, `r OL_fig`).
+The remaining six--$NPP$, $ANPP_{stem}$, $ANPP_{branch}$, $BNPP_{coarse}$, **$R_{soil-het}$, and $R_{soil-het}$**-–displayed no significant relationship to stand age.
 
@@
 ```
+*(paragraph on young vs mature)*
+
+Differences in C fluxes across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests.
+The single exception was $ANPP_{stem}$, for which temperate broadleaf and conifer forests had similar flux rates than tropical forests.
+Notably, and in contrast to the lack of biome differences in $NEP$ for mature forests (Fig.
+`r flux_age_trends`), the tendency for temperate forests to have greater fluxes than boreal forests held for $NEP$ in regrowth forests (tropical forests excluded because of insufficient data).
+
+"Closure" and internal consistency of the C flux budget was less successful for young than mature forests (Figs. `r TrB_age_trends`-`r BoN_age_trends`).
+Summed regression equations for $R_{soil-het}$ and $R_{root}$ were generally very close to $R_{soil}$.
+We calculated $R_{auto-ag}$ as the difference between $R_{eco}$ and $R_{soil}$ (except for tropical forests, which had insufficient $R_{eco}$ data), effectively guaranteeing near-closure of the CO~2~ efflux (respiration) portion of the budget (negative values in Figs.
+`r TrB_age_trends`-`r BoN_age_trends`).
+In contrast, the  CO~2~ influx portion of the budget generally did not "close": the sum of $R_{auto}$ ($R_{root}+R_{auto-ag}$, as described above) and components of $NPP$ consistently fell short of $GPP$, particularly in in young stands (range across forest types and ages: 0.9-7.6 Mg C ha^-1^ yr^-1^).
+Moreover, there was not consistent budget closure among the components of $NPP$, and substantially different age trends resulting from the sum of components versus total $NPP$ (Figs.
+`r TrB_age_trends`-`r BoN_age_trends`).
+
+In terms of C stocks, ten variables (all but standing deadwood, $DW_{standing}$) had sufficient data to test for age trends (Table 1, Figs. `r stock_age_trends`, `r all_stock_age_trends`).
+All of these displayed a significant overall increase with with $log10[stand.age]$.
+There were sufficient data to model age $\times$ biome interactions were also significant for all ten of these C stock variables (Table S2), with living C stocks tending to accumulate more rapidly during the early stages of forest regrowth in tropical forests (Figs.
+`r stock_age_trends`, `r all_stock_age_trends`).
+In the case of two non-living C stocks ($DW_{down}$ and $OL$), age $\times$ biome interactions were such that 
+Specifically, $DW_{down}$ declined with age in temperate and boreal forests, compared to an increase with age in tropical forests (Figs. `r stock_age_trends`, `r DW_down_fig`).
+Similarly, $OL$ declined slightly with age in temperate broadleaf forests, contrasting an increase in the other three biomes (Figs. `r stock_age_trends`, `r OL_fig`).
+
+*(paragraph on young vs mature)*
 

### 2021-01-19 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m

@@ -400,4 +400,15 @@ cd(working_dir)
 savename=char(strcat(biomes(b),'_age_trends'));
 print(savename, '-dpng', '-r300')
 
+
+%figure(10)
+%if b>1
+%plot(age, GPP-in_sum); hold on;
+%end
+
 end
+
+%figure 10, cont
+%xlabel ('age')
+%ylabel ('GPP-in.sum')
+%legend ('TeB', 'TeN', 'BoN')

### 2021-01-19 manuscript/ERL_global_C_review.Rmd
updated age trends by biome
#63

@@
 Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
-Asterisks indicate variables whose age trends were calculated based on other variables (*young and mature forests; **young forests only; ***mature forests only), as follows: $R_{auto-ag}=R_{auto}-R{root}$, where $R_{auto}=NPP(1/CUE-1)$, where $CUE$ is TEMPORARILY 0.5; $ANPP_{woody}=max(0,ANPP-ANPP_{foliage})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
+Asterisks indicate variables whose age trends were calculated based on other variables (* young and mature forests; ** young forests only; *** mature forests only), as follows: $R_{auto-ag}=R_{auto}-R{root}$, where $R_{auto}=NPP(1/CUE-1)$, where $CUE$ is TEMPORARILY 0.5; $ANPP_{woody}=max(0,ANPP-ANPP_{foliage})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
 age_trends_caption_generic= "Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
-Asterisks indicate variables whose age trends were calculated based on other variables (*young and mature forests; **young forests only; ***mature forests only), as follows: $R_{auto-ag}=R_{eco}-R{soil}$; ANPP_{woody}=min(ANPP_{stem},ANPP_{woody}); $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
+Asterisks indicate variables whose age trends were calculated based on other variables (* young and mature forests; ** young forests only; *** mature forests only), as follows: $R_{auto-ag}=R_{eco}-R{soil}$; ANPP_{woody}=min(ANPP_{stem},ANPP_{woody}); $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
 TeB_age_trends_caption = paste("Figure 10 | Age trends in C cycling in temperate broadleaf forests.", age_trends_caption_generic) 

### 2021-01-19 manuscript/ERL_global_C_review.Rmd
update CUE

--- 2021-01-19 manuscript/ERL_global_C_review.Rmd
@@
 Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
-Asterisks indicate variables whose age trends were calculated based on other variables (*young and mature forests; **young forests only; ***mature forests only), as follows: $R_{auto-ag}=R_{auto}-R{root}$, where $R_{auto}=NPP(1/CUE-1)$, where $CUE$ is TEMPORARILY 0.5; $ANPP_{woody}=max(0,ANPP-ANPP_{foliage})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
+Asterisks indicate variables whose age trends were calculated based on other variables (* young and mature forests; ** young forests only; *** mature forests only), as follows: $R_{auto-ag}=R_{auto}-R{root}$, where $R_{auto}=NPP(1/CUE-1)$, where $CUE$=0.46 (Collati et al.
+2020); $ANPP_{woody}=max(0,ANPP-ANPP_{foliage})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
 age_trends_caption_generic= "Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
-Asterisks indicate variables whose age trends were calculated based on other variables (*young and mature forests; **young forests only; ***mature forests only), as follows: $R_{auto-ag}=R_{eco}-R{soil}$; ANPP_{woody}=min(ANPP_{stem},ANPP_{woody}); $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
+Asterisks indicate variables whose age trends were calculated based on other variables (* young and mature forests; ** young forests only; *** mature forests only), as follows: $R_{auto-ag}=R_{eco}-R{soil}$; $ANPP_{woody}=min(ANPP_{stem},ANPP_{woody})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
 TeB_age_trends_caption = paste("Figure 10 | Age trends in C cycling in temperate broadleaf forests.", age_trends_caption_generic) 
@@
 ```
+
+---
+nocite: | 
+@collalti_forest_2020
+---
 
@@
 Of these, ten increased significantly with $log10[age]$: $NEP$, $GPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{foliage}$, $ANPP_{litterfal}$, $BNPP$, $BNPP_{fine}$, $R_{eco}$, and $R_{root}$.
-The remaining six--$NPP$, $ANPP_{stem}$, $ANPP_{branch}$, $BNPP_{coarse}$, **$R_{soil-het}$, and $R_{soil-het}$**-–displayed no significant relationship to stand age.
+The remaining six--$NPP$, $ANPP_{stem}$, $ANPP_{branch}$, $BNPP_{coarse}$, $R_{soil}$, and $R_{het-soil}$-–displayed no significant relationship to stand age.
 
@@
 ```
-*(paragraph on young vs mature)*
 
@@
 `r TrB_age_trends`-`r BoN_age_trends`).
+
+There were also some substantial discrepancies between 
+*(paragraph on young vs mature)*
 

### 2021-01-19 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m

@@ -148,8 +148,8 @@ NEP_calc=GPP-R_eco;
 
 
 %options for CUE. lat one in list is the one that will be used.
-CUE=0.679-0.153*log10(age); %from DeLucia et al. 2007 (note that equation given mistakenly leaves of the log10(age))
-CUE=0.5;
+CUE=0.679-0.153*log10(age); %from DeLucia et al. 2007 (note that equation given mistakenly leaves of the log10(age)). Collati et al. 2020 gives an update of this.
+CUE=0.46; %from Collati et al. 2020
 
 if b>1  %extratropical forests
     ANPP_woody_calc=min(ANPP_stem, ANPP_woody);

### 2021-01-19 manuscript/response_to_reviews.Rmd
a bit of cleanup
small updates

--- 2021-01-14 manuscript/response_to_reviews.Rmd
@@
 
-## REFEREE REPORT(S):
+## LETTER TO EDITOR:
+
+The most substantive changes included the following:
+
+- Changed the figure set in accordance with comments from R3. Specifically, we
+(1) added a schematic figure, somewhat in the style of @odum_strategy_1969, to help with conceptual framing of the review
+(2) added figures synthesizing changes in fluxes through time for each biome;
+(3) removed maps from age trend diagrams (available in SI) from age trend diagrams, which makes the figures less busy/ easier to read; and
+(4) moved C cycle diagrams for young forests to SI;
+
+- Improved conceptual framework and "review" nature of the work by adding background information to the introduction, including the schematic figure mentioned above.
+We also shifted emphasis more towards the ecological questions (trends across biomes and with age), while retaining (with less emphasis) the analysis of database representativeness.
+
+\newpage
+
+## REFEREE REPORT(S) & RESPONSES:
 
@@
 
+**The 147 refers to *total* annual terrestrial photosynthesis, including non-forests.
+Table S4 breaks this down by biome.
+The >69 Gt C/yr figure was obtained by summing the forest biomes, not including savannas.**
+
 b. Line 81: "rare exceptions that span regions or continents" is repeated twice.
+
+**Fixed.**
 
@@
 
-*text edit*
+**Fixed.**
 
@@
 
-*straightforward--just a bit of text edit.*
+**Fixed.**
 

### 2021-01-19 manuscript/ERL_global_C_review.Rmd
small updates

@@
 *Background.* Earth’s climate is closely linked to forests, which strongly influence atmospheric carbon dioxide (CO~2~) and climate by their impact on the global carbon (C) cycle.
-However, efforts to incorporate forests into climate models and CO~2~ accounting frameworks have been constrained by a lack of accessible, global-scale data on how C cycling varies across forest types and stand ages.
+However, efforts to incorporate forests into climate models and CO~2~ accounting frameworks have been constrained by a lack of accessible, global-scale synthesis on how C cycling varies across forest types and stand ages.
 
@@
 The majority of flux variables, together with most live biomass pools, increased significantly with stand age.
-Importantly, there was generally good closure of C budgets, *i.e.,* internal consistency in the *ForC* data.
+There was generally good closure of C budgets for mature forests, whereas age trends and C budget closure in young forests remain less clearly resolved.
 
 *Discussion.* As climate change accelerates, understanding and managing the carbon dynamics of forests is critical to forecasting, mitigation, and adaptation.
-This synthetic and internally consistent global overview of C stocks and fluxes across biomes and stand ages will help to advance these efforts.
+This synthetic global overview of C stocks and fluxes across biomes and stand ages will help to advance these efforts.
 
@@
 Forest C cycling varies enormously across biomes, which are artificial categorical distinctions convenient for categorizing the world's forests according to major differences in climate, vegetation, *etc.*
-Since *(Humboldt?)*, it has been recognized that climate plays a dominant role in shaping differences among forests.
+Since the early 19th century, it has been recognized that climate plays a dominant role in shaping differences among forests on a global scale [@humboldt_essay_1807; @holdridge_determination_1947].
 Early quantitative syntheses revealed that C fluxes such as net primary productivity ($NPP$) decrease dramatically with latitude, or increase across gradients of mean annual temperature and, to a lesser extent, precipitation [@lieth_primary_1973].
@@
 
-While current data and technology allow increasingly fine-scale resolution for understanding differences across forests, cateogorization into biomes remains useful for applications where data do not yet support robust analysis across finer-scale axes of variation.
+While current data and technology allow increasingly fine-scale resolution for understanding differences across forests, categorization into biomes remains useful for applications where data do not yet support robust analysis across finer-scale axes of variation.
 

### 2021-01-20 manuscript/ERL_global_C_review.Rmd
worked on schematic and its coordination with results figures
#61 , #64

@@
 Since the early 19th century, it has been recognized that climate plays a dominant role in shaping differences among forests on a global scale [@humboldt_essay_1807; @holdridge_determination_1947].
+*C stocks*
 Early quantitative syntheses revealed that C fluxes such as net primary productivity ($NPP$) decrease dramatically with latitude, or increase across gradients of mean annual temperature and, to a lesser extent, precipitation [@lieth_primary_1973].
@@
 There has been little synthesis of differences in variables other than C stocks or their accumulation rates.
-Given the important role of secondary forests in the current and future global C cycle, concrete understanding of age trends in C fluxes and stocks and how these vary across biomes is critical to better understandign of the global C cycle.
+Given the important role of secondary forests in the current and future global C cycle, concrete understanding of age trends in C fluxes and stocks and how these vary across biomes is critical to better understanding of the global C cycle.
 
-```{r schematic, eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 1 | DRAFT Schematic diagram summarizing current understanding of biome differences and age trends in forest C cycling.
-Plotted values are realistic with respect to the results of this synthesis, but schematic is not intended to be quantitatively precise.
-", out.width='100%', fig.pos='H'}
+```{r schematic, eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 1 | Schematic diagram summarizing current understanding of biome differences and age trends in forest C cycling.
+Variables are defined in Table 1.
+Age trends, which represent idealized dynamics following a disturbance that removes all living and non-living vegetation, are an updated version of the classic figure from Odum (1969), with heavy lines corresponding to those in Odum's figure and $NEP$ corresponding to Odum's 'net production'.
+Here, $NEP$ conists primarily of woody aboveground net primary production ($ANPP_{woody}$), while $ANPP_{woody.turnover}$ is the sum of woody mortality and branch turnover ", out.width='100%', fig.pos='H'}
 

### 2021-01-20 manuscript/tables_figures/stacked_age_trends/plotBarStackGroups.m

@@ -44,9 +44,13 @@ set(gca,'XTick',1:NumGroupsPerAxis);
 set(gca,'XTickLabelMode','manual');
 set(gca,'XTickLabel',groupLabels);
 
-%set(h(1,1),'FaceColor',[11 102 35]/255); %KAT
-%for n=1:size(stackData,3)  %KAT
-%    set(h(2,n),'FaceColor',bar_colors(n,:)); %KAT
-%end %KAT
+for n=1:size(stackData,3)  %KAT
+    set(h(1,n),'FaceColor',bar_colors(n,:)); %KAT
+end %KAT
+
+set(h(2,1),'FaceColor',bar_colors(1,:)); %KAT
+set(h(2,2),'FaceColor',[.7 0 .7]); %KAT
+
+
 
 end 

### 2021-01-20 manuscript/tables_figures/stacked_age_trends/schematic.m

@@ -30,11 +30,11 @@ biome_names={'tropical', 'temperate', 'boreal'};
 %FLUXES
 % 1. by age
 %for stacked plot
-NEP=2*log10(age)-.037*(age-1); %placeholder (peaks around 2 at 23 yrs age)
+NEP=2.4*log10(age)-.04*(age-1); %placeholder (peaks around 2 at 23 yrs age)
 GPP= NEP + (1.99+1.44)* log10(age);
 ANPP_foliage = 1.66* log10(age) +.18;
 ANPP_woody = 2.31* log10(age) -.76;
-ANPP_woody_turnover=max(0,ANPP_woody-NEP);
+ANPP_woody_turnover=1.7*log10(age)-1; %max(0,ANPP_woody-NEP);
 BNPP=0.97*log10(age)+.67;
 NPP=ANPP_foliage+ANPP_woody+BNPP;
 R_auto = NPP.*(1+.001*age); %insufficient data
@@ -83,65 +83,81 @@ mature_stocks = [stocks(:,end)'; Stock_ratio_temp_trop*stocks(:,end)' ; Stock_ra
 
 %% ~~~~~~~PLOTTING~~~~~~~~~~
 
+%set facecolors:
+facecolor_in_fluxes= [0.2 0 .7;...  %' 'R_{auto}'
+                    0 .7 1;...  %'BNPP'
+                    0 .527 .2;... %'ANPP_{foliage}',
+                    150/255 1 0 ;...%'ANPP_{woody.turnover}*'
+                    255/256 255/256 0]; % NEP
+facecolor_stocks= [%0.1 0 .4;...  %'B_{root-coarse}*', ' 
+                    %0 .7 1;... % 'B_{root-fine}',
+                    %0 .527 .27 ;... % B_{foliage}',
+                    170/255 1 0;... % 'B_{ag-wood}*', 
+                    %1 .8 0;...  % 'DW_{standing}*', 
+                    1 99/255 99/255 ... % 'DW_{down}',
+                    %; 0.7 0 .7 ;... %'OL'
+                    ];
+
 figure (1)
 
 subplot (2,3,1)
 %stack as components of GPP and Reco, outline those.
-plotBarStackGroups(mature_fluxes, biome_names); 
+plotBarStackGroups(mature_fluxes, biome_names, facecolor_in_fluxes); 
 t = title('biome differences');
[…]

### 2021-01-20 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m

@@ -169,8 +169,8 @@ BNPP_coarse_calc=max(0,BNPP-BNPP_fine);
 
 % 1.1.3 group for plotting
 % "in" (GPP components) 
-in_flux_names = {'R_{auto-ag}*', 'R_{root}',  'BNPP',  'ANPP_{woody}**','ANPP_{foliage}'};
-in_fluxes = [ R_auto_ag_calc; R_root; BNPP;  ANPP_woody_calc ;ANPP_foliage]; %matrix with all fluxes for stacked plot
+in_flux_names = {'R_{auto-ag}*', 'R_{root}',  'BNPP', 'ANPP_{foliage}', 'ANPP_{woody}**'};
+in_fluxes = [ R_auto_ag_calc; R_root; BNPP;ANPP_foliage ;  ANPP_woody_calc ]; %matrix with all fluxes for stacked plot
 
 in_sum = sum(in_fluxes,1);
 
@@ -222,7 +222,7 @@ R_eco_mature=ForC_variable_averages.mean(R_eco_index);
 R_auto_ag_mature_calc=R_eco_mature-R_soil_mature;
 
 %for grouped stacked plot (%THIS NEEDS TO MATCH YOUNG STANDS, ABOVE)
-mature_in_fluxes= [ R_auto_ag_mature_calc R_root_mature  BNPP_mature ANPP_woody_mature ANPP_foliage_mature ]; %in fluxes
+mature_in_fluxes= [ R_auto_ag_mature_calc R_root_mature  BNPP_mature ANPP_foliage_mature  ANPP_woody_mature ]; %in fluxes
 mature_out_fluxes= [-R_auto_ag_mature_calc -R_root_mature  -R_het_soil_mature ]; % out fluxes
 mature_in_fluxes_matrix(b,1:length(mature_in_fluxes)) =mature_in_fluxes;
 mature_out_fluxes_matrix(b,1:length(mature_out_fluxes)) =mature_out_fluxes;
@@ -265,8 +265,8 @@ B_tot_calc=B_ag+B_root_coarse_calc+B_root_fine;
 B_ag_wood_calc=max(0,B_ag-B_foliage);
 
 % 2.2 group for plotting
-stock_names = { 'B_{root-coarse}*', 'B_{root-fine}','B_{ag-wood}*', 'B_{foliage}','DW_{standing}**', 'DW_{down}***', 'OL'};
[…]

### 2021-01-21 manuscript/ERL_global_C_review.Rmd
worked on intro

@@
 schematic =1
+schematic_fluxes_biome="1a"
+  schematic_fluxes_age="1b"
+  schematic_stocks_biome="1c"
+  schematic_stocks_age="1d"
+  schematic_fs_age = "1b,d"
 figure_map=2
@@
 Further standardization has been made possible through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; @davies_forestgeo_2021], NEON [@schimel_neon_2007], and FLUXNET [@baldocchi_fluxnet_2001; @novick_ameriflux_2018].
-Remote sensing technology has become increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass: @hu_mapping_2016; @spawn_harmonized_2020; gross primary productivity, $GPP$: @li_mapping_2019; @saatchi_benchmark_2011], yet measurement and validation of most forest C stocks and fluxes necessarily requires intensive on-the-ground data collection.
+Remote sensing technology has become increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass, $B_{ag}$: @saatchi_benchmark_2011; @hu_mapping_2016; @spawn_harmonized_2020; gross primary productivity, $GPP$: @li_mapping_2019], yet measurement and validation of most forest C stocks and fluxes necessarily requires intensive on-the-ground data collection.
 
@@
 Since the early 19th century, it has been recognized that climate plays a dominant role in shaping differences among forests on a global scale [@humboldt_essay_1807; @holdridge_determination_1947].
-*C stocks*
-Early quantitative syntheses revealed that C fluxes such as net primary productivity ($NPP$) decrease dramatically with latitude, or increase across gradients of mean annual temperature and, to a lesser extent, precipitation [@lieth_primary_1973].
+Global scale data synthesis has revealed that C fluxes including $GPP$, net primary productivity ($NPP$), and soil respiration ($R_{soil}$) decrease with latitude or, correspondingly, increase with mean annual temperature and, to a lesser extent, precipitation [Fig.
+`r schematic_fluxes_biome` **REFS**; @lieth_primary_1973; @luyssaert_co2_2007; @banbury_morgan_global_nodate].
+C stocks of mature forests show less directional variation (Fig. `r schematic_fluxes_biome`).
+On average, $B_{ag}$ tends to decrease with latitude, but not as dramatically as fluxes, and with the highest $B_{ag}$ forests in relatively cool, moist temperate regions [**REFS**; @keith_re-evaluation_2009].
+In contrast, standing and downed dead wood ($DW_{standing}$ and $DW_{down}$, respectively) and the organic layer ($OL$) tend to accumulate more in colder climates where decomposition is slow relative to $NPP$ [**REFS**].
 
-@banbury_morgan_global_nodate
-
-Recently, machine learning draws on global maps in combination with ground-based data to create global maps of C cycling [e.g., REFS; @cook-patton_mapping_2020].
-
-While current data and technology allow increasingly fine-scale resolution for understanding differences across forests, categorization into biomes remains useful for applications where data do not yet support robust analysis across finer-scale axes of variation.
+Correlative analyses relating C cycle variables to climate and other environmental variables have recently been taken to a new level through use of machine-learning algorithms that relate ground-based C cycle data to *global data on climate/soils/satellite data*, making it possible to create fine-scale global maps of C cycling [e.g., **REFS**; @cook-patton_mapping_2020].
+*This approach can be particularly effective when paired with satellite data ...
+(e.g., aboveground biomass: Saatchi et al 2011, Hu et al 2016, Spawn et al 2020, gross primary productivity, GPP: Li and Xiao 2019, ).*
+Yet, any such analysis is constrained by the quality and coverage of ground-based estimates of forest C fluxes or stocks. 
+While estimates of some variables (e.g., $B_{ag}$, $GPP$, $NPP$, $R_{soil}$) are widely available, many remain poorly characterized (e.g., $DW$; autotrophic respiration, $R_{auto}$) --even at the coarse resolution of biomes.
+This is a critical limitation not only for understanding forest C cycling, but also for quantifying climate change mitigation across forest biomes or ecozones [e.g., **REFS**].
 
 ### Age trends and their variation across biomes
-In 1969, Eugene Odum published ... [@odum_strategy_1969].
-Although the conceptualization of the C cycle in this paper is simplistic by current standards, the paper was foundational... [@corman_foundations_2019].
-...
-*(Need for a synthetic update.)*
+
+Stand age is another important axis of variation in forest C cycling (Fig. `r schematic_fs_age`).
+In 1969, E.P. Odum's "The Strategy of Ecosystem Development" laid out predictions as to how forest energy flows and organic matter stocks vary with stand age  [@odum_strategy_1969].
+Although the conceptualization of the C cycle in this paper is simplistic by current standards, the paper was foundational in framing the theory around which research on the subject still revolves [@corman_foundations_2019], and the basic framework still holds, *albiet* with modest modifications [@andersonteixeira_altered_2013].
+Following stand-clearing disturbance, $GPP$, $NPP$, and biomass of leaves ($B_{foliage}$) and fine roots ($B_{root-fine}$) increase rapidly and thereafter remain relatively stable ($B_{foliage}$,$B_{root-fine}$, sometimes $GPP$) or decline slightly [$NPP$, sometimes $GPP$; e.g., *Goulden et al.
+2011, ....
+refs in * @andersonteixeira_altered_2013].
+The decline in $NPP$ occurs because $R_{auto}$ increases relative to $GPP$ as forests age, corresponding to declining carbon use efficiency with stand age [@delucia_forest_2007; @collalti_forest_2020].
+Heterotrophic respiration, most of which originates from the soil ($R_{het-soil}$) remains relatively constant with stand age *[Law et al., 2003; Pregitzer & Euskirchen, 2004; Goulden et al., 2011]*, with the result that net ecosystem production ($NEP=GPP-R_{eco}$, where $R_{eco}$ is total ecosystem respiration) is initially negative, increases to a maximum at intermediate ages, and declines thereafter *[Law et al., 2003; Pregitzer & Euskirchen, 2004; Zhou et al., 2006; Baldocchi, 2008; Luyssaert et al., 2008; Amiro et al., 2010; Goulden et al., 2011]*.
+The result is that biomass accumulates rapidly in young forests, followed by a slow decline to near zero in old forests *[Lichstein et al., 2009; Yang et al., 2011;
+Hember et al., 2012]*.
+While these trends have been subject of fairly recent review [@andersonteixeira_altered_2013], *there is need for a synthetic, quantitaitive review.*
 
@@
 @anderson_temperature-dependence_2006
-@chazdon_carbon_2016
+@chazdon_carbon_2016, 
+Tepley et al., 
 @cook-patton_mapping_2020
@@
 Age trends, which represent idealized dynamics following a disturbance that removes all living and non-living vegetation, are an updated version of the classic figure from Odum (1969), with heavy lines corresponding to those in Odum's figure and $NEP$ corresponding to Odum's 'net production'.
-Here, $NEP$ conists primarily of woody aboveground net primary production ($ANPP_{woody}$), while $ANPP_{woody.turnover}$ is the sum of woody mortality and branch turnover ", out.width='100%', fig.pos='H'}
+Here, $NEP$ conists primarily of woody aboveground net primary production ($ANPP_{woody}$), while $ANPP_{woody.turnover}$ is the sum of woody mortality and branch turnover.
+", out.width='100%', fig.pos='H'}
 

### 2021-01-21 manuscript/ERL_global_C_review.Rmd
work on intro

@@
 
-In the past *couple of/several* decades, researchers have started asking how age trends--mostly in biomass accumulation-- vary across biomes.
-*(look up early work on this--citations in @anderson_temperature-dependence_2006)*
-@anderson_temperature-dependence_2006
-@chazdon_carbon_2016, 
-Tepley et al., 
-@cook-patton_mapping_2020
-There has been little synthesis of differences in variables other than C stocks or their accumulation rates.
-Given the important role of secondary forests in the current and future global C cycle, concrete understanding of age trends in C fluxes and stocks and how these vary across biomes is critical to better understanding of the global C cycle.
-
 ```{r schematic, eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 1 | Schematic diagram summarizing current understanding of biome differences and age trends in forest C cycling.
@@
 
-Here, conduct a data-based review of carbon cycling from a stand to global level, and by biome and stand age, using the largest global compilation of forest carbon data, which is available in our open source Global Carbon Forest database (*ForC*; Fig.
+In the past *couple of/several* decades, researchers have started asking how age trends--mostly in $B_{ag}$ or total biomass ($B_{tot}$) accumulation-- vary across biomes.
+Early research on this theme showed that biomass accumulation rates during secondary succession increase with temperature on a global scale [**REFS**; @anderson_temperature-dependence_2006] and *with precipitation in the neotropics* [**REFS**; @chazdon_carbon_2016].
+Most recently, @cook-patton_mapping_2020 reinforced these earlier findings with a much larger dataset and crated a high-resolution global map of estimated potential C accumulation rates.
+However, there has been little synthesis of cross-biome differences in variables other than biomass and its accumulation rate (but see @cook-patton_mapping_2020 for $DW$, $OL$, and soil C accumulation in young stands).
+Given the important role of secondary forests in the current and future global C cycle, concrete understanding of age trends in C fluxes and stocks and how these vary across biomes is critical to better understanding of the global C cycle.
+Accurate estimates of C sequestration rates by regrowth forests are also critical for national greenhouse gas accounting under the IPCC framework [**REFS**] and to quantifying the value of regrwoth forests for climate change mitigation [@anderson-teixeira_greenhouse_2011; @goldstein_protecting_2020].
+
+Here, we conduct a data-based review of carbon cycling from a stand to global level, and by biome and stand age, using the largest global compilation of forest carbon data, which is available in our open source Global Carbon Forest database (*ForC*; Fig.
 `r figure_map`).

### 2021-01-21 manuscript/ERL_global_C_review.Rmd
worked on manuscript

@@
 Since publication of ForC v2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [*SRDB* v4, `r n_SRDB` records; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database [*GROA* v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020].
-We have also added data from individual publications, with a particular focus on productivity [e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @lutz_global_2018, @johnson_climate_2018].
+We have also added data from individual publications, with a particular focus on productivity [e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @lutz_global_2018; @johnson_climate_2018].
 The database version used for this analysis has been tagged as a new release on Github (v3.0) and assigned a DOI through Zenodo (DOI: TBD).
@@
 
-  knitr::include_graphics("tables_figures/stacked_age_trends/Tropical broadleaf_age_trends.png")
+  knitr::include_graphics("tables_figures/stacked_age_trends/TrB_age_trends.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/stacked_age_trends/Temperate broadleaf_age_trends.png")
+  knitr::include_graphics("tables_figures/stacked_age_trends/TeB_age_trends.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/stacked_age_trends/Temperate conifer_age_trends.png")
+  knitr::include_graphics("tables_figures/stacked_age_trends/TeN_age_trends.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/stacked_age_trends/Boreal conifer_age_trends.png")
+  knitr::include_graphics("tables_figures/stacked_age_trends/BoN_age_trends.png")
 ```
@@
 `r TrB_age_trends`-`r BoN_age_trends`).
-
-There were also some substantial discrepancies between 
-*(paragraph on young vs mature)*
+ALthough age trends of young forests often converged towards mature forest averages, there were also some  discrepancies between young forest trends and mature forest averages (Figs.
+`r flux_age_trends`, `r TrB_age_trends`-`r BoN_age_trends`, `r all_age_trends`), most notably including a tendency for higher fluxes in regrowth boreal forests than in their mature counterparts (Fig.
+`r BoN_age_trends`).
 
@@
 Similarly, $OL$ declined slightly with age in temperate broadleaf forests, contrasting an increase in the other three biomes (Figs. `r stock_age_trends`, `r OL_fig`).
-
-*(paragraph on young vs mature)*
+Again, there were some discrepencies between young forest trends and mature forests, most notably including generally higher C stocks in mature forests relative to their 100-year counterparts, particularly for temperate conifer forests (*again, likely a geographic representation issue?*) and, to a lesser extent, tropical broadleaf forests.
 
@@
 
-*ForC* v3.0 provided unprecedented coverage of most major variables, yielding an internally consistent picture of C cycling in the world's major forest biomes.
+*ForC* v3.0 provided unprecedented coverage of most major variables, yielding a broad picture of C cycling in the world's major forest biomes.
 Carbon cycling rates generally increased from boreal to tropical regions and with stand age.
@@
 In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs. 2-5, `r stock_age_trends`).
-The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Table 1; Figs. `r flux_age_trends`- `r stock_age_trends`, `r all_age_trends`).
+The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Table 1; Figs. `r flux_age_trends`- `r BoN_age_trends`, `r all_age_trends`).
 Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks and $NEP$ of mature forests are correlated with a different set of factors.
+Together, these results refine and expand out understanding of C cycling in mature forests, while providing the first global-scale analysis of age trends in multiple forest C cycling stocks and fluxes (Figs.
+`r TrB_age_trends`-`r BoN_age_trends`).
 
@@
 
-The large number of C cycle variables covered by ForC, and the general consistency between them, provide confidence that our overall reported means provide accurate and useful baselines for analysis -- with the caveats that they are unlikely to be accurate representations of C cycling for any particular forest, and that these sample means almost certainly do not represent true biome means (particularly for temperate conifer forests where high-biomass stands are over-represented in *ForC*).
+The large number of C cycle variables covered by ForC, and the general consistency among them, provide confidence that our overall reported mature forest means provide accurate and useful baselines for analysis -- with the caveats that they are unlikely to be accurate representations of C cycling for any particular forest, and that these sample means almost certainly do not represent true biome means (particularly for temperate conifer forests where high-biomass stands are over-represented in *ForC*).
 
@@
 `r diagram_TeN_mature`) than of methodological accuracy.
-The overall high degree of closure implies that *ForC* gives a consistent picture of C cycling within biomes. 
+The overall high degree of closure implies that *ForC* gives a consistent picture of C cycling within biomes for mature forests. 
 This is an important and useful test, because it allows for consistency checks within the C cycle, for example leveraging separate and independently-measured fluxes to constrain errors in another [@phillips_value_2017; @williams_impacts_2014; @harmon_heterotrophic_2011], or producing internally consistent global data products [@wang_golum-cnp_2018].
+
+In contrast, age trends for young forests generally remain less clearly defined, in large part because their data records remain somewhat sparse for most variables (*i.e.*, have low representation of different geographical regions for any given age).
+While this analysis provides a first analysis of age trends in forest C cycling for multiple variables at a global scale, improved resolution of these trends will require additional data.
 
@@
 
-Our study reveals that most C fluxes quickly increase to a plateau as stands age (Fig.
-`r flux_age_trends`), consistent with current understanding of age trends in forest C cycling [e.g., @andersonteixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
+Our study reveals that most C fluxes quickly increase to a plateau as stands age (Figs.
+`r flux_age_trends`, `r TrB_age_trends`-`r BoN_age_trends`), consistent with current understanding of age trends in forest C cycling [Fig.
+`r schematic`; e.g., @andersonteixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
 While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_{auto}$) would be minimal immediately following a stand-clearing disturbance.
@@
 
+###  Citations to add
+
+@harris_global_2021
+
 ## Acknowledgements

### 2021-01-21 manuscript/tables_figures/stacked_age_trends/schematic.m

@@ -32,12 +32,13 @@ biome_names={'tropical', 'temperate', 'boreal'};
 %for stacked plot
 NEP=2.4*log10(age)-.04*(age-1); %placeholder (peaks around 2 at 23 yrs age)
 GPP= NEP + (1.99+1.44)* log10(age);
-ANPP_foliage = 1.66* log10(age) +.18;
-ANPP_woody = 2.31* log10(age) -.76;
+ANPP_foliage = 1.66* log10(age) +.18 - .01*(age-1);
+ANPP_woody = 2.31* log10(age) -.76 - .02*(age-1);
 ANPP_woody_turnover=1.7*log10(age)-1; %max(0,ANPP_woody-NEP);
-BNPP=0.97*log10(age)+.67;
+BNPP=0.97*log10(age)+.67 - .01*(age-1);
 NPP=ANPP_foliage+ANPP_woody+BNPP;
-R_auto = NPP.*(1+.001*age); %insufficient data
+CUE=0.679-0.153*log10(age); %from DeLucia et al. 2007 (note that equation given mistakenly leaves of the log10(age)). Collati et al. 2020 gives an update of this.
+R_auto= max(0,NPP.*(1./CUE-1)); %Rauto=NPP*(1/CUE-1), and CUE equation is above
 R_het=ANPP_foliage+ANPP_woody_turnover+BNPP;
 
 

### 2021-01-21 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m

@@ -24,6 +24,7 @@ end
 %prepare some variables
 variables = unique(model_summaries.Variable);
 biomes= {'Tropical broadleaf','Temperate broadleaf', 'Temperate conifer', 'Boreal conifer'};
+biome_codes= {'TrB','TeB', 'TeN', 'BoN'};
 %{'BiomeTropical broadleaf','BiomeTemperate broadleaf','BiomeTemperate conifer','BiomeBoreal conifer'}
 params_matrix=cell2table(num2cell(NaN*ones(length(variables), 10)),'VariableNames',{'variable', 'beta' 'betaTrB' 'betaTeB' 'betaTeN' 'betaBoN' 'intTrB' 'intTeB' 'intTeN' 'intBoN'});
 params_matrix.variable=variables;
@@ -345,7 +346,7 @@ facecolor_in_fluxes= [0.3 0 1;...  %'R_{auto-ag}*'
                     0.1 0 .4;...  % 'R_{root}', 
                     0 .7 1;...  %'BNPP'
                     0 .527 .27;... %'ANPP_{foliage}',
-                    170/255 1 0 ]...%'ANPP_{woody}*'
+                    170/255 1 0 ];...%'ANPP_{woody}*'
                 
 facecolor_out_fluxes= [0.3 0 1;...  %'R_{auto-ag}*'
                     0.1 0 .4;...  % 'R_{root}', 
@@ -397,7 +398,7 @@ legend (stock_names, 'Location', 'BestOutside');
 
 %% ~~~~~~~ SAVE FIGURE~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 cd(working_dir)
-savename=char(strcat(biomes(b),'_age_trends'));
+savename=char(strcat(biome_codes(b),'_age_trends'));
 print(savename, '-dpng', '-r300')
 
 

### 2021-01-22 manuscript/ERL_global_C_review.Rmd
incorporate @bpbond 's comments

@@
 
-*Review Results/ Synthesis.*  *ForC v3.0* yielded a comprehensive picture of C cycling in the world's major forest biomes.
-The rate of C cycling generally increased from boreal to tropical regions in both mature and regrowth forests, whereas C stocks showed less directional variation.
+*Review Results/ Synthesis.*   The rate of C cycling generally increased from boreal to tropical regions in both mature and regrowth forests, whereas C stocks showed less directional variation.
 The majority of flux variables, together with most live biomass pools, increased significantly with stand age.
@@
 *Discussion.* As climate change accelerates, understanding and managing the carbon dynamics of forests is critical to forecasting, mitigation, and adaptation.
-This synthetic global overview of C stocks and fluxes across biomes and stand ages will help to advance these efforts.
+This comprehensive and synthetic global overview of C stocks and fluxes across biomes and stand ages will help to advance these efforts.
 
@@
 A robust understanding of forest impacts on global C cycling is essential. 
-Annual gross CO~2~ sequestration in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
-Most of this enormous C sequestration is counterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] and fire emissions [@van_der_werf_global_2017].
-In recent years, the remaining CO~2~ sink averaged 3.2 $\pm$ 0.6 GtC yr^-1^ from 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
+Annual gross sequestration in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
+Most of this enormous C sequestration is counterbalanced by releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] and fire emissions [@van_der_werf_global_2017].
+In recent years, the remaining C sink averaged 3.2 $\pm$ 0.6 Gt C yr^-1^ from 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
 However, deforestation, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reduces the net forest sink to ~1.1-2.2 Gt C yr^-1^ across Earth's forests [@friedlingstein_global_2019].
@@
 Over this time, an increasingly refined conceptual understanding of the elements of ecosystem C cycles has developed, as a growing number of variables have been defined [e.g., @chapin_reconciling_2006], along with appropriate measurement methods [e.g., @clark_measuring_2001].
-New technology has also enabled researches to directly measure an expanding set of variables, notably including the development of continuous measurements of soil CO~2~ efflux (**REF-BEN**) and ecosystem-atmosphere CO~2~ exchange (**REFS**).
-Standardization of estimation techniques has also increased; for example, there has been increasing standardization of the biomass allometries strongly influence estimates of any C cycle variable including woody biomass or changes therein [e.g., @chojnacky_updated_2014; @chave_improved_2014].
+New technology has also enabled researchers to directly measure an expanding set of variables, notably including the development of continuous measurements of soil CO~2~ efflux [@kuzyakov_sources_2006] and ecosystem-atmosphere CO~2~ exchange [@baldocchi_fluxnet_2001].
+Measurement techniques have been increasingly standardized; for example, of the biomass allometries that strongly influence estimates of most C cycle variables [e.g., @chojnacky_updated_2014; @chave_improved_2014].
 Further standardization has been made possible through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; @davies_forestgeo_2021], NEON [@schimel_neon_2007], and FLUXNET [@baldocchi_fluxnet_2001; @novick_ameriflux_2018].
@@
 Since the early 19th century, it has been recognized that climate plays a dominant role in shaping differences among forests on a global scale [@humboldt_essay_1807; @holdridge_determination_1947].
-Global scale data synthesis has revealed that C fluxes including $GPP$, net primary productivity ($NPP$), and soil respiration ($R_{soil}$) decrease with latitude or, correspondingly, increase with mean annual temperature and, to a lesser extent, precipitation [Fig.
-`r schematic_fluxes_biome` **REFS**; @lieth_primary_1973; @luyssaert_co2_2007; @banbury_morgan_global_nodate].
+Global scale data syntheses have shown that C fluxes including $GPP$, net primary productivity ($NPP$), and soil respiration ($R_{soil}$) decrease with latitude or, correspondingly, increase with mean annual temperature and, to a lesser extent, precipitation [Fig.
+`r schematic_fluxes_biome` **REFS**; @lieth_primary_1973;  @luyssaert_co2_2007; @hursh_sensitivity_2017; @banbury_morgan_global_nodate].
 C stocks of mature forests show less directional variation (Fig. `r schematic_fluxes_biome`).
 On average, $B_{ag}$ tends to decrease with latitude, but not as dramatically as fluxes, and with the highest $B_{ag}$ forests in relatively cool, moist temperate regions [**REFS**; @keith_re-evaluation_2009].
-In contrast, standing and downed dead wood ($DW_{standing}$ and $DW_{down}$, respectively) and the organic layer ($OL$) tend to accumulate more in colder climates where decomposition is slow relative to $NPP$ [**REFS**].
+In contrast, standing and downed dead wood ($DW_{standing}$ and $DW_{down}$, respectively) and the organic layer ($OL$) tend to accumulate more in colder climates where decomposition is slow relative to $NPP$ [@harmon_ecology_1986; @allen_global_2002].
 
-Correlative analyses relating C cycle variables to climate and other environmental variables have recently been taken to a new level through use of machine-learning algorithms that relate ground-based C cycle data to *global data on climate/soils/satellite data*, making it possible to create fine-scale global maps of C cycling [e.g., **REFS**; @cook-patton_mapping_2020].
+Correlative analyses relating C cycle variables to climate and other environmental variables have recently been taken to a new level through use of machine-learning algorithms that relate ground-based C cycle data to *global data on climate/soils/satellite data*, making it possible to create fine-scale global maps of C cycling [e.g., **REFS**;@warner_spatial_2019; @cook-patton_mapping_2020].
 *This approach can be particularly effective when paired with satellite data ...
 (e.g., aboveground biomass: Saatchi et al 2011, Hu et al 2016, Spawn et al 2020, gross primary productivity, GPP: Li and Xiao 2019, ).*
-Yet, any such analysis is constrained by the quality and coverage of ground-based estimates of forest C fluxes or stocks. 
+Any such analysis is however constrained by the quality and coverage of ground-based estimates of forest C fluxes or stocks. 
 While estimates of some variables (e.g., $B_{ag}$, $GPP$, $NPP$, $R_{soil}$) are widely available, many remain poorly characterized (e.g., $DW$; autotrophic respiration, $R_{auto}$) --even at the coarse resolution of biomes.
@@
 In 1969, E.P. Odum's "The Strategy of Ecosystem Development" laid out predictions as to how forest energy flows and organic matter stocks vary with stand age  [@odum_strategy_1969].
-Although the conceptualization of the C cycle in this paper is simplistic by current standards, the paper was foundational in framing the theory around which research on the subject still revolves [@corman_foundations_2019], and the basic framework still holds, *albiet* with modest modifications [@andersonteixeira_altered_2013].
+Although the conceptualization of the C cycle in this paper is simplistic by current standards, the paper was foundational in framing the theory around which research on the subject still revolves [@corman_foundations_2019], and the basic framework still holds, albeit with modest modifications [@andersonteixeira_altered_2013].
 Following stand-clearing disturbance, $GPP$, $NPP$, and biomass of leaves ($B_{foliage}$) and fine roots ($B_{root-fine}$) increase rapidly and thereafter remain relatively stable ($B_{foliage}$,$B_{root-fine}$, sometimes $GPP$) or decline slightly [$NPP$, sometimes $GPP$; e.g., *Goulden et al.
@@
 
-In the past *couple of/several* decades, researchers have started asking how age trends--mostly in $B_{ag}$ or total biomass ($B_{tot}$) accumulation-- vary across biomes.
+In the past several decades, researchers have started asking how age trends--mostly in $B_{ag}$ or total biomass ($B_{tot}$) accumulation-- vary across biomes.
 Early research on this theme showed that biomass accumulation rates during secondary succession increase with temperature on a global scale [**REFS**; @anderson_temperature-dependence_2006] and *with precipitation in the neotropics* [**REFS**; @chazdon_carbon_2016].

### 2021-01-22 manuscript/response_to_reviews.Rmd
working on specific reviewer comments

@@
 
-*This just requires adding some explanation.*
+**In addition to addressing the specific questions below, we have added a link to a spreadsheet recording the datasets that have been incorporated, which is maintained in the open-access GitHub repository: "A record of data sets added to *ForC* is available at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv."**
 
@@
 
-*This just requires adding some explanation.*
+**We now clarify as follows:**
+
+**FLUXNET:**
+
+**Methods: "Third, we manually incorporated records of annual $NEP$, $GPP$, and $R_{eco}$ from the FLUXNET2015 dataset [@pastorello_fluxnet2015_2020], treating these records as authoritative when they duplicated earlier records (Appendix S1)."**
+
+**Appendix S1: "For eddy-covariance variables ($NEP$, $GPP$, $R_{eco}$), we retained the record associated with the most recent publication [most often @pastorello_fluxnet2015_2020], as these data are commonly re-assessed using new analysis methods."**
+
+**SRDB:**
+
+**Methods: "First, we imported (via R script) the Global Database of Soil Respiration Database [*SRDB* v4, `r n_SRDB` records; @bond-lamberty_global_2010], and corrections and improvements to *SRDB* arising from this process were incorporated in *SRDB* v5 [@jian_restructured_2020]."**
 
@@
 
-*This just requires adding some explanation.*
+**All records analyzed here have been fully incorporated into the ForC format and treated as described.
+We believe this should now be clearer where we describe the importation of new datasets: "Since publication of ForC v2.0, we imported three large additional databases [SRDB, GROA, FLUXNET15] into *ForC* via a combination of R scripts and manual edits.
+..."**
 
@@
 
-*This just requires some re-wording. We are building on/ updating results of Luyssaert.*
+**We don't see this as circular; we are reaffirming earlier results with an expanded dataset.
+To make it clear that we are not claiming to provide an independent test, we have edited this text as follows: "For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude (or increase with temperature) on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
+The consistency with which this occurs across numerous fluxes is not surprising, particularly given commonality in the data analyzed or used for calibration, but has never been simultaneously assessed across such a large number of variables [but see @banbury_morgan_global_nodate for nine autotrophic fluxes]."**
 

### 2021-01-22 manuscript/ERL_global_C_SI.Rmd
working on specific reviewer comments

@@
 If duplicates remained (within *ForC v2.0* or *SRDB*), we selected a record for retention based on citation year, keeping the oldest (presumably original) citations for most variables. 
-For eddy-covariance variables ($NEP$, $GPP$, $R_{eco}$), we retained the record associated with the most recent publication, as these data are commonly re-assessed using new analysis methods.
+For eddy-covariance variables ($NEP$, $GPP$, $R_{eco}$), we retained the record associated with the most recent publication [most often @pastorello_fluxnet2015_2020], as these data are commonly re-assessed using new analysis methods.
 If publication years were the same, we randomly selected one record.

### 2021-01-22 manuscript/ERL_global_C_review.Rmd
working on specific reviewer comments

@@
 Hember et al., 2012]*.
-While these trends have been subject of fairly recent review [@andersonteixeira_altered_2013], *there is need for a synthetic, quantitaitive review.*
+While these trends have been subject of fairly recent review [@andersonteixeira_altered_2013], *there is need for a synthetic, quantitative review.*
 
@@
 The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age.
-As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible [83% of records in v.2.0; @anderson-teixeira_forc_2018].
+As such, there has been a focus on incorporating data from regrowth forests [*e.g.*, @anderson_temperature-dependence_2006; @martin_carbon_2013; @bonner_meta-analytical_2013] and obtaining stand age data when possible [83% of records in v.2.0; @anderson-teixeira_forc_2018].
 Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], which represented roughly one-third of records in *ForC* v2.0 [@anderson-teixeira_forc_2018].
-Since publication of ForC v2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [*SRDB* v4, `r n_SRDB` records; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database [*GROA* v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020].
+Since publication of ForC v2.0, we imported three large additional databases into *ForC* via a combination of R scripts and manual edits. 
+First, we imported (via R script) the Global Database of Soil Respiration Database [*SRDB* v4, `r n_SRDB` records; @bond-lamberty_global_2010], and corrections and improvements to *SRDB* arising from this process were incorporated in *SRDB* v5 [@jian_restructured_2020].
+Second, we imported (via R script) the Global Reforestation Opportunity Assessment database [*GROA* v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020], which itself had drawn on an earlier version of *ForC*.
+Because all records in *GROA* were checked against original publications, these records were given priority over duplicates in *ForC*.
+Third, we manually incorporated records of annual $NEP$, $GPP$, and $R_{eco}$ from the FLUXNET2015 dataset [@pastorello_fluxnet2015_2020], treating these records as authoritative when they duplicated earlier records (Appendix S1).
 We have also added data from individual publications, with a particular focus on productivity [e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @lutz_global_2018; @johnson_climate_2018].
+A record of data sets added to *ForC* is available at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv.
 The database version used for this analysis has been tagged as a new release on Github (v3.0) and assigned a DOI through Zenodo (DOI: TBD).
@@
 For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude (or increase with temperature) on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
-The consistency with which this occurs across numerous fluxes is not surprising, but has never been simultaneously assessed across such a large number of variables [but see @banbury_morgan_global_nodate for nine autotrophic fluxes].
+The consistency with which this occurs across numerous fluxes is not surprising, particularly given commonality in the data analyzed or used for calibration, but has never been simultaneously assessed across such a large number of variables [but see @banbury_morgan_global_nodate for nine autotrophic fluxes].
 

### 2021-01-22 manuscript/response_to_reviews.Rmd
working on responses to reviews

@@
 
-**Methods: "First, we imported (via R script) the Global Database of Soil Respiration Database [*SRDB* v4, `r n_SRDB` records; @bond-lamberty_global_2010], and corrections and improvements to *SRDB* arising from this process were incorporated in *SRDB* v5 [@jian_restructured_2020]."**
+**Methods: "First, we imported (via R script) the Global Database of Soil Respiration Database [*SRDB* v4, 9488 records; @bond-lamberty_global_2010], and corrections and improvements to *SRDB* arising from this process were incorporated in *SRDB* v5 [@jian_restructured_2020]."**
 
@@
 
-*This just requires adding some explanation.*
+**We acknowledge and discuss lack of closure in various places throughout the results and discussion.
+To ensure that this is clear from the figures, we have added the following statement to the captions of Figures 3-6 (formerly 2-5): "Note that variables differ in geographical representation, resulting in potential imbalances (Figs.
+S5-S30).
+Probability that estimates reflect true biome means scales with the number of distinct geographical areas represented." We have also added a similar statement to the captions of Figures S1-S4 (formerly 8-11).
+We also make a related statement in the captions of new figures 9-12: "Note that there remain substantial uncertainties as to the functional form of age trends and discrepancies in closure among related variables."**
 
@@
 
-*This just requires adding some explanation.*
+**This is partially addressed by the response above, and covered in the results and discussion.
+In addition, to help explain the fact that lack of closure is particularly pronounced in temperate conifer forests, we have included errorbars in on the new schematic figure (Fig.
+1) indicating that temperate forests have particularly high variation in C stocks.**
 
@@
 
-*This just requires adding some explanation. Note that R2 and R3 have very different interpretations of the "closure"*
+**We have addressed this by adding the following statement to the captions of Figures 3-6 (formerly 2-5): "Probability that estimates reflect true biome means scales with the number of distinct geographical areas represented."**
 
 3. There is a general sloppiness to the manuscript that makes me think the authors did not proofread the manuscript carefully.
-
-*This is true. Need to fix all points below. None difficult.*
 
@@
 
+**We have added an explanation.**
+
 d. Line 162: Clarify what is meant by "the minimum diameter breast height (DBH) threshold for tree census was <10cm"?
+
+**We have reworded this sentence for improved clarity: "Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), trees were censused down to a minimum diameter breast height (DBH) threshold of 10 cm or less."**
 
@@
 The table uses "BoN", but the note refers to "B".
+
+**We have fixed both of these.**
 

### 2021-01-22 manuscript/ERL_global_C_review.Rmd
working on responses to reviews

@@
 
+schematics_caption_generic ="Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
+Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
+Note that variables differ in geographical repressentation, resulting in potential imbalances (Figs.
+S5-S30).
+Probability that estimates reflect true biome means scales with the number of distinct geographical areas represented.
+Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
+To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.
+Asterisk after variable name indicates lack of C cycle closure."
+TrB_mature_caption = paste( "Figure 3 | C cycle diagram for mature tropical broadleaf forests. ", schematics_caption_generic)  
+TeB_mature_caption = paste( "Figure 4 | C cycle diagram for mature temperate broadleaf forests. ", schematics_caption_generic)  
+TeN_mature_caption = paste( "Figure 5 | C cycle diagram for mature temperate conifer forests. ", schematics_caption_generic)  
+BoN_mature_caption = paste( "Figure 6 | C cycle diagram for mature boreaal conifer forests. ", schematics_caption_generic)
+
 TrB_age_trends_caption = "Figure 9 | Age trends in C cycling in tropical broadleaf forests.
@@
 Asterisks indicate variables whose age trends were calculated based on other variables (* young and mature forests; ** young forests only; *** mature forests only), as follows: $R_{auto-ag}=R_{auto}-R{root}$, where $R_{auto}=NPP(1/CUE-1)$, where $CUE$=0.46 (Collati et al.
-2020); $ANPP_{woody}=max(0,ANPP-ANPP_{foliage})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
+2020); $ANPP_{woody}=max(0,ANPP-ANPP_{foliage})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$.
+Note that there remain substantial uncertainties as to the functional form of age trends and discrepancies in closure among related variables."
 age_trends_caption_generic= "Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
-Asterisks indicate variables whose age trends were calculated based on other variables (* young and mature forests; ** young forests only; *** mature forests only), as follows: $R_{auto-ag}=R_{eco}-R{soil}$; $ANPP_{woody}=min(ANPP_{stem},ANPP_{woody})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$."
+Asterisks indicate variables whose age trends were calculated based on other variables (* young and mature forests; ** young forests only; *** mature forests only), as follows: $R_{auto-ag}=R_{eco}-R{soil}$; $ANPP_{woody}=min(ANPP_{stem},ANPP_{woody})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$.
+Note that there remain substantial uncertainties as to the functional form of age trends and discrepencies in closure among related variables."
 TeB_age_trends_caption = paste("Figure 10 | Age trends in C cycling in temperate broadleaf forests.", age_trends_caption_generic) 
@@
 
----
-nocite: | 
-@collalti_forest_2020
----
-
 \newpage
@@
 C stocks of mature forests show less directional variation (Fig. `r schematic_fluxes_biome`).
-On average, $B_{ag}$ tends to decrease with latitude, but not as dramatically as fluxes, and with the highest $B_{ag}$ forests in relatively cool, moist temperate regions [**REFS**; @keith_re-evaluation_2009].
+On average, $B_{ag}$ tends to decrease with latitude, but not as dramatically as fluxes, and with the highest $B_{ag}$ forests in relatively cool, moist temperate regions [@keith_re-evaluation_2009; @smithwick_potential_2002; @hu_mapping_2016].
 In contrast, standing and downed dead wood ($DW_{standing}$ and $DW_{down}$, respectively) and the organic layer ($OL$) tend to accumulate more in colder climates where decomposition is slow relative to $NPP$ [@harmon_ecology_1986; @allen_global_2002].
@@
 Symbols are colored according to the number of records at each site.
-Underlying map shows coverage of evergreen, deciduous, and mixed forests (from SYNMAP; Jung et al.
-2006) and biomes.
+Underlying map shows coverage of evergreen, deciduous, and mixed forests (shading differences; data from Jung et al.
+2006) and biomes (color differences).
 Distribution of sites, plots, and records among biomes is shown in the inset.", out.width='100%', fig.pos='H'}
@@
 ```
+---
+nocite: | 
+@jung_exploiting_2006
+---
 
@@
 Furthermore, for $NPP$, aboveground $NPP$ ($ANPP$), and the litterfall component of $ANPP$ ($ANPP_{litterfall}$), *ForC* variables specifying inclusion of different components were combined (*e.g.*, measurements including or excluding fruit and flower production and herbivory).
-Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum diameter breast height (DBH) threshold for tree census was $\le$ 10cm. 
+Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), trees were censused down to a minimum diameter breast height (DBH) threshold of 10 cm or less. 
 All records were measured directly or derived from field measurements (as opposed to modeled).
@@
   pack_rows("Stocks", 24, 34, colnum = 2, hline_before = FALSE) %>%
-kableExtra::footnote(symbol = c("Tr: Tropical, TeB: Temperate Broadleaf, TeN: Temperate Needleleaf, B: Boreal, n.s.: no significant differences, n.t.: not tested ", "+ or -: significant positive or negative trend, xB: significant age x biome interaction, n.s.: no significant age trend, n.t.: not tested"), footnote_as_chunk = FALSE)
+kableExtra::footnote(symbol = c("TrB: Tropical, TeB: Temperate Broadleaf, TeN: Temperate Needleleaf, BoN: Boreal, n.s.: no significant differences, n.t.: not tested ", "+ or -: significant positive or negative trend, xB: significant age x biome interaction, n.s.: no significant age trend, n.t.: not tested"), footnote_as_chunk = FALSE)
 ```
@@
 \blandscape
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 3 | C cycle diagram for mature tropical broadleaf forests.
-Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.
-Asterisk after variable name indicates lack of C cycle closure.", out.width='90%', fig.pos='H'}
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = TrB_mature_caption, out.width='90%', fig.pos='H'}
 
@@
 
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 4 | C cycle diagram for mature temperate broadleaf forests.
-Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.
-Asterisk after variable name indicates lack of C cycle closure.", out.width='100%', fig.pos='H'}
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = TeB_mature_caption, out.width='100%', fig.pos='H'}
 
@@
 
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 5 | C cycle diagram for mature temperate conifer forests.
-Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.
-Asterisk after variable name indicates lack of C cycle closure.", out.width='100%', fig.pos='H'}
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = TeN_mature_caption, out.width='100%', fig.pos='H'}
 
@@
 
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 6 | C cycle diagram for mature boreal conifer forests.
-Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
-Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.
-Asterisk after variable name indicates lack of C cycle closure.", out.width='100%', fig.pos='H'}
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = BoN_mature_caption, out.width='100%', fig.pos='H'}
 
@@
 
-The large number of C cycle variables covered by ForC, and the general consistency among them, provide confidence that our overall reported mature forest means provide accurate and useful baselines for analysis -- with the caveats that they are unlikely to be accurate representations of C cycling for any particular forest, and that these sample means almost certainly do not represent true biome means (particularly for temperate conifer forests where high-biomass stands are over-represented in *ForC*).
+The large number of C cycle variables covered by ForC, and the general consistency among them, provide confidence that our overall reported mature forest means provide useful baselines for analysis -- with the caveats that they are unlikely to be accurate representations of C cycling for any particular forest, and that these sample means almost certainly do not represent true biome means (particularly for temperate conifer forests where high-biomass stands are over-represented in *ForC*).
 
@@
 For aboveground biomass, which is the variable in *ForC* with broadest geographical representation, the modest trend of declining biomass from tropical to boreal regions mirrors observations from spaceborne lidar that reveal a decline in aboveground biomass (for all forests, including secondary) with latitude across the N hemisphere [@hu_mapping_2016].
-The highest- biomass forests on Earth are, however, found in coastal temperate climates of both the southern and northern hemisphere [@keith_re-evaluation_2009; @smithwick_potential_2002; @hu_mapping_2016].
+The highest- biomass forests on Earth are, however, found in coastal temperate climates of both the southern and northern hemisphere [Figs.
+`r schematic`, `r stock_age_trends`a; @keith_re-evaluation_2009; @smithwick_potential_2002; @hu_mapping_2016].
 Disproportionate representation of forests in one such region--the US Pacific Northwest--inflated estimates of temperate conifer fluxes and stocks for some variables and was responsible for all of the anomalous results described here (e.g., lack of complete C budget closure, anomalous trend across biomes for $BNPP_{coarse}$).

### 2021-01-22 manuscript/tables_figures/stacked_age_trends/schematic.m

@@ -81,6 +81,7 @@ Stock_ratio_boreal_trop=0.7;
 
 mature_stocks = [stocks(:,end)'; Stock_ratio_temp_trop*stocks(:,end)' ; Stock_ratio_boreal_trop*stocks(:,end)'];
 
+stocks_std = sum(mature_stocks, 2)'.*[.4 1 .4];
 
 %% ~~~~~~~PLOTTING~~~~~~~~~~
 
@@ -134,10 +135,13 @@ ylim([0, max(in_sum)+1]);
 legend ([{'R_{het}'}, in_flux_names, {'GPP' 'Reco'}], 'Location', 'BestOutside');
 
 subplot (2,3,4)
-h=bar(mature_stocks, 'stacked');
+h=bar(mature_stocks, 'stacked'); hold on;
     for n=1:size(mature_stocks,2)
         h(n).FaceColor= facecolor_stocks(n,:);
     end
+er=errorbar([1:3],sum(mature_stocks, 2), 0.5*stocks_std, stocks_std);
+er.Color = [0 0 0];                            
+er.LineStyle = 'none'; 
 ylabel ('C  stocks (Mg C ha^{-1})')
 set(gca, 'XTickLabel', {'Tropical' 'Temperate' 'Boreal'})
 xtickangle(30)

### 2021-01-22 manuscript/response_to_reviews.Rmd
finished responses to R2

@@
 
+**It appears that reference to ForC-simplified (essentially an automatically generated rearrangement of ForC that is purged of duplicates and organized for easy analysis) makes the process more confusing, and isn't necessary, so we eliminated mention of it.**
+
 g. Line 212: The authors refer to ForC, where in the previous sentence they referred to ForC v3.0, and previously used ForC-simplified. Clarify what database is used in the analyses.
+
+**As in the response above, we have eliminated unnecessary/confusing mention of ForC-simplified.
+The version used is ForC v3.0 (to be released with DOI upon acceptance of this manuscript).
+We do not feel it is necessary to state the version number every time ForC is mentione.**
 
@@
 
-*This requires careful consideration and response*
+**We'd define this as a "quantitative review"--*i.e.*, one that quantitatively synthesizes previous studies.
+However, we agree that the previous version did little to critically review the corresponding literature.
+To improve in this area, we have expanded and reorganized the introduction into four sub-sections ("Forests in the global C cycle: current and future", "Evolution of forest C cycle research", "Biome differences", and "Age trends and their variation across biomes").
+The latter three sections are mostly or entirely new relative to the previous version, and briefly review the development and current status of the science on each topic.
+We have also added a schematic figure (Fig.
+1) to summarize current understanding of each topic and help with conceptual framing of the review.
+We also shifted emphasis more towards the ecological questions (trends across biomes and with age), while retaining (with less emphasis) the analysis of database representativeness.**
 

### 2021-01-22 manuscript/ERL_global_C_review.Rmd
finished responses to R2

@@
 
-To facilitate analyses, we created a simplified version of ForC, *ForC-simplified* (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which we analyzed here.
-In generating *ForC-simplified*, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ [@ipcc_global_2018].
-Duplicate or otherwise conflicting records were reconciled as described in Appendix S1, resulting in a total of `r ForC_simplified_duplicates_removed` records (`r round(ForC_simplified_duplicates_removed/nrow(ForC_measurements)*100,digits=1)`% size of total database).
+All measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ [@ipcc_global_2018].
+Duplicate or otherwise conflicting records were purged as described in Appendix S1, resulting in a total of `r ForC_simplified_duplicates_removed` records (`r round(ForC_simplified_duplicates_removed/nrow(ForC_measurements)*100,digits=1)`% size of total database).
 Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand initiation event.
-Specifically, we removed all plots flagged as managed in ForC-simplified (`r round(n_managed_in_ForC_simplified/ForC_simplified_duplicates_removed*100,digits=1)`%).
-This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.
-Plots flagged as disturbed in ForC-simplified (`r round(n_disturbed_in_ForC_simplified/ForC_simplified_duplicates_removed*100,digits=1)`%) included stands that had undergone any notable anthropogenic thinning or partial harvest.
+Specifically, we removed plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized” (`r round(n_managed_in_ForC_simplified/ForC_simplified_duplicates_removed*100,digits=1)`% of duplicate-purged records).
+We also removed stands that had undergone any notable anthropogenic thinning or partial harvest (`r round(n_disturbed_in_ForC_simplified/ForC_simplified_duplicates_removed*100,digits=1)`% of duplicate-purged records).
 We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods.

### 2021-01-22 manuscript/response_to_reviews.Rmd
worked on response to reviews

@@
 The version used is ForC v3.0 (to be released with DOI upon acceptance of this manuscript).
-We do not feel it is necessary to state the version number every time ForC is mentione.**
+We do not feel it is necessary to state the version number every time ForC is mentioned.**
 
@@
 
-**Thank you**
+**Thank you.**
 
@@
 
-(*add some broad-brush hypotheses/ frameworks*)
+**We agree that this covers a lot of ground, which corresponds with our vision of producing a comprehensive synthesis of broad trends.
+We envision that smaller questions can (and hopefully will) be broken off for more detailed analysis in the future.**
+
+**At the same time, we agreed that the manuscript could be improved with more of a conceptual framework.
+To improve in this area, we have expanded and reorganized the introduction into four sub-sections ("Forests in the global C cycle: current and future", "Evolution of forest C cycle research", "Biome differences", and "Age trends and their variation across biomes").
+The latter three sections are mostly or entirely new relative to the previous version, and briefly review the development and current status of the science on each topic.
+We have also added a schematic figure (Fig.
+1) to summarize current understanding of each topic and help with conceptual framing of the review.
+We also shifted emphasis more towards the ecological questions (trends across biomes and with age), while retaining (with less emphasis) the analysis of database representativeness.**
 
@@
 
-**Regarding reference to Odum, ...**, *add citation to Odum*
+**We greatly appreciated the suggestion to engage with Odum's framework.
+We now cite and discuss @odum_strategy_1969 in the introduction, and designed the age trends portion of the schematic figure to correspond to Odum's classic Fig.
+1 (updated with more details and current terminology).**
 
@@
 
-**Regarding the NEP result for mature forests, ...** *(Note that R4 also found this surprising. I think an Odum-esque diagram can help clarify.)*
+**Regarding the NEP result for mature forests, we agree that a better conceptual framework (now implemented) should help set the stage for explaining this result.
+We have modified the discussion paragraph on this to cite the schematic figure and also to note that this result does agree with @luyssaert_co2_2007 (although that study did not standardize for age).
+The paragraph now reads as follows:**
+
+**"The notable exception to the pattern of fluxes decreasing from tropical to boreal regions is $NEP$, which showed no significant differences across biomes (Fig.
+6f).
+Unlike the other C flux variables, $NEP$ does not characterize the rate at which C cycles through the ecosystem, but is the balance between C sequestration ($GPP$) and respiratory losses ($R_{eco}$) and represents net CO~2~ sequestration (or release) by the ecosystem (Fig.
+1).
+$NEP$ tends to be relatively small in mature forest stands (discussed further below), which accumulate carbon slowly relative to younger stands, if at all [@luyssaert_old-growth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
+It is therefore unsurprising that there are no pronounced differences across biomes [in agreement with @luyssaert_co2_2007], suggesting that variation in $NEP$ of mature forests is controlled less by climate and more by other factors including moderate disturbances [@curtis_forest_2018] or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition; @wilson_stability_2016]."**
 
@@
 
-**This is a good point, but unlikely to have much influence at this relatively coarse scale of analysis. ...**
+**This is an excellent question, and one that is of great interest to the lead author [e.g., @andersonteixeira_altered_2013]; however, we think it is unlikely to have much influence at this relatively coarse scale of analysis, especially given that much of the data in ForC was collected in the 20th century.
+Sorting this out would require carefully designed studies, and probably would not be possible using ForC data at this point.**
 

### 2021-01-22 manuscript/ERL_global_C_review.Rmd
worked on response to reviews

@@
 
-The notable exception to the pattern of fluxes decreasing from tropical to boreal regions is $NEP$ (Fig. `r flux_age_trends`), which showed no significant differences across biomes.
-Unlike the other C flux variables, $NEP$ does not characterize the rate at which C cycles through the ecosystem, but is the balance between C sequestration ($GPP$) and respiratory losses ($R_{eco}$) and represents net CO~2~ sequestration (or release) by the ecosystem.
+The notable exception to the pattern of fluxes decreasing from tropical to boreal regions is $NEP$, which showed no significant differences across biomes (Fig. `r flux_age_trends`f).
+Unlike the other C flux variables, $NEP$ does not characterize the rate at which C cycles through the ecosystem, but is the balance between C sequestration ($GPP$) and respiratory losses ($R_{eco}$) and represents net CO~2~ sequestration (or release) by the ecosystem (Fig.
+`r schematic`).
 $NEP$ tends to be relatively small in mature forest stands (discussed further below), which accumulate carbon slowly relative to younger stands, if at all [@luyssaert_old-growth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
-It is therefore unsurprising that there are no pronounced differences across biomes, suggesting that variation in $NEP$ of mature forests is controlled less by climate and more by other factors including moderate disturbances [@curtis_forest_2018] or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition; @wilson_stability_2016].
+It is therefore unsurprising that there are no pronounced differences across biomes [in agreement with @luyssaert_co2_2007], suggesting that variation in $NEP$ of mature forests is controlled less by climate and more by other factors including moderate disturbances [@curtis_forest_2018] or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition; @wilson_stability_2016].
 

### 2021-01-22 manuscript/ERL_global_C_review.Rmd
worked on HM feedback

@@
 
-*Background.* Earth’s climate is closely linked to forests, which strongly influence atmospheric carbon dioxide (CO~2~) and climate by their impact on the global carbon (C) cycle.
+*Background.* Forests are major components of the global carbon (C) cycle and thereby strongly influence atmospheric carbon dioxide (CO2) and climate.
 However, efforts to incorporate forests into climate models and CO~2~ accounting frameworks have been constrained by a lack of accessible, global-scale synthesis on how C cycling varies across forest types and stand ages.
@@
 *Methods/Design.* Here, we draw from the Global Forest Carbon Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation.
-Specifically, we use `r n_records` *ForC* records from `r n_areas_tot` geographic locations representing `r nrow(T1)` C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests.
-For regrowth forests, we quantify age trends for all variables with sufficient data.
+Specifically, we use `r n_records` *ForC* records for `r nrow(T1)` C cycle vaariables from `r n_areas_tot` geographic locations to characterize ensemble C budgets for four broad forest types -- tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga.
+We include estimates for both mature and regrowth (age <100 years) forests, and quantify trends with stand age in regrowth forests for all variables with sufficient data.
 
 *Review Results/ Synthesis.*   The rate of C cycling generally increased from boreal to tropical regions in both mature and regrowth forests, whereas C stocks showed less directional variation.
-The majority of flux variables, together with most live biomass pools, increased significantly with stand age.
-There was generally good closure of C budgets for mature forests, whereas age trends and C budget closure in young forests remain less clearly resolved.
+Net ecosystem production of mature forests was indistinguishable across biomes.
+The majority of flux variables, together with most live biomass pools, increased significantly with stand age when fit with logarithmic functions.
 
@@
 This requires accessible, comprehensive, and large-scale databases with global coverage, which runs contrary to the traditional way forest C stocks and fluxes have been measured and published.
-Large-scale synthesis is critical to benchmarking model performance with global data (Clark et al 2017, Luo et al 2012), quantifying the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], and using book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
+Large-scale synthesis is critical to benchmarking model performance with global data (Clark et al 2017, Luo et al 2012), quantifying the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], and using book-keeping methods to quantify actual or potential exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
 
@@
 A robust understanding of forest impacts on global C cycling is essential. 
-Annual gross sequestration in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
-Most of this enormous C sequestration is counterbalanced by releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] and fire emissions [@van_der_werf_global_2017].
-In recent years, the remaining C sink averaged 3.2 $\pm$ 0.6 Gt C yr^-1^ from 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
-However, deforestation, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reduces the net forest sink to ~1.1-2.2 Gt C yr^-1^ across Earth's forests [@friedlingstein_global_2019].
+Total annual photosynthesis in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], more than seven times the average annual fossil fuel emissions during 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
+Most of this enormous C uptake is counterbalanced by releases to the atmosphere through ecosystem respiration ($R_{eco}$) and fire, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] and fire [@van_der_werf_global_2017].
+In recent years, total forest C uptake has exceeded releases, such that globally forests have been a C sink.
+Considering only areas remaining in forest, this C sink has averaged 3.2 $\pm$ 0.6 Gt C yr^-1^ for 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
+However, deforestation, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reduces the net forest sink to ~1.1-2.2 Gt C yr^-1^ [@friedlingstein_global_2019].
 
@@
 Alongside these conceptual and methodological developments, there has been a proliferation of measurements across the world's forests. 
-The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles, along with variation in data formats, units, measurement methods, *etc.*
-To use these data to address global-scale questions, researchers began synthesizing data into increasingly large databases [e.g., @lieth_primary_1973; @luyssaert_co2_2007;  @bond-lamberty_global_2010; @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018; @cook-patton_mapping_2020].
-The current largest, most comprehensive database on forest C cycling is *ForC* [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], which contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables), with the different variables capturing the unique ecosystem pools (e.g., woody, foliage, and root biomass; dead wood) and flux types (e.g, gross and net primary productivity; soil, root, and ecosystem respiration).
+The result of decades of research on forest C cycling is tens of thousands of records distributed across thousands of scientific articles, varying in data formats, units, measurement methods, *etc.*
+To address global-scale questions, researchers began synthesizing data into increasingly large databases [e.g., @lieth_primary_1973; @luyssaert_co2_2007;  @bond-lamberty_global_2010; @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018; @cook-patton_mapping_2020].
+The current largest, most comprehensive database on forest C cycling is *ForC* [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], which contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables), with different variables capturing distinct ecosystem pools (e.g., woody, foliage, and root biomass; dead wood) and flux types (e.g, gross and net primary productivity; soil, root, and ecosystem respiration).
 These data are ground-based measurements, and *ForC* contains associated data required for interpretation (*e.g.*, stand history, measurement methods).
@@
 
-Forest C cycling varies enormously across biomes, which are artificial categorical distinctions convenient for categorizing the world's forests according to major differences in climate, vegetation, *etc.*
+Forest C cycling varies enormously across biomes, which cateogrize the world's forests according to major differences in climate, vegetation, *etc.* 
 Since the early 19th century, it has been recognized that climate plays a dominant role in shaping differences among forests on a global scale [@humboldt_essay_1807; @holdridge_determination_1947].
@@
 C stocks of mature forests show less directional variation (Fig. `r schematic_fluxes_biome`).
-On average, $B_{ag}$ tends to decrease with latitude, but not as dramatically as fluxes, and with the highest $B_{ag}$ forests in relatively cool, moist temperate regions [@keith_re-evaluation_2009; @smithwick_potential_2002; @hu_mapping_2016].
+On average, aboveground biomass ($B_{ag}$) tends to decrease with latitude, but not as dramatically as fluxes, and with the highest $B_{ag}$ forests in relatively cool, moist temperate regions [@keith_re-evaluation_2009; @smithwick_potential_2002; @hu_mapping_2016].
 In contrast, standing and downed dead wood ($DW_{standing}$ and $DW_{down}$, respectively) and the organic layer ($OL$) tend to accumulate more in colder climates where decomposition is slow relative to $NPP$ [@harmon_ecology_1986; @allen_global_2002].
@@
 Hember et al., 2012]*.
-While these trends have been subject of fairly recent review [@andersonteixeira_altered_2013], *there is need for a synthetic, quantitative review.*
+While these trends have been subject of fairly recent qualitative review [@andersonteixeira_altered_2013], there is need for a synthetic, quantitative review taking advantage of the greatly expanded data now available.
 
@@
 In the past several decades, researchers have started asking how age trends--mostly in $B_{ag}$ or total biomass ($B_{tot}$) accumulation-- vary across biomes.
-Early research on this theme showed that biomass accumulation rates during secondary succession increase with temperature on a global scale [**REFS**; @anderson_temperature-dependence_2006] and *with precipitation in the neotropics* [**REFS**; @chazdon_carbon_2016].
+Early research on this theme showed that biomass accumulation rates during secondary succession increase with temperature on a global scale [**REFS**; @anderson_temperature-dependence_2006] and *with precipitation in the neotropics* [**REFS**; @poorter_biomass_2016; @chazdon_carbon_2016].
 Most recently, @cook-patton_mapping_2020 reinforced these earlier findings with a much larger dataset and crated a high-resolution global map of estimated potential C accumulation rates.
@@
 `r figure_map`).
-Our primary goal is to provide a comprehensive synthesis on broad trends in forest C cycling that can serve as a foundation for improved understanding of global forest C cycling and highlight where key sources of uncertainty still reside.
+Our goal is to provide a comprehensive synthesis on broad trends in forest C cycling that can serve as a foundation for improved understanding of global forest C cycling and highlight where key sources of uncertainty still reside.
 
@@
 We have also added data from individual publications, with a particular focus on productivity [e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @lutz_global_2018; @johnson_climate_2018].
-A record of data sets added to *ForC* is available at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv.
+A record of data sets added to *ForC* over the course of its development is available at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv.
 The database version used for this analysis has been tagged as a new release on Github (v3.0) and assigned a DOI through Zenodo (DOI: TBD).
@@
 These different flux and stock variables represent different pools (e.g., aboveground biomass, root biomass, dead wood) and levels of combination (e.g., total net primary productivity, $NPP$, versus the individual elements of $NPP$ such as foliage, roots, and branches).
-Note that two flux variables, aboveground heterotropic ($R_{het-ag}$) and total ($R_{het}$) respiration, were included for conceptual completeness but had no records in *ForC* (Table 1).
+Note that two flux variables, aboveground heterotropic respiration ($R_{het-ag}$) and total respiration ($R_{het}$), were included for conceptual completeness but had no records in *ForC* (Table 1).
 Records for our focal variables represented `r round(n_focal_variable_records_eligible_for_inclusion/records_eligible_for_inclusion*100,digits=1)`% of the total records eligible for inclusion.
@@
 Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), trees were censused down to a minimum diameter breast height (DBH) threshold of 10 cm or less. 
-All records were measured directly or derived from field measurements (as opposed to modeled).
+All records were measured directly or derived from field measurements.
 
@@
 Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis.
-We defined leaf type (broadleaf / needleleaf) was based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
+We defined leaf type (broadleaf / needleleaf) based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
 For young tropical forests imported from *GROA* but not yet classified by leaf type, we assumed that all were broadleaf, consistent with the rarity of naturally regenerating needleleaf forests in the tropics.

### 2021-01-22 manuscript/tables_figures/stacked_age_trends/schematic.m

@@ -107,8 +107,9 @@ subplot (2,3,1)
 plotBarStackGroups(mature_fluxes, biome_names, facecolor_in_fluxes); 
 t = title('biome differences');
 ylabel ('C fluxes (Mg C ha^{-1} yr^{-1})');
-set(gca, 'XTick', []); %ticks off
 set(gca, 'YTick', []); %ticks off
+set(gca, 'XTickLabel', {'Tropical' 'Temperate' 'Boreal'})
+xtickangle(30)
 
 subplot (2,3,2:3)
 %flux plot:
@@ -129,6 +130,7 @@ plot(age, in_sum, '-b', 'LineWidth', 4); hold on;
 plot(age, out_sum, '--r', 'LineWidth', 4);
 
 t = title('age trends'); 
+xlabel ('stand age');
 set(gca, 'XTick', []); %ticks off
 set(gca, 'YTick', []); %ticks off
 ylim([0, max(in_sum)+1]);
@@ -143,9 +145,9 @@ er=errorbar([1:3],sum(mature_stocks, 2), 0.5*stocks_std, stocks_std);
 er.Color = [0 0 0];                            
 er.LineStyle = 'none'; 
 ylabel ('C  stocks (Mg C ha^{-1})')
+set(gca, 'YTick', []); %ticks off
 set(gca, 'XTickLabel', {'Tropical' 'Temperate' 'Boreal'})
 xtickangle(30)
-set(gca, 'YTick', []); %ticks off
 
 subplot (2,3,5:6) %stocks by age:
 

### 2021-01-23 manuscript/ERL_global_C_review.Rmd
combined stacked area results plots into one, per suggestion of @hmullerlandau
#64

@@
 stock_age_trends=8
-TrB_age_trends=9
-TeB_age_trends=10
-TeN_age_trends=11
-BoN_age_trends=12
+age_trends_all=9
 
@@
 
-TrB_age_trends_caption = "Figure 9 | Age trends in C cycling in tropical broadleaf forests.
+age_trends_all_caption= "Figure 9 | Age trends in C cycling.
 Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
-Asterisks indicate variables whose age trends were calculated based on other variables (* young and mature forests; ** young forests only; *** mature forests only), as follows: $R_{auto-ag}=R_{auto}-R{root}$, where $R_{auto}=NPP(1/CUE-1)$, where $CUE$=0.46 (Collati et al.
-2020); $ANPP_{woody}=max(0,ANPP-ANPP_{foliage})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$.
-Note that there remain substantial uncertainties as to the functional form of age trends and discrepancies in closure among related variables."
-age_trends_caption_generic= "Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
-Asterisks indicate variables whose age trends were calculated based on other variables (* young and mature forests; ** young forests only; *** mature forests only), as follows: $R_{auto-ag}=R_{eco}-R{soil}$; $ANPP_{woody}=min(ANPP_{stem},ANPP_{woody})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$.
+Asterisks indicate variables whose age trends were calculated based on other variables (* young and mature forests; ** young forests only; *** mature forests only), as follows.
+For all forests: $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$, $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$, $DW_{standing}=max(0, DW_{tot}-DW_{down})$.
+For tropical forests: $ANPP_{woody}=max(0,ANPP-ANPP_{foliage})$, $R_{auto-ag}=R_{auto}-R{root}$, where $R_{auto}=NPP(1/CUE-1)$ and $CUE$=0.46 (Collati et al.
+2020).
+for non-tropical forests: $ANPP_{woody}=min(ANPP_{stem},ANPP_{woody})$, $R_{auto-ag}=R_{eco}-R{soil}$, .
 Note that there remain substantial uncertainties as to the functional form of age trends and discrepencies in closure among related variables."
-TeB_age_trends_caption = paste("Figure 10 | Age trends in C cycling in temperate broadleaf forests.", age_trends_caption_generic) 
-TeN_age_trends_caption = paste("Figure 11 | Age trends in C cycling in temperate conifer forests.", age_trends_caption_generic) 
-BoN_age_trends_caption = paste("Figure 12 | Age trends in C cycling in boreal conifer forests.", age_trends_caption_generic)
 
@@
 
-C fluxes commonly increased significantly with stand age (Tables 1, S2, Figs. `r flux_age_trends`, `r TrB_age_trends`-`r BoN_age_trends`, `r all_age_trends`).
+C fluxes commonly increased significantly with stand age (Tables 1, S2, Figs. `r flux_age_trends`, `r age_trends_all`, `r all_age_trends`).
 *ForC* contained 16 C flux variables with sufficient data for analyses of age trends in young forests (see Methods).
@@
 
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap =TrB_age_trends_caption, out.width='100%', fig.pos='H'}
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap =age_trends_all_caption, out.width='100%', fig.pos='H'}
 
-  knitr::include_graphics("tables_figures/stacked_age_trends/TrB_age_trends.png")
-```
-
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap =TeB_age_trends_caption, out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics("tables_figures/stacked_age_trends/TeB_age_trends.png")
-```
-
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap =TeN_age_trends_caption, out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics("tables_figures/stacked_age_trends/TeN_age_trends.png")
-```
-
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap =BoN_age_trends_caption, out.width='100%', fig.pos='H'}
-
-  knitr::include_graphics("tables_figures/stacked_age_trends/BoN_age_trends.png")
+  knitr::include_graphics("tables_figures/stacked_age_trends/age_trends_all.png")
 ```
@@
 
-"Closure" and internal consistency of the C flux budget was less successful for young than mature forests (Figs. `r TrB_age_trends`-`r BoN_age_trends`).
+"Closure" and internal consistency of the C flux budget was less successful for young than mature forests (Figs. `r age_trends_all`).
 Summed regression equations for $R_{soil-het}$ and $R_{root}$ were generally very close to $R_{soil}$.
 We calculated $R_{auto-ag}$ as the difference between $R_{eco}$ and $R_{soil}$ (except for tropical forests, which had insufficient $R_{eco}$ data), effectively guaranteeing near-closure of the CO~2~ efflux (respiration) portion of the budget (negative values in Figs.
-`r TrB_age_trends`-`r BoN_age_trends`).
+`r age_trends_all`).
 In contrast, the  CO~2~ influx portion of the budget generally did not "close": the sum of $R_{auto}$ ($R_{root}+R_{auto-ag}$, as described above) and components of $NPP$ consistently fell short of $GPP$, particularly in in young stands (range across forest types and ages: 0.9-7.6 Mg C ha^-1^ yr^-1^).
 Moreover, there was not consistent budget closure among the components of $NPP$, and substantially different age trends resulting from the sum of components versus total $NPP$ (Figs.
-`r TrB_age_trends`-`r BoN_age_trends`).
+`r age_trends_all`).
 ALthough age trends of young forests often converged towards mature forest averages, there were also some  discrepancies between young forest trends and mature forest averages (Figs.
-`r flux_age_trends`, `r TrB_age_trends`-`r BoN_age_trends`, `r all_age_trends`), most notably including a tendency for higher fluxes in regrowth boreal forests than in their mature counterparts (Fig.
-`r BoN_age_trends`).
+`r flux_age_trends`, `r age_trends_all`, `r all_age_trends`), most notably including a tendency for higher fluxes in regrowth boreal forests than in their mature counterparts (Fig.
+`r age_trends_all`).
 
@@
 In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs. 2-5, `r stock_age_trends`).
-The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Table 1; Figs. `r flux_age_trends`- `r BoN_age_trends`, `r all_age_trends`).
+The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Table 1; Figs. `r flux_age_trends`- `r age_trends_all`, `r all_age_trends`).
 Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks and $NEP$ of mature forests are correlated with a different set of factors.
 Together, these results refine and expand out understanding of C cycling in mature forests, while providing the first global-scale analysis of age trends in multiple forest C cycling stocks and fluxes (Figs.
-`r TrB_age_trends`-`r BoN_age_trends`).
+`r age_trends_all`).
 
@@
 Our study reveals that most C fluxes quickly increase to a plateau as stands age (Figs.
-`r flux_age_trends`, `r TrB_age_trends`-`r BoN_age_trends`), consistent with current understanding of age trends in forest C cycling [Fig.
+`r flux_age_trends`, `r age_trends_all`), consistent with current understanding of age trends in forest C cycling [Fig.
 `r schematic`; e.g., @andersonteixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].

### 2021-01-23 manuscript/tables_figures/stacked_age_trends/schematic.m

@@ -6,21 +6,6 @@ clf; clear; close all;
 set(0,'DefaultLegendAutoUpdate','off');
 
 
-%plot settings
-
-    %subplot positions
-    plot_width=0.32;
-    plot_height=0.15;
-    plot_space_vertical=(1-plot_height*5)/5-.01;
-    plot_space_horizontal=(1-plot_width*2.5)/2-.05;
-
-    %color
-    facecolor_cat=[.1 .4 .9; 0.8 0 0;  .9 .527 .527 ; 1 .9 .9 ]; %blue/ red
-
-    
-    %font
-    fs= 10; %font size
-
 
 
 %% ~~~~~~~CREATE DATA FOR PLOTTING~~~~~~~~~~

### 2021-01-23 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m

@@ -10,6 +10,45 @@ figures_dir='/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/ma
 working_dir='/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/manuscript/tables_figures/stacked_age_trends';
 ForC_dir='/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC';
 
+%plot settings
+    fig_dims=[20 20 700 800];
+
+    %subplot positions
+    plot_width=0.33;  %plot_width+legend_width+left_marg=.5
+    legend_width=0.1;
+    plot_height=0.19;
+    left_marg=.07;
+    lower_marg=.05;
+    plot_space_vertical=.05;
+    plot_space_horizontal=.025;
+    pos_TrB_f=[left_marg lower_marg+3*(plot_height+plot_space_vertical) plot_width+legend_width plot_height];
+    pos_TeB_f=[left_marg lower_marg+2*(plot_height+plot_space_vertical) plot_width plot_height];
+    pos_TeN_f=[left_marg lower_marg+1*(plot_height+plot_space_vertical) plot_width plot_height];
+    pos_BoN_f=[left_marg lower_marg+0*(plot_height+plot_space_vertical) plot_width plot_height];
+    pos_TrB_s=[left_marg+plot_width+legend_width+plot_space_horizontal lower_marg+3*(plot_height+plot_space_vertical) plot_width+legend_width plot_height];
+    pos_TeB_s=[left_marg+plot_width+legend_width+plot_space_horizontal lower_marg+2*(plot_height+plot_space_vertical) plot_width plot_height];
+    pos_TeN_s=[left_marg+plot_width+legend_width+plot_space_horizontal lower_marg+1*(plot_height+plot_space_vertical) plot_width plot_height];
+    pos_BoN_s=[left_marg+plot_width+legend_width+plot_space_horizontal lower_marg+0*(plot_height+plot_space_vertical) plot_width plot_height];
+
+    %colors
+    facecolor_in_fluxes= [0.3 0 1;...  %'R_{auto-ag}*'
+                        0.1 0 .4;...  % 'R_{root}', 
[…]

### 2021-01-23 manuscript/response_to_reviews.Rmd
work on HM comments

@@
 (1) added a schematic figure, somewhat in the style of @odum_strategy_1969, to help with conceptual framing of the review
-(2) added figures synthesizing changes in fluxes through time for each biome;
+(2) added a figure synthesizing changes in fluxes through time for each biome;
 (3) removed maps from age trend diagrams (available in SI) from age trend diagrams, which makes the figures less busy/ easier to read; and
@@
 3. There is a general sloppiness to the manuscript that makes me think the authors did not proofread the manuscript carefully.
+
+**We have proofread more carefully this time.**
 

### 2021-01-23 manuscript/ERL_global_C_review.Rmd
work on HM comments

@@
 Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
-Note that variables differ in geographical repressentation, resulting in potential imbalances (Figs.
+Note that variables differ in geographical representation, resulting in potential imbalances (Figs.
 S5-S30).
@@
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.
-Asterisk after variable name indicates lack of C cycle closure."
+To illustrate the magnitude of different fluxes, arrow width is proportional to the square root of the corresponding flux.
+An asterisk after a variable name indicates lack of C cycle closure."
 TrB_mature_caption = paste( "Figure 3 | C cycle diagram for mature tropical broadleaf forests. ", schematics_caption_generic)  
@@
 
-There were sufficient data to assess mature forest biome differences for 15 flux variables, and significant differences among biomes were detected for 12 variables (Table 1).
+There were sufficient data to assess differences among biomes in mature forest values for 15 flux variables, and significant differences among biomes were detected for 12 variables (Table 1).
 In all of these cases--including C fluxes into, within, and out of the ecosystem--C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs.
@@
 Differences between tropical and boreal forests were always significant, with temperate forests intermediate and significantly different from one or both.
-Fluxes tended to be numerically greater in temperate broadleaf than conifer forests, but the difference was never statistically significant.
+Fluxes tended to be numerically greater in temperate broadleaf than temperate conifer forests, but the difference was never statistically significant.
 This pattern held for the following variables: $GPP$, $NPP$, $ANPP$, $ANPP_{stem}$,$ANPP_{branch}$, $ANPP_{foliage}$, $BNPP$, $R_{eco}$, $R_{root}$, $R_{soil}$, and $R_{het-soil}$. 
@@
 ```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEP$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
@@
 ```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
 In each panel, the left scatterplot  shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
@@
 Differences in C fluxes across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests.
-The single exception was $ANPP_{stem}$, for which temperate broadleaf and conifer forests had similar flux rates than tropical forests.
+The single exception was $ANPP_{stem}$, for which temperate broadleaf and conifer forests had flux rates similar to tropical forests.
 Notably, and in contrast to the lack of biome differences in $NEP$ for mature forests (Fig.
@@
 
+*work on this par:*
+
 *ForC* v3.0 provided unprecedented coverage of most major variables, yielding a broad picture of C cycling in the world's major forest biomes.
 Carbon cycling rates generally increased from boreal to tropical regions and with stand age.
-Specifically, the major C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for regrowth as well as mature forests (Figs.
+Specifically, most C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for regrowth as well as mature forests (Figs.
 `r flux_age_trends`- `r stock_age_trends`).
-In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs. 2-5, `r stock_age_trends`).
-The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Table 1; Figs. `r flux_age_trends`- `r age_trends_all`, `r all_age_trends`).
-Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks and $NEP$ of mature forests are correlated with a different set of factors.
-Together, these results refine and expand out understanding of C cycling in mature forests, while providing the first global-scale analysis of age trends in multiple forest C cycling stocks and fluxes (Figs.
+The notable exception was mature forest $NEP$, which, as the difference between $GPP$ and $R_{eco}$, was indistinguishable across biomes.
+There was also little directional variation in mean mature forest C stocks across biomes, although maximum values for stocks including live or standing woody biomass ($B_{tot}$, $B_{ag]$, $B_{ag-wood}$, $B_{root}$, $B_{root_coarse}$, $DW_{tot}$, $DW_{standing}$) consistently occurred in temperate biomes (Figs.
+`r schematic`, `r `all_diagrams_mature`, `r stock_age_trends`).
+Consistent with theory and previous studies (Fig.
+`r schematic`), the majority of flux variables, together with most live biomass pools, increased significantly with stand age (Table 1; Figs.
+`r flux_age_trends`- `r age_trends_all`, `r all_age_trends`).
+Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks and $NEP$ of mature forests, which are defined by the differences between in- and out- fluxes, do not vary systematically across biomes.
+Together, these results refine and expand out understanding of C cycling in mature forests, while providing the first global-scale analysis of age trends in multiple forest C stocks and fluxes (Figs.
 `r age_trends_all`).

### 2021-01-23 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m

@@ -14,13 +14,13 @@ ForC_dir='/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC';
     fig_dims=[20 20 700 800];
 
     %subplot positions
-    plot_width=0.33;  %plot_width+legend_width+left_marg=.5
-    legend_width=0.1;
+    plot_width=0.295;  %plot_width+legend_width+left_marg=.5
+    legend_width=0.145;
     plot_height=0.19;
-    left_marg=.07;
+    left_marg=.057;
     lower_marg=.05;
     plot_space_vertical=.05;
-    plot_space_horizontal=.025;
+    plot_space_horizontal=left_marg;
     pos_TrB_f=[left_marg lower_marg+3*(plot_height+plot_space_vertical) plot_width+legend_width plot_height];
     pos_TeB_f=[left_marg lower_marg+2*(plot_height+plot_space_vertical) plot_width plot_height];
     pos_TeN_f=[left_marg lower_marg+1*(plot_height+plot_space_vertical) plot_width plot_height];

### 2021-01-23 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m
move legends

@@ -14,21 +14,22 @@ ForC_dir='/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC';
     fig_dims=[20 20 700 800];
 
     %subplot positions
-    plot_width=0.295;  %plot_width+legend_width+left_marg=.5
+    plot_width=0.4;  %plot_width+legend_width+left_marg=.5
     legend_width=0.145;
-    plot_height=0.19;
+    plot_height=0.17;
+    legend_height=.1
     left_marg=.057;
     lower_marg=.05;
     plot_space_vertical=.05;
     plot_space_horizontal=left_marg;
-    pos_TrB_f=[left_marg lower_marg+3*(plot_height+plot_space_vertical) plot_width+legend_width plot_height];
+    pos_TrB_f=[left_marg lower_marg+3*(plot_height+plot_space_vertical) plot_width plot_height+legend_height];
     pos_TeB_f=[left_marg lower_marg+2*(plot_height+plot_space_vertical) plot_width plot_height];
     pos_TeN_f=[left_marg lower_marg+1*(plot_height+plot_space_vertical) plot_width plot_height];
     pos_BoN_f=[left_marg lower_marg+0*(plot_height+plot_space_vertical) plot_width plot_height];
-    pos_TrB_s=[left_marg+plot_width+legend_width+plot_space_horizontal lower_marg+3*(plot_height+plot_space_vertical) plot_width+legend_width plot_height];
-    pos_TeB_s=[left_marg+plot_width+legend_width+plot_space_horizontal lower_marg+2*(plot_height+plot_space_vertical) plot_width plot_height];
-    pos_TeN_s=[left_marg+plot_width+legend_width+plot_space_horizontal lower_marg+1*(plot_height+plot_space_vertical) plot_width plot_height];
-    pos_BoN_s=[left_marg+plot_width+legend_width+plot_space_horizontal lower_marg+0*(plot_height+plot_space_vertical) plot_width plot_height];
+    pos_TrB_s=[left_marg+plot_width+plot_space_horizontal lower_marg+3*(plot_height+plot_space_vertical) plot_width plot_height+legend_height];
+    pos_TeB_s=[left_marg+plot_width+plot_space_horizontal lower_marg+2*(plot_height+plot_space_vertical) plot_width plot_height];
[…]

### 2021-01-23 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m
formatting
#64

@@ -14,14 +14,13 @@ ForC_dir='/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC';
     fig_dims=[20 20 700 800];
 
     %subplot positions
-    plot_width=0.4;  %plot_width+legend_width+left_marg=.5
-    legend_width=0.145;
+    plot_width=0.41;  
     plot_height=0.17;
-    legend_height=.1
-    left_marg=.057;
+    legend_height=.1;
+    left_marg=.09;
     lower_marg=.05;
     plot_space_vertical=.05;
-    plot_space_horizontal=left_marg;
+    plot_space_horizontal=.07;
     pos_TrB_f=[left_marg lower_marg+3*(plot_height+plot_space_vertical) plot_width plot_height+legend_height];
     pos_TeB_f=[left_marg lower_marg+2*(plot_height+plot_space_vertical) plot_width plot_height];
     pos_TeN_f=[left_marg lower_marg+1*(plot_height+plot_space_vertical) plot_width plot_height];
@@ -404,8 +403,7 @@ plot (114, -R_eco_mature, 'ok', 'MarkerFaceColor', 'r'); hold on;
     end
     h2bl.FaceColor= facecolor_out_fluxes(3,:);
 xlim([0 119])
-t = title(biomes(b));
-ylabel ('C  fluxes (Mg C ha^{-1} yr^{-1})')
+ylabel ({'\bf ',char(biomes(b)), '\rm C  fluxes (Mg C ha^{-1} yr^{-1})'})
 
 if b==4
 xlabel ('stand age (years) ......... MATURE STANDS');
@@ -440,7 +438,7 @@ h2=area(112:116, mature_stocks.*ones(5,size(mature_stocks,1))); hold on;
 %plot(age, B_tot, '-k', 'LineWidth', 3);
 %plot(age, B_tot_calc, '--k', 'LineWidth', 3);
 
-ylabel ('C  stocks (Mg C ha^{-1})')
+ylabel ({'C  stocks (Mg C ha^{-1})'})
 
 xlim([0 119])
 if b==4

### 2021-01-23 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m
formatting
#64

@@ -375,22 +375,11 @@ h2b=area(112:116, mature_out_fluxes.*ones(5,size(mature_out_fluxes,1)),'HandleVi
 if b~= 1 
     plot(age, GPP, '-b', 'LineWidth', 2,'HandleVisibility','off'); hold on; % eddy flux: insufficient data for tropics
     plot(age, -R_eco, '-r', 'LineWidth', 2,'HandleVisibility','off'); % eddy flux: insufficient data for tropics
-    %plot(age, -R_root-R_het_soil-R_auto_ag_calc, '--r', 'LineWidth', 3); % eddy flux: insufficient data for tropics
     plot(age, NEP, '-y', 'LineWidth', 2,'HandleVisibility','off');hold on; % eddy flux: insufficient data for tropics
-    %plot(age, NEP_calc, '--y', 'LineWidth', 3);hold on;
 end
 plot (114, GPP_mature, 'ok', 'MarkerFaceColor', 'b'); hold on;
 plot (114, NEP_mature, 'ok', 'MarkerFaceColor', 'y'); hold on;
 plot (114, -R_eco_mature, 'ok', 'MarkerFaceColor', 'r'); hold on;
-
-
-%plot(age, in_sum, '--k', 'LineWidth', 3); hold on;
-%plot(age, R_auto_calc+NPP, '-k', 'LineWidth', 3);hold on;
-%plot(age, -R_auto_ag_calc-R_soil, '-y', 'LineWidth', 3);hold on;
-%plot(age, -R_auto_ag_calc-R_root-R_het_soil, '--y', 'LineWidth', 3);hold on;
-
-
-
 
 %set facecolor:
     for n=1:length(mature_in_fluxes)
@@ -428,7 +417,6 @@ end
 h1=area (age, stocks'); hold on;
 h2=area(112:116, mature_stocks.*ones(5,size(mature_stocks,1))); hold on;
 
-
 %set facecolor:
     for n=1:length(mature_stocks)
         h1(n).FaceColor= facecolor_stocks(n,:);
@@ -438,9 +426,10 @@ h2=area(112:116, mature_stocks.*ones(5,size(mature_stocks,1))); hold on;
 %plot(age, B_tot, '-k', 'LineWidth', 3);
 %plot(age, B_tot_calc, '--k', 'LineWidth', 3);
 
-ylabel ({'C  stocks (Mg C ha^{-1})'})
-
[…]

### 2021-01-23 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m
formatting
#64

@@ -15,11 +15,11 @@ ForC_dir='/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC';
 
     %subplot positions
     plot_width=0.41;  
-    plot_height=0.17;
+    plot_height=0.19;
     legend_height=.1;
     left_marg=.09;
     lower_marg=.05;
-    plot_space_vertical=.05;
+    plot_space_vertical=.027;
     plot_space_horizontal=.07;
     pos_TrB_f=[left_marg lower_marg+3*(plot_height+plot_space_vertical) plot_width plot_height+legend_height];
     pos_TeB_f=[left_marg lower_marg+2*(plot_height+plot_space_vertical) plot_width plot_height];
@@ -395,7 +395,7 @@ xlim([0 119])
 ylabel ({'\bf ',char(biomes(b)), '\rm C  fluxes (Mg C ha^{-1} yr^{-1})'})
 
 if b==4
-xlabel ('stand age (years) ......... MATURE STANDS');
+xlabel ('stand age (years)');
 end
 
 if b==1
@@ -429,9 +429,10 @@ h2=area(112:116, mature_stocks.*ones(5,size(mature_stocks,1))); hold on;
 ylim([0 sum(mature_stocks)+10])
 xlim([0 119])
 
+
 ylabel ({'C  stocks (Mg C ha^{-1})'})
 if b==4
-xlabel ('stand age (years) ......... MATURE STANDS');
+xlabel ('stand age (years)');
 end
 
 
@@ -452,6 +453,16 @@ end
 %xlabel ('age')
 %ylabel ('GPP-in.sum')
 %legend ('TeB', 'TeN', 'BoN')
+
+%add "mature" annotation on x-axes:
+annotation('textbox', [.453, .7, 0, 0], 'string', 'mature')
+annotation('textbox', [.93, .7, 0, 0], 'string', 'mature')
+annotation('textbox', [.453, .485, 0, 0], 'string', 'mature')
+annotation('textbox', [.93, .485, 0, 0], 'string', 'mature')
+annotation('textbox', [.453, .265, 0, 0], 'string', 'mature')
+annotation('textbox', [.93, .265, 0, 0], 'string', 'mature')
+annotation('textbox', [.453, .049, 0, 0], 'string', 'mature')
[…]

### 2021-01-23 manuscript/ERL_global_C_review.Rmd
finish (mostly) HM comments
all comments are either taken care of or transferred to issues/ notes in manuscript

@@
 All of these displayed a significant overall increase with with $log10[stand.age]$.
-There were sufficient data to model age $\times$ biome interactions were also significant for all ten of these C stock variables (Table S2), with living C stocks tending to accumulate more rapidly during the early stages of forest regrowth in tropical forests (Figs.
+Age $\times$ biome interactions were also significant for all ten of these C stock variables (Table S2), with living C stocks tending to accumulate more rapidly during the early stages of forest regrowth in tropical forests (Figs.
 `r stock_age_trends`, `r all_stock_age_trends`).
-In the case of two non-living C stocks ($DW_{down}$ and $OL$), age $\times$ biome interactions were such that 
+In the case of two non-living C stocks ($DW_{down}$ and $OL$), age $\times$ biome interactions were such that age trends were positive in some biomes and negative in others.
 Specifically, $DW_{down}$ declined with age in temperate and boreal forests, compared to an increase with age in tropical forests (Figs. `r stock_age_trends`, `r DW_down_fig`).
@@
 The notable exception was mature forest $NEP$, which, as the difference between $GPP$ and $R_{eco}$, was indistinguishable across biomes.
-There was also little directional variation in mean mature forest C stocks across biomes, although maximum values for stocks including live or standing woody biomass ($B_{tot}$, $B_{ag]$, $B_{ag-wood}$, $B_{root}$, $B_{root_coarse}$, $DW_{tot}$, $DW_{standing}$) consistently occurred in temperate biomes (Figs.
+There was also little directional variation in mean mature forest C stocks across biomes, although maximum values for stocks including live or standing woody biomass ($B_{tot}$, $B_{ag}$, $B_{ag-wood}$, $B_{root}$, $B_{root_coarse}$, $DW_{tot}$, $DW_{standing}$) consistently occurred in temperate biomes (Figs.
 `r schematic`, `r `all_diagrams_mature`, `r stock_age_trends`).
@@
 
+*HM: I recommend putting this later in the discussion.  *
+
 The large number of C cycle variables covered by ForC, and the general consistency among them, provide confidence that our overall reported mature forest means provide useful baselines for analysis -- with the caveats that they are unlikely to be accurate representations of C cycling for any particular forest, and that these sample means almost certainly do not represent true biome means (particularly for temperate conifer forests where high-biomass stands are over-represented in *ForC*).
+
+In this analysis, the C cycle budgets for mature forests (Figs.
+`r all_diagrams_mature`) generally “close”–-that is, the sums of component variables do not differ from the larger fluxes by more than one standard deviation.
+On the one hand, this reflects the general fact that ecosystem-scale measurements tend to close the C budget more easily and consistently than, for example, for energy balance [@stoy_data-driven_2013].
+On the other, however, *ForC* derives data from multiple heterogeneous sources, and standard deviations within each biome are high; as a result, the standard for C closure is relatively loose [*c.f.* @houghton_terrestrial_2020].
+The one instance where the C budgets doesn’t close is likely due to differences in the representation of forest types (*i.e.*, disproportionate representation of US Pacific NW for  $B_{root-coarse}$ relative to  $B_{root}$; Fig.
+`r diagram_TeN_mature`) rather than issues of methodological accuracy.
+The overall high degree of closure implies that *ForC* gives a consistent picture of C cycling within biomes for mature forests. 
+This is an important and useful test, because it allows for consistency checks within the C cycle, for example leveraging separate and independently-measured fluxes to constrain errors in another [@phillips_value_2017; @williams_impacts_2014; @harmon_heterotrophic_2011], or producing internally consistent global data products [@wang_golum-cnp_2018].
+
+In contrast, age trends for young forests generally remain less clearly defined, in large part because their data records remain somewhat sparse for most variables (*i.e.*, have low representation of different geographical regions for any given age).
+While this analysis provides a first analysis of age trends in forest C cycling for multiple variables at a global scale, improved resolution of these trends will require additional data.
 
@@
 *ForC* is not intended to replace databases that are specialized for particular parts of the C cycle analyses, e.g., aboveground biomass [@spawn_harmonized_2020], land-atmosphere fluxes [@baldocchi_fluxnet_2001], soil respiration [@jian_restructured_2020], or the human footprint in global forests [@magnani_human_2007].
-
-In this analysis, the C cycle budgets for mature forests (Figs.
-`r all_diagrams_mature`) generally “close”–-that is, the sums of component variables do not differ from the larger fluxes by more than one standard deviation.
-On the one hand, this reflects the general fact that ecosystem-scale measurements tend to close the C budget more easily and consistently than, for example, for energy balance [@stoy_data-driven_2013].
-On the other, however, *ForC* derives data from multiple heterogeneous sources, and standard deviations within each biome are high; as a result, the standard for C closure is relatively loose [*c.f.* @houghton_terrestrial_2020].
-Nonetheless, the lack of closure, in the one instance where it occurs, is probably more reflective of differences in the representation of forest types (*i.e.*, disproportionate representation of US Pacific NW for  $B_{root-coarse}$ relative to  $B_{root}$; Fig.
-`r diagram_TeN_mature`) than of methodological accuracy.
-The overall high degree of closure implies that *ForC* gives a consistent picture of C cycling within biomes for mature forests. 
-This is an important and useful test, because it allows for consistency checks within the C cycle, for example leveraging separate and independently-measured fluxes to constrain errors in another [@phillips_value_2017; @williams_impacts_2014; @harmon_heterotrophic_2011], or producing internally consistent global data products [@wang_golum-cnp_2018].
-
-In contrast, age trends for young forests generally remain less clearly defined, in large part because their data records remain somewhat sparse for most variables (*i.e.*, have low representation of different geographical regions for any given age).
-While this analysis provides a first analysis of age trends in forest C cycling for multiple variables at a global scale, improved resolution of these trends will require additional data.
 
@@
 
-Whereas biomass can be remotely sensed and receives significant research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
+Whereas aboveground biomass can be remotely sensed [albeit with significant uncertainties; @ploton_spatial_2020] and receives significant research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
 Although these stocks can be important--exceeding 100 Mg C ha^-1^ in some stands (Figs.
@@
 Biomass is the largest C stock in most forests, and most of the emphasis has traditionally been on this variable. 
-Remote-sensing driven biomass estimates [e.g., @saatchi_benchmark_2011], calibrated based on high-quality ground-based data [@schepaschenko_forest_2019; @chave_ground_2019], are well suited for this task.
-Note, however, that factors such as stand age and disturbance history are difficult, if possible, to detect remotely, and can only be characterized for very recent decades [@hansen_high-resolution_2013; @song_global_2018; @curtis_classifying_2018].
+Remote-sensing driven aboveground biomass estimates [e.g., @saatchi_benchmark_2011], calibrated based on high-quality ground-based data [@schepaschenko_forest_2019; @chave_ground_2019], provide the most promising approach, but significant uncertainties remain [@ploton_spatial_2020].
+Note, however, that factors such as stand age and disturbance history are difficult, if not impossible, to detect remotely, and can only be characterized for very recent decades [@hansen_high-resolution_2013; @song_global_2018; @curtis_classifying_2018].
 Ground-based data such as *ForC* are therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020, and thus constraining variables such as carbon sink potential [@luyssaert_old-growth_2008].

### 2021-01-23 manuscript/tables_figures/stacked_age_trends/schematic.m
formatting
#61

@@ -5,9 +5,35 @@
 clf; clear; close all;
 set(0,'DefaultLegendAutoUpdate','off');
 
+%plot settings
+    %subplot positions
+    plot_width_biome=0.27; 
+    plot_width_age_f=0.57; 
+    plot_width_age_s=0.47;
+    plot_height=0.37;
+    left_marg=.09;
+    lower_marg=.1;
+    plot_space_vertical=.1;
+    plot_space_horizontal=.04;
+    pos_biome_f=[left_marg lower_marg+1*(plot_height+plot_space_vertical) plot_width_biome plot_height];
+    pos_biome_s=[left_marg lower_marg+0*(plot_height+plot_space_vertical) plot_width_biome plot_height];
+    pos_age_f=[left_marg+plot_width_biome+plot_space_horizontal lower_marg+1*(plot_height+plot_space_vertical) plot_width_age_f plot_height];
+    pos_age_s=[left_marg+plot_width_biome+plot_space_horizontal lower_marg+0*(plot_height+plot_space_vertical) plot_width_age_s plot_height];
 
-
-
+%set facecolors:
+    facecolor_in_fluxes= [0.2 0 .7;...  %' 'R_{auto}'
+                        0 .7 1;...  %'BNPP'
+                        0 .527 .2;... %'ANPP_{foliage}',
+                        150/255 1 0 ;...%'ANPP_{woody.turnover}*'
+                        255/256 255/256 0]; % NEP
+    facecolor_stocks= [%0.1 0 .4;...  %'B_{root-coarse}*', ' 
+                        %0 .7 1;... % 'B_{root-fine}',
+                        %0 .527 .27 ;... % B_{foliage}',
+                        170/255 1 0;... % 'B_{ag-wood}*', 
+                        %1 .8 0;...  % 'DW_{standing}*', 
+                        1 99/255 99/255 ... % 'DW_{down}',
+                        %; 0.7 0 .7 ;... %'OL'
+                        ];  
 %% ~~~~~~~CREATE DATA FOR PLOTTING~~~~~~~~~~
 age = linspace (1,100,100);
 biome_names={'tropical', 'temperate', 'boreal'};
@@ -70,24 +96,11 @@ stocks_std = sum(mature_stocks, 2)'.*[.4 1 .4];
 
[…]

### 2021-01-23 manuscript/tables_figures/stacked_age_trends/schematic.m
differentiate blue shades
#61

@@ -21,7 +21,7 @@ set(0,'DefaultLegendAutoUpdate','off');
     pos_age_s=[left_marg+plot_width_biome+plot_space_horizontal lower_marg+0*(plot_height+plot_space_vertical) plot_width_age_s plot_height];
 
 %set facecolors:
-    facecolor_in_fluxes= [0.2 0 .7;...  %' 'R_{auto}'
+    facecolor_in_fluxes= [0.2 0 .5;...  %' 'R_{auto}'
                         0 .7 1;...  %'BNPP'
                         0 .527 .2;... %'ANPP_{foliage}',
                         150/255 1 0 ;...%'ANPP_{woody.turnover}*'

### 2021-01-23 manuscript/tables_figures/stacked_age_trends/schematic.m
formatting
#61

@@ -124,8 +124,8 @@ h2=area (95:100, out_fluxes(:, 95:100)' ,'HandleVisibility','off'); hold on;
     h2(1).EdgeColor= facecolor_in_fluxes(1,:);
     h2(2).FaceColor= [.7 0 .7];
     
-plot(age, in_sum, '-b', 'LineWidth', 4); hold on;
-plot(age, out_sum, '--r', 'LineWidth', 4);
+plot(age, in_sum, '-b', 'LineWidth', 3.4); hold on;
+plot(age, out_sum, '--r', 'LineWidth', 3.4);
 
 t = title('age trends'); 
 xlabel ('stand age');
@@ -143,6 +143,7 @@ h=bar(mature_stocks, 'stacked'); hold on;
 er=errorbar([1:3],sum(mature_stocks, 2), 0.5*stocks_std, stocks_std);
 er.Color = [0 0 0];                            
 er.LineStyle = 'none'; 
+er.LineWidth= 1.7;
 ylabel ('C  stocks (Mg C ha^{-1})')
 set(gca, 'YTick', []); %ticks off
 set(gca, 'XTickLabel', {'Tropical' 'Temperate' 'Boreal'})
@@ -152,7 +153,7 @@ subplot ('Position', pos_age_s) %stocks by age:
 
 h=area (age, stocks'); %, 'LineStyle','-'); 
 hold on;
-plot(age, biomass, '-k', 'LineWidth', 4);
+plot(age, biomass, '-k', 'LineWidth', 3.4);
 xlabel ('stand age');
 set(gca, 'XTick', []); %ticks off
 set(gca, 'YTick', []); %ticks off

### 2021-01-24 manuscript/ERL_global_C_review.Rmd
"finish" schematic
#61

@@
 Variables are defined in Table 1.
-Age trends, which represent idealized dynamics following a disturbance that removes all living and non-living vegetation, are an updated version of the classic figure from Odum (1969), with heavy lines corresponding to those in Odum's figure and $NEP$ corresponding to Odum's 'net production'.
+Age trends, which represent idealized dynamics following a disturbance that removes all living vegetation, are an updated version of the classic figure from Odum (1969), with heavy lines corresponding to those in Odum's figure and $NEP$ corresponding to Odum's 'net production'.
 Here, $NEP$ conists primarily of woody aboveground net primary production ($ANPP_{woody}$), while $ANPP_{woody.turnover}$ is the sum of woody mortality and branch turnover.
-", out.width='100%', fig.pos='H'}
+Dotted lines refer to decomposition of potential 'legacy' organic material produced prior to the disturbance and remaining at the site (e.g., standing and fallen dead wood, $DW_{tot}$).", out.width='100%', fig.pos='H'}
 

### 2021-01-24 manuscript/tables_figures/stacked_age_trends/schematic.m

@@ -9,7 +9,7 @@ set(0,'DefaultLegendAutoUpdate','off');
     %subplot positions
     plot_width_biome=0.27; 
     plot_width_age_f=0.57; 
-    plot_width_age_s=0.47;
+    plot_width_age_s=0.515;
     plot_height=0.37;
     left_marg=.09;
     lower_marg=.1;
@@ -27,13 +27,13 @@ set(0,'DefaultLegendAutoUpdate','off');
                         150/255 1 0 ;...%'ANPP_{woody.turnover}*'
                         255/256 255/256 0]; % NEP
     facecolor_stocks= [%0.1 0 .4;...  %'B_{root-coarse}*', ' 
-                        %0 .7 1;... % 'B_{root-fine}',
-                        %0 .527 .27 ;... % B_{foliage}',
+                        0 .7 1;... % 'B_{root-fine}',
+                        0 .527 .27 ;... % B_{foliage}',
                         170/255 1 0;... % 'B_{ag-wood}*', 
                         %1 .8 0;...  % 'DW_{standing}*', 
-                        1 99/255 99/255 ... % 'DW_{down}',
-                        %; 0.7 0 .7 ;... %'OL'
-                        ];  
+                        1 99/255 99/255; ... % 'DW_{down}',
+                        0.7 0 .7 ];... %'OL'
+                     
 %% ~~~~~~~CREATE DATA FOR PLOTTING~~~~~~~~~~
 age = linspace (1,100,100);
 biome_names={'tropical', 'temperate', 'boreal'};
@@ -41,7 +41,7 @@ biome_names={'tropical', 'temperate', 'boreal'};
 %FLUXES
 % 1. by age
 %for stacked plot
-NEP=2.4*log10(age)-.04*(age-1); %placeholder (peaks around 2 at 23 yrs age)
+NEP=2.4*log10(age)-.04*(age-1); 
 GPP= NEP + (1.99+1.44)* log10(age);
 ANPP_foliage = 1.66* log10(age) +.18 - .01*(age-1);
 ANPP_woody = 2.31* log10(age) -.76 - .02*(age-1);
[…]

### 2021-01-24 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m

@@ -467,3 +467,5 @@ annotation('textbox', [.93, .049, 0, 0], 'string', 'mature')
 %% ~~~~~~~ SAVE FIGURE~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 cd(working_dir)
 print('age_trends_all', '-dpng', '-r300')
+
+save params_matrix

### 2021-01-24 manuscript/response_to_reviews.Rmd
finished responses to R3
I think this is good to go.... But see issue #57

@@
 (2) added a figure synthesizing changes in fluxes through time for each biome;
-(3) removed maps from age trend diagrams (available in SI) from age trend diagrams, which makes the figures less busy/ easier to read; and
+(3) removed maps from age trend diagrams (maps still available in SI), which makes the figures less busy/ easier to read; and
 (4) moved C cycle diagrams for young forests to SI;
@@
 
-*[Issue #57](https://github.com/forc-db/ERL-review/issues/57)*
+**An analysis of this type would need to be fairly involved, including a separate analysis for each variable.
+While do-able (we have previously published a plot similar to that described for the entire database), such an analysis would only begin to cover important aspects of representativeness (MAT and MAP, but not seasonality, soils, floristic composition, etc.).
+We feel that, in this case, the insights gained from the analysis would not justify the strain on time and resources that it would require.
+We have, however, given additional attention to the issue of representativeness in the revision.
+Specifically, we have done the following:**
+
+- **Set the stage for discussion of within-biome variability by including error bars on our schematic figure (Fig. 1).**
+
+- **Ensured that the issue is clearly and transparently presented in the figures--for example, by adding the following statement to the captions of Figures 3-6 (formerly 2-5): "Note that variables differ in geographical representation, resulting in potential imbalances (Figs.
+S5-S30).
+Probability that estimates reflect true biome means scales with the number of distinct geographical areas represented." **
+
+- **Revisited our presentation of representativeness issues in the results and discussion, revising text to ensure clear communication of the concept that biomed differences were influenced by geographical representation within the biome.
+For example, our discussion of stock differences across biome differences now reads: "...For $B_{tot}$ and $B_{ag}$, tropical broadleaf forests had the highest mean biomass and boreal forests the lowest, with temperate broadleaf and needleleaf ($B_{ag}$ only) intermediate.
+However, maximum values for stocks including live or standing woody biomass ($B_{tot}$, $B_{ag}$, $B_{ag-wood}$, $B_{root}$, $B_{root_coarse}$, $DW_{tot}$, $DW_{standing}$) consistently occurred in temperate biomes (Figs.
+1, 8, S20-S30).
+For variables that were disproportionately sampled in such high-biomass forests ($B_{ag-wood}$, $B_{foliage}$, and $B_{root-coarse}$; disproportionately sampled in  the US Pacific Northwest), temperate conifer forests had significantly higher stocks than the other biomes."**
 
@@
 
-- *move young forest C budgets to SI?*
+**We have done a major re-arrangment of figures, including the following: (1) added a schematic figure, somewhat in the style of @odum_strategy_1969, to help with conceptual framing of the review; (2) added a figure synthesizing changes in fluxes through time for each biome; (3) removed maps from age trend diagrams (maps still available in SI), which makes the figures less busy/ easier to read; and (4) moved C cycle diagrams for young forests to SI.
+The result is that (1) the main manuscript contains only the more informative illustrated C budgets (mature forests); (2) the figures showing age trends and biome differences in 6 C flux and stock variables (Figs.
+7-8, formerly 6-7) are much easier to read; and (3) the new figure 9 provides an overall synthesis of results, focusing on the most trusted variables.**
 
-- *remove maps from age trend figures?*
-
-**Regarding the suggestion of showing age of old forests as a scatterplot, this is not possible give unknown ages of most mature tropical forests, as explained above.**
-
-**Regarding the proposed figure rearrangement, ...** *[issue #58](https://github.com/forc-db/ERL-review/issues/58)*
+**Regarding the suggestion of showing age of old forests as a scatterplot, this is unfortunately not possible give unknown ages of most mature tropical forests, as explained above.**
 
@@
 
-*No changes here; just respond....*
+**We prefer not to separate forests of known age into categories, but rather to represent age as a continuous variable, when possible.
+In large part this is motivated by the fact that appropriate age thresholds for various categories would vary across biomes, given differences in biomass accumulation rates.
+While our ideal would be to have no categorical divisions, unknown stand ages--particularly in the tropics--force us to divide the data into the categories used here.**
 
-**We prefer not to separate forests of known age into categories, but rather to represent age as a continuous variable. ...**
+**To provide a better representation of age trends, we created new figure 9.
+Although it does not break forests into additional age categories, it does address what seems to be at the core of this concern, which is a concise representation of differences across the full range of stand ages.**
 
@@
 
-*(No! Too much work!)*
-
-*I'm inclined to place less--not more--emphasis on representativeness: [Issue #60](https://github.com/forc-db/ERL-review/issues/60) *
+**We agree that such an analysis would be valuable, but time and resource constraints place it beyond the scope of what is currently feasible in this already very large analysis.
+We believe that the current consistency check is valuable as a first pass for identifying cases of extreme mis-matches, and therefore retain it as is.
+However, we agree that it is not the ideal test, and that it is not very stringent (given high standard deviations) and have therefore toned down claims that the budgets are "closed"/"internally consistent" throughout the manuscript (e.g., removed from abstract).
+We note that the discussion already stated that our criteria for closure was relatively loose (and this statement is unchanged in the revised version): "On the other, however, *ForC* derives data from multiple heterogeneous sources, and standard deviations within each biome are high; as a result, the standard for C closure is relatively loose [*c.f.* @houghton_terrestrial_2020]."**
 

### 2021-01-24 manuscript/ERL_global_C_review.Rmd
finished responses to R3
I think this is good to go.... But see issue #57

@@
 *Methods/Design.* Here, we draw from the Global Forest Carbon Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation.
-Specifically, we use `r n_records` *ForC* records for `r nrow(T1)` C cycle vaariables from `r n_areas_tot` geographic locations to characterize ensemble C budgets for four broad forest types -- tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga.
+Specifically, we use `r n_records` *ForC* records for `r nrow(T1)` C cycle variables from `r n_areas_tot` geographic locations to characterize ensemble C budgets for four broad forest types -- tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga.
 We include estimates for both mature and regrowth (age <100 years) forests, and quantify trends with stand age in regrowth forests for all variables with sufficient data.
@@
 Here, $NEP$ conists primarily of woody aboveground net primary production ($ANPP_{woody}$), while $ANPP_{woody.turnover}$ is the sum of woody mortality and branch turnover.
-Dotted lines refer to decomposition of potential 'legacy' organic material produced prior to the disturbance and remaining at the site (e.g., standing and fallen dead wood, $DW_{tot}$).", out.width='100%', fig.pos='H'}
+Dotted lines refer to decomposition of potential 'legacy' organic material produced prior to the disturbance and remaining at the site (e.g., standing and fallen dead wood, $DW_{tot}$; soil organic matter).
+Error bars on C stocks plot represent within-biome variability, wherein mean biomass is highest in the tropics, but maximum biomass is highest in temperate regions.", out.width='100%', fig.pos='H'}
 
@@
 
-Data were summarized to produce schematics of C cycling across the eight biome by age group combinations identified above.
+Data were summarized to produce schematics of C cycling for mature forets of each biome.
 To obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot.
@@
 When the effect of stand age was significant at p $\le$ 0.05 and when each biome had records for stands of at least 10 different ages, a biome $\times$ stand.age interaction was included in the model.
+We note that the logarithmic function fit in this analysis does not always correspond to theoretical expectations (Fig.
+`r schematic`); however, data limitations did not support fitting of functions with more parameters or reliable comparison of different functional forms.
+Within the data constraints, we deemed a logarithmic function to be the most appropriate functional form for the majority of variables.
 
@@
 
-For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets were generally consistent.
+For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets met our criteria for budget "closure".
 That is, component variables summed to within one standard deviation of their respective aggregate variables in all but one instance.
@@
 Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).
-Individual figures for each flux with sufficient data are given in the Supplement (Figs.
+Individual figures for each flux with sufficient data, along with maps showing geographic distribution of the data, are given in the Supplement (Figs.
 S4-S19).", out.width='100%', fig.pos='H'}
@@
 C stocks had less consistent patterns across biomes (Figs. `r stock_age_trends`, `r all_stock_age_trends`).
-For $B_{tot}$ and $B_{ag}$, tropical broadleaf forests had the highest biomass and boreal forests the lowest, with temperate broadleaf and needleleaf ($B_{ag}$ only) intermediate.
-For three variables that had been disproportionately sampled in the high-biomass forests of the US Pacific Northwest ($B_{ag-wood}$, $B_{foliage}$, and $B_{root-coarse}$), temperate conifer forests had significantly higher stocks than the other biomes, which were not significantly different from one another.
+For $B_{tot}$ and $B_{ag}$, tropical broadleaf forests had the highest mean biomass and boreal forests the lowest, with temperate broadleaf and needleleaf ($B_{ag}$ only) intermediate.
+However, maximum values for stocks including live or standing woody biomass ($B_{tot}$, $B_{ag}$, $B_{ag-wood}$, $B_{root}$, $B_{root_coarse}$, $DW_{tot}$, $DW_{standing}$) consistently occurred in temperate biomes (Figs.
+`r schematic`, `r stock_age_trends`, `r all_stock_age_trends`).
+For variables that were disproportionately sampled in such high-biomass forests ($B_{ag-wood}$, $B_{foliage}$, and $B_{root-coarse}$; disproportionately sampled in  the US Pacific Northwest), temperate conifer forests had significantly higher stocks than the other biomes.
 
@@
 Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).
-Individual figures for each stock with sufficient data are given in the Supplement (Figs.
+Individual figures for each stock with sufficient data, along with maps showing geographic distribution of the data, are given in the Supplement (Figs.
 S20-S30).", out.width='100%', fig.pos='H'}
@@
 The notable exception was mature forest $NEP$, which, as the difference between $GPP$ and $R_{eco}$, was indistinguishable across biomes.
-There was also little directional variation in mean mature forest C stocks across biomes, although maximum values for stocks including live or standing woody biomass ($B_{tot}$, $B_{ag}$, $B_{ag-wood}$, $B_{root}$, $B_{root_coarse}$, $DW_{tot}$, $DW_{standing}$) consistently occurred in temperate biomes (Figs.
+There was also little directional variation in mean mature forest C stocks across biomes, although maximum values for the majority of stocks (all including live or standing woody biomass) occurred in temperate biomes (Figs.
 `r schematic`, `r `all_diagrams_mature`, `r stock_age_trends`).
@@
 `r diagram_TeN_mature`) rather than issues of methodological accuracy.
-The overall high degree of closure implies that *ForC* gives a consistent picture of C cycling within biomes for mature forests. 
+The overall high degree of closure implies that *ForC* gives an at least roughly consistent picture of C cycling within biomes for mature forests. 
 This is an important and useful test, because it allows for consistency checks within the C cycle, for example leveraging separate and independently-measured fluxes to constrain errors in another [@phillips_value_2017; @williams_impacts_2014; @harmon_heterotrophic_2011], or producing internally consistent global data products [@wang_golum-cnp_2018].

### 2021-01-24 manuscript/response_to_reviews.Rmd
some rearrangmenet in discussion, main comment of R4
- discussion rearrangment suggested by @hmullerlandau (move section on C variable coverage and budget closure to later in the discussion)
- added paragraph on lack of representation of degraded and fragmented forests

@@
 To make it clear that we are not claiming to provide an independent test, we have edited this text as follows: "For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude (or increase with temperature) on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
-The consistency with which this occurs across numerous fluxes is not surprising, particularly given commonality in the data analyzed or used for calibration, but has never been simultaneously assessed across such a large number of variables [but see @banbury_morgan_global_nodate for nine autotrophic fluxes]."**
+This consistency is not not surprising, particularly given commonality in the data analyzed or used for calibration."**
 
@@
 
-*add discussion of the more intact stands studied here, vs what actually exists. straightforward.*
+**Thank you.
+We have added the following paragraph (citing the references given above) on lack of representation of fragmented and degraded forests in the Discussion section "C variable coverage and budget closure":**
+
+**Importantly, *ForC* and the analyses presented here cover the forests that have received research attention, which are not a representative sample of the world's existing forests--geographically or in terms of human impacts [@martin_mapping_2012].
+Geographically, all variables are poorly covered in Africa and Siberia (Fig.
+1), a common problem in the carbon-cycle community [@xu_contribution_2016; @schimel_effect_2015].
+In terms of human impacts, research efforts tend to focus on interior forest ecosystems [@martin_mapping_2012], often in permanently protected areas [e.g., @davies_forestgeo_2021].
+Studies of regrowth forests tend to focus on sites where recurring anthropogenic disturbance is not a confounding factor.
+Yet, fragmentation and degradation impact a large and growing proportion of Earth's forests [@fao_and_unep_state_2020].
+Fragmentation and the creation of edges strongly impacts forest C cycling [e.g., @chaplin-kramer_degradation_2015; @remy_strong_2016; @reinmann_edge_2017; @smith_evidence_2019; @reinmann_urbanization_2020;  @ordway_carbon_2020].Partial logging and other forms of non- stand clearing anthropogenic disturbance also strongly impact forest C cycling [e.g., @huang_long-term_2010; @piponiot_carbon_2016] but are under-studied [@sist_tropical_2015] and excluded from this analysis.
+Fragmented and degraded forests do not fit the idealized conceptual framework around which this review is structured (Fig.
+`r schematic`), yet their representation in models, sustainability assessments, and C accounting systems is critical to accurate accounting of C cycling in Earth's forests [e.g., @huang_long-term_2010; @reinmann_edge_2017; @smith_evidence_2019; @piponiot_can_2019].
+Additional research and synthesis are needed to fill critical gaps in our understanding of forest C cycling [e.g., @sist_tropical_2015].**
 
@@
 
-*straightforward--just a bit of text edit.*
+**Reworded to read, "Notably, net carbon sequestration ($NEP$) exhibits an overall increase with age across the first 100 years of stand development...".**
 

### 2021-01-24 manuscript/ERL_global_C_review.Rmd
some rearrangmenet in discussion, main comment of R4
- discussion rearrangment suggested by @hmullerlandau (move section on C variable coverage and budget closure to later in the discussion)
- added paragraph on lack of representation of degraded and fragmented forests

@@
 
-*work on this par:*
-
 *ForC* v3.0 provided unprecedented coverage of most major variables, yielding a broad picture of C cycling in the world's major forest biomes.
-Carbon cycling rates generally increased from boreal to tropical regions and with stand age.
+Carbon cycling rates generally increased from boreal to tropical regions and with stand age (Figs. `r schematic`, `r all_age_trends`).
 Specifically, most C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for regrowth as well as mature forests (Figs.
-`r flux_age_trends`- `r stock_age_trends`).
-The notable exception was mature forest $NEP$, which, as the difference between $GPP$ and $R_{eco}$, was indistinguishable across biomes.
+`r schematic`, `r flux_age_trends`- `r stock_age_trends`, `r age_trends_all`).
+The notable exception was mature forest $NEP$, which, as the difference between $GPP$ and $R_{eco}$, was indistinguishable across biomes (Fig. `r flux_age_trends`f).
 There was also little directional variation in mean mature forest C stocks across biomes, although maximum values for the majority of stocks (all including live or standing woody biomass) occurred in temperate biomes (Figs.
@@
 
-### C variable coverage and budget closure
-
-*HM: I recommend putting this later in the discussion.  *
-
-The large number of C cycle variables covered by ForC, and the general consistency among them, provide confidence that our overall reported mature forest means provide useful baselines for analysis -- with the caveats that they are unlikely to be accurate representations of C cycling for any particular forest, and that these sample means almost certainly do not represent true biome means (particularly for temperate conifer forests where high-biomass stands are over-represented in *ForC*).
-
-In this analysis, the C cycle budgets for mature forests (Figs.
-`r all_diagrams_mature`) generally “close”–-that is, the sums of component variables do not differ from the larger fluxes by more than one standard deviation.
-On the one hand, this reflects the general fact that ecosystem-scale measurements tend to close the C budget more easily and consistently than, for example, for energy balance [@stoy_data-driven_2013].
-On the other, however, *ForC* derives data from multiple heterogeneous sources, and standard deviations within each biome are high; as a result, the standard for C closure is relatively loose [*c.f.* @houghton_terrestrial_2020].
-The one instance where the C budgets doesn’t close is likely due to differences in the representation of forest types (*i.e.*, disproportionate representation of US Pacific NW for  $B_{root-coarse}$ relative to  $B_{root}$; Fig.
-`r diagram_TeN_mature`) rather than issues of methodological accuracy.
-The overall high degree of closure implies that *ForC* gives an at least roughly consistent picture of C cycling within biomes for mature forests. 
-This is an important and useful test, because it allows for consistency checks within the C cycle, for example leveraging separate and independently-measured fluxes to constrain errors in another [@phillips_value_2017; @williams_impacts_2014; @harmon_heterotrophic_2011], or producing internally consistent global data products [@wang_golum-cnp_2018].
-
-In contrast, age trends for young forests generally remain less clearly defined, in large part because their data records remain somewhat sparse for most variables (*i.e.*, have low representation of different geographical regions for any given age).
-While this analysis provides a first analysis of age trends in forest C cycling for multiple variables at a global scale, improved resolution of these trends will require additional data.
-
-There are of course notable holes in the ForC variable coverage (Fig. `r figure_map`) that limit the scope of our inferences here. 
-Notably, *ForC* currently has sparse--if any--coverage of fluxes to herbivores and higher consumers, along with the woody mortality ($M_{woody}$) and dead wood stocks (Table 1, Figs.
-`r DW_tot_fig`-`r DW_down_fig`).
-Geographically, all variables are poorly covered in Africa and Siberia, a common problem in the carbon-cycle community [@xu_contribution_2016; @schimel_effect_2015]. 
-*ForC* does not include soil carbon, which is covered by other efforts [e.g., @kochy_global_2015]. 
-*ForC* is not intended to replace databases that are specialized for particular parts of the C cycle analyses, e.g., aboveground biomass [@spawn_harmonized_2020], land-atmosphere fluxes [@baldocchi_fluxnet_2001], soil respiration [@jian_restructured_2020], or the human footprint in global forests [@magnani_human_2007].
-
 ### C cycling across biomes
@@
 For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude (or increase with temperature) on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
-The consistency with which this occurs across numerous fluxes is not surprising, particularly given commonality in the data analyzed or used for calibration, but has never been simultaneously assessed across such a large number of variables [but see @banbury_morgan_global_nodate for nine autotrophic fluxes].
+This consistency is not not surprising, particularly given commonality in the data analyzed or used for calibration.
+The finding that these patterns hold consistently across across numerous fluxes, while unsurprising, is novel to this analysis [but see @banbury_morgan_global_nodate for nine autotrophic fluxes].
 
@@
 
-Notably, net carbon sequestration ($NEP$) increases with age up to the 100-yr threshold examined here, with more pronounced patterns in temperate than boreal forests (Fig. `r flux_age_trends`). 
+Notably, net carbon sequestration ($NEP$) exhibits an overall increase with age across the first 100 years of stand development, with more pronounced patterns in temperate than boreal forests (Fig.
+`r flux_age_trends`).
 This finding is largely consistent with, but built from a far larger dataset than, previous studies showing an increase in $NEP$ across relatively young stand ages [@pregitzer_carbon_2004; @baldocchi_fluxnet_2001; @luyssaert_old-growth_2008].
@@
 Again, further study and synthesis of non-living C stocks across biomes and stand ages will be valuable to giving a more comprehensive picture.
+
+### C variable coverage and budget closure
+
+The large number of C cycle variables covered by ForC, and the general consistency among them, provide confidence that our overall reported mature forest means provide useful baselines for analysis -- with the caveats that they are unlikely to be accurate representations of C cycling for any particular forest, and that these sample means almost certainly do not represent true biome means (particularly for temperate conifer forests where high-biomass stands are over-represented in *ForC*).
+
+In this analysis, the C cycle budgets for mature forests (Figs.
+`r all_diagrams_mature`) generally “close”–-that is, the sums of component variables do not differ from the larger fluxes by more than one standard deviation.
+On the one hand, this reflects the general fact that ecosystem-scale measurements tend to close the C budget more easily and consistently than, for example, for energy balance [@stoy_data-driven_2013].
+On the other, however, *ForC* derives data from multiple heterogeneous sources, and standard deviations within each biome are high; as a result, the standard for C closure is relatively loose [*c.f.* @houghton_terrestrial_2020].
+The one instance where the C budgets doesn’t close is likely due to differences in the representation of forest types (*i.e.*, disproportionate representation of US Pacific NW for  $B_{root-coarse}$ relative to  $B_{root}$; Fig.
+`r diagram_TeN_mature`) rather than issues of methodological accuracy.
+The overall high degree of closure implies that *ForC* gives an at least roughly consistent picture of C cycling within biomes for mature forests. 
+This is an important and useful test, because it allows for consistency checks within the C cycle, for example leveraging separate and independently-measured fluxes to constrain errors in another [@phillips_value_2017; @williams_impacts_2014; @harmon_heterotrophic_2011], or producing internally consistent global data products [@wang_golum-cnp_2018].
+
+In contrast, age trends for young forests generally remain less clearly defined, in large part because their data records remain somewhat sparse for most variables (*i.e.*, have low representation of different geographical regions for any given age).
+While this analysis provides a first analysis of age trends in forest C cycling for multiple variables at a global scale, improved resolution of these trends will require additional data.
+
+There are of course notable holes in the ForC variable coverage that limit the scope of our inferences here. 
+Notably, *ForC* currently has sparse--if any--coverage of fluxes to herbivores and higher consumers, along with the woody mortality ($M_{woody}$) and dead wood stocks (Table 1, Figs.
+`r DW_tot_fig`-`r DW_down_fig`).
+*ForC* does not include soil carbon, which is covered by other efforts [e.g., @kochy_global_2015]. 
+*ForC* is not intended to replace databases that are specialized for particular parts of the C cycle analyses, e.g., aboveground biomass [@spawn_harmonized_2020], land-atmosphere fluxes [@baldocchi_fluxnet_2001], soil respiration [@jian_restructured_2020], or the human footprint in global forests [@magnani_human_2007].
+
+Importantly, *ForC* and the analyses presented here cover the forests that have received research attention, which are not a representative sample of the world's existing forests--geographically or in terms of human impacts [@martin_mapping_2012].
+Geographically, all variables are poorly covered in Africa and Siberia (Fig. `r figure_map`), a common problem in the carbon-cycle community [@xu_contribution_2016; @schimel_effect_2015]. 
+In terms of human impacts, research efforts tend to focus on interior forest ecosystems [@martin_mapping_2012], often in permanently protected areas [e.g., @davies_forestgeo_2021].
+Studies of regrowth forests tend to focus on sites where recurring anthropogenic disturbance is not a confounding factor.
+Yet, fragmentation and degradation impact a large and growing proportion of Earth's forests [@fao_and_unep_state_2020].
+Fragmentation and the creation of edges strongly impacts forest C cycling [e.g., @chaplin-kramer_degradation_2015; @remy_strong_2016; @reinmann_edge_2017; @smith_evidence_2019; @reinmann_urbanization_2020;  @ordway_carbon_2020].
+Partial logging and other forms of non- stand clearing anthropogenic disturbance also strongly impact forest C cycling [e.g., @huang_long-term_2010; @piponiot_carbon_2016] but are under-studied [@sist_tropical_2015] and excluded from this analysis.
+Fragmented and degraded forests do not fit the idealized conceptual framework around which this review is structured (Fig.
+`r schematic`), yet their representation in models, sustainability assessments, and C accounting systems is critical to accurate accounting of C cycling in Earth's forests [e.g., @huang_long-term_2010; @reinmann_edge_2017; @smith_evidence_2019; @piponiot_can_2019].
+Additional research and synthesis are needed to fill critical gaps in our understanding of forest C cycling [e.g., @sist_tropical_2015].
 
@@
 Thanks to all researchers whose data are included in *ForC* and this analysis, to Jennifer McGarvey and Ian McGregor for help with the database, and to Norbert Kunert for helpful discussion.
+Three anonymous reviewers provided helpful feedback. 
 Funding sources included a Smithsonian Scholarly Studies grant to KAT and HML and a Smithsonian Working Land and Seascapes grant to KAT.

### 2021-01-24 manuscript/response_to_reviews.Rmd
fixed knit

@@
 Fragmented and degraded forests do not fit the idealized conceptual framework around which this review is structured (Fig.
-`r schematic`), yet their representation in models, sustainability assessments, and C accounting systems is critical to accurate accounting of C cycling in Earth's forests [e.g., @huang_long-term_2010; @reinmann_edge_2017; @smith_evidence_2019; @piponiot_can_2019].
+1), yet their representation in models, sustainability assessments, and C accounting systems is critical to accurate accounting of C cycling in Earth's forests [e.g., @huang_long-term_2010; @reinmann_edge_2017; @smith_evidence_2019; @piponiot_can_2019].
 Additional research and synthesis are needed to fill critical gaps in our understanding of forest C cycling [e.g., @sist_tropical_2015].**

### 2021-01-24 manuscript/response_to_reviews.Rmd
finished response to reviews! (...pending final check)

@@
 
-**Thank you.
-We have added the following paragraph (citing the references given above) on lack of representation of fragmented and degraded forests in the Discussion section "C variable coverage and budget closure":**
+**Thank you. We have added the following paragraph (citing the references given above) on representation of forest types in the Discussion section "C variable coverage and budget closure":**
 
@@
 1), yet their representation in models, sustainability assessments, and C accounting systems is critical to accurate accounting of C cycling in Earth's forests [e.g., @huang_long-term_2010; @reinmann_edge_2017; @smith_evidence_2019; @piponiot_can_2019].
-Additional research and synthesis are needed to fill critical gaps in our understanding of forest C cycling [e.g., @sist_tropical_2015].**
+Finally, plantation forests account for approximately 3% of Earth's forests [@fao_and_unep_state_2020] but are not included in this analysis.
+While it is known that these tend to accumulate biomass faster than naturally regenerating forests [@anderson_temperature-dependence_2006; @bonner_meta-analytical_2013], their global scale C cycling patterns remain less clearly understood [c.f.
+@cook-patton_mapping_2020].
+Additional research and synthesis are needed to fill these critical gaps in our understanding of forest C cycling.**
 
@@
 
-*straightforward--just a bit of text*
+**We have added a statement on plantation forests in the paragraph representation of forest types: Finally, plantation forests account for approximately 3% of Earth's forests [@fao_and_unep_state_2020] but are not included in this analysis.
+While it is known that these tend to accumulate biomass faster than naturally regenerating forests [@anderson_temperature-dependence_2006; @bonner_meta-analytical_2013], their global scale C cycling patterns remain less clearly understood [c.f.
+@cook-patton_mapping_2020]."**
 
@@
 
-*straightforward--just a bit of text edit.*
+**We have reworded this sentence to read, "Age $\times$ biome interactions were also significant for all ten of these C stock variables (Table S2), with living C stocks tending to accumulate more rapidly during the early stages of forest regrowth in tropical forests (Figs.
+8, S20-S30)."**
 
@@
 
-*It strikes me that both R3 and R4 find this surprising. I don't. An Odum-esque conceptual diagram + some discussion should help to explain this.*
+**We have added the following sentence to the first paragraph of the discussion: "The notable exception was mature forest $NEP$, which, as the difference between $GPP$ and $R_{eco}$, was statistically indistinguishable across biomes (Fig.
+7f)." Given that there is no signficant difference among biomes, we prefer not to discuss temperate broadleaf forests as higher.**
 
@@
 
-*straightforward--just a bit of text edit.*
+**We have added the following statement to the methods section: "We did not analyze soil carbon, which is not a focus of the *ForC* database." We also now note this in the first paragraph of the discussion "There was also little directional variation in mean mature forest C stocks (biomass, dead wood, and organic layer) across biomes, although maximum values for the majority of stocks (all including live or standing woody biomass) occurred in temperate biomes (Figs.
+1, 2-5, 8)."**
 
@@
 
-*straightforward--just a bit of text edit.*
+**We have modified sentences in this paragraph to address this: "The notable exception to the pattern of fluxes decreasing from tropical to boreal regions is $NEP$, which showed no significant differences across biomes, albeit with the highest mean in temperate broadleaf forests (Fig.
+7f).
+...
+The fact that mature temperate broadleaf forests have a higher mean than the other biomes may reflect the fact that most of these forests are older secondary forests that, while classified here as mature, are still accumulating carbon [@curtis_forest_2018].**
 
@@
 
-*straightforward--just a bit of text edit.*
+**This question is now addressed earlier in the discussion (see response to comment on L322-323).
+The higher NEP of mature temperate forests contributes to their value for climate change mitigation, but does not make them more valuable than higher-biomass tropical forests (for example) that have already sequestered C that would become committed to eventual release as CO~2~ emissions if the forest were to be cleared (with the rapidity of that release depending on mode of clearing, wood product decay...).
+That trade-off is explored to some extent in one of the references given in this paragraph [@anderson-teixeira_greenhouse_2011], but seems too complex to get into here.**
 

### 2021-01-24 manuscript/ERL_global_C_review.Rmd
finished response to reviews! (...pending final check)

@@
 These different flux and stock variables represent different pools (e.g., aboveground biomass, root biomass, dead wood) and levels of combination (e.g., total net primary productivity, $NPP$, versus the individual elements of $NPP$ such as foliage, roots, and branches).
+We did not analyze soil carbon, which is not a focus of the *ForC* database.
 Note that two flux variables, aboveground heterotropic respiration ($R_{het-ag}$) and total respiration ($R_{het}$), were included for conceptual completeness but had no records in *ForC* (Table 1).
@@
 `r age_trends_all`).
-ALthough age trends of young forests often converged towards mature forest averages, there were also some  discrepancies between young forest trends and mature forest averages (Figs.
+Although age trends of young forests often converged towards mature forest averages, there were also some  discrepancies between young forest trends and mature forest averages (Figs.
 `r flux_age_trends`, `r age_trends_all`, `r all_age_trends`), most notably including a tendency for higher fluxes in regrowth boreal forests than in their mature counterparts (Fig.
@@
 `r schematic`, `r flux_age_trends`- `r stock_age_trends`, `r age_trends_all`).
-The notable exception was mature forest $NEP$, which, as the difference between $GPP$ and $R_{eco}$, was indistinguishable across biomes (Fig. `r flux_age_trends`f).
-There was also little directional variation in mean mature forest C stocks across biomes, although maximum values for the majority of stocks (all including live or standing woody biomass) occurred in temperate biomes (Figs.
-`r schematic`, `r `all_diagrams_mature`, `r stock_age_trends`).
+The notable exception was mature forest $NEP$, which, as the difference between $GPP$ and $R_{eco}$, was statistically indistinguishable across biomes (Fig. `r flux_age_trends`f).
+There was also little directional variation in mean mature forest C stocks (biomass, dead wood, and organic layer) across biomes, although maximum values for the majority of stocks (all including live or standing woody biomass) occurred in temperate biomes (Figs.
+`r schematic`, `r all_diagrams_mature`, `r stock_age_trends`).
 Consistent with theory and previous studies (Fig.
@@
 
-The notable exception to the pattern of fluxes decreasing from tropical to boreal regions is $NEP$, which showed no significant differences across biomes (Fig. `r flux_age_trends`f).
+The notable exception to the pattern of fluxes decreasing from tropical to boreal regions is $NEP$, which showed no significant differences across biomes, albeit with the highest mean in temperate broadleaf forests (Fig.
+`r flux_age_trends`f).
 Unlike the other C flux variables, $NEP$ does not characterize the rate at which C cycles through the ecosystem, but is the balance between C sequestration ($GPP$) and respiratory losses ($R_{eco}$) and represents net CO~2~ sequestration (or release) by the ecosystem (Fig.
 `r schematic`).
-$NEP$ tends to be relatively small in mature forest stands (discussed further below), which accumulate carbon slowly relative to younger stands, if at all [@luyssaert_old-growth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
+$NEP$ tends to be relatively small in mature forest stands, which accumulate carbon slowly relative to younger stands (discussed further below), if at all [Fig.
+`r schematic`; @luyssaert_old-growth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
 It is therefore unsurprising that there are no pronounced differences across biomes [in agreement with @luyssaert_co2_2007], suggesting that variation in $NEP$ of mature forests is controlled less by climate and more by other factors including moderate disturbances [@curtis_forest_2018] or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition; @wilson_stability_2016].
+The fact that mature temperate broadleaf forests have a higher mean than the other biomes may reflect the fact that most of these forests are older secondary forests that, while classified here as mature, are still accumulating carbon [@curtis_forest_2018].
 
@@
 7a, S16-S22; @anderson_temperature-dependence_2006; @cook-patton_mapping_2020].
-Note, though, that $NEP$ includes not only $\Delta B_{tot}$, but also changes in $DW_{tot}$, $OL$, and soil carbon, and biome differences in the accumulation rates of these variables have not been detected, in part because these variables do not consistently increase with stand age [Figs.
-`r stock_age_trends`, `r DW_tot_fig`-`r OL_fig`, and see discussion below; @cook-patton_mapping_2020].
+Note, though, that $NEP$ includes not only $\Delta B_{tot}$, but also changes in $DW_{tot}$, $OL$, and soil carbon (not analyzed here), and biome differences in the accumulation rates of these variables have not been detected, in part because these variables do not consistently increase with stand age [Figs.
+`r schematic`, `r stock_age_trends`, `r DW_tot_fig`-`r OL_fig`, and see discussion below; @cook-patton_mapping_2020].
 
@@
 `r schematic`), yet their representation in models, sustainability assessments, and C accounting systems is critical to accurate accounting of C cycling in Earth's forests [e.g., @huang_long-term_2010; @reinmann_edge_2017; @smith_evidence_2019; @piponiot_can_2019].
-Additional research and synthesis are needed to fill critical gaps in our understanding of forest C cycling [e.g., @sist_tropical_2015].
+Finally, plantation forests account for approximately 3% of Earth's forests [@fao_and_unep_state_2020] but are not included in this analysis. 
+While it is known that these tend to accumulate biomass faster than naturally regenerating forests [@anderson_temperature-dependence_2006; @bonner_meta-analytical_2013], their global scale C cycling patterns remain less clearly understood [c.f.
+@cook-patton_mapping_2020].
+Additional research and synthesis are needed to fill these critical gaps in our understanding of forest C cycling.
 

### 2021-01-24 manuscript/ERL_global_C_review.Rmd
references

@@
 Global scale data syntheses have shown that C fluxes including $GPP$, net primary productivity ($NPP$), and soil respiration ($R_{soil}$) decrease with latitude or, correspondingly, increase with mean annual temperature and, to a lesser extent, precipitation [Fig.
-`r schematic_fluxes_biome` **REFS**; @lieth_primary_1973;  @luyssaert_co2_2007; @hursh_sensitivity_2017; @banbury_morgan_global_nodate].
+`r schematic_fluxes_biome`; e.g., @lieth_primary_1973;  @luyssaert_co2_2007; @hursh_sensitivity_2017; @banbury_morgan_global_nodate].
 C stocks of mature forests show less directional variation (Fig. `r schematic_fluxes_biome`).
@@
 
-Correlative analyses relating C cycle variables to climate and other environmental variables have recently been taken to a new level through use of machine-learning algorithms that relate ground-based C cycle data to *global data on climate/soils/satellite data*, making it possible to create fine-scale global maps of C cycling [e.g., **REFS**;@warner_spatial_2019; @cook-patton_mapping_2020].
-*This approach can be particularly effective when paired with satellite data ...
-(e.g., aboveground biomass: Saatchi et al 2011, Hu et al 2016, Spawn et al 2020, gross primary productivity, GPP: Li and Xiao 2019, ).*
-Any such analysis is however constrained by the quality and coverage of ground-based estimates of forest C fluxes or stocks. 
+Correlative analyses relating C cycle variables to climate and other environmental variables have recently been taken to a new level through use of machine-learning algorithms that relate ground-based C cycle data to global maps of environmental covariates, making it possible to create fine-scale global maps of C cycling [e.g., @warner_spatial_2019; @cook-patton_mapping_2020].
+This approach can be particularly effective when paired with satellite measurements that correlate to C cycle variables of interest; for example, solar-induced chlorophyll fluorescence is useful for fine-scale mapping of gross primary productivity [$GPP$; @li_mapping_2019], while LiDAR, radar, and optical imagery is being used to estimate $B_{ag}$ at regional to global scales [e.g., @saatchi_benchmark_2011; @hu_mapping_2016].
+Any such analysis is however constrained by the quality and coverage of ground-based estimates of forest C fluxes or stocks [e.g., @schepaschenko_forest_2019]. 
 While estimates of some variables (e.g., $B_{ag}$, $GPP$, $NPP$, $R_{soil}$) are widely available, many remain poorly characterized (e.g., $DW$; autotrophic respiration, $R_{auto}$) --even at the coarse resolution of biomes.
-This is a critical limitation not only for understanding forest C cycling, but also for quantifying climate change mitigation across forest biomes or ecozones [e.g., **REFS**].
+This is a critical limitation not only for understanding forest C cycling, but also for quantifying forest-based climate change mitigation across forest biomes or ecozones [e.g., @ipcc_2019_2019].
 
@@
 Although the conceptualization of the C cycle in this paper is simplistic by current standards, the paper was foundational in framing the theory around which research on the subject still revolves [@corman_foundations_2019], and the basic framework still holds, albeit with modest modifications [@andersonteixeira_altered_2013].
-Following stand-clearing disturbance, $GPP$, $NPP$, and biomass of leaves ($B_{foliage}$) and fine roots ($B_{root-fine}$) increase rapidly and thereafter remain relatively stable ($B_{foliage}$,$B_{root-fine}$, sometimes $GPP$) or decline slightly [$NPP$, sometimes $GPP$; e.g., *Goulden et al.
-2011, ....
-refs in * @andersonteixeira_altered_2013].
+Following stand-clearing disturbance, $GPP$, $NPP$, and biomass of leaves ($B_{foliage}$) and fine roots ($B_{root-fine}$) increase rapidly and thereafter remain relatively stable ($B_{foliage}$,$B_{root-fine}$, sometimes $GPP$) or decline slightly [$NPP$, sometimes $GPP$; e.g., @law_changes_2003; @pregitzer_carbon_2004; @amiro_ecosystem_2010; @goulden_patterns_2011].
 The decline in $NPP$ occurs because $R_{auto}$ increases relative to $GPP$ as forests age, corresponding to declining carbon use efficiency with stand age [@delucia_forest_2007; @collalti_forest_2020].
-Heterotrophic respiration, most of which originates from the soil ($R_{het-soil}$) remains relatively constant with stand age *[Law et al., 2003; Pregitzer & Euskirchen, 2004; Goulden et al., 2011]*, with the result that net ecosystem production ($NEP=GPP-R_{eco}$, where $R_{eco}$ is total ecosystem respiration) is initially negative, increases to a maximum at intermediate ages, and declines thereafter *[Law et al., 2003; Pregitzer & Euskirchen, 2004; Zhou et al., 2006; Baldocchi, 2008; Luyssaert et al., 2008; Amiro et al., 2010; Goulden et al., 2011]*.
-The result is that biomass accumulates rapidly in young forests, followed by a slow decline to near zero in old forests *[Lichstein et al., 2009; Yang et al., 2011;
-Hember et al., 2012]*.
+Heterotrophic respiration, most of which originates from the soil ($R_{het-soil}$) remains relatively constant with stand age [@law_changes_2003; @pregitzer_carbon_2004; @goulden_patterns_2011], with the result that net ecosystem production ($NEP=GPP-R_{eco}$, where $R_{eco}$ is total ecosystem respiration) is initially negative, increases to a maximum at intermediate ages, and declines thereafter [@law_changes_2003; @pregitzer_carbon_2004; @amiro_ecosystem_2010; @goulden_patterns_2011; @luyssaert_old-growth_2008].
+The result is that biomass accumulates rapidly in young forests, followed by a slow decline to near zero in old forests [e.g., @lichstein_biomass_2009; @yang_carbon_2011].
 While these trends have been subject of fairly recent qualitative review [@andersonteixeira_altered_2013], there is need for a synthetic, quantitative review taking advantage of the greatly expanded data now available.
@@
 In the past several decades, researchers have started asking how age trends--mostly in $B_{ag}$ or total biomass ($B_{tot}$) accumulation-- vary across biomes.
-Early research on this theme showed that biomass accumulation rates during secondary succession increase with temperature on a global scale [**REFS**; @anderson_temperature-dependence_2006] and *with precipitation in the neotropics* [**REFS**; @poorter_biomass_2016; @chazdon_carbon_2016].
+Early research on this theme showed that biomass accumulation rates during secondary succession increase with temperature on a global scale [@johnson_post-disturbance_2000; @anderson_temperature-dependence_2006] and with water availability in the neotropics [@poorter_biomass_2016].
 Most recently, @cook-patton_mapping_2020 reinforced these earlier findings with a much larger dataset and crated a high-resolution global map of estimated potential C accumulation rates.
@@
 Given the important role of secondary forests in the current and future global C cycle, concrete understanding of age trends in C fluxes and stocks and how these vary across biomes is critical to better understanding of the global C cycle.
-Accurate estimates of C sequestration rates by regrowth forests are also critical for national greenhouse gas accounting under the IPCC framework [**REFS**] and to quantifying the value of regrwoth forests for climate change mitigation [@anderson-teixeira_greenhouse_2011; @goldstein_protecting_2020].
+Accurate estimates of C sequestration rates by regrowth forests are also critical for national greenhouse gas accounting under the IPCC framework [@ipcc_2019_2019; @requena_suarez_estimating_2019] and to quantifying the value of regrowth forests for climate change mitigation [@anderson-teixeira_greenhouse_2011; @goldstein_protecting_2020].
 
@@
 This finding is largely consistent with, but built from a far larger dataset than, previous studies showing an increase in $NEP$ across relatively young stand ages [@pregitzer_carbon_2004; @baldocchi_fluxnet_2001; @luyssaert_old-growth_2008].
-However, $NEP$ has been observed to decline from intermediate to old stands [@luyssaert_old-growth_2008], and the $NEP$ estimated by our model for 100-year-old temperate conifer stands (~5 Mg C ha^-1^ yr^-1^) exceeds the mean of mature forests in the same biome (0.7 Mg C ha^-1^ yr^-1^; Fig.
+However, $NEP$ has been observed to decline from intermediate to old stands [ @law_changes_2003; @luyssaert_old-growth_2008], and the $NEP$ estimated by our model for 100-year-old temperate conifer stands (~5 Mg C ha^-1^ yr^-1^) exceeds the mean of mature forests in the same biome (0.7 Mg C ha^-1^ yr^-1^; Fig.
 `r diagram_TeN_mature`).

### 2021-01-24 manuscript/ERL_global_C_review.Rmd
references

@@
 This requires accessible, comprehensive, and large-scale databases with global coverage, which runs contrary to the traditional way forest C stocks and fluxes have been measured and published.
-Large-scale synthesis is critical to benchmarking model performance with global data (Clark et al 2017, Luo et al 2012), quantifying the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], and using book-keeping methods to quantify actual or potential exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
+Large-scale synthesis is critical to benchmarking model performance with global data [@clark_field_2017; @luo_framework_2012], quantifying the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], and using book-keeping methods to quantify actual or potential exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
 
@@
 Most of this enormous C uptake is counterbalanced by releases to the atmosphere through ecosystem respiration ($R_{eco}$) and fire, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] and fire [@van_der_werf_global_2017].
-In recent years, total forest C uptake has exceeded releases, such that globally forests have been a C sink.
+In recent years, total forest C uptake has exceeded releases, such that globally forests have been a C sink [@harris_global_2021].
 Considering only areas remaining in forest, this C sink has averaged 3.2 $\pm$ 0.6 Gt C yr^-1^ for 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
-However, deforestation, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reduces the net forest sink to ~1.1-2.2 Gt C yr^-1^ [@friedlingstein_global_2019].
+However, deforestation, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reduces the net forest sink to ~1.1-2.2 Gt C yr^-1^ [@friedlingstein_global_2019; @harris_global_2021].
 
@@
 
-###  Citations to add
-
-@harris_global_2021
-
 ## Acknowledgements

### 2021-01-24 manuscript/response_to_reviews.Rmd
letter to editor
#67

@@
 
+Dear Editor:
+
+We are please to submit a revised version of our manuscript, "Carbon cycling in mature and regrowth forests globally" (ERL-109898), for consideration for publication in *Environmental Research Letters*.
+We have addressed all points raised by the reviewers, as detailed below.
 The most substantive changes included the following:
 
-- Changed the figure set in accordance with comments from R3. Specifically, we
-(1) added a schematic figure, somewhat in the style of @odum_strategy_1969, to help with conceptual framing of the review
-(2) added a figure synthesizing changes in fluxes through time for each biome;
-(3) removed maps from age trend diagrams (maps still available in SI), which makes the figures less busy/ easier to read; and
-(4) moved C cycle diagrams for young forests to SI;
+- We changed the figure set in accordance with comments from R3.
+Specifically, we (1) added a schematic figure, somewhat in the style of @odum_strategy_1969, to help with conceptual framing of the review; (2) added a figure synthesizing changes in fluxes through time for each biome; (3) removed maps from age trend diagrams (maps still available in SI), which makes the figures less busy/ easier to read; and (4) moved C cycle diagrams for young forests to SI;
 
-- Improved conceptual framework and "review" nature of the work by adding background information to the introduction, including the schematic figure mentioned above.
+- We improved conceptual framework and "review" nature of the work by adding background information to the introduction, including the schematic figure mentioned above.
 We also shifted emphasis more towards the ecological questions (trends across biomes and with age), while retaining (with less emphasis) the analysis of database representativeness.
+
+Thank you for considering this revised version, and we look forward to your response.
+
+Sincerely,
+
+Kristina Anderson-Teixeira (on behalf of all coauthors)
 

### 2021-01-25 manuscript/response_to_reviews.Rmd
various small fixes

@@
 **We acknowledge and discuss lack of closure in various places throughout the results and discussion.
-To ensure that this is clear from the figures, we have added the following statement to the captions of Figures 3-6 (formerly 2-5): "Note that variables differ in geographical representation, resulting in potential imbalances (Figs.
-S5-S30).
-Probability that estimates reflect true biome means scales with the number of distinct geographical areas represented." We have also added a similar statement to the captions of Figures S1-S4 (formerly 8-11).
+To ensure that this is clear from the figures, we have added the following statement to the captions of Figures 3-6 (formerly 2-5): "Mean component fluxes do not necessarily add up to the mean total fluxes because different sets of sites are included depending on availability of data (Figs.
+S5-S30)." We have also added a similar statement to the captions of Figures S1-S4 (formerly 8-11).
 We also make a related statement in the captions of new figures 9-12: "Note that there remain substantial uncertainties as to the functional form of age trends and discrepancies in closure among related variables."**
@@
 
-**We have addressed this by adding the following statement to the captions of Figures 3-6 (formerly 2-5): "Probability that estimates reflect true biome means scales with the number of distinct geographical areas represented."**
+**We have done two things to address this comment.
+First, to ensure accurate interpretation, we added the following statement to the captions of Figures 3-6 (formerly 2-5): "Mean component fluxes do not necessarily add up to the mean total fluxes because different sets of sites are included depending on availability of data (Figs.
+S5-S30)." Second, we have added the following guidance on using the data to the discussion: "*ForC*’s tens of thousands of records are readily available in a standardized format, along with all code used in the analyses presented here, and we recommend that researchers use these resources to identify and summarize data specific the analysis at hand."**
 
@@
 
-**The 147 refers to *total* annual terrestrial photosynthesis, including non-forests.
-Table S4 breaks this down by biome.
-The >69 Gt C/yr figure was obtained by summing the forest biomes, not including savannas.**
+**The 147 Pg C/year figure refers to *total* annual terrestrial photosynthesis, including non-forests.
+Table S4 in @badgley_terrestrial_2019 breaks this down by biome.
+The >69 Gt C/yr figure was obtained by summing the forest biomes, not including savannas/ mixed biomes.**
 

### 2021-01-25 manuscript/ERL_global_C_SI.Rmd
working on responses to reviews
various small fixes

@@
 Presented are observed ranges, where geographically distinct areas are treated as the unit of replication.
+Note that variables differ in geographical and age representation, resulting in potential imbalances (Figs.
+S5-S30).
 All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks).
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-To illustrate the magnitude of different fluxes, arrow size is proportional to the square root of corresponding flux.", out.width='87%', fig.pos='H'}
+To illustrate the magnitude of different fluxes, arrow width is proportional to the square root of the mean of the corresponding flux.", out.width='87%', fig.pos='H'}
 
@@
 Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
-In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
-The fitted line indicates the effect of age on flux (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant age x biome interaction.
+In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.

### 2021-01-25 manuscript/ERL_global_C_review.Rmd
various small fixes

@@
 Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
-Note that variables differ in geographical representation, resulting in potential imbalances (Figs.
-S5-S30).
-Probability that estimates reflect true biome means scales with the number of distinct geographical areas represented.
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
 To illustrate the magnitude of different fluxes, arrow width is proportional to the square root of the corresponding flux.
+Mean component fluxes do not necessarily add up to the mean total fluxes because different sets of sites are included depending on availability of data (Figs.
+S5-S30).
 An asterisk after a variable name indicates lack of C cycle closure."
@@
 ```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEP$.
-In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
-The fitted line indicates the effect of age on flux (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant age x biome interaction.
+In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
 The boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.
-In each panel, the left scatterplot  shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of age and biome.
-The fitted line indicates the effect of age on flux (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant age x biome interaction.
+In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
 The boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.
@@
 To ensure that models are giving the right answers for the right reasons [@sulman_multiple_2018], it is important to benchmark against multiple components of the C cycle that are internally consistent with each other [@collier_international_2018; @wang_golum-cnp_2018].
-*ForC*’s tens of thousands of records are readily available in a standardized format, and our analyses here indicate that their internal consistency is reasonably high.
+*ForC*’s tens of thousands of records are readily available in a standardized format, along with all code used in the analyses presented here, and we recommend that researchers use these resources to identify and summarize data specific the analysis at hand.
 Integration of *ForC* with models will be valuable to improving the accuracy and reliability of models [@fer_beyond_2021].

### 2021-01-25 manuscript/tables_figures/stacked_age_trends/schematic.m

@@ -145,7 +145,7 @@ h=bar(mature_stocks, 'stacked'); hold on;
     for n=1:size(mature_stocks,2)
         h(n).FaceColor= facecolor_stocks(n,:);
     end
-er=errorbar([1:3],sum(mature_stocks, 2), 0.5*stocks_std, stocks_std);
+er=errorbar([1:3],sum(mature_stocks, 2), 50-sum(mature_stocks, 2), stocks_std);
 er.Color = [0 0 0];                            
 er.LineStyle = 'none'; 
 er.LineWidth= 1.7;

### 2021-01-25 manuscript/ERL_global_C_review.Rmd
proofread through Figs 3-6

@@
 S5-S30).
-An asterisk after a variable name indicates lack of C cycle closure."
+"
 TrB_mature_caption = paste( "Figure 3 | C cycle diagram for mature tropical broadleaf forests. ", schematics_caption_generic)  
 TeB_mature_caption = paste( "Figure 4 | C cycle diagram for mature temperate broadleaf forests. ", schematics_caption_generic)  
-TeN_mature_caption = paste( "Figure 5 | C cycle diagram for mature temperate conifer forests. ", schematics_caption_generic)  
+TeN_mature_caption = paste( "Figure 5 | C cycle diagram for mature temperate conifer forests.
+", schematics_caption_generic, "The temperate conifer biome in particular is subject to high variability, with highest fluxes and stocks in the high-biomass forests of the US Pacific Northwest.
+An asterisk after a variable name indicates lack of C cycle closure.")
 BoN_mature_caption = paste( "Figure 6 | C cycle diagram for mature boreaal conifer forests. ", schematics_caption_generic)
@@
 
-*Background.* Forests are major components of the global carbon (C) cycle and thereby strongly influence atmospheric carbon dioxide (CO2) and climate.
+*Background.* Forests are major components of the global carbon (C) cycle and thereby strongly influence atmospheric carbon dioxide (CO~2~) and climate.
 However, efforts to incorporate forests into climate models and CO~2~ accounting frameworks have been constrained by a lack of accessible, global-scale synthesis on how C cycling varies across forest types and stand ages.
@@
 *Methods/Design.* Here, we draw from the Global Forest Carbon Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation.
-Specifically, we use `r n_records` *ForC* records for `r nrow(T1)` C cycle variables from `r n_areas_tot` geographic locations to characterize ensemble C budgets for four broad forest types -- tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga.
-We include estimates for both mature and regrowth (age <100 years) forests, and quantify trends with stand age in regrowth forests for all variables with sufficient data.
+Specifically, we use `r n_records` *ForC* records for `r nrow(T1)` C cycle variables from `r n_areas_tot` geographic locations to characterize ensemble C budgets for four broad forest types -- tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and boreal.
+We calculaate statistics for both mature and regrowth (age <100 years) forests, and quantify trends with stand age in regrowth forests for all variables with sufficient data.
 
-*Review Results/ Synthesis.*   The rate of C cycling generally increased from boreal to tropical regions in both mature and regrowth forests, whereas C stocks showed less directional variation.
+*Review Results/ Synthesis.*   The rate of C cycling generally decreased from tropical to boreal regions in both mature and regrowth forests, whereas C stocks showed less directional variation.
 Net ecosystem production of mature forests was indistinguishable across biomes.
@@
 Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide [CO~2~; @bonan_forests_2008; @friedlingstein_global_2019; @ipcc_global_2018].
-Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climatecarbon_2006; @krause_large_2018; @bonan_model_2019; @di_vittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of understanding on how C cycling varies across forest types and in relation to stand history.
+Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climatecarbon_2006; @krause_large_2018; @bonan_model_2019; @di_vittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011; @ipcc_2019_2019] can be traced to lack of understanding on how C cycling varies across forest types and in relation to stand history.
 This requires accessible, comprehensive, and large-scale databases with global coverage, which runs contrary to the traditional way forest C stocks and fluxes have been measured and published.
@@
 A robust understanding of forest impacts on global C cycling is essential. 
-Total annual photosynthesis in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], more than seven times the average annual fossil fuel emissions during 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
+Total annual photosynthesis in forests (gross primary productivity, $GPP$) is estimated at approximately 69 Gt C yr^-1^ [@badgley_terrestrial_2019], more than seven times the average annual fossil fuel emissions during 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
 Most of this enormous C uptake is counterbalanced by releases to the atmosphere through ecosystem respiration ($R_{eco}$) and fire, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] and fire [@van_der_werf_global_2017].
@@
 
-The future of the current forest C sink is dependent both upon forest responses to climate change itself and human land use decisions, which will feedback and strongly influence the course of climate change.
+The future of the current forest C sink is dependent both upon forest responses to climate change itself and human land use decisions, which will feedback and strongly influence the course of climate change [@friedlingstein_climatecarbon_2006].
 Regrowing forests in particular will play an important role [@pugh_role_2019], as almost two-thirds of the world’s forests were secondary as of 2010 [@fao_global_2010].
 As anthropogenic and climate-driven disturbances impact an growing proportion of Earth's forests [@andela_human-driven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests is increasingly important [@andersonteixeira_altered_2013].
-Although age trends in aboveground biomass have been relatively well-studied and synthesized globally [@cook-patton_mapping_2020], a relative dearth of data and synthesis on other C stocks and fluxes in secondary forests points to an under-filled need to characterize age-related trends in forest C cycling.
+Although age trends in aboveground biomass have been well-studied and synthesized globally [@cook-patton_mapping_2020], a relative dearth of data and synthesis on other C stocks and fluxes in secondary forests points to an under-filled need to characterize age-related trends in forest C cycling.
 Such understanding is particularly critical for reducing uncertainty regarding the potential for carbon uptake and climate change mitigation by regrowth forests [@krause_large_2018; @cook-patton_mapping_2020].
@@
 New technology has also enabled researchers to directly measure an expanding set of variables, notably including the development of continuous measurements of soil CO~2~ efflux [@kuzyakov_sources_2006] and ecosystem-atmosphere CO~2~ exchange [@baldocchi_fluxnet_2001].
-Measurement techniques have been increasingly standardized; for example, of the biomass allometries that strongly influence estimates of most C cycle variables [e.g., @chojnacky_updated_2014; @chave_improved_2014].
+Measurement techniques have been increasingly standardized; for example, of the biomass allometries that strongly influence estimates of most C cycle variables [e.g., @chave_improved_2014].
 Further standardization has been made possible through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; @davies_forestgeo_2021], NEON [@schimel_neon_2007], and FLUXNET [@baldocchi_fluxnet_2001; @novick_ameriflux_2018].
@@
 The current largest, most comprehensive database on forest C cycling is *ForC* [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], which contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables), with different variables capturing distinct ecosystem pools (e.g., woody, foliage, and root biomass; dead wood) and flux types (e.g, gross and net primary productivity; soil, root, and ecosystem respiration).
-These data are ground-based measurements, and *ForC* contains associated data required for interpretation (*e.g.*, stand history, measurement methods).
-Since its most recent publication [@anderson-teixeira_forc_2018], *ForC* has grown `r round((nrow(ForC_measurements) - 17367)/17367*100, digits=0)`% through the incorporation of two additional large databases that also synthesized published forest C data: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010; @jian_restructured_2020] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020].
+These data represent ground-based measurements, and *ForC* contains associated data required for interpretation (*e.g.*, stand history, measurement methods).
+Since its most recent publication [*ForC v2.0-Ecology*; @anderson-teixeira_forc_2018], *ForC* has grown `r round((nrow(ForC_measurements) - 17367)/17367*100, digits=0)`%, primarily through the incorporation of two additional large databases that also synthesized published forest C data: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010; @jian_restructured_2020] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020].
 Following these additions, *ForC* currently contains `r nrow(ForC_measurements)` records from `r nrow(ForC_plots)` plots and `r n_areas` distinct geographic areas representing all forested biogeographic and climate zones, making it ideal for assessing how forest C cycling varies across biomes and with respect to stand age.
@@
 Since the early 19th century, it has been recognized that climate plays a dominant role in shaping differences among forests on a global scale [@humboldt_essay_1807; @holdridge_determination_1947].
-Global scale data syntheses have shown that C fluxes including $GPP$, net primary productivity ($NPP$), and soil respiration ($R_{soil}$) decrease with latitude or, correspondingly, increase with mean annual temperature and, to a lesser extent, precipitation [Fig.
+Global scale data syntheses have shown that C fluxes including $GPP$, net primary productivity ($NPP$), and soil respiration ($R_{soil}$) decrease with latitude or, correspondingly, increase with mean annual temperature [Fig.
 `r schematic_fluxes_biome`; e.g., @lieth_primary_1973;  @luyssaert_co2_2007; @hursh_sensitivity_2017; @banbury_morgan_global_nodate].
@@
 On average, aboveground biomass ($B_{ag}$) tends to decrease with latitude, but not as dramatically as fluxes, and with the highest $B_{ag}$ forests in relatively cool, moist temperate regions [@keith_re-evaluation_2009; @smithwick_potential_2002; @hu_mapping_2016].
-In contrast, standing and downed dead wood ($DW_{standing}$ and $DW_{down}$, respectively) and the organic layer ($OL$) tend to accumulate more in colder climates where decomposition is slow relative to $NPP$ [@harmon_ecology_1986; @allen_global_2002].
+In contrast, standing and downed dead wood ($DW_{standing}$ and $DW_{down}$, respectively, summing to $DW_{tot}$) and the organic layer ($OL$) tend to accumulate more in colder climates where decomposition is slow relative to $NPP$ [@harmon_ecology_1986; @allen_global_2002].
 
 Correlative analyses relating C cycle variables to climate and other environmental variables have recently been taken to a new level through use of machine-learning algorithms that relate ground-based C cycle data to global maps of environmental covariates, making it possible to create fine-scale global maps of C cycling [e.g., @warner_spatial_2019; @cook-patton_mapping_2020].
-This approach can be particularly effective when paired with satellite measurements that correlate to C cycle variables of interest; for example, solar-induced chlorophyll fluorescence is useful for fine-scale mapping of gross primary productivity [$GPP$; @li_mapping_2019], while LiDAR, radar, and optical imagery is being used to estimate $B_{ag}$ at regional to global scales [e.g., @saatchi_benchmark_2011; @hu_mapping_2016].
+This approach can be particularly effective when paired with satellite measurements that correlate to C cycle variables of interest; for example, solar-induced chlorophyll fluorescence is useful for fine-scale mapping of gross primary productivity [$GPP$; @li_mapping_2019], while LiDAR, radar, and optical imagery are being used to estimate $B_{ag}$ at regional to global scales [e.g., @saatchi_benchmark_2011; @hu_mapping_2016].
 Any such analysis is however constrained by the quality and coverage of ground-based estimates of forest C fluxes or stocks [e.g., @schepaschenko_forest_2019]. 
-While estimates of some variables (e.g., $B_{ag}$, $GPP$, $NPP$, $R_{soil}$) are widely available, many remain poorly characterized (e.g., $DW$; autotrophic respiration, $R_{auto}$) --even at the coarse resolution of biomes.
+While estimates of some variables (e.g., $B_{ag}$, $GPP$, $NPP$, $R_{soil}$) are widely available, many remain poorly characterized (e.g., $DW_{tot}$; $OL$; autotrophic respiration, $R_{auto}$) --even at the coarse resolution of biomes.
 This is a critical limitation not only for understanding forest C cycling, but also for quantifying forest-based climate change mitigation across forest biomes or ecozones [e.g., @ipcc_2019_2019].
@@
 
-Stand age is another important axis of variation in forest C cycling (Fig. `r schematic_fs_age`).
+Stand age is another important axis of variation in forest C cycling (Fig. `r schematic`).
 In 1969, E.P. Odum's "The Strategy of Ecosystem Development" laid out predictions as to how forest energy flows and organic matter stocks vary with stand age  [@odum_strategy_1969].
-Although the conceptualization of the C cycle in this paper is simplistic by current standards, the paper was foundational in framing the theory around which research on the subject still revolves [@corman_foundations_2019], and the basic framework still holds, albeit with modest modifications [@andersonteixeira_altered_2013].
-Following stand-clearing disturbance, $GPP$, $NPP$, and biomass of leaves ($B_{foliage}$) and fine roots ($B_{root-fine}$) increase rapidly and thereafter remain relatively stable ($B_{foliage}$,$B_{root-fine}$, sometimes $GPP$) or decline slightly [$NPP$, sometimes $GPP$; e.g., @law_changes_2003; @pregitzer_carbon_2004; @amiro_ecosystem_2010; @goulden_patterns_2011].
+Although the conceptualization of the C cycle in this paper was simplistic by current standards, the paper was foundational in framing the theory around which research on the subject still revolves [@corman_foundations_2019], and the basic framework still holds, albeit with modest modifications [Fig.
+`r schematic`; @andersonteixeira_altered_2013].
+Following stand-clearing disturbance, $GPP$, $NPP$, and biomass of leaves ($B_{foliage}$) and fine roots ($B_{root-fine}$) increase rapidly and thereafter remain relatively stable ($B_{foliage}$, $B_{root-fine}$, sometimes $GPP$) or decline slightly [$NPP$, sometimes $GPP$; e.g., @law_changes_2003; @pregitzer_carbon_2004; @amiro_ecosystem_2010; @goulden_patterns_2011].
 The decline in $NPP$ occurs because $R_{auto}$ increases relative to $GPP$ as forests age, corresponding to declining carbon use efficiency with stand age [@delucia_forest_2007; @collalti_forest_2020].
@@
 
-In the past several decades, researchers have started asking how age trends--mostly in $B_{ag}$ or total biomass ($B_{tot}$) accumulation-- vary across biomes.
+In the past few decades, researchers have started asking how age trends--mostly in $B_{ag}$ or total biomass ($B_{tot}$) accumulation-- vary across biomes.
 Early research on this theme showed that biomass accumulation rates during secondary succession increase with temperature on a global scale [@johnson_post-disturbance_2000; @anderson_temperature-dependence_2006] and with water availability in the neotropics [@poorter_biomass_2016].
@@
 Second, we imported (via R script) the Global Reforestation Opportunity Assessment database [*GROA* v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020], which itself had drawn on an earlier version of *ForC*.
-Because all records in *GROA* were checked against original publications, these records were given priority over duplicates in *ForC*.
-Third, we manually incorporated records of annual $NEP$, $GPP$, and $R_{eco}$ from the FLUXNET2015 dataset [@pastorello_fluxnet2015_2020], treating these records as authoritative when they duplicated earlier records (Appendix S1).
+Because all records in *GROA* were checked against original publications, these records were given priority over duplicates in *ForC* (Appendix S1).
+Third, we incorporated records of annual $NEP$, $GPP$, and $R_{eco}$ from the FLUXNET2015 dataset [@pastorello_fluxnet2015_2020], treating these records as authoritative when they duplicated earlier records (Appendix S1).
 We have also added data from individual publications, with a particular focus on productivity [e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @lutz_global_2018; @johnson_climate_2018].
@@
 We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods.
-We removed all plots for which no stand history information had been retrieved (`r round(n_no_history/ForC_simplified_duplicates_removed*100,digits=1)`%).
+We removed all plots for which no stand history information had been retrieved (`r round(n_no_history/ForC_simplified_duplicates_removed*100,digits=1)`% of duplicate-purged records).
 In total, this resulted in `r records_eligible_for_inclusion` records (`r round(records_eligible_for_inclusion/nrow(ForC_measurements)*100,digits=1)`% of the records in the database) being eligible for inclusion in the analysis.
@@
 We also excluded records of stand age = 0 year (`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
-In total, our analysis retained `r round(n_records/n_focal_variable_records_eligible_for_inclusion*100, digits=1)` of the focal variable records for forests of known age.
+In total, our analysis retained `n_records` records. 
 Numbers of records by biome and age class are given in Table S1.
@@
 
-To test for age trends in young (<100yrs) forests, we employed a mixed effects model with biome and log10[stand.age] as fixed effects and plot nested within geographic.area as a random effect on the intercept.
+To test for age trends in young (<100yrs) forests, we employed a mixed effects model with biome and log10[stand age] as fixed effects and plot nested within geographic area as a random effect on the intercept.
 This analysis was run for variables with records for at least three distinct geographic areas in more than one biome, excluding any biomes that failed this criteria (Table 1). 
@@
 
-To facilitate the accessibility of our results and data, and to allow for rapid updates as additional data become available, we have automated all database manipulation, analyses, and figure production in R [@r_core_team_r_2020].
+To facilitate the accessibility of our results and data, and to allow for rapid updates as additional data become available, we automated all database manipulation, analyses, and figure production in R [@r_core_team_r_2020].
 
@@
 These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas.
-Of the 23 flux and 11 stock variables mapped in these diagrams, *ForC* contained sufficient mature forest data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ 7 distinct geographic areas) for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests.
+Of the 23 flux and 11 stock variables mapped in our C cycle diagrams (Figs.
+`r all_diagrams_mature`, `r all_diagrams_young`), *ForC* contained sufficient mature forest data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ 7 distinct geographic areas) for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests.
 For regrowth forests (<100 yrs), *ForC* contained sufficient data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ 3 distinct geographic areas) for `r n_fluxes_with_ge3_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge3_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.
@@
 
-Average C cycles for mature tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major natural or anthropogenic disturbance are presented in Figures 2-5 (and available in tabular format in the *ForC* release accompanying this publication:  `ForC/numbers_and_facts/ForC_variable_averages_per_Biome.csv`).
+Average C cycles for mature tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major natural or anthropogenic disturbance are presented in Figures `r all_diagrams_mature` (and available in tabular format in the *ForC* release accompanying this publication:  `ForC/numbers_and_facts/ForC_variable_averages_per_Biome.csv`).
 
@@
 In the temperate conifer biome, the average composite measure of root biomass ($B_{root}$) was less than the combined average value of coarse and fine root biomass ($B_{root-coarse}$ and $B_{root-fine}$, respectively).
-This lack of closure was driven by very high estimates of $B_{root-coarse}$ from high-biomass forests of the US Pacific Northwest.
+This lack of closure was driven by very high estimates of $B_{root-coarse}$ from high-biomass forests of the US Pacific Northwest (Fig. `r B_root_coarse_fig`).
 

### 2021-01-25 manuscript/response_to_reviews.Rmd
update letter

@@
 - We changed the figure set in accordance with comments from R3.
-Specifically, we (1) added a schematic figure, somewhat in the style of @odum_strategy_1969, to help with conceptual framing of the review; (2) added a figure synthesizing changes in fluxes through time for each biome; (3) removed maps from age trend diagrams (maps still available in SI), which makes the figures less busy/ easier to read; and (4) moved C cycle diagrams for young forests to SI;
+Specifically, we (1) added a schematic figure, somewhat in the style of @odum_strategy_1969, to help with conceptual framing of the review; (2) added a figure synthesizing changes in fluxes through time for each biome; (3) removed maps from age trend diagrams (maps still available in supplementary information), which makes the figures less busy/ easier to read; and (4) moved C cycle diagrams for young forests to the supplementary information.
 

### 2021-01-25 manuscript/ERL_global_C_SI.Rmd
change title to match online submission

@@
 ---
-title: "Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database"
+title: "Supplement for: Carbon cycling in mature and regrowth forests globally"
 output: 

### 2021-01-25 manuscript/ERL_global_C_review.Rmd
change title to match online submission

@@
 
-**Title:** Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the Global Forest Carbon (ForC) database
+**Title:** Carbon cycling in mature and regrowth forests globally
 

### 2021-01-25 manuscript/response_to_reviews.Rmd
small update

@@
 **We agree that such an analysis would be valuable, but time and resource constraints place it beyond the scope of what is currently feasible in this already very large analysis.
-We believe that the current consistency check is valuable as a first pass for identifying cases of extreme mis-matches, and therefore retain it as is.
+We also note that the focus is now shifted towards the ecological questions (trends across biomes and with age), dropping as a main focus the original question on database representativeness and C cycle closure.**
+
+**We believe that the current consistency check is valuable as a first pass for identifying cases of extreme mis-matches, and therefore retain it as is.
 However, we agree that it is not the ideal test, and that it is not very stringent (given high standard deviations) and have therefore toned down claims that the budgets are "closed"/"internally consistent" throughout the manuscript (e.g., removed from abstract).

### 2021-01-25 manuscript/ERL_global_C_review.Rmd
proofreed through p. 16

@@
 The decline in $NPP$ occurs because $R_{auto}$ increases relative to $GPP$ as forests age, corresponding to declining carbon use efficiency with stand age [@delucia_forest_2007; @collalti_forest_2020].
-Heterotrophic respiration, most of which originates from the soil ($R_{het-soil}$) remains relatively constant with stand age [@law_changes_2003; @pregitzer_carbon_2004; @goulden_patterns_2011], with the result that net ecosystem production ($NEP=GPP-R_{eco}$, where $R_{eco}$ is total ecosystem respiration) is initially negative, increases to a maximum at intermediate ages, and declines thereafter [@law_changes_2003; @pregitzer_carbon_2004; @amiro_ecosystem_2010; @goulden_patterns_2011; @luyssaert_old-growth_2008].
+Heterotrophic respiration, most of which originates from the soil ($R_{het-soil}$) remains relatively constant with stand age [@law_changes_2003; @pregitzer_carbon_2004; @goulden_patterns_2011], with the result that net ecosystem production ($NEP=GPP-R_{eco}$, where $R_{eco}$ is total ecosystem respiration) is initially negative, increases to a maximum at intermediate ages, and declines--typically to a small positive value--thereafter [@law_changes_2003; @pregitzer_carbon_2004; @amiro_ecosystem_2010; @goulden_patterns_2011; @luyssaert_old-growth_2008].
 The result is that biomass accumulates rapidly in young forests, followed by a slow decline to near zero in old forests [e.g., @lichstein_biomass_2009; @yang_carbon_2011].
@@
 `r ANPP_litterfall_fig` and `r BNPP_fine_fig`, respectively), the same general trends applied but were not statistically significant.
-Another exception was for $BNPP_{root-coarse}$, where all records came from high-biomass forests in the US Pacific Northwest, resulting in marginally higher values for the temperate conifer biome (Table 1, Fig.
+Another exception was for $BNPP_{root-coarse}$, where all records came from high-biomass forests in the US Pacific Northwest, resulting in marginally higher values for the temperate conifer biome (Fig.
 `r BNPP_coarse_fig`; differences significant in mixed effects model but not in post-hoc pairwise comparison).
@@
 `r flux_age_trends`, `r NEP_fig`).
+For all biomes, $NEP$ was positive, with 95% confidence intervals excluding zero.
 

### 2021-01-25 manuscript/ERL_global_C_SI.Rmd
cleanup

@@
 If duplicates remained (within *ForC v2.0* or *SRDB*), we selected a record for retention based on citation year, keeping the oldest (presumably original) citations for most variables. 
-For eddy-covariance variables ($NEP$, $GPP$, $R_{eco}$), we retained the record associated with the most recent publication [most often @pastorello_fluxnet2015_2020], as these data are commonly re-assessed using new analysis methods.
+For eddy-covariance variables ($NEP$, $GPP$, $R_{eco}$), we retained the record associated with the most recent publication (most often FLUXNET2015), as these data are commonly re-assessed using new analysis methods.
 If publication years were the same, we randomly selected one record.
@@
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
-To illustrate the magnitude of different fluxes, arrow width is proportional to the square root of the mean of the corresponding flux.", out.width='87%', fig.pos='H'}
+To illustrate the magnitude of different fluxes, arrow width is proportional to the square root of the mean of the corresponding flux.", out.width='84%', fig.pos='H'}
 
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $NEP$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $GPP$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $NPP$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $ANPP$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $ANPP_{woody}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $ANPP_{stem}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $ANPP_{foliage}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $ANPP_{litterfall}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $BNPP$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $BNPP_{coarse}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $BNPP_{fine}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $R_{eco}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $R_{root}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $R_{soil}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $R_{het-soil}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $B_{tot}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $B_{ag}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $B_{ag-wood}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $B_{foliage}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $B_{root}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $B_{root-coarse}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $B_{root-fine}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $DW_{tot}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $DW_{standing}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $DW_{down}$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
@@
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Age trends and biome differences for $OL$.
-Map shows data sources ($x$ and $o$ indicate young and mature stands, respectively).
+Map shows data sources ('$\\times$' and '$\\circ$' indicate young and mature stands, respectively).
 In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
+The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\\times$ biome interaction.
 Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.

### 2021-01-25 manuscript/ERL_global_C_review.Rmd
minor clean-up: results

@@
 Individual figures for each flux with sufficient data, along with maps showing geographic distribution of the data, are given in the Supplement (Figs.
-S4-S19).", out.width='100%', fig.pos='H'}
+S5-S19).", out.width='100%', fig.pos='H'}
 
@@
 
+Biome differences were less consistent across C stocks than fluxes (Figs. `r stock_age_trends`, `r all_stock_age_trends`).
 There were sufficient data to assess mature forest biome differences for nine stock variables, and significant differences among biomes were detected for five variables ($B_{tot}$, $B_{ag}$, $B_{ag-wood}$, $B_{foliage}$, $B_{root-coarse}$; Table 1).
-C stocks had less consistent patterns across biomes (Figs. `r stock_age_trends`, `r all_stock_age_trends`).
-For $B_{tot}$ and $B_{ag}$, tropical broadleaf forests had the highest mean biomass and boreal forests the lowest, with temperate broadleaf and needleleaf ($B_{ag}$ only) intermediate.
-However, maximum values for stocks including live or standing woody biomass ($B_{tot}$, $B_{ag}$, $B_{ag-wood}$, $B_{root}$, $B_{root_coarse}$, $DW_{tot}$, $DW_{standing}$) consistently occurred in temperate biomes (Figs.
+For $B_{tot}$ and $B_{ag}$, tropical broadleaf forests had the highest mean biomass and boreal forests the lowest, with intermediate means for temperate broadleaf and needleleaf forests (temperate needleleaf excluded from $B_{tot}$ analysis becuase of insufficient data; Figs.
+`r B_tot_fig`, `r B_ag_fig`).
+However, maximum values for these variables -- along with all other  stocks including live or standing woody biomass ($B_{ag-wood}$, $B_{root}$, $B_{root-coarse}$, $DW_{tot}$, $DW_{standing}$) -- consistently occurred in temperate biomes (Figs.
 `r schematic`, `r stock_age_trends`, `r all_stock_age_trends`).
@@
 Of these, ten increased significantly with $log10[age]$: $NEP$, $GPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{foliage}$, $ANPP_{litterfal}$, $BNPP$, $BNPP_{fine}$, $R_{eco}$, and $R_{root}$.
-The remaining six--$NPP$, $ANPP_{stem}$, $ANPP_{branch}$, $BNPP_{coarse}$, $R_{soil}$, and $R_{het-soil}$-–displayed no significant relationship to stand age.
+The remaining six -- $NPP$, $ANPP_{stem}$, $ANPP_{branch}$, $BNPP_{coarse}$, $R_{soil}$, and $R_{het-soil}$ -- displayed no significant relationship to stand age.
+
+Differences in C fluxes across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests (Table 1, Figs.
+`r flux_age_trends`, `r all_age_trends`).
+The single exception was $ANPP_{stem}$, for which temperate broadleaf and conifer forests had flux rates similar to tropical forests.
+Notably, and in contrast to the lack of biome differences in $NEP$ for mature forests (Fig.
+`r flux_age_trends`), the tendency for temperate forests to have greater fluxes than boreal forests held for $NEP$ in regrowth forests (tropical forests excluded because of insufficient data).
 
@@
 
-Differences in C fluxes across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests.
-The single exception was $ANPP_{stem}$, for which temperate broadleaf and conifer forests had flux rates similar to tropical forests.
-Notably, and in contrast to the lack of biome differences in $NEP$ for mature forests (Fig.
-`r flux_age_trends`), the tendency for temperate forests to have greater fluxes than boreal forests held for $NEP$ in regrowth forests (tropical forests excluded because of insufficient data).
-
 "Closure" and internal consistency of the C flux budget was less successful for young than mature forests (Figs. `r age_trends_all`).
 Summed regression equations for $R_{soil-het}$ and $R_{root}$ were generally very close to $R_{soil}$.
-We calculated $R_{auto-ag}$ as the difference between $R_{eco}$ and $R_{soil}$ (except for tropical forests, which had insufficient $R_{eco}$ data), effectively guaranteeing near-closure of the CO~2~ efflux (respiration) portion of the budget (negative values in Figs.
+In assessing the C budget of young forests, we calculated $R_{auto-ag}$ as the difference between $R_{eco}$ and $R_{soil}$ (except for tropical forests, which had insufficient $R_{eco}$ data), effectively guaranteeing near-closure of the CO~2~ efflux (respiration) portion of the budget (negative values in Figs.
 `r age_trends_all`).
@@
 Age $\times$ biome interactions were also significant for all ten of these C stock variables (Table S2), with living C stocks tending to accumulate more rapidly during the early stages of forest regrowth in tropical forests (Figs.
-`r stock_age_trends`, `r all_stock_age_trends`).
+`r stock_age_trends`, `r age_trends_all`, `r all_stock_age_trends`).
 In the case of two non-living C stocks ($DW_{down}$ and $OL$), age $\times$ biome interactions were such that age trends were positive in some biomes and negative in others.
-Specifically, $DW_{down}$ declined with age in temperate and boreal forests, compared to an increase with age in tropical forests (Figs. `r stock_age_trends`, `r DW_down_fig`).
-Similarly, $OL$ declined slightly with age in temperate broadleaf forests, contrasting an increase in the other three biomes (Figs. `r stock_age_trends`, `r OL_fig`).
-Again, there were some discrepencies between young forest trends and mature forests, most notably including generally higher C stocks in mature forests relative to their 100-year counterparts, particularly for temperate conifer forests (*again, likely a geographic representation issue?*) and, to a lesser extent, tropical broadleaf forests.
+Specifically, $DW_{down}$ declined with age in temperate and boreal forests, compared to an increase with age in tropical forests (Figs. `r stock_age_trends`,`r age_trends_all`, `r DW_down_fig`).
+Similarly, $OL$ declined slightly with age in temperate broadleaf forests, contrasting an increase in the other three biomes (Figs. `r stock_age_trends`, `r age_trends_all`,`r OL_fig`).
+Again, there were some discrepancies between young forest trends and mature forests, most notably including generally higher C stocks in mature forests relative to their 100-year counterparts, particularly for temperate conifer forests (with discrepancies again driven by differences in geographic representation) and, to a lesser extent, tropical broadleaf forests (`r age_trends_all`).
 

### 2021-01-25 manuscript/ERL_global_C_review.Rmd
fix-up through line 548

@@
 Similarly, $OL$ declined slightly with age in temperate broadleaf forests, contrasting an increase in the other three biomes (Figs. `r stock_age_trends`, `r age_trends_all`,`r OL_fig`).
-Again, there were some discrepancies between young forest trends and mature forests, most notably including generally higher C stocks in mature forests relative to their 100-year counterparts, particularly for temperate conifer forests (with discrepancies again driven by differences in geographic representation) and, to a lesser extent, tropical broadleaf forests (`r age_trends_all`).
+Again, there were some discrepancies between young forest trends and mature forests, most notably including generally higher C stocks in mature forests relative to their 100-year counterparts, particularly for temperate conifer forests (with discrepancies again driven by differences in geographic representation) and, to a lesser extent, tropical broadleaf forests (FIg.
+`r age_trends_all`).
 
@@
 *ForC* v3.0 provided unprecedented coverage of most major variables, yielding a broad picture of C cycling in the world's major forest biomes.
-Carbon cycling rates generally increased from boreal to tropical regions and with stand age (Figs. `r schematic`, `r all_age_trends`).
+Carbon cycling rates generally increased from boreal to tropical regions and with stand age (Figs. `r schematic`, `r age_trends_all`).
 Specifically, most C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for regrowth as well as mature forests (Figs.
@@
 Our analysis reveals that carbon cycling is most rapid in the tropics and slowest in boreal regions, including C fluxes into ($GPP$), within (e.g., $NPP$ and its components), and out of (e.g., $R_{soil}$, $R_{eco}$) the ecosystem.
-For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude (or increase with temperature) on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
-This consistency is not not surprising, particularly given commonality in the data analyzed or used for calibration.
-The finding that these patterns hold consistently across across numerous fluxes, while unsurprising, is novel to this analysis [but see @banbury_morgan_global_nodate for nine autotrophic fluxes].
+For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude -- or increase with temperature -- on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
+This consistency is not surprising, particularly given commonality in the data analyzed or used for calibration.
+The finding that these patterns hold consistently across across numerous fluxes, while consistent with theoretical expectations (Fig.
+`r schematic`), is novel to this analysis [but see @banbury_morgan_global_nodate for nine autotrophic fluxes].
 
@@
 `r flux_age_trends`f).
-Unlike the other C flux variables, $NEP$ does not characterize the rate at which C cycles through the ecosystem, but is the balance between C sequestration ($GPP$) and respiratory losses ($R_{eco}$) and represents net CO~2~ sequestration (or release) by the ecosystem (Fig.
+Unlike the other C flux variables, $NEP$ does not characterize the rate at which C cycles through the ecosystem, but, as the balance between $GPP$ and $R_{eco}$, represents net CO~2~ sequestration (or release) by the ecosystem (Fig.
 `r schematic`).
-$NEP$ tends to be relatively small in mature forest stands, which accumulate carbon slowly relative to younger stands (discussed further below), if at all [Fig.
+$NEP$ tends to be relatively small in mature forest stands, which accumulate carbon slowly relative to younger stands, if at all [Fig.
 `r schematic`; @luyssaert_old-growth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
-It is therefore unsurprising that there are no pronounced differences across biomes [in agreement with @luyssaert_co2_2007], suggesting that variation in $NEP$ of mature forests is controlled less by climate and more by other factors including moderate disturbances [@curtis_forest_2018] or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition; @wilson_stability_2016].
+It is therefore consistent with theory -- and with previous research [@luyssaert_co2_2007] -- that there are no pronounced differences across biomes. 
+Rather, variation in $NEP$ of mature forests appears to be controlled less by climate and more by other factors including moderate disturbances [@curtis_forest_2018] or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition; @wilson_stability_2016].
 The fact that mature temperate broadleaf forests have a higher mean than the other biomes may reflect the fact that most of these forests are older secondary forests that, while classified here as mature, are still accumulating carbon [@curtis_forest_2018].
@@
 `r flux_age_trends`, `r NEP_fig`).
-This is consistent with findings that live biomass accumulation rates (*e.g.*, $\Delta B_{ag}$ or $\Delta B_{tot}$) during early secondary succession decrease with latitude [Figs.
-7a, S16-S22; @anderson_temperature-dependence_2006; @cook-patton_mapping_2020].
-Note, though, that $NEP$ includes not only $\Delta B_{tot}$, but also changes in $DW_{tot}$, $OL$, and soil carbon (not analyzed here), and biome differences in the accumulation rates of these variables have not been detected, in part because these variables do not consistently increase with stand age [Figs.
+This is consistent with findings that live biomass accumulation rates ($\Delta B_{ag}$ or $\Delta B_{tot}$) during early secondary succession decrease with latitude [Figs.
+`r stock_age_trends`a, `r all_stock_age_trends`; @anderson_temperature-dependence_2006; @cook-patton_mapping_2020].
+Note, though, that $NEP$ includes not only $\Delta B_{tot}$, but also changes in $DW_{tot}$, $OL$, and soil carbon (not analyzed here). 
+Biome differences in the accumulation rates of $DW$, $OL$, and soil C  have not been detected, in part because these variables do not consistently increase with stand age [Figs.
 `r schematic`, `r stock_age_trends`, `r DW_tot_fig`-`r OL_fig`, and see discussion below; @cook-patton_mapping_2020].
 
-For regrowth forests, little is known about cross-biome differences in carbon fluxes, and we are not aware of any previous large-scale comparisons of C fluxes that have been limited to regrowth forests.
+For regrowth forests, little was previously known about cross-biome differences in carbon fluxes, and we are not aware of any previous large-scale comparisons of C fluxes that have been limited to regrowth forests.
 Thus, this analysis was the first to examine flux trends in regrowth forests across biomes.
 The observed tendency for young forest fluxes to decrease from tropical to boreal regions paralleled patterns in mature forests (Figs.
-`r flux_age_trends`, `r all_flux_age_trends`), suggesting that regrowth forests follow latitudinal trends in carbon cycling similar to those of mature forests [e.g., @banbury_morgan_global_nodate].
+`r flux_age_trends`, `r age_trends_all`, `r all_flux_age_trends`), suggesting that regrowth forests follow latitudinal trends in carbon cycling similar to those of mature forests [e.g., @banbury_morgan_global_nodate].
 

### 2021-01-25 manuscript/ERL_global_C_review.Rmd
refining discussion (up to "C variable coverage and budget closure")

@@
 
-In contrast to C fluxes and biomass accumulation rates in regrowth forests, stocks showed less systematic variation across biomes.
+In contrast to C fluxes and biomass accumulation rates in regrowth forests, stocks showed less systematic variation across biomes (c.f. Fig. `r schematic`).
 For aboveground biomass, which is the variable in *ForC* with broadest geographical representation, the modest trend of declining biomass from tropical to boreal regions mirrors observations from spaceborne lidar that reveal a decline in aboveground biomass (for all forests, including secondary) with latitude across the N hemisphere [@hu_mapping_2016].
@@
 `r stock_age_trends`, `r DW_tot_fig`-`r DW_down_fig`; but see below for age trends), pointing to a need for more widespread quantification of both standing and downed deadwood.
-*ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that proportionally slower decomposition in colder climates results in more buildup of organic matter [@allen_global_2002; @andersonteixeira_differential_2011].
+*ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that proportionally slower decomposition in colder climates results in more buildup of organic matter [Fig.
+`r schematic`; @allen_global_2002; @andersonteixeira_differential_2011].
 Further research on non-living C stocks in the world's forests will be essential to completing the picture.
@@
 While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_{auto}$) would be minimal immediately following a stand-clearing disturbance.
-These would be expected to increase rapidly with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Fig. `r stock_age_trends`).
-In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$) are expected to be non-zero following stand-clearing disturbance, although these may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots and litter [@ribeiro-kumara_how_2020; @maurer_carbon_2016; @bond-lamberty_contribution_2004].
-In this study, we detect no significant age trends in either variable.
+These would be expected to increase rapidly with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Figs.
+`r schematic`, `r stock_age_trends`).
+In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$)--and therefore $R_{eco}$ are expected to be non-zero following stand-clearing disturbance (Fig.
+`r schematic`), although these may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots and litter [@ribeiro-kumara_how_2020; @maurer_carbon_2016; @bond-lamberty_contribution_2004].
+In this study, we detect no significant age trends in either $R_{het-soil}$ or $R_{soil}$, consistent with previous findings [@law_changes_2003;@pregitzer_carbon_2004;@goulden_patterns_2011].
 
@@
 This finding is largely consistent with, but built from a far larger dataset than, previous studies showing an increase in $NEP$ across relatively young stand ages [@pregitzer_carbon_2004; @baldocchi_fluxnet_2001; @luyssaert_old-growth_2008].
-However, $NEP$ has been observed to decline from intermediate to old stands [ @law_changes_2003; @luyssaert_old-growth_2008], and the $NEP$ estimated by our model for 100-year-old temperate conifer stands (~5 Mg C ha^-1^ yr^-1^) exceeds the mean of mature forests in the same biome (0.7 Mg C ha^-1^ yr^-1^; Fig.
-`r diagram_TeN_mature`).
-A decrease in $NEP$ is consistent with the observed deceleration of biomass accumulation as stands age, although both biomass and non-living C stocks will often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands [@luyssaert_old-growth_2008; @mcgarvey_carbon_2014; @lichstein_biomass_2009].
+However, $NEP$ has been observed to decline from intermediate to old stands [@law_changes_2003; @luyssaert_old-growth_2008], whereas the $NEP$ values estimated by our models for 100-year-old stands were not systematically different from those of mature stands (lower for temperate broadleaf, higher for temperate conifer, and equal for boreal; Fig.
+`r age_trends_all`).
+The lack of a consistent age-related decrease may be driven by differences in geographical representation across age classes or by the fitting of an inappropriate functional form.
+A decrease in $NEP$ would be consistent with the observed deceleration of C accumulation as stands age (Fig.
+`r age_trends_all`), although both biomass and non-living C stocks will often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands [@luyssaert_old-growth_2008; @mcgarvey_carbon_2014; @lichstein_biomass_2009].
+Additional data, including on age trends of $DW$, $OL$, and SOC will be important to parsing the timing and extend of an age-related $NEP$ decrease across forest biomes.
 
-In terms of stocks, our study reveals consistent increases in live biomass stocks with stand age--a pattern that is well-known and expected [e.g., @lichstein_biomass_2009; @yang_carbon_2011]--and more variable age trends in deadwood and $OL$.
-The latter are particularly sensitive to the type of disturbance, where disturbances that remove most organic material (e.g., logging, agriculture) result in negligible deadwood in young stands, followed by a buildup over time [tropical stands in Fig.
-`r stock_age_trends`; e.g., @vargas_biomass_2008].
-In contrast, natural disturbances (e.g., fire, drought) can produce large amounts of deadwood (mostly $DW_{standing}$) that slowly decomposes as the stand recovers, resulting in declines across young stand ages [e.g., temperate and boreal stands in Fig.
-`r stock_age_trends`; e.g., @carmona_coarse_2002].
+In terms of stocks, our study reveals consistent increases in live biomass stocks with stand age--a pattern that is well-known and expected [e.g., @lichstein_biomass_2009; @yang_carbon_2011]--and more variable age trends in deadwood and $OL$ (Fig.
+`r age_trends_all`).
+The latter are particularly sensitive to the type of disturbance, where disturbances that remove most organic material (e.g., logging, agriculture) result in negligible deadwood in young stands, followed by a buildup over time [tropical stands in Figs.
+`r stock_age_trends`, `r age_trends_all`; e.g., @vargas_biomass_2008].
+In contrast, natural disturbances (e.g., fire, drought) can produce large amounts of deadwood (mostly $DW_{standing}$) that slowly decomposes as the stand recovers, resulting in declines across young stand ages [e.g., temperate and boreal stands in Figs.
+`r stock_age_trends`, `r age_trends_all`; e.g., @carmona_coarse_2002].
 Again, further study and synthesis of non-living C stocks across biomes and stand ages will be valuable to giving a more comprehensive picture.

### 2021-01-25 manuscript/ERL_global_C_review.Rmd
final pass on discussion

@@
 
-Our study reveals that most C fluxes quickly increase to a plateau as stands age (Figs.
+Our study reveals that most C fluxes quickly increase and then decelerate as stands age (Figs.
 `r flux_age_trends`, `r age_trends_all`), consistent with current understanding of age trends in forest C cycling [Fig.
 `r schematic`; e.g., @andersonteixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
-While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_{auto}$) would be minimal immediately following a stand-clearing disturbance.
-These would be expected to increase rapidly with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Figs.
-`r schematic`, `r stock_age_trends`).
+While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_{auto}$) would be minimal immediately following a stand-clearing disturbance (Fig.
+`r schematic`).
+These would be expected to increase rapidly, along with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Figs.
+`r schematic`, `r flux_age_trends`-`r age_trends_all`).
 In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$)--and therefore $R_{eco}$ are expected to be non-zero following stand-clearing disturbance (Fig.
@@
 `r flux_age_trends`).
-This finding is largely consistent with, but built from a far larger dataset than, previous studies showing an increase in $NEP$ across relatively young stand ages [@pregitzer_carbon_2004; @baldocchi_fluxnet_2001; @luyssaert_old-growth_2008].
+This finding is consistent with previous studies showing an increase in $NEP$ across relatively young stand ages [@pregitzer_carbon_2004; @baldocchi_fluxnet_2001; @luyssaert_old-growth_2008].
 However, $NEP$ has been observed to decline from intermediate to old stands [@law_changes_2003; @luyssaert_old-growth_2008], whereas the $NEP$ values estimated by our models for 100-year-old stands were not systematically different from those of mature stands (lower for temperate broadleaf, higher for temperate conifer, and equal for boreal; Fig.
 `r age_trends_all`).
-The lack of a consistent age-related decrease may be driven by differences in geographical representation across age classes or by the fitting of an inappropriate functional form.
+This lack of a consistent age-related decrease may be driven by differences in geographical representation across age classes or by the fitting of an inappropriate functional form.
 A decrease in $NEP$ would be consistent with the observed deceleration of C accumulation as stands age (Fig.
 `r age_trends_all`), although both biomass and non-living C stocks will often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands [@luyssaert_old-growth_2008; @mcgarvey_carbon_2014; @lichstein_biomass_2009].
-Additional data, including on age trends of $DW$, $OL$, and SOC will be important to parsing the timing and extend of an age-related $NEP$ decrease across forest biomes.
+Additional data, including on age trends of deadwood, the organic layer, and soil C will be important to parsing the timing and extend of an age-related $NEP$ decrease across forest biomes.
 
-In terms of stocks, our study reveals consistent increases in live biomass stocks with stand age--a pattern that is well-known and expected [e.g., @lichstein_biomass_2009; @yang_carbon_2011]--and more variable age trends in deadwood and $OL$ (Fig.
+In terms of stocks, our study reveals consistent increases in live biomass stocks with stand age, a pattern that is well-known and expected [e.g., @lichstein_biomass_2009; @yang_carbon_2011], contrasting with more variable age trends in deadwood and the organic layer (Fig.
 `r age_trends_all`).
-The latter are particularly sensitive to the type of disturbance, where disturbances that remove most organic material (e.g., logging, agriculture) result in negligible deadwood in young stands, followed by a buildup over time [tropical stands in Figs.
+The latter are particularly sensitive to the type of disturbance, where disturbances that remove most organic material (e.g., logging, agriculture) result in negligible deadwood in young stands, followed by a buildup over time [e.g., tropical stands in Figs.
 `r stock_age_trends`, `r age_trends_all`; e.g., @vargas_biomass_2008].
 In contrast, natural disturbances (e.g., fire, drought) can produce large amounts of deadwood (mostly $DW_{standing}$) that slowly decomposes as the stand recovers, resulting in declines across young stand ages [e.g., temperate and boreal stands in Figs.
-`r stock_age_trends`, `r age_trends_all`; e.g., @carmona_coarse_2002].
+`r stock_age_trends`, `r age_trends_all`; @carmona_coarse_2002].
 Again, further study and synthesis of non-living C stocks across biomes and stand ages will be valuable to giving a more comprehensive picture.
@@
 
-The large number of C cycle variables covered by ForC, and the general consistency among them, provide confidence that our overall reported mature forest means provide useful baselines for analysis -- with the caveats that they are unlikely to be accurate representations of C cycling for any particular forest, and that these sample means almost certainly do not represent true biome means (particularly for temperate conifer forests where high-biomass stands are over-represented in *ForC*).
+The large number of C cycle variables covered by ForC, and the relatively high consistency among them (Figs.
+`r all_diagrams_mature`, `r age_trends_all`), provide confidence that our reported mature forest means provide useful baselines for analysis -- with the caveats that they are unlikely to be accurate representations of C cycling for any particular forest, and that these sample means almost certainly do not represent true biome means (particularly for temperate conifer forests where high-biomass stands are over-represented in *ForC*).
 
-In this analysis, the C cycle budgets for mature forests (Figs.
-`r all_diagrams_mature`) generally “close”–-that is, the sums of component variables do not differ from the larger fluxes by more than one standard deviation.
+In this analysis, the C cycle budgets for mature forests usually come close to closure–-that is, the sums of component variables do not differ from the larger fluxes by more than one standard deviation (Figs.
+`r all_diagrams_mature`, `r age_trends_all`).
 On the one hand, this reflects the general fact that ecosystem-scale measurements tend to close the C budget more easily and consistently than, for example, for energy balance [@stoy_data-driven_2013].
 On the other, however, *ForC* derives data from multiple heterogeneous sources, and standard deviations within each biome are high; as a result, the standard for C closure is relatively loose [*c.f.* @houghton_terrestrial_2020].
-The one instance where the C budgets doesn’t close is likely due to differences in the representation of forest types (*i.e.*, disproportionate representation of US Pacific NW for  $B_{root-coarse}$ relative to  $B_{root}$; Fig.
+The one instance where the C budgets does not close according to our criteria is likely due to differences in the representation of forest types (*i.e.*, disproportionate representation of US Pacific NW for $B_{root-coarse}$ relative to  $B_{root}$; Fig.
 `r diagram_TeN_mature`) rather than issues of methodological accuracy.
@@
 
-In contrast, age trends for young forests generally remain less clearly defined, in large part because their data records remain somewhat sparse for most variables (*i.e.*, have low representation of different geographical regions for any given age).
-While this analysis provides a first analysis of age trends in forest C cycling for multiple variables at a global scale, improved resolution of these trends will require additional data.
+In contrast, age trends for young forests generally remain less clearly defined, in large part because their data records remain relatively sparse (*i.e.*, have low representation of different geographical regions for any given age) for most variables, particularly in the tropics [@anderson-teixeira_carbon_2016].
+While this review provides a first analysis of age trends in forest C cycling for multiple variables at a global scale, improved resolution of these trends will require additional data.
 
 There are of course notable holes in the ForC variable coverage that limit the scope of our inferences here. 
-Notably, *ForC* currently has sparse--if any--coverage of fluxes to herbivores and higher consumers, along with the woody mortality ($M_{woody}$) and dead wood stocks (Table 1, Figs.
+Notably, *ForC* currently has sparse--if any--coverage of fluxes to herbivores and higher consumers, along with the woody mortality ($M_{woody}$) and $DW$ (Table 1, Figs.
 `r DW_tot_fig`-`r DW_down_fig`).
@@
 
-Importantly, *ForC* and the analyses presented here cover the forests that have received research attention, which are not a representative sample of the world's existing forests--geographically or in terms of human impacts [@martin_mapping_2012].
+Importantly, *ForC* and the analyses presented here cover the forests that have received research attention, which are not a representative sample of the world's existing forests, either geographically or in terms of human impacts [@martin_mapping_2012].
 Geographically, all variables are poorly covered in Africa and Siberia (Fig. `r figure_map`), a common problem in the carbon-cycle community [@xu_contribution_2016; @schimel_effect_2015]. 
@@
 Fragmentation and the creation of edges strongly impacts forest C cycling [e.g., @chaplin-kramer_degradation_2015; @remy_strong_2016; @reinmann_edge_2017; @smith_evidence_2019; @reinmann_urbanization_2020;  @ordway_carbon_2020].
-Partial logging and other forms of non- stand clearing anthropogenic disturbance also strongly impact forest C cycling [e.g., @huang_long-term_2010; @piponiot_carbon_2016] but are under-studied [@sist_tropical_2015] and excluded from this analysis.
+Partial logging and other forms of non- stand clearing anthropogenic disturbance also alter forest C cycling [e.g., @huang_long-term_2010; @piponiot_carbon_2016], but are under-studied [@sist_tropical_2015] and excluded from this analysis.
 Fragmented and degraded forests do not fit the idealized conceptual framework around which this review is structured (Fig.
@@
 To ensure that models are giving the right answers for the right reasons [@sulman_multiple_2018], it is important to benchmark against multiple components of the C cycle that are internally consistent with each other [@collier_international_2018; @wang_golum-cnp_2018].
-*ForC*’s tens of thousands of records are readily available in a standardized format, along with all code used in the analyses presented here, and we recommend that researchers use these resources to identify and summarize data specific the analysis at hand.
+*ForC*’s tens of thousands of records are readily available in a standardized format, along with all code used in the analyses presented here, and we recommend that researchers use these resources to identify and summarize data specific to the analysis at hand.
 Integration of *ForC* with models will be valuable to improving the accuracy and reliability of models [@fer_beyond_2021].
@@
 It is also interesting to consider the complementary utility of global-scale but spatially discontinuous databases such as ForC and remote wall-to-wall remote sensing products. 
-The latter provide unparalleled insight into aboveground carbon stocks, but less constraint on belowground stocks or carbon fluxes in general [@bondlamberty_estimating_2016; @anav_spatiotemporal_2015].
+The latter provide insight into aboveground carbon stocks, but less constraint on belowground stocks or carbon fluxes in general [@bondlamberty_estimating_2016; @anav_spatiotemporal_2015]. 
 Combining observational data and remote observations may provide a much more comprehensive and accurate picture of global forest C cycling, particularly when used in formal data assimilation systems [@konings_global_2019; @liu_detecting_2018].
@@
 ### Conclusions
-As climate change accelerates, understanding and managing the carbon dynamics of forests--notably including dynamics and fluxes that cannot be observed by satellites--is critical to forecasting, mitigation, and adaptation.
+As climate change accelerates, understanding and managing the carbon dynamics of forests-- including dynamics and fluxes that cannot be observed by satellites--is critical to forecasting, mitigation, and adaptation.
 The C data in *ForC*, as summarized here, will be valuable to these efforts.

### 2021-01-25 manuscript/ERL_global_C_review.Rmd
clarify data availability statement, add error bars to age trends figure
closes #56

@@
 Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
+Error bars on mature forest flux estimates indicate $\\pm$ 1 standard deviation.
 Asterisks indicate variables whose age trends were calculated based on other variables (* young and mature forests; ** young forests only; *** mature forests only), as follows.
@@
 2020).
-for non-tropical forests: $ANPP_{woody}=min(ANPP_{stem},ANPP_{woody})$, $R_{auto-ag}=R_{eco}-R{soil}$, .
+For non-tropical forests: $ANPP_{woody}=min(ANPP_{stem},ANPP_{woody})$, $R_{auto-ag}=R_{eco}-R{soil}$.
 Note that there remain substantial uncertainties as to the functional form of age trends and discrepencies in closure among related variables."
@@
 ## Data availability statement
+The data that support the findings of this study are openly available.
 Materials required to fully reproduce these analyses, including data, R scripts, and image files, are archived in Zenodo (DOI: TBD].

### 2021-01-25 manuscript/tables_figures/stacked_age_trends/stacked_area_results.m

@@ -227,6 +227,7 @@ fluxes = [in_fluxes; -1*out_fluxes];
 % 1.2.1 define mature fluxes 
 NEP_index=find(strcmp(ForC_variable_averages.variable_diagram,'NEP') + strcmp(ForC_variable_averages.Biome,strcat(biomes(b),' MATURE'))==2);
 NEP_mature=ForC_variable_averages.mean(NEP_index);
+NEP_std=ForC_variable_averages.std(NEP_index);
 
 R_root_index=find(strcmp(ForC_variable_averages.variable_diagram,'R_root') + strcmp(ForC_variable_averages.Biome,strcat(biomes(b),' MATURE'))==2);
 R_root_mature=ForC_variable_averages.mean(R_root_index);
@@ -248,6 +249,7 @@ ANPP_stem_mature=ForC_variable_averages.mean(ANPP_stem_index);
 
 GPP_index=find(strcmp(ForC_variable_averages.variable_diagram,'GPP') + strcmp(ForC_variable_averages.Biome,strcat(biomes(b),' MATURE'))==2);
 GPP_mature=ForC_variable_averages.mean(GPP_index);
+GPP_std=ForC_variable_averages.std(GPP_index);
 
 R_soil_index=find(strcmp(ForC_variable_averages.variable_diagram,'R_soil') + strcmp(ForC_variable_averages.Biome,strcat(biomes(b),' MATURE'))==2);
 R_soil_mature=ForC_variable_averages.mean(R_soil_index);
@@ -257,6 +259,7 @@ R_het_soil_mature=ForC_variable_averages.mean(R_het_soil_index);
 
 R_eco_index=find(strcmp(ForC_variable_averages.variable_diagram,'R_eco') + strcmp(ForC_variable_averages.Biome,strcat(biomes(b),' MATURE'))==2);
 R_eco_mature=ForC_variable_averages.mean(R_eco_index);
+R_eco_std=ForC_variable_averages.std(R_eco_index);
 
 % 1.2.2 calculations
 R_auto_ag_mature_calc=R_eco_mature-R_soil_mature;
@@ -378,8 +381,11 @@ if b~= 1
     plot(age, NEP, '-y', 'LineWidth', 2,'HandleVisibility','off');hold on; % eddy flux: insufficient data for tropics
 end
[…]

### 2021-01-25 manuscript/ERL_global_C_review.Rmd
cleanup, knitting for submission, create submission files

@@
 schematic =1
-schematic_fluxes_biome="1a"
-  schematic_fluxes_age="1b"
-  schematic_stocks_biome="1c"
-  schematic_stocks_age="1d"
-  schematic_fs_age = "1b,d"
 figure_map=2
@@
 Global scale data syntheses have shown that C fluxes including $GPP$, net primary productivity ($NPP$), and soil respiration ($R_{soil}$) decrease with latitude or, correspondingly, increase with mean annual temperature [Fig.
-`r schematic_fluxes_biome`; e.g., @lieth_primary_1973;  @luyssaert_co2_2007; @hursh_sensitivity_2017; @banbury_morgan_global_nodate].
-C stocks of mature forests show less directional variation (Fig. `r schematic_fluxes_biome`).
+`r schematic`; e.g., @lieth_primary_1973;  @luyssaert_co2_2007; @hursh_sensitivity_2017; @banbury_morgan_global_nodate].
+C stocks of mature forests show less directional variation (Fig. `r schematic`).
 On average, aboveground biomass ($B_{ag}$) tends to decrease with latitude, but not as dramatically as fluxes, and with the highest $B_{ag}$ forests in relatively cool, moist temperate regions [@keith_re-evaluation_2009; @smithwick_potential_2002; @hu_mapping_2016].
@@
 ```
----
-nocite: | 
-@jung_exploiting_2006
----
 

### 2021-01-25 manuscript/ERL_global_C_review.Rmd
produce doc with highlighted changes, fix typos in submission version and .Rmd
#67

@@
 Specifically, we use `r n_records` *ForC* records for `r nrow(T1)` C cycle variables from `r n_areas_tot` geographic locations to characterize ensemble C budgets for four broad forest types -- tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and boreal.
-We calculaate statistics for both mature and regrowth (age <100 years) forests, and quantify trends with stand age in regrowth forests for all variables with sufficient data.
+We calculate statistics for both mature and regrowth (age <100 years) forests, and quantify trends with stand age in regrowth forests for all variables with sufficient data.
 
@@
 We also excluded records of stand age = 0 year (`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
-In total, our analysis retained `n_records` records. 
+In total, our analysis retained `r n_records` records. 
 Numbers of records by biome and age class are given in Table S1.
 
-Data were summarized to produce schematics of C cycling for mature forets of each biome.
+Data were summarized to produce schematics of C cycling for mature forests of each biome.
 To obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot.
@@
 The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
-The boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes.
+The boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
 Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).

### 2021-02-24 manuscript/ERL_global_C_SI.Rmd
knit in new age trend figures
#69

@@
 
-#Age diagrams
+#Age diagrams - for these, need to rename organic.layer and woody.mortality, replacing period with underscore.
 #do.call(file.remove, list(list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ERL-review/manuscript/tables_figures/age_trends/", full.names = TRUE))) # first remove old files
@@
 
-  knitr::include_graphics("tables_figures/age_trends/NEP.png")
+  knitr::include_graphics("tables_figures/age_trends/NEP_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/GPP.png")
+  knitr::include_graphics("tables_figures/age_trends/GPP_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/NPP.png")
+  knitr::include_graphics("tables_figures/age_trends/NPP_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/ANPP.png")
+  knitr::include_graphics("tables_figures/age_trends/ANPP_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/ANPP_woody.png")
+  knitr::include_graphics("tables_figures/age_trends/ANPP_woody_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/ANPP_stem.png")
+  knitr::include_graphics("tables_figures/age_trends/ANPP_stem_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/ANPP_foliage.png")
+  knitr::include_graphics("tables_figures/age_trends/ANPP_foliage_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/ANPP_litterfall.png")
+  knitr::include_graphics("tables_figures/age_trends/ANPP_litterfall_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/BNPP.png")
+  knitr::include_graphics("tables_figures/age_trends/BNPP_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/BNPP_coarse.png")
+  knitr::include_graphics("tables_figures/age_trends/BNPP_coarse_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/BNPP_fine.png")
+  knitr::include_graphics("tables_figures/age_trends/BNPP_fine_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/R_eco.png")
+  knitr::include_graphics("tables_figures/age_trends/R_eco_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/R_root.png")
+  knitr::include_graphics("tables_figures/age_trends/R_root_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/R_soil.png")
+  knitr::include_graphics("tables_figures/age_trends/R_soil_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/R_het_soil.png")
+  knitr::include_graphics("tables_figures/age_trends/R_het_soil_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/biomass.png")
+  knitr::include_graphics("tables_figures/age_trends/biomass_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/biomass_ag.png")
+  knitr::include_graphics("tables_figures/age_trends/biomass_ag_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/biomass_ag_woody.png")
+  knitr::include_graphics("tables_figures/age_trends/biomass_ag_woody_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/biomass_foliage.png")
+  knitr::include_graphics("tables_figures/age_trends/biomass_foliage_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/biomass_root.png")
+  knitr::include_graphics("tables_figures/age_trends/biomass_root_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/biomass_root_coarse.png")
+  knitr::include_graphics("tables_figures/age_trends/biomass_root_coarse_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/biomass_root_fine.png")
+  knitr::include_graphics("tables_figures/age_trends/biomass_root_fine_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/deadwood.png")
+  knitr::include_graphics("tables_figures/age_trends/deadwood_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/deadwood_standing.png")
+  knitr::include_graphics("tables_figures/age_trends/deadwood_standing_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/deadwood_down.png")
+  knitr::include_graphics("tables_figures/age_trends/deadwood_down_with_map.png")
 ```
@@
 
-  knitr::include_graphics("tables_figures/age_trends/organic_layer.png")
+  knitr::include_graphics("tables_figures/age_trends/organic_layer_with_map.png")
 ```

### 2021-02-24 manuscript/response_to_reviews.Rmd
address all reviewer concerns

--- 2021-01-25 manuscript/response_to_reviews.Rmd
@@
 We have addressed all points raised by the reviewers, as detailed below.
-The most substantive changes included the following:
-
-- We changed the figure set in accordance with comments from R3.
-Specifically, we (1) added a schematic figure, somewhat in the style of @odum_strategy_1969, to help with conceptual framing of the review; (2) added a figure synthesizing changes in fluxes through time for each biome; (3) removed maps from age trend diagrams (maps still available in SI), which makes the figures less busy/ easier to read; and (4) moved C cycle diagrams for young forests to SI;
-
-- We improved conceptual framework and "review" nature of the work by adding background information to the introduction, including the schematic figure mentioned above.
-We also shifted emphasis more towards the ecological questions (trends across biomes and with age), while retaining (with less emphasis) the analysis of database representativeness.
 
@@
 
-# Referee: 2
+### Referee: 4
 
 COMMENTS TO THE AUTHOR(S)
+I am satisfied with the revised manuscript and feel it is now acceptable for publication.
 
-The manuscript provides an important contribution to carbon cycle research.
-It provides an update to a previously published data base of forest carbon stocks and fluxes (ForC) and analyses the dataset across forests biomes and in relation to stand age.
-This is a valuable dataset for assessing the current state of the carbon cycle and comparing across biomes.
-The dataset presents carbon cycle modelers with an opportunity to rigorously compare their models with observations.
-Figures 2-5, 8-11 are excellent summaries of the dataset.
-Yet I have some concerns and clarifications that must be addressed.
+I did find a couple of typos (e.g., L142: change "crated" to "created") and suggest the authors give the manuscript a once over.
 
-1.
-It is not clear what the various sources of data are.
-Perhaps this is because the authors have already described data sources in previous documentation of the dataset.
-However, I think some summary is needed for readers new to the dataset.
-I need to know what I am looking at when I see the various numbers in Figures 2-5, 8-11.
+**We fixed the noted typo and proofread the entire manuscript.**
 
-**In addition to addressing the specific questions below, we have added a link to a spreadsheet recording the datasets that have been incorporated, which is maintained in the open-access GitHub repository: "A record of data sets added to *ForC* is available at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv."**
-
-a.
-For example, is Fluxnet the source of GPP and Reco?
-The authors state that "ForC amalgamates numerous intermediary data sets" (line 119) and cite Luyssaert et al (2007).
-Is that the source of GPP and Reco, or have the fluxes been updated to newer Fluxnet data products?
-Similarly for the soil respiration data.
-Is this the data from Bond-Lamberty and Thomson (2010) or has it been updated?
-
-**We now clarify as follows:**
-
-**FLUXNET:**
-
-**Methods: "Third, we manually incorporated records of annual $NEP$, $GPP$, and $R_{eco}$ from the FLUXNET2015 dataset [@pastorello_fluxnet2015_2020], treating these records as authoritative when they duplicated earlier records (Appendix S1)."**
-
-**Appendix S1: "For eddy-covariance variables ($NEP$, $GPP$, $R_{eco}$), we retained the record associated with the most recent publication [most often @pastorello_fluxnet2015_2020], as these data are commonly re-assessed using new analysis methods."**
-
-**SRDB:**
-
-**Methods: "First, we imported (via R script) the Global Database of Soil Respiration Database [*SRDB* v4, 9488 records; @bond-lamberty_global_2010], and corrections and improvements to *SRDB* arising from this process were incorporated in *SRDB* v5 [@jian_restructured_2020]."**
-
-b.
-I need information on how Figures 2-5, 8-11 were prepared.
-Carbon stocks and fluxes such as NPP, litterfall, etc.
-are probably plot data.
-There is a detailed discussion of how the data were filtered to remove, for example, disturbance effects (lines 140-149).
-The GPP, Reco, and soil respiration databases seem to me to be different and may not be collocated with the plot data.
-Was disturbance filtering applied to the GPP, Reco, and soil respiration databases, too?
-
-**All records analyzed here have been fully incorporated into the ForC format and treated as described.
-We believe this should now be clearer where we describe the importation of new datasets: "Since publication of ForC v2.0, we imported three large additional databases [SRDB, GROA, FLUXNET15] into *ForC* via a combination of R scripts and manual edits.
-..."**
-
-c. If Luyssaert et al (2007) is a source of data, is there circularity in the comparison with latitudinal trends found in other datasets, for which Luyssaert et al is cited (lines 317-318)?
-
-**We don't see this as circular; we are reaffirming earlier results with an expanded dataset.
-To make it clear that we are not claiming to provide an independent test, we have edited this text as follows: "For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude (or increase with temperature) on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
-This consistency is not not surprising, particularly given commonality in the data analyzed or used for calibration."**
-
-2.
-Figures 2-5, 8-11 are excellent summaries of the dataset.
-Yet there is much unexplained in the figure and the reported data values.
-No guidance is given on how to use the data, and especially how to resolve discrepancies in the data.
-
-a.
-For example, Figure 2 lists foliage and woody aboveground biomass in tropical broadleaf forests.
-These do not sum to aboveground biomass (3.53 + 125.42 = 128.95 Mg C/ha vs 146.69).
-Nor do coarse root biomass and fine root biomass sum to root biomass (23.15 + 9.29 = 32.44 vs.
-21.86).
-Total biomass (147.96) is not the sum of aboveground (146.69) and root (21.86).
-
-**We acknowledge and discuss lack of closure in various places throughout the results and discussion.
-To ensure that this is clear from the figures, we have added the following statement to the captions of Figures 3-6 (formerly 2-5): "Mean component fluxes do not necessarily add up to the mean total fluxes because different sets of sites are included depending on availability of data (Figs.
-S5-S30)." We have also added a similar statement to the captions of Figures S1-S4 (formerly 8-11).
-We also make a related statement in the captions of new figures 9-12: "Note that there remain substantial uncertainties as to the functional form of age trends and discrepancies in closure among related variables."**
-
-b.
-Figure 4 for temperate conifer forests has large discrepancies in carbon stocks: e.g., 315.61 Mg C/ha for woody aboveground biomass, but only 167.46 for aboveground biomass and 226.58 for total biomass.
-The authors have a detailed discussion about lack of closure for root biomass, but do not discuss aboveground biomass.
-Presumably, this is because the data fall within the closure criteria, but what are we to make of the inconsistent data?
-
-**This is partially addressed by the response above, and covered in the results and discussion.
-In addition, to help explain the fact that lack of closure is particularly pronounced in temperate conifer forests, we have included errorbars in on the new schematic figure (Fig.
-1) indicating that temperate forests have particularly high variation in C stocks.**
-
-I understand that the data comes from many different sources and so may not always be compatible and that they have large standard deviations, but how are we to interpret and use the various data entries?
-For example, what has higher confidence: the component variables or the aggregate variables?
-What is the best estimate for aboveground biomass or root biomass?
-The authors take great pride in that component variables sum to within one standard deviation of the aggregate variables in all but one instance (lines 225-226, 301-302, 309-310).
-This is somewhat remarkable!
-But users of the dataset need guidance on how to interpret and use the data.
-
-**We have done two things to address this comment.
-First, to ensure accurate interpretation, we added the following statement to the captions of Figures 3-6 (formerly 2-5): "Mean component fluxes do not necessarily add up to the mean total fluxes because different sets of sites are included depending on availability of data (Figs.
-S5-S30)." Second, we have added the following guidance on using the data to the discussion: "*ForC*’s tens of thousands of records are readily available in a standardized format, along with all code used in the analyses presented here, and we recommend that researchers use these resources to identify and summarize data specific the analysis at hand."**
-
-3. There is a general sloppiness to the manuscript that makes me think the authors did not proofread the manuscript carefully.
-
-**We have proofread more carefully this time.**
-
-a.
-Line 58: The authors say GPP is estimated to be >69 Gt C/year, and cite Badgley et al.
-(2019).
-This statement is factually correct, but is quite misleading.
-Badgley et al.
-"estimate global annual terrestrial photosynthesis to be 147 Pg C/year (95% credible interval 131-163 Pg C/year)".
-The value of 69 Gt C is less than one-half that reported by Badgley.
-What are the authors trying to say?
-Such a large discrepancy does not instill confidence in the other numbers reported in the manuscript.
-
-**The 147 Pg C/year figure refers to *total* annual terrestrial photosynthesis, including non-forests.
-Table S4 in @badgley_terrestrial_2019 breaks this down by biome.
-The >69 Gt C/yr figure was obtained by summing the forest biomes, not including savannas/ mixed biomes.**
-
-b. Line 81: "rare exceptions that span regions or continents" is repeated twice.
-
-**Fixed.**
-
-c. Figure 1: Explain the gray scale for "forest cover". It seems like this refers to gradations of the various biomes, but the figure caption does not provide an explanation.
-
-**We have added an explanation.**
-
-d. Line 162: Clarify what is meant by "the minimum diameter breast height (DBH) threshold for tree census was <10cm"?
-
-**We have reworded this sentence for improved clarity: "Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), trees were censused down to a minimum diameter breast height (DBH) threshold of 10 cm or less."**
-
-e.
-Table 1: (i) The definition of GPP refers to NEE, but this in never defined.
-The preceding entry defines NEP not NEE.
-(ii) The description of the biome labels is inaccurate.
-The table uses "TrB" but the note refers to "Tr".
-The table uses "BoN", but the note refers to "B".
-
-**We have fixed both of these.**
-
-f.
-Line 210: I am confused by the statement that of the 39762 records in ForC v3.0, 11923 were included in this study.
-Previously (lines 135-149), the authors described creating "ForC-simplified" with 17349 records.
-Clarify the difference between ForC-simplified and ForC v3.0.
-
-**It appears that reference to ForC-simplified (essentially an automatically generated rearrangement of ForC that is purged of duplicates and organized for easy analysis) makes the process more confusing, and isn't necessary, so we eliminated mention of it.**
-
-g. Line 212: The authors refer to ForC, where in the previous sentence they referred to ForC v3.0, and previously used ForC-simplified. Clarify what database is used in the analyses.
-
-**As in the response above, we have eliminated unnecessary/confusing mention of ForC-simplified.
-The version used is ForC v3.0 (to be released with DOI upon acceptance of this manuscript).
-We do not feel it is necessary to state the version number every time ForC is mentioned.**
-
-4.
-This is not a review article in the traditional sense that a review is a critical assessment of recent papers in the field of carbon cycle research or identifies future research priorities.
-Instead, the paper documents a database and shows the utility of the database for carbon cycle research.
-My comments on the review aspect of the manuscript are:
-
-a. Yes, there is a need for this database; but
-
-b. No, this is not a critical and authoritative review of the carbon cycle; and
-
-c. The authors reference many other publications and datasets, but they do not critically evaluate their own dataset or other datasets.
-
-**We'd define this as a "quantitative review"--*i.e.*, one that quantitatively synthesizes previous studies.
-However, we agree that the previous version did little to critically review the corresponding literature.
-To improve in this area, we have expanded and reorganized the introduction into four sub-sections ("Forests in the global C cycle: current and future", "Evolution of forest C cycle research", "Biome differences", and "Age trends and their variation across biomes").
-The latter three sections are mostly or entirely new relative to the previous version, and briefly review the development and current status of the science on each topic.
-We have also added a schematic figure (Fig.
-1) to summarize current understanding of each topic and help with conceptual framing of the review.
-We also shifted emphasis more towards the ecological questions (trends across biomes and with age), while retaining (with less emphasis) the analysis of database representativeness.**
-
-# Referee: 3
+### Referee: 2
 
 COMMENTS TO THE AUTHOR(S)
+The authors have satisfactorily addressed my comments
 
-This papers uses a database recently created and updated by the author team – For C – to understand how forest carbon stocks and pools varies across broad biome classifications (e.g.
-boreal, temperature, and tropical) and stand age.
-The find that the rate of C cycling is faster in warmer climates, and that many C fluxes and pools increase with stand age (at least up to 100 years of age).
-
-There were several things I really liked about this paper.
-First, I applaud the author’s ambition in creating (and maintaining) this database, which has already been used in multiple high-profile papers.
-It is a novel idea to curate all the C fluxes and pools together in one virtual location, and the carbon cycle budgets illustrated in Figures 2-5, and 8-11, will likely function as useful “reality checks” against which both empirical and modeling results can be assessed.
-I also appreciated the focus on understanding how carbon cycles varies with stand age, as this is an important unknown that limits understanding of the usefulness of reforestation as a natural climate solution (among other unknowns).
-Overall I found the manuscript to be clearly written.
-
-**Thank you.**
-
-However, I also found that many aspects of the paper gave me considerable pause.
-
-First, the paper covers an awful lot of ground.
-It strikes me that each of the three research questions (bottom of page 4) could easily motivate an entire paper on their own.
-By attempting to address all three in one (relatively short) manuscript, it is not possible to discuss in any detail the mechanisms and processes that determine the results.
-There are no process-oriented hypotheses or frameworks against which the results are evaluated.
-
-**We agree that this covers a lot of ground, which corresponds with our vision of producing a comprehensive synthesis of broad trends.
-We envision that smaller questions can (and hopefully will) be broken off for more detailed analysis in the future.**
-
-**At the same time, we agreed that the manuscript could be improved with more of a conceptual framework.
-To improve in this area, we have expanded and reorganized the introduction into four sub-sections ("Forests in the global C cycle: current and future", "Evolution of forest C cycle research", "Biome differences", and "Age trends and their variation across biomes").
-The latter three sections are mostly or entirely new relative to the previous version, and briefly review the development and current status of the science on each topic.
-We have also added a schematic figure (Fig.
-1) to summarize current understanding of each topic and help with conceptual framing of the review.
-We also shifted emphasis more towards the ecological questions (trends across biomes and with age), while retaining (with less emphasis) the analysis of database representativeness.**
-
-For example, concerning expectations about how various C cycles and pools vary with stand age?
-There is a fairly extensive literature on this topic … Odum’s classic paper on the topic (“The Strategy of Ecosystem Development, 1969) has been cited thousands of times.
-But this literature is not referenced or cited in the current manuscript.
-Odum’s hypothesis suggests that NEP (arguably the most important flux, at least from a climate mitigation perspective) should increase initially with stand age and then eventually decrease as forests continue to mature.
-However, this framework/literature is not referenced, and the way the results are presented make it difficult to understand whether those expectations were borne out in the data.
-On that note, I didn’t understand why the authors chose to show flux trends explicitly as a function of age up to 100 years (Figure 6), but then group all the forests >100 years old into a single “bar).
-Why not plot the mature forests explicitly on the flux versus age axis?
-This would allow a clearer assessment of whether flux trends with age are really linear.
-It would also help the reader understand better one of the most striking results from this manuscript: that NEP of mature forests is indistinguishable across biomes.
-The authors describe this as unsurprising, but I think it is a bit unexpected, especially given the results of some of the more synthetic work from FLUXNET (e.g.
-Luyssaert et al.
-2007).
-The authors explain this result in a couple of sentences (Page 25, lines 22-26), speculating that the result is driver by “moderate disturbances” or “disequilibrium of Rsoil relative to C inputs).
-A deeper dive into the results, combined with some mechanistic grounding, might reveal to what extent this result represents a climate-age interaction that is predicted from the existing theory.
-
-**We greatly appreciated the suggestion to engage with Odum's framework.
-We now cite and discuss @odum_strategy_1969 in the introduction, and designed the age trends portion of the schematic figure to correspond to Odum's classic Fig.
-1 (updated with more details and current terminology).**
-
-**Regarding the question of showing flux trends as a function of age beyond 100 years, this unfortunately doesn't make sense within the context of the database because tropical forests can rarely be aged beyond 100 years (if that).
-This is because tropical trees rarely form annual rings, which are used to age (older) extratropical forests.
-It would not make sense to treat tropical and extra-tropical forests differently.**
-
-**Regarding the NEP result for mature forests, we agree that a better conceptual framework (now implemented) should help set the stage for explaining this result.
-We have modified the discussion paragraph on this to cite the schematic figure and also to note that this result does agree with @luyssaert_co2_2007 (although that study did not standardize for age).
-The paragraph now reads as follows:**
-
-**"The notable exception to the pattern of fluxes decreasing from tropical to boreal regions is $NEP$, which showed no significant differences across biomes (Fig.
-6f).
-Unlike the other C flux variables, $NEP$ does not characterize the rate at which C cycles through the ecosystem, but is the balance between C sequestration ($GPP$) and respiratory losses ($R_{eco}$) and represents net CO~2~ sequestration (or release) by the ecosystem (Fig.
-1).
-$NEP$ tends to be relatively small in mature forest stands (discussed further below), which accumulate carbon slowly relative to younger stands, if at all [@luyssaert_old-growth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
-It is therefore unsurprising that there are no pronounced differences across biomes [in agreement with @luyssaert_co2_2007], suggesting that variation in $NEP$ of mature forests is controlled less by climate and more by other factors including moderate disturbances [@curtis_forest_2018] or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition; @wilson_stability_2016]."**
-
-Finally, I wondered about the interacting effects of changing climate (rising CO2, warming temperatures) and stand age, especially in determining trends in the pools.
-Mature forests will have experienced a much wider range of climate conditions than younger stands.
-How does this complicate the comparison of live biomass across forest of different age?
-
-**This is an excellent question, and one that is of great interest to the lead author [e.g., @andersonteixeira_altered_2013]; however, we think it is unlikely to have much influence at this relatively coarse scale of analysis, especially given that much of the data in ForC was collected in the 20th century.
-Sorting this out would require carefully designed studies, and probably would not be possible using ForC data at this point.**
-
-Second, I had some questions about the representativeness of the dataset.
-While I appreciate that the authors choose to use distinct geographic areas as the unit of analysis (avoiding some issues of pseudo-replication from many observations from a single site), I still wondered about the extent to which the distinct geographic areas were representative of the climate space within each biome.
-For example, if mean annual climate versus presentation for all boreal, temperate and tropical locations are shown in a scatterplot (for example, using reanalysis data), and then mean annual temperature and precipation of the observations are shown on top, how much of the “climate space” is covered by the dataset?
-
-**An analysis of this type would need to be fairly involved, including a separate analysis for each variable.
-While do-able (we have previously published a plot similar to that described for the entire database), such an analysis would only begin to cover important aspects of representativeness (MAT and MAP, but not seasonality, soils, floristic composition, etc.).
-We feel that, in this case, the insights gained from the analysis would not justify the strain on time and resources that it would require.
-We have, however, given additional attention to the issue of representativeness in the revision.
-Specifically, we have done the following:**
-
-- **Set the stage for discussion of within-biome variability by including error bars on our schematic figure (Fig. 1).**
-
-- **Ensured that the issue is clearly and transparently presented in the figures--for example, by adding the following statement to the captions of Figures 3-6 (formerly 2-5): "Note that variables differ in geographical representation, resulting in potential imbalances (Figs.
-S5-S30).
-Probability that estimates reflect true biome means scales with the number of distinct geographical areas represented." **
-
-- **Revisited our presentation of representativeness issues in the results and discussion, revising text to ensure clear communication of the concept that biomed differences were influenced by geographical representation within the biome.
-For example, our discussion of stock differences across biome differences now reads: "...For $B_{tot}$ and $B_{ag}$, tropical broadleaf forests had the highest mean biomass and boreal forests the lowest, with temperate broadleaf and needleleaf ($B_{ag}$ only) intermediate.
-However, maximum values for stocks including live or standing woody biomass ($B_{tot}$, $B_{ag}$, $B_{ag-wood}$, $B_{root}$, $B_{root_coarse}$, $DW_{tot}$, $DW_{standing}$) consistently occurred in temperate biomes (Figs.
-1, 8, S20-S30).
-For variables that were disproportionately sampled in such high-biomass forests ($B_{ag-wood}$, $B_{foliage}$, and $B_{root-coarse}$; disproportionately sampled in  the US Pacific Northwest), temperate conifer forests had significantly higher stocks than the other biomes."**
-
-Third, I found the presentation of the results made it difficult to see clearly the major differences in C fluxes and pools across biomes and age classes.
-The illustrated C budgets (the majority of the figures, 8 in total) are visually very appealing, but the reader has to do a lot of flipping back and forth to see how any particular flux or pool varies across biomes and age class.
-Figure 6 is more synthetic, but each panel is very small and the differences from one group to the next are hard to see.
-It is also difficult to compare results for young forests (as scatterplots) with the box plots for the mature forests.
-My advice is to move some of the budgets to SI, and include in the main manuscript more figures that clearly illustrate the most interesting trends with biome, and to allow an expansion of the results in Figure 6 (for example, by first showing scatterplots of all forests, young and old, as a function of stand age), and then perhaps another that is a box plot comparison of mature versus old forests in each biome.
-
-**We have done a major re-arrangment of figures, including the following: (1) added a schematic figure, somewhat in the style of @odum_strategy_1969, to help with conceptual framing of the review; (2) added a figure synthesizing changes in fluxes through time for each biome; (3) removed maps from age trend diagrams (maps still available in SI), which makes the figures less busy/ easier to read; and (4) moved C cycle diagrams for young forests to SI.
-The result is that (1) the main manuscript contains only the more informative illustrated C budgets (mature forests); (2) the figures showing age trends and biome differences in 6 C flux and stock variables (Figs.
-7-8, formerly 6-7) are much easier to read; and (3) the new figure 9 provides an overall synthesis of results, focusing on the most trusted variables.**
-
-**Regarding the suggestion of showing age of old forests as a scatterplot, this is unfortunately not possible give unknown ages of most mature tropical forests, as explained above.**
-
-A few other comments:
-
-Page 6, first paragraph: I wondered about the extent to which filtering the data for “managed” affected the results.
-In the Eastern US, for example, its difficult to find any forests on public land that aren’t managed to some extent (for example, through periodic selective harvests), and many of them regenerated from “planted” stands back in the 1930s.
-I would be curious to know if including “managed” forests substantially altered results.
-
-The binary age classification (two ages, less than 100 years or greater than 100 years) was difficult to accept, as forests that are in the 80-100 year age range are often considered to be mature (at least in the temperate zone).
-I realize that even with a dataset as rich as For C, data availability will limit stratification into too many bins.
-Nonetheless, at least for some measurements, I wonder if it's possible to consider a greater number of age classes (for example, young, maturing, and mature).
-
-**We prefer not to separate forests of known age into categories, but rather to represent age as a continuous variable, when possible.
-In large part this is motivated by the fact that appropriate age thresholds for various categories would vary across biomes, given differences in biomass accumulation rates.
-While our ideal would be to have no categorical divisions, unknown stand ages--particularly in the tropics--force us to divide the data into the categories used here.**
-
-**To provide a better representation of age trends, we created new figure 9.
-Although it does not break forests into additional age categories, it does address what seems to be at the core of this concern, which is a concise representation of differences across the full range of stand ages.**
-
-The consistency check (e.g.
-do component fluxes/pools sum to within one standard deviation of the aggregate flux or pool) seems like it is destined to provide a favorable assessment of the degree of closure so long as the data within each grouping represent a wide range of natural climatic and soil variability.
-As long as the aggregate variable has a large standard deviation, the results are likely guaranteed to be “consistent.” It seems this metric would benefit from some simulations (perhaps with artificially generated data) to understand exactly how poor closure needs to be at the site-level to generate an inconsistent results when aggregated across sites.
-
-**We agree that such an analysis would be valuable, but time and resource constraints place it beyond the scope of what is currently feasible in this already very large analysis.
-We believe that the current consistency check is valuable as a first pass for identifying cases of extreme mis-matches, and therefore retain it as is.
-However, we agree that it is not the ideal test, and that it is not very stringent (given high standard deviations) and have therefore toned down claims that the budgets are "closed"/"internally consistent" throughout the manuscript (e.g., removed from abstract).
-We note that the discussion already stated that our criteria for closure was relatively loose (and this statement is unchanged in the revised version): "On the other, however, *ForC* derives data from multiple heterogeneous sources, and standard deviations within each biome are high; as a result, the standard for C closure is relatively loose [*c.f.* @houghton_terrestrial_2020]."**
-
-# Referee: 4
+### Referee: 3
 
 COMMENTS TO THE AUTHOR(S)
-General:
-This paper presents the findings of a rather comprehensive modeling analysis of C fluxes and stocks in the world’s major forest biomes.
-An important strength of this work is its reliance on a fairly large empirical data set to calculate a comprehensive suite of fluxes and stocks to close (or come close) C budgets in these systems.
-Another unique component of this work is the contrasts of young vs mature forests in each biome.
-An important contribution of this work is its highlighting of gaps in data (e.g., deadwood) and non-random distribution of empirical data (though this, of course is well-known) and how these factors influence C accounting efforts.
-Overall, the paper is well written with beautiful figures and the analyses and data sets are sound.
-I think this work makes an important and timely contribution to fields of C cycle science, forest ecology, and ecosystem ecology and is likely to be of interest to many ERL readers.
-I have no fundamental concerns with this manuscript.
-However, in addition to the detailed comments/suggestions below, I think it would greatly benefit the paper if the authors could include some discussion of how the forest ecosystems they characterize here compare to the forest ecosystems that actually exist.
-The data here represent generally interior forest ecosystems, which of course are incredibly important.
-But, work over the last 5 years or so highlights the extent to which forest fragmentation influences a large proportion of the world’s forests (Haddad et al., 2015).
-Fragmentation and the creation of edges has been shown to have important implications for C stocks and fluxes with regional and global implications (e.g., (Chaplin-Kramer et al., 2015; Remy et al., 2016; Reinmann & Hutyra, 2017; Smith et al., 2019; Ordway & Asner, 2020; Reinmann et al., 2020: FULL REFS BELOW).
-Logging and other forms of management also influence a large proportion of the world’s forests.
-I am not suggesting this be included in modeling efforts here, but in placing this work in the broader context of C stocks and fluxes of the world’s forests I think it would do the scientific community a great service to more explicitly recognize what is being modeled and perhaps the proportion of the world’s forests these data might represent…even if discussed in a qualitative sense.
+I find the authors did a thorough and good job responding to concerns from the last round of review.
+I appreciate the re-organization of the introduction, and the changes made to the figures (including the inclusion of a conceptual figure).
+This is a nice paper.
+I have just two suggestions for minor amendments before final submission:
 
-**Thank you. We have added the following paragraph (citing the references given above) on representation of forest types in the Discussion section "C variable coverage and budget closure":**
+a) It'd be nice to return explicitly to the Odum hypotheses more clearly in the discussion.
+It seems that the results, as they are, do not support the prediction that NEP becomes near-zero in the most mature ecosystems (noting that Odum predicted this would happen in the southeastern US by 100 years of age).
+I think its important to highlight this fact explicitly (though possible some of this result is driven by CO2 fertilization effects).
 
-**Importantly, *ForC* and the analyses presented here cover the forests that have received research attention, which are not a representative sample of the world's existing forests--geographically or in terms of human impacts [@martin_mapping_2012].
-Geographically, all variables are poorly covered in Africa and Siberia (Fig.
-1), a common problem in the carbon-cycle community [@xu_contribution_2016; @schimel_effect_2015].
-In terms of human impacts, research efforts tend to focus on interior forest ecosystems [@martin_mapping_2012], often in permanently protected areas [e.g., @davies_forestgeo_2021].
-Studies of regrowth forests tend to focus on sites where recurring anthropogenic disturbance is not a confounding factor.
-Yet, fragmentation and degradation impact a large and growing proportion of Earth's forests [@fao_and_unep_state_2020].
-Fragmentation and the creation of edges strongly impacts forest C cycling [e.g., @chaplin-kramer_degradation_2015; @remy_strong_2016; @reinmann_edge_2017; @smith_evidence_2019; @reinmann_urbanization_2020;  @ordway_carbon_2020].Partial logging and other forms of non- stand clearing anthropogenic disturbance also strongly impact forest C cycling [e.g., @huang_long-term_2010; @piponiot_carbon_2016] but are under-studied [@sist_tropical_2015] and excluded from this analysis.
-Fragmented and degraded forests do not fit the idealized conceptual framework around which this review is structured (Fig.
-1), yet their representation in models, sustainability assessments, and C accounting systems is critical to accurate accounting of C cycling in Earth's forests [e.g., @huang_long-term_2010; @reinmann_edge_2017; @smith_evidence_2019; @piponiot_can_2019].
-Finally, plantation forests account for approximately 3% of Earth's forests [@fao_and_unep_state_2020] but are not included in this analysis.
-While it is known that these tend to accumulate biomass faster than naturally regenerating forests [@anderson_temperature-dependence_2006; @bonner_meta-analytical_2013], their global scale C cycling patterns remain less clearly understood [c.f.
-@cook-patton_mapping_2020].
-Additional research and synthesis are needed to fill these critical gaps in our understanding of forest C cycling.**
+**We have modified the paragraph in the discussion that discusses age trends in $NEP$ to refer back to the Odum hypothesis. It now reads as follows:**
 
-Introduction:
+**Notably, net carbon sequestration (NEP) exhibits an overall increase with age across the first 100 years of stand development, with more pronounced patterns in temperate than boreal forests (Fig.
+7).
+This finding is consistent with previous studies showing an increase in NEP across relatively young stand ages (Pregitzer and Euskirchen 2004, Baldocchi et al 2001, Luyssaert et al 2008).
+However, NEP is theoretically expected to peak in intermediate-aged stands and thereafter decline, consistent with decelerating C accumulation as stands age (Fig.
+9; Odum 1969), and such declines have been observed (Law et al 2003, Luyssaert et al 2008).
+The fact that NEP values estimated by our models for 100-year-old stands were not systematically different from those of mature stands (lower for temperate broadleaf, higher for temperate conifer, and equal for boreal; Fig.
+9) may be driven by differences in geographical representation across age classes or by the fitting of an inappropriate functional form.
+Moreover, both biomass and non-living C stocks often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands (Luyssaert et al 2008, McGarvey et al 2014, Lichstein et al 2009).
+Additional data, including on age trends of deadwood, the organic layer, and soil C will be important to parsing the timing and extend of an age-related NEP decrease across forest biomes.**
 
-L81: “exceptions” in sentence twice.
+Figures 7 and 8 could be made wider (if not necessarily taller), so they take up the full width of the page, which might allow the authors to inch up the font size. Some of that text is really tiny.
 
-**Fixed.**
-
-L97: Typo “Since the its most”
-
-**Fixed.**
-
-Methods:
-
-L144: If there is a non-trivial proportion of the world’s forests (especially in certain biomes) is plantation or planted forests, does removing such plots from the dataset bias the results of a global modeling product?
-
-**We have added a statement on plantation forests in the paragraph representation of forest types: Finally, plantation forests account for approximately 3% of Earth's forests [@fao_and_unep_state_2020] but are not included in this analysis.
-While it is known that these tend to accumulate biomass faster than naturally regenerating forests [@anderson_temperature-dependence_2006; @bonner_meta-analytical_2013], their global scale C cycling patterns remain less clearly understood [c.f.
-@cook-patton_mapping_2020]."**
-
-Results:
-
-L243-245: I think this is per unit forest area, correct? If so, I think it would be helpful to specify that here. Also, the reader should be referred to Figs 2-5, not just 5, right?
-
-**Fixed.**
-
-L270-271: The wording of the sentence “There were sufficient data to model….” Seems a bit awkward. Should it read “…WHICH were also significant…”?
-
-**We have reworded this sentence to read, "Age $\times$ biome interactions were also significant for all ten of these C stock variables (Table S2), with living C stocks tending to accumulate more rapidly during the early stages of forest regrowth in tropical forests (Figs.
-8, S20-S30)."**
-
-Discussion:
-You might consider reiterating in the first paragraph of the Discussion section that your findings indicate that Temperate Broadleaf Forests are actually the most productive forest biome in terms of NEP.
-I think this is a surprising finding (we normally think of Tropical forests as being more productive) that warrants attention.
-
-**We have added the following sentence to the first paragraph of the discussion: "The notable exception was mature forest $NEP$, which, as the difference between $GPP$ and $R_{eco}$, was statistically indistinguishable across biomes (Fig.
-7f)." Given that there is no signficant difference among biomes, we prefer not to discuss temperate broadleaf forests as higher.**
-
-L297: The fact that ForC does not include soil C is important.
-It is not a flaw in the model or the approach here, but throughout you discuss C stocks, which many would interpret as being inclusive of soil C.
-I think the authors should consider clearly indicating in the Methods section, and perhaps reiterating in the Results and Discussion, that stocks here are defined as litter layer, biomass, and necromass, but excludes C in soil.
-
-**We have added the following statement to the methods section: "We did not analyze soil carbon, which is not a focus of the *ForC* database." We also now note this in the first paragraph of the discussion "There was also little directional variation in mean mature forest C stocks (biomass, dead wood, and organic layer) across biomes, although maximum values for the majority of stocks (all including live or standing woody biomass) occurred in temperate biomes (Figs.
-1, 2-5, 8)."**
-
-L322-323: As you point out in the Results section, while there are no statistically sig differences in NEP across biomes, there are large differences in the means.
-Is this an artifact of the data sources used (i.e.
-distribution and number of sites with empirical data)?
-Can you speak a little more to this point in this section?
-
-**We have modified sentences in this paragraph to address this: "The notable exception to the pattern of fluxes decreasing from tropical to boreal regions is $NEP$, which showed no significant differences across biomes, albeit with the highest mean in temperate broadleaf forests (Fig.
-7f).
-...
-The fact that mature temperate broadleaf forests have a higher mean than the other biomes may reflect the fact that most of these forests are older secondary forests that, while classified here as mature, are still accumulating carbon [@curtis_forest_2018].**
-
-L382: You might consider changing text to “increases with age AT LEAST up to the 100-yr threshold examined here” so that it does not come across as suggesting that NEP only increases for the first hundred years of stand development, which we know is not true.
-
-**Reworded to read, "Notably, net carbon sequestration ($NEP$) exhibits an overall increase with age across the first 100 years of stand development...".**
-
-L445-450: I come back to the data in Figs 2-5 and text in the results section (L243-245), which indicates no sig differences in NEP, but highest means in temperate forests.
-Of course, the high biomass in tropical forests makes them critical to protect from a C storage perspective, but if NEP (i.e.
-rates of C sequestration) are highest, at least as a mean, in temperate broadleaf forests how should those ecosystems factor into conservation priorities.
-Related to this point, is the high NEP of temperate forests driven by the relatively young nature of temperate broadleaf forests in the eastern U.S., where a lot of data exist?
-
-**This question is now addressed earlier in the discussion (see response to comment on L322-323).
-The higher NEP of mature temperate forests contributes to their value for climate change mitigation, but does not make them more valuable than higher-biomass tropical forests (for example) that have already sequestered C that would become committed to eventual release as CO~2~ emissions if the forest were to be cleared (with the rapidity of that release depending on mode of clearing, wood product decay...).
-That trade-off is explored to some extent in one of the references given in this paragraph [@anderson-teixeira_greenhouse_2011], but seems too complex to get into here.**
-
-References mentioned above:
-
-Chaplin-Kramer R., Ramler I., Sharp R., Haddad N.
-M., Gerber J.
-S., West P.
-C., … King H.
-(2015).
-Degradation in carbon stocks near tropical forest edges.
-Nature Communications, 6, 1–6.
-https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1038%2Fncomms10158&amp;data=04%7C01%7Cteixeirak%40si.edu%7Cd27c53d6ec434c8ae42108d8a03017cf%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637435475707545147%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C1000&amp;sdata=maYY3mHl%2F%2FX0jgkAOK5Kc9K1TpaoTVkFM52HsOxeMjU%3D&amp;reserved=0.
-
-Haddad N.
-M., Brudvig L.
-a., Clobert J., Davies K.
-F., Gonzalez A., Holt W.
-M., … Townshend J.
-R.
-(2015).
-Habitat fragmentation and its lasting impact on Earth’s ecosystems.
-Science Advances, (March), e1500052.
-https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.4028%2Fwww.scientific.net%2FAMM.315.108&amp;data=04%7C01%7Cteixeirak%40si.edu%7Cd27c53d6ec434c8ae42108d8a03017cf%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637435475707545147%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C1000&amp;sdata=1L35xf2RdBj6Y9PjsYhSY5tct4RhnJLE8Jd50QJdzBM%3D&amp;reserved=0.
-
-Ordway E.
-M.
-& Asner G.
-P.
-(2020).
-Carbon declines along tropical forest edges correspond to heterogeneous effects on canopy structure and function.
-Proceedings of the National Academy of Sciences of the United States of America, 117(14), 7863–7870.
-https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1073%2Fpnas.1914420117&amp;data=04%7C01%7Cteixeirak%40si.edu%7Cd27c53d6ec434c8ae42108d8a03017cf%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637435475707545147%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C1000&amp;sdata=DnTMc4LCIoRcHMLUEAcpvNOn%2FeJLO3%2Bh8pXVYSUb%2BI4%3D&amp;reserved=0.
-
-Reinmann A.
-B.
-& Hutyra L.
-R.
-(2017).
-Edge effects enhance carbon uptake and its vulnerability to climate change in temperate broadleaf forests.
-Proceedings of the National Academy of Sciences, 114(1), 107–112.
-https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1073%2Fpnas.1612369114&amp;data=04%7C01%7Cteixeirak%40si.edu%7Cd27c53d6ec434c8ae42108d8a03017cf%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637435475707545147%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C1000&amp;sdata=GncJJ%2FLa2f96c0tjkl5RZZv2JLJFoAoQnOTm0b%2BYyNo%3D&amp;reserved=0.
-
-Reinmann A. B., Smith I. A., Thompson J. R. & Hutyra L. R. (2020). Urbanization and fragmentation mediate temperate forest carbon cycle response to climate. Environmental Research Letters, 15, 114036.
-
-Remy E., Wuyts K., Boeckx P., Ginzburg S., Gundersen P., Demey A., … Verheyen K.
-(2016).
-Strong gradients in nitrogen and carbon stocks at temperate forest edges.
-Forest Ecology and Management, 376(2016), 45–58.
-https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1016%2Fj.foreco.2016.05.040&amp;data=04%7C01%7Cteixeirak%40si.edu%7Cd27c53d6ec434c8ae42108d8a03017cf%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637435475707545147%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C1000&amp;sdata=u2ylmC4K4VIxHWamrCnk%2BNfu%2FsZDUkSPfABgfN%2BCzT4%3D&amp;reserved=0.
-
-Smith I.
-A., Hutyra L.
-R., Reinmann A.
-B.
-& Thompson J.
-R.
-(2019).
-Evidence for Edge Enhancements of Soil Respiration in Temperate Forests Geophysical Research Letters.
-Geophysical Research Letters, 46, 1–10.
-https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1029%2F2019GL082459&amp;data=04%7C01%7Cteixeirak%40si.edu%7Cd27c53d6ec434c8ae42108d8a03017cf%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637435475707545147%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C1000&amp;sdata=92jjRZSUzQrNX6SGlV%2BZRfHlUN%2B6AL96%2FJIl4PIK3a8%3D&amp;reserved=0.
-
-\newpage
-## References
-
-```{r eval=FALSE, echo=FALSE}
-# All references should be presented in alphabetical order and should also be bundled in ways that will help the reader to find those that are germane to each of the themes, topics or issues that the authors have used to organize the review.
-Such a matrix will be included as supplementary material.
-ERL has no standardized format for subdividing references by cross-cutting themes, but guidance from editors will be helpful to authors.
-```
+**We agree that these figures should be full-page width. They were displayed at a smaller size than intended because of low resolution, which we have now fixed.**

### 2021-02-24 manuscript/ERL_global_C_review.Rmd
work on submission files
address all reviewer concerns

@@
 Early research on this theme showed that biomass accumulation rates during secondary succession increase with temperature on a global scale [@johnson_post-disturbance_2000; @anderson_temperature-dependence_2006] and with water availability in the neotropics [@poorter_biomass_2016].
-Most recently, @cook-patton_mapping_2020 reinforced these earlier findings with a much larger dataset and crated a high-resolution global map of estimated potential C accumulation rates.
+Most recently, @cook-patton_mapping_2020 reinforced these earlier findings with a much larger dataset and created a high-resolution global map of estimated potential C accumulation rates.
 However, there has been little synthesis of cross-biome differences in variables other than biomass and its accumulation rate (but see @cook-patton_mapping_2020 for $DW$, $OL$, and soil C accumulation in young stands).
@@
 This finding is consistent with previous studies showing an increase in $NEP$ across relatively young stand ages [@pregitzer_carbon_2004; @baldocchi_fluxnet_2001; @luyssaert_old-growth_2008].
-However, $NEP$ has been observed to decline from intermediate to old stands [@law_changes_2003; @luyssaert_old-growth_2008], whereas the $NEP$ values estimated by our models for 100-year-old stands were not systematically different from those of mature stands (lower for temperate broadleaf, higher for temperate conifer, and equal for boreal; Fig.
-`r age_trends_all`).
-This lack of a consistent age-related decrease may be driven by differences in geographical representation across age classes or by the fitting of an inappropriate functional form.
-A decrease in $NEP$ would be consistent with the observed deceleration of C accumulation as stands age (Fig.
-`r age_trends_all`), although both biomass and non-living C stocks will often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands [@luyssaert_old-growth_2008; @mcgarvey_carbon_2014; @lichstein_biomass_2009].
+However, $NEP$ is theoretically expected to peak in intermediate-aged stands and thereafter decline, consistent with decelerating C accumulation as stands age [Fig.
+`r age_trends_all`; @odum_strategy_1969], and such declines have been observed [@law_changes_2003; @luyssaert_old-growth_2008].
+The fact that $NEP$ values estimated by our models for 100-year-old stands were not systematically different from those of mature stands (lower for temperate broadleaf, higher for temperate conifer, and equal for boreal; Fig.
+`r age_trends_all`) may be driven by differences in geographical representation across age classes or by the fitting of an inappropriate functional form.
+Moreover, both biomass and non-living C stocks often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands [@luyssaert_old-growth_2008; @mcgarvey_carbon_2014; @lichstein_biomass_2009].
 Additional data, including on age trends of deadwood, the organic layer, and soil C will be important to parsing the timing and extend of an age-related $NEP$ decrease across forest biomes.
@@
 
-It is also interesting to consider the complementary utility of global-scale but spatially discontinuous databases such as ForC and remote wall-to-wall remote sensing products. 
+It is also interesting to consider the complementary utility of global-scale but spatially discontinuous databases such as ForC and wall-to-wall remote sensing products. 
 The latter provide insight into aboveground carbon stocks, but less constraint on belowground stocks or carbon fluxes in general [@bondlamberty_estimating_2016; @anav_spatiotemporal_2015]. 

### 2021-02-25 manuscript/ERL_global_C_review.Rmd
working on Helene's comments
#70

@@
 Specifically, we use `r n_records` *ForC* records for `r nrow(T1)` C cycle variables from `r n_areas_tot` geographic locations to characterize ensemble C budgets for four broad forest types -- tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and boreal.
-We calculate statistics for both mature and regrowth (age <100 years) forests, and quantify trends with stand age in regrowth forests for all variables with sufficient data.
+We calculate means and standard deviations for both mature and regrowth (age <100 years) forests, and quantify trends with stand age in regrowth forests for all variables with sufficient data.
 
-*Review Results/ Synthesis.*   The rate of C cycling generally decreased from tropical to boreal regions in both mature and regrowth forests, whereas C stocks showed less directional variation.
-Net ecosystem production of mature forests was indistinguishable across biomes.
-The majority of flux variables, together with most live biomass pools, increased significantly with stand age when fit with logarithmic functions.
+*Review Results/ Synthesis.*   The rate of C cycling generally decreased from tropical to temperate to boreal in both mature and regrowth forests, whereas C stocks showed less directional variation.
+Mature forest net ecosystem production did not differ significantly among biomes.
+The majority of flux variables, together with most live biomass pools, increased significantly with the logarithm of stand age.
 
 *Discussion.* As climate change accelerates, understanding and managing the carbon dynamics of forests is critical to forecasting, mitigation, and adaptation.
-This comprehensive and synthetic global overview of C stocks and fluxes across biomes and stand ages will help to advance these efforts.
+This comprehensive and synthetic global overview of C stocks and fluxes across biomes and stand ages contributes to these efforts.
 
@@
 A robust understanding of forest impacts on global C cycling is essential. 
-Total annual photosynthesis in forests (gross primary productivity, $GPP$) is estimated at approximately 69 Gt C yr^-1^ [@badgley_terrestrial_2019], more than seven times the average annual fossil fuel emissions during 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
-Most of this enormous C uptake is counterbalanced by releases to the atmosphere through ecosystem respiration ($R_{eco}$) and fire, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] and fire [@van_der_werf_global_2017].
+Total annual photosynthesis in forests (gross primary productivity, $GPP$) is estimated at approximately 69 Gt C yr^-1^ [@badgley_terrestrial_2019].
+Most of this enormous C uptake is counterbalanced by releases to the atmosphere through ecosystem respiration ($R_{eco}$) and fire.
 In recent years, total forest C uptake has exceeded releases, such that globally forests have been a C sink [@harris_global_2021].
@@
 However, deforestation, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reduces the net forest sink to ~1.1-2.2 Gt C yr^-1^ [@friedlingstein_global_2019; @harris_global_2021].
+Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thus central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; @cavaleri_urgent_2015].
 
-The future of the current forest C sink is dependent both upon forest responses to climate change itself and human land use decisions, which will feedback and strongly influence the course of climate change [@friedlingstein_climatecarbon_2006].
-Regrowing forests in particular will play an important role [@pugh_role_2019], as almost two-thirds of the world’s forests were secondary as of 2010 [@fao_global_2010].
-As anthropogenic and climate-driven disturbances impact an growing proportion of Earth's forests [@andela_human-driven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests is increasingly important [@andersonteixeira_altered_2013].
-Although age trends in aboveground biomass have been well-studied and synthesized globally [@cook-patton_mapping_2020], a relative dearth of data and synthesis on other C stocks and fluxes in secondary forests points to an under-filled need to characterize age-related trends in forest C cycling.
-Such understanding is particularly critical for reducing uncertainty regarding the potential for carbon uptake and climate change mitigation by regrowth forests [@krause_large_2018; @cook-patton_mapping_2020].
-Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thus central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; @cavaleri_urgent_2015].
+The future of the current forest C sink is dependent both upon forest responses to climate change and human land use decisions, with land use change itself strongly influencing the course of climate change [@friedlingstein_climatecarbon_2006].
+Regrowing forests (i.e., secondary forests) will play a particularly important role [@pugh_role_2019], as almost two-thirds of the world’s forests were secondary as of 2010 [@fao_global_2010].
+As anthropogenic and climate-driven disturbances impact a growing proportion of Earth's forests [@andela_human-driven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests is increasingly important [@andersonteixeira_altered_2013].
+Although age trends in aboveground biomass have been well-studied and synthesized globally [@cook-patton_mapping_2020], there is a relative dearth of data and synthesis on other C stocks and fluxes in secondary forests.
+An understanding of age-related trends in forest C cycling is particularly critical for reducing uncertainty regarding the potential for carbon uptake and climate change mitigation by regrowth forests [@krause_large_2018; @cook-patton_mapping_2020].
 
@@
 
-For more than half a century, researchers have sought to understand how forest carbon cycling varies across stands, including those of different biomes [e.g., @lieth_primary_1973; @luyssaert_co2_2007] and stand ages [e.g., @odum_strategy_1969; @luyssaert_old-growth_2008].
+For more than half a century, researchers have sought to understand how forest carbon cycling varies across stands, including among biomes [e.g., @lieth_primary_1973; @luyssaert_co2_2007] and with stand age [e.g., @odum_strategy_1969; @luyssaert_old-growth_2008].
 Over this time, an increasingly refined conceptual understanding of the elements of ecosystem C cycles has developed, as a growing number of variables have been defined [e.g., @chapin_reconciling_2006], along with appropriate measurement methods [e.g., @clark_measuring_2001].
-New technology has also enabled researchers to directly measure an expanding set of variables, notably including the development of continuous measurements of soil CO~2~ efflux [@kuzyakov_sources_2006] and ecosystem-atmosphere CO~2~ exchange [@baldocchi_fluxnet_2001].
+New technology has also enabled researchers to directly measure an expanding set of variables, notably including continuous measurements of soil CO~2~ efflux [@kuzyakov_sources_2006] and ecosystem-atmosphere CO~2~ exchange [@baldocchi_fluxnet_2001].
 Measurement techniques have been increasingly standardized; for example, of the biomass allometries that strongly influence estimates of most C cycle variables [e.g., @chave_improved_2014].
@@
 
-Forest C cycling varies enormously across biomes, which cateogrize the world's forests according to major differences in climate, vegetation, *etc.* 
-Since the early 19th century, it has been recognized that climate plays a dominant role in shaping differences among forests on a global scale [@humboldt_essay_1807; @holdridge_determination_1947].
+Forest C cycling varies enormously across biomes, categories that encapsulate major differences in climate and vegetation.
+The dominant role of climate in shaping global variation among forests has been recognized since the early 19^th^ century [@humboldt_essay_1807; @holdridge_determination_1947].
 Global scale data syntheses have shown that C fluxes including $GPP$, net primary productivity ($NPP$), and soil respiration ($R_{soil}$) decrease with latitude or, correspondingly, increase with mean annual temperature [Fig.
@@
 
-Correlative analyses relating C cycle variables to climate and other environmental variables have recently been taken to a new level through use of machine-learning algorithms that relate ground-based C cycle data to global maps of environmental covariates, making it possible to create fine-scale global maps of C cycling [e.g., @warner_spatial_2019; @cook-patton_mapping_2020].
-This approach can be particularly effective when paired with satellite measurements that correlate to C cycle variables of interest; for example, solar-induced chlorophyll fluorescence is useful for fine-scale mapping of gross primary productivity [$GPP$; @li_mapping_2019], while LiDAR, radar, and optical imagery are being used to estimate $B_{ag}$ at regional to global scales [e.g., @saatchi_benchmark_2011; @hu_mapping_2016].
-Any such analysis is however constrained by the quality and coverage of ground-based estimates of forest C fluxes or stocks [e.g., @schepaschenko_forest_2019]. 
-While estimates of some variables (e.g., $B_{ag}$, $GPP$, $NPP$, $R_{soil}$) are widely available, many remain poorly characterized (e.g., $DW_{tot}$; $OL$; autotrophic respiration, $R_{auto}$) --even at the coarse resolution of biomes.
-This is a critical limitation not only for understanding forest C cycling, but also for quantifying forest-based climate change mitigation across forest biomes or ecozones [e.g., @ipcc_2019_2019].
+Phenomenological analyses relating C stocks and fluxes to climate and other environmental variables have recently been taken to a new level through use of machine-learning algorithms that relate ground-based C cycle data to global maps of environmental covariates [e.g., @warner_spatial_2019; @cook-patton_mapping_2020].
+The resulting models enable construction of fine-scale global maps of estimated C cycling variables.
+This approach can be particularly effective when it integrates satellite measurements that correlate with C cycle variables of interest; for example, solar-induced chlorophyll fluorescence is useful for fine-scale mapping of gross primary productivity [$GPP$; @li_mapping_2019], while LiDAR, radar, and optical imagery are being used to model $B_{ag}$ at regional to global scales [e.g., @saatchi_benchmark_2011; @hu_mapping_2016].
+However, all such analyses are ultimately constrained by the quality and coverage of ground-based estimates of forest C fluxes or stocks to train models [e.g., @schepaschenko_forest_2019]. 
+While estimates of some variables (e.g., $B_{ag}$, $GPP$, $NPP$, $R_{soil}$) are widely available, many remain poorly characterized (e.g., $DW_{tot}$; $OL$; autotrophic respiration, $R_{auto}$) even at the coarse resolution of biomes.
+This is a critical limitation not only for understanding forest C cycling, but also for quantifying forest-based climate change mitigation potential across forest biomes or ecozones [e.g., @ipcc_2019_2019].
 
@@
 `r schematic`; @andersonteixeira_altered_2013].
-Following stand-clearing disturbance, $GPP$, $NPP$, and biomass of leaves ($B_{foliage}$) and fine roots ($B_{root-fine}$) increase rapidly and thereafter remain relatively stable ($B_{foliage}$, $B_{root-fine}$, sometimes $GPP$) or decline slightly [$NPP$, sometimes $GPP$; e.g., @law_changes_2003; @pregitzer_carbon_2004; @amiro_ecosystem_2010; @goulden_patterns_2011].
+Following stand-clearing disturbance, $GPP$, $NPP$, and biomass of leaves ($B_{foliage}$) and fine roots ($B_{root-fine}$) initially increase rapidly and thereafter remain relatively stable ($B_{foliage}$, $B_{root-fine}$, sometimes $GPP$) or decline slightly [$NPP$, sometimes $GPP$; e.g., @law_changes_2003; @pregitzer_carbon_2004; @amiro_ecosystem_2010; @goulden_patterns_2011].
 The decline in $NPP$ occurs because $R_{auto}$ increases relative to $GPP$ as forests age, corresponding to declining carbon use efficiency with stand age [@delucia_forest_2007; @collalti_forest_2020].
 Heterotrophic respiration, most of which originates from the soil ($R_{het-soil}$) remains relatively constant with stand age [@law_changes_2003; @pregitzer_carbon_2004; @goulden_patterns_2011], with the result that net ecosystem production ($NEP=GPP-R_{eco}$, where $R_{eco}$ is total ecosystem respiration) is initially negative, increases to a maximum at intermediate ages, and declines--typically to a small positive value--thereafter [@law_changes_2003; @pregitzer_carbon_2004; @amiro_ecosystem_2010; @goulden_patterns_2011; @luyssaert_old-growth_2008].
-The result is that biomass accumulates rapidly in young forests, followed by a slow decline to near zero in old forests [e.g., @lichstein_biomass_2009; @yang_carbon_2011].
+The result is that biomass accumulation is rapid in young forests, followed by a slow decline to near zero in old forests [e.g., @lichstein_biomass_2009; @yang_carbon_2011].
 While these trends have been subject of fairly recent qualitative review [@andersonteixeira_altered_2013], there is need for a synthetic, quantitative review taking advantage of the greatly expanded data now available.
@@
 Age trends, which represent idealized dynamics following a disturbance that removes all living vegetation, are an updated version of the classic figure from Odum (1969), with heavy lines corresponding to those in Odum's figure and $NEP$ corresponding to Odum's 'net production'.
-Here, $NEP$ conists primarily of woody aboveground net primary production ($ANPP_{woody}$), while $ANPP_{woody.turnover}$ is the sum of woody mortality and branch turnover.
+Here, $NEP$ consists primarily of woody aboveground net primary production ($ANPP_{woody}$), while $ANPP_{woody.turnover}$ is the sum of woody mortality and branch turnover.
 Dotted lines refer to decomposition of potential 'legacy' organic material produced prior to the disturbance and remaining at the site (e.g., standing and fallen dead wood, $DW_{tot}$; soil organic matter).
@@
 We did not analyze soil carbon, which is not a focus of the *ForC* database.
-Note that two flux variables, aboveground heterotropic respiration ($R_{het-ag}$) and total respiration ($R_{het}$), were included for conceptual completeness but had no records in *ForC* (Table 1).
+Note that two flux variables, aboveground heterotropic respiration ($R_{het-ag}$) and total heterotropic respiration ($R_{het}$), were included for conceptual completeness but had no records in *ForC* (Table 1).
 Records for our focal variables represented `r round(n_focal_variable_records_eligible_for_inclusion/records_eligible_for_inclusion*100,digits=1)`% of the total records eligible for inclusion.
@@
 Specifically, net ecosystem exchange [measured by eddy-covariance; @baldocchi_fluxnet_2001] and biometric estimates of $NEP$ were combined into the single variable $NEP$ (Table 1).
-Furthermore, for $NPP$, aboveground $NPP$ ($ANPP$), and the litterfall component of $ANPP$ ($ANPP_{litterfall}$), *ForC* variables specifying inclusion of different components were combined (*e.g.*, measurements including or excluding fruit and flower production and herbivory).
+Furthermore, for $NPP$, aboveground $NPP$ ($ANPP$), and the litterfall component of $ANPP$ ($ANPP_{litterfall}$), we combined ForC variables specifying inclusion or exclusion of minor components (*e.g.*, measurements including or excluding fruit production, flower production, and herbivory).
 Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), trees were censused down to a minimum diameter breast height (DBH) threshold of 10 cm or less. 
-All records were measured directly or derived from field measurements.
+All records were based on ground-based field measurements.
 
@@
 
-We grouped forests into four broad biome types based on climate zones and dominant vegetation type (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
-Climate zones (Fig. `r figure_map`) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
+We grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
+The climate component of the biome definitions (Fig. `r figure_map`) was based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
 We defined the tropical biome as including all equatorial (A) zones, temperate biomes as including all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and the boreal biome as including the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd).
-Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis.
-We defined leaf type (broadleaf / needleleaf) based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
+Forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis.
+We distinguished broadleaf and needleleaf forests based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
 For young tropical forests imported from *GROA* but not yet classified by leaf type, we assumed that all were broadleaf, consistent with the rarity of naturally regenerating needleleaf forests in the tropics.
-We also classified forests as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication).
-Assigning stands to these groupings required the exclusion of records for which *ForC* lacked geographic coordinates (`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database) or records of stand age (`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
+We classified forests as “young”  if stand age was less than 100 years,  or “mature” if stand age was older or if they were described as “mature”, “old growth”, “intact”, or “undisturbed” in the original publication.
+Assigning stands to these groupings required the exclusion of records for which *ForC* lacked geographic coordinates (`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database) or records of stand age or forest maturity (`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
 We also excluded records of stand age = 0 year (`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).

### 2021-02-26 manuscript/ERL_global_C_SI.tex
working on Helene's comments
#70

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Supplement for: Carbon cycling in mature and regrowth forests globally},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{float} \usepackage{caption} \captionsetup[table]{font=footnotesize} \captionsetup[figure]{font=footnotesize} \captionsetup[figure]{labelformat=empty} \captionsetup[table]{labelformat=empty} \usepackage{pdflscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}}
\usepackage{setspace}\onehalfspacing
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Supplement for: Carbon cycling in mature and regrowth forests
globally}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\raggedright

\newpage

\hypertarget{appendix-s1.-duplicates-and-conflicting-records-within-forc}{%
\subsection{Appendix S1. Duplicates and Conflicting Records within
ForC}\label{appendix-s1.-duplicates-and-conflicting-records-within-forc}}

\hypertarget{status-of-duplicates-and-conflicting-records-within-forc}{%
\subsubsection{Status of duplicates and conflicting records within
ForC}\label{status-of-duplicates-and-conflicting-records-within-forc}}

\emph{ForC} v3.0 contains potential duplicates that have arisen from
importing intermediary data sets with overlap in data. In particular,
the recent imports of \emph{SRDB} and \emph{GROA} resulted in the
introduction of numerous potential duplicates. \emph{SRDB} and
\emph{ForC} were developed independently (but with different foci),
resulting in numerous duplicates. The \emph{GROA} database was developed
semi-independently, incorporating \emph{ForC} data partway through its
development. For records retaining the same site names, \emph{GROA}
duplicates were not imported into \emph{ForC} (although they were
checked against corresponding \emph{ForC} records, and differences
reconciled). There were, however, some records entered into \emph{GROA}
independently of \emph{ForC}, resulting in mismatched names.

Prior to the merging of the three databases (\emph{ForC} v2.0,
\emph{SRDB}, and \emph{GROA}), there were potential duplicates within
the \emph{ForC} and \emph{SRDB} portions, the latter an artifact of how
data were recorded in \emph{SRDB} and imported into \emph{ForC.}
\emph{GROA} has no known duplicates, with several steps having been
taken to ensure independence of the records within the database.

To the extent possible, we used automated scripts in R to detect and
remove duplicate records in the creation of \emph{ForC-simplified}.

\hypertarget{handling-within-plot-duplicates}{%
\subsubsection{Handling within-plot
duplicates}\label{handling-within-plot-duplicates}}

Duplicate records within a plot and taken in the same year were
reconciled as follows. Replicate measurements (\emph{i.e.}, replicates
from within a single study) were averaged. Records that subsumed
others---- \emph{i.e.}, the time period included that of \(\ge\) 2 other
records or dates were unknown and therefore conflicted with \(\ge\) 2
other records---were removed. For each group of duplicate
records---\emph{i.e.}, measurements of the same variable in the same
plot at the same time---one record was assigned precedence as follows.
If duplicates were exact matches, one was dropped. When measurement
periods overlapped or were not specified, precedence was given first to
records representing longer measurement periods (\emph{i.e.}, multi-year
mean over a single year) and then to more recently published values. We
manually reviewed duplicates that differed only in methodology,
assigning precedence to the record employing a more comprehensive
approach (\emph{e.g.}, inclusion of understory, lianas, or bamboo as
opposed to just trees) or using a favored methodology.

\hypertarget{handling-duplicate-sites}{%
\subsubsection{Handling duplicate
sites}\label{handling-duplicate-sites}}

To handle cases where duplicate records had different site and/or plot
names (and often slightly different geographic coordinates), we detected
potential duplicates and, for those that had not yet been resolved,
included only one in \emph{ForC-simplified} and the analyses that were
based on it. Potential duplicates were defined as records for the same
variable made in the same year (when known) on stands of similar age and
in the same geographic area (see Methods). Because records of stand age
often lack precision and were sometimes recorded differently in
different intermediary databases, we considered records to be potential
duplicates when ages fell within one of several broad age groups:
\textless10, 10-50, 50-100, and \textgreater100 years (or mature and of
undefined age). We then generated groups of potential duplicates grouped
by variable, geographic area, stand age grouping, and measurement year.

Within each group, we flagged records as potential duplicates--and
excluded them from \emph{ForC-simplified} until they can be
reconciled--as follows. Because \emph{GROA} is believed to contain no
duplicates, all records originating from \emph{GROA} were assumed
independent, and any other records within the group flagged as potential
duplicates. If all records originated from \emph{ForC v2.0} or
\emph{SRDB}, we gave precedence to \emph{ForC} records, flagging
\emph{SRDB} records as potential duplicates. If duplicates remained
(within \emph{ForC v2.0} or \emph{SRDB}), we selected a record for
retention based on citation year, keeping the oldest (presumably
original) citations for most variables. For eddy-covariance variables
(\(NEP\), \(GPP\), \(R_{eco}\)), we retained the record associated with
the most recent publication (most often FLUXNET2015), as these data are
commonly re-assessed using new analysis methods. If publication years
were the same, we randomly selected one record.

This approach was conservative in that it removed some legitimately
independent records. Future refinement of this process and manual review
of potential duplicates will be needed to include
legitimately-independent records in ForC-simplified and in analyses.

\newpage

\hypertarget{table-s1.-numbers-of-records-by-biome-and-age-class}{%
\subsection{Table S1. Numbers of records by biome and age
class}\label{table-s1.-numbers-of-records-by-biome-and-age-class}}

\begin{table}
\centering\begingroup\fontsize{10}{12}\selectfont

\begin{tabular}{lrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{n records} \\
\cmidrule(l{3pt}r{3pt}){2-3}
Biome & Mature & Young\\
\midrule
\addlinespace[1em]
\multicolumn{3}{l}{\textbf{Boreal climate zones}}\\
\hspace{1em}Boreal broadleaf & 3 & 326\\
\textbf{\hspace{1em}Boreal conifer} & \textbf{491} & \textbf{919}\\
\hspace{1em}Boreal Other & 46 & 151\\
\addlinespace[1em]
\multicolumn{3}{l}{\textbf{Excluded climate zones}}\\
\hspace{1em}Other broadleaf & 6 & 111\\
\hspace{1em}Other conifer & 30 & 45\\
\hspace{1em}Other Other & 6 & 29\\
\addlinespace[1em]
\multicolumn{3}{l}{\textbf{Temperate climate zones}}\\
\textbf{\hspace{1em}Temperate broadleaf} & \textbf{650} & \textbf{3159}\\
\textbf{\hspace{1em}Temperate conifer} & \textbf{563} & \textbf{2730}\\
\hspace{1em}Temperate Other & 159 & 2363\\
\addlinespace[1em]
\multicolumn{3}{l}{\textbf{Tropical climate zones}}\\
\textbf{\hspace{1em}Tropical broadleaf} & \textbf{835} & \textbf{2576}\\
\hspace{1em}Tropical conifer & 3 & 1\\
\hspace{1em}Tropical Other & 8 & 0\\
\bottomrule
\end{tabular}
\endgroup{}
\end{table}

Sample sizes refer to data set after merging of duplicates, removal of
stands with no age or history data, and removal of managed and disturbed
stands. For vegetation type, ``Other'' refers to stands that are mixed
broadleaf/ conifer or that have not been classified. Focal biomes are
indicated in bold.

\newpage

\hypertarget{table-s2.-model-parameter-estimates-for-age-trends-and-biome-differences-in-young-forests}{%
\subsection{Table S2. Model parameter estimates for age trends and biome
differences in young
forests}\label{table-s2.-model-parameter-estimates-for-age-trends-and-biome-differences-in-young-forests}}

\begingroup\fontsize{7}{9}\selectfont

\begin{longtable}{lllll}
\toprule
Variable & Parameter & Estimate & SE & $t_{value}$\\
\midrule
\endfirsthead
\multicolumn{5}{@{}l}{\textit{(continued)}}\\
\toprule
Variable & Parameter & Estimate & SE & $t_{value}$\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$NEP$ & log10(stand.age) & 0.58 & 0.76 & 0.77\\
\hspace{1em} & BiomeTemperate broadleaf & 1.33 & 1.26 & 1.05\\
\hspace{1em} & BiomeTemperate conifer & -2.52 & 0.7 & -3.59\\
\hspace{1em} & BiomeBoreal conifer & 0.2 & 1.05 & 0.19\\
\hspace{1em} & log10(stand.age):BiomeTemperate conifer & 3.35 & 0.88 & 3.83\\
\hspace{1em} & log10(stand.age):BiomeBoreal conifer & -0.16 & 0.97 & -0.16\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$GPP$ & log10(stand.age) & 1.99 & 2.31 & 0.86\\
\hspace{1em} & BiomeTemperate broadleaf & 11.76 & 3.86 & 3.05\\
\hspace{1em} & BiomeTemperate conifer & 11.15 & 2.08 & 5.35\\
\hspace{1em} & BiomeBoreal conifer & 5.23 & 3.96 & 1.32\\
\hspace{1em} & log10(stand.age):BiomeTemperate conifer & 1.44 & 2.69 & 0.54\\
\hspace{1em} & log10(stand.age):BiomeBoreal conifer & 0.13 & 3.31 & 0.04\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$NPP$ & log10(stand.age) & -0.27 & 0.68 & -0.4\\
\hspace{1em} & BiomeTropical broadleaf & 12.05 & 1.83 & 6.59\\
\hspace{1em} & BiomeTemperate broadleaf & 7.29 & 1.31 & 5.58\\
\hspace{1em} & BiomeTemperate conifer & 5.63 & 1.33 & 4.22\\
\hspace{1em} & BiomeBoreal conifer & 4.34 & 1.56 & 2.79\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$ANPP$ & log10(stand.age) & 6.48 & 0.95 & 6.83\\
\hspace{1em} & BiomeTropical broadleaf & -2.56 & 1.54 & -1.66\\
\hspace{1em} & BiomeTemperate broadleaf & 4.86 & 1.32 & 3.69\\
\hspace{1em} & BiomeTemperate conifer & 1.37 & 1.18 & 1.16\\
\hspace{1em} & BiomeBoreal conifer & 2.93 & 2.75 & 1.06\\
\hspace{1em} & log10(stand.age):BiomeTemperate broadleaf & -6.76 & 1.24 & -5.45\\
\hspace{1em} & log10(stand.age):BiomeTemperate conifer & -5.29 & 1.24 & -4.28\\
\hspace{1em} & log10(stand.age):BiomeBoreal conifer & -6.8 & 1.91 & -3.55\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$ANPP_{woody}$ & log10(stand.age) & 2.31 & 0.68 & 3.4\\
\hspace{1em} & BiomeTemperate broadleaf & -1.21 & 1.28 & -0.94\\
\hspace{1em} & BiomeTemperate conifer & -0.76 & 0.91 & -0.83\\
\hspace{1em} & BiomeBoreal conifer & -2.34 & 1.21 & -1.94\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$ANPP_{stem}$ & log10(stand.age) & 0.06 & 0.46 & 0.14\\
\hspace{1em} & BiomeTropical broadleaf & 2.25 & 0.96 & 2.33\\
\hspace{1em} & BiomeTemperate broadleaf & 2.68 & 0.72 & 3.73\\
\hspace{1em} & BiomeTemperate conifer & 1.93 & 0.71 & 2.7\\
\hspace{1em} & BiomeBoreal conifer & 0.72 & 0.92 & 0.78\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$ANPP_{branch}$ & log10(stand.age) & 0.24 & 0.23 & 1.08\\
\hspace{1em} & BiomeTemperate broadleaf & 0.18 & 0.45 & 0.4\\
\hspace{1em} & BiomeTemperate conifer & -0.01 & 0.35 & -0.03\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$ANPP_{foliage}$ & log10(stand.age) & 1.66 & 0.2 & 8.12\\
\hspace{1em} & BiomeTropical broadleaf & 0.24 & 0.6 & 0.41\\
\hspace{1em} & BiomeTemperate broadleaf & -0.61 & 0.4 & -1.54\\
\hspace{1em} & BiomeTemperate conifer & -1.26 & 0.39 & -3.23\\
\hspace{1em} & BiomeBoreal conifer & -2 & 0.46 & -4.34\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$ANPP_{litterfall}$ & log10(stand.age) & 0.74 & 1.84 & 0.4\\
\hspace{1em} & BiomeTropical broadleaf & 2.97 & 2.88 & 1.03\\
\hspace{1em} & BiomeTemperate broadleaf & 0.68 & 3.34 & 0.2\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$ANPP_{repro}$ & - & - & - & -\\
\hspace{1em}$ANPP_{folivory}$ & - & - & - & -\\
\hspace{1em}$M_{woody}$ & - & - & - & -\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$BNPP$ & log10(stand.age) & 0.97 & 0.39 & 2.5\\
\hspace{1em} & BiomeTropical broadleaf & 2.85 & 0.88 & 3.24\\
\hspace{1em} & BiomeTemperate broadleaf & 0.66 & 0.7 & 0.94\\
\hspace{1em} & BiomeTemperate conifer & 0.67 & 0.65 & 1.03\\
\hspace{1em} & BiomeBoreal conifer & -0.16 & 0.8 & -0.2\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$BNPP_{coarse}$ & log10(stand.age) & 0.04 & 0.09 & 0.41\\
\hspace{1em} & BiomeTemperate broadleaf & 0.42 & 0.2 & 2.14\\
\hspace{1em} & BiomeTemperate conifer & 0.52 & 0.16 & 3.17\\
\hspace{1em} & BiomeBoreal conifer & 0.26 & 0.21 & 1.22\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$BNPP_{fine}$ & log10(stand.age) & 1.46 & 0.32 & 4.56\\
\hspace{1em} & BiomeTropical broadleaf & 2.6 & 0.61 & 4.29\\
\hspace{1em} & BiomeTemperate broadleaf & -0.87 & 0.54 & -1.62\\
\hspace{1em} & BiomeTemperate conifer & -0.14 & 0.5 & -0.29\\
\hspace{1em} & BiomeBoreal conifer & -1.05 & 0.63 & -1.66\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$R_{eco}$ & log10(stand.age) & 0.91 & 0.6 & 1.52\\
\hspace{1em} & BiomeTropical broadleaf & 22.74 & 2.28 & 9.95\\
\hspace{1em} & BiomeTemperate broadleaf & 10.5 & 1.36 & 7.71\\
\hspace{1em} & BiomeTemperate conifer & 11.08 & 1.17 & 9.5\\
\hspace{1em} & BiomeBoreal conifer & 6.02 & 1.56 & 3.87\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$R_{auto}$ & - & - & - & -\\
\hspace{1em}$R_{auto-ag}$ & - & - & - & -\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$R_{root}$ & log10(stand.age) & 1.04 & 0.5 & 2.07\\
\hspace{1em} & BiomeTropical broadleaf & 6.31 & 1.01 & 6.27\\
\hspace{1em} & BiomeTemperate broadleaf & 2 & 0.89 & 2.25\\
\hspace{1em} & BiomeTemperate conifer & 2.39 & 0.86 & 2.79\\
\hspace{1em} & BiomeBoreal conifer & -0.15 & 1.2 & -0.13\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$R_{soil}$ & log10(stand.age) & 0.31 & 0.33 & 0.93\\
\hspace{1em} & BiomeTropical broadleaf & 12.55 & 0.85 & 14.83\\
\hspace{1em} & BiomeTemperate broadleaf & 8.99 & 0.64 & 13.99\\
\hspace{1em} & BiomeTemperate conifer & 7.96 & 0.6 & 13.26\\
\hspace{1em} & BiomeBoreal conifer & 4.98 & 1.03 & 4.81\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$R_{het-soil}$ & log10(stand.age) & 0.18 & 0.42 & 0.42\\
\hspace{1em} & BiomeTropical broadleaf & 5.88 & 0.75 & 7.84\\
\hspace{1em} & BiomeTemperate broadleaf & 4.15 & 0.74 & 5.64\\
\hspace{1em} & BiomeTemperate conifer & 3.88 & 0.68 & 5.72\\
\hspace{1em} & BiomeBoreal conifer & 3.29 & 0.94 & 3.49\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$B_{tot}$ & log10(stand.age) & 53.84 & 12.49 & 4.31\\
\hspace{1em} & BiomeTropical broadleaf & -8.52 & 16.39 & -0.52\\
\hspace{1em} & BiomeTemperate broadleaf & -42.69 & 33.86 & -1.26\\
\hspace{1em} & BiomeTemperate conifer & -33.81 & 23.78 & -1.42\\
\hspace{1em} & BiomeBoreal conifer & -70.4 & 88.23 & -0.8\\
\hspace{1em} & log10(stand.age):BiomeTemperate broadleaf & 27.83 & 23 & 1.21\\
\hspace{1em} & log10(stand.age):BiomeTemperate conifer & 22.27 & 19.25 & 1.16\\
\hspace{1em} & log10(stand.age):BiomeBoreal conifer & 17.33 & 50.55 & 0.34\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$B_{ag}$ & log10(stand.age) & 56.94 & 1.89 & 30.11\\
\hspace{1em} & BiomeTropical broadleaf & -16.72 & 3.4 & -4.92\\
\hspace{1em} & BiomeTemperate broadleaf & -59.33 & 5.87 & -10.11\\
\hspace{1em} & BiomeTemperate conifer & -42.82 & 5.24 & -8.17\\
\hspace{1em} & BiomeBoreal conifer & -39.58 & 13.99 & -2.83\\
\hspace{1em} & log10(stand.age):BiomeTemperate broadleaf & 13.16 & 4.04 & 3.26\\
\hspace{1em} & log10(stand.age):BiomeTemperate conifer & -1.5 & 3.69 & -0.41\\
\hspace{1em} & log10(stand.age):BiomeBoreal conifer & -10.67 & 8.09 & -1.32\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$B_{ag-wood}$ & log10(stand.age) & 43.25 & 28.16 & 1.54\\
\hspace{1em} & BiomeTropical broadleaf & -11.68 & 37.64 & -0.31\\
\hspace{1em} & BiomeTemperate broadleaf & -39.81 & 40.45 & -0.98\\
\hspace{1em} & BiomeTemperate conifer & -138.39 & 42.31 & -3.27\\
\hspace{1em} & BiomeBoreal conifer & -74.75 & 154.24 & -0.48\\
\hspace{1em} & log10(stand.age):BiomeTemperate broadleaf & 29.83 & 38.33 & 0.78\\
\hspace{1em} & log10(stand.age):BiomeTemperate conifer & 100.06 & 40.6 & 2.46\\
\hspace{1em} & log10(stand.age):BiomeBoreal conifer & 23.72 & 93.18 & 0.25\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$B_{foliage}$ & log10(stand.age) & 2.73 & 0.57 & 4.76\\
\hspace{1em} & BiomeTropical broadleaf & 0.65 & 0.86 & 0.76\\
\hspace{1em} & BiomeTemperate broadleaf & 0.93 & 0.8 & 1.16\\
\hspace{1em} & BiomeTemperate conifer & -4.34 & 1.57 & -2.77\\
\hspace{1em} & BiomeBoreal conifer & -1.29 & 4.42 & -0.29\\
\hspace{1em} & log10(stand.age):BiomeTemperate broadleaf & -2.04 & 0.72 & -2.83\\
\hspace{1em} & log10(stand.age):BiomeTemperate conifer & 3.55 & 1.17 & 3.03\\
\hspace{1em} & log10(stand.age):BiomeBoreal conifer & -0.61 & 2.57 & -0.24\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$B_{root}$ & log10(stand.age) & 10.34 & 0.93 & 11.14\\
\hspace{1em} & BiomeTropical broadleaf & -1.47 & 1.53 & -0.96\\
\hspace{1em} & BiomeTemperate broadleaf & -8.2 & 1.69 & -4.84\\
\hspace{1em} & BiomeTemperate conifer & -9.72 & 1.59 & -6.13\\
\hspace{1em} & BiomeBoreal conifer & -5.86 & 3.98 & -1.47\\
\hspace{1em} & log10(stand.age):BiomeTemperate broadleaf & 2.04 & 1.4 & 1.45\\
\hspace{1em} & log10(stand.age):BiomeTemperate conifer & 0.81 & 1.36 & 0.59\\
\hspace{1em} & log10(stand.age):BiomeBoreal conifer & -0.19 & 2.49 & -0.08\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$B_{root-coarse}$ & log10(stand.age) & 5.78 & 1.01 & 5.7\\
\hspace{1em} & BiomeTropical broadleaf & -0.5 & 4.71 & -0.11\\
\hspace{1em} & BiomeTemperate broadleaf & 5.65 & 3.07 & 1.84\\
\hspace{1em} & BiomeTemperate conifer & -16.45 & 8.23 & -2\\
\hspace{1em} & BiomeBoreal conifer & -61.67 & 8.54 & -7.22\\
\hspace{1em} & log10(stand.age):BiomeTemperate broadleaf & -5.56 & 1.64 & -3.39\\
\hspace{1em} & log10(stand.age):BiomeTemperate conifer & 15.29 & 5.89 & 2.6\\
\hspace{1em} & log10(stand.age):BiomeBoreal conifer & 33.3 & 4.28 & 7.79\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$B_{root-fine}$ & log10(stand.age) & 0.96 & 0.37 & 2.62\\
\hspace{1em} & BiomeTropical broadleaf & 1.37 & 0.67 & 2.05\\
\hspace{1em} & BiomeTemperate broadleaf & 2.24 & 0.5 & 4.46\\
\hspace{1em} & BiomeTemperate conifer & 5.94 & 0.49 & 12.11\\
\hspace{1em} & BiomeBoreal conifer & 0.51 & 2.29 & 0.22\\
\hspace{1em} & log10(stand.age):BiomeTemperate broadleaf & -0.81 & 0.45 & -1.81\\
\hspace{1em} & log10(stand.age):BiomeTemperate conifer & -3.27 & 0.42 & -7.75\\
\hspace{1em} & log10(stand.age):BiomeBoreal conifer & 0.14 & 1.32 & 0.11\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$DW_{tot}$ & log10(stand.age) & 2.6 & 2.86 & 0.91\\
\hspace{1em} & BiomeTropical broadleaf & 2.1 & 4.04 & 0.52\\
\hspace{1em} & BiomeTemperate broadleaf & 18.37 & 9.48 & 1.94\\
\hspace{1em} & log10(stand.age):BiomeTemperate broadleaf & -1.49 & 5.38 & -0.28\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$DW_{standing}$ & - & - & - & -\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$DW_{down}$ & log10(stand.age) & 2.87 & 1.72 & 1.67\\
\hspace{1em} & BiomeTropical broadleaf & 4.26 & 3.59 & 1.19\\
\hspace{1em} & BiomeTemperate broadleaf & 15.46 & 12.05 & 1.28\\
\hspace{1em} & BiomeTemperate conifer & 45.37 & 6.7 & 6.77\\
\hspace{1em} & BiomeBoreal conifer & 45.22 & 11.02 & 4.1\\
\hspace{1em} & log10(stand.age):BiomeTemperate broadleaf & -9.13 & 7.81 & -1.17\\
\hspace{1em} & log10(stand.age):BiomeTemperate conifer & -13.93 & 3.46 & -4.03\\
\hspace{1em} & log10(stand.age):BiomeBoreal conifer & -22.37 & 7.06 & -3.17\\
\addlinespace[1em]
\multicolumn{4}{l}{\textbf{}}\\
\hspace{1em}$OL$ & log10(stand.age) & 0.46 & 0.84 & 0.55\\
\hspace{1em} & BiomeTropical broadleaf & 3.5 & 2.32 & 1.51\\
\hspace{1em} & BiomeTemperate broadleaf & 20.09 & 3.53 & 5.69\\
\hspace{1em} & BiomeTemperate conifer & -1.69 & 3.05 & -0.55\\
\hspace{1em} & BiomeBoreal conifer & 28.53 & 6.53 & 4.37\\
\hspace{1em} & log10(stand.age):BiomeTemperate broadleaf & -4.84 & 2.24 & -2.16\\
\hspace{1em} & log10(stand.age):BiomeTemperate conifer & 6.56 & 1.73 & 3.79\\
\hspace{1em} & log10(stand.age):BiomeBoreal conifer & 4.77 & 4.14 & 1.15\\*
\end{longtable}
\endgroup{}

\begin{landscape}

\hypertarget{figure-s1.-c-cycle-diagram-for-young-tropical-broadleaf-forests.}{%
\subsection{Figure S1. C cycle diagram for young tropical broadleaf
forests.}\label{figure-s1.-c-cycle-diagram-for-young-tropical-broadleaf-forests.}}

\begin{figure}[H]

{\centering \includegraphics[width=0.84\linewidth]{tables_figures/C_cycle_diagrams/Tropical_broadleaf_YOUNG}

}

\caption{C cycle diagram for young tropical broadleaf forests. Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1. Presented are observed ranges, where geographically distinct areas are treated as the unit of replication. Note that variables differ in geographical and age representation, resulting in potential imbalances (Figs. S5-S30). All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks). Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data. To illustrate the magnitude of different fluxes, arrow width is proportional to the square root of the mean of the corresponding flux.}\label{fig:unnamed-chunk-4}
\end{figure}

\hypertarget{figure-s2.-c-cycle-diagram-for-young-temperate-broadleaf-forests.}{%
\subsection{Figure S2. C cycle diagram for young temperate broadleaf
forests.}\label{figure-s2.-c-cycle-diagram-for-young-temperate-broadleaf-forests.}}

\begin{figure}[H]

{\centering \includegraphics[width=0.84\linewidth]{tables_figures/C_cycle_diagrams/Temperate_broadleaf_YOUNG}

}

\caption{C cycle diagram for young temperate broadleaf forests. Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1. Presented are observed ranges, where geographically distinct areas are treated as the unit of replication. Note that variables differ in geographical and age representation, resulting in potential imbalances (Figs. S5-S30). All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks). Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data. To illustrate the magnitude of different fluxes, arrow width is proportional to the square root of the mean of the corresponding flux.}\label{fig:unnamed-chunk-5}
\end{figure}

\hypertarget{figure-s3.-c-cycle-diagram-for-young-temperate-conifer-forests.}{%
\subsection{Figure S3. C cycle diagram for young temperate conifer
forests.}\label{figure-s3.-c-cycle-diagram-for-young-temperate-conifer-forests.}}

\begin{figure}[H]

{\centering \includegraphics[width=0.84\linewidth]{tables_figures/C_cycle_diagrams/Temperate_conifer_YOUNG}

}

\caption{ C cycle diagram for young temperate conifer forests. Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1. Presented are observed ranges, where geographically distinct areas are treated as the unit of replication. Note that variables differ in geographical and age representation, resulting in potential imbalances (Figs. S5-S30). All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks). Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data. To illustrate the magnitude of different fluxes, arrow width is proportional to the square root of the mean of the corresponding flux.}\label{fig:unnamed-chunk-6}
\end{figure}

\hypertarget{figure-s4.-c-cycle-diagram-for-young-boreal-conifer-forests.}{%
\subsection{Figure S4. C cycle diagram for young boreal conifer
forests.}\label{figure-s4.-c-cycle-diagram-for-young-boreal-conifer-forests.}}

\begin{figure}[H]

{\centering \includegraphics[width=0.84\linewidth]{tables_figures/C_cycle_diagrams/Boreal_conifer_YOUNG}

}

\caption{C cycle diagram for young boreal conifer forests. Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1. Presented are observed ranges, where geographically distinct areas are treated as the unit of replication. Note that variables differ in geographical and age representation, resulting in potential imbalances (Figs. S5-S30). All units are Mg C ha$^{-1}$ yr$^{-1}$ (fluxes) or Mg C ha$^{-1}$ (stocks). Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data. To illustrate the magnitude of different fluxes, arrow width is proportional to the square root of the mean of the corresponding flux.}\label{fig:unnamed-chunk-7}
\end{figure}
\end{landscape}

\newpage

\hypertarget{figure-s5.-age-trends-and-biome-differences-for-nep}{%
\subsection{\texorpdfstring{Figure S5. Age trends and biome differences
for
\(NEP\)}{Figure S5. Age trends and biome differences for NEP}}\label{figure-s5.-age-trends-and-biome-differences-for-nep}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/NEP_with_map}

}

\caption{Age trends and biome differences for $NEP$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-8}
\end{figure}

\newpage

\hypertarget{figure-s6.-age-trends-and-biome-differences-for-gpp}{%
\subsection{\texorpdfstring{Figure S6. Age trends and biome differences
for
\(GPP\)}{Figure S6. Age trends and biome differences for GPP}}\label{figure-s6.-age-trends-and-biome-differences-for-gpp}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/GPP_with_map}

}

\caption{Age trends and biome differences for $GPP$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-9}
\end{figure}

\newpage

\hypertarget{figure-s7.-age-trends-and-biome-differences-for-npp}{%
\subsection{\texorpdfstring{Figure S7. Age trends and biome differences
for
\(NPP\)}{Figure S7. Age trends and biome differences for NPP}}\label{figure-s7.-age-trends-and-biome-differences-for-npp}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/NPP_with_map}

}

\caption{Age trends and biome differences for $NPP$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-10}
\end{figure}

\newpage

\hypertarget{figure-s8.-age-trends-and-biome-differences-for-anpp}{%
\subsection{\texorpdfstring{Figure S8. Age trends and biome differences
for
\(ANPP\)}{Figure S8. Age trends and biome differences for ANPP}}\label{figure-s8.-age-trends-and-biome-differences-for-anpp}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/ANPP_with_map}

}

\caption{Age trends and biome differences for $ANPP$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-11}
\end{figure}

\newpage

\hypertarget{figure-s9.-age-trends-and-biome-differences-for-anpp_woody}{%
\subsection{\texorpdfstring{Figure S9. Age trends and biome differences
for
\(ANPP_{woody}\)}{Figure S9. Age trends and biome differences for ANPP\_\{woody\}}}\label{figure-s9.-age-trends-and-biome-differences-for-anpp_woody}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/ANPP_woody_with_map}

}

\caption{Age trends and biome differences for $ANPP_{woody}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-12}
\end{figure}

\newpage

\hypertarget{figure-s10.-age-trends-and-biome-differences-for-anpp_stem}{%
\subsection{\texorpdfstring{Figure S10. Age trends and biome differences
for
\(ANPP_{stem}\)}{Figure S10. Age trends and biome differences for ANPP\_\{stem\}}}\label{figure-s10.-age-trends-and-biome-differences-for-anpp_stem}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/ANPP_stem_with_map}

}

\caption{Age trends and biome differences for $ANPP_{stem}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-13}
\end{figure}

\newpage

\hypertarget{figure-s11.-age-trends-and-biome-differences-for-anpp_foliage}{%
\subsection{\texorpdfstring{Figure S11. Age trends and biome differences
for
\(ANPP_{foliage}\)}{Figure S11. Age trends and biome differences for ANPP\_\{foliage\}}}\label{figure-s11.-age-trends-and-biome-differences-for-anpp_foliage}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/ANPP_foliage_with_map}

}

\caption{Age trends and biome differences for $ANPP_{foliage}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-14}
\end{figure}

\newpage

\hypertarget{figure-s12.-age-trends-and-biome-differences-for-anpp_litterfall}{%
\subsection{\texorpdfstring{Figure S12. Age trends and biome differences
for
\(ANPP_{litterfall}\)}{Figure S12. Age trends and biome differences for ANPP\_\{litterfall\}}}\label{figure-s12.-age-trends-and-biome-differences-for-anpp_litterfall}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/ANPP_litterfall_with_map}

}

\caption{Age trends and biome differences for $ANPP_{litterfall}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-15}
\end{figure}

\newpage

\hypertarget{figure-s13.-age-trends-and-biome-differences-for-bnpp}{%
\subsection{\texorpdfstring{Figure S13. Age trends and biome differences
for
\(BNPP\)}{Figure S13. Age trends and biome differences for BNPP}}\label{figure-s13.-age-trends-and-biome-differences-for-bnpp}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/BNPP_with_map}

}

\caption{Age trends and biome differences for $BNPP$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-16}
\end{figure}

\newpage

\hypertarget{figure-s14.-age-trends-and-biome-differences-for-bnpp_coarse}{%
\subsection{\texorpdfstring{Figure S14. Age trends and biome differences
for
\(BNPP_{coarse}\)}{Figure S14. Age trends and biome differences for BNPP\_\{coarse\}}}\label{figure-s14.-age-trends-and-biome-differences-for-bnpp_coarse}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/BNPP_coarse_with_map}

}

\caption{Age trends and biome differences for $BNPP_{coarse}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-17}
\end{figure}

\newpage

\hypertarget{figure-s15.-age-trends-and-biome-differences-for-bnpp_fine}{%
\subsection{\texorpdfstring{Figure S15. Age trends and biome differences
for
\(BNPP_{fine}\)}{Figure S15. Age trends and biome differences for BNPP\_\{fine\}}}\label{figure-s15.-age-trends-and-biome-differences-for-bnpp_fine}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/BNPP_fine_with_map}

}

\caption{Age trends and biome differences for $BNPP_{fine}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-18}
\end{figure}

\newpage

\hypertarget{figure-s16.-age-trends-and-biome-differences-for-r_eco}{%
\subsection{\texorpdfstring{Figure S16. Age trends and biome differences
for
\(R_{eco}\)}{Figure S16. Age trends and biome differences for R\_\{eco\}}}\label{figure-s16.-age-trends-and-biome-differences-for-r_eco}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/R_eco_with_map}

}

\caption{Age trends and biome differences for $R_{eco}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-19}
\end{figure}

\newpage

\hypertarget{figure-s17.-age-trends-and-biome-differences-for-r_root}{%
\subsection{\texorpdfstring{Figure S17. Age trends and biome differences
for
\(R_{root}\)}{Figure S17. Age trends and biome differences for R\_\{root\}}}\label{figure-s17.-age-trends-and-biome-differences-for-r_root}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/R_root_with_map}

}

\caption{Age trends and biome differences for $R_{root}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-20}
\end{figure}

\newpage

\hypertarget{figure-s18.-age-trends-and-biome-differences-for-r_soil}{%
\subsection{\texorpdfstring{Figure S18. Age trends and biome differences
for
\(R_{soil}\)}{Figure S18. Age trends and biome differences for R\_\{soil\}}}\label{figure-s18.-age-trends-and-biome-differences-for-r_soil}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/R_soil_with_map}

}

\caption{Age trends and biome differences for $R_{soil}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-21}
\end{figure}

\newpage

\hypertarget{figure-s19.-age-trends-and-biome-differences-for-r_het-soil}{%
\subsection{\texorpdfstring{Figure S19. Age trends and biome differences
for
\(R_{het-soil}\)}{Figure S19. Age trends and biome differences for R\_\{het-soil\}}}\label{figure-s19.-age-trends-and-biome-differences-for-r_het-soil}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/R_het_soil_with_map}

}

\caption{Age trends and biome differences for $R_{het-soil}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-22}
\end{figure}

\newpage

\hypertarget{figure-s20.-age-trends-and-biome-differences-for-b_tot}{%
\subsection{\texorpdfstring{Figure S20. Age trends and biome differences
for
\(B_{tot}\)}{Figure S20. Age trends and biome differences for B\_\{tot\}}}\label{figure-s20.-age-trends-and-biome-differences-for-b_tot}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/biomass_with_map}

}

\caption{Age trends and biome differences for $B_{tot}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-23}
\end{figure}

\newpage

\hypertarget{figure-s21.-age-trends-and-biome-differences-for-b_ag}{%
\subsection{\texorpdfstring{Figure S21. Age trends and biome differences
for
\(B_{ag}\)}{Figure S21. Age trends and biome differences for B\_\{ag\}}}\label{figure-s21.-age-trends-and-biome-differences-for-b_ag}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/biomass_ag_with_map}

}

\caption{Age trends and biome differences for $B_{ag}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-24}
\end{figure}

\newpage

\hypertarget{figure-s22.-age-trends-and-biome-differences-for-b_ag-wood}{%
\subsection{\texorpdfstring{Figure S22. Age trends and biome differences
for
\(B_{ag-wood}\)}{Figure S22. Age trends and biome differences for B\_\{ag-wood\}}}\label{figure-s22.-age-trends-and-biome-differences-for-b_ag-wood}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/biomass_ag_woody_with_map}

}

\caption{Age trends and biome differences for $B_{ag-wood}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-25}
\end{figure}

\newpage

\hypertarget{figure-s23.-age-trends-and-biome-differences-for-b_foliage}{%
\subsection{\texorpdfstring{Figure S23. Age trends and biome differences
for
\(B_{foliage}\)}{Figure S23. Age trends and biome differences for B\_\{foliage\}}}\label{figure-s23.-age-trends-and-biome-differences-for-b_foliage}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/biomass_foliage_with_map}

}

\caption{Age trends and biome differences for $B_{foliage}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-26}
\end{figure}

\newpage

\hypertarget{figure-s24.-age-trends-and-biome-differences-for-b_root}{%
\subsection{\texorpdfstring{Figure S24. Age trends and biome differences
for
\(B_{root}\)}{Figure S24. Age trends and biome differences for B\_\{root\}}}\label{figure-s24.-age-trends-and-biome-differences-for-b_root}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/biomass_root_with_map}

}

\caption{Age trends and biome differences for $B_{root}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-27}
\end{figure}

\newpage

\hypertarget{figure-s25.-age-trends-and-biome-differences-for-b_root-coarse}{%
\subsection{\texorpdfstring{Figure S25. Age trends and biome differences
for
\(B_{root-coarse}\)}{Figure S25. Age trends and biome differences for B\_\{root-coarse\}}}\label{figure-s25.-age-trends-and-biome-differences-for-b_root-coarse}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/biomass_root_coarse_with_map}

}

\caption{Age trends and biome differences for $B_{root-coarse}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-28}
\end{figure}

\newpage

\hypertarget{figure-s26.-age-trends-and-biome-differences-for-b_root-fine}{%
\subsection{\texorpdfstring{Figure S26. Age trends and biome differences
for
\(B_{root-fine}\)}{Figure S26. Age trends and biome differences for B\_\{root-fine\}}}\label{figure-s26.-age-trends-and-biome-differences-for-b_root-fine}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/biomass_root_fine_with_map}

}

\caption{Age trends and biome differences for $B_{root-fine}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-29}
\end{figure}

\newpage

\hypertarget{figure-s27.-age-trends-and-biome-differences-for-dw_tot}{%
\subsection{\texorpdfstring{Figure S27. Age trends and biome differences
for
\(DW_{tot}\)}{Figure S27. Age trends and biome differences for DW\_\{tot\}}}\label{figure-s27.-age-trends-and-biome-differences-for-dw_tot}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/deadwood_with_map}

}

\caption{Age trends and biome differences for $DW_{tot}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-30}
\end{figure}

\newpage

\hypertarget{figure-s28.-age-trends-and-biome-differences-for-dw_standing}{%
\subsection{\texorpdfstring{Figure S28. Age trends and biome differences
for
\(DW_{standing}\)}{Figure S28. Age trends and biome differences for DW\_\{standing\}}}\label{figure-s28.-age-trends-and-biome-differences-for-dw_standing}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/deadwood_standing_with_map}

}

\caption{Age trends and biome differences for $DW_{standing}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-31}
\end{figure}

\newpage

\hypertarget{figure-s29.-age-trends-and-biome-differences-for-dw_down}{%
\subsection{\texorpdfstring{Figure S29. Age trends and biome differences
for
\(DW_{down}\)}{Figure S29. Age trends and biome differences for DW\_\{down\}}}\label{figure-s29.-age-trends-and-biome-differences-for-dw_down}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/deadwood_down_with_map}

}

\caption{Age trends and biome differences for $DW_{down}$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-32}
\end{figure}

\newpage

\hypertarget{figure-s30.-age-trends-and-biome-differences-for-ol}{%
\subsection{\texorpdfstring{Figure S30. Age trends and biome differences
for
\(OL\)}{Figure S30. Age trends and biome differences for OL}}\label{figure-s30.-age-trends-and-biome-differences-for-ol}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/age_trends/organic_layer_with_map}

}

\caption{Age trends and biome differences for $OL$. Map shows data sources ('$\times$' and '$\circ$' indicate young and mature stands, respectively). In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) $\times$ biome interaction. Boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).}\label{fig:unnamed-chunk-33}
\end{figure}

\end{document}

### 2021-02-26 manuscript/ERL_global_C_review.Rmd
working on Helene's comments
#70

@@
 schematics_caption_generic ="Arrows indicate fluxes (Mg C ha$^{-1}$ yr$^{-1}$); boxes indicate stocks (Mg C ha$^{-1}$), with variables as defined in Table 1.
-Presented are mean ± std, where geographically distinct areas are treated as the unit of replication.
+Presented are mean ± std over geographically distinct areas (clusters of plots within 25 km of each other).
 Dashed shape outlines indicate variables with records from <7 distinct geographic areas, and dashed arrows indicate fluxes with no data.
@@
 S5-S30).
-"
+Mature forests are defined  as ≥ 100 years old and with no known major natural or anthropogenic disturbance in that time."
 TrB_mature_caption = paste( "Figure 3 | C cycle diagram for mature tropical broadleaf forests. ", schematics_caption_generic)  
@@
 
-Data were summarized to produce schematics of C cycling for mature forests of each biome.
-To obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot.
-Values were then averaged across geographically distinct areas, defined as plots clustered within 25 km of one another [*sensu* @anderson-teixeira_forc_2018], weighting by area sampled if available for all records.
-This step was taken to avoid pseudo-replication.
+We calculated means and standard deviations of each mature forest C cycle variable by biome over geographically distinct areas to produce biome-specific schematics.
+We first averaged any repeated measurements within a plot.
+To avoid pseudo-replication, we then averaged multiple measurements within geographically distinct areas, defined as plots clustered within 25 km of one another  [*sensu* @anderson-teixeira_forc_2018], weighting by area sampled if available for all records.
+Finally, we took means and standard deviations over geographic areas.
 
@@
 Henceforth, we refer to the variables on the left side of the equation as "aggregate" fluxes or stocks, and those that are summed as "component" fluxes or stocks, noting that the same variable can take both aggregate and component positions in different relationships.
-We considered the C budget for a given relationship "closed" when component variables summed to within one standard deviation of the aggregate variable.
+We considered the C budget for a given relationship “closed” when the means of component variables summed to within one standard deviation of the mean of the aggregate variable.
 
 To test for differences across mature forest biomes, we also examined how stand age impacted fluxes and stocks, employing a mixed effects model ['lmer' function in 'lme4' R package; @bates_fitting_2015] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
-When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
-This analysis was run for variables with records for at least seven distinct geographic areas in more than one biome, excluding any biomes that failed this criteria (Table 1).
+When biome had a significant effect, we used a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
+This analysis was run for variables and biomes with records for at least seven distinct geographic areas per biome (Table 1).
 
 To test for age trends in young (<100yrs) forests, we employed a mixed effects model with biome and log10[stand age] as fixed effects and plot nested within geographic area as a random effect on the intercept.
-This analysis was run for variables with records for at least three distinct geographic areas in more than one biome, excluding any biomes that failed this criteria (Table 1). 
+This analysis was run for variables and biomes with records for at least three distinct geographic areas per biome, excluding any biomes that failed this criteria (Table 1).
 When the effect of stand age was significant at p $\le$ 0.05 and when each biome had records for stands of at least 10 different ages, a biome $\times$ stand.age interaction was included in the model.
@@
 
-Average C cycles for mature tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major natural or anthropogenic disturbance are presented in Figures `r all_diagrams_mature` (and available in tabular format in the *ForC* release accompanying this publication:  `ForC/numbers_and_facts/ForC_variable_averages_per_Biome.csv`).
+Average C cycles for mature tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests are presented in Figures `r all_diagrams_mature` (and available in tabular format in the *ForC* release accompanying this publication:  `ForC/numbers_and_facts/ForC_variable_averages_per_Biome.csv`).
 
@@
 In the temperate conifer biome, the average composite measure of root biomass ($B_{root}$) was less than the combined average value of coarse and fine root biomass ($B_{root-coarse}$ and $B_{root-fine}$, respectively).
-This lack of closure was driven by very high estimates of $B_{root-coarse}$ from high-biomass forests of the US Pacific Northwest (Fig. `r B_root_coarse_fig`).
+This lack of closure was driven by very high estimates of $B_{root-coarse}$ from high-biomass forests of the US Pacific Northwest, a geographic region with a disproportionately large number of records for this variable (Fig.
+`r B_root_coarse_fig`).
 
@@
 
-There were sufficient data to assess differences among biomes in mature forest values for 15 flux variables, and significant differences among biomes were detected for 12 variables (Table 1).
-In all of these cases--including C fluxes into, within, and out of the ecosystem--C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs.
+There were sufficient data to assess differences among biomes in mature forest values for 15 flux variables, and 12 of these variables exhibited statistically significant differences among biomes (Table 1).
+In all cases of significant differences (including C fluxes into, within, and out of the ecosystem), C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs.
 `r flux_age_trends`, `r all_flux_age_trends`).
@@
 Fluxes tended to be numerically greater in temperate broadleaf than temperate conifer forests, but the difference was never statistically significant.
-This pattern held for the following variables: $GPP$, $NPP$, $ANPP$, $ANPP_{stem}$,$ANPP_{branch}$, $ANPP_{foliage}$, $BNPP$, $R_{eco}$, $R_{root}$, $R_{soil}$, and $R_{het-soil}$. 
+This pattern held for 11 of the 12 variables with significant biome effects: $GPP$, $NPP$, $ANPP$, $ANPP_{stem}$,$ANPP_{branch}$, $ANPP_{foliage}$, $BNPP$, $R_{eco}$, $R_{root}$, $R_{soil}$, and $R_{het-soil}$.
 For two of the variables without significant differences among biomes ($ANPP_{litterfall}$ and $BNPP_{fine}$; Figs.
 `r ANPP_litterfall_fig` and `r BNPP_fine_fig`, respectively), the same general trends applied but were not statistically significant.
-Another exception was for $BNPP_{root-coarse}$, where all records came from high-biomass forests in the US Pacific Northwest, resulting in marginally higher values for the temperate conifer biome (Fig.
-`r BNPP_coarse_fig`; differences significant in mixed effects model but not in post-hoc pairwise comparison).
 
@@
 For all biomes, $NEP$ was positive, with 95% confidence intervals excluding zero.
+$BNPP_{root-coarse}$ also exhibited significant differences among biomes with the highest means outside the tropics, in this case in temperate conifer forest, a biome for which all records came from high-biomass forests in the US Pacific Northwest (Fig.
+`r BNPP_coarse_fig`; differences significant in mixed effects model but not in post-hoc pairwise comparison).
 
 ```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 7 | Age trends and biome differences in some of the major C fluxes: (a) $GPP$, (b) $NPP$, (c) $ANPP$, (d) $R_{soil}$, (e) $R_{eco}$, and (f) $NEP$.
-In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
-The boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
-Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).
+The scatterplots show age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
+The fitted lines indicate the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction (interaction effects were tested only if the main age effect was significant and data were available for at least 10 stand ages per biome--i.e., for $GPP$, $ANPP$, $R_{soil}$, and $NEP$).
+The boxplots illustrate variation among biomes in mature forests, with different letters indicating significant differences between biomes.
+Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or tests of differences across biomes (mature forests, indicated by a dash instead of a letter above the box plot).
 Individual figures for each flux with sufficient data, along with maps showing geographic distribution of the data, are given in the Supplement (Figs.
@@
 `r schematic`, `r stock_age_trends`, `r all_stock_age_trends`).
-For variables that were disproportionately sampled in such high-biomass forests ($B_{ag-wood}$, $B_{foliage}$, and $B_{root-coarse}$; disproportionately sampled in  the US Pacific Northwest), temperate conifer forests had significantly higher stocks than the other biomes.
+For variables for which temperate conifer forest records were disproportionately from high-biomass forests in the US Pacific Northwest ($B_{ag-wood}$, $B_{foliage}$, and $B_{root-coarse}$), temperate conifer forests had significantly higher stocks than other biomes.
 
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood.
-In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction.
-The boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes.
-Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests).
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | Age trends and biome differences in some of the major forest C stocks: (a) $B_{ag}$, (b) $B_{foliage}$, (c) $B_{root-fine}$, (d) $DW_{standing}$, (d) $DW_{down}$, and (d) $OL$.
+The scatterplots show age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
+The fitted lines indicate the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction (all variables but $DW_{down}$).
+The boxplots illustrate variation among biomes in mature forests, with different letters indicating significant differences between biomes.
+Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or tests of differences across biomes (mature forests, indicated by a dash instead of a letter above the box plot).
 Individual figures for each stock with sufficient data, along with maps showing geographic distribution of the data, are given in the Supplement (Figs.
@@
 
-Differences in C fluxes across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests (Table 1, Figs.
+Differences among biomes in regrowth forest C fluxes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests (Table 1, Figs.
 `r flux_age_trends`, `r all_age_trends`).
@@
 `r schematic`, `r flux_age_trends`- `r stock_age_trends`, `r age_trends_all`).
-The notable exception was mature forest $NEP$, which, as the difference between $GPP$ and $R_{eco}$, was statistically indistinguishable across biomes (Fig. `r flux_age_trends`f).
+The notable exception was mature forest $NEP$ (the difference between $GPP$ and $R_{eco}$) which was statistically indistinguishable across biomes (Fig. `r flux_age_trends`f).
 There was also little directional variation in mean mature forest C stocks (biomass, dead wood, and organic layer) across biomes, although maximum values for the majority of stocks (all including live or standing woody biomass) occurred in temperate biomes (Figs.
@@
 
-Our analysis reveals that carbon cycling is most rapid in the tropics and slowest in boreal regions, including C fluxes into ($GPP$), within (e.g., $NPP$ and its components), and out of (e.g., $R_{soil}$, $R_{eco}$) the ecosystem.
-For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude -- or increase with temperature -- on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
+Our analysis reveals that carbon cycling is most rapid in the tropics and slowest in boreal regions.
+For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude and increase with temperature on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
 This consistency is not surprising, particularly given commonality in the data analyzed or used for calibration.
-The finding that these patterns hold consistently across across numerous fluxes, while consistent with theoretical expectations (Fig.
+The finding that these patterns hold consistently across numerous fluxes, while consistent with theoretical expectations (Fig.
 `r schematic`), is novel to this analysis [but see @banbury_morgan_global_nodate for nine autotrophic fluxes].
@@
 In contrast to C fluxes and biomass accumulation rates in regrowth forests, stocks showed less systematic variation across biomes (c.f. Fig. `r schematic`).
-For aboveground biomass, which is the variable in *ForC* with broadest geographical representation, the modest trend of declining biomass from tropical to boreal regions mirrors observations from spaceborne lidar that reveal a decline in aboveground biomass (for all forests, including secondary) with latitude across the N hemisphere [@hu_mapping_2016].
-The highest- biomass forests on Earth are, however, found in coastal temperate climates of both the southern and northern hemisphere [Figs.
-`r schematic`, `r stock_age_trends`a; @keith_re-evaluation_2009; @smithwick_potential_2002; @hu_mapping_2016].
+For aboveground biomass, which is the variable in *ForC* with the broadest geographical representation, the modest trend of declining biomass from tropical to boreal regions mirrors observations from spaceborne lidar that reveal a decline in aboveground biomass (for all forests, including secondary) with latitude across the N hemisphere [@hu_mapping_2016].
+The highest-biomass forests on Earth are, however, found in coastal temperate climates of both the southern and northern hemisphere [Figs.
+`r schematic`, `r stock_age_trends`a; @keith_re-evaluation_2009; @smithwick_potential_2002; @hu_mapping_2016; @larjavaara_temperature_2012].
 Disproportionate representation of forests in one such region--the US Pacific Northwest--inflated estimates of temperate conifer fluxes and stocks for some variables and was responsible for all of the anomalous results described here (e.g., lack of complete C budget closure, anomalous trend across biomes for $BNPP_{coarse}$).
-Thus, biome differences should always be interpreted relative to the geographic distribution of sampling, which only rarely covers the majority of forested area within a biome.
+Thus, biome differences should always be interpreted relative to the geographic distribution of sampling, which only rarely adequately represents the majority of forested area within a biome.
 
-Whereas aboveground biomass can be remotely sensed [albeit with significant uncertainties; @ploton_spatial_2020] and receives significant research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
+Whereas aboveground biomass can be remotely sensed [albeit with significant uncertainties; @ploton_spatial_2020] and receives significant research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) carbon across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
 Although these stocks can be important--exceeding 100 Mg C ha^-1^ in some stands (Figs.
@@
 
-Our study reveals that most C fluxes quickly increase and then decelerate as stands age (Figs.
+Our study reveals that most autotrophic C fluxes quickly increase and then decelerate as stands age (Figs.
 `r flux_age_trends`, `r age_trends_all`), consistent with current understanding of age trends in forest C cycling [Fig.

### 2021-02-27 manuscript/ERL_global_C_review.Rmd
fixed reference keys
slightly different reference keys under updated computer system

@@
 Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide [CO~2~; @bonan_forests_2008; @friedlingstein_global_2019; @ipcc_global_2018].
-Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climatecarbon_2006; @krause_large_2018; @bonan_model_2019; @di_vittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011; @ipcc_2019_2019] can be traced to lack of understanding on how C cycling varies across forest types and in relation to stand history.
+Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climate_2006; @krause_large_2018; @bonan_model_2019; @divittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011; @ipcc_2019_2019] can be traced to lack of understanding on how C cycling varies across forest types and in relation to stand history.
 This requires accessible, comprehensive, and large-scale databases with global coverage, which runs contrary to the traditional way forest C stocks and fluxes have been measured and published.
@@
 
-The future of the current forest C sink is dependent both upon forest responses to climate change and human land use decisions, with land use change itself strongly influencing the course of climate change [@friedlingstein_climatecarbon_2006].
+The future of the current forest C sink is dependent both upon forest responses to climate change and human land use decisions, with land use change itself strongly influencing the course of climate change [@friedlingstein_climate_2006].
 Regrowing forests (i.e., secondary forests) will play a particularly important role [@pugh_role_2019], as almost two-thirds of the world’s forests were secondary as of 2010 [@fao_global_2010].
-As anthropogenic and climate-driven disturbances impact a growing proportion of Earth's forests [@andela_human-driven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests is increasingly important [@andersonteixeira_altered_2013].
+As anthropogenic and climate-driven disturbances impact a growing proportion of Earth's forests [@andela_humandriven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests is increasingly important [@anderson-teixeira_altered_2013].
 Although age trends in aboveground biomass have been well-studied and synthesized globally [@cook-patton_mapping_2020], there is a relative dearth of data and synthesis on other C stocks and fluxes in secondary forests.
@@
 
-For more than half a century, researchers have sought to understand how forest carbon cycling varies across stands, including among biomes [e.g., @lieth_primary_1973; @luyssaert_co2_2007] and with stand age [e.g., @odum_strategy_1969; @luyssaert_old-growth_2008].
+For more than half a century, researchers have sought to understand how forest carbon cycling varies across stands, including among biomes [e.g., @lieth_primary_1973; @luyssaert_co2_2007] and with stand age [e.g., @odum_strategy_1969; @luyssaert_oldgrowth_2008].
 Over this time, an increasingly refined conceptual understanding of the elements of ecosystem C cycles has developed, as a growing number of variables have been defined [e.g., @chapin_reconciling_2006], along with appropriate measurement methods [e.g., @clark_measuring_2001].
@@
 Measurement techniques have been increasingly standardized; for example, of the biomass allometries that strongly influence estimates of most C cycle variables [e.g., @chave_improved_2014].
-Further standardization has been made possible through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015; @davies_forestgeo_2021], NEON [@schimel_neon_2007], and FLUXNET [@baldocchi_fluxnet_2001; @novick_ameriflux_2018].
+Further standardization has been made possible through research networks such as ForestGEO [@anderson-teixeira_ctfsforestgeo_2015; @davies_forestgeo_2021], NEON [@schimel_neon_2007], and FLUXNET [@baldocchi_fluxnet_2001; @novick_ameriflux_2018].
 Remote sensing technology has become increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass, $B_{ag}$: @saatchi_benchmark_2011; @hu_mapping_2016; @spawn_harmonized_2020; gross primary productivity, $GPP$: @li_mapping_2019], yet measurement and validation of most forest C stocks and fluxes necessarily requires intensive on-the-ground data collection.
@@
 Global scale data syntheses have shown that C fluxes including $GPP$, net primary productivity ($NPP$), and soil respiration ($R_{soil}$) decrease with latitude or, correspondingly, increase with mean annual temperature [Fig.
-`r schematic`; e.g., @lieth_primary_1973;  @luyssaert_co2_2007; @hursh_sensitivity_2017; @banbury_morgan_global_nodate].
+`r schematic`; e.g., @lieth_primary_1973;  @luyssaert_co2_2007; @hursh_sensitivity_2017; @banburymorgan_global_inpress].
 C stocks of mature forests show less directional variation (Fig. `r schematic`).
-On average, aboveground biomass ($B_{ag}$) tends to decrease with latitude, but not as dramatically as fluxes, and with the highest $B_{ag}$ forests in relatively cool, moist temperate regions [@keith_re-evaluation_2009; @smithwick_potential_2002; @hu_mapping_2016].
+On average, aboveground biomass ($B_{ag}$) tends to decrease with latitude, but not as dramatically as fluxes, and with the highest $B_{ag}$ forests in relatively cool, moist temperate regions [@keith_reevaluation_2009; @smithwick_potential_2002; @hu_mapping_2016].
 In contrast, standing and downed dead wood ($DW_{standing}$ and $DW_{down}$, respectively, summing to $DW_{tot}$) and the organic layer ($OL$) tend to accumulate more in colder climates where decomposition is slow relative to $NPP$ [@harmon_ecology_1986; @allen_global_2002].
@@
 Although the conceptualization of the C cycle in this paper was simplistic by current standards, the paper was foundational in framing the theory around which research on the subject still revolves [@corman_foundations_2019], and the basic framework still holds, albeit with modest modifications [Fig.
-`r schematic`; @andersonteixeira_altered_2013].
+`r schematic`; @anderson-teixeira_altered_2013].
 Following stand-clearing disturbance, $GPP$, $NPP$, and biomass of leaves ($B_{foliage}$) and fine roots ($B_{root-fine}$) initially increase rapidly and thereafter remain relatively stable ($B_{foliage}$, $B_{root-fine}$, sometimes $GPP$) or decline slightly [$NPP$, sometimes $GPP$; e.g., @law_changes_2003; @pregitzer_carbon_2004; @amiro_ecosystem_2010; @goulden_patterns_2011].
 The decline in $NPP$ occurs because $R_{auto}$ increases relative to $GPP$ as forests age, corresponding to declining carbon use efficiency with stand age [@delucia_forest_2007; @collalti_forest_2020].
-Heterotrophic respiration, most of which originates from the soil ($R_{het-soil}$) remains relatively constant with stand age [@law_changes_2003; @pregitzer_carbon_2004; @goulden_patterns_2011], with the result that net ecosystem production ($NEP=GPP-R_{eco}$, where $R_{eco}$ is total ecosystem respiration) is initially negative, increases to a maximum at intermediate ages, and declines--typically to a small positive value--thereafter [@law_changes_2003; @pregitzer_carbon_2004; @amiro_ecosystem_2010; @goulden_patterns_2011; @luyssaert_old-growth_2008].
+Heterotrophic respiration, most of which originates from the soil ($R_{het-soil}$) remains relatively constant with stand age [@law_changes_2003; @pregitzer_carbon_2004; @goulden_patterns_2011], with the result that net ecosystem production ($NEP=GPP-R_{eco}$, where $R_{eco}$ is total ecosystem respiration) is initially negative, increases to a maximum at intermediate ages, and declines--typically to a small positive value--thereafter [@law_changes_2003; @pregitzer_carbon_2004; @amiro_ecosystem_2010; @goulden_patterns_2011; @luyssaert_oldgrowth_2008].
 The result is that biomass accumulation is rapid in young forests, followed by a slow decline to near zero in old forests [e.g., @lichstein_biomass_2009; @yang_carbon_2011].
-While these trends have been subject of fairly recent qualitative review [@andersonteixeira_altered_2013], there is need for a synthetic, quantitative review taking advantage of the greatly expanded data now available.
+While these trends have been subject of fairly recent qualitative review [@anderson-teixeira_altered_2013], there is need for a synthetic, quantitative review taking advantage of the greatly expanded data now available.
 
@@
 In the past few decades, researchers have started asking how age trends--mostly in $B_{ag}$ or total biomass ($B_{tot}$) accumulation-- vary across biomes.
-Early research on this theme showed that biomass accumulation rates during secondary succession increase with temperature on a global scale [@johnson_post-disturbance_2000; @anderson_temperature-dependence_2006] and with water availability in the neotropics [@poorter_biomass_2016].
+Early research on this theme showed that biomass accumulation rates during secondary succession increase with temperature on a global scale [@johnson_postdisturbance_2000; @anderson_temperaturedependence_2006] and with water availability in the neotropics [@poorter_biomass_2016].
 Most recently, @cook-patton_mapping_2020 reinforced these earlier findings with a much larger dataset and created a high-resolution global map of estimated potential C accumulation rates.
@@
 Given the important role of secondary forests in the current and future global C cycle, concrete understanding of age trends in C fluxes and stocks and how these vary across biomes is critical to better understanding of the global C cycle.
-Accurate estimates of C sequestration rates by regrowth forests are also critical for national greenhouse gas accounting under the IPCC framework [@ipcc_2019_2019; @requena_suarez_estimating_2019] and to quantifying the value of regrowth forests for climate change mitigation [@anderson-teixeira_greenhouse_2011; @goldstein_protecting_2020].
+Accurate estimates of C sequestration rates by regrowth forests are also critical for national greenhouse gas accounting under the IPCC framework [@ipcc_2019_2019; @requenasuarez_estimating_2019] and to quantifying the value of regrowth forests for climate change mitigation [@anderson-teixeira_greenhouse_2011; @goldstein_protecting_2020].
 
@@
 The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age.
-As such, there has been a focus on incorporating data from regrowth forests [*e.g.*, @anderson_temperature-dependence_2006; @martin_carbon_2013; @bonner_meta-analytical_2013] and obtaining stand age data when possible [83% of records in v.2.0; @anderson-teixeira_forc_2018].
+As such, there has been a focus on incorporating data from regrowth forests [*e.g.*, @anderson_temperaturedependence_2006; @martin_carbon_2013; @bonner_metaanalytical_2013] and obtaining stand age data when possible [83% of records in v.2.0; @anderson-teixeira_forc_2018].
 Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], which represented roughly one-third of records in *ForC* v2.0 [@anderson-teixeira_forc_2018].
@@
 First, we imported (via R script) the Global Database of Soil Respiration Database [*SRDB* v4, `r n_SRDB` records; @bond-lamberty_global_2010], and corrections and improvements to *SRDB* arising from this process were incorporated in *SRDB* v5 [@jian_restructured_2020].
-Second, we imported (via R script) the Global Reforestation Opportunity Assessment database [*GROA* v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020], which itself had drawn on an earlier version of *ForC*.
+Second, we imported (via R script) the Global Reforestation Opportunity Assessment database [*GROA* v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forcdb_2020], which itself had drawn on an earlier version of *ForC*.
 Because all records in *GROA* were checked against original publications, these records were given priority over duplicates in *ForC* (Appendix S1).
@@
 
-To facilitate the accessibility of our results and data, and to allow for rapid updates as additional data become available, we automated all database manipulation, analyses, and figure production in R [@r_core_team_r_2020].
+To facilitate the accessibility of our results and data, and to allow for rapid updates as additional data become available, we automated all database manipulation, analyses, and figure production in R [@rcoreteam_language_2020].
 
@@
 Our analysis reveals that carbon cycling is most rapid in the tropics and slowest in boreal regions.
-For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude and increase with temperature on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
+For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude and increase with temperature on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banburymorgan_global_inpress].
 This consistency is not surprising, particularly given commonality in the data analyzed or used for calibration.
 The finding that these patterns hold consistently across numerous fluxes, while consistent with theoretical expectations (Fig.
-`r schematic`), is novel to this analysis [but see @banbury_morgan_global_nodate for nine autotrophic fluxes].
+`r schematic`), is novel to this analysis [but see @banburymorgan_global_inpress for nine autotrophic fluxes].
 
@@
 $NEP$ tends to be relatively small in mature forest stands, which accumulate carbon slowly relative to younger stands, if at all [Fig.
-`r schematic`; @luyssaert_old-growth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
+`r schematic`; @luyssaert_oldgrowth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
 It is therefore consistent with theory -- and with previous research [@luyssaert_co2_2007] -- that there are no pronounced differences across biomes. 
@@
 This is consistent with findings that live biomass accumulation rates ($\Delta B_{ag}$ or $\Delta B_{tot}$) during early secondary succession decrease with latitude [Figs.
-`r stock_age_trends`a, `r all_stock_age_trends`; @anderson_temperature-dependence_2006; @cook-patton_mapping_2020].
+`r stock_age_trends`a, `r all_stock_age_trends`; @anderson_temperaturedependence_2006; @cook-patton_mapping_2020].
 Note, though, that $NEP$ includes not only $\Delta B_{tot}$, but also changes in $DW_{tot}$, $OL$, and soil carbon (not analyzed here). 
@@
 The observed tendency for young forest fluxes to decrease from tropical to boreal regions paralleled patterns in mature forests (Figs.
-`r flux_age_trends`, `r age_trends_all`, `r all_flux_age_trends`), suggesting that regrowth forests follow latitudinal trends in carbon cycling similar to those of mature forests [e.g., @banbury_morgan_global_nodate].
+`r flux_age_trends`, `r age_trends_all`, `r all_flux_age_trends`), suggesting that regrowth forests follow latitudinal trends in carbon cycling similar to those of mature forests [e.g., @banburymorgan_global_inpress].
 
@@
 The highest-biomass forests on Earth are, however, found in coastal temperate climates of both the southern and northern hemisphere [Figs.
-`r schematic`, `r stock_age_trends`a; @keith_re-evaluation_2009; @smithwick_potential_2002; @hu_mapping_2016; @larjavaara_temperature_2012].
+`r schematic`, `r stock_age_trends`a; @keith_reevaluation_2009; @smithwick_potential_2002; @hu_mapping_2016; @larjavaara_temperature_2012].
 Disproportionate representation of forests in one such region--the US Pacific Northwest--inflated estimates of temperate conifer fluxes and stocks for some variables and was responsible for all of the anomalous results described here (e.g., lack of complete C budget closure, anomalous trend across biomes for $BNPP_{coarse}$).
@@
 *ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that proportionally slower decomposition in colder climates results in more buildup of organic matter [Fig.
-`r schematic`; @allen_global_2002; @andersonteixeira_differential_2011].
+`r schematic`; @allen_global_2002; @anderson-teixeira_differential_2011].
 Further research on non-living C stocks in the world's forests will be essential to completing the picture.
@@
 `r flux_age_trends`, `r age_trends_all`), consistent with current understanding of age trends in forest C cycling [Fig.
-`r schematic`; e.g., @andersonteixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
+`r schematic`; e.g., @anderson-teixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
 While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_{auto}$) would be minimal immediately following a stand-clearing disturbance (Fig.
@@
 `r flux_age_trends`).
-This finding is consistent with previous studies showing an increase in $NEP$ across relatively young stand ages [@pregitzer_carbon_2004; @baldocchi_fluxnet_2001; @luyssaert_old-growth_2008].
+This finding is consistent with previous studies showing an increase in $NEP$ across relatively young stand ages [@pregitzer_carbon_2004; @baldocchi_fluxnet_2001; @luyssaert_oldgrowth_2008].
 However, $NEP$ is theoretically expected to peak in intermediate-aged stands and thereafter decline, consistent with decelerating C accumulation as stands age [Fig.
-`r age_trends_all`; @odum_strategy_1969], and such declines have been observed [@law_changes_2003; @luyssaert_old-growth_2008].
+`r age_trends_all`; @odum_strategy_1969], and such declines have been observed [@law_changes_2003; @luyssaert_oldgrowth_2008].
 The fact that $NEP$ values estimated by our models for 100-year-old stands were not systematically different from those of mature stands (lower for temperate broadleaf, higher for temperate conifer, and equal for boreal; Fig.
 `r age_trends_all`) may be driven by differences in geographical representation across age classes or by the fitting of an inappropriate functional form.
-Moreover, both biomass and non-living C stocks often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands [@luyssaert_old-growth_2008; @mcgarvey_carbon_2014; @lichstein_biomass_2009].
+Moreover, both biomass and non-living C stocks often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands [@luyssaert_oldgrowth_2008; @mcgarvey_carbon_2014; @lichstein_biomass_2009].
 Additional data, including on age trends of deadwood, the organic layer, and soil C will be important to parsing the timing and extend of an age-related $NEP$ decrease across forest biomes.
@@
 `r all_diagrams_mature`, `r age_trends_all`).
-On the one hand, this reflects the general fact that ecosystem-scale measurements tend to close the C budget more easily and consistently than, for example, for energy balance [@stoy_data-driven_2013].
+On the one hand, this reflects the general fact that ecosystem-scale measurements tend to close the C budget more easily and consistently than, for example, for energy balance [@stoy_datadriven_2013].
 On the other, however, *ForC* derives data from multiple heterogeneous sources, and standard deviations within each biome are high; as a result, the standard for C closure is relatively loose [*c.f.* @houghton_terrestrial_2020].
@@
 The overall high degree of closure implies that *ForC* gives an at least roughly consistent picture of C cycling within biomes for mature forests. 
-This is an important and useful test, because it allows for consistency checks within the C cycle, for example leveraging separate and independently-measured fluxes to constrain errors in another [@phillips_value_2017; @williams_impacts_2014; @harmon_heterotrophic_2011], or producing internally consistent global data products [@wang_golum-cnp_2018].
+This is an important and useful test, because it allows for consistency checks within the C cycle, for example leveraging separate and independently-measured fluxes to constrain errors in another [@phillips_value_2017; @williams_impacts_2014; @harmon_heterotrophic_2011], or producing internally consistent global data products [@wang_golumcnp_2018].
 
@@
 Studies of regrowth forests tend to focus on sites where recurring anthropogenic disturbance is not a confounding factor.
-Yet, fragmentation and degradation impact a large and growing proportion of Earth's forests [@fao_and_unep_state_2020].
+Yet, fragmentation and degradation impact a large and growing proportion of Earth's forests [@faoandunep_state_2020].
 Fragmentation and the creation of edges strongly impacts forest C cycling [e.g., @chaplin-kramer_degradation_2015; @remy_strong_2016; @reinmann_edge_2017; @smith_evidence_2019; @reinmann_urbanization_2020;  @ordway_carbon_2020].
-Partial logging and other forms of non- stand clearing anthropogenic disturbance also alter forest C cycling [e.g., @huang_long-term_2010; @piponiot_carbon_2016], but are under-studied [@sist_tropical_2015] and excluded from this analysis.
+Partial logging and other forms of non- stand clearing anthropogenic disturbance also alter forest C cycling [e.g., @huang_longterm_2010; @piponiot_carbon_2016], but are under-studied [@sist_tropical_2015] and excluded from this analysis.
 Fragmented and degraded forests do not fit the idealized conceptual framework around which this review is structured (Fig.
-`r schematic`), yet their representation in models, sustainability assessments, and C accounting systems is critical to accurate accounting of C cycling in Earth's forests [e.g., @huang_long-term_2010; @reinmann_edge_2017; @smith_evidence_2019; @piponiot_can_2019].
-Finally, plantation forests account for approximately 3% of Earth's forests [@fao_and_unep_state_2020] but are not included in this analysis. 
-While it is known that these tend to accumulate biomass faster than naturally regenerating forests [@anderson_temperature-dependence_2006; @bonner_meta-analytical_2013], their global scale C cycling patterns remain less clearly understood [c.f.
+`r schematic`), yet their representation in models, sustainability assessments, and C accounting systems is critical to accurate accounting of C cycling in Earth's forests [e.g., @huang_longterm_2010; @reinmann_edge_2017; @smith_evidence_2019; @piponiot_can_2019].
+Finally, plantation forests account for approximately 3% of Earth's forests [@faoandunep_state_2020] but are not included in this analysis. 
+While it is known that these tend to accumulate biomass faster than naturally regenerating forests [@anderson_temperaturedependence_2006; @bonner_metaanalytical_2013], their global scale C cycling patterns remain less clearly understood [c.f.
 @cook-patton_mapping_2020].
@@
 
-The future of forest C cycling [@song_meta-analysis_2019] will shape trends in atmospheric CO~2~ and the course of climate change [@schimel_effect_2015].
+The future of forest C cycling [@song_metaanalysis_2019] will shape trends in atmospheric CO~2~ and the course of climate change [@schimel_effect_2015].
 Our findings, and more generally the data contained in *ForC* and summarized here, can help to meet two major challenges.
@@
 First, improved representation of forest C cycling in models is essential to improving predictions of the future course of climate change, for the simple reason that by definition future projections extend our existing observations and understanding to conditions that do not currently exist on Earth [@mcdowell_predicting_2018; @bonan_climate_2018; @gustafson_extrapolating_2018].
-To ensure that models are giving the right answers for the right reasons [@sulman_multiple_2018], it is important to benchmark against multiple components of the C cycle that are internally consistent with each other [@collier_international_2018; @wang_golum-cnp_2018].
+To ensure that models are giving the right answers for the right reasons [@sulman_multiple_2018], it is important to benchmark against multiple components of the C cycle that are internally consistent with each other [@collier_international_2018; @wang_golumcnp_2018].
 *ForC*’s tens of thousands of records are readily available in a standardized format, along with all code used in the analyses presented here, and we recommend that researchers use these resources to identify and summarize data specific to the analysis at hand.
-Integration of *ForC* with models will be valuable to improving the accuracy and reliability of models [@fer_beyond_2021].
+Integration of *ForC* with models will be valuable to improving the accuracy and reliability of models [@fer_ecosystem_2021].
 
 Second, *ForC* can serve as a pipeline through which information can flow efficiently from forest researchers to decision-makers working to implement forest conservation strategies at global, national, or landscape scales.
-This is already happening: *ForC* has contributed to updating the IPCC guidelines for carbon accounting in forests [@ipcc_2019_2019; @requena_suarez_estimating_2019], mapping C accumulation potential from natural forest regrowth globally [@cook-patton_mapping_2020], and informing ecosystem conservation priorities [@goldstein_protecting_2020].
+This is already happening: *ForC* has contributed to updating the IPCC guidelines for carbon accounting in forests [@ipcc_2019_2019; @requenasuarez_estimating_2019], mapping C accumulation potential from natural forest regrowth globally [@cook-patton_mapping_2020], and informing ecosystem conservation priorities [@goldstein_protecting_2020].
 
 It is also interesting to consider the complementary utility of global-scale but spatially discontinuous databases such as ForC and wall-to-wall remote sensing products. 
-The latter provide insight into aboveground carbon stocks, but less constraint on belowground stocks or carbon fluxes in general [@bondlamberty_estimating_2016; @anav_spatiotemporal_2015]. 
+The latter provide insight into aboveground carbon stocks, but less constraint on belowground stocks or carbon fluxes in general [@bond-lamberty_estimating_2016; @anav_spatiotemporal_2015]. 
 Combining observational data and remote observations may provide a much more comprehensive and accurate picture of global forest C cycling, particularly when used in formal data assimilation systems [@konings_global_2019; @liu_detecting_2018].
@@
 Remote-sensing driven aboveground biomass estimates [e.g., @saatchi_benchmark_2011], calibrated based on high-quality ground-based data [@schepaschenko_forest_2019; @chave_ground_2019], provide the most promising approach, but significant uncertainties remain [@ploton_spatial_2020].
-Note, however, that factors such as stand age and disturbance history are difficult, if not impossible, to detect remotely, and can only be characterized for very recent decades [@hansen_high-resolution_2013; @song_global_2018; @curtis_classifying_2018].
-Ground-based data such as *ForC* are therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020, and thus constraining variables such as carbon sink potential [@luyssaert_old-growth_2008].
+Note, however, that factors such as stand age and disturbance history are difficult, if not impossible, to detect remotely, and can only be characterized for very recent decades [@hansen_highresolution_2013; @song_global_2018; @curtis_classifying_2018].
+Ground-based data such as *ForC* are therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020, and thus constraining variables such as carbon sink potential [@luyssaert_oldgrowth_2008].
 
 In contrast, carbon allocation within forest ecosystems and respiration fluxes cannot be remotely sensed. 
-Efforts such as the Global Carbon Project [@friedlingstein_global_2019] and NASA’s Carbon Monitoring System [@liu_detecting_2018] typically compute respiration as residuals of all other terms [@bondlamberty_estimating_2016; @harmon_heterotrophic_2011].
+Efforts such as the Global Carbon Project [@friedlingstein_global_2019] and NASA’s Carbon Monitoring System [@liu_detecting_2018] typically compute respiration as residuals of all other terms [@bond-lamberty_estimating_2016; @harmon_heterotrophic_2011].
 This means that the errors on respiration outputs are likely to be large and certainly poorly constrained, offering a unique opportunity for databases such as ForC and SRDB [@jian_restructured_2020] to provide observational benchmarks.
-For example, @konings_global_2019 produced a unique top-down estimate of global heterotrophic respiration that can both be compared with extant bottom-up estimates [@bondlamberty_new_2018] and used as an internal consistency check on other parts of the carbon cycle [@phillips_value_2017].
+For example, @konings_global_2019 produced a unique top-down estimate of global heterotrophic respiration that can both be compared with extant bottom-up estimates [@bond-lamberty_new_2018] and used as an internal consistency check on other parts of the carbon cycle [@phillips_value_2017].
 

### 2021-02-27 manuscript/ERL_global_C_review.Rmd
finish Helene's changes
#70

@@
 
-Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide [CO~2~; @bonan_forests_2008; @friedlingstein_global_2019; @ipcc_global_2018].
+Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide [CO~2~, @bonan_forests_2008; @friedlingstein_global_2019; @ipcc_global_2018].
 Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climate_2006; @krause_large_2018; @bonan_model_2019; @divittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011; @ipcc_2019_2019] can be traced to lack of understanding on how C cycling varies across forest types and in relation to stand history.
@@
 `r schematic`, `r flux_age_trends`-`r age_trends_all`).
-In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$)--and therefore $R_{eco}$ are expected to be non-zero following stand-clearing disturbance (Fig.
-`r schematic`), although these may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots and litter [@ribeiro-kumara_how_2020; @maurer_carbon_2016; @bond-lamberty_contribution_2004].
-In this study, we detect no significant age trends in either $R_{het-soil}$ or $R_{soil}$, consistent with previous findings [@law_changes_2003;@pregitzer_carbon_2004;@goulden_patterns_2011].
+In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$) -- and therefore $R_{eco}$ -- are expected to be non-zero following stand-clearing disturbance (Fig.
+`r schematic`).
+These may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots, $DW$, and $OL$ [@ribeiro-kumara_how_2020; @maurer_carbon_2016; @bond-lamberty_contribution_2004], with the latter being strongly dependent upon the type of stand initiating disturbance (discussed below).
+In this study, we detect no significant overall age trends in either $R_{het-soil}$ or $R_{soil}$, consistent with previous findings [@law_changes_2003;@pregitzer_carbon_2004;@goulden_patterns_2011].
 
@@
 `r age_trends_all`).
-The latter are particularly sensitive to the type of disturbance, where disturbances that remove most organic material (e.g., logging, agriculture) result in negligible deadwood in young stands, followed by a buildup over time [e.g., tropical stands in Figs.
+The latter depend strongly on the type of disturbance. 
+Disturbances that remove most woody material (e.g., clearcut logging, agriculture) result in negligible deadwood in young stands, followed by a buildup over time [e.g., tropical stands in Figs.
 `r stock_age_trends`, `r age_trends_all`; e.g., @vargas_biomass_2008].
-In contrast, natural disturbances (e.g., fire, drought) can produce large amounts of deadwood (mostly $DW_{standing}$) that slowly decomposes as the stand recovers, resulting in declines across young stand ages [e.g., temperate and boreal stands in Figs.
+In contrast, natural disturbances (e.g., fire, drought, typhoons/hurricanes) can produce large amounts of deadwood (mostly $DW_{standing}$) that slowly decomposes as the stand recovers, resulting in declines across young stand ages [e.g., temperate and boreal stands in Figs.
 `r stock_age_trends`, `r age_trends_all`; @carmona_coarse_2002].
-Again, further study and synthesis of non-living C stocks across biomes and stand ages will be valuable to giving a more comprehensive picture.
+Further study and synthesis of non-living C stocks across biomes, stand ages, and disturbance types will be valuable to giving a more comprehensive picture.
 
@@
 
-In contrast, age trends for young forests generally remain less clearly defined, in large part because their data records remain relatively sparse (*i.e.*, have low representation of different geographical regions for any given age) for most variables, particularly in the tropics [@anderson-teixeira_carbon_2016].
+In contrast, age trends for young forests generally remain less clearly defined. 
+In large part, this is because their data records remain relatively sparse (*i.e.*, have low representation of different geographical regions for any given age) for most variables, particularly in the tropics [@anderson-teixeira_carbon_2016].
+In addition, variation in the type and severity of stand-initiating disturbances introduces significant heterogeneity in both initial values and age trends of C cycle variables [e.g., @cook-patton_mapping_2020]
 While this review provides a first analysis of age trends in forest C cycling for multiple variables at a global scale, improved resolution of these trends will require additional data.
@@
 There are of course notable holes in the ForC variable coverage that limit the scope of our inferences here. 
-Notably, *ForC* currently has sparse--if any--coverage of fluxes to herbivores and higher consumers, along with the woody mortality ($M_{woody}$) and $DW$ (Table 1, Figs.
-`r DW_tot_fig`-`r DW_down_fig`).
+Notably, *ForC* currently has sparse--if any--coverage of fluxes to herbivores and higher consumers, along with woody mortality ($M_{woody}$) and $DW$ (Table 1, Figs. `r DW_tot_fig`-`r DW_down_fig`). 
+We note that there are considerable opportunities to expand data on $M_{woody}$ and $DW_{standing}$ through calculations from existing forest census data.
 *ForC* does not include soil carbon, which is covered by other efforts [e.g., @kochy_global_2015]. 
@@
 Remote-sensing driven aboveground biomass estimates [e.g., @saatchi_benchmark_2011], calibrated based on high-quality ground-based data [@schepaschenko_forest_2019; @chave_ground_2019], provide the most promising approach, but significant uncertainties remain [@ploton_spatial_2020].
-Note, however, that factors such as stand age and disturbance history are difficult, if not impossible, to detect remotely, and can only be characterized for very recent decades [@hansen_highresolution_2013; @song_global_2018; @curtis_classifying_2018].
+Note, however, that factors such as stand age and disturbance history are difficult, if not impossible, to detect remotely, and can only be characterized for recent decades [@hansen_highresolution_2013; @song_global_2018; @curtis_classifying_2018].
 Ground-based data such as *ForC* are therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020, and thus constraining variables such as carbon sink potential [@luyssaert_oldgrowth_2008].
@@
 This means that the errors on respiration outputs are likely to be large and certainly poorly constrained, offering a unique opportunity for databases such as ForC and SRDB [@jian_restructured_2020] to provide observational benchmarks.
-For example, @konings_global_2019 produced a unique top-down estimate of global heterotrophic respiration that can both be compared with extant bottom-up estimates [@bond-lamberty_new_2018] and used as an internal consistency check on other parts of the carbon cycle [@phillips_value_2017].
+For example, @konings_global_2019 produced a top-down estimate of global heterotrophic respiration that can both be compared with extant bottom-up estimates [@bond-lamberty_new_2018] and used as an internal consistency check on other parts of the carbon cycle [@phillips_value_2017].
 
@@
 As climate change accelerates, understanding and managing the carbon dynamics of forests-- including dynamics and fluxes that cannot be observed by satellites--is critical to forecasting, mitigation, and adaptation.
-The C data in *ForC*, as summarized here, will be valuable to these efforts.
+The C data in *ForC*, as summarized here, are valuable to these efforts.
 Notably, the fact that tropical forests tend to have both the highest rates of C sequestration in young stands [Fig.
@@
 `r flux_age_trends`), and the highest mean biomass [Fig.
-`r stock_age_trends`; Table `r variables_table`; @hu_mapping_2016; @jian_restructured_2020] reinforces the concept that conservation and restoration of these forests is a priority for climate change mitigation, along with high-biomass old-growth temperate stands [@grassi_key_2017; @goldstein_protecting_2020].
-It is also important to note the trade-off in climate mitigation potential of restoration of young forests, with high rates of CO~2~ sequestration [$NEP$; @cook-patton_mapping_2020], versus conservation and management of mature forests, with low $NEP$ but high C stocks that could not be recovered on a time scale relevant to climate change mitigation [@goldstein_protecting_2020].
+`r stock_age_trends`; Table `r variables_table`; @hu_mapping_2016; @jian_restructured_2020] reinforces the idea that conservation and restoration of these forests is a priority for climate change mitigation, along with high-biomass old-growth temperate stands [@grassi_key_2017; @goldstein_protecting_2020].
+It is also important to note the trade-off in climate mitigation potential of restoration of young forests, with high rates of CO~2~ sequestration [$NEP$; @cook-patton_mapping_2020], versus conservation and management of mature forests, with low $NEP$ but high C stocks that, if lost through disturbance, could not be recovered on time scales most relevant to avoiding dangerous climate change [i.e., @goldstein_protecting_2020].
 Generally speaking, the conservation of mature forests will yield greater climate benefits [@anderson-teixeira_greenhouse_2011], but both approaches are critical to avoiding catastrophic climate change [@ipcc_global_2018].

### 2021-02-27 manuscript/ERL_global_C_review.Rmd
fixed citation parentheses

@@
 
-Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide [CO~2~, @bonan_forests_2008; @friedlingstein_global_2019; @ipcc_global_2018].
+Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide, CO~2~ [@bonan_forests_2008; @friedlingstein_global_2019; @ipcc_global_2018].
 Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climate_2006; @krause_large_2018; @bonan_model_2019; @divittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011; @ipcc_2019_2019] can be traced to lack of understanding on how C cycling varies across forest types and in relation to stand history.
@@
 Further standardization has been made possible through research networks such as ForestGEO [@anderson-teixeira_ctfsforestgeo_2015; @davies_forestgeo_2021], NEON [@schimel_neon_2007], and FLUXNET [@baldocchi_fluxnet_2001; @novick_ameriflux_2018].
-Remote sensing technology has become increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass, $B_{ag}$: @saatchi_benchmark_2011; @hu_mapping_2016; @spawn_harmonized_2020; gross primary productivity, $GPP$: @li_mapping_2019], yet measurement and validation of most forest C stocks and fluxes necessarily requires intensive on-the-ground data collection.
+Remote sensing technology has become increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass, $B_{ag}$: @saatchi_benchmark_2011; @hu_mapping_2016; @spawn_harmonized_2020, $GPP$: @li_mapping_2019], yet measurement and validation of most forest C stocks and fluxes necessarily requires intensive on-the-ground data collection.
 
@@
 These data represent ground-based measurements, and *ForC* contains associated data required for interpretation (*e.g.*, stand history, measurement methods).
-Since its most recent publication [*ForC v2.0-Ecology*; @anderson-teixeira_forc_2018], *ForC* has grown `r round((nrow(ForC_measurements) - 17367)/17367*100, digits=0)`%, primarily through the incorporation of two additional large databases that also synthesized published forest C data: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010; @jian_restructured_2020] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020].
+Since its most recent publication [*ForC v2.0-Ecology*, @anderson-teixeira_forc_2018], *ForC* has grown `r round((nrow(ForC_measurements) - 17367)/17367*100, digits=0)`%, primarily through the incorporation of two additional large databases that also synthesized published forest C data: the Global Soil Respiration Database [SRDB, @bond-lamberty_global_2010; @jian_restructured_2020] and the Global Reforestation Opportunity Assessment database [GROA, @cook-patton_mapping_2020].
 Following these additions, *ForC* currently contains `r nrow(ForC_measurements)` records from `r nrow(ForC_plots)` plots and `r n_areas` distinct geographic areas representing all forested biogeographic and climate zones, making it ideal for assessing how forest C cycling varies across biomes and with respect to stand age.
@@
 The resulting models enable construction of fine-scale global maps of estimated C cycling variables.
-This approach can be particularly effective when it integrates satellite measurements that correlate with C cycle variables of interest; for example, solar-induced chlorophyll fluorescence is useful for fine-scale mapping of gross primary productivity [$GPP$; @li_mapping_2019], while LiDAR, radar, and optical imagery are being used to model $B_{ag}$ at regional to global scales [e.g., @saatchi_benchmark_2011; @hu_mapping_2016].
+This approach can be particularly effective when it integrates satellite measurements that correlate with C cycle variables of interest; for example, solar-induced chlorophyll fluorescence is useful for fine-scale mapping of gross primary productivity [$GPP$, @li_mapping_2019], while LiDAR, radar, and optical imagery are being used to model $B_{ag}$ at regional to global scales [e.g., @saatchi_benchmark_2011; @hu_mapping_2016].
 However, all such analyses are ultimately constrained by the quality and coverage of ground-based estimates of forest C fluxes or stocks to train models [e.g., @schepaschenko_forest_2019]. 
@@
 Although the conceptualization of the C cycle in this paper was simplistic by current standards, the paper was foundational in framing the theory around which research on the subject still revolves [@corman_foundations_2019], and the basic framework still holds, albeit with modest modifications [Fig.
-`r schematic`; @anderson-teixeira_altered_2013].
+`r schematic`, @anderson-teixeira_altered_2013].
 Following stand-clearing disturbance, $GPP$, $NPP$, and biomass of leaves ($B_{foliage}$) and fine roots ($B_{root-fine}$) initially increase rapidly and thereafter remain relatively stable ($B_{foliage}$, $B_{root-fine}$, sometimes $GPP$) or decline slightly [$NPP$, sometimes $GPP$; e.g., @law_changes_2003; @pregitzer_carbon_2004; @amiro_ecosystem_2010; @goulden_patterns_2011].
@@
 
-This review synthesizes data from the *ForC* database [Fig. `r figure_map`; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
+This review synthesizes data from the *ForC* database [Fig. `r figure_map`, https://github.com/forc-db/ForC, @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
 *ForC* amalgamates numerous intermediary data sets [*e.g.*, @luyssaert_co2_2007; @bond-lamberty_global_2010; @cook-patton_mapping_2020] and original studies.
@@
 The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age.
-As such, there has been a focus on incorporating data from regrowth forests [*e.g.*, @anderson_temperaturedependence_2006; @martin_carbon_2013; @bonner_metaanalytical_2013] and obtaining stand age data when possible [83% of records in v.2.0; @anderson-teixeira_forc_2018].
+As such, there has been a focus on incorporating data from regrowth forests [*e.g.*, @anderson_temperaturedependence_2006; @martin_carbon_2013; @bonner_metaanalytical_2013] and obtaining stand age data when possible [83% of records in v.2.0, @anderson-teixeira_forc_2018].
 Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], which represented roughly one-third of records in *ForC* v2.0 [@anderson-teixeira_forc_2018].
 Since publication of ForC v2.0, we imported three large additional databases into *ForC* via a combination of R scripts and manual edits. 
-First, we imported (via R script) the Global Database of Soil Respiration Database [*SRDB* v4, `r n_SRDB` records; @bond-lamberty_global_2010], and corrections and improvements to *SRDB* arising from this process were incorporated in *SRDB* v5 [@jian_restructured_2020].
-Second, we imported (via R script) the Global Reforestation Opportunity Assessment database [*GROA* v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forcdb_2020], which itself had drawn on an earlier version of *ForC*.
+First, we imported (via R script) the Global Database of Soil Respiration Database [*SRDB* v4, `r n_SRDB` records, @bond-lamberty_global_2010], and corrections and improvements to *SRDB* arising from this process were incorporated in *SRDB* v5 [@jian_restructured_2020].
+Second, we imported (via R script) the Global Reforestation Opportunity Assessment database [*GROA* v1.0, `r n_GROA` records, @cook-patton_mapping_2020; @anderson-teixeira_forcdb_2020], which itself had drawn on an earlier version of *ForC*.
 Because all records in *GROA* were checked against original publications, these records were given priority over duplicates in *ForC* (Appendix S1).
@@
 For this analysis, we combined some of ForC’s specific variables into more broadly defined variables. 
-Specifically, net ecosystem exchange [measured by eddy-covariance; @baldocchi_fluxnet_2001] and biometric estimates of $NEP$ were combined into the single variable $NEP$ (Table 1).
+Specifically, net ecosystem exchange [measured by eddy-covariance, @baldocchi_fluxnet_2001] and biometric estimates of $NEP$ were combined into the single variable $NEP$ (Table 1).
 Furthermore, for $NPP$, aboveground $NPP$ ($ANPP$), and the litterfall component of $ANPP$ ($ANPP_{litterfall}$), we combined ForC variables specifying inclusion or exclusion of minor components (*e.g.*, measurements including or excluding fruit production, flower production, and herbivory).
@@
 Forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis.
-We distinguished broadleaf and needleleaf forests based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
+We distinguished broadleaf and needleleaf forests based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP, @jung_exploiting_2006].
 For young tropical forests imported from *GROA* but not yet classified by leaf type, we assumed that all were broadleaf, consistent with the rarity of naturally regenerating needleleaf forests in the tropics.
@@
 
-To test for differences across mature forest biomes, we also examined how stand age impacted fluxes and stocks, employing a mixed effects model ['lmer' function in 'lme4' R package; @bates_fitting_2015] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
+To test for differences across mature forest biomes, we also examined how stand age impacted fluxes and stocks, employing a mixed effects model ['lmer' function in 'lme4' R package, @bates_fitting_2015] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
 When biome had a significant effect, we used a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
@@
 $NEP$ tends to be relatively small in mature forest stands, which accumulate carbon slowly relative to younger stands, if at all [Fig.
-`r schematic`; @luyssaert_oldgrowth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
+`r schematic`, @luyssaert_oldgrowth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
 It is therefore consistent with theory -- and with previous research [@luyssaert_co2_2007] -- that there are no pronounced differences across biomes. 
-Rather, variation in $NEP$ of mature forests appears to be controlled less by climate and more by other factors including moderate disturbances [@curtis_forest_2018] or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition; @wilson_stability_2016].
+Rather, variation in $NEP$ of mature forests appears to be controlled less by climate and more by other factors including moderate disturbances [@curtis_forest_2018] or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition, @wilson_stability_2016].
 The fact that mature temperate broadleaf forests have a higher mean than the other biomes may reflect the fact that most of these forests are older secondary forests that, while classified here as mature, are still accumulating carbon [@curtis_forest_2018].
@@
 This is consistent with findings that live biomass accumulation rates ($\Delta B_{ag}$ or $\Delta B_{tot}$) during early secondary succession decrease with latitude [Figs.
-`r stock_age_trends`a, `r all_stock_age_trends`; @anderson_temperaturedependence_2006; @cook-patton_mapping_2020].
+`r stock_age_trends`a, `r all_stock_age_trends`, @anderson_temperaturedependence_2006; @cook-patton_mapping_2020].
 Note, though, that $NEP$ includes not only $\Delta B_{tot}$, but also changes in $DW_{tot}$, $OL$, and soil carbon (not analyzed here). 
 Biome differences in the accumulation rates of $DW$, $OL$, and soil C  have not been detected, in part because these variables do not consistently increase with stand age [Figs.
-`r schematic`, `r stock_age_trends`, `r DW_tot_fig`-`r OL_fig`, and see discussion below; @cook-patton_mapping_2020].
+`r schematic`, `r stock_age_trends`, `r DW_tot_fig`-`r OL_fig`, and see discussion below, @cook-patton_mapping_2020].
 
@@
 The highest-biomass forests on Earth are, however, found in coastal temperate climates of both the southern and northern hemisphere [Figs.
-`r schematic`, `r stock_age_trends`a; @keith_reevaluation_2009; @smithwick_potential_2002; @hu_mapping_2016; @larjavaara_temperature_2012].
+`r schematic`, `r stock_age_trends`a, @keith_reevaluation_2009; @smithwick_potential_2002; @hu_mapping_2016; @larjavaara_temperature_2012].
 Disproportionate representation of forests in one such region--the US Pacific Northwest--inflated estimates of temperate conifer fluxes and stocks for some variables and was responsible for all of the anomalous results described here (e.g., lack of complete C budget closure, anomalous trend across biomes for $BNPP_{coarse}$).
@@
 
-Whereas aboveground biomass can be remotely sensed [albeit with significant uncertainties; @ploton_spatial_2020] and receives significant research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) carbon across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
+Whereas aboveground biomass can be remotely sensed [albeit with significant uncertainties, @ploton_spatial_2020] and receives significant research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) carbon across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
 Although these stocks can be important--exceeding 100 Mg C ha^-1^ in some stands (Figs.
@@
 *ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that proportionally slower decomposition in colder climates results in more buildup of organic matter [Fig.
-`r schematic`; @allen_global_2002; @anderson-teixeira_differential_2011].
+`r schematic`, @allen_global_2002; @anderson-teixeira_differential_2011].
 Further research on non-living C stocks in the world's forests will be essential to completing the picture.
@@
 However, $NEP$ is theoretically expected to peak in intermediate-aged stands and thereafter decline, consistent with decelerating C accumulation as stands age [Fig.
-`r age_trends_all`; @odum_strategy_1969], and such declines have been observed [@law_changes_2003; @luyssaert_oldgrowth_2008].
+`r age_trends_all`, @odum_strategy_1969], and such declines have been observed [@law_changes_2003; @luyssaert_oldgrowth_2008].
 The fact that $NEP$ values estimated by our models for 100-year-old stands were not systematically different from those of mature stands (lower for temperate broadleaf, higher for temperate conifer, and equal for boreal; Fig.
@@
 In contrast, natural disturbances (e.g., fire, drought, typhoons/hurricanes) can produce large amounts of deadwood (mostly $DW_{standing}$) that slowly decomposes as the stand recovers, resulting in declines across young stand ages [e.g., temperate and boreal stands in Figs.
-`r stock_age_trends`, `r age_trends_all`; @carmona_coarse_2002].
+`r stock_age_trends`, `r age_trends_all`, @carmona_coarse_2002].
 Further study and synthesis of non-living C stocks across biomes, stand ages, and disturbance types will be valuable to giving a more comprehensive picture.
@@
 Notably, the fact that tropical forests tend to have both the highest rates of C sequestration in young stands [Fig.
-`r stock_age_trends`; @cook-patton_mapping_2020], fueled by their generally high C flux rates (Table 1; Fig.
+`r stock_age_trends`, @cook-patton_mapping_2020], fueled by their generally high C flux rates (Table 1; Fig.
 `r flux_age_trends`), and the highest mean biomass [Fig.
-`r stock_age_trends`; Table `r variables_table`; @hu_mapping_2016; @jian_restructured_2020] reinforces the idea that conservation and restoration of these forests is a priority for climate change mitigation, along with high-biomass old-growth temperate stands [@grassi_key_2017; @goldstein_protecting_2020].
-It is also important to note the trade-off in climate mitigation potential of restoration of young forests, with high rates of CO~2~ sequestration [$NEP$; @cook-patton_mapping_2020], versus conservation and management of mature forests, with low $NEP$ but high C stocks that, if lost through disturbance, could not be recovered on time scales most relevant to avoiding dangerous climate change [i.e., @goldstein_protecting_2020].
+`r stock_age_trends`, Table `r variables_table`, @hu_mapping_2016; @jian_restructured_2020] reinforces the idea that conservation and restoration of these forests is a priority for climate change mitigation, along with high-biomass old-growth temperate stands [@grassi_key_2017; @goldstein_protecting_2020].
+It is also important to note the trade-off in climate mitigation potential of restoration of young forests, with high rates of CO~2~ sequestration [$NEP$, @cook-patton_mapping_2020], versus conservation and management of mature forests, with low $NEP$ but high C stocks that, if lost through disturbance, could not be recovered on time scales most relevant to avoiding dangerous climate change [i.e., @goldstein_protecting_2020].
 Generally speaking, the conservation of mature forests will yield greater climate benefits [@anderson-teixeira_greenhouse_2011], but both approaches are critical to avoiding catastrophic climate change [@ipcc_global_2018].

### 2021-02-28 manuscript/ERL_global_C_review.Rmd
working on Abby's changes

@@
 Specifically, we use `r n_records` *ForC* records for `r nrow(T1)` C cycle variables from `r n_areas_tot` geographic locations to characterize ensemble C budgets for four broad forest types -- tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and boreal.
-We calculate means and standard deviations for both mature and regrowth (age <100 years) forests, and quantify trends with stand age in regrowth forests for all variables with sufficient data.
+We calculate means and standard deviations for both mature and regrowth (age <100 years) forests and quantify trends with stand age in regrowth forests for all variables with sufficient data.
 
-*Review Results/ Synthesis.*   The rate of C cycling generally decreased from tropical to temperate to boreal in both mature and regrowth forests, whereas C stocks showed less directional variation.
+*Review Results/ Synthesis.*  C cycling rates generally decreased from tropical to temperate to boreal in both mature and regrowth forests, whereas C stocks showed less directional variation.
 Mature forest net ecosystem production did not differ significantly among biomes.
@@
 
-Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide, CO~2~ [@bonan_forests_2008; @friedlingstein_global_2019; @ipcc_global_2018].
+Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide, CO~2~ [@bonan_forests_2008; @ipcc_global_2018; @friedlingstein_global_2019].
 Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climate_2006; @krause_large_2018; @bonan_model_2019; @divittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011; @ipcc_2019_2019] can be traced to lack of understanding on how C cycling varies across forest types and in relation to stand history.
 This requires accessible, comprehensive, and large-scale databases with global coverage, which runs contrary to the traditional way forest C stocks and fluxes have been measured and published.
-Large-scale synthesis is critical to benchmarking model performance with global data [@clark_field_2017; @luo_framework_2012], quantifying the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], and using book-keeping methods to quantify actual or potential exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
+Large-scale synthesis is critical to benchmarking model performance with global data [@luo_framework_2012; @clark_field_2017], quantifying the role of forests in the global C cycle [*e.g.*, @pan_large_2011], and using book-keeping methods to quantify actual or potential exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
 
@@
 Most of this enormous C uptake is counterbalanced by releases to the atmosphere through ecosystem respiration ($R_{eco}$) and fire.
-In recent years, total forest C uptake has exceeded releases, such that globally forests have been a C sink [@harris_global_2021].
-Considering only areas remaining in forest, this C sink has averaged 3.2 $\pm$ 0.6 Gt C yr^-1^ for 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
+In recent years, total forest C uptake has exceeded releases, such that forests globally have been a C sink [@harris_global_2021].
+This C sink has averaged 3.2 $\pm$ 0.6 Gt C yr^-1^ for 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions, when considering only areas remaining forested [@friedlingstein_global_2019].
 However, deforestation, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reduces the net forest sink to ~1.1-2.2 Gt C yr^-1^ [@friedlingstein_global_2019; @harris_global_2021].
-Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thus central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; @cavaleri_urgent_2015].
+Understanding, modeling, and managing forest-atmosphere CO2 exchange is central to mitigating climate change [@cavaleri_urgent_2015; @grassi_key_2017; @griscom_natural_2017].
 
@@
 Although age trends in aboveground biomass have been well-studied and synthesized globally [@cook-patton_mapping_2020], there is a relative dearth of data and synthesis on other C stocks and fluxes in secondary forests.
-An understanding of age-related trends in forest C cycling is particularly critical for reducing uncertainty regarding the potential for carbon uptake and climate change mitigation by regrowth forests [@krause_large_2018; @cook-patton_mapping_2020].
+Understanding age-related trends in forest C cycling is particularly critical for reducing uncertainty regarding the potential for carbon uptake and climate change mitigation by regrowth forests [@krause_large_2018; @cook-patton_mapping_2020].
 
@@
 New technology has also enabled researchers to directly measure an expanding set of variables, notably including continuous measurements of soil CO~2~ efflux [@kuzyakov_sources_2006] and ecosystem-atmosphere CO~2~ exchange [@baldocchi_fluxnet_2001].
-Measurement techniques have been increasingly standardized; for example, of the biomass allometries that strongly influence estimates of most C cycle variables [e.g., @chave_improved_2014].
+Measurement techniques have been increasingly standardized, such as the biomass allometries that strongly influence estimates of most C cycle variables [e.g., @chave_improved_2014].
 Further standardization has been made possible through research networks such as ForestGEO [@anderson-teixeira_ctfsforestgeo_2015; @davies_forestgeo_2021], NEON [@schimel_neon_2007], and FLUXNET [@baldocchi_fluxnet_2001; @novick_ameriflux_2018].
-Remote sensing technology has become increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass, $B_{ag}$: @saatchi_benchmark_2011; @hu_mapping_2016; @spawn_harmonized_2020, $GPP$: @li_mapping_2019], yet measurement and validation of most forest C stocks and fluxes necessarily requires intensive on-the-ground data collection.
+Remote sensing technology has become increasingly useful for global- or regional-scale estimates of a few critical variables, including aboveground biomass [$B_{ag}$, @saatchi_benchmark_2011; @hu_mapping_2016; @spawn_harmonized_2020] and $GPP$ [@li_mapping_2019].
+Yet, measurement and validation of most forest C stocks and fluxes necessarily requires intensive on-the-ground data collection.
 

### 2021-02-28 manuscript/ERL_global_C_review.Rmd
Abby's comments, through results
#70

@@
 To illustrate the magnitude of different fluxes, arrow width is proportional to the square root of the corresponding flux.
-Mean component fluxes do not necessarily add up to the mean total fluxes because different sets of sites are included depending on availability of data (Figs.
+Mean component fluxes do not necessarily add up to the mean total fluxes because different sets of sites are included depending on data availability (Figs.
 S5-S30).
@@
 age_trends_all_caption= "Figure 9 | Age trends in C cycling.
-Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle.
+The selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all C cycle elements.
 Error bars on mature forest flux estimates indicate $\\pm$ 1 standard deviation.
@@
 For non-tropical forests: $ANPP_{woody}=min(ANPP_{stem},ANPP_{woody})$, $R_{auto-ag}=R_{eco}-R{soil}$.
-Note that there remain substantial uncertainties as to the functional form of age trends and discrepencies in closure among related variables."
+Note that there remain substantial uncertainties in the functional form of age trends and discrepancies in closure among related variables."
 
@@
 This requires accessible, comprehensive, and large-scale databases with global coverage, which runs contrary to the traditional way forest C stocks and fluxes have been measured and published.
-Large-scale synthesis is critical to benchmarking model performance with global data [@luo_framework_2012; @clark_field_2017], quantifying the role of forests in the global C cycle [*e.g.*, @pan_large_2011], and using book-keeping methods to quantify actual or potential exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
+Large-scale synthesis is critical to benchmarking model performance with global data [@luo_framework_2012; @clark_field_2017], quantifying the role of forests in the global C cycle [e.g., @pan_large_2011], and using book-keeping methods to quantify actual or potential exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
 
@@
 Alongside these conceptual and methodological developments, there has been a proliferation of measurements across the world's forests. 
-The result of decades of research on forest C cycling is tens of thousands of records distributed across thousands of scientific articles, varying in data formats, units, measurement methods, *etc.*
-To address global-scale questions, researchers began synthesizing data into increasingly large databases [e.g., @lieth_primary_1973; @luyssaert_co2_2007;  @bond-lamberty_global_2010; @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018; @cook-patton_mapping_2020].
+The result of decades of research on forest C cycling is tens of thousands of records distributed across thousands of scientific articles, varying in data formats, units, measurement methods, etc.
+To address questions at a global scale, researchers began synthesizing data into increasingly large databases [e.g., @lieth_primary_1973; @luyssaert_co2_2007;  @bond-lamberty_global_2010; @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018; @cook-patton_mapping_2020].
 The current largest, most comprehensive database on forest C cycling is *ForC* [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], which contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables), with different variables capturing distinct ecosystem pools (e.g., woody, foliage, and root biomass; dead wood) and flux types (e.g, gross and net primary productivity; soil, root, and ecosystem respiration).
-These data represent ground-based measurements, and *ForC* contains associated data required for interpretation (*e.g.*, stand history, measurement methods).
+These data represent ground-based measurements, and *ForC* contains associated data required for interpretation (e.g., stand history, measurement methods).
 Since its most recent publication [*ForC v2.0-Ecology*, @anderson-teixeira_forc_2018], *ForC* has grown `r round((nrow(ForC_measurements) - 17367)/17367*100, digits=0)`%, primarily through the incorporation of two additional large databases that also synthesized published forest C data: the Global Soil Respiration Database [SRDB, @bond-lamberty_global_2010; @jian_restructured_2020] and the Global Reforestation Opportunity Assessment database [GROA, @cook-patton_mapping_2020].
@@
 C stocks of mature forests show less directional variation (Fig. `r schematic`).
-On average, aboveground biomass ($B_{ag}$) tends to decrease with latitude, but not as dramatically as fluxes, and with the highest $B_{ag}$ forests in relatively cool, moist temperate regions [@keith_reevaluation_2009; @smithwick_potential_2002; @hu_mapping_2016].
+On average, aboveground biomass ($B_{ag}$) tends to decrease with latitude, but not as dramatically as fluxes, and with the highest $B_{ag}$ forests in relatively cool, moist temperate regions [@smithwick_potential_2002; @keith_reevaluation_2009; @hu_mapping_2016].
 In contrast, standing and downed dead wood ($DW_{standing}$ and $DW_{down}$, respectively, summing to $DW_{tot}$) and the organic layer ($OL$) tend to accumulate more in colder climates where decomposition is slow relative to $NPP$ [@harmon_ecology_1986; @allen_global_2002].
@@
 While estimates of some variables (e.g., $B_{ag}$, $GPP$, $NPP$, $R_{soil}$) are widely available, many remain poorly characterized (e.g., $DW_{tot}$; $OL$; autotrophic respiration, $R_{auto}$) even at the coarse resolution of biomes.
-This is a critical limitation not only for understanding forest C cycling, but also for quantifying forest-based climate change mitigation potential across forest biomes or ecozones [e.g., @ipcc_2019_2019].
+This is a critical limitation for understanding forest C cycling and quantifying forest-based climate change mitigation potential across forest biomes or ecozones [e.g., @ipcc_2019_2019].
 
@@
 The decline in $NPP$ occurs because $R_{auto}$ increases relative to $GPP$ as forests age, corresponding to declining carbon use efficiency with stand age [@delucia_forest_2007; @collalti_forest_2020].
-Heterotrophic respiration, most of which originates from the soil ($R_{het-soil}$) remains relatively constant with stand age [@law_changes_2003; @pregitzer_carbon_2004; @goulden_patterns_2011], with the result that net ecosystem production ($NEP=GPP-R_{eco}$, where $R_{eco}$ is total ecosystem respiration) is initially negative, increases to a maximum at intermediate ages, and declines--typically to a small positive value--thereafter [@law_changes_2003; @pregitzer_carbon_2004; @amiro_ecosystem_2010; @goulden_patterns_2011; @luyssaert_oldgrowth_2008].
+Heterotrophic respiration, most of which originates from the soil ($R_{het-soil}$) remains relatively constant with stand age [@law_changes_2003; @pregitzer_carbon_2004; @goulden_patterns_2011], with the result that net ecosystem production ($NEP=GPP-R_{eco}$, where $R_{eco}$ is total ecosystem respiration) is initially negative, increases to a maximum at intermediate ages, and declines--typically to a small positive value--thereafter [@law_changes_2003; @pregitzer_carbon_2004; @luyssaert_oldgrowth_2008; @amiro_ecosystem_2010; @goulden_patterns_2011].
 The result is that biomass accumulation is rapid in young forests, followed by a slow decline to near zero in old forests [e.g., @lichstein_biomass_2009; @yang_carbon_2011].
-While these trends have been subject of fairly recent qualitative review [@anderson-teixeira_altered_2013], there is need for a synthetic, quantitative review taking advantage of the greatly expanded data now available.
+While these trends have been the subject of fairly recent qualitative review [@anderson-teixeira_altered_2013], there is need for a synthetic, quantitative review taking advantage of the greatly expanded data now available.
 
@@
 Early research on this theme showed that biomass accumulation rates during secondary succession increase with temperature on a global scale [@johnson_postdisturbance_2000; @anderson_temperaturedependence_2006] and with water availability in the neotropics [@poorter_biomass_2016].
-Most recently, @cook-patton_mapping_2020 reinforced these earlier findings with a much larger dataset and created a high-resolution global map of estimated potential C accumulation rates.
-However, there has been little synthesis of cross-biome differences in variables other than biomass and its accumulation rate (but see @cook-patton_mapping_2020 for $DW$, $OL$, and soil C accumulation in young stands).
-Given the important role of secondary forests in the current and future global C cycle, concrete understanding of age trends in C fluxes and stocks and how these vary across biomes is critical to better understanding of the global C cycle.
-Accurate estimates of C sequestration rates by regrowth forests are also critical for national greenhouse gas accounting under the IPCC framework [@ipcc_2019_2019; @requenasuarez_estimating_2019] and to quantifying the value of regrowth forests for climate change mitigation [@anderson-teixeira_greenhouse_2011; @goldstein_protecting_2020].
+@cook-patton_mapping_2020 reinforced these earlier findings with a much larger dataset and created a high-resolution global map of estimated potential C accumulation rates.
+However, there has been little synthesis of cross-biome differences in variables other than biomass and its accumulation rate [but see @cook-patton_mapping_2020 for $DW$, $OL$, and soil C accumulation in young stands].
+Given the important role of secondary forests in the current and future global C cycle, a concrete understanding of age trends in C fluxes and stocks and how these vary across biomes is critical to better understanding the global C cycle.
+Accurate estimates of C sequestration rates by regrowth forests are also critical for national greenhouse gas accounting under the IPCC framework [@ipcc_2019_2019; @requenasuarez_estimating_2019] and quantifying the value of regrowth forests for climate change mitigation [@anderson-teixeira_greenhouse_2011; @goldstein_protecting_2020].
 
-Here, we conduct a data-based review of carbon cycling from a stand to global level, and by biome and stand age, using the largest global compilation of forest carbon data, which is available in our open source Global Carbon Forest database (*ForC*; Fig.
-`r figure_map`).
+Here, we conduct a data-based review of carbon cycling from a stand to global level, and by biome and stand age, using our open-source Global Carbon Forest database (*ForC*; Fig.
+`r figure_map`), which is the largest global compilation of forest carbon data.
 Our goal is to provide a comprehensive synthesis on broad trends in forest C cycling that can serve as a foundation for improved understanding of global forest C cycling and highlight where key sources of uncertainty still reside.
@@
 This review synthesizes data from the *ForC* database [Fig. `r figure_map`, https://github.com/forc-db/ForC, @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
-*ForC* amalgamates numerous intermediary data sets [*e.g.*, @luyssaert_co2_2007; @bond-lamberty_global_2010; @cook-patton_mapping_2020] and original studies.
+*ForC* amalgamates numerous intermediary data sets [e.g., @luyssaert_co2_2007; @bond-lamberty_global_2010; @cook-patton_mapping_2020] and original studies.
 Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records.
 The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age.
-As such, there has been a focus on incorporating data from regrowth forests [*e.g.*, @anderson_temperaturedependence_2006; @martin_carbon_2013; @bonner_metaanalytical_2013] and obtaining stand age data when possible [83% of records in v.2.0, @anderson-teixeira_forc_2018].
+As such, there has been a focus on incorporating data from regrowth forests [e.g., @anderson_temperaturedependence_2006; @martin_carbon_2013; @bonner_metaanalytical_2013] and obtaining stand age data when possible [83% of records in *ForC* v.2.0, @anderson-teixeira_forc_2018].
 Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], which represented roughly one-third of records in *ForC* v2.0 [@anderson-teixeira_forc_2018].
 Since publication of ForC v2.0, we imported three large additional databases into *ForC* via a combination of R scripts and manual edits. 
-First, we imported (via R script) the Global Database of Soil Respiration Database [*SRDB* v4, `r n_SRDB` records, @bond-lamberty_global_2010], and corrections and improvements to *SRDB* arising from this process were incorporated in *SRDB* v5 [@jian_restructured_2020].
+First, we imported (via R script) the Global Soil Respiration Database [*SRDB* v4, `r n_SRDB` records, @bond-lamberty_global_2010], with corrections and improvements to *SRDB* arising from this process incorporated into *SRDB* v5 [@jian_restructured_2020].
 Second, we imported (via R script) the Global Reforestation Opportunity Assessment database [*GROA* v1.0, `r n_GROA` records, @cook-patton_mapping_2020; @anderson-teixeira_forcdb_2020], which itself had drawn on an earlier version of *ForC*.
@@
 Third, we incorporated records of annual $NEP$, $GPP$, and $R_{eco}$ from the FLUXNET2015 dataset [@pastorello_fluxnet2015_2020], treating these records as authoritative when they duplicated earlier records (Appendix S1).
-We have also added data from individual publications, with a particular focus on productivity [e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @lutz_global_2018; @johnson_climate_2018].
+We have also added data from individual publications, focusing on productivity [e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @johnson_climate_2018; @lutz_global_2018].
 A record of data sets added to *ForC* over the course of its development is available at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv.
@@
 We did not analyze soil carbon, which is not a focus of the *ForC* database.
-Note that two flux variables, aboveground heterotropic respiration ($R_{het-ag}$) and total heterotropic respiration ($R_{het}$), were included for conceptual completeness but had no records in *ForC* (Table 1).
+Note that two flux variables, aboveground heterotrophic respiration ($R_{het-ag}$) and total heterotrophic respiration ($R_{het}$), were included for conceptual completeness but had no records in *ForC* (Table 1).
 Records for our focal variables represented `r round(n_focal_variable_records_eligible_for_inclusion/records_eligible_for_inclusion*100,digits=1)`% of the total records eligible for inclusion.
-For this analysis, we combined some of ForC’s specific variables into more broadly defined variables. 
+For this analysis, we combined some specific variables from *ForC* into more broadly defined variables. 
 Specifically, net ecosystem exchange [measured by eddy-covariance, @baldocchi_fluxnet_2001] and biometric estimates of $NEP$ were combined into the single variable $NEP$ (Table 1).
-Furthermore, for $NPP$, aboveground $NPP$ ($ANPP$), and the litterfall component of $ANPP$ ($ANPP_{litterfall}$), we combined ForC variables specifying inclusion or exclusion of minor components (*e.g.*, measurements including or excluding fruit production, flower production, and herbivory).
-Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), trees were censused down to a minimum diameter breast height (DBH) threshold of 10 cm or less. 
+Furthermore, for $NPP$, aboveground $NPP$ ($ANPP$), and the litterfall component of $ANPP$ ($ANPP_{litterfall}$), we combined ForC variables specifying inclusion or exclusion of minor components (e.g., measurements including or excluding fruit production, flower production, and herbivory).
+Throughout ForC, for all measurements drawing from tree census data (e.g., biomass, productivity), trees were censused down to a minimum diameter breast height (DBH) threshold of 10 cm or less. 
 All records were based on ground-based field measurements.
@@
 We classified forests as “young”  if stand age was less than 100 years,  or “mature” if stand age was older or if they were described as “mature”, “old growth”, “intact”, or “undisturbed” in the original publication.
-Assigning stands to these groupings required the exclusion of records for which *ForC* lacked geographic coordinates (`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database) or records of stand age or forest maturity (`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
+Assigning stands to these groupings required excluding records for which *ForC* lacked geographic coordinates (`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in the full database) or records of stand age or forest maturity (`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in the full database).
 We also excluded records of stand age = 0 year (`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
 In total, our analysis retained `r n_records` records. 
-Numbers of records by biome and age class are given in Table S1.
+Numbers of records by biome and age class are provided in Table S1.
 
-We calculated means and standard deviations of each mature forest C cycle variable by biome over geographically distinct areas to produce biome-specific schematics.
+We calculated the means and standard deviations of each mature forest C cycle variable by biome over geographically distinct areas to produce biome-specific schematics.
 We first averaged any repeated measurements within a plot.
@@
 We tested whether the C budgets described above "closed"--*i.e.*, whether they were internally consistent.
-Specifically, we first defined relationships among variables: for example, $NEP =	GPP - R_{eco}$, $BNPP = BNPP_{coarse} + BNPP_{fine}$, $DW_{tot} = DW_{standing} + DW_{down}$).
+Specifically, we first defined relationships among variables (e.g., $NEP =	GPP - R_{eco}$, $BNPP = BNPP_{coarse} + BNPP_{fine}$, $DW_{tot} = DW_{standing} + DW_{down}$).
 Henceforth, we refer to the variables on the left side of the equation as "aggregate" fluxes or stocks, and those that are summed as "component" fluxes or stocks, noting that the same variable can take both aggregate and component positions in different relationships.
@@
 
-To test for differences across mature forest biomes, we also examined how stand age impacted fluxes and stocks, employing a mixed effects model ['lmer' function in 'lme4' R package, @bates_fitting_2015] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
-When biome had a significant effect, we used a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
+To test for differences across mature forest biomes, we also examined how stand age impacted fluxes and stocks, employing a mixed effects model ['lmer' function in 'lme4' R package, @bates_fitting_2015] with biome as a fixed effect and plot nested within geographic area as random effects on the intercept.
+When biome had a significant effect, we used Tukey's pairwise comparison to see which biomes were significantly different from one another. 
 This analysis was run for variables and biomes with records for at least seven distinct geographic areas per biome (Table 1).
@@
 To test for age trends in young (<100yrs) forests, we employed a mixed effects model with biome and log10[stand age] as fixed effects and plot nested within geographic area as a random effect on the intercept.
-This analysis was run for variables and biomes with records for at least three distinct geographic areas per biome, excluding any biomes that failed this criteria (Table 1).
-When the effect of stand age was significant at p $\le$ 0.05 and when each biome had records for stands of at least 10 different ages, a biome $\times$ stand.age interaction was included in the model.
+This analysis was run for variables and biomes with records for at least three distinct geographic areas per biome, excluding any biomes that failed this criterion (Table 1).
+When the effect of stand age was significant at p $\le$ 0.05, and when each biome had records for stands of at least ten different ages, a biome $\times$ stand age interaction was included in the model.
 We note that the logarithmic function fit in this analysis does not always correspond to theoretical expectations (Fig.
-`r schematic`); however, data limitations did not support fitting of functions with more parameters or reliable comparison of different functional forms.
-Within the data constraints, we deemed a logarithmic function to be the most appropriate functional form for the majority of variables.
+`r schematic`); however, data limitations did not support fitting of functions with more parameters or reliable comparisons of different functional forms.
+Within the data constraints, we deemed a logarithmic function to be the appropriate functional form for most variables.
 
@@
 `r flux_age_trends`, `r all_flux_age_trends`).
-Differences between tropical and boreal forests were always significant, with temperate forests intermediate and significantly different from one or both.
+Differences between tropical and boreal forests were consistently significant, with temperate forests intermediate and significantly different from one or both.
 Fluxes tended to be numerically greater in temperate broadleaf than temperate conifer forests, but the difference was never statistically significant.
@@
 The scatterplots show age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
-The fitted lines indicate the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction (interaction effects were tested only if the main age effect was significant and data were available for at least 10 stand ages per biome--i.e., for $GPP$, $ANPP$, $R_{soil}$, and $NEP$).
+The fitted lines indicate the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction (interaction effects were tested only if the main age effect was significant and data were available for at least ten stand ages per biome--i.e., for $GPP$, $ANPP$, $R_{soil}$, and $NEP$).
 The boxplots illustrate variation among biomes in mature forests, with different letters indicating significant differences between biomes.
@@
 There were sufficient data to assess mature forest biome differences for nine stock variables, and significant differences among biomes were detected for five variables ($B_{tot}$, $B_{ag}$, $B_{ag-wood}$, $B_{foliage}$, $B_{root-coarse}$; Table 1).
-For $B_{tot}$ and $B_{ag}$, tropical broadleaf forests had the highest mean biomass and boreal forests the lowest, with intermediate means for temperate broadleaf and needleleaf forests (temperate needleleaf excluded from $B_{tot}$ analysis becuase of insufficient data; Figs.
+For $B_{tot}$ and $B_{ag}$, tropical broadleaf forests had the highest mean biomass and boreal forests the lowest, with intermediate means for temperate broadleaf and needleleaf forests (temperate needleleaf excluded from $B_{tot}$ analysis because of insufficient data; Figs.
 `r B_tot_fig`, `r B_ag_fig`).
@@
 
-```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | Age trends and biome differences in some of the major forest C stocks: (a) $B_{ag}$, (b) $B_{foliage}$, (c) $B_{root-fine}$, (d) $DW_{standing}$, (d) $DW_{down}$, and (d) $OL$.
+```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | Age trends and biome differences in some of the major forest C stocks: (a) $B_{ag}$, (b) $B_{foliage}$, (c) $B_{root-fine}$, (d) $DW_{standing}$, (e) $DW_{down}$, and (f) $OL$.
 The scatterplots show age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome.
@@
 C fluxes commonly increased significantly with stand age (Tables 1, S2, Figs. `r flux_age_trends`, `r age_trends_all`, `r all_age_trends`).
-*ForC* contained 16 C flux variables with sufficient data for analyses of age trends in young forests (see Methods).
+*ForC* contained 16 C flux variables with sufficient data to analyize age trends in young forests (see Methods).
 Of these, ten increased significantly with $log10[age]$: $NEP$, $GPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{foliage}$, $ANPP_{litterfal}$, $BNPP$, $BNPP_{fine}$, $R_{eco}$, and $R_{root}$.
@@
 
-"Closure" and internal consistency of the C flux budget was less successful for young than mature forests (Figs. `r age_trends_all`).
+"Closure" and internal consistency of the C flux budget were less successful for young than mature forests (Figs. `r age_trends_all`).
 Summed regression equations for $R_{soil-het}$ and $R_{root}$ were generally very close to $R_{soil}$.
@@
 `r age_trends_all`).
-In contrast, the  CO~2~ influx portion of the budget generally did not "close": the sum of $R_{auto}$ ($R_{root}+R_{auto-ag}$, as described above) and components of $NPP$ consistently fell short of $GPP$, particularly in in young stands (range across forest types and ages: 0.9-7.6 Mg C ha^-1^ yr^-1^).
+In contrast, the  CO~2~ influx portion of the budget generally did not "close": the sum of $R_{auto}$ ($R_{root}+R_{auto-ag}$, as described above) and components of $NPP$ consistently fell short of $GPP$, particularly in young stands (range across forest types and ages: 0.9-7.6 Mg C ha^-1^ yr^-1^).
 Moreover, there was not consistent budget closure among the components of $NPP$, and substantially different age trends resulting from the sum of components versus total $NPP$ (Figs.
 `r age_trends_all`).
-Although age trends of young forests often converged towards mature forest averages, there were also some discrepancies between young forest trends and mature forest averages (Figs.
-`r flux_age_trends`, `r age_trends_all`, `r all_age_trends`), most notably including a tendency for higher fluxes in regrowth boreal forests than in their mature counterparts (Fig.
-`r age_trends_all`).
+Although age trends of young forests often converged towards mature forest averages, there were some discrepancies, most notably including a tendency for higher fluxes in regrowth boreal forests than in their mature counterparts (Figs.
+`r flux_age_trends`, `r age_trends_all`, `r all_age_trends`).
 
 In terms of C stocks, ten variables (all but standing deadwood, $DW_{standing}$) had sufficient data to test for age trends (Table 1, Figs. `r stock_age_trends`, `r all_stock_age_trends`).
-All of these displayed a significant overall increase with with $log10[stand.age]$.
+All of these displayed a significant overall increase with $log10[stand.age]$.
 Age $\times$ biome interactions were also significant for all ten of these C stock variables (Table S2), with living C stocks tending to accumulate more rapidly during the early stages of forest regrowth in tropical forests (Figs.
@@
 In the case of two non-living C stocks ($DW_{down}$ and $OL$), age $\times$ biome interactions were such that age trends were positive in some biomes and negative in others.
-Specifically, $DW_{down}$ declined with age in temperate and boreal forests, compared to an increase with age in tropical forests (Figs. `r stock_age_trends`,`r age_trends_all`, `r DW_down_fig`).
+Specifically, $DW_{down}$ declined with age in temperate and boreal forests, compared to an increase in tropical forests (Figs. `r stock_age_trends`,`r age_trends_all`, `r DW_down_fig`).
 Similarly, $OL$ declined slightly with age in temperate broadleaf forests, contrasting an increase in the other three biomes (Figs. `r stock_age_trends`, `r age_trends_all`,`r OL_fig`).
-Again, there were some discrepancies between young forest trends and mature forests, most notably including generally higher C stocks in mature forests relative to their 100-year counterparts, particularly for temperate conifer forests (with discrepancies again driven by differences in geographic representation) and, to a lesser extent, tropical broadleaf forests (FIg.
+Again, there were some discrepancies between young forest trends and mature forests, most notably including generally higher C stocks in mature forests relative to their 100-year counterparts, particularly for temperate conifer forests (with discrepancies again driven by differences in geographic representation) and, to a lesser extent, tropical broadleaf forests (Fig.
 `r age_trends_all`).

### 2021-03-01 manuscript/ERL_global_C_review.Rmd
finish integrating @aferson's edits
#70

@@
 Carbon cycling rates generally increased from boreal to tropical regions and with stand age (Figs. `r schematic`, `r age_trends_all`).
-Specifically, most C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for regrowth as well as mature forests (Figs.
+Specifically, most C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for both regrowth and mature forests (Figs.
 `r schematic`, `r flux_age_trends`- `r stock_age_trends`, `r age_trends_all`).
 The notable exception was mature forest $NEP$ (the difference between $GPP$ and $R_{eco}$) which was statistically indistinguishable across biomes (Fig. `r flux_age_trends`f).
-There was also little directional variation in mean mature forest C stocks (biomass, dead wood, and organic layer) across biomes, although maximum values for the majority of stocks (all including live or standing woody biomass) occurred in temperate biomes (Figs.
+There was also little directional variation in mean mature forest C stocks (biomass, dead wood, and organic layer) across biomes.
+However, maximum values for most stocks (all including live or standing woody biomass) occurred in temperate biomes (Figs.
 `r schematic`, `r all_diagrams_mature`, `r stock_age_trends`).
@@
 `r flux_age_trends`- `r age_trends_all`, `r all_age_trends`).
-Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks and $NEP$ of mature forests, which are defined by the differences between in- and out- fluxes, do not vary systematically across biomes.
-Together, these results refine and expand out understanding of C cycling in mature forests, while providing the first global-scale analysis of age trends in multiple forest C stocks and fluxes (Figs.
+Thus, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling.
+In contrast, C stocks and $NEP$ of mature forests, which are defined by the differences between in- and out- fluxes, do not vary systematically across biomes.
+Together, these results refine and expand our understanding of C cycling in mature forests, while providing the first global-scale analysis of age trends in multiple forest C stocks and fluxes (Figs.
 `r age_trends_all`).
@@
 `r schematic`, @luyssaert_oldgrowth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
-It is therefore consistent with theory -- and with previous research [@luyssaert_co2_2007] -- that there are no pronounced differences across biomes. 
+The lack of pronounced differences across biomes is therefore consistent with both theory and previous research [e.g., @luyssaert_co2_2007]. 
 Rather, variation in $NEP$ of mature forests appears to be controlled less by climate and more by other factors including moderate disturbances [@curtis_forest_2018] or disequilibrium of $R_{soil}$ relative to C inputs [e.g., in peatlands where anoxic conditions inhibit decomposition, @wilson_stability_2016].
@@
 In contrast to C fluxes and biomass accumulation rates in regrowth forests, stocks showed less systematic variation across biomes (c.f. Fig. `r schematic`).
-For aboveground biomass, which is the variable in *ForC* with the broadest geographical representation, the modest trend of declining biomass from tropical to boreal regions mirrors observations from spaceborne lidar that reveal a decline in aboveground biomass (for all forests, including secondary) with latitude across the N hemisphere [@hu_mapping_2016].
-The highest-biomass forests on Earth are, however, found in coastal temperate climates of both the southern and northern hemisphere [Figs.
-`r schematic`, `r stock_age_trends`a, @keith_reevaluation_2009; @smithwick_potential_2002; @hu_mapping_2016; @larjavaara_temperature_2012].
-Disproportionate representation of forests in one such region--the US Pacific Northwest--inflated estimates of temperate conifer fluxes and stocks for some variables and was responsible for all of the anomalous results described here (e.g., lack of complete C budget closure, anomalous trend across biomes for $BNPP_{coarse}$).
+For aboveground biomass, which is the variable in *ForC* with the broadest geographical representation, the modest trend of declining biomass from tropical to boreal regions mirrors observations from spaceborne lidar that reveal a decline in aboveground biomass (for all forests, including secondary) with latitude across the Northern hemisphere [@hu_mapping_2016].
+The highest-biomass forests on Earth are, however, found in coastal temperate climates of both the southern and northern hemispheres [Figs.
+`r schematic`, `r stock_age_trends`a, @smithwick_potential_2002; @keith_reevaluation_2009; @larjavaara_temperature_2012; @hu_mapping_2016].
+Disproportionate representation of forests in one such region -- the US Pacific Northwest -- inflated estimates of temperate conifer fluxes and stocks for some variables and was responsible for all the anomalous results described here (e.g., lack of complete C budget closure, an anomalous trend across biomes for $BNPP_{coarse}$).
 Thus, biome differences should always be interpreted relative to the geographic distribution of sampling, which only rarely adequately represents the majority of forested area within a biome.
 
-Whereas aboveground biomass can be remotely sensed [albeit with significant uncertainties, @ploton_spatial_2020] and receives significant research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) carbon across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
-Although these stocks can be important--exceeding 100 Mg C ha^-1^ in some stands (Figs.
+Whereas aboveground biomass can be remotely sensed [albeit with significant uncertainties, @ploton_spatial_2020] and receives substantial research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) carbon across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
+Although these stocks can be important, exceeding 100 Mg C ha^-1^ in some stands (Figs.
 `r stock_age_trends`, `r DW_tot_fig`-`r DW_down_fig`), this study is the first to synthesize deadwood data on a global scale [but see @cook-patton_mapping_2020 for young forests].
@@
 `r flux_age_trends`, `r age_trends_all`), consistent with current understanding of age trends in forest C cycling [Fig.
-`r schematic`; e.g., @anderson-teixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
+`r schematic`; e.g., @magnani_human_2007; @amiro_ecosystem_2010; @anderson-teixeira_altered_2013].
 While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_{auto}$) would be minimal immediately following a stand-clearing disturbance (Fig.
@@
 `r schematic`).
-These may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots, $DW$, and $OL$ [@ribeiro-kumara_how_2020; @maurer_carbon_2016; @bond-lamberty_contribution_2004], with the latter being strongly dependent upon the type of stand initiating disturbance (discussed below).
-In this study, we detect no significant overall age trends in either $R_{het-soil}$ or $R_{soil}$, consistent with previous findings [@law_changes_2003;@pregitzer_carbon_2004;@goulden_patterns_2011].
+These may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots, $DW$, and $OL$ [@bond-lamberty_contribution_2004; @maurer_carbon_2016; @ribeiro-kumara_how_2020], with the latter being strongly dependent upon the type of stand initiating disturbance (discussed below).
+This study detects no significant overall age trends in either $R_{het-soil}$ or $R_{soil}$, consistent with previous findings [@law_changes_2003;@pregitzer_carbon_2004;@goulden_patterns_2011].
 
@@
 `r flux_age_trends`).
-This finding is consistent with previous studies showing an increase in $NEP$ across relatively young stand ages [@pregitzer_carbon_2004; @baldocchi_fluxnet_2001; @luyssaert_oldgrowth_2008].
+This finding is consistent with previous studies showing an increase in $NEP$ across relatively young stand ages [@baldocchi_fluxnet_2001; @pregitzer_carbon_2004;  @luyssaert_oldgrowth_2008].
 However, $NEP$ is theoretically expected to peak in intermediate-aged stands and thereafter decline, consistent with decelerating C accumulation as stands age [Fig.
@@
 `r age_trends_all`) may be driven by differences in geographical representation across age classes or by the fitting of an inappropriate functional form.
-Moreover, both biomass and non-living C stocks often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands [@luyssaert_oldgrowth_2008; @mcgarvey_carbon_2014; @lichstein_biomass_2009].
+Moreover, both biomass and non-living C stocks often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands [@luyssaert_oldgrowth_2008; @lichstein_biomass_2009; @mcgarvey_carbon_2014].
 Additional data, including on age trends of deadwood, the organic layer, and soil C will be important to parsing the timing and extend of an age-related $NEP$ decrease across forest biomes.
@@
 In contrast, natural disturbances (e.g., fire, drought, typhoons/hurricanes) can produce large amounts of deadwood (mostly $DW_{standing}$) that slowly decomposes as the stand recovers, resulting in declines across young stand ages [e.g., temperate and boreal stands in Figs.
-`r stock_age_trends`, `r age_trends_all`, @carmona_coarse_2002].
-Further study and synthesis of non-living C stocks across biomes, stand ages, and disturbance types will be valuable to giving a more comprehensive picture.
+`r stock_age_trends`, `r age_trends_all`; e.g., @carmona_coarse_2002].
+Further study and synthesis of non-living C stocks across biomes, stand ages, and disturbance types will be valuable in giving a more comprehensive picture.
 
@@
 The large number of C cycle variables covered by ForC, and the relatively high consistency among them (Figs.
-`r all_diagrams_mature`, `r age_trends_all`), provide confidence that our reported mature forest means provide useful baselines for analysis -- with the caveats that they are unlikely to be accurate representations of C cycling for any particular forest, and that these sample means almost certainly do not represent true biome means (particularly for temperate conifer forests where high-biomass stands are over-represented in *ForC*).
+`r all_diagrams_mature`, `r age_trends_all`), provide confidence that our reported mature forest means provide useful baselines for analysis.
+However, there is wide variation around these means, implying that any given stand could deviate substantially from them, and the sample means presented here probably do not represent true biome means (particularly for temperate conifer forests where high-biomass stands are over-represented in *ForC*).
 
@@
 On the other, however, *ForC* derives data from multiple heterogeneous sources, and standard deviations within each biome are high; as a result, the standard for C closure is relatively loose [*c.f.* @houghton_terrestrial_2020].
-The one instance where the C budgets does not close according to our criteria is likely due to differences in the representation of forest types (*i.e.*, disproportionate representation of US Pacific NW for $B_{root-coarse}$ relative to  $B_{root}$; Fig.
+The one instance where the C budgets do not close according to our criteria is likely due to differences in the representation of forest types (*i.e.*, disproportionate representation of US Pacific NW for $B_{root-coarse}$ relative to  $B_{root}$; Fig.
 `r diagram_TeN_mature`) rather than issues of methodological accuracy.
-The overall high degree of closure implies that *ForC* gives an at least roughly consistent picture of C cycling within biomes for mature forests. 
-This is an important and useful test, because it allows for consistency checks within the C cycle, for example leveraging separate and independently-measured fluxes to constrain errors in another [@phillips_value_2017; @williams_impacts_2014; @harmon_heterotrophic_2011], or producing internally consistent global data products [@wang_golumcnp_2018].
+The overall high degree of closure implies that *ForC* gives a roughly consistent picture of C cycling within biomes for mature forests. 
+This is an important and useful test because it allows for consistency checks within the C cycle, for example leveraging separate and independently measured fluxes to constrain errors in another [@williams_impacts_2014; @harmon_heterotrophic_2011; @phillips_value_2017], or producing internally consistent global data products [@wang_golumcnp_2018].
 
@@
 In large part, this is because their data records remain relatively sparse (*i.e.*, have low representation of different geographical regions for any given age) for most variables, particularly in the tropics [@anderson-teixeira_carbon_2016].
-In addition, variation in the type and severity of stand-initiating disturbances introduces significant heterogeneity in both initial values and age trends of C cycle variables [e.g., @cook-patton_mapping_2020]
-While this review provides a first analysis of age trends in forest C cycling for multiple variables at a global scale, improved resolution of these trends will require additional data.
+In addition, variation in the type and severity of stand-initiating disturbances introduces significant heterogeneity in both initial values and age trends of C cycle variables [e.g., @cook-patton_mapping_2020].
+While this review provides the first analysis of age trends in forest C cycling for multiple variables at a global scale, improved resolution of these trends will require additional data.
 
-There are of course notable holes in the ForC variable coverage that limit the scope of our inferences here. 
+There are, of course, notable holes in the ForC variable coverage that limit the scope of our inferences here. 
 Notably, *ForC* currently has sparse--if any--coverage of fluxes to herbivores and higher consumers, along with woody mortality ($M_{woody}$) and $DW$ (Table 1, Figs. `r DW_tot_fig`-`r DW_down_fig`). 
@@
 Importantly, *ForC* and the analyses presented here cover the forests that have received research attention, which are not a representative sample of the world's existing forests, either geographically or in terms of human impacts [@martin_mapping_2012].
-Geographically, all variables are poorly covered in Africa and Siberia (Fig. `r figure_map`), a common problem in the carbon-cycle community [@xu_contribution_2016; @schimel_effect_2015]. 
+Geographically, all variables are poorly covered in Africa and Siberia (Fig. `r figure_map`), a common problem in the carbon-cycle community [@schimel_effect_2015; @xu_contribution_2016]. 
 In terms of human impacts, research efforts tend to focus on interior forest ecosystems [@martin_mapping_2012], often in permanently protected areas [e.g., @davies_forestgeo_2021].
@@
 Yet, fragmentation and degradation impact a large and growing proportion of Earth's forests [@faoandunep_state_2020].
-Fragmentation and the creation of edges strongly impacts forest C cycling [e.g., @chaplin-kramer_degradation_2015; @remy_strong_2016; @reinmann_edge_2017; @smith_evidence_2019; @reinmann_urbanization_2020;  @ordway_carbon_2020].
+Fragmentation and the creation of edges strongly impact forest C cycling [e.g., @chaplin-kramer_degradation_2015; @remy_strong_2016; @reinmann_edge_2017; @smith_evidence_2019;  @ordway_carbon_2020; @reinmann_urbanization_2020].
 Partial logging and other forms of non- stand clearing anthropogenic disturbance also alter forest C cycling [e.g., @huang_longterm_2010; @piponiot_carbon_2016], but are under-studied [@sist_tropical_2015] and excluded from this analysis.
 Fragmented and degraded forests do not fit the idealized conceptual framework around which this review is structured (Fig.
-`r schematic`), yet their representation in models, sustainability assessments, and C accounting systems is critical to accurate accounting of C cycling in Earth's forests [e.g., @huang_longterm_2010; @reinmann_edge_2017; @smith_evidence_2019; @piponiot_can_2019].
+`r schematic`), yet their representation in models, sustainability assessments, and C accounting systems is critical to accurate accounting of C cycling in Earth's forests [e.g., @huang_longterm_2010; @reinmann_edge_2017; @piponiot_can_2019; @smith_evidence_2019].
 Finally, plantation forests account for approximately 3% of Earth's forests [@faoandunep_state_2020] but are not included in this analysis. 
@@
 The future of forest C cycling [@song_metaanalysis_2019] will shape trends in atmospheric CO~2~ and the course of climate change [@schimel_effect_2015].
-Our findings, and more generally the data contained in *ForC* and summarized here, can help to meet two major challenges.
+Our findings, and more generally the data contained in *ForC* and summarized here, can help meet two significant challenges.
 
-First, improved representation of forest C cycling in models is essential to improving predictions of the future course of climate change, for the simple reason that by definition future projections extend our existing observations and understanding to conditions that do not currently exist on Earth [@mcdowell_predicting_2018; @bonan_climate_2018; @gustafson_extrapolating_2018].
+First, improved representation of forest C cycling in models is essential to improving predictions of the future course of climate change. 
+By definition, future projections extend our existing observations and understanding to conditions that do not currently exist on Earth [@bonan_climate_2018; @gustafson_extrapolating_2018; @mcdowell_predicting_2018].
 To ensure that models are giving the right answers for the right reasons [@sulman_multiple_2018], it is important to benchmark against multiple components of the C cycle that are internally consistent with each other [@collier_international_2018; @wang_golumcnp_2018].
-*ForC*’s tens of thousands of records are readily available in a standardized format, along with all code used in the analyses presented here, and we recommend that researchers use these resources to identify and summarize data specific to the analysis at hand.
-Integration of *ForC* with models will be valuable to improving the accuracy and reliability of models [@fer_ecosystem_2021].
+*ForC*’s tens of thousands of records are readily available in a standardized format, along with all code used in the analyses presented here
+We recommend that researchers use these resources to identify and summarize data specific to the analysis at hand. 
+Integration of *ForC* with predictive models will be valuable to improving model accuracy and reliability [@fer_ecosystem_2021].
 
-Second, *ForC* can serve as a pipeline through which information can flow efficiently from forest researchers to decision-makers working to implement forest conservation strategies at global, national, or landscape scales.
+Second, *ForC* can serve as a pipeline through which information can flow efficiently from forest researchers to decision-makers and practitioners working to implement forest conservation strategies at global, national, or landscape scales.
 This is already happening: *ForC* has contributed to updating the IPCC guidelines for carbon accounting in forests [@ipcc_2019_2019; @requenasuarez_estimating_2019], mapping C accumulation potential from natural forest regrowth globally [@cook-patton_mapping_2020], and informing ecosystem conservation priorities [@goldstein_protecting_2020].
@@
 It is also interesting to consider the complementary utility of global-scale but spatially discontinuous databases such as ForC and wall-to-wall remote sensing products. 
-The latter provide insight into aboveground carbon stocks, but less constraint on belowground stocks or carbon fluxes in general [@bond-lamberty_estimating_2016; @anav_spatiotemporal_2015]. 
-Combining observational data and remote observations may provide a much more comprehensive and accurate picture of global forest C cycling, particularly when used in formal data assimilation systems [@konings_global_2019; @liu_detecting_2018].
+The latter provide insight, with substantial uncertainty, into aboveground carbon stocks and $GPP$, but are less useful for constraining belowground stocks or carbon fluxes in general [@anav_spatiotemporal_2015; @bond-lamberty_estimating_2016].
+Combining observational data and remote observations may provide a much more comprehensive and accurate picture of global forest C cycling, particularly when used in formal data assimilation systems [@liu_detecting_2018;@konings_global_2019].
 Biomass is the largest C stock in most forests, and most of the emphasis has traditionally been on this variable. 
-Remote-sensing driven aboveground biomass estimates [e.g., @saatchi_benchmark_2011], calibrated based on high-quality ground-based data [@schepaschenko_forest_2019; @chave_ground_2019], provide the most promising approach, but significant uncertainties remain [@ploton_spatial_2020].
-Note, however, that factors such as stand age and disturbance history are difficult, if not impossible, to detect remotely, and can only be characterized for recent decades [@hansen_highresolution_2013; @song_global_2018; @curtis_classifying_2018].
+Remote-sensing driven aboveground biomass estimates [e.g., @saatchi_benchmark_2011], calibrated based on high-quality ground-based data [@chave_ground_2019; @schepaschenko_forest_2019], provide the most promising approach, but significant uncertainties remain [@ploton_spatial_2020].
+Note, however, that factors such as stand age and disturbance history are difficult, if not impossible, to detect remotely, and can only be characterized for recent decades [@hansen_highresolution_2013; @curtis_classifying_2018; @song_global_2018].
 Ground-based data such as *ForC* are therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020, and thus constraining variables such as carbon sink potential [@luyssaert_oldgrowth_2008].
@@
 In contrast, carbon allocation within forest ecosystems and respiration fluxes cannot be remotely sensed. 
-Efforts such as the Global Carbon Project [@friedlingstein_global_2019] and NASA’s Carbon Monitoring System [@liu_detecting_2018] typically compute respiration as residuals of all other terms [@bond-lamberty_estimating_2016; @harmon_heterotrophic_2011].
+Efforts such as the Global Carbon Project [@friedlingstein_global_2019] and NASA’s Carbon Monitoring System [@liu_detecting_2018] typically compute respiration as residuals of all other terms [@harmon_heterotrophic_2011; @bond-lamberty_estimating_2016].
 This means that the errors on respiration outputs are likely to be large and certainly poorly constrained, offering a unique opportunity for databases such as ForC and SRDB [@jian_restructured_2020] to provide observational benchmarks.
@@
 ### Conclusions
-As climate change accelerates, understanding and managing the carbon dynamics of forests-- including dynamics and fluxes that cannot be observed by satellites--is critical to forecasting, mitigation, and adaptation.
+As climate change accelerates, understanding and managing the carbon dynamics of forests-- including stocks and fluxes that satellites cannot observe--is critical to forecasting, mitigation, and adaptation.
 The C data in *ForC*, as summarized here, are valuable to these efforts.

### 2021-03-01 manuscript/tables_figures/stacked_age_trends/schematic.m
added letters to schematic
#71

@@ -111,6 +111,7 @@ ylabel ('C fluxes (Mg C ha^{-1} yr^{-1})');
 set(gca, 'YTick', []); %ticks off
 set(gca, 'XTickLabel', {'Tropical' 'Temperate' 'Boreal'})
 xtickangle(30)
+text(.05,.9,'A.', 'Units','Normalized' ) 
 box on
 
 subplot ('Position', pos_age_f)
@@ -139,6 +140,7 @@ set(gca, 'YTick', []); %ticks off
 ylim([0, max(in_sum)+1]);
 legend ([{'R_{het}'}, in_flux_names, {'GPP' 'R_{eco}' 'legacy R_{eco}'}], 'Location', 'BestOutside');
 legend('Boxoff')
+text(.05,.9,'B.', 'Units','Normalized' ) 
 
 subplot ('Position', pos_biome_s)
 h=bar(mature_stocks, 'stacked'); hold on;
@@ -153,6 +155,7 @@ ylabel ('C  stocks (Mg C ha^{-1})')
 set(gca, 'YTick', []); %ticks off
 set(gca, 'XTickLabel', {'Tropical' 'Temperate' 'Boreal'})
 xtickangle(30)
+text(.05,.9,'C.', 'Units','Normalized' ) 
 
 subplot ('Position', pos_age_s) %stocks by age:
 
@@ -171,6 +174,7 @@ set(gca, 'YTick', []); %ticks off
 
 legend ([stock_names,{'B_{tot}', 'legacy DW_{tot}'}], 'Location', 'BestOutside');
 legend('Boxoff')
+text(.05,.9,'D.', 'Units','Normalized' ) 
 
 %% ~~~~~~~ SAVE FIGURE~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 print('schematic', '-dpng', '-r600')

### 2021-03-01 manuscript/ERL_global_C_review.Rmd
add letters to references of Fig. 1
#71

@@
 Global scale data syntheses have shown that C fluxes including $GPP$, net primary productivity ($NPP$), and soil respiration ($R_{soil}$) decrease with latitude or, correspondingly, increase with mean annual temperature [Fig.
-`r schematic`; e.g., @lieth_primary_1973;  @luyssaert_co2_2007; @hursh_sensitivity_2017; @banburymorgan_global_inpress].
-C stocks of mature forests show less directional variation (Fig. `r schematic`).
+`r schematic`a; e.g., @lieth_primary_1973;  @luyssaert_co2_2007; @hursh_sensitivity_2017; @banburymorgan_global_inpress].
+C stocks of mature forests show less directional variation (Fig. `r schematic`c).
 On average, aboveground biomass ($B_{ag}$) tends to decrease with latitude, but not as dramatically as fluxes, and with the highest $B_{ag}$ forests in relatively cool, moist temperate regions [@smithwick_potential_2002; @keith_reevaluation_2009; @hu_mapping_2016].
@@
 
-Stand age is another important axis of variation in forest C cycling (Fig. `r schematic`).
+Stand age is another important axis of variation in forest C cycling (Fig. `r schematic`b,d).
 In 1969, E.P. Odum's "The Strategy of Ecosystem Development" laid out predictions as to how forest energy flows and organic matter stocks vary with stand age  [@odum_strategy_1969].
 Although the conceptualization of the C cycle in this paper was simplistic by current standards, the paper was foundational in framing the theory around which research on the subject still revolves [@corman_foundations_2019], and the basic framework still holds, albeit with modest modifications [Fig.
-`r schematic`, @anderson-teixeira_altered_2013].
+`r schematic`b, @anderson-teixeira_altered_2013].
 Following stand-clearing disturbance, $GPP$, $NPP$, and biomass of leaves ($B_{foliage}$) and fine roots ($B_{root-fine}$) initially increase rapidly and thereafter remain relatively stable ($B_{foliage}$, $B_{root-fine}$, sometimes $GPP$) or decline slightly [$NPP$, sometimes $GPP$; e.g., @law_changes_2003; @pregitzer_carbon_2004; @amiro_ecosystem_2010; @goulden_patterns_2011].
@@
 When the effect of stand age was significant at p $\le$ 0.05, and when each biome had records for stands of at least ten different ages, a biome $\times$ stand age interaction was included in the model.
-We note that the logarithmic function fit in this analysis does not always correspond to theoretical expectations (Fig.
-`r schematic`); however, data limitations did not support fitting of functions with more parameters or reliable comparisons of different functional forms.
+We note that the logarithmic function fit in this analysis does not always correspond to theoretical expectations, particularly for $NEP$ (Fig.
+`r schematic`b); however, data limitations did not support fitting of functions with more parameters or reliable comparisons of different functional forms.
 Within the data constraints, we deemed a logarithmic function to be the appropriate functional form for most variables.
@@
 However, maximum values for these variables -- along with all other  stocks including live or standing woody biomass ($B_{ag-wood}$, $B_{root}$, $B_{root-coarse}$, $DW_{tot}$, $DW_{standing}$) -- consistently occurred in temperate biomes (Figs.
-`r schematic`, `r stock_age_trends`, `r all_stock_age_trends`).
+`r schematic`c, `r stock_age_trends`, `r all_stock_age_trends`).
 For variables for which temperate conifer forest records were disproportionately from high-biomass forests in the US Pacific Northwest ($B_{ag-wood}$, $B_{foliage}$, and $B_{root-coarse}$), temperate conifer forests had significantly higher stocks than other biomes.
@@
 Specifically, most C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for both regrowth and mature forests (Figs.
-`r schematic`, `r flux_age_trends`- `r stock_age_trends`, `r age_trends_all`).
+`r schematic`a, `r flux_age_trends`- `r stock_age_trends`, `r age_trends_all`).
 The notable exception was mature forest $NEP$ (the difference between $GPP$ and $R_{eco}$) which was statistically indistinguishable across biomes (Fig. `r flux_age_trends`f).
@@
 However, maximum values for most stocks (all including live or standing woody biomass) occurred in temperate biomes (Figs.
-`r schematic`, `r all_diagrams_mature`, `r stock_age_trends`).
+`r schematic`c, `r all_diagrams_mature`, `r stock_age_trends`).
 Consistent with theory and previous studies (Fig.
-`r schematic`), the majority of flux variables, together with most live biomass pools, increased significantly with stand age (Table 1; Figs.
+`r schematic`b,d), the majority of autotrophic C fluxes, together with most live biomass pools, increased significantly with stand age (Table 1; Figs.
 `r flux_age_trends`- `r age_trends_all`, `r all_age_trends`).
@@
 The finding that these patterns hold consistently across numerous fluxes, while consistent with theoretical expectations (Fig.
-`r schematic`), is novel to this analysis [but see @banburymorgan_global_inpress for nine autotrophic fluxes].
+`r schematic`a), is novel to this analysis [but see @banburymorgan_global_inpress for nine autotrophic fluxes].
 
@@
 Unlike the other C flux variables, $NEP$ does not characterize the rate at which C cycles through the ecosystem, but, as the balance between $GPP$ and $R_{eco}$, represents net CO~2~ sequestration (or release) by the ecosystem (Fig.
-`r schematic`).
+`r schematic`a-b).
 $NEP$ tends to be relatively small in mature forest stands, which accumulate carbon slowly relative to younger stands, if at all [Fig.
-`r schematic`, @luyssaert_oldgrowth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
+`r schematic`a-b, @luyssaert_oldgrowth_2008; @amiro_ecosystem_2010; @besnard_quantifying_2018].
 The lack of pronounced differences across biomes is therefore consistent with both theory and previous research [e.g., @luyssaert_co2_2007]. 
@@
 Biome differences in the accumulation rates of $DW$, $OL$, and soil C  have not been detected, in part because these variables do not consistently increase with stand age [Figs.
-`r schematic`, `r stock_age_trends`, `r DW_tot_fig`-`r OL_fig`, and see discussion below, @cook-patton_mapping_2020].
+`r schematic`d, `r stock_age_trends`, `r DW_tot_fig`-`r OL_fig`, and see discussion below, @cook-patton_mapping_2020].
 
@@
 
-In contrast to C fluxes and biomass accumulation rates in regrowth forests, stocks showed less systematic variation across biomes (c.f. Fig. `r schematic`).
+In contrast to C fluxes and biomass accumulation rates in regrowth forests, stocks showed less systematic variation across biomes (c.f. Fig. `r schematic`c).
 For aboveground biomass, which is the variable in *ForC* with the broadest geographical representation, the modest trend of declining biomass from tropical to boreal regions mirrors observations from spaceborne lidar that reveal a decline in aboveground biomass (for all forests, including secondary) with latitude across the Northern hemisphere [@hu_mapping_2016].
 The highest-biomass forests on Earth are, however, found in coastal temperate climates of both the southern and northern hemispheres [Figs.
-`r schematic`, `r stock_age_trends`a, @smithwick_potential_2002; @keith_reevaluation_2009; @larjavaara_temperature_2012; @hu_mapping_2016].
+`r schematic`c, `r stock_age_trends`a, @smithwick_potential_2002; @keith_reevaluation_2009; @larjavaara_temperature_2012; @hu_mapping_2016].
 Disproportionate representation of forests in one such region -- the US Pacific Northwest -- inflated estimates of temperate conifer fluxes and stocks for some variables and was responsible for all the anomalous results described here (e.g., lack of complete C budget closure, an anomalous trend across biomes for $BNPP_{coarse}$).
@@
 *ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that proportionally slower decomposition in colder climates results in more buildup of organic matter [Fig.
-`r schematic`, @allen_global_2002; @anderson-teixeira_differential_2011].
+`r schematic`c, @allen_global_2002; @anderson-teixeira_differential_2011].
 Further research on non-living C stocks in the world's forests will be essential to completing the picture.
@@
 `r flux_age_trends`, `r age_trends_all`), consistent with current understanding of age trends in forest C cycling [Fig.
-`r schematic`; e.g., @magnani_human_2007; @amiro_ecosystem_2010; @anderson-teixeira_altered_2013].
+`r schematic`b; e.g., @magnani_human_2007; @amiro_ecosystem_2010; @anderson-teixeira_altered_2013].
 While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_{auto}$) would be minimal immediately following a stand-clearing disturbance (Fig.
-`r schematic`).
+`r schematic`b).
 These would be expected to increase rapidly, along with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Figs.
-`r schematic`, `r flux_age_trends`-`r age_trends_all`).
+`r schematic`b,d, `r flux_age_trends`-`r age_trends_all`).
 In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$) -- and therefore $R_{eco}$ -- are expected to be non-zero following stand-clearing disturbance (Fig.
-`r schematic`).
+`r schematic`b).
 These may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots, $DW$, and $OL$ [@bond-lamberty_contribution_2004; @maurer_carbon_2016; @ribeiro-kumara_how_2020], with the latter being strongly dependent upon the type of stand initiating disturbance (discussed below).

### 2021-03-01 manuscript/ERL_global_C_review.Rmd
adjust Fig 1
#71

@@
 
-```{r schematic, eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 1 | Schematic diagram summarizing current understanding of biome differences and age trends in forest C cycling.
+```{r schematic, eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 1 | Schematic diagram summarizing current understanding of biome differences (a,c) and age trends (b,d) in forest C cycling.
 Variables are defined in Table 1.
 Age trends, which represent idealized dynamics following a disturbance that removes all living vegetation, are an updated version of the classic figure from Odum (1969), with heavy lines corresponding to those in Odum's figure and $NEP$ corresponding to Odum's 'net production'.
-Here, $NEP$ consists primarily of woody aboveground net primary production ($ANPP_{woody}$), while $ANPP_{woody.turnover}$ is the sum of woody mortality and branch turnover.
+*Here, positive $NEP$ of young forests is typically dominated by woody net primary production ($NPP_{woody} = ANPP_{woody} + BNPP_{coarse}$).
+As forests age and biomass accumulation slows, $NPP_{woody}$ approaches equilibrium with woody turnover ($M_{woody}$ + $ANPP_{branch}$ + coarse root turnover), and $NEP$ may be dominated by changes in dead wood or soil organic carbon.
 Dotted lines refer to decomposition of potential 'legacy' organic material produced prior to the disturbance and remaining at the site (e.g., standing and fallen dead wood, $DW_{tot}$; soil organic matter).

### 2021-03-01 manuscript/tables_figures/stacked_age_trends/schematic.m

@@ -9,7 +9,7 @@ set(0,'DefaultLegendAutoUpdate','off');
     %subplot positions
     plot_width_biome=0.27; 
     plot_width_age_f=0.57; 
-    plot_width_age_s=0.515;
+    plot_width_age_s=0.507;
     plot_height=0.37;
     left_marg=.09;
     lower_marg=.1;
@@ -46,16 +46,16 @@ GPP= NEP + (1.99+1.44)* log10(age);
 ANPP_foliage = 1.66* log10(age) +.18 - .01*(age-1);
 ANPP_woody = 2.31* log10(age) -.76 - .02*(age-1);
 ANPP_woody_turnover=1.7*log10(age)-1; %max(0,ANPP_woody-NEP);
-BNPP=0.97*log10(age)+.67 - .01*(age-1);
-NPP=ANPP_foliage+ANPP_woody+BNPP;
+BNPP_fine=1.46*log10(age) - .01*(age-1);
+NPP=ANPP_foliage+ANPP_woody+BNPP_fine; %missing BNPP_coarse, but equations for BNPP_fine and BNPP are very similar
 CUE=0.679-0.153*log10(age); %from DeLucia et al. 2007 (note that equation given mistakenly leaves of the log10(age)). Collati et al. 2020 gives an update of this.
 R_auto= max(0,NPP.*(1./CUE-1)); %Rauto=NPP*(1/CUE-1), and CUE equation is above
-R_het=ANPP_foliage+ANPP_woody_turnover+BNPP;
+R_het=ANPP_foliage+ANPP_woody_turnover+BNPP_fine;
 
 
 %categorize by "in" (GPP components, including NEP) and "out" (Reco components)
-in_fluxes = [R_auto; BNPP; ANPP_foliage;  ANPP_woody_turnover; NEP]; %matrix with all fluxes for stacked plot
-in_flux_names = {'R_{auto}', 'BNPP', 'ANPP_{foliage}', 'ANPP_{woody.turnover}',  'NEP'};
+in_fluxes = [R_auto; BNPP_fine; ANPP_foliage;  ANPP_woody_turnover; NEP]; %matrix with all fluxes for stacked plot
+in_flux_names = {'R_{auto}', 'BNPP_{fine}', 'ANPP_{foliage}', 'woody turnover*',  'NEP (incl. NPP_{woody})*'};
 out_fluxes = [R_auto; R_het];
 out_flux_names = {'R_{auto}', 'R_{het}'};
 

### 2021-03-01 manuscript/ERL_global_C_review.Rmd
proofread
#70

@@
 S5-S30).
-Mature forests are defined  as ≥ 100 years old and with no known major natural or anthropogenic disturbance in that time."
+Mature forests are defined  as $\\ge$ 100 years old and with no known major natural or anthropogenic disturbance in that time."
 TrB_mature_caption = paste( "Figure 3 | C cycle diagram for mature tropical broadleaf forests. ", schematics_caption_generic)  
@@
 Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide, CO~2~ [@bonan_forests_2008; @ipcc_global_2018; @friedlingstein_global_2019].
-Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climate_2006; @krause_large_2018; @bonan_model_2019; @divittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011; @ipcc_2019_2019] can be traced to lack of understanding on how C cycling varies across forest types and in relation to stand history.
-This requires accessible, comprehensive, and large-scale databases with global coverage, which runs contrary to the traditional way forest C stocks and fluxes have been measured and published.
-Large-scale synthesis is critical to benchmarking model performance with global data [@luo_framework_2012; @clark_field_2017], quantifying the role of forests in the global C cycle [e.g., @pan_large_2011], and using book-keeping methods to quantify actual or potential exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
+Despite the centrality of forest C cycling in regulating atmospheric CO~2~, gaps in our understanding of how C cycling varies across forest types and in relation to stand history underly important uncertainties in climate models [@friedlingstein_climate_2006; @krause_large_2018; @bonan_model_2019; @divittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011; @ipcc_2019_2019].
+Improved understanding of forest C cycling globally require accessible, comprehensive, and large-scale databases with worldwide coverage, which runs contrary to the traditional way forest C stocks and fluxes have been measured and published.
+Large-scale synthesis is critical to benchmarking model performance with global data [@luo_framework_2012; @clark_field_2017; @fer_ecosystem_2021], quantifying the role of forests in the global C cycle [e.g., @pan_large_2011;@harris_global_2021], and using book-keeping methods to quantify actual or potential exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].
 
@@
 For more than half a century, researchers have sought to understand how forest carbon cycling varies across stands, including among biomes [e.g., @lieth_primary_1973; @luyssaert_co2_2007] and with stand age [e.g., @odum_strategy_1969; @luyssaert_oldgrowth_2008].
-Over this time, an increasingly refined conceptual understanding of the elements of ecosystem C cycles has developed, as a growing number of variables have been defined [e.g., @chapin_reconciling_2006], along with appropriate measurement methods [e.g., @clark_measuring_2001].
+Over this time, an increasingly refined conceptual understanding of the elements of ecosystem C cycles has developed, as a growing number of variables have been defined, along with appropriate measurement methods [e.g., @clark_measuring_2001; @chapin_reconciling_2006].
 New technology has also enabled researchers to directly measure an expanding set of variables, notably including continuous measurements of soil CO~2~ efflux [@kuzyakov_sources_2006] and ecosystem-atmosphere CO~2~ exchange [@baldocchi_fluxnet_2001].
@@
 Remote sensing technology has become increasingly useful for global- or regional-scale estimates of a few critical variables, including aboveground biomass [$B_{ag}$, @saatchi_benchmark_2011; @hu_mapping_2016; @spawn_harmonized_2020] and $GPP$ [@li_mapping_2019].
-Yet, measurement and validation of most forest C stocks and fluxes necessarily requires intensive on-the-ground data collection.
+Yet, measurement and validation of most forest C stocks and fluxes requires intensive on-the-ground data collection.
 
@@
 C stocks of mature forests show less directional variation (Fig. `r schematic`c).
-On average, aboveground biomass ($B_{ag}$) tends to decrease with latitude, but not as dramatically as fluxes, and with the highest $B_{ag}$ forests in relatively cool, moist temperate regions [@smithwick_potential_2002; @keith_reevaluation_2009; @hu_mapping_2016].
+On average, aboveground biomass ($B_{ag}$) tends to decrease with latitude, but not as dramatically as fluxes, and with the highest biomass forests in relatively cool, moist temperate regions [@smithwick_potential_2002; @keith_reevaluation_2009; @hu_mapping_2016].
 In contrast, standing and downed dead wood ($DW_{standing}$ and $DW_{down}$, respectively, summing to $DW_{tot}$) and the organic layer ($OL$) tend to accumulate more in colder climates where decomposition is slow relative to $NPP$ [@harmon_ecology_1986; @allen_global_2002].
@@
 However, all such analyses are ultimately constrained by the quality and coverage of ground-based estimates of forest C fluxes or stocks to train models [e.g., @schepaschenko_forest_2019]. 
-While estimates of some variables (e.g., $B_{ag}$, $GPP$, $NPP$, $R_{soil}$) are widely available, many remain poorly characterized (e.g., $DW_{tot}$; $OL$; autotrophic respiration, $R_{auto}$) even at the coarse resolution of biomes.
+While estimates of some variables (e.g., $B_{ag}$, $GPP$, $NPP$, $R_{soil}$) are widely available, many remain poorly characterized (e.g., $DW_{tot}$; $OL$; autotrophic respiration, $R_{auto}$), even at the coarse resolution of biomes.
 This is a critical limitation for understanding forest C cycling and quantifying forest-based climate change mitigation potential across forest biomes or ecozones [e.g., @ipcc_2019_2019].
@@
 The decline in $NPP$ occurs because $R_{auto}$ increases relative to $GPP$ as forests age, corresponding to declining carbon use efficiency with stand age [@delucia_forest_2007; @collalti_forest_2020].
-Heterotrophic respiration, most of which originates from the soil ($R_{het-soil}$) remains relatively constant with stand age [@law_changes_2003; @pregitzer_carbon_2004; @goulden_patterns_2011], with the result that net ecosystem production ($NEP=GPP-R_{eco}$, where $R_{eco}$ is total ecosystem respiration) is initially negative, increases to a maximum at intermediate ages, and declines--typically to a small positive value--thereafter [@law_changes_2003; @pregitzer_carbon_2004; @luyssaert_oldgrowth_2008; @amiro_ecosystem_2010; @goulden_patterns_2011].
+Heterotrophic respiration, most of which originates from the soil ($R_{het-soil}$), remains relatively constant with stand age [@law_changes_2003; @pregitzer_carbon_2004; @goulden_patterns_2011]. 
+As a result, net ecosystem production ($NEP=GPP-R_{eco}$, where $R_{eco}$ is total ecosystem respiration) is initially negative, increases to a maximum at intermediate ages, and thereafter declines -- typically to a small positive value [@law_changes_2003; @pregitzer_carbon_2004; @luyssaert_oldgrowth_2008; @amiro_ecosystem_2010; @goulden_patterns_2011].
 The result is that biomass accumulation is rapid in young forests, followed by a slow decline to near zero in old forests [e.g., @lichstein_biomass_2009; @yang_carbon_2011].
@@
 Variables are defined in Table 1.
-Age trends, which represent idealized dynamics following a disturbance that removes all living vegetation, are an updated version of the classic figure from Odum (1969), with heavy lines corresponding to those in Odum's figure and $NEP$ corresponding to Odum's 'net production'.
-*Here, positive $NEP$ of young forests is typically dominated by woody net primary production ($NPP_{woody} = ANPP_{woody} + BNPP_{coarse}$).
+Age trends, which represent idealized dynamics following a disturbance that removes all living vegetation, are an updated version of the classic figure from Odum (1969), with heavy lines (b,d) corresponding to those in Odum's figure 1 and $NEP$ corresponding to Odum's 'net production' (b).
+*Positive $NEP$ of young forests is typically dominated by woody NPP ($NPP_{woody} = ANPP_{woody} + BNPP_{coarse}$).
 As forests age and biomass accumulation slows, $NPP_{woody}$ approaches equilibrium with woody turnover ($M_{woody}$ + $ANPP_{branch}$ + coarse root turnover), and $NEP$ may be dominated by changes in dead wood or soil organic carbon.
@@
 
-In the past few decades, researchers have started asking how age trends--mostly in $B_{ag}$ or total biomass ($B_{tot}$) accumulation-- vary across biomes.
+In the past few decades, researchers have started asking how age trends -- mostly in $B_{ag}$ or total biomass ($B_{tot}$) accumulation -- vary across biomes.
 Early research on this theme showed that biomass accumulation rates during secondary succession increase with temperature on a global scale [@johnson_postdisturbance_2000; @anderson_temperaturedependence_2006] and with water availability in the neotropics [@poorter_biomass_2016].
@@
 
-Here, we conduct a data-based review of carbon cycling from a stand to global level, and by biome and stand age, using our open-source Global Carbon Forest database (*ForC*; Fig.
-`r figure_map`), which is the largest global compilation of forest carbon data.
+Here, we conduct a data-based review of carbon cycling from a stand to global level, and by biome and stand age, using our open-source Global Carbon Forest database (*ForC*; Fig. `r figure_map`).
 Our goal is to provide a comprehensive synthesis on broad trends in forest C cycling that can serve as a foundation for improved understanding of global forest C cycling and highlight where key sources of uncertainty still reside.
@@
 A record of data sets added to *ForC* over the course of its development is available at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv.
-The database version used for this analysis has been tagged as a new release on Github (v3.0) and assigned a DOI through Zenodo (DOI: TBD).
+The database version used for this analysis has been tagged as a new release on Github (v3.0) and assigned a DOI through Zenodo (DOI: 10.5281/zenodo.4571538).
 
@@
 Assigning stands to these groupings required excluding records for which *ForC* lacked geographic coordinates (`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in the full database) or records of stand age or forest maturity (`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in the full database).
-We also excluded records of stand age = 0 year (`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
+We also excluded records with stand age of zero years (`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
 In total, our analysis retained `r n_records` records. 
@@
 C fluxes commonly increased significantly with stand age (Tables 1, S2, Figs. `r flux_age_trends`, `r age_trends_all`, `r all_age_trends`).
-*ForC* contained 16 C flux variables with sufficient data to analyize age trends in young forests (see Methods).
+*ForC* contained 16 C flux variables with sufficient data to analyze age trends in young forests (see Methods).
 Of these, ten increased significantly with $log10[age]$: $NEP$, $GPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{foliage}$, $ANPP_{litterfal}$, $BNPP$, $BNPP_{fine}$, $R_{eco}$, and $R_{root}$.
@@
 In terms of C stocks, ten variables (all but standing deadwood, $DW_{standing}$) had sufficient data to test for age trends (Table 1, Figs. `r stock_age_trends`, `r all_stock_age_trends`).
-All of these displayed a significant overall increase with $log10[stand.age]$.
+All of these displayed a significant overall increase with the logarithm of stand age.
 Age $\times$ biome interactions were also significant for all ten of these C stock variables (Table S2), with living C stocks tending to accumulate more rapidly during the early stages of forest regrowth in tropical forests (Figs.
@@
 *ForC* v3.0 provided unprecedented coverage of most major variables, yielding a broad picture of C cycling in the world's major forest biomes.
-Carbon cycling rates generally increased from boreal to tropical regions and with stand age (Figs. `r schematic`, `r age_trends_all`).
-Specifically, most C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for both regrowth and mature forests (Figs.
-`r schematic`a, `r flux_age_trends`- `r stock_age_trends`, `r age_trends_all`).
-The notable exception was mature forest $NEP$ (the difference between $GPP$ and $R_{eco}$) which was statistically indistinguishable across biomes (Fig. `r flux_age_trends`f).
-There was also little directional variation in mean mature forest C stocks (biomass, dead wood, and organic layer) across biomes.
-However, maximum values for most stocks (all including live or standing woody biomass) occurred in temperate biomes (Figs.
-`r schematic`c, `r all_diagrams_mature`, `r stock_age_trends`).
+Carbon cycling rates generally decreased from tropical to boreal climates in both mature and regrowth forests (Figs. `r schematic`, `r flux_age_trends`-`r age_trends_all`).
+In contrast,  mature forest C stocks (biomass, dead wood, and organic layer) and $NEP$, which are defined by the differences between in- and out- fluxes, exhibited little systematic variation across biomes (Figs.
+`r schematic`, `r all_diagrams_mature`, `r stock_age_trends`).
 Consistent with theory and previous studies (Fig.
@@
 `r flux_age_trends`- `r age_trends_all`, `r all_age_trends`).
-Thus, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling.
-In contrast, C stocks and $NEP$ of mature forests, which are defined by the differences between in- and out- fluxes, do not vary systematically across biomes.
-Together, these results refine and expand our understanding of C cycling in mature forests, while providing the first global-scale analysis of age trends in multiple forest C stocks and fluxes (Figs.
+Together, these results refine and expand our understanding of C cycling in mature forests, while providing the first global-scale analysis of age trends in multiple forest C stocks and fluxes (Fig.
 `r age_trends_all`).
@@
 
-Our analysis reveals that carbon cycling is most rapid in the tropics and slowest in boreal regions.
+Our analysis revealed that most C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for both regrowth and mature forests (Figs.
+`r schematic`a, `r flux_age_trends`- `r age_trends_all`).
 For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude and increase with temperature on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banburymorgan_global_inpress].
 This consistency is not surprising, particularly given commonality in the data analyzed or used for calibration.
-The finding that these patterns hold consistently across numerous fluxes, while consistent with theoretical expectations (Fig.
+The finding that these patterns hold consistently across numerous fluxes, while aligning with theoretical expectations (Fig.
 `r schematic`a), is novel to this analysis [but see @banburymorgan_global_inpress for nine autotrophic fluxes].
 
-The notable exception to the pattern of fluxes decreasing from tropical to boreal regions is $NEP$, which showed no significant differences across biomes, albeit with the highest mean in temperate broadleaf forests (Fig.
+The notable exception to the pattern of fluxes decreasing from tropical to boreal regions was $NEP$ (the difference between $GPP$ and $R_{eco}$), which showed no significant differences across biomes, albeit with the highest mean in temperate broadleaf forests (Fig.
 `r flux_age_trends`f).
-Unlike the other C flux variables, $NEP$ does not characterize the rate at which C cycles through the ecosystem, but, as the balance between $GPP$ and $R_{eco}$, represents net CO~2~ sequestration (or release) by the ecosystem (Fig.
-`r schematic`a-b).
+Unlike the other C flux variables, $NEP$ does not characterize the rate at which C cycles through the ecosystem, but, as the balance between $GPP$ and $R_{eco}$, represents net CO~2~ sequestration (or release) by the ecosystem.
 $NEP$ tends to be relatively small in mature forest stands, which accumulate carbon slowly relative to younger stands, if at all [Fig.
@@
 However, $NEP$ is theoretically expected to peak in intermediate-aged stands and thereafter decline, consistent with decelerating C accumulation as stands age [Fig.
-`r age_trends_all`, @odum_strategy_1969], and such declines have been observed [@law_changes_2003; @luyssaert_oldgrowth_2008].
+`r age_trends_all`, @odum_strategy_1969], and such declines have been documented [@law_changes_2003; @luyssaert_oldgrowth_2008].
 The fact that $NEP$ values estimated by our models for 100-year-old stands were not systematically different from those of mature stands (lower for temperate broadleaf, higher for temperate conifer, and equal for boreal; Fig.
@@
 
-In terms of stocks, our study reveals consistent increases in live biomass stocks with stand age, a pattern that is well-known and expected [e.g., @lichstein_biomass_2009; @yang_carbon_2011], contrasting with more variable age trends in deadwood and the organic layer (Fig.
-`r age_trends_all`).
-The latter depend strongly on the type of disturbance. 
+In terms of stocks, our study reveals consistent increases in live biomass stocks with stand age, a pattern that is well-known and expected [e.g., @lichstein_biomass_2009; @yang_carbon_2011].
+This contrasts with more variable age trends in deadwood and the organic layer (Fig. `r age_trends_all`), which depend strongly on the type of disturbance. 
 Disturbances that remove most woody material (e.g., clearcut logging, agriculture) result in negligible deadwood in young stands, followed by a buildup over time [e.g., tropical stands in Figs.
-`r stock_age_trends`, `r age_trends_all`; e.g., @vargas_biomass_2008].
+`r stock_age_trends`, `r age_trends_all`, @vargas_biomass_2008].
 In contrast, natural disturbances (e.g., fire, drought, typhoons/hurricanes) can produce large amounts of deadwood (mostly $DW_{standing}$) that slowly decomposes as the stand recovers, resulting in declines across young stand ages [e.g., temperate and boreal stands in Figs.
-`r stock_age_trends`, `r age_trends_all`; e.g., @carmona_coarse_2002].
+`r stock_age_trends`, `r age_trends_all`, @carmona_coarse_2002].
 Further study and synthesis of non-living C stocks across biomes, stand ages, and disturbance types will be valuable in giving a more comprehensive picture.
@@
 `r all_diagrams_mature`, `r age_trends_all`), provide confidence that our reported mature forest means provide useful baselines for analysis.
-However, there is wide variation around these means, implying that any given stand could deviate substantially from them, and the sample means presented here probably do not represent true biome means (particularly for temperate conifer forests where high-biomass stands are over-represented in *ForC*).
+However, there is wide variation around these means, implying that any given stand could deviate substantially, and the sample means presented here probably do not represent true biome means (particularly for temperate conifer forests where high-biomass stands are over-represented in *ForC*).
 
@@
 There are, of course, notable holes in the ForC variable coverage that limit the scope of our inferences here. 
-Notably, *ForC* currently has sparse--if any--coverage of fluxes to herbivores and higher consumers, along with woody mortality ($M_{woody}$) and $DW$ (Table 1, Figs. `r DW_tot_fig`-`r DW_down_fig`). 
+*ForC* currently has sparse--if any--coverage of fluxes to herbivores and higher consumers, along with woody mortality ($M_{woody}$) and $DW$ (Table 1, Figs. `r DW_tot_fig`-`r DW_down_fig`). 
 We note that there are considerable opportunities to expand data on $M_{woody}$ and $DW_{standing}$ through calculations from existing forest census data.
@@
 To ensure that models are giving the right answers for the right reasons [@sulman_multiple_2018], it is important to benchmark against multiple components of the C cycle that are internally consistent with each other [@collier_international_2018; @wang_golumcnp_2018].
-*ForC*’s tens of thousands of records are readily available in a standardized format, along with all code used in the analyses presented here
+*ForC*’s tens of thousands of records are readily available in a standardized format, along with all code used in the analyses presented here.
 We recommend that researchers use these resources to identify and summarize data specific to the analysis at hand. 
@@
 Note, however, that factors such as stand age and disturbance history are difficult, if not impossible, to detect remotely, and can only be characterized for recent decades [@hansen_highresolution_2013; @curtis_classifying_2018; @song_global_2018].
-Ground-based data such as *ForC* are therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020, and thus constraining variables such as carbon sink potential [@luyssaert_oldgrowth_2008].
+Ground-based data such as those in *ForC* are therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020, and thus constraining variables such as carbon sink potential [@luyssaert_oldgrowth_2008].
 

### 2021-03-01 manuscript/submissions/2021_03_01/response_to_reviews.Rmd
finalized/ moved response to reviews

--- 2021-02-24 manuscript/response_to_reviews.Rmd
@@
 
-**We fixed the noted typo and proofread the entire manuscript.**
+**We fixed the noted typo and proofread the entire manuscript a few times.**
 
@@
 7).
-This finding is consistent with previous studies showing an increase in NEP across relatively young stand ages (Pregitzer and Euskirchen 2004, Baldocchi et al 2001, Luyssaert et al 2008).
+This finding is consistent with previous studies showing an increase in NEP across relatively young stand ages (Baldocchi et al 2001, Pregitzer and Euskirchen 2004, Luyssaert et al 2008).
 However, NEP is theoretically expected to peak in intermediate-aged stands and thereafter decline, consistent with decelerating C accumulation as stands age (Fig.
-9; Odum 1969), and such declines have been observed (Law et al 2003, Luyssaert et al 2008).
+9, Odum 1969), and such declines have been documented (Law et al 2003, Luyssaert et al 2008).
 The fact that NEP values estimated by our models for 100-year-old stands were not systematically different from those of mature stands (lower for temperate broadleaf, higher for temperate conifer, and equal for boreal; Fig.
 9) may be driven by differences in geographical representation across age classes or by the fitting of an inappropriate functional form.
-Moreover, both biomass and non-living C stocks often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands (Luyssaert et al 2008, McGarvey et al 2014, Lichstein et al 2009).
+Moreover, both biomass and non-living C stocks often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands (Luyssaert et al 2008, Lichstein et al 2009, McGarvey et al 2014).
 Additional data, including on age trends of deadwood, the organic layer, and soil C will be important to parsing the timing and extend of an age-related NEP decrease across forest biomes.**

### 2021-03-01 manuscript/ERL_global_C_review.Rmd
working on files for review

@@
 # KNIT IN FIGURES? 
-generate_figures=TRUE
+generate_figures=FALSE
 #note: table fails when knitting in word
@@
 An asterisk after a variable name indicates lack of C cycle closure.")
-BoN_mature_caption = paste( "Figure 6 | C cycle diagram for mature boreaal conifer forests. ", schematics_caption_generic)
+BoN_mature_caption = paste( "Figure 6 | C cycle diagram for mature boreal conifer forests. ", schematics_caption_generic)
 
